### 2022-05-01 README.Rmd
Initial commit

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# streamFind

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of streamFind is to ...

## Installation

You can install the development version of streamFind like so:

``` r
# FILL THIS IN! HOW CAN PEOPLE INSTALL YOUR DEV PACKAGE?
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(streamFind)
## basic example code
```

What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:

```{r cars}
summary(cars)
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.

### 2022-07-01 Handbook_backend_streamFind.Rmd
Large Framework update 01

---
title: "Handbook streamFind: Backend"
author: "Ricardo Cunha"
date: '2022-06-15'
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 9) #results = "asis"
```
<br> <br>

***

```{r libraries, include=FALSE}

library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(plotly)

devtools::load_all()

```

# Objective

# Install

## Install dependencies

## Install streamFind

# Resources

Trimmed example files are included in the streamFind package and are loaded below. The trimming was done with the function `trimSpectraFilesMZR` from streamFind which uses the package [mzR](https://bioconductor.org/packages/release/bioc/html/mzR.html).

```{r resource_files}
r_path <- system.file(package = "streamFind", dir = "extdata")
files <- list.files(r_path, full.names = TRUE)
```

# S4 classes

The streamFind package uses S4 class objects to define the data structure. The S4 classes, respective methods and usage are presented and exemplified through a demonstration of the basic workflow (pre-processing) for HRMS data. In short, the basic pre-processing workflow includes the following recommended steps:  (1) project creation, (2) data conversion (when needed), (3) replicate analyses assignment, (4) peak finding (or peak picking), alignment, grouping and isotopic and adduct annotation across analyses and (5) feature quality evaluation and (6) filtering.

The semantics in streamFind package is described in details in section x. Yet, a brief summary is given below for the most relevant terms:  
  * __trace__ correspond to a raw data signal (e.g., mass trace in MS data) obtained from a given analysis file;  
  * __peak__ correspond to a cluster of traces (e.g., integrated chromatographic MS peak) within an analysis file;  
  * __feature__ correspond to a group of corresponding peaks across analysis files;

## streamProject

The project creation is demonstrated below for three HRMS files (triplicate) where a set of substances were spiked. The project path can be assign as argument but left as default (i.e., the working directory) for this example.
The initial S4 class created in the `newStreamProject` function is _streamProject_.

```{r newStreamProject_call, message=FALSE, results='hide'}
exS1 <- newStreamProject(files = files[1:3])
```

## msData

Because the added files were all with mzML format, the sub-class is directly defined as _msData_, adding additional structure for further processing of MS data. Other data formats/structure will lead to different sub-classes (e.g., _uvData_ and _ramanData_) but is not yet implemented within streamFind.

```{r msData_streamProject_classes}
is(exS1)
```

The _msData_ structure includes the _streamProject_ class plus the `features` slot. The project, date and path slots correspond to the project title, creation date and directory in disk, respectively. The analyses slot is a list with a _msAnalysis_ S4 class object per MS file added. The structure of the _msAnalysis_ class is described in the next sub-chapter.

The `features` `r # TODO make msFeatures after peaks extraction and storage`` 
which includes the _msFeatures_ S4 class(presented in section X).

 * _msData_ structure:

```{r structure_msData}
str(exS1, max.level = 2)
```

## msAnalysis

The _msAnalysis_ class is structured as shown below. Note that the S4 method `getAnalyses` for _msData_ was used to obtain the _msAnalysis_ of the first analysis as defined by the second argument. In general, access and plotting S4 methods can also be applied directly to the _msAnalysis_ class besides _msData_ (as described in \@ref(data-access-and-visualization)) but is not part of the basic workflow and mostly used for advanced data visualization and access of low level data (e.g., traces).

The slots analysis, file and replicate correspond to the analysis name, the full file path and the replicate group name, respectively. The metadata slot is a list of objects with various informative entries of the analysis. The metadata list can be expanded with additional information about the analysis/sample (e.g., sample location, weather conditions, process sensor data, etc.). Below, the `getMetadata` S4 method of _msData_ is used to display the metadata of the first file. Note, `getMetadata` also works for _msAnalysis_ S4 class as shown below in the commented code. The slot parameters is used to store processing steps (e.g., peak picking) as _settingsSet_ S4 class objects, including the processing step name/ID, the algorithm used and the respective settings applied. The parameters can be applied for history track but also for re-run the processing when required. The spectra and chromatograms slots are used to store low level data (i.e., mass traces) from the MS files. Finally, the peaks slot is used to store the chromatographic peaks after peak picking. As no other processing step was applied the spectra, chromatograms and peaks tables are empty for now.

 * _msAnalysis_ structure:

```{r msAnalysis_structure}
exA1 <- getAnalyses(exS1, 1)
str(exA1, max.level = 2)
```

* metadata of _msAnalysis_:

```{r metadata_structure}
t(metadata(exS1, analyses = 1))

#applied directly to the msAnalysis S4 class
t(metadata(exA1))
```

# Low level data access and plotting

The tools for low level access and plotting (i.e., traces and chromatograms) share a target search approach. The data search is based on pre-defined time and mass (inc. _m/z_) targets. For traces (i.e., raw MS mass traces), time (seconds) and _m/z_ (Da) ranges are given for accessing and plotting. Access to mass traces is relevant for further development of data processing functionalists and for visualization of raw data to support evaluation of processing steps (e.g., peaks wrongly not found during peak picking).

## Defining targets

As aforementioned, access to raw data is based on defining time and _m/z_ targets. The function `makeTargets` is used for comprehensively build targets. As shown below, diverse ways can be used to assemble targets, similarly returning a uniform `data.table`. Accessing and plotting S4 methods in streamFind use the `makeTargets` function to collect data. Therefore, the arguments `mz`, `rt`, `ppm`, `sec` and `id` are present in most S4 methods for accessing and plotting data as demonstrated in the following sections.

```{r make_targets}
#case 1
mz01 <- c(247.1651, 239.0628)
rt01 <- c(839, 937)
id01 <- c("target1", "target2")
targets01 <- makeTargets(mz = mz01, rt = rt01, ppm = 20, sec = 60, id = id01)

targets01

#case 2
mz02 <- data.frame(mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652))
rt02 <- data.frame(rtmin = c(809, 907), rtmax = c(869, 967))
targets02 <- makeTargets(mz = mz02, rt = rt02)

targets02

#case 3
mz03 <- data.frame(
  #id = c("target1", "target2"),
  mz = c(247.1651, 239.0628),
  rt = c(839, 937)
)
targets03 <- makeTargets(mz = mz03, ppm = 20, sec = 60)

targets03

#case 4
mz04 <- data.frame(
  id = c("target1", "target2"),
  mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652),
  rtmin = c(809, 907), rtmax = c(869, 967)
)
targets04 <- makeTargets(mz = mz04)

targets04
```

## EICs

`r # TODO remove zeros when returning EIC table?`
`r # TODO add check for spectra/chroms is the object before querying the data?`

Extracted ion chromatograms (EICs) are obtained either from loaded spectra in the _msData_ or _msAnalysis_ objects or from querying traces directly in the raw data file (mzML or mzXML). The latter is less disk memory intense as data is loaded from raw files on demand. The S4 method `EICs` is used to extract MS1 data from raw data files.

```{r EICs_method}
#EIC of targets from first analysis using msAnalysis class
eic01 <- EICs(exA1, mz = targets04)
head(eic01, 6)

#EIC of targets from second analyses using msData class
eic02 <- EICs(exS1, analyses = 2, mz = targets04)
head(eic02, 6)
```

The obtained EICs can be plotted with `plotEICs` S4 method, as shown below. The object for `plotEICs` can be a `data.table` as resulting from the `EICs` method or the actual _msData_ or _msAnalysis_ objects, using the same arguments as in `EICs` for collecting the data (i.e., to define the targets).

```{r plotEICs_static}
#plot the produced EIC data.table
plotEICs(eic01)
```

```{r plotEICs_interactive}
#using the msData class object and targets but with interactive plot
plotEICs(exS1, analyses = NULL, mz = targets04, colorBy = "analyses", interactive = TRUE)

#the same plot as the first for data.table but interactive
plotEICs(exA1, mz = targets04, interactive = TRUE)
```

## TICs

Total ion chromatograms (TICs) are obtained either by pre-loading and adding the chromatogram data to the _msAnalysis_, extracting the data in a `TICs` call (S4 method) or loading all spectra and applying the sum of intensities to each spectrum. Note, TIC is already available as chromatogram in mzML data from the major HRMS vendors. The S4 method TICs first evaluates if the TIC chromatogram is available in the mzML and when not, the spectra data is extracted and used instead.

```{r TICs_method}
# TIC for both analyses in the msData class object
tic01 <- TICs(exS1)
tic01

#TIC of the msAnalysis object
head(TICs(exA1), 6)
```

Similarly to EICs, the TICs can be plotted with `plotTICs` S4 method, as shown below.

```{r plotTICs_method}
#using a data.table as obatined by TICs
plotTICs(tic01, interactive = TRUE)

#using the msData object
plotTICs(exS1, colorBy = "replicates")

#using the msAnalysis
plotTICs(exA1)
```

## XICs

A more informative method of obtaining EICs is via the method `XICs` which extract a the three dimensional (_m/Z_, rt and intensity) chromatograms from given targets. In the example below, the _m/z_ dimension is also included in the resulting table to improve inspection of traces.

```{r XICs_method}
#using the msData class object
xic01 <- XICs(exS1, mz = targets04)
xic01

#with a msAnalysis class object
xic02 <- XICs(exA1, mz = mz01, rt = rt01, ppm = 20, sec = 60)
head(xic02, 3)
```

Similarly, the XICs can be plotted with the S4 method `plotXICs`, as shown below. Other arguments are available for plotting `XICs`, including a target area for the expected targets. As shown in the plot below, the red mark is defined by the arguments `targetsMark` (a table with columns mz and rt to define the expected _m/z_ and retention times) and `ppmMark` and `secMark` to define the mass (in ppm) and time (in seconds) range of the target squares. The `numberRows` is used to control the number of rows of the grouped plot. The concept for the XIC plot is adapted from the R package [MSnbase](https://bioconductor.org/packages/release/bioc/html/MSnbase.html).

```{r plotXICs_method}
#using the output of XICs S4 method (a data.table)
plotXICs(xic01, legendNames = c("target number 1", "target number 2"), plotTargetMark = TRUE,
 targetsMark = targets04[, c("mz", "rt")], ppmMark = 5, secMark = 10, numberRows = 2)

#using the msData class and plotting the first target from the first analysis
plotXICs(exS1[1], mz = targets04[1, ])

#for the msAnalysis, the second target with wide m/z and rt ranges
plotXICs(exA1, mz = mz01[2], rt = rt01[2], ppm = 500, sec = 120)
```

`r # TODO add MSn low level access`

# Basic workflow

The basic workflow consists of essential steps for processing of raw data.

`r # TODO make a scheme of the basic workflow, maybe from the PPT`

## Project creation

The project creation is demonstrated below for six HRMS data files corresponding to blank and wastewater influent samples, both measured in triplicate. The argument `files` is the main input and can either be a list of full path files, a table with file (i.e., full file path), replicate (i.e., the name of the analysis replicate group for each file) and blank (i.e., the name of the associated blank analysis replicate group for each file), or the `analysisInfo` data.frame from the package [patRoon](https://github.com/rickhelmus/patRoon). The project path can be assign as argument `path` and is the directory where all the project files (i.e., scripts, cache databases, objects and results) will be stored.

```{r newStreamProject_basic_workflow, message=FALSE, results='hide'}
exS2 <- newStreamProject(files = files[4:9], path = getwd(), title = "Project Example")
```

## Assign replicate names

An essential aspect of environmental analysis is to operate with sample/analysis replicates. The replicates in _msData_ or _msAnalysis_ can be obtained by the `replicates` S4 method as shown below. The setter (assignment) for the replicate names is demonstrated below with the method `replicates<-` that takes a vector of character strings with the same length as the number of analyses in a _msData_ object or length one for a single analysis in _msAnalysis_. Alternatively, a table with a replicate character column or a character vector can be given in the files or replicates arguments, respectively.

```{r assign_replicates}
#getter for replicate names in msData
replicates(exS2)

#setter for replicate names in msData
replicates(exS2) <- c(rep("blank", 3),rep("influent", 3))  
replicates(exS2)

#setter for replicate name in msAnalysis
exA2 <- getAnalyses(exS2, 4)
replicates(exA2) <- "wastewater"

#getter for replicate name in msAnalysis
replicates(exA2)
```

## Assign blanks

The blank subtraction is another crucial aspect in environmental analysis. The getter and setter for blank subtraction works the same way as replicates. See examples below. The blank analysis replicate is then assign to the respective analysis. In the example, the first replicate group is assign to the influent samples. Note that the blank will also be assigned to itself. Although not shown, different blank analysis replicates can be assigned to different analysis. The subtraction is then applied per analysis considering the assigned blank replicate. Getting and assigning blank replicates from an _msAnalysis_ is also possible. However, assigning a blank without the context of a _streamProject_ and without the structure of _msData_ or similar level classes is not recommended. For cross project analysis, a subset of the analyses (_msAnalysis_ objects) in a given _msData_ can be concatenated to another _msData_, carrying the assigned blank replicate for each analysis as well as all the other analysis information (see section cross-project analysis for more information).

```{r assign_blanks}
#getter for blank replicates in each sample in a msData object
blanks(exS2)

#setter for the blank analysis replicate of msData object
blanks(exS2) <- rep("blank", 6)
blanks(exS2)

#getter for msAnalysis
blanks(exA2)
```

## Other getter S4 methods

```{r other_S4_methods}
#getter for analysis names in msdata
analyses(exS2)

#getter for analysis name in msAnalysis
analyses(exA2)

#getter for polarity of analyses in a msData
polarities(exS2)

#getter for polarity of an msAnalysis
polarities(exA2)

#subsetting an msData with `[`-method for the blank samples
exS2[1:3]
```

## Peak picking

The initial processing step of the basic workflow is the peak picking. Yet, other steps might be necessary prior the pick picking, such as data conversion, data calibration, etc. These are discussed later in section X. The basic workflow in streamFind is based on [patRoon](https://github.com/rickhelmus/patRoon) and most of the output in streamFind can be used/converted to enable the use of native [patRoon](https://github.com/rickhelmus/patRoon) functions and methods.

### Processing parameters

For each processing steps, parameter settings are often used. In streamFind, the parameters used for a given processing step are added/stored as _settings_ S4 class objects in each _msAnalysis_ (or other class of the same level, such as _uvAnalysis_ or _ramanAnalysis_). Each _settings_ object contains the algorithm and list of parameters with the respective values. The example below demonstrates the workflow to perform peak picking either by adding the parameters during the call for `peakPicking` or by initially add the parameters to each _msAnalysis_ and then run the `peakPicking`, which looks for the parameters in each _msAnalysis_. If different settings are applied among the _msAnalysis_ objects, these are used. This means that different processing parameters can be applied within the same function call. If parameters are added during the function call but they already exist in the _msAnalysis_, these are overwritten by the settings used in the function call.

### Create and add parameters

The function `createSettings` is use to assemble the _settings_ object, as shown below. Then, the `addParameters` method is used to either add parameters to each analysis in a _msData_ object or directly to a _msAnalysis_ object. The `getParameters` can be used to check which exist in the _msData_ and _msAnalysis_.

```{r create_processing_parameters}

exS2_pp <- exS2

param <- xcms::CentWaveParam(
  ppm = 15, peakwidth = c(5, 60),
  snthresh = 10, prefilter = c(6, 10000),
  mzCenterFun = "mean", integrate = 2,
  mzdiff = -0.0001, fitgauss = TRUE,
  noise = 250, verboseColumns = TRUE,
  firstBaselineCheck = FALSE,
  extendLengthMSW = TRUE
)

#creating the settings S4 class for peak picking
settings_pp <- createSettings(
  call = "peakPicking",
  algorithm = "xcms3",
  settings = param
)

# the class of settings 
is(settings_pp)

# add the settings to all analyses in msData
exS2_pp <- addParameters(exS2_pp, settings = settings_pp, where = "analyses")

# get the parameters in the third analysis of the msData object
getParameters(exS2_pp, where = "analyses", analyses = 3)

# add the settings to the msAnalysis object
exA2 <- addParameters(exA2, settings = settings_pp)

# get the parameters of the msAnalysis object
getParameters(exA2, call = "peakPicking")
```

### Processing peak picking

```{r peakPicking}

#peak picking call using the stored parameters
exS2_pp <- peakPicking(exS2_pp)

# peaks added for each analysis, as shown by column peaks of show method (below)
exS2_pp

#alternatively, parameters can be added during the function call
exS2_pp_2 <- peakPicking(exS2, algorithm = "xcms3", settings = param)

# parameters are added to each analysis after running peak picking
summary(getParameters(exS2_pp_2, where = "analyses", analyses = 1))

exS2_pp_2

# running peak picking for an msAnalysis object (settings were already added above)
exA2 <- peakPicking(exA2)

# msAnalysis with peaks
exA2
```

### Using multiple settings

As aforementioned, different settings for the same processing step can exist in a _msData_. Below we add different settings for the influent samples, as an example. The use case for such functionality can be connected to optimization of processing steps, for example.

```{r peakPicking_multiple_settings}

exS2_pp_3 <- exS2

# Settings for using openms instead of xcms
settings_pp_2 <- createSettings(
  call = "peakPicking",
  algorithm = "openms",
  settings = list(
    noiseThrInt = 500,
    chromSNR = 10,
    chromFWHM = 10,
    mzPPM = 15,
    reEstimateMTSD = TRUE,
    traceTermCriterion = "sample_rate",
    traceTermOutliers = 5,
    minSampleRate = 1,
    minTraceLength = 3,
    maxTraceLength = -1,
    widthFiltering = "fixed",
    minFWHM = 2,
    maxFWHM = 40,
    traceSNRFiltering = FALSE
  )
)

exS2_pp_3 <- addParameters(exS2_pp_3, settings = settings_pp, where = "analyses")

# adding different settings to analysis 4 to 6
exS2_pp_3 <- addParameters(exS2_pp_3, settings = settings_pp_2, where = "analyses", analyses = 4:6)

# the algorithm of the influent samples is different
sapply(analyses(exS2_pp_3), function(x) getAlgorithm(getParameters(exS2_pp_3, where = "analyses", analyses = x)[[1]]))

#running peak picking
exS2_pp_3 <- peakPicking(exS2_pp_3)

# as the peak parameters with "openms" as algorithm were less stringent, the influent samples have more peaks then when "xcms3" is used 
exS2_pp_3

```

### Export and Import settings

The _settings_ object can be exported as a JSON or rds file and then imported for other projects where the same parameter settings are to be applied. The methods `exportSettings` and `importSettings` are used, as shown below.

```{r export_import_settings}

#export settings as rds
exportSettings(settings_pp, name = "settings_pp", format = "rds", path = getwd())

#export settings as JSON
exportSettings(settings_pp_2, name = "settings_pp_2", format = "json", path = getwd())

#import settings as rds
settings_pp_imported <- importSettings(file = paste0(getwd(), "/settings_pp.rds"))
all.equal(settings_pp, settings_pp_imported)

#import settings as JSON
settings_pp_2_imported <- importSettings(file = paste0(getwd(), "/settings_pp_2.json"))
all.equal(settings_pp_2, settings_pp_2_imported)
```

### Inspecting peaks

The access and visualization of peaks uses the calculated/estimated time and _m/z_ dimensions for collecting and plotting the correspondent mass traces. Examples are shown below to extract and plot peaks from both _msData_ and _msAnalysis_.

```{r}
#access to peaks in msData based on built targets as used for extracting EICs
peaks(exS2_pp[c(1, 4)], mz = mz01[1], rt = rt01[1], ppm = 10, sec = 30)

#access to peak in msAnalysis
peaks(exA2, mz = targets04)

#plotting peaks in msData
plotPeaks(exS2_pp, mz = targets04[2, ], interactive = TRUE, colorBy = "analyses")

#plotting peaks from msAnalysis
plotPeaks(exA2, mz = targets04, interactive = FALSE)

#map plot for peak time and m/z dimensions for msAnalysis
mapPeaks(exA2, mz = targets04[1, ], xlim = 30, ylim = 0.001)

##map plot for peak time and m/z dimensions for msData
mapPeaks(exS2_pp, mz = targets04[2, ], colorBy = "replicates", xlim = 30, ylim = 0.001)
```

## Alingment and Grouping

After peak picking, a common following step is to group the peaks across analyses. A retention time alignment can be applied to correct elution deviations across analyses. The result of peak grouping and alignment is added to the _msFeatures_ S4 class object already present in the _msData_.

### Grouping settings

The processing parameters for grouping and alignment are added similarly to the processing settings for peak picking but the argument `where` is set to "features", as shown below.
The settings are added to the slot parameters of the _msFeatures_ object and can be obtained by the method `getParameters`, as shown below.

```{r creating_grouping_parameters}

param_g <- list(
  rtalign = TRUE,
  loadRawData = TRUE,
  groupParam = xcms::PeakDensityParam(
    sampleGroups = "holder",
    bw = 3,
    minFraction = 0.5,
    minSamples = 1,
    binSize = 0.008,
    maxFeatures = 100),
  preGroupParam = xcms::PeakDensityParam(
    sampleGroups = "holder",
    bw = 5,
    minFraction = 0.5,
    minSamples = 1,
    binSize = 0.008,
    maxFeatures = 100),
  retAlignParam = xcms::PeakGroupsParam(
    minFraction = 1,
    extraPeaks = 0,
    smooth = "loess",
    span = 0.3,
    family = "gaussian")
)

settings_pg <- createSettings(
  call = "peakGrouping",
  algorithm = "xcms3",
  settings = param_g
)

# add processing parameter for peakGrouping to features in the msData
exS2_pp <- addParameters(exS2_pp, settings = settings_pg, where = "features")

# get the processing parameters applied to features
getParameters(exS2_pp, where = "features")
```

### Processing grouping and alignment

The data processing for grouping and alignment follows the same principle as peak picking. Either the settings are added as arguments or the added parameters in the _msFeatures_ object are used in the function call. Below, both cases are demonstrated below.

```{r processing_grouping, message=FALSE}
#using the added parameters
exS2_pg <- peakGrouping(exS2_pp)

#adding the parameters as arguments
exS2_pg_2 <- peakGrouping(exS2_pp, algorithm = "xcms3", settings = param_g)
```

### Inspecting features

Features can be accessed via the method `features` applied to a _msData_ object. The argument `complete` can be set to `TRUE` for a complete list of features information. Also, the argument `average` can be set to `FALSE` for returning the intensity of the feature in each analysis.

```{r inspecting_features}
#getter for feature intensities and intensity deviations
features(exS2_pg, mz = targets04)

#getter for all the feature details by setting complete to TRUE
t(features(exS2_pg_2, mz = targets04[1, ], complete = TRUE))

#getter for features with intensities for each sample by setting average to FALSE
t(features(exS2_pg_2, mz = targets04[1, ], average = FALSE))

#plotting features
plotFeatures(exS2_pg, mz = targets04, colorBy = "targets", interactive = FALSE)

#plotting features for replicates and interactive
plotFeatures(exS2_pg, mz = targets04, colorBy = "replicates", interactive = TRUE)

#plot the individual peaks in features and for each analysis shows the time deviation.
plotFeaturePeaks(exS2_pg, mz = targets04)
```

```{r}

```

# Data access and visualization

Data access and visualization assets are implemented for different data levels: low level data (i.e., mass traces), peaks and features. Additionally, basic and advanced tools are available for accessing, plotting and exporting data (e.g., chromatograms), statistical results (e.g., PCA) and other workflow dependent data/results.

## S4 method

```{r demoCode, echo=FALSE, eval=FALSE, include=FALSE}

# trimSpectraFilesMZR(files = choose.files(), rtr = c(500, 1000), mzr = c(200, 800), onlyMS1 = TRUE)
# 
# mz_05 <- data.frame(
#   retmin = c(809, 907), retmax = c(869, 967),
#   mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652)
# )
# 
# #test <- patRoon::getEICs(msa_1@file, mz_05)
# 
# test <- extractEICs(object, mz = targets04)
# test_xic <- extractXICs(object, mz = targets04)

# projectInfo(msd)
# path(msd)
# files(msd)
# analyses(msd)
# replicates(msd)
# blanks(msd)
# polarities(msd)
# analysisInfo(msd)
# 
# replicates(msd) <- c("s1", "s2")
# blanks(msd) <- c(NA_character_, "s2")
# 
# validObject(msd)
# 
# files(msa_1)
# analyses(msa_1)
# replicates(msa_1)
# blanks(msa_1)
# polarities(msa_1)
# 
# replicates(msa_1) <- "s1"
# blanks(msa_1) <- "s2"

# object <- msd
# 
# loadMS1 <- function(object, analysis = NULL, what = "MS1") {
#   
#   
#   test <- RaMS::grabMSdata(files(object)[1], grab_what = "MS1", rtrange = c(min(rt_02), max(rt_02)))
#   
#   ms2 <- test$MS2
#   
#   ms2[premz > 273 & premz < 274, ]
#   
#   View(ms2)
# 
# }

```

## Peaks

## Features

## Workflows

# Instructions

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents.
For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
Diverse cheatsheets can be found in <https://www.rstudio.com/resources/cheatsheets/>.
This section gives quick instructions to display text, code, tables and figures.

## Text

Text in R markdown has several inline codes for costumization.
The book in <https://bookdown.org/yihui/rmarkdown/> offers a detailed guideline.
Moreover, a cheatsheet can be downdloaded with the following link <https://raw.githubusercontent.com/rstudio/cheatsheets/main/rmarkdown.pdf>.

## Code

Code is automaticelly embedded unless `echo` chunk option is set to `FALSE`.
Evaluation of code in a given chunk can be skiped by setting `eval` to `FALSE`.
Code chunks should always be unically named.

```{r exCode}

#Example of code chunk

list(a = 1:6, b = 1:10)

```

<br>

## Tables

The `knitr::kable` function can be used for displaying tables, see table \@ref(tab:exTable). Consult <https://haozhu233.github.io/kableExtra/awesome_table_in_html.html> for other costumization options.
Table caption is added with the argument `caption` of the `kable` function.

```{r exTable}

df <- data.frame(name = c("a", "b", "c", "d", "e", "f"), col1 = 1:6, col2 = 1:6)

knitr::kable(df, caption = "Example of table caption.") %>%
kable_styling(
  font_size = 12,
  bootstrap_options = c("striped", "hover", "condensed", "responsive"),
  fixed_thead = TRUE
)

```

<br>

## Figures

Figures/plots can also by embedded using diverse packages.
The packages `ggplot2` and `plotly` (interactive) are recommended as they are versatile and easy to use.
Cheatsheet for `ggplot2` can be found in <https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-visualization.pdf>.
For `plotly` the website <https://plotly.com/r/> offers examples and a cheatsheet can be downdloaded in <https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf>.
Note that the `echo = FALSE` parameter can be added to the code chunk to prevent printing of the R code that generate the plots.

### Example `ggplot2`

```{r exGgplot, fig.cap="Plot caption example"}

ggplot(pressure, aes(temperature, pressure)) +
geom_point() +
theme_bw()

```

### Example `plotly`

```{r exPlotly, fig.cap="Plot caption example"}

plot_ly(pressure, x = ~temperature, y = ~pressure, type = "scatter", mode = "markers+lines")

```

<br> <br>

# Signature

The report was produced on `r Sys.time()`. The data and report are stored in of the workstation *`r Sys.info()["nodename"]`* under the user *`r Sys.info()["user"]`*.

<br>

***
Insitut für Energie- und Umwelttechnik e.V. (IUTA)  
<br>
For further questions, please contact Ricardo Cunha ([cunha@iuta.de](cunha@iuta.de)).

<style>
body, p {
  color: black;
  font-family: Arial;
  text-align: justify;
  font-size: 12pt;
}
h1{
  font-size: 26pt;
  font-style: bold;
}
h2{
  font-size:20pt;
  font-style: bold;
}
h3{
  font-size:16pt;
  font-style: bold;
}
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: #239B56;
}
caption {
  color: black;
  font-size: 1.0em;
}
</style>

### 2022-07-07 Handbook_backend_streamFind.Rmd
Large framework update 04

@@
 
+## Add metadata
+
+`r # TODO make addMetadata method for analyses`
+
 # Low level data access and plotting
@@
 _m/z_) targets.
-For traces (i.e., raw MS mass traces), time (seconds) and _m/z_ (Da) ranges are given for accessing and plotting.
-Access to mass traces is relevant for further development of data processing functionalists and for visualization of raw data to support evaluation of processing steps (e.g., peaks wrongly not found during peak picking).
+For traces (i.e., raw MS mass traces), time (in seconds) and _m/z_ (in Da) ranges are given for accessing and plotting.
+Access to mass traces is relevant for further development of data processing steps/functions and for visualization of raw data, easing the evaluation of processing steps (e.g., peaks wrongly not found during peak picking).
 
@@
 
-`r # TODO add MSn low level access`
+# Loading raw data
+
+The S4 method applied above query the data from the raw data files as that is less disk memory intense.
+However, for certain cases the raw data might be needed entirely.
+Therefore, the data can be loaded and stored as a simple _data.table_ in the respective _msAnalysis_ (or other classes of the same level).
+Below we demonstrate how to load (via `loadRawData` S4 method) and access (via `spectra` or `chromatograms`) the raw spectra and chromatograms to/from an _msAnalaysis_ and _msData_ object.
+Both S4 methods (`spectra` or `chromatograms`) return a _data.table_ where further operations can be applied to subset the data, as demonstrated below.
+
+```{r loadRawData}
+#load raw data to an msAnalysis object
+exA1 <- loadRawData(exA1)
+
+#example of accessing the six most intense fragments of the first target with msAnalysis
+s_data <- spectra(exA1)
+s_data <- s_data[preMZ >= targets04$mzmin[1] & preMZ <= targets04$mzmax[1], ]
+s_data <- s_data[rt >= targets04$rtmin[1] & rt <= targets04$rtmax[1], ]
+s_data <- s_data[order(intensity, decreasing = TRUE), ]
+s_data[1:6, ]
+
+#load raw data to an msData object.
+exS1 <- loadRawData(exS1, analyses = NULL)
+
+#example of accessing the two most intense fragments of the first target with msData
+s_data_2 <- spectra(exS1)
+s_data_2 <- s_data_2[preMZ >= targets04$mzmin[1] & preMZ <= targets04$mzmax[1], ]
+s_data_2 <- s_data_2[rt >= targets04$rtmin[1] & rt <= targets04$rtmax[1], ]
+s_data_2 <- s_data_2[order(intensity, decreasing = TRUE), ]
+s_data_2[1:4, ]
+```
+
+`r # TODO create method to access chromatograms in msAnalysis and msData`
 
@@
   ppm = 15, peakwidth = c(5, 60),
-  snthresh = 10, prefilter = c(6, 10000),
+  snthresh = 10, prefilter = c(6, 1500),
   mzCenterFun = "mean", integrate = 2,
   mzdiff = -0.0001, fitgauss = TRUE,
-  noise = 250, verboseColumns = TRUE,
+  noise = 500, verboseColumns = TRUE,
   firstBaselineCheck = FALSE,
@@
 #alternatively, parameters can be added during the function call
-exS2_pp_2 <- peakPicking(exS2, algorithm = "xcms3", settings = param)
+exS2_pp_2 <- peakPicking(exS2, settings = settings_pp)
 
@@
 
-## Alingment and Grouping
+## Alingment and grouping
 
@@
 ```{r creating_grouping_parameters}
-
 param_g <- list(
@@
 #adding the parameters as arguments
-exS2_pg_2 <- peakGrouping(exS2_pp, algorithm = "xcms3", settings = param_g)
+exS2_pg_2 <- peakGrouping(exS2_pp, settings = settings_pg)
+```
+
+### msFeatures class
+
+The structure of the _msFeatures_ is shown below.
+The slot `analyses` is the `data.table` as obtained by the `analysisInfo`, containing basic information about the analyses, such as file, replicate and associeted blank.
+The intensity slot has the intensity of the feature in each analysis (i.e., peak intensity in each analysis).
+The slot metadata has other information about each feature, such as average _m/z_ and retention time.
+
+`r # TODO add annotation description`
+
+The slot parameters has the list of settings used to obtain/process the features.
+
+```{r msFeatures_class}
+#accessing directly the msFeatures slot, not recommended
+str(exS2_pg@features, max.level = 2)
 ```
@@
 
-#getter for all the feature details by setting complete to TRUE
+#getter for all the feature metadata by setting complete to TRUE
 t(features(exS2_pg_2, mz = targets04[1, ], complete = TRUE))
@@
 #getter for features with intensities for each sample by setting average to FALSE
-t(features(exS2_pg_2, mz = targets04[1, ], average = FALSE))
+features(exS2_pg_2, mz = targets04[1, ], average = FALSE)
 
@@
 
-```{r}
+### Alignment results
 
+The alignment results can be inspected with the `plotAlignment` function. However, the time adjustment results are only available when using the algorithm "xcms3" for alignment.
+
+```{r plot_alignment}
+plotAlignment(exS2_pg)
 ```
 
-# Data access and visualization
+## Recursive integration
 
-Data access and visualization assets are implemented for different data levels: low level data (i.e., mass traces), peaks and features.
-Additionally, basic and advanced tools are available for accessing, plotting and exporting data (e.g., chromatograms), statistical results (e.g., PCA) and other workflow dependent data/results.
+As peak picking can result in false negatives and to ensure that false positives are less likely, a recursive integration is often performed for filling missing peaks in a feature.
+The recursive integration consists of extracting the mass traces in the feature region from analyses not represented, returning the hipotetical peak intensity.
+The exact collection of mass traces and calculation of the peak intensity (or height) is dependent on the algorithm applied.
+Below, an example of the recursive integration workflow using the function [fillChromPeaks](https://rdrr.io/bioc/xcms/man/fillChromPeaks.html) from the package [xcms](https://bioconductor.org/packages/release/bioc/html/xcms.html) is shown.
+The function `peakFilling` is used for recursive integration and a _settings_ S4 class object is given to define the algorithm and respective parameters.
+A _settings_ object can be obtained with `fillingSettingsDefaultXCMS` for a default _ChromPeakAreaParam_ from xcms.
+Note that the peakFilling settings are added to the parameters slot of the _msData_ object.
 
-## S4 method
+```{r recursive_integration_settings}
+#default settings S4 class with ChromPeakAreaParam from xcms as parameter settings
+settings_fl <- fillingSettingsDefaultXCMS()
+
+settings_fl
+```
+
+```{r recursive_integration_procesing, results='hide', message=FALSE}
+#filling peaks for msData
+exS2_pg_fl <- peakFilling(exS2_pg, settings = settings_fl)
+```
+
+## Inspecting of filled peaks
+
+To inspect filled peaks, the function `plotFeaturePeaks` can be used as shown below.
+The white dots represent filled peaks, meaning that the intensity value was not obtained from peak picking.
+The recursive integration is relevant to further minimize false negative features by reinforce the presence of the feature in the relevant analyses but not in the blank analysis replicate.
+Also, on the other end, for reducing false positives by confirming the presence of the features at similar level in the blank replicate analysis, as shown below.
+The second feature, at 935 seconds, peaks were picked only for the influent but the blank has very similar level as confirmed by recursive integration, which most likely is related to peaks from noise.
+Applying filtering by blank subtraction further during the basic data processing workflow, would remove the irrelevant features, such as the example shown below.
+
+```{r recursive_integration_inspection}
+# inspecting the filled peaks for a given mz and rt target
+plotFeaturePeaks(exS2_pg_fl, mz = 441.1670, rt = 916, ppm = 20, sec = 60)
+```
+
+## Annotation
+
+`r # TODO improve annotation of features/peaks``
+
+## Quality
+
+## Filter
+
+# MSn
+
+Often MS acquisition includes operation in tandem mode, where fragmentation spectra (i.e., MSn, where n is the level of fragmentation) are acquired.
+The most common fragmentation data is MS/MS or MS2 either acquired via data dependent or independent acquisition.
+
+## Processing settings
+
+Settings are needed to extract and average the fragmentation data.
+Below an example of settings used for accessing and averaging MS2 data is given.
+The resulting _settings_ object is then used further in the call to `MS2s` or `plotMS2s`.
+
+```{r create_settings_msn}
+#the parameter settings to obtained and process MSn spectra
+param_msn <- list(
+  isolationTimeWindow = 0,
+  isolationMassWindow = 1.3,
+  clusteringMethod = "distance",
+  clusteringUnit = "ppm",
+  clusteringWindow = 10,
+  minIntensityPre = 200,
+  minIntensityPost = 300,
+  asPatRoon = FALSE, #when TRUE, is only used by running generateMS2
+  mergeVoltages = TRUE,
+  mergeBy = "analyses"
+)
+
+#creating the settings S4 class
+settings_msn <- createSettings(
+  call = "extractMSn",
+  algorithm = "streamFind",
+  settings = param_msn
+)
+
+# the class of settings 
+is(settings_msn)
+```
+
+## Access and plot MS2
+
+The method `MS2s` can be used to collect MS2 data from given targets, following the same principle as `makeTargets` within `EICs`.
+Similarly, `plotMS2s` can be used to access and directly plot MS2 data from targets, as shown below.
+A more general function for accessing the data is `extractMSn`, where further fragmentation levels can be accessed.
+When using streamFind basic workflow, the use of `MS2s` and `plotMS2s` is recommended.
+All the S4 methods/function can be applied for _msAnalysis_ and _msData_ objects.
+
+```{r extract_msn}
+#head of MS2 data for target 1 in a msAnalysis object
+head(MS2s(exA1, mz = targets04[1, ], settings = settings_msn))
+
+#ploting the MS2 data of the first target in a msAnalysis object
+plotMS2s(exA1, mz = targets04[1, ], settings = settings_msn, interactive = FALSE)
+
+#most intense MS2 data traces for both targets in a msData object
+s_data_3 <- MS2s(exS1, mz = targets04, settings = settings_msn)
+s_data_3 <- s_data_3[order(intensity, decreasing = TRUE), ]
+head(s_data_3)
+
+#ploting the MS2 data of the first target in a msAnalysis object
+plotMS2s(exS1, mz = targets04, settings = settings_msn, interactive = TRUE, colorBy = "targets")
+```
+
+## MS2 for features/peaks
+
+` # TODO create function for collecting MS2 data for peaks/features`
+
+# Workflows
+
+# Miscellaneous
 
@@
 
-## Peaks
-
-## Features
-
-## Workflows
-
 # Instructions
@@
 
-```{r exCode}
+```{r exCode, echo=FALSE, eval=FALSE, include=FALSE}
 
@@
 
-```{r exTable}
+```{r exTable, echo=FALSE, eval=FALSE, include=FALSE}
 
-df <- data.frame(name = c("a", "b", "c", "d", "e", "f"), col1 = 1:6, col2 = 1:6)
-
-knitr::kable(df, caption = "Example of table caption.") %>%
-kable_styling(
-  font_size = 12,
-  bootstrap_options = c("striped", "hover", "condensed", "responsive"),
-  fixed_thead = TRUE
-)
+# df <- data.frame(name = c("a", "b", "c", "d", "e", "f"), col1 = 1:6, col2 = 1:6)
+# 
+# knitr::kable(df, caption = "Example of table caption.") %>%
+# kable_styling(
+#   font_size = 12,
+#   bootstrap_options = c("striped", "hover", "condensed", "responsive"),
+#   fixed_thead = TRUE
+# )
 

### 2022-07-07 R/S4_class_004_msData.R

@@ -42,9 +42,6 @@ msData_validity <- function(object) {
 #'
 #' @export
 #'
-#' @importClassesFrom patRoon featuresSIRIUS featureGroupsSIRIUS componentsCliqueMS
-#' @importFrom purrr quietly
-#'
 #' @md
 setClass("msData",
   slots = c(
@@ -101,6 +98,8 @@ setMethod("show", "msData", function(object) {
 #' @template args-single-analyses
 #'
 #' @export
+#'
+#' @aliases getAnalyses,msData,msData-method
 #'
 setMethod("getAnalyses", "msData", function(object, analyses = NULL) {
 
@@ -178,6 +177,8 @@ setMethod("[", c("msData", "ANY", "missing", "missing"), function(x, i, ...) {
 #'
 #' @export
 #'
+#' @aliases polarities,msData,msData-method
+#'
 setMethod("polarities", "msData", function(object) {
   mt <- metadata(object, which = "polarity")
   mt_v <- mt$polarity
@@ -225,16 +226,18 @@ setMethod("EICs", "msData", function(object,
 #'
 #' @export
 #'
+#' @aliases plotEICs,msData,msData-method
+#'
 setMethod("plotEICs", "msData", function(object,
                                          analyses = NULL,
-                                         mz = NULL, ppm = 20,
-                                         rt = NULL, sec = 30, id = NULL,
+                                         mz = NULL, rt = NULL,
+                                         ppm = 20, sec = 30, id = NULL,
                                          colorBy = "targets",
                                          legendNames = NULL,
                                          title = NULL,
                                          interactive = FALSE) {
 
-  eic <- extractEICs(object, analyses, mz, rt, ppm, id)
+  eic <- extractEICs(object, analyses = analyses, mz = mz, rt = rt, ppm = ppm, sec = sec, id = id)
 
   return(plotEICs(eic, analyses = NULL, colorBy, legendNames, title, interactive))
 })
@@ -248,6 +251,8 @@ setMethod("plotEICs", "msData", function(object,
 #' @export
 #'
 #' @importFrom data.table `:=` setcolorder
+#'
+#' @aliases TICs,msData,msData-method
 #'
 setMethod("TICs", "msData", function(object, analyses = NULL) {
 
@@ -288,6 +293,8 @@ setMethod("TICs", "msData", function(object, analyses = NULL) {
 #'
 #' @export
 #'
+#' @aliases plotTICs,msData,msData-method
+#'
 setMethod("plotTICs", "msData", function(object,
[…]

### 2022-07-15 Handbook_backend_streamFind.Rmd
Large framework update 05

@@
 author: "Ricardo Cunha"
-date: '2022-06-15'
 output:
@@
     theme: paper
+    css: "Handbook_backend_streamFind.css"
+editor_options: 
+  chunk_output_type: console
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 9) #results = "asis"
+knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 9, cache = TRUE)
 ```
@@
 devtools::load_all()
-
 ```
@@
 param <- xcms::CentWaveParam(
-  ppm = 15, peakwidth = c(5, 60),
-  snthresh = 10, prefilter = c(6, 1500),
+  ppm = 10, peakwidth = c(5, 60),
+  snthresh = 5, prefilter = c(4, 800),
   mzCenterFun = "mean", integrate = 2,
   mzdiff = -0.0001, fitgauss = TRUE,
-  noise = 500, verboseColumns = TRUE,
+  noise = 250, verboseColumns = TRUE,
   firstBaselineCheck = FALSE,
@@
     bw = 3,
-    minFraction = 0.5,
-    minSamples = 1,
+    minFraction = 0.6,
+    minSamples = 2,
     binSize = 0.008,
@@
     bw = 5,
-    minFraction = 0.5,
-    minSamples = 1,
+    minFraction = 1,
+    minSamples = 3,
     binSize = 0.008,
@@
 
-## Inspecting of filled peaks
+### Inspection of filled peaks
 
@@
 
+Annotation of isotopes and adducts is crucial not only to prioritize the relevant features but also to deliver essential information for further identification, such as charge, isotopic pattern and adduct yield.
+
+```{r annotation_processing, results='hide'}
+#creating settings for using CAMERA
+settings_an <- createSettings(
+  call = "peakAnnotation",
+  algorithm = "camera",
+  settings = list(
+    ionization = NA_character_,
+    onlyIsotopes = FALSE,
+    minSize = 1,
+    relMinReplicates = 1,
+    extraOpts = list(
+      sigma = 6,
+      perfwhm = 0.35,
+      cor_eic_th = 0.3,
+      graphMethod = "hcs",
+      pval = 0.05,
+      calcCiS = TRUE,
+      calcIso = TRUE,
+      calcCaS = TRUE,
+      maxcharge = 3,
+      maxiso = 5,
+      ppm = 15,
+      mzabs = 0.008,
+      rules =  data.table::fread(system.file("rules/primary_adducts_pos.csv", package = "CAMERA"), header = TRUE),
+      multiplier = 3,
+      max_peaks = 500,
+      intval = "maxo"
+    )
+  )
+)
+
+exS2_pg_fl_an <- peakAnnotation(exS2_pg_fl, settings = settings_an)
+```
+
+The results of annotation are stored under annotation of the _msFeatures_ S4 class and matches the [components](https://rickhelmus.github.io/patRoon/reference/components-class.html) S4 class of the [patRoon](https://github.com/rickhelmus/patRoon) package.
+This means that the feature components can be accessed via the `getAnnotation` method for an _msData_ object and inspected also with native methods from patRoon.
+
+### Inspection of annotation
+
+The annotation can be inspected with the `annotation` S4 method for the _msData_ object, as shown below.
+The method `plotAnnotation` can then be used to visualize the annotation of features, as shown below.
+
+```{r}
+#isotopic and adduct features annotated with a given target mz and rt
+t(annotation(exS2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE))
+
+#plotting the annotation of each feature in a given target
+plotAnnotation(exS2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE)
+```
+
 `r # TODO improve annotation of features/peaks``
 
+## Filter data
+
+Data filtration is crucial for proper prioritization of relevant peaks/features. In streamFind, two levels of data filtration is implemented.
+
+### Sub-seeting with '['
+
+The first is based on data sub-setting using the method `[` and is permanent, meaning that the peaks/features are completely removed from the _msAnalysis_/_msFeatures_.
+An example is shown below for data filtration using the `[` method for different classes.
+
+```{r}
+#sub-setting the msData class on analyses (the 3 first)
+exS2_pg[1:3, ]
+
+#sub-setting the msData class on analys2s and features (the 3 last analyses and features annotated with a target)
+target_s <- annotation(exS2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE)
+exS2_pg_fl_an[4:6, target_s$id]
+
+#sub-seeting the msData class on analys2s and features by index
+exS2_pg_fl_an[1:2, 1:5]
+
+#sub-setting an msAnalysis class on the first 5 peaks
+exA2[1:5]
+```
+
+### Filter methods
+
+The second method uses filter methods (i.e., `filterPeaks` and `filterFeatures`) and is conservative, meaning that the data is conserved but tagged as filtered (i.e., the _filtered_ column is changed to `TRUE` and the respective filter tag is added to column _filter_, as shown below).
+Both filtered peaks and features can be permanently removed by running the methods `removeFilteredPeaks` or `removeFilteredFeatures`, respectively.
+Note that, removing features permanently does not remove filtered peaks but peaks not represented by features are filtered with "grouping" tag.
+When removing peaks, the features left without peaks representation are removed as well.
+Also, the methods `peaks` and `features` have the argument _filtered_ to return filtered peaks/features.
+The default for _filtered_ argument is `TRUE`, meaning that filtered peaks/features are returned by default for methods `peaks` and `features`.
+
+`r # TODO add filtered argument to peaks and features methods`
+
+```{r}
+
+```
+
 ## Quality
 
-## Filter
+The quality of the data is crucial to improve the prioritization of peaks/features and the relevance of the results.
+In streamFind, two functions are available to calculate quality of peaks/features: `calculateSNR` and `calculateMetaClean` to estimate the signal-to-noise ration and several fitting parameters of each peak/feature, respectively.
+The latter is based on the [MetaClean](https://rdrr.io/cran/MetaClean/) R package and is applied via patRoon.
+It is important to emphasize that peaks/features should be filtered with basic filters (e.g., blank subtraction, minimum intensity, etc.) before applying quality calculation has the computational demand for large number of peaks/features is high.
+
+```{r}
+
+exS2_pg <- calculateSNR(exS2_pg, targetsID = features(exS2_pg)[["id"]][1:5])
+
+features(exS2_pg, complete = TRUE)[1:5, ]
+
+```
 
@@
 
+# object <- exS2_pg_fl
+# settings <- annotationSettingsDefaultCAMERA()
+# settings <- annotationSettingsDefaultRAMClustR()
+# object <- addParameters(object, where = "features", settings)
+# 
+# 
+# 
+# findFGroup(comp, "M239_R936_497")
+# as.data.frame(comp[702, ])
+# 
+# mapPeaks(object, mz = data.table(mzmin = 208, mzmax = 220, rtmin = 635, rtmax = 645))
+# 
+# plotXICs(object, analyses = 1:2, mz = 262.0450, ppm = 40, rt = 936, sec = 120)
+# 
+# View(features(object, mz = 441.1670, ppm = 20, rt = 916, sec = 10, complete = TRUE))
+# View(features(object, complete = TRUE)[monoiso %in% "m441.168_d0.7_r916_t6_f3225", ])
+
 # trimSpectraFilesMZR(files = choose.files(), rtr = c(500, 1000), mzr = c(200, 800), onlyMS1 = TRUE)
@@
   text-align: justify;
-  font-size: 12pt;
+  font-size: 14pt;
 }

### 2022-07-15 R/S4_000_generics.R

@@ -67,7 +67,7 @@ setGeneric("loadRawData", function(object, ...) standardGeneric("loadRawData"))
 #' spectra
 #'
 #' @noRd
-setGeneric("spectra", function(object, ...) standardGeneric("spectra"))
+setGeneric("spectra", function(obj, ...) standardGeneric("spectra"))
 
 #' EICs
 #'
@@ -139,21 +139,25 @@ setGeneric("plotFeatures", function(object, ...) standardGeneric("plotFeatures")
 #' @noRd
 setGeneric("mapFeatures", function(object, ...) standardGeneric("mapFeatures"))
 
+#' filterFeatures
+#'
+#' @noRd
+setGeneric("filterFeatures", function(object, ...) standardGeneric("filterFeatures"))
+
 #' hasAdjustedRetentionTime
 #'
 #' @noRd
 setGeneric("hasAdjustedRetentionTime", function(object) standardGeneric("hasAdjustedRetentionTime"))
 
-#' components
+#' annotation
 #'
 #' @noRd
-setGeneric("components", function(object, ...) standardGeneric("components"))
+setGeneric("annotation", function(object, ...) standardGeneric("annotation"))
 
-#' plotComponents
+#' plotAnnotation
 #'
 #' @noRd
-setGeneric("plotComponents", function(object, ...) standardGeneric("plotComponents"))
-
+setGeneric("plotAnnotation", function(object, ...) standardGeneric("plotAnnotation"))
 
 #' addParameters
 #'

### 2022-07-19 handbooks/Handbook_backend_streamFind.Rmd
Large Framework update 06

--- 2022-07-15 Handbook_backend_streamFind.Rmd
@@
 ---
-title: "Handbook streamFind: Backend"
+title: "Handbook streamFind: backend"
 author: "Ricardo Cunha"
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 9, cache = TRUE)
+knitr::opts_chunk$set(echo = TRUE, fig.align = "center", cache = TRUE) # fig.width = 9,
 ```
@@
 
-```{r libraries, include=FALSE}
+```{r libraries, include=FALSE, cache=FALSE}
 
@@
 library(plotly)
-
-devtools::load_all()
+library(streamFind)
+
+#devtools::load_all("../")
 ```
@@
 
+The streamFind package is a backend (code based) and frontend ([shiny](https://shiny.rstudio.com/) app) platform for assembly of modular workflows to support spectrometric and spectroscopic data processing.
+The major focus of streamFind is data processing for environmental and quality studies of the water cycle.
+The streamFind package aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data.
+The streamFind package can be used by academics but also by technicians due to the comprehensive documentation and well categorized set of integrated functionalities (modules).
+This handbook focuses on describing and demonstrating the backend of the streamFind package.
+
 # Install
@@
 
-Trimmed example files are included in the streamFind package and are loaded below.
-The trimming was done with the function `trimSpectraFilesMZR` from streamFind which uses the package [mzR](https://bioconductor.org/packages/release/bioc/html/mzR.html).
-
-```{r resource_files}
+## Files
+
+The streamFind package includes example files, which are located  in the folder `system.file(package = "streamFind", dir = "extdata")`.
+Note, the MS files (.mzML) were trimmed with the function `trimSpectraFilesMZR` to reduce their size.
+The files available are described in Table \@ref(tab:description-files).
+
+```{r resource-files, cache=FALSE}
+#path to files
 r_path <- system.file(package = "streamFind", dir = "extdata")
+
+#list of file full paths
 files <- list.files(r_path, full.names = TRUE)
@@
 
-# S4 classes
-
-The streamFind package uses S4 class objects to define the data structure.
-The S4 classes, respective methods and usage are presented and exemplified through a demonstration of the basic workflow (pre-processing) for HRMS data.
-In short, the basic pre-processing workflow includes the following recommended steps:  (1) project creation, (2) data conversion (when needed), (3) replicate analyses assignment, (4) peak finding (or peak picking), alignment, grouping and isotopic and adduct annotation across analyses and (5) feature quality evaluation and (6) filtering.
-
-The semantics in streamFind package is described in details in section x. Yet, a brief summary is given below for the most relevant terms:  
+```{r description-files, echo=FALSE}
+df_files_desc <- data.frame(
+  file_name = basename(files),
+  type = c(rep("MS/MS", 9)),
+  polarity = c(rep("positive", 9)),
+  description = c(
+    rep("Basic centroided MS data spiked with chemicals.", 3),
+    rep("Blank centroided MS data spiked with internal standards.", 3),
+    rep("Wastewater secondary effluent centroided MS data spiked with internal standards.", 3)
+  )
+)
+
+knitr::kable(df_files_desc, caption = "Files included in the streamFind package.") %>%
+kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE)
+```
+
+## Spiked chemicals
+
+`r # TODO add a list of spiked chemical in the package` List the chemicals spiked in the MS data files.
+
+# Objects
+
+The streamFind package uses S4 class objects.
+The S4 classes, main methods and usage are presented and exemplified in this section.
+An overview of the S4 classes and their hierarchy is presented in Figure \@ref(fig:classes-overview) using MS analyses as an example.
+Other data types, such as from UV and Raman analyses, would lead to different classes (e.g., _uvData_ and _uvAnalysis_ and _ramanData_ and _ramanAnalysis_, respectively) but the hierarchy is similar as shown for MS data.
+
+```{r classes-overview, echo=FALSE, fig.cap="Overview of the S4 classes and their hierarchy with emphasis on MS analyses.", out.width="120%"}
+knitr::include_graphics(paste0(getwd(), "/hb_backend_figures/classes_overview.png"))
+```
+
+## Basic terminalogy
+
+In streamFind, the terms listed below are consistently applied throughout the objects and functionalities.
+
+* __project__ is the project title;
+* __analysis__ is the name or the actual analysis file;
+* __replicate__ is the name of an analysis replicate group;
+* __blank__ is the name of a blank analysis replicate group ot be used for blank (blind) subtration;
 * __trace__ correspond to a raw data signal (e.g., mass trace in MS data) obtained from a given analysis file;  
+* __spectra__ in MS is a table with a spectrum for each cycle time (i.e., retention time);
+* __EIC__ is an extracted ion chromatograms (retention time vs intensity);
+* __TIC__ is a total ion chromatograms (retention time vs intensity);
+* __XIC__ is a three dimensional ion chromatogram (_m/z_ vs retention time vs intensity);
 * __peak__ correspond to a cluster of traces (e.g., integrated chromatographic MS peak) within an analysis file;
 * __feature__ correspond to a group of corresponding peaks across analysis files;
-
-## streamProject
+* __MS2__ correspond to second order fragmentation data as acquired in tandem MS;
+(...)
+
+## streamProject class
 
@@
 
-```{r newStreamProject_call, message=FALSE, results='hide'}
+```{r newStreamProject-call, message=FALSE, results='hide'}
 exS1 <- newStreamProject(files = files[1:3])
@@
 
-## msData
+## msData class
 
@@
 
-## msAnalysis
+## msAnalysis class
 
@@
 Note that the S4 method `getAnalyses` for _msData_ was used to obtain the _msAnalysis_ of the first analysis as defined by the second argument.
-In general, access and plotting S4 methods can also be applied directly to the _msAnalysis_ class besides _msData_ (as described in \@ref(data-access-and-visualization)) but is not part of the basic workflow and mostly used for advanced data visualization and access of low level data (e.g., traces).
+In general, access and plotting S4 methods can also be applied directly to the _msAnalysis_ class besides _msData_ (as described in \@ref(low-level-data-access-and-plotting)) but is not part of the basic workflow and mostly used for advanced data visualization and access of low level data (e.g., traces).
 
@@
 ```{r metadata_structure}
-t(metadata(exS1, analyses = 1))
+t(streamFind::metadata(exS1, analyses = 1))
 
 #applied directly to the msAnalysis S4 class
-t(metadata(exA1))
+t(streamFind::metadata(exA1))
 ```
@@
 
-The basic workflow consists of essential steps for processing of raw data.
-
-`r # TODO make a scheme of the basic workflow, maybe from the PPT`
+The basic workflow (Figure \@ref(fig:workflow-scheme)) includes the following steps: (1) project creation, (2) data conversion (when required/applicable), (3) assignment of analysis replicate names and respective blank analysis replicates, (4) peak finding (or peak picking), (5) peak alignment and grouping into features across analyses, (6) filling of peaks missing in features, (7) annotation of isotopic and adduct features and (8) feature quality evaluation and (9) filtering.
+The usage of the S4 class objects within the basic workflow (pre-processing) for HRMS data is presented in this section with examples.
+
+```{r workflow-scheme, echo=FALSE, fig.cap="Basic workflow scheme."}
+knitr::include_graphics(paste0(getwd(), "/hb_backend_figures/workflow_scheme.png"))
+```
 
@@
 
-```{r}
+```{r inspecting_peaks}
 #access to peaks in msData based on built targets as used for extracting EICs
@@
 
-```{r annotation_processing, results='hide'}
+```{r annotation_processing, results='hide', message=FALSE}
 #creating settings for using CAMERA
@@
 
-```{r}
+```{r annotation_results}
 #isotopic and adduct features annotated with a given target mz and rt
-t(annotation(exS2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE))
+t(streamFind::annotation(exS2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE))
 
@@
 
-Data filtration is crucial for proper prioritization of relevant peaks/features. In streamFind, two levels of data filtration is implemented.
+Data filtration is crucial for proper prioritization of relevant peaks/features.
+In streamFind, two levels of data filtration is implemented: (1) data sub-setting (permanent) and (2) filter processing steps for peaks and/or features (tagged but not removed).
 
@@
 
-```{r}
+```{r filter_subsetting}
 #sub-setting the msData class on analyses (the 3 first)
@@
 #sub-setting the msData class on analys2s and features (the 3 last analyses and features annotated with a target)
-target_s <- annotation(exS2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE)
+target_s <- streamFind::annotation(exS2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE)
 exS2_pg_fl_an[4:6, target_s$id]
 
-#sub-seeting the msData class on analys2s and features by index
+#sub-seeting the msData class on analyses ([i, ]) and features ([, j]) by index
 exS2_pg_fl_an[1:2, 1:5]
@@
 
-### Filter methods
-
-The second method uses filter methods (i.e., `filterPeaks` and `filterFeatures`) and is conservative, meaning that the data is conserved but tagged as filtered (i.e., the _filtered_ column is changed to `TRUE` and the respective filter tag is added to column _filter_, as shown below).
+### Filter functions
+
+The second method uses filter functions (i.e., `filterPeaks` and `filterFeatures`) and is conservative, meaning that the data is conserved but tagged as filtered (i.e., the _filtered_ column is changed to `TRUE` and the respective filter tag is added to column _filter_, as shown below).
 Both filtered peaks and features can be permanently removed by running the methods `removeFilteredPeaks` or `removeFilteredFeatures`, respectively.
@@
 Also, the methods `peaks` and `features` have the argument _filtered_ to return filtered peaks/features.
-The default for _filtered_ argument is `TRUE`, meaning that filtered peaks/features are returned by default for methods `peaks` and `features`.
-
-`r # TODO add filtered argument to peaks and features methods`
-
-```{r}
+The default for the _filtered_ argument is `TRUE`, meaning that filtered peaks/features are returned by default for methods `peaks` and `features`.
+Simmilarly, to peakPicking, peak Grouping, etc., the filter methods use a _msSettings_ S4 class to store the filtering parameters.
+Also, the filtering _msSettings_ object when applied is stored in the parameters section opf the respective class (i.e., `filterPeaks` settings are stored in the _msAnalysis_ and `filterFeatures` settings are stored in the _msFeatures_).
+
+#### Filter features
+
+```{r filter_features}
+
+filters <- list(
+  minIntensity = 5000,
+  blankThreshold = 3,
+  maxReplicateIntensityDeviation = 30,
+  minReplicateAbundance = 3,
+  excludeIsotopes = TRUE,
+  excludeAdducts = TRUE
+)
+
+settings_filFeat <- createSettings(
+  call = "filterFeatures",
+  algorithm = "filter",
+  settings = filters
+)
+
+exS2_pg_fl_an_filFeat <- addParameters(exS2_pg_fl_an, where = "features", settings = settings_filFeat)
+
+exS2_pg_fl_an_filFeat <- filterFeatures(exS2_pg_fl_an_filFeat)
+
+#number of features not filtered
+nrow(features(exS2_pg_fl_an, filtered = FALSE))
+
+#number of features not filtered after applying the filters
+nrow(features(exS2_pg_fl_an_filFeat, filtered = FALSE))
+
+#remove permanently all filtered features from the msData object
+exS2_pg_fl_an_filFeat_rem <- removeFilteredFeatures(exS2_pg_fl_an_filFeat, which =  "all")
+
+exS2_pg_fl_an_filFeat_rem
+```
+
+#### Filter peaks
+
+```{r filter_peaks}
 
@@
 
-```{r}
-
-exS2_pg <- calculateSNR(exS2_pg, targetsID = features(exS2_pg)[["id"]][1:5])
-
-features(exS2_pg, complete = TRUE)[1:5, ]
+```{r quality_sn}
+
+# exS2_pg <- calculateSNR(exS2_pg, targetsID = features(exS2_pg)[["id"]][1:5])
+# 
+# features(exS2_pg, complete = TRUE)[1:5, ]
 
@@
 
-```{r exTable, echo=FALSE, eval=FALSE, include=FALSE}
-
-# df <- data.frame(name = c("a", "b", "c", "d", "e", "f"), col1 = 1:6, col2 = 1:6)
-# 
-# knitr::kable(df, caption = "Example of table caption.") %>%
-# kable_styling(
-#   font_size = 12,
-#   bootstrap_options = c("striped", "hover", "condensed", "responsive"),
-#   fixed_thead = TRUE
-# )
+```{r exTable, echo=TRUE, eval=TRUE, include=TRUE}
+
+df <- data.frame(name = c("a", "b", "c", "d", "e", "f"), col1 = 1:6, col2 = 1:6)
+
+knitr::kable(df, caption = "Example of table caption.") %>%
+kable_styling(
+  font_size = 12,
+  bootstrap_options = c("striped", "hover", "condensed", "responsive"),
+  fixed_thead = TRUE
+)
 

### 2022-07-20 handbooks/Handbook_backend_streamFind.Rmd
Large framework update 07

@@
 ```{r libraries, include=FALSE, cache=FALSE}
-
 library(knitr)
@@
 library(streamFind)
-
-#devtools::load_all("../")
 ```
@@
 The streamFind package is a backend (code based) and frontend ([shiny](https://shiny.rstudio.com/) app) platform for assembly of modular workflows to support spectrometric and spectroscopic data processing.
-The major focus of streamFind is data processing for environmental and quality studies of the water cycle.
+The major focus of streamFind is data processing for environmental and quality studies.
 The streamFind package aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data.
@@
 
-The streamFind package uses S4 class objects.
-The S4 classes, main methods and usage are presented and exemplified in this section.
+The streamFind package uses S4 class objects (e.g., _streamProject_, _msData_, _msAnalysis_).
+The S4 classes and their main getter and setter methods are presented and exemplified in this section.
 An overview of the S4 classes and their hierarchy is presented in Figure \@ref(fig:classes-overview) using MS analyses as an example.
@@
 
-```{r classes-overview, echo=FALSE, fig.cap="Overview of the S4 classes and their hierarchy with emphasis on MS analyses.", out.width="120%"}
+```{r classes-overview, echo=FALSE, fig.cap="Overview of the S4 classes and their hierarchy with emphasis on MS analyses.", out.width="95%"}
 knitr::include_graphics(paste0(getwd(), "/hb_backend_figures/classes_overview.png"))
@@
 
-## Basic terminalogy
-
-In streamFind, the terms listed below are consistently applied throughout the objects and functionalities.
-
-* __project__ is the project title;
-* __analysis__ is the name or the actual analysis file;
-* __replicate__ is the name of an analysis replicate group;
-* __blank__ is the name of a blank analysis replicate group ot be used for blank (blind) subtration;
-* __trace__ correspond to a raw data signal (e.g., mass trace in MS data) obtained from a given analysis file;  
-* __spectra__ in MS is a table with a spectrum for each cycle time (i.e., retention time);
-* __EIC__ is an extracted ion chromatograms (retention time vs intensity);
-* __TIC__ is a total ion chromatograms (retention time vs intensity);
-* __XIC__ is a three dimensional ion chromatogram (_m/z_ vs retention time vs intensity);
-* __peak__ correspond to a cluster of traces (e.g., integrated chromatographic MS peak) within an analysis file;
-* __feature__ correspond to a group of corresponding peaks across analysis files;
-* __MS2__ correspond to second order fragmentation data as acquired in tandem MS;
+## Nomenclature
+
+The nomenclature listed below is consistently applied throughout the objects and arguments of functionalities in the streamFind package.
+Other less relevant terms are used but not listed.
+For further information please consult the specific documentation of each object or functionality when required using the `?x` method, where x is the name of the S4 class, S4 method or function.
+
+* _streamProject_ refers to the broader S4 class applied in the streamFind package and holds basic project information, see \@ref(streamproject-class);
+* _msData_ refers to the S4 subclass of _streamProject_ when MS analyses (i.e., MS files) are added to the project, see \@ref(msdata-class);
+* _msAnalysis_ refers to the S4 class of the object generated for each MS file added to the _msData_ object in the slot "analyses";
+* `projectInfo` is the method to obtain the basic information (i.e., title, date and path) of a _streamProject_ object;
+* `analyses` is the name of the slot holding the analyses (as list) and is also a method to get the analysis names in the _streamProject_;
+* `replicate` is the name of an analysis replicate group and the method to get the replicate names;
+* `blank` is the name of a blank analysis replicate group to be used for blank (blind) subtraction and the method to get blank replicate names;
+* __traces__ corresponds to a raw data signal (e.g., mass trace in MS data) obtained from a given analysis file;  
+* `spectra` for MS analyses is the method to get a table with a spectrum for each cycle time (i.e., retention time
+`r # TODO add chromatograms entry and make method`
+* __EIC__ is an extracted ion chromatogram (retention time vs intensity) and can be obtained with the method `EICs`;
+* __TIC__ is a total ion chromatograms (retention time vs intensity) and can be obtained with the method `TICs`;
+* __XIC__ is a three dimensional ion chromatogram (_m/z_ vs retention time vs intensity) and can be obtained with the method `XICs`;
+* __peak__ corresponds to a cluster of traces (e.g., integrated chromatographic MS peak) within an analysis file and can be accessed with the method `peaks`;
+* __feature__ corresponds to a group of corresponding peaks across analysis files/analyses and can be accessed with the method `features`;
+* __MS2__ corresponds to second order fragmentation data as acquired in tandem MS or MS/MS mode and can be accessed with the method `MS2s`.
+
 (...)
@@
 
-The project creation is demonstrated below for three HRMS files (triplicate) where a set of substances were spiked.
-The project path can be assign as argument but left as default (i.e., the working directory) for this example.
+The project creation is demonstrated below for the three MS files (triplicate) where a set of chemicals were spiked (see \@ref(spiked-chemicals)).
+The project path can be assign as argument but for the example below it was left as default (i.e., the working directory).
 The initial S4 class created in the `newStreamProject` function is _streamProject_.
+The subclass is then defined according to the type of files added.
+Note that multiple file types are not possible as a subclass cannot be defined.
 
 ```{r newStreamProject-call, message=FALSE, results='hide'}
-exS1 <- newStreamProject(files = files[1:3])
+sp1 <- newStreamProject(files = files[1:3], title = "Example_01")
 ```
@@
 
-Because the added files were all with mzML format, the sub-class is directly defined as _msData_, adding additional structure for further processing of MS data.
-Other data formats/structure will lead to different sub-classes (e.g., _uvData_ and _ramanData_) but is not yet implemented within streamFind.
+Because the added files were all with mzML format (i.e., MS files), the subclass is directly defined as _msData_.
+The structure for further processing data is then set for MS data.
+Other data formats/structure will lead to different subclasses (e.g., _uvData_ and _ramanData_) but is not yet implemented within the streamFind package.
 
 ```{r msData_streamProject_classes}
-is(exS1)
+is(sp1)
 ```
@@
 The structure of the _msAnalysis_ class is described in the next sub-chapter.
-
-The `features` `r # TODO make msFeatures after peaks extraction and storage`` 
-which includes the _msFeatures_ S4 class(presented in section X).
+The `features` slot is a _msFeatures_ S4 class object to hold the results from alignment, grouping and annotation of peaks across analyses.
 
@@
 ```{r structure_msData}
-str(exS1, max.level = 2)
+str(sp1, max.level = 2)
 ```
@@
 ```{r msAnalysis_structure}
-exA1 <- getAnalyses(exS1, 1)
-str(exA1, max.level = 2)
+a1 <- getAnalyses(sp1, 1)
+str(a1, max.level = 2)
 ```
@@
 ```{r metadata_structure}
-t(streamFind::metadata(exS1, analyses = 1))
+t(streamFind::metadata(sp1, analyses = 1))
 
 #applied directly to the msAnalysis S4 class
-t(streamFind::metadata(exA1))
+t(streamFind::metadata(a1))
 ```
@@
 
-# Low level data access and plotting
-
-The tools for low level access and plotting (i.e., traces and chromatograms) share a target search approach.
+# Raw data access
+
+The tools for raw data access and plotting (i.e., spectra and chromatograms) share a target search approach.
 The data search is based on pre-defined time and mass (inc.
@@
 
-As aforementioned, access to raw data is based on defining time and _m/z_ targets.
+As aforementioned, access to raw data is based on defining _m/z_ and retention time targets.
 The function `makeTargets` is used for comprehensively build targets.
-As shown below, diverse ways can be used to assemble targets, similarly returning a uniform `data.table`.
-Accessing and plotting S4 methods in streamFind use the `makeTargets` function to collect data.
+As shown below, multiple ways can be used to assemble targets but similarly returning a uniform `data.table` class object with a target list.
+Accessing and plotting S4 methods in streamFind use the `makeTargets` function to collect raw data from the analysis files.
 Therefore, the arguments `mz`, `rt`, `ppm`, `sec` and `id` are present in most S4 methods for accessing and plotting data as demonstrated in the following sections.
@@
 #case 1
-mz01 <- c(247.1651, 239.0628)
-rt01 <- c(839, 937)
-id01 <- c("target1", "target2")
-targets01 <- makeTargets(mz = mz01, rt = rt01, ppm = 20, sec = 60, id = id01)
-
-targets01
+mz1 <- c(247.1651, 239.0628)
+rt1 <- c(839, 937)
+id1 <- c("target1", "target2")
+targets1 <- makeTargets(mz = mz1, rt = rt1, ppm = 20, sec = 60, id = id1)
+
+targets1
 
 #case 2
-mz02 <- data.frame(mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652))
-rt02 <- data.frame(rtmin = c(809, 907), rtmax = c(869, 967))
-targets02 <- makeTargets(mz = mz02, rt = rt02)
-
-targets02
+mz2 <- data.frame(mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652))
+rt2 <- data.frame(rtmin = c(809, 907), rtmax = c(869, 967))
+targets2 <- makeTargets(mz = mz2, rt = rt2)
+
+targets2
 
 #case 3
-mz03 <- data.frame(
-  #id = c("target1", "target2"),
+mz3 <- data.frame(
+  id = c("target1", "target2"),
   mz = c(247.1651, 239.0628),
@@
 )
-targets03 <- makeTargets(mz = mz03, ppm = 20, sec = 60)
-
-targets03
+targets3 <- makeTargets(mz = mz3, ppm = 20, sec = 60)
+
+targets3
 
 #case 4
-mz04 <- data.frame(
+mz4 <- data.frame(
   id = c("target1", "target2"),
@@
 )
-targets04 <- makeTargets(mz = mz04)
-
-targets04
+targets4 <- makeTargets(mz = mz4)
+
+targets4
 ```
@@
 #EIC of targets from first analysis using msAnalysis class
-eic01 <- EICs(exA1, mz = targets04)
-head(eic01, 6)
+eic1 <- EICs(a1, mz = targets4)
+head(eic1, 6)
 
 #EIC of targets from second analyses using msData class
-eic02 <- EICs(exS1, analyses = 2, mz = targets04)
-head(eic02, 6)
+eic2 <- EICs(sp1, analyses = 2, mz = targets4)
+head(eic2, 6)
 ```
@@
 #plot the produced EIC data.table
-plotEICs(eic01)
+plotEICs(eic1)
 ```
@@
 #using the msData class object and targets but with interactive plot
-plotEICs(exS1, analyses = NULL, mz = targets04, colorBy = "analyses", interactive = TRUE)
+plotEICs(sp1, analyses = NULL, mz = targets4, colorBy = "analyses", interactive = TRUE)
 
 #the same plot as the first for data.table but interactive
-plotEICs(exA1, mz = targets04, interactive = TRUE)
+plotEICs(a1, mz = targets4, interactive = TRUE)
 ```
@@
 # TIC for both analyses in the msData class object
-tic01 <- TICs(exS1)
-tic01
+tic1 <- TICs(sp1)
+tic1
 
 #TIC of the msAnalysis object
-head(TICs(exA1), 6)
+head(TICs(a1), 6)
 ```
@@
 #using a data.table as obatined by TICs
-plotTICs(tic01, interactive = TRUE)
+plotTICs(tic1, interactive = TRUE)
 
 #using the msData object
-plotTICs(exS1, colorBy = "replicates")
+plotTICs(sp1, colorBy = "replicates")
 
 #using the msAnalysis
-plotTICs(exA1)
+plotTICs(a1)
 ```
@@
 #using the msData class object
-xic01 <- XICs(exS1, mz = targets04)
-xic01
+xic1 <- XICs(sp1, mz = targets4)
+xic1
 
 #with a msAnalysis class object
-xic02 <- XICs(exA1, mz = mz01, rt = rt01, ppm = 20, sec = 60)
-head(xic02, 3)
+xic2 <- XICs(a1, mz = mz1, rt = rt1, ppm = 20, sec = 60)
+head(xic2, 3)
 ```
@@
 #using the output of XICs S4 method (a data.table)
-plotXICs(xic01, legendNames = c("target number 1", "target number 2"), plotTargetMark = TRUE,
- targetsMark = targets04[, c("mz", "rt")], ppmMark = 5, secMark = 10, numberRows = 2)
+plotXICs(xic1, legendNames = c("target number 1", "target number 2"), plotTargetMark = TRUE,
+ targetsMark = targets4[, c("mz", "rt")], ppmMark = 5, secMark = 10, numberRows = 2)
 
 #using the msData class and plotting the first target from the first analysis
-plotXICs(exS1[1], mz = targets04[1, ])
+plotXICs(sp1[1], mz = targets4[1, ])
 
 #for the msAnalysis, the second target with wide m/z and rt ranges
-plotXICs(exA1, mz = mz01[2], rt = rt01[2], ppm = 500, sec = 120)
+plotXICs(a1, mz = mz1[2], rt = rt1[2], ppm = 500, sec = 120)
 ```
@@
 #load raw data to an msAnalysis object
-exA1 <- loadRawData(exA1)
+a1 <- loadRawData(a1)
 
 #example of accessing the six most intense fragments of the first target with msAnalysis
-s_data <- spectra(exA1)
-s_data <- s_data[preMZ >= targets04$mzmin[1] & preMZ <= targets04$mzmax[1], ]
-s_data <- s_data[rt >= targets04$rtmin[1] & rt <= targets04$rtmax[1], ]
+s_data <- spectra(a1)
+s_data <- s_data[preMZ >= targets4$mzmin[1] & preMZ <= targets4$mzmax[1], ]
+s_data <- s_data[rt >= targets4$rtmin[1] & rt <= targets4$rtmax[1], ]
 s_data <- s_data[order(intensity, decreasing = TRUE), ]
@@
 #load raw data to an msData object.
-exS1 <- loadRawData(exS1, analyses = NULL)
+sp1 <- loadRawData(sp1, analyses = NULL)
 
 #example of accessing the two most intense fragments of the first target with msData
-s_data_2 <- spectra(exS1)
-s_data_2 <- s_data_2[preMZ >= targets04$mzmin[1] & preMZ <= targets04$mzmax[1], ]
-s_data_2 <- s_data_2[rt >= targets04$rtmin[1] & rt <= targets04$rtmax[1], ]
+s_data_2 <- spectra(sp1)
+s_data_2 <- s_data_2[preMZ >= targets4$mzmin[1] & preMZ <= targets4$mzmax[1], ]
+s_data_2 <- s_data_2[rt >= targets4$rtmin[1] & rt <= targets4$rtmax[1], ]
 s_data_2 <- s_data_2[order(intensity, decreasing = TRUE), ]
@@
 
-The basic workflow (Figure \@ref(fig:workflow-scheme)) includes the following steps: (1) project creation, (2) data conversion (when required/applicable), (3) assignment of analysis replicate names and respective blank analysis replicates, (4) peak finding (or peak picking), (5) peak alignment and grouping into features across analyses, (6) filling of peaks missing in features, (7) annotation of isotopic and adduct features and (8) feature quality evaluation and (9) filtering.
+The basic workflow (Figure \@ref(fig:workflow-scheme) left column) includes the following steps: (1) project creation, (2) data conversion (when required/applicable), (3) assignment of analysis replicate names and respective blank analysis replicates, (4) peak finding (or peak picking), (5) peak alignment and grouping into features across analyses, (6) filling of peaks missing in features, (7) annotation of isotopic and adduct features and (8) feature quality evaluation and (9) filtering.
 The usage of the S4 class objects within the basic workflow (pre-processing) for HRMS data is presented in this section with examples.
+The right column in Figure \@ref(fig:workflow-scheme illustrates the implementation of modular functionalities for assembly of data processing workflows.
+The backend for modular assembly and framework of workflows in streamFind is described and demonstrated in section \@ref(workflows).
 
@@
 #access to peaks in msData based on built targets as used for extracting EICs
-peaks(exS2_pp[c(1, 4)], mz = mz01[1], rt = rt01[1], ppm = 10, sec = 30)
+peaks(exS2_pp[c(1, 4)], mz = mz1[1], rt = rt1[1], ppm = 10, sec = 30)
 
 #access to peak in msAnalysis
-peaks(exA2, mz = targets04)
+peaks(exA2, mz = targets4)
 
 #plotting peaks in msData
-plotPeaks(exS2_pp, mz = targets04[2, ], interactive = TRUE, colorBy = "analyses")
+plotPeaks(exS2_pp, mz = targets4[2, ], interactive = TRUE, colorBy = "analyses")
 
 #plotting peaks from msAnalysis
-plotPeaks(exA2, mz = targets04, interactive = FALSE)
+plotPeaks(exA2, mz = targets4, interactive = FALSE)
 
 #map plot for peak time and m/z dimensions for msAnalysis
-mapPeaks(exA2, mz = targets04[1, ], xlim = 30, ylim = 0.001)
+mapPeaks(exA2, mz = targets4[1, ], xlim = 30, ylim = 0.001)
 
 ##map plot for peak time and m/z dimensions for msData
-mapPeaks(exS2_pp, mz = targets04[2, ], colorBy = "replicates", xlim = 30, ylim = 0.001)
+mapPeaks(exS2_pp, mz = targets4[2, ], colorBy = "replicates", xlim = 30, ylim = 0.001)
 ```
@@
 #getter for feature intensities and intensity deviations
-features(exS2_pg, mz = targets04)
+features(exS2_pg, mz = targets4)
 
 #getter for all the feature metadata by setting complete to TRUE
-t(features(exS2_pg_2, mz = targets04[1, ], complete = TRUE))
+t(features(exS2_pg_2, mz = targets4[1, ], complete = TRUE))
 
 #getter for features with intensities for each sample by setting average to FALSE
-features(exS2_pg_2, mz = targets04[1, ], average = FALSE)
+features(exS2_pg_2, mz = targets4[1, ], average = FALSE)
 
 #plotting features
-plotFeatures(exS2_pg, mz = targets04, colorBy = "targets", interactive = FALSE)
+plotFeatures(exS2_pg, mz = targets4, colorBy = "targets", interactive = FALSE)
 
 #plotting features for replicates and interactive
-plotFeatures(exS2_pg, mz = targets04, colorBy = "replicates", interactive = TRUE)
+plotFeatures(exS2_pg, mz = targets4, colorBy = "replicates", interactive = TRUE)
 
 #plot the individual peaks in features and for each analysis shows the time deviation.
-plotFeaturePeaks(exS2_pg, mz = targets04)
+plotFeaturePeaks(exS2_pg, mz = targets4)
 ```
@@
 #head of MS2 data for target 1 in a msAnalysis object
-head(MS2s(exA1, mz = targets04[1, ], settings = settings_msn))
+head(MS2s(a1, mz = targets4[1, ], settings = settings_msn))
 
 #ploting the MS2 data of the first target in a msAnalysis object
-plotMS2s(exA1, mz = targets04[1, ], settings = settings_msn, interactive = FALSE)
+plotMS2s(a1, mz = targets4[1, ], settings = settings_msn, interactive = FALSE)
 
 #most intense MS2 data traces for both targets in a msData object
-s_data_3 <- MS2s(exS1, mz = targets04, settings = settings_msn)
+s_data_3 <- MS2s(sp1, mz = targets4, settings = settings_msn)
 s_data_3 <- s_data_3[order(intensity, decreasing = TRUE), ]
@@
 #ploting the MS2 data of the first target in a msAnalysis object
-plotMS2s(exS1, mz = targets04, settings = settings_msn, interactive = TRUE, colorBy = "targets")
+plotMS2s(sp1, mz = targets4, settings = settings_msn, interactive = TRUE, colorBy = "targets")
 ```
@@
 # 
-# test <- extractEICs(object, mz = targets04)
-# test_xic <- extractXICs(object, mz = targets04)
+# test <- extractEICs(object, mz = targets4)
+# test_xic <- extractXICs(object, mz = targets4)
 
@@
 
-# Instructions
-
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents.
-For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-Diverse cheatsheets can be found in <https://www.rstudio.com/resources/cheatsheets/>.
-This section gives quick instructions to display text, code, tables and figures.
-
-## Text
-
-Text in R markdown has several inline codes for costumization.
-The book in <https://bookdown.org/yihui/rmarkdown/> offers a detailed guideline.
-Moreover, a cheatsheet can be downdloaded with the following link <https://raw.githubusercontent.com/rstudio/cheatsheets/main/rmarkdown.pdf>.
-
-## Code
-
-Code is automaticelly embedded unless `echo` chunk option is set to `FALSE`.
-Evaluation of code in a given chunk can be skiped by setting `eval` to `FALSE`.
-Code chunks should always be unically named.
-
-```{r exCode, echo=FALSE, eval=FALSE, include=FALSE}
-
-#Example of code chunk
-
-list(a = 1:6, b = 1:10)
-
-```
-
-<br>
-
-## Tables
-
-The `knitr::kable` function can be used for displaying tables, see table \@ref(tab:exTable).
-Consult <https://haozhu233.github.io/kableExtra/awesome_table_in_html.html> for other costumization options.
-Table caption is added with the argument `caption` of the `kable` function.
-
-```{r exTable, echo=TRUE, eval=TRUE, include=TRUE}
-
-df <- data.frame(name = c("a", "b", "c", "d", "e", "f"), col1 = 1:6, col2 = 1:6)
-
-knitr::kable(df, caption = "Example of table caption.") %>%
-kable_styling(
-  font_size = 12,
-  bootstrap_options = c("striped", "hover", "condensed", "responsive"),
-  fixed_thead = TRUE
-)
-
-```
-
-<br>
-
-## Figures
-
-Figures/plots can also by embedded using diverse packages.
-The packages `ggplot2` and `plotly` (interactive) are recommended as they are versatile and easy to use.
-Cheatsheet for `ggplot2` can be found in <https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-visualization.pdf>.
-For `plotly` the website <https://plotly.com/r/> offers examples and a cheatsheet can be downdloaded in <https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf>.
-Note that the `echo = FALSE` parameter can be added to the code chunk to prevent printing of the R code that generate the plots.
-
-### Example `ggplot2`
-
-```{r exGgplot, fig.cap="Plot caption example"}
-
-ggplot(pressure, aes(temperature, pressure)) +
-geom_point() +
-theme_bw()
-
-```
-
-### Example `plotly`
-
-```{r exPlotly, fig.cap="Plot caption example"}
-
-plot_ly(pressure, x = ~temperature, y = ~pressure, type = "scatter", mode = "markers+lines")
-
-```
-
-<br> <br>
-
-# Signature
-
-The report was produced on `r Sys.time()`. The data and report are stored in of the workstation *`r Sys.info()["nodename"]`* under the user *`r Sys.info()["user"]`*.
-
-<br>
-
 ***
@@
 For further questions, please contact Ricardo Cunha ([cunha@iuta.de](cunha@iuta.de)).
-
-<style>
-body, p {
-  color: black;
-  font-family: Arial;
-  text-align: justify;
-  font-size: 14pt;
-}
-h1{
-  font-size: 26pt;
-  font-style: bold;
-}
-h2{
-  font-size:20pt;
-  font-style: bold;
-}
-h3{
-  font-size:16pt;
-  font-style: bold;
-}
-.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
-    background-color: #239B56;
-}
-caption {
-  color: black;
-  font-size: 1.0em;
-}
-</style>

### 2022-07-21 handbooks/Handbook_backend_streamFind.Rmd
LFU08 and new MS files and spiked chemicals

@@
 The streamFind package includes example files, which are located  in the folder `system.file(package = "streamFind", dir = "extdata")`.
+The files available are described in Table \@ref(tab:description-files).
 Note, the MS files (.mzML) were trimmed with the function `trimSpectraFilesMZR` to reduce their size.
-The files available are described in Table \@ref(tab:description-files).
+The centroided Agilent Q-TOF files were trimmed with retention time between 900 and 1350 seconds, _m/z_ for MS1 data between 200 and 450 Da and _m/z_ for MS2 between 35 and 450 Da.
+Additionally, MS1 and MS2 traces below 100 and 50 counts were removed, respectively.
+The profile Agilent Q-TOF files were trimmed with retention time between 1000 and 1200 seconds, _m/z_ for MS1 data between 200 and 300 Da and _m/z_ for MS2 between 35 and 300 Da.
+Additionally, MS1 and MS2 traces below 1 count were removed to exclude empty spectra.
+The files are used within the document for demonstration of the backend.
 
@@
 #list of file full paths
-files <- list.files(r_path, full.names = TRUE)
+files <- list.files(r_path, pattern = ".mzML", full.names = TRUE)
 ```
@@
   file_name = basename(files),
-  type = c(rep("MS/MS", 9)),
-  polarity = c(rep("positive", 9)),
+  device = c(rep("Agilent Q-TOF", 24)),
+  type = c(rep("centroided", 3), rep("profile", 3), rep("profile", 18)),
+  mode = c(rep("MS/MS", 24)),
+  polarity = c(rep("positive", 6),
+               rep("negative", 3), rep("positive",3),
+               rep("negative", 3), rep("positive",3),
+               rep("negative", 3), rep("positive",3)),
   description = c(
     rep("Basic centroided MS data spiked with chemicals.", 3),
-    rep("Blank centroided MS data spiked with internal standards.", 3),
-    rep("Wastewater secondary effluent centroided MS data spiked with internal standards.", 3)
+    rep("Basic profile MS data spiked with chemicals.", 3),
+    rep("Blank MS data spiked with internal standards.", 6),
+    rep("Wastewater secondary effluent MS data spiked with internal standards.", 6),
+    rep("Wastewater secondary effluent treated with ozonated strong water MS data spiked with internal standards.", 6)
   )
@@
 
-`r # TODO add a list of spiked chemical in the package` List the chemicals spiked in the MS data files.
+The chemicals spiked in the files are described below in Table \@ref(tab:show-table-chemicals).
+The internal standards (IS) are spiked to all Agilent Q-TOF files.
+The chemical standards (S) were only spiked for the first 6  Agilent Q-TOF files not to the blanks nor wastewater analysis files.
+Note that not all IS and S are visible in the profile Agilent Q-TOF files as a narrower trimming was applied.
+
+```{r load-table-chemicals}
+db <- paste0(r_path, "/spiked_chemicals_hrms.csv")
+db <- data.table::fread(db)
+```
+
+```{r show-table-chemicals, echo=FALSE}
+db$ionization <- "positive"
+db[, `:=`("mz_pos" = neutralMass + 1.0073, "mz_neg" = neutralMass - 1.0073 )]
+db[grepl("neg", db$comment), ionization := "both"]
+data.table::setnames(db, "mixIUTA", "tag")
+db[tag %in% 1, tag := "S"]
+knitr::kable(db[, .(name, PubChem, formula, neutralMass, rt, ionization, tag)], caption = "Chemicals spiked.") %>%
+kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE)
+```
 
@@
 mz1 <- c(247.1651, 239.0628)
-rt1 <- c(839, 937)
+rt1 <- c(1075, 1157)
 id1 <- c("target1", "target2")
@@
 mz2 <- data.frame(mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652))
-rt2 <- data.frame(rtmin = c(809, 907), rtmax = c(869, 967))
+rt2 <- data.frame(rtmin = c(1045, 1127), rtmax = c(1105, 1187))
 targets2 <- makeTargets(mz = mz2, rt = rt2)
@@
   mz = c(247.1651, 239.0628),
-  rt = c(839, 937)
+  rt = c(1075, 1157)
 )
@@
   mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652),
-  rtmin = c(809, 907), rtmax = c(869, 967)
+  rtmin = c(1045, 1127), rtmax = c(1105, 1187)
 )
@@
 ```{r newStreamProject_basic_workflow, message=FALSE, results='hide'}
-exS2 <- newStreamProject(files = files[4:9], path = getwd(), title = "Project Example")
+sp2 <- newStreamProject(files = files[c(10:12, 16:18)], path = getwd(), title = "Project Example")
 ```
@@
 #getter for replicate names in msData
-replicates(exS2)
+replicates(sp2)
 
 #setter for replicate names in msData
-replicates(exS2) <- c(rep("blank", 3),rep("influent", 3))  
-replicates(exS2)
+replicates(sp2) <- c(rep("blank", 3),rep("influent", 3))  
+replicates(sp2)
 
 #setter for replicate name in msAnalysis
-exA2 <- getAnalyses(exS2, 4)
-replicates(exA2) <- "wastewater"
+a2 <- getAnalyses(sp2, 4)
+replicates(a2) <- "wastewater"
 
 #getter for replicate name in msAnalysis
-replicates(exA2)
+replicates(a2)
 ```
@@
 #getter for blank replicates in each sample in a msData object
-blanks(exS2)
+blanks(sp2)
 
 #setter for the blank analysis replicate of msData object
-blanks(exS2) <- rep("blank", 6)
-blanks(exS2)
+blanks(sp2) <- rep("blank", 6)
+blanks(sp2)
 
 #getter for msAnalysis
-blanks(exA2)
+blanks(a2)
 ```
@@
 #getter for analysis names in msdata
-analyses(exS2)
+analyses(sp2)
 
 #getter for analysis name in msAnalysis
-analyses(exA2)
+analyses(a2)
 
 #getter for polarity of analyses in a msData
-polarities(exS2)
+polarities(sp2)
 
 #getter for polarity of an msAnalysis
-polarities(exA2)
+polarities(a2)
 
 #subsetting an msData with `[`-method for the blank samples
-exS2[1:3]
+sp2[1:3]
 ```
@@
 
-exS2_pp <- exS2
+sp2_pp <- sp2
 
@@
 # add the settings to all analyses in msData
-exS2_pp <- addParameters(exS2_pp, settings = settings_pp, where = "analyses")
+sp2_pp <- addParameters(sp2_pp, settings = settings_pp, where = "analyses")
 
 # get the parameters in the third analysis of the msData object
-getParameters(exS2_pp, where = "analyses", analyses = 3)
+getParameters(sp2_pp, where = "analyses", analyses = 3)
 
 # add the settings to the msAnalysis object
-exA2 <- addParameters(exA2, settings = settings_pp)
+a2 <- addParameters(a2, settings = settings_pp)
 
 # get the parameters of the msAnalysis object
-getParameters(exA2, call = "peakPicking")
+getParameters(a2, call = "peakPicking")
 ```
@@
 #peak picking call using the stored parameters
-exS2_pp <- peakPicking(exS2_pp)
+sp2_pp <- peakPicking(sp2_pp)
 
 # peaks added for each analysis, as shown by column peaks of show method (below)
-exS2_pp
+sp2_pp
 
 #alternatively, parameters can be added during the function call
-exS2_pp_2 <- peakPicking(exS2, settings = settings_pp)
+sp2_pp_2 <- peakPicking(sp2, settings = settings_pp)
 
 # parameters are added to each analysis after running peak picking
-summary(getParameters(exS2_pp_2, where = "analyses", analyses = 1))
-
-exS2_pp_2
+summary(getParameters(sp2_pp_2, where = "analyses", analyses = 1))
+
+sp2_pp_2
 
 # running peak picking for an msAnalysis object (settings were already added above)
-exA2 <- peakPicking(exA2)
+a2 <- peakPicking(a2)
 
 # msAnalysis with peaks
-exA2
+a2
 ```
@@
 
-exS2_pp_3 <- exS2
+sp2_pp_3 <- sp2
 
@@
 
-exS2_pp_3 <- addParameters(exS2_pp_3, settings = settings_pp, where = "analyses")
+sp2_pp_3 <- addParameters(sp2_pp_3, settings = settings_pp, where = "analyses")
 
 # adding different settings to analysis 4 to 6
-exS2_pp_3 <- addParameters(exS2_pp_3, settings = settings_pp_2, where = "analyses", analyses = 4:6)
+sp2_pp_3 <- addParameters(sp2_pp_3, settings = settings_pp_2, where = "analyses", analyses = 4:6)
 
 # the algorithm of the influent samples is different
-sapply(analyses(exS2_pp_3), function(x) getAlgorithm(getParameters(exS2_pp_3, where = "analyses", analyses = x)[[1]]))
+sapply(analyses(sp2_pp_3), function(x) getAlgorithm(getParameters(sp2_pp_3, where = "analyses", analyses = x)[[1]]))
 
 #running peak picking
-exS2_pp_3 <- peakPicking(exS2_pp_3)
+sp2_pp_3 <- peakPicking(sp2_pp_3)
 
 # as the peak parameters with "openms" as algorithm were less stringent, the influent samples have more peaks then when "xcms3" is used 
-exS2_pp_3
+sp2_pp_3
 
@@
 #access to peaks in msData based on built targets as used for extracting EICs
-peaks(exS2_pp[c(1, 4)], mz = mz1[1], rt = rt1[1], ppm = 10, sec = 30)
+peaks(sp2_pp[c(1, 4)], mz = mz1[1], rt = rt1[1], ppm = 10, sec = 30)
 
 #access to peak in msAnalysis
-peaks(exA2, mz = targets4)
+peaks(a2, mz = targets4)
 
 #plotting peaks in msData
-plotPeaks(exS2_pp, mz = targets4[2, ], interactive = TRUE, colorBy = "analyses")
+plotPeaks(sp2_pp, mz = targets4[2, ], interactive = TRUE, colorBy = "analyses")
 
 #plotting peaks from msAnalysis
-plotPeaks(exA2, mz = targets4, interactive = FALSE)
+plotPeaks(a2, mz = targets4, interactive = FALSE)
 
 #map plot for peak time and m/z dimensions for msAnalysis
-mapPeaks(exA2, mz = targets4[1, ], xlim = 30, ylim = 0.001)
+mapPeaks(a2, mz = targets4[1, ], xlim = 30, ylim = 0.001)
 
 ##map plot for peak time and m/z dimensions for msData
-mapPeaks(exS2_pp, mz = targets4[2, ], colorBy = "replicates", xlim = 30, ylim = 0.001)
+mapPeaks(sp2_pp, mz = targets4[2, ], colorBy = "replicates", xlim = 30, ylim = 0.001)
 ```
@@
 # add processing parameter for peakGrouping to features in the msData
-exS2_pp <- addParameters(exS2_pp, settings = settings_pg, where = "features")
+sp2_pp <- addParameters(sp2_pp, settings = settings_pg, where = "features")
 
 # get the processing parameters applied to features
-getParameters(exS2_pp, where = "features")
+getParameters(sp2_pp, where = "features")
 ```
@@
 #using the added parameters
-exS2_pg <- peakGrouping(exS2_pp)
+sp2_pg <- peakGrouping(sp2_pp)
 
 #adding the parameters as arguments
-exS2_pg_2 <- peakGrouping(exS2_pp, settings = settings_pg)
+sp2_pg_2 <- peakGrouping(sp2_pp, settings = settings_pg)
 ```
@@
 #accessing directly the msFeatures slot, not recommended
-str(exS2_pg@features, max.level = 2)
+str(sp2_pg@features, max.level = 2)
 ```
@@
 #getter for feature intensities and intensity deviations
-features(exS2_pg, mz = targets4)
+features(sp2_pg, mz = targets4)
 
 #getter for all the feature metadata by setting complete to TRUE
-t(features(exS2_pg_2, mz = targets4[1, ], complete = TRUE))
+t(features(sp2_pg_2, mz = targets4[1, ], complete = TRUE))
 
 #getter for features with intensities for each sample by setting average to FALSE
-features(exS2_pg_2, mz = targets4[1, ], average = FALSE)
+features(sp2_pg_2, mz = targets4[1, ], average = FALSE)
 
 #plotting features
-plotFeatures(exS2_pg, mz = targets4, colorBy = "targets", interactive = FALSE)
+plotFeatures(sp2_pg, mz = targets4, colorBy = "targets", interactive = FALSE)
 
 #plotting features for replicates and interactive
-plotFeatures(exS2_pg, mz = targets4, colorBy = "replicates", interactive = TRUE)
+plotFeatures(sp2_pg, mz = targets4, colorBy = "replicates", interactive = TRUE)
 
 #plot the individual peaks in features and for each analysis shows the time deviation.
-plotFeaturePeaks(exS2_pg, mz = targets4)
+plotFeaturePeaks(sp2_pg, mz = targets4)
 ```
@@
 ```{r plot_alignment}
-plotAlignment(exS2_pg)
+plotAlignment(sp2_pg)
 ```
@@
 #filling peaks for msData
-exS2_pg_fl <- peakFilling(exS2_pg, settings = settings_fl)
+sp2_pg_fl <- peakFilling(sp2_pg, settings = settings_fl)
 ```
@@
 # inspecting the filled peaks for a given mz and rt target
-plotFeaturePeaks(exS2_pg_fl, mz = 441.1670, rt = 916, ppm = 20, sec = 60)
+plotFeaturePeaks(sp2_pg_fl, mz = 441.1670, rt = 916, ppm = 20, sec = 60)
 ```
@@
 
-exS2_pg_fl_an <- peakAnnotation(exS2_pg_fl, settings = settings_an)
+sp2_pg_fl_an <- peakAnnotation(sp2_pg_fl, settings = settings_an)
 ```
@@
 #isotopic and adduct features annotated with a given target mz and rt
-t(streamFind::annotation(exS2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE))
+t(streamFind::annotation(sp2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE))
 
 #plotting the annotation of each feature in a given target
-plotAnnotation(exS2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE)
+plotAnnotation(sp2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE)
 ```
@@
 #sub-setting the msData class on analyses (the 3 first)
-exS2_pg[1:3, ]
+sp2_pg[1:3, ]
 
 #sub-setting the msData class on analys2s and features (the 3 last analyses and features annotated with a target)
-target_s <- streamFind::annotation(exS2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE)
-exS2_pg_fl_an[4:6, target_s$id]
+target_s <- streamFind::annotation(sp2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE)
+sp2_pg_fl_an[4:6, target_s$id]
 
 #sub-seeting the msData class on analyses ([i, ]) and features ([, j]) by index
-exS2_pg_fl_an[1:2, 1:5]
+sp2_pg_fl_an[1:2, 1:5]
 
 #sub-setting an msAnalysis class on the first 5 peaks
-exA2[1:5]
+a2[1:5]
 ```
@@
 
-exS2_pg_fl_an_filFeat <- addParameters(exS2_pg_fl_an, where = "features", settings = settings_filFeat)
-
-exS2_pg_fl_an_filFeat <- filterFeatures(exS2_pg_fl_an_filFeat)
+sp2_pg_fl_an_filFeat <- addParameters(sp2_pg_fl_an, where = "features", settings = settings_filFeat)
+
+sp2_pg_fl_an_filFeat <- filterFeatures(sp2_pg_fl_an_filFeat)
 
 #number of features not filtered
-nrow(features(exS2_pg_fl_an, filtered = FALSE))
+nrow(features(sp2_pg_fl_an, filtered = FALSE))
 
 #number of features not filtered after applying the filters
-nrow(features(exS2_pg_fl_an_filFeat, filtered = FALSE))
+nrow(features(sp2_pg_fl_an_filFeat, filtered = FALSE))
 
 #remove permanently all filtered features from the msData object
-exS2_pg_fl_an_filFeat_rem <- removeFilteredFeatures(exS2_pg_fl_an_filFeat, which =  "all")
-
-exS2_pg_fl_an_filFeat_rem
+sp2_pg_fl_an_filFeat_rem <- removeFilteredFeatures(sp2_pg_fl_an_filFeat, which =  "all")
+
+sp2_pg_fl_an_filFeat_rem
 ```
@@
 
-# exS2_pg <- calculateSNR(exS2_pg, targetsID = features(exS2_pg)[["id"]][1:5])
-# 
-# features(exS2_pg, complete = TRUE)[1:5, ]
+# sp2_pg <- calculateSNR(sp2_pg, targetsID = features(sp2_pg)[["id"]][1:5])
+# 
+# features(sp2_pg, complete = TRUE)[1:5, ]
 
@@
 
-# object <- exS2_pg_fl
+# object <- sp2_pg_fl
 # settings <- annotationSettingsDefaultCAMERA()

### 2022-07-23 handbooks/Handbook_backend_streamFind.Rmd
LFU09 and changed ms file names

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, fig.align = "center", cache = TRUE) # fig.width = 9,
+knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 9, cache = TRUE)
 ```
@@
 
+For installation of the streamFind package, it is recommended to firstly ensure that the dependencies are currently installed.
+Several existing R packages and external functionalities are used in streamFind for various processing steps.
+A major dependency of streamFind is the [patRoon](https://github.com/rickhelmus/patRoon) R package.`r # TODO add reference` The patRoon package combines several functionalities for basic and advanced data processing and can be used interchangeably with streamFind, see \@ref(patroon-interchangeability) for more information.
+Most of the tools that patRoon depends on are equally used in streamFind.
+In the next sub-chapter, the mandatory and optional dependencies are described with installation instructions.
+For a standalone and self-contained usage option, the streamFind docker image is recommended (see \@ref(docker-image)), avoiding the need to installing all required dependencies.
+
 ## Install dependencies
 
+### R and RTools
+
+[R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) are essential dependencies.
+The `r R.version.string` can be obtained in https://cran.r-project.org/.
+The RTools can be downloaded in https://cran.r-project.org/bin/windows/Rtools/.
+Make sure to download the right version for the R version installed.
+Installation instructions are given in both sources.
+
+### Third party software
+
+The external third party dependencies recommended are listed in the section [other dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies) from the patRoon [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites).
+Below we list the mandatory and really recommended software and resources.
+
+* [Java JDK](https://www.oracle.com/java/technologies/downloads/)  
+* [ProteoWizard](https://proteowizard.sourceforge.io/)  
+* [OpenMS](https://www.openms.de/downloads/)  
+* [MetFrag CL](http://ipb-halle.github.io/MetFrag/projects/metfragcl/)
+* [PubChemLite](https://zenodo.org/record/6503754#.YtqxE3bP1hE)
+* [SIRIUS](https://bio.informatik.uni-jena.de/software/sirius/)  
+* [BioTransformer](https://bitbucket.org/djoumbou/biotransformer/downloads/)  
+* [SAFD](https://bitbucket.org/SSamanipour/safd.jl/src/master/)  
+* [Open Babel](http://openbabel.org/wiki/Main_Page)
+
+### R packages
+
+The following code lines are adapted from the patRoon [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites) to ensure that both patRoon and other mutual R dependencies are installed properly.
+
+```{r r-packages, eval=FALSE}
+install.packages(c("BiocManager", "remotes"))
+
+BiocManager::install(c("mzR", "xcms","CAMERA"))
+
+remotes::install_github("blosloos/enviPick")
+
+BiocManager::install("ropls")
+remotes::install_github("rickhelmus/KPIC2")
+
+BiocManager::install("InterpretMSSpectrum")
+remotes::install_github("cbroeckl/RAMClustR")
+
+remotes::install_github("blosloos/nontargetData")
+remotes::install_github("blosloos/nontarget")
+
+remotes::install_github("rickhelmus/cliqueMS")
+
+# only needed for Bruker DataAnalysis integration
+remotes::install_github("BSchamberger/RDCOMClient")
+
+BiocManager::install(c("BiocStyle", "Rgraphviz")) 
+remotes::install_github("KelseyChetnik/MetaClean")
+
+remotes::install_github("rickhelmus/patRoon", upgrade = "never", dependencies = TRUE)
+```
+
+After installation of the patRoon R package, the location of each third party software should be given in an `.Rprofile` file as suggested in section [other dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies) the patRoon [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites).
+For facilitation the same code lines are given below.
+The function `patRoon::verifyDependencies()` from patRoon can be used to verify if all dependencies are found.
+
+```{r options-third-party-software, eval=FALSE}
+options(patRoon.path.pwiz = "C:/ProteoWizard") # location of ProteoWizard installation folder
+options(patRoon.path.SIRIUS = "C:/sirius-win64-3.5.1") # location where SIRIUS was extracted
+options(patRoon.path.OpenMS = "/usr/local/bin") # directory with the OpenMS binaries
+options(patRoon.path.pngquant = "~/pngquant") # directory containing pngquant binary
+options(patRoon.path.MetFragCL = "~/MetFragCommandLine-2.4.8.jar") # full location to the jar file
+options(patRoon.path.MetFragCompTox = "C:/CompTox_17March2019_SelectMetaData.csv") # full location to desired CompTox CSV file
+options(patRoon.path.MetFragPubChemLite = "~/PubChemLite_exposomics_20220429.csv") # full location to desired PubChemLite CSV file
+options(patRoon.path.BioTransformer = "~/biotransformer/biotransformer-3.0.1.jar")
+options(patRoon.path.obabel = "C:/Program Files/OpenBabel-3.0.0") # directory with Open Babel binaries
+```
+
 ## Install streamFind
+
+The streamFind package can be installed from the GitHub repository https://github.com/ricardobachertdacunha/streamFind with the following code line.
+
+```{r install-streamfind, eval=FALSE}
+remotes::install_github("ricardobachertdacunha/streamFind", dependencies = TRUE)
+```
+
+## Docker image
 
@@
 df_files_desc <- data.frame(
+  N. = seq_len(length(files)),
   file_name = basename(files),
@@
 knitr::kable(df_files_desc, caption = "Files included in the streamFind package.") %>%
-kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE)
+  column_spec(column = 2, width = "100px") %>%
+  column_spec(column = 3, width = "60px") %>%
+  kable_styling(font_size = 11.5, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>% #"condensed",
+  scroll_box(width = "100%", height = "600px")
 ```
@@
 db[tag %in% 1, tag := "S"]
-knitr::kable(db[, .(name, PubChem, formula, neutralMass, rt, ionization, tag)], caption = "Chemicals spiked.") %>%
-kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE)
+
+knitr::kable(db[, .(name, CAS, formula, neutralMass, rt, ionization, tag)], caption = "Chemicals spiked.") %>%
+  kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>% #"condensed",
+  scroll_box(width = "100%", height = "600px")
 ```
@@
 
-```{r classes-overview, echo=FALSE, fig.cap="Overview of the S4 classes and their hierarchy with emphasis on MS analyses.", out.width="95%"}
+```{r classes-overview, echo=FALSE, fig.cap="Overview of the S4 classes and their hierarchy with emphasis on MS analyses.", out.width="100%"}
 knitr::include_graphics(paste0(getwd(), "/hb_backend_figures/classes_overview.png"))
@@
 * __traces__ corresponds to a raw data signal (e.g., mass trace in MS data) obtained from a given analysis file;
-* `spectra` for MS analyses is the method to get a table with a spectrum for each cycle time (i.e., retention time
-`r # TODO add chromatograms entry and make method`
+* `spectra` for MS analyses is the method to get a table with a spectrum for each cycle time (i.e., retention time);
 * __EIC__ is an extracted ion chromatogram (retention time vs intensity) and can be obtained with the method `EICs`;
@@
 
-(...)
+(...) `r # TODO add chromatograms entry and make method`
 
@@
 
-```{r newStreamProject-call, message=FALSE, results='hide'}
+```{r newstreamproject-call, message=FALSE, results='hide'}
 sp1 <- newStreamProject(files = files[1:3], title = "Example_01")
@@
 
-```{r msData_streamProject_classes}
+```{r msdata_streamproject_classes}
 is(sp1)
@@
 The `features` slot is a _msFeatures_ S4 class object to hold the results from alignment, grouping and annotation of peaks across analyses.
+Further information for _msAnalysis_ is given in the next sub-chapter (\@ref(msanalysis-class))  and for _msFeatures_ is given in the sub-chapter \@ref(msfeatures-class), after creation of features by performing alignment and grouping of peaks across analyses.
 
@@
 
-```{r structure_msData}
+```{r structure_msdata}
 str(sp1, max.level = 2)
@@
 Note that the S4 method `getAnalyses` for _msData_ was used to obtain the _msAnalysis_ of the first analysis as defined by the second argument.
-In general, access and plotting S4 methods can also be applied directly to the _msAnalysis_ class besides _msData_ (as described in \@ref(low-level-data-access-and-plotting)) but is not part of the basic workflow and mostly used for advanced data visualization and access of low level data (e.g., traces).
+In general, access and plotting S4 methods can also be applied directly to the _msAnalysis_ class besides _msData_ (as described in \@ref(raw-data-access)) but is not part of the basic workflow and mostly used for advanced data visualization and access of low level data (e.g., traces).
 
@@
 
-```{r msAnalysis_structure}
+```{r msanalysis_structure}
 a1 <- getAnalyses(sp1, 1)
@@
 ```
+
+### Create msAnalysis
+
+`r # TODO make method to open a file as xAnalysis``
+
+### Handle metadata
 
@@
 
-```{r EICs_method}
+```{r eics-method}
 #EIC of targets from first analysis using msAnalysis class
@@
 
-```{r plotEICs_static}
+```{r ploteics-static}
 #plot the produced EIC data.table
@@
 
-```{r plotEICs_interactive}
+```{r ploteics-interactive}
 #using the msData class object and targets but with interactive plot
@@
 
-```{r TICs_method}
+```{r tics_method}
 # TIC for both analyses in the msData class object
@@
 
-```{r plotTICs_method}
+```{r plotticss-method}
 #using a data.table as obatined by TICs
@@
 
-```{r XICs_method}
+```{r xics-method}
 #using the msData class object
@@
 
-```{r plotXICs_method}
+```{r plotxics-method}
 #using the output of XICs S4 method (a data.table)
@@
 
-```{r loadRawData}
+```{r loadrawdata}
 #load raw data to an msAnalysis object
@@
 
-```{r newStreamProject_basic_workflow, message=FALSE, results='hide'}
+```{r newstreamproject-basicworkflow, message=FALSE, results='hide'}
 sp2 <- newStreamProject(files = files[c(10:12, 16:18)], path = getwd(), title = "Project Example")
 ```
+
+## Manage analyses
+
+### Add analyses
+
+### Remove analyses
 
@@
 
-```{r assign_replicates}
+```{r assign-replicates}
 #getter for replicate names in msData
@@
 
-```{r assign_blanks}
+```{r assign-blanks}
 #getter for blank replicates in each sample in a msData object
@@
 
-```{r other_S4_methods}
+```{r other-s4methods}
 #getter for analysis names in msdata
@@
 
-```{r create_processing_parameters}
-
+```{r create-processing-parameters}
 sp2_pp <- sp2
@@
 
-```{r peakPicking}
-
+```{r peakpicking-msdata-run, results='hide', message=FALSE}
 #peak picking call using the stored parameters
 sp2_pp <- peakPicking(sp2_pp)
-
+```
+
+```{r peakpicking-msdata-show}
 # peaks added for each analysis, as shown by column peaks of show method (below)
 sp2_pp
-
+```
+
+```{r peakpicking-msdata-run2, results='hide', message=FALSE}
 #alternatively, parameters can be added during the function call
 sp2_pp_2 <- peakPicking(sp2, settings = settings_pp)
-
+```
+
+```{r peakpicking-msdata-show2, }
 # parameters are added to each analysis after running peak picking
@@
 sp2_pp_2
-
+```
+
+```{r peakpicking-msanalysis-run, results='hide', message=FALSE}
 # running peak picking for an msAnalysis object (settings were already added above)
 a2 <- peakPicking(a2)
-
+```
+
+```{r peakpicking-msanalysis-show}
 # msAnalysis with peaks
@@
 
-```{r peakPicking_multiple_settings}
-
+```{r peakpicking-multiple-settings}
 sp2_pp_3 <- sp2
@@
 sapply(analyses(sp2_pp_3), function(x) getAlgorithm(getParameters(sp2_pp_3, where = "analyses", analyses = x)[[1]]))
-
+```
+
+```{r peakpicking-multiple-run, results='hide', message=FALSE}
 #running peak picking
 sp2_pp_3 <- peakPicking(sp2_pp_3)
-
+```
+
+```{r peakpicking-multiple-show}
 # as the peak parameters with "openms" as algorithm were less stringent, the influent samples have more peaks then when "xcms3" is used 
 sp2_pp_3
-
 ```
@@
 
-```{r export_import_settings}
+Export settings:
+
+```{r export-settings}
 
@@
 exportSettings(settings_pp_2, name = "settings_pp_2", format = "json", path = getwd())
-
+```
+
+Import settings:
+
+```{r import-settings}
 #import settings as rds
@@
 
-```{r inspecting_peaks}
+```{r inspecting-peaks}
 #access to peaks in msData based on built targets as used for extracting EICs
@@
 
-```{r creating_grouping_parameters}
+```{r creating-grouping-parameters}
 param_g <- list(
@@
 
-```{r processing_grouping, message=FALSE}
-#using the added parameters
+```{r processing-grouping-run, results='hide', message=FALSE}
+#using the added parameter settings
 sp2_pg <- peakGrouping(sp2_pp)
-
-#adding the parameters as arguments
+```
+
+```{r processing-grouping-show}
+#features added to msData
+sp2_pg
+```
+
+```{r processing_grouping-run2, results='hide', message=FALSE}
+#adding the parameter settings as arguments
 sp2_pg_2 <- peakGrouping(sp2_pp, settings = settings_pg)
+```
+
+```{r processing-grouping-show2}
+#features added to msData
+sp2_pg_2
 ```
@@
 
-```{r msFeatures_class}
+```{r msfeatures-class}
 #accessing directly the msFeatures slot, not recommended
@@
 
-```{r inspecting_features}
+```{r inspecting-features}
 #getter for feature intensities and intensity deviations
@@
 
-```{r plot_alignment}
+```{r plot-alignment}
 plotAlignment(sp2_pg)
@@
 
-```{r recursive_integration_settings}
+```{r recursive-integration-settings}
 #default settings S4 class with ChromPeakAreaParam from xcms as parameter settings
@@
 
-```{r recursive_integration_procesing, results='hide', message=FALSE}
+```{r recursive-integration-run, results='hide', message=FALSE}
 #filling peaks for msData
@@
 
-```{r recursive_integration_inspection}
+```{r recursive-integration-inspection}
 # inspecting the filled peaks for a given mz and rt target
@@
 
-```{r annotation_processing, results='hide', message=FALSE}
+```{r creating-annotation-parameters, results='hide', message=FALSE}
 #creating settings for using CAMERA
@@
 )
-
+```
+
+```{r annotation-run, results='hide', message=FALSE}
 sp2_pg_fl_an <- peakAnnotation(sp2_pg_fl, settings = settings_an)
@@
 
-```{r annotation_results}
+```{r annotation-results}
 #isotopic and adduct features annotated with a given target mz and rt
-t(streamFind::annotation(sp2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE))
+streamFind::annotation(sp2_pg_fl_an, mz = targets4[1, ], ppm = 20, rt = NULL, sec = 60, all = FALSE)
 
 #plotting the annotation of each feature in a given target
-plotAnnotation(sp2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE)
+plotAnnotation(sp2_pg_fl_an, mz = targets4[1, ], ppm = 20, rt = NULL, sec = 60, all = FALSE)
 ```
@@
 
-```{r filter_subsetting}
+```{r filter-subsetting}
 #sub-setting the msData class on analyses (the 3 first)
@@
 #sub-setting the msData class on analys2s and features (the 3 last analyses and features annotated with a target)
-target_s <- streamFind::annotation(sp2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE)
+target_s <- streamFind::annotation(sp2_pg_fl_an, mz = targets4[1, ], ppm = 20, rt = NULL, sec = 5, all = FALSE)
 sp2_pg_fl_an[4:6, target_s$id]
@@
 
-```{r filter_features}
+```{r filter-features}
 
@@
 
-```{r filter_peaks}
+```{r filter-peaks}
 
@@
 
-```{r quality_sn}
+```{r quality-sn}
 
@@
 
-```{r create_settings_msn}
+```{r create-settings-msn}
 #the parameter settings to obtained and process MSn spectra
@@
 
-```{r extract_msn}
+```{r extract-ms2}
 #head of MS2 data for target 1 in a msAnalysis object
@@
 
+## patRoon interchangeability
+
 # Workflows
@@
 
-```{r demoCode, echo=FALSE, eval=FALSE, include=FALSE}
+```{r democode, echo=FALSE, eval=FALSE, include=FALSE}
 

### 2022-07-25 handbooks/Handbook_backend_streamFind.Rmd
LFU10

@@
 
-The project creation is demonstrated below for the three MS files (triplicate) where a set of chemicals were spiked (see \@ref(spiked-chemicals)).
-The project path can be assign as argument but for the example below it was left as default (i.e., the working directory).
+The project creation is demonstrated below for the three MS files (triplicate, 1 to 3 in \@ref(tab:description-files)) where a set of chemicals were spiked (see \@ref(spiked-chemicals) with tag IS and S).
+The project path can be assigned as argument but for the example below it was left as default (i.e., the working directory).
 The initial S4 class created in the `newStreamProject` function is _streamProject_.
@@
 ```{r newstreamproject-call, message=FALSE, results='hide'}
+#New streamProject with the files 1 to 3
 sp1 <- newStreamProject(files = files[1:3], title = "Example_01")
@@
 The structure for further processing data is then set for MS data.
-Other data formats/structure will lead to different subclasses (e.g., _uvData_ and _ramanData_) but is not yet implemented within the streamFind package.
+Other data formats/structure will lead to different sub classes (e.g., _uvData_ and _ramanData_) but is not yet implemented within the streamFind package.
 
@@
 Note that the S4 method `getAnalyses` for _msData_ was used to obtain the _msAnalysis_ of the first analysis as defined by the second argument.
-In general, access and plotting S4 methods can also be applied directly to the _msAnalysis_ class besides _msData_ (as described in \@ref(raw-data-access)) but is not part of the basic workflow and mostly used for advanced data visualization and access of low level data (e.g., traces).
+In general, access to raw data (i.e., traces and chromatograms) as well as respective plotting S4 methods can be applied to both the _msAnalysis_ and _msData_ (as described in \@ref(raw-data-access)).
 
 The slots analysis, file and replicate correspond to the analysis name, the full file path and the replicate group name, respectively.
-The metadata slot is a list of objects with various informative entries of the analysis.
+The metadata slot is a list of objects with informative entries of the analysis file.
 The metadata list can be expanded with additional information about the analysis/sample (e.g., sample location, weather conditions, process sensor data, etc.).
-Below, the `getMetadata` S4 method of _msData_ is used to display the metadata of the first file.
-Note, `getMetadata` also works for _msAnalysis_ S4 class as shown below in the commented code.
-The slot parameters is used to store processing steps (e.g., peak picking) as _settingsSet_ S4 class objects, including the processing step name/ID, the algorithm used and the respective settings applied.
-The parameters can be applied for history track but also for re-run the processing when required.
-The spectra and chromatograms slots are used to store low level data (i.e., mass traces) from the MS files.
+The slot parameters is used to store processing steps (e.g., peak picking) as _settings_ S4 class objects, including the processing step name/ID, the algorithm used and the respective settings applied.
+The parameters can be used for history track but also for re-run the processing steps when required.
+The spectra and chromatograms slots are used to store low level data (i.e., mass traces either in profile or centroided modes) from the MS files.
 Finally, the peaks slot is used to store the chromatographic peaks after peak picking.
-As no other processing step was applied the spectra, chromatograms and peaks tables are empty for now.
+As no other processing step was applied the spectra, chromatograms and peaks tables in the _msAnalysis_ are empty for now.
 
@@
 
-### Create msAnalysis
-
-`r # TODO make method to open a file as xAnalysis``
+### Create an msAnalysis
+
+An _msAnalysis_ object can be created without the need of running `newStreamProject`.
+The method `newAnalysis` can be used for initiating the _msAnalysis_ when a full path of an MS file (i.e., with extension mzML or other MS format) is given.
+Similar to `newStreamProject`, the format of the file given as file argument dictates the class of the analysis object returned from the `newAnalysis` function.
+
+```{r create-msanalysis}
+#creating an msAnalysis object for the 4th file, corresponding to profile MS data
+aProf <- newAnalysis(file = files[4])
+
+is(aProf)
+```
+
+### Add msAnalysis to msData
+
+As the method `getAnalysis` can be used to obtain an _msAnalysis_ from an _msData_ object, the method `addAnalyses` is used to add extra _msAnalysis_ objects to _msData_.
+Below the files 4 to 6 are added to the already created _msData_ (containing files 1 to 3).
+Note that both `getAnalyses` and `addAnalyses` can be applied to equivalent classes to _msData_ (i.e., _uvData_ and _ramanData_)
+
+```{r add-msanalysis}
+
+#all profile MS files
+aProfs <- lapply(files[4:6], function(x) newAnalysis(x))
+
+#adding the profile analyses to msData
+sp1 <- addAnalyses(sp1, analyses = aProfs)
+
+#show added files in msData
+sp1
+```
 
 ### Handle metadata
+
+#### Get metadata
+
+Below, the `getMetadata` S4 method of _msData_ is used to display the metadata of the first file. Note, `getMetadata` also works for _msAnalysis_ S4 class as shown below in the commented code.
 
@@
 
-## Add metadata
+#### Add metadata
 

### 2022-07-29 handbooks/Handbook_backend_streamFind.Rmd
LFU11

@@
 
-* [Java JDK](https://www.oracle.com/java/technologies/downloads/)  
-* [ProteoWizard](https://proteowizard.sourceforge.io/)  
-* [OpenMS](https://www.openms.de/downloads/)  
-* [MetFrag CL](http://ipb-halle.github.io/MetFrag/projects/metfragcl/)
-* [PubChemLite](https://zenodo.org/record/6503754#.YtqxE3bP1hE)
-* [SIRIUS](https://bio.informatik.uni-jena.de/software/sirius/)  
-* [BioTransformer](https://bitbucket.org/djoumbou/biotransformer/downloads/)  
-* [SAFD](https://bitbucket.org/SSamanipour/safd.jl/src/master/)  
-* [Open Babel](http://openbabel.org/wiki/Main_Page)
+* [ProteoWizard](https://proteowizard.sourceforge.io/), required for conversion of vendor MS formats to mzML/mzXML;  
+* [Java JDK](https://www.oracle.com/java/technologies/downloads/), mandatory for patRoon when plotting structure using MetFrag;  
+* [OpenMS](https://www.openms.de/downloads/); esternal tool for peak picking, grouping and alignment;  
+* [MetFrag CL](http://ipb-halle.github.io/MetFrag/projects/metfragcl/), used for formula annotation in patRoon;  
+* [PubChemLite](https://zenodo.org/record/6503754#.YtqxE3bP1hE), database used with MetFrag in patRoon;  
+* [SIRIUS](https://bio.informatik.uni-jena.de/software/sirius/), tool for peak picking and formula annotation in patRoon;    
+* [BioTransformer](https://bitbucket.org/djoumbou/biotransformer/downloads/), used for prediction of transformation products in patRoon;  
+* [SAFD](https://bitbucket.org/SSamanipour/safd.jl/src/master/), tool for peak picking with data in profile mode;  
+* [Open Babel](http://openbabel.org/wiki/Main_Page), used in patRoon for validation of chemicals properties (e.g., InChI and formulae);
 
@@
   description = c(
-    rep("Basic centroided MS data spiked with chemicals.", 3),
-    rep("Basic profile MS data spiked with chemicals.", 3),
-    rep("Blank MS data spiked with internal standards.", 6),
-    rep("Wastewater secondary effluent MS data spiked with internal standards.", 6),
-    rep("Wastewater secondary effluent treated with ozonated strong water MS data spiked with internal standards.", 6)
+    rep("Basic centroided MS data spiked with chemical and internal standards (CS and IS, respectively).", 3),
+    rep("Basic profile MS data spiked with CS and IS.", 3),
+    rep("Blank MS data spiked with IS.", 6),
+    rep("Wastewater secondary effluent MS data spiked with IS.", 6),
+    rep("Wastewater secondary effluent treated with ozonated strong water MS data spiked with IS.", 6)
   )
@@
   column_spec(column = 3, width = "60px") %>%
-  kable_styling(font_size = 11.5, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>% #"condensed",
+  kable_styling(font_size = 11.5, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>%
   scroll_box(width = "100%", height = "600px")
@@
 data.table::setnames(db, "mixIUTA", "tag")
-db[tag %in% 1, tag := "S"]
-
-knitr::kable(db[, .(name, CAS, formula, neutralMass, rt, ionization, tag)], caption = "Chemicals spiked.") %>%
-  kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>% #"condensed",
+db[tag %in% 1, tag := "CS"]
+
+knitr::kable(db[, .(name, CAS, formula, neutralMass, rt, ionization, tag)], caption = "Chemical standards spiked.") %>%
+  kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>%
   scroll_box(width = "100%", height = "600px")
@@
 
-The streamFind package uses S4 class objects (e.g., _streamProject_, _msData_, _msAnalysis_).
+The streamFind package uses S4 class objects (e.g., _streamSet_, _msData_, _msAnalysis_, etc.) for structuring data.
 The S4 classes and their main getter and setter methods are presented and exemplified in this section.
@@
 
-* _streamProject_ refers to the broader S4 class applied in the streamFind package and holds basic project information, see \@ref(streamproject-class);
-* _msData_ refers to the S4 subclass of _streamProject_ when MS analyses (i.e., MS files) are added to the project, see \@ref(msdata-class);
+* _streamSet_ refers to the broader S4 class applied in the streamFind package and holds basic information for an analysis set, see \@ref(streamset-class);
+* _msData_ refers to the S4 subclass of _streamSet_ when MS analyses (i.e., MS files) are added to the set, see \@ref(msdata-class);
 * _msAnalysis_ refers to the S4 class of the object generated for each MS file added to the _msData_ object in the slot "analyses";
-* `projectInfo` is the method to obtain the basic information (i.e., title, date and path) of a _streamProject_ object;
-* `analyses` is the name of the slot holding the analyses (as list) and is also a method to get the analysis names in the _streamProject_;
+* `setInfo` is the method to obtain the basic information (i.e., title, date and path) of a _streamSet_ object;
+* `analyses` is the name of the slot holding the analyses (as list) and is also a method to get the analysis names in the _streamSet_;
 * `replicate` is the name of an analysis replicate group and the method to get the replicate names;
@@
 
-## streamProject class
-
-The project creation is demonstrated below for the three MS files (triplicate, 1 to 3 in \@ref(tab:description-files)) where a set of chemicals were spiked (see \@ref(spiked-chemicals) with tag IS and S).
-The project path can be assigned as argument but for the example below it was left as default (i.e., the working directory).
-The initial S4 class created in the `newStreamProject` function is _streamProject_.
+## streamSet class
+
+The set creation is demonstrated below for the three MS files (triplicate, 1 to 3 in \@ref(tab:description-files)) where a set of chemicals were spiked (see \@ref(spiked-chemicals) with tag IS and S).
+The set path can be assigned as argument but for the example below it was left as default (i.e., the working directory).
+The initial S4 class created in the `newStreamSet` function is _streamSet_.
 The subclass is then defined according to the type of files added.
@@
 
-```{r newstreamproject-call, message=FALSE, results='hide'}
-#New streamProject with the files 1 to 3
-sp1 <- newStreamProject(files = files[1:3], title = "Example_01")
+```{r newstreamset-call, message=FALSE, results='hide'}
+#New streamSet with the files 1 to 3
+sp1 <- newStreamSet(files = files[1:3], title = "Example_01")
 ```
@@
 
-```{r msdata_streamproject_classes}
+```{r msdata_streamset_classes}
 is(sp1)
@@
 
-The _msData_ structure includes the _streamProject_ class plus the `features` slot.
-The project, date and path slots correspond to the project title, creation date and directory in disk, respectively.
+The _msData_ structure includes the _streamSet_ class plus the `features` slot.
+The title, date and path slots correspond to the set title, creation date and directory in disk, respectively.
 The analyses slot is a list with a _msAnalysis_ S4 class object per MS file added.
@@
 Note that the S4 method `getAnalyses` for _msData_ was used to obtain the _msAnalysis_ of the first analysis as defined by the second argument.
-In general, access to raw data (i.e., traces and chromatograms) as well as respective plotting S4 methods can be applied to both the _msAnalysis_ and _msData_ (as described in \@ref(raw-data-access)).
+In general, access to raw data (i.e., traces and chromatograms) as well as respective plotting S4 methods can be applied to both the _msAnalysis_ and _msData_ (as described in \@ref(accessing-raw-data)).
 
@@
 
-An _msAnalysis_ object can be created without the need of running `newStreamProject`.
+An _msAnalysis_ object can be created without the need of running `newStreamSet`.
 The method `newAnalysis` can be used for initiating the _msAnalysis_ when a full path of an MS file (i.e., with extension mzML or other MS format) is given.
-Similar to `newStreamProject`, the format of the file given as file argument dictates the class of the analysis object returned from the `newAnalysis` function.
-
-```{r create-msanalysis}
+Similar to `newStreamSet`, the format of the file given as file argument dictates the class of the analysis object returned from the `newAnalysis` function.
+
+```{r create-msanalysis, message=FALSE, results='hide'}
 #creating an msAnalysis object for the 4th file, corresponding to profile MS data
 aProf <- newAnalysis(file = files[4])
-
-is(aProf)
+```
+
+```{r create-msanalysis-show}
+aProf
 ```
@@
 Below the files 4 to 6 are added to the already created _msData_ (containing files 1 to 3).
-Note that both `getAnalyses` and `addAnalyses` can be applied to equivalent classes to _msData_ (i.e., _uvData_ and _ramanData_)
+Note that both `getAnalyses` and `addAnalyses` can be applied to equivalent classes to _msData_ (i.e., _uvData_ and _ramanData_).
+Also, note that when adding analyses to an _msData_ object with features (i.e., with peaks grouped across analyses), the _msFeatures_ content is cleared as the grouping is invalidated with new _msAnalysis_ objects.
 
@@
 
-Below, the `getMetadata` S4 method of _msData_ is used to display the metadata of the first file. Note, `getMetadata` also works for _msAnalysis_ S4 class as shown below in the commented code.
+Below, the `getMetadata` S4 method of _msData_ is used to display the metadata. Note, `getMetadata` also works for _msAnalysis_ S4 class as shown below.
 
@@
 ```{r metadata_structure}
-t(streamFind::metadata(sp1, analyses = 1))
+#metadata of the fourth analyses in the msData
+t(streamFind::metadata(sp1, analyses = 4))
 
@@
 
-`r # TODO make addMetadata method for analyses`
-
-# Raw data access
+Additional metadata can be added with the `addMetadata` method for both _msAnalysis_ and _msData_.
+Note that for _msData_ the metadata input needs to be a list of vectors with the same length as the number of analyses or a data.frame/data.table with the number of rows as the nmumber of analyses, as shown below.
+
+add metadata to an _msAnalysis_ object (a1):
+
+```{r add-metadata-msanalysis}
+#as a named vector
+meta_vec <- c("water_as_vec", TRUE)
+names(meta_vec) <- c("matrix", "spiked")
+
+a1 <- addMetadata(a1, metadata = meta_vec)
+
+metadata(a1, which = names(meta_vec))
+
+#as data.frame (would also work with data.table)
+meta_df <- data.frame(matrix = "water_as_df", spiked = TRUE)
+
+#add overwrite to TRUE, as the metadata name is already in the msAnalysis
+a1 <- addMetadata(a1, metadata = meta_df, overwrite = TRUE)
+
+metadata(a1, which = names(meta_vec))
+```
+
+add metadata to an _msData_ object (sp1):
+
+```{r}
+
+#as a named vector
+meta_1 <- c("water", TRUE, "centroid")
+names(meta_1) <- c("matrix", "spiked", "datatype")
+
+meta_2 <- c("water", TRUE, "profile")
+names(meta_2) <- c("matrix", "spiked", "datatype")
+
+meta_vec_list <- c(rep(list(meta_1), 3), rep(list(meta_2), 3))
+
+#optional, but when numbered list the order is taken as is
+names(meta_vec_list) <- analyses(sp1)
+
+sp1 <- addMetadata(sp1, metadata = meta_vec_list)
+
+metadata(sp1, which = names(meta_1))
+
+#as a data.frame (would also work with data.table)
+meta_df_2 <- data.frame(
+  analysis = analyses(sp1),
+  matrix = rep("water_as_df", 6),
+  spiked = rep(TRUE, 6),
+  datatype = c(rep("centroid", 3), rep("profile", 3))
+)
+
+# Note, the analysis column doesn't need to be added
+# but the order of the rows will be used as is.
+# when names of the analyses are in analysis column
+# that order is taken to match the analysis in the msData object.
+
+#add overwrite to TRUE, as the metadata name is already in the analyses of msData
+sp1 <- addMetadata(sp1, metadata = meta_df_2, overwrite = TRUE)
+
+metadata(sp1, which = colnames(meta_df_2))
+```
+
+# Accessing raw data
 
@@
 head(eic1, 6)
-
+```
+
+```{r eics-method-2}
 #EIC of targets from second analyses using msData class
@@
 #using the msData class object and targets but with interactive plot
-plotEICs(sp1, analyses = NULL, mz = targets4, colorBy = "analyses", interactive = TRUE)
-
+plotEICs(sp1, analyses = 1:3, mz = targets4, colorBy = "analyses", interactive = TRUE)
+```
+
+```{r ploteics-interactive-2}
 #the same plot as the first for data.table but interactive
@@
 # TIC for both analyses in the msData class object
-tic1 <- TICs(sp1)
+tic1 <- TICs(sp1, analyses = 1:3)
 tic1
-
+```
+
+```{r tics_method-2}
 #TIC of the msAnalysis object
@@
 plotTICs(tic1, interactive = TRUE)
-
+```
+
+```{r plotticss-method-2}
 #using the msData object
-plotTICs(sp1, colorBy = "replicates")
-
+plotTICs(sp1, analyses = 1:3, colorBy = "replicates")
+```
+
+```{r plotticss-method-3}
 #using the msAnalysis
@@
 ```{r xics-method}
-#using the msData class object
-xic1 <- XICs(sp1, mz = targets4)
+#using the msData class object for second target with wide ppm range
+xic1 <- XICs(sp1, analyses = 3:4, mz = mz1[2], rt = rt1[2], ppm = 50, sec = 30)
 xic1
-
-#with a msAnalysis class object
-xic2 <- XICs(a1, mz = mz1, rt = rt1, ppm = 20, sec = 60)
+```
+
+```{r xics-method-2}
+#with a msAnalysis class object for targets in data.frame
+xic2 <- XICs(a1, mz = targets4)
 head(xic2, 3)
@@
 ```{r plotxics-method}
-#using the output of XICs S4 method (a data.table)
+#using the output of XICs S4 method (a data.table) for plotting centroided and profile data
 plotXICs(xic1, legendNames = c("target number 1", "target number 2"), plotTargetMark = TRUE,
  targetsMark = targets4[, c("mz", "rt")], ppmMark = 5, secMark = 10, numberRows = 2)
-
+```
+
+```{r plotxics-method-2}
 #using the msData class and plotting the first target from the first analysis
 plotXICs(sp1[1], mz = targets4[1, ])
-
+```
+
+```{r plotxics-method-3}
 #for the msAnalysis, the second target with wide m/z and rt ranges
@@
 
-```{r loadrawdata}
+```{r loadrawdata-msanalysis}
 #load raw data to an msAnalysis object
@@
 s_data[1:6, ]
-
-#load raw data to an msData object.
-sp1 <- loadRawData(sp1, analyses = NULL)
+```
+
+```{r loadrawdata-msdata}
+#load raw data to the first three samples in an msData object.
+sp1 <- loadRawData(sp1, analyses = 1:3)
 
@@
 
+# Converting raw data
+
+Regarding MS data, the format used in streamFind is .mzML or .mzXML either centroided or profile.
+However, both mentioned formats do not correspond to the formats of actual MS system vendors.
+[msConvert](https://proteowizard.sourceforge.io/download.html) from [ProteoWizard](https://proteowizard.sourceforge.io/) can be used to convert the main MS formats (and others) to .mzML or .mzXML, including data centroiding when required.
+The msConvert GUI tool can be used for conversion.
+However, if automation is required, the [command line](https://proteowizard.sourceforge.io/tools/msconvert.html) of msConvert is used within the function `convertFiles` in streamFind, as shown below.
+The function `compatibleFileFormatsForConversion()` can be used to check which vendor formats are currently possible to convert in streamFind through ProteoWizard.
+
+```{r compatible-vendor-formats, echo=FALSE}
+vfor <- streamFind::compatibleFileFormatsForConversion()
+knitr::kable(vfor, caption = "Vendor formats possible to be converted to mzML/mzXML.") %>%
+  kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>%
+  scroll_box(width = "100%", height = "600px")
+```
+
+```{r converting-raw-data, eval = FALSE}
+##Not run
+
+#vector with full paths to vendor files
+file <- full_path_to_vendor_file/s
+
+#option to perform centroiding of MS1 and MS2 data
+opts_list <- list(filter = "peakPicking vendor msLevel=1-2")
+
+#convert the file
+convertFiles(file, type = "ms", outfile = "mzML", outdir = NULL, opts_list = opts_list)
+
+#Note, when outdir is NULL the directory of the original file will be used for the new file.
+```
+
 # Basic workflow
 
-The basic workflow (Figure \@ref(fig:workflow-scheme) left column) includes the following steps: (1) project creation, (2) data conversion (when required/applicable), (3) assignment of analysis replicate names and respective blank analysis replicates, (4) peak finding (or peak picking), (5) peak alignment and grouping into features across analyses, (6) filling of peaks missing in features, (7) annotation of isotopic and adduct features and (8) feature quality evaluation and (9) filtering.
+The basic workflow (Figure \@ref(fig:workflow-scheme) left column) includes the following steps: (1) set creation, (2) data conversion (when required/applicable), (3) assignment of analysis replicate names and respective blank analysis replicates, (4) peak finding (or peak picking), (5) peak alignment and grouping into features across analyses, (6) filling of peaks missing in features, (7) annotation of isotopic and adduct features and (8) feature quality evaluation and (9) filtering.
 The usage of the S4 class objects within the basic workflow (pre-processing) for HRMS data is presented in this section with examples.
@@
 
-## Project creation
-
-The project creation is demonstrated below for six HRMS data files corresponding to blank and wastewater influent samples, both measured in triplicate.
+## Creation of analysis set
+
+The analysis set creation is demonstrated below for six HRMS data files corresponding to blank and wastewater influent samples, both measured in triplicate.
 The argument `files` is the main input and can either be a list of full path files, a table with file (i.e., full file path), replicate (i.e., the name of the analysis replicate group for each file) and blank (i.e., the name of the associated blank analysis replicate group for each file), or the `analysisInfo` data.frame from the package [patRoon](https://github.com/rickhelmus/patRoon).
-The project path can be assign as argument `path` and is the directory where all the project files (i.e., scripts, cache databases, objects and results) will be stored.
-
-```{r newstreamproject-basicworkflow, message=FALSE, results='hide'}
-sp2 <- newStreamProject(files = files[c(10:12, 16:18)], path = getwd(), title = "Project Example")
+The set path can be assign as argument `path` and is the directory where all the set files (i.e., scripts, cache databases, objects and results) will be stored by default.
+
+```{r newstreamset-basicworkflow, message=FALSE, results='hide'}
+sp2 <- newStreamSet(files = files[c(10:12, 16:18)], path = getwd(), title = "Project Example")
+```
+
+```{r newstreamset-object-show}
+sp2
 ```
@@
 
-### Add analyses
-
-### Remove analyses
-
-## Assign replicate names
+Before preforming data processing, the analysis names (i.e., replicate name and associated blank replicate name) and metadata can be amended/extended.
+
+### Assign replicate names
 
@@
 
-## Assign blanks
+### Assign blanks
 
@@
 Although not shown, different blank analysis replicates can be assigned to different analysis.
-The subtraction is then applied per analysis considering the assigned blank replicate.
+The blank subtraction (\@ref(filter-functions)) is then applied per analysis replicate considering the assigned blank replicate.
 Getting and assigning blank replicates from an _msAnalysis_ is also possible.
-However, assigning a blank without the context of a _streamProject_ and without the structure of _msData_ or similar level classes is not recommended.
-For cross project analysis, a subset of the analyses (_msAnalysis_ objects) in a given _msData_ can be concatenated to another _msData_, carrying the assigned blank replicate for each analysis as well as all the other analysis information (see section cross-project analysis for more information).
+However, assigning a blank without the context of a _streamSet_ and without the structure of _msData_ or similar level classes is not recommended.
+For cross project analysis, a subset of the analyses (_msAnalysis_ objects) in a given _msData_ can be concatenated to another _msData_, carrying the assigned blank replicate for each analysis as well as all the other analysis information (see `r # TODO make section cross project analysis``section cross project analysis for more information).
 
@@
 
-## Other getter S4 methods
+### Other getter S4 methods
 
@@
 
-#subsetting an msData with `[`-method for the blank samples
+#sub-setting an msData with `[`-method for the blank analyses
 sp2[1:3]
@@
 summary(getParameters(sp2_pp_2, where = "analyses", analyses = 1))
-
+```
+
+```{r peakpicking-msdata-show-2}
 sp2_pp_2
@@
 
-The _settings_ object can be exported as a JSON or rds file and then imported for other projects where the same parameter settings are to be applied.
+The _settings_ object can be exported as a JSON or rds file and then imported for other sets where the same parameter settings are to be applied.
 The methods `exportSettings` and `importSettings` are used, as shown below.
@@
 
-```{r inspecting-peaks}
+For a _msData_ object:
+
+```{r inspecting-peaks-msdata, eval=FALSE}
 #access to peaks in msData based on built targets as used for extracting EICs
 peaks(sp2_pp[c(1, 4)], mz = mz1[1], rt = rt1[1], ppm = 10, sec = 30)
-
+```
+
+```{r inspecting-peaks-msdata-table}
+#access to peaks in msData based on built targets as used for extracting EICs
+pks_1 <- peaks(sp2_pp[c(1, 4)], mz = mz1[1], rt = rt1[1], ppm = 10, sec = 30)
+
+knitr::kable(pks_1, caption = "Peaks data.table from msData using the peaks S4 method.") %>%
+  kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>%
+  scroll_box(width = "100%", height = "600px")
+```
+
+```{r inspecting-peaks-msdata-plot}
+#plotting peaks in msData
+plotPeaks(sp2_pp, mz = targets4[2, ], interactive = TRUE, colorBy = "analyses")
+```
+
+```{r inspecting-peaks-msdata-map}
+#map plot for peak time and m/z dimensions for msData
+mapPeaks(sp2_pp, mz = targets4[2, ], colorBy = "replicates", xlim = 30, ylim = 0.001)
+```
+
+For a _msAnalysis_ object:
+
+```{r inspecting-peaks-msAnalysis}
 #access to peak in msAnalysis
 peaks(a2, mz = targets4)
-
-#plotting peaks in msData
-plotPeaks(sp2_pp, mz = targets4[2, ], interactive = TRUE, colorBy = "analyses")
-
+```
+
+```{r inspecting-peaks-msAnalysis-plot}
 #plotting peaks from msAnalysis
 plotPeaks(a2, mz = targets4, interactive = FALSE)
-
+```
+
+```{r inspecting-peaks-msAnalysis-map}
 #map plot for peak time and m/z dimensions for msAnalysis
 mapPeaks(a2, mz = targets4[1, ], xlim = 30, ylim = 0.001)
-
-##map plot for peak time and m/z dimensions for msData
-mapPeaks(sp2_pp, mz = targets4[2, ], colorBy = "replicates", xlim = 30, ylim = 0.001)
 ```
@@
 
-#sub-setting the msData class on analys2s and features (the 3 last analyses and features annotated with a target)
+#sub-setting the msData class on analyses and features (the 3 last analyses and features annotated with a target)
 target_s <- streamFind::annotation(sp2_pg_fl_an, mz = targets4[1, ], ppm = 20, rt = NULL, sec = 5, all = FALSE)
@@
 
-# projectInfo(msd)
+# setInfo(msd)
 # path(msd)

### 2022-08-04 handbooks/Handbook_backend_streamFind.Rmd
LFU12

@@
   N. = seq_len(length(files)),
+  file_type = rep("MS", 6 + 18),
   file_name = basename(files),
   device = c(rep("Agilent Q-TOF", 24)),
-  type = c(rep("centroided", 3), rep("profile", 3), rep("profile", 18)),
-  mode = c(rep("MS/MS", 24)),
+  data_type = c(rep("centroid", 3), rep("profile", 3), rep("centroid", 18)),
+  ac_mode = c(rep("MS/MS", 24)),
   polarity = c(rep("positive", 6),
@@
 knitr::kable(df_files_desc, caption = "Files included in the streamFind package.") %>%
-  column_spec(column = 2, width = "100px") %>%
-  column_spec(column = 3, width = "60px") %>%
+  column_spec(column = 3, width = "100px") %>%
+  column_spec(column = 4, width = "60px") %>%
   kable_styling(font_size = 11.5, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>%
@@
 db[tag %in% 1, tag := "CS"]
-
-knitr::kable(db[, .(name, CAS, formula, neutralMass, rt, ionization, tag)], caption = "Chemical standards spiked.") %>%
+db[, in_file := "1-6"]
+db[tag %in% "IS", in_file := "1-24"]
+
+knitr::kable(db[, .(name, CAS, formula, neutralMass, rt, ionization, tag, in_file)], caption = "Chemical standards spiked.") %>%
   kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>%
@@
 #adding the profile analyses to msData
-sp1 <- addAnalyses(sp1, analyses = aProfs)
+sp1 <- addAnalyses(sp1, analysisList = aProfs)
 
@@
 #metadata of the fourth analyses in the msData
-t(streamFind::metadata(sp1, analyses = 4))
+t(metadata(sp1, analyses = 4))
 
 #applied directly to the msAnalysis S4 class
-t(streamFind::metadata(a1))
+t(metadata(a1))
 ```
@@
 ```{r}
-
 #as a named vector
@@
 #optional, but when numbered list the order is taken as is
-names(meta_vec_list) <- analyses(sp1)
+names(meta_vec_list) <- analysisNames(sp1)
 
@@
 meta_df_2 <- data.frame(
-  analysis = analyses(sp1),
+  analysis = analysisNames(sp1),
   matrix = rep("water_as_df", 6),
@@
 The function `makeTargets` is used for comprehensively build targets.
-As shown below, multiple ways can be used to assemble targets but similarly returning a uniform `data.table` class object with a target list.
+As shown below for Carbamazepin d10 and Diuron d6, multiple ways can be used to assemble targets but similarly returning a uniform `data.table` class object with a target list.
 Accessing and plotting S4 methods in streamFind use the `makeTargets` function to collect raw data from the analysis files.
@@
 
+```{r carb-diu}
+carb <- db[name %in% "Carbamazepin d10", .(name, neutralMass, rt)]
+carb$mz <- carb$neutralMass + 1.0073
+carb
+
+diu <- db[name %in% "Diuron d6", .(name, neutralMass, rt)]
+diu$mz <- diu$neutralMass + 1.0073
+diu
+```
+
 ```{r make_targets}
 #case 1
-mz1 <- c(247.1651, 239.0628)
+mz1 <- c(247.1650, 239.0620)
 rt1 <- c(1075, 1157)
 id1 <- c("target1", "target2")
-targets1 <- makeTargets(mz = mz1, rt = rt1, ppm = 20, sec = 60, id = id1)
+targets1 <- makeTargets(mz = mz1, rt = rt1, ppm = 20, sec = 30, id = id1)
 
@@
 #case 2
-mz2 <- data.frame(mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652))
+mz2 <- data.frame(mzmin = c(247.1601, 239.0572), mzmax = c(247.1699, 239.0668))
 rt2 <- data.frame(rtmin = c(1045, 1127), rtmax = c(1105, 1187))
@@
   id = c("target1", "target2"),
-  mz = c(247.1651, 239.0628),
+  mz = c(247.1650, 239.0620),
   rt = c(1075, 1157)
 )
-targets3 <- makeTargets(mz = mz3, ppm = 20, sec = 60)
+targets3 <- makeTargets(mz = mz3, ppm = 20, sec = 30)
 
@@
   id = c("target1", "target2"),
-  mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652),
+  mzmin = c(247.1601, 239.0572), mzmax = c(247.1699, 239.0668),
   rtmin = c(1045, 1127), rtmax = c(1105, 1187)
@@
 
+```{r test-targets}
+#test all equal target 1
+t1 <- rbind(targets1[1, ], targets2[1,], targets3[1, ], targets4[1, ])
+all(round(apply(t1[, 2:7], 2, sd), digits = 4) == 0)
+
+#test all equal target 2
+t2 <- rbind(targets1[2, ], targets2[2,], targets3[2, ], targets4[2, ])
+all(round(apply(t2[, 2:7], 2, sd), digits = 4) == 0)
+```
+
 ## EICs
-
-`r # TODO remove zeros when returning EIC table?`
-`r # TODO add check for spectra/chroms is the object before querying the data?`
 
@@
 ```{r eics-method-2}
-#EIC of targets from second analyses using msData class
-eic2 <- EICs(sp1, analyses = 2, mz = targets4)
+#EIC of targets from analyses 1 to 3 using msData class
+eic2 <- EICs(sp1, analyses = 1:3, mz = targets4)
 head(eic2, 6)
@@
 ```{r ploteics-interactive-2}
-#the same plot as the first for data.table but interactive
+#using msAnalysis
 plotEICs(a1, mz = targets4, interactive = TRUE)
+```
+
+```{r ploteics-interactive-3}
+#using msData centroid (file 3) vs profile (file 4); note that the higher intensity for profile data is due to the sum of the traces within the same spectrum (i.e., with the same retention time)
+plotEICs(sp1, analyses = 3:4, mz = targets3, colorBy = "replicates", interactive = TRUE)
 ```
@@
 #using the msData class object for second target with wide ppm range
-xic1 <- XICs(sp1, analyses = 3:4, mz = mz1[2], rt = rt1[2], ppm = 50, sec = 30)
+xic1 <- XICs(sp1, analyses = 3:4, mz = mz1[2], rt = rt1[2], ppm = 500, sec = 30)
 xic1
@@
 #for the msAnalysis, the second target with wide m/z and rt ranges
-plotXICs(a1, mz = mz1[2], rt = rt1[2], ppm = 500, sec = 120)
+plotXICs(a1, mz = mz1[2], rt = rt1[2], ppm = 50, sec = 120)
 ```
@@
 
-The S4 method applied above query the data from the raw data files as that is less disk memory intense.
-However, for certain cases the raw data might be needed entirely.
-Therefore, the data can be loaded and stored as a simple _data.table_ in the respective _msAnalysis_ (or other classes of the same level).
-Below we demonstrate how to load (via `loadRawData` S4 method) and access (via `spectra` or `chromatograms`) the raw spectra and chromatograms to/from an _msAnalaysis_ and _msData_ object.
+The S4 methods applied above parse the data from the raw data files which is less disk memory intense but slower.
+Alternatively, the raw data can be loaded to (each) _msAnalysis_  and from their is faster available.
+However, when loaded, the raw data has a higher memory footprint.
+Therefore, the data can be optionally loaded and stored as a simple _data.table_ in the respective _msAnalysis_ (other classes of the same level might have the raw data directly loaded if the data has a lower memory footprint).
+Below we demonstrate how to load (via `loadRawData` S4 method) and access (via `spectra` or `chromatograms`) the raw spectra and chromatograms from an _msAnalaysis_ and _msData_ object.
 Both S4 methods (`spectra` or `chromatograms`) return a _data.table_ where further operations can be applied to subset the data, as demonstrated below.
+The standard _data.table_ operations can be applied, see documentation in https://cran.r-project.org/web/packages/data.table/ for more information and code examples/instructions.
+A cheat sheet for _data.table_ operations can be downloaded in https://raw.githubusercontent.com/rstudio/cheatsheets/main/datatable.pdf.
+In streamFind, the _data.table_ is always the preferred format for table objects.
 
 ```{r loadrawdata-msanalysis}
-#load raw data to an msAnalysis object
-a1 <- loadRawData(a1)
-
+#load raw data to an msAnalysis object, while setting a minimum intensity value for both MS1 and MS2 traces
+a1 <- loadRawData(a1, minIntensityMS1 = 200, minIntensityMS2 = 100)
+```
+
+```{r loadrawdata-msanalysis-show}
+a1
+```
+
+```{r loadrawdata-msdata}
+#load raw data to the first three samples in an msData object.
+sp1 <- loadRawData(sp1, analyses = 1:3)
+```
+
+```{r loadrawdata-msdata-show}
+sp1
+```
+
+## Spectra
+
+```{r spectra-msanalysis}
 #example of accessing the six most intense fragments of the first target with msAnalysis
@@
 
-```{r loadrawdata-msdata}
-#load raw data to the first three samples in an msData object.
-sp1 <- loadRawData(sp1, analyses = 1:3)
-
+```{r spectra-msdata}
 #example of accessing the two most intense fragments of the first target with msData
@@
 
+## Chromatograms
+
 `r # TODO create method to access chromatograms in msAnalysis and msData`
@@
   kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>%
-  scroll_box(width = "100%", height = "600px")
+  scroll_box(width = "100%", height = "300px")
 ```
@@
 An essential aspect of environmental analysis is to operate with sample/analysis replicates.
-The replicates in _msData_ or _msAnalysis_ can be obtained by the `replicates` S4 method as shown below.
-The setter (assignment) for the replicate names is demonstrated below with the method `replicates<-` that takes a vector of character strings with the same length as the number of analyses in a _msData_ object or length one for a single analysis in _msAnalysis_.
+The replicates in _msData_ or _msAnalysis_ can be obtained by the `replicateNames` S4 method as shown below.
+The setter (assignment) for the replicate names is demonstrated below with the method `replicateNames<-` that takes a vector of character strings with the same length as the number of analyses in a _msData_ object or length one for a single analysis in _msAnalysis_.
 Alternatively, a table with a replicate character column or a character vector can be given in the files or replicates arguments, respectively.
 
-```{r assign-replicates}
+```{r assign-replicatenames}
 #getter for replicate names in msData
-replicates(sp2)
+replicateNames(sp2)
 
 #setter for replicate names in msData
-replicates(sp2) <- c(rep("blank", 3),rep("influent", 3))  
-replicates(sp2)
+replicateNames(sp2) <- c(rep("blank", 3),rep("influent", 3))  
+replicateNames(sp2)
 
@@
 a2 <- getAnalyses(sp2, 4)
-replicates(a2) <- "wastewater"
+replicateNames(a2) <- "wastewater"
 
 #getter for replicate name in msAnalysis
-replicates(a2)
-```
-
-### Assign blanks
+replicateNames(a2)
+```
+
+### Assign blankReplicateNames
 
@@
 
-```{r assign-blanks}
+```{r assign-blankreplicatenames}
 #getter for blank replicates in each sample in a msData object
-blanks(sp2)
+blankReplicateNames(sp2)
 
 #setter for the blank analysis replicate of msData object
-blanks(sp2) <- rep("blank", 6)
-blanks(sp2)
+blankReplicateNames(sp2) <- rep("blank", 6)
+blankReplicateNames(sp2)
 
 #getter for msAnalysis
-blanks(a2)
+blankReplicateNames(a2)
 ```
@@
 #getter for analysis names in msdata
-analyses(sp2)
+analysisNames(sp2)
 
 #getter for analysis name in msAnalysis
-analyses(a2)
+analysisNames(a2)
 
@@
 param <- xcms::CentWaveParam(
-  ppm = 10, peakwidth = c(5, 60),
+  ppm = 12, peakwidth = c(5, 40),
   snthresh = 5, prefilter = c(4, 800),
@@
 # the algorithm of the influent samples is different
-sapply(analyses(sp2_pp_3), function(x) getAlgorithm(getParameters(sp2_pp_3, where = "analyses", analyses = x)[[1]]))
+sapply(analysisNames(sp2_pp_3), function(x) getAlgorithm(getParameters(sp2_pp_3, where = "analyses", analyses = x)[[1]]))
 ```
@@
 ```{r peakpicking-multiple-show}
-# as the peak parameters with "openms" as algorithm were less stringent, the influent samples have more peaks then when "xcms3" is used 
+# as the peak parameters with "openms" as algorithm were more stringent, the influent analyses have less peaks when compared to "xcms3" 
 sp2_pp_3
@@
 #access to peaks in msData based on built targets as used for extracting EICs
-pks_1 <- peaks(sp2_pp[c(1, 4)], mz = mz1[1], rt = rt1[1], ppm = 10, sec = 30)
-
-knitr::kable(pks_1, caption = "Peaks data.table from msData using the peaks S4 method.") %>%
-  kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>%
-  scroll_box(width = "100%", height = "600px")
+peaks(sp2_pp[c(1, 4)], mz = mz1[1], rt = rt1[1], ppm = 10, sec = 30)
 ```
@@
 
+# knitr::kable(pks_1, caption = "Peaks data.table from msData using the peaks S4 method.") %>%
+#   kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>%
+#   scroll_box(width = "100%", height = "600px")
+
 # object <- sp2_pg_fl
@@
 # path(msd)
-# files(msd)
-# analyses(msd)
-# replicates(msd)
-# blanks(msd)
+# filePaths(msd)
+# analysisNames(msd)
+# replicateNames(msd)
+# blankReplicateNames(msd)
 # polarities(msd)
@@
 # 
-# replicates(msd) <- c("s1", "s2")
-# blanks(msd) <- c(NA_character_, "s2")
+# replicateNames(msd) <- c("s1", "s2")
+# blankReplicateNames(msd) <- c(NA_character_, "s2")
 # 
@@
 # 
-# files(msa_1)
-# analyses(msa_1)
-# replicates(msa_1)
-# blanks(msa_1)
+# filePaths(msa_1)
+# analysisNames(msa_1)
+# replicateNames(msa_1)
+# blankReplicateNames(msa_1)
 # polarities(msa_1)
 # 
-# replicates(msa_1) <- "s1"
-# blanks(msa_1) <- "s2"
+# replicateNames(msa_1) <- "s1"
+# blankReplicateNames(msa_1) <- "s2"
 
@@
 #   
-#   test <- RaMS::grabMSdata(files(object)[1], grab_what = "MS1", rtrange = c(min(rt_02), max(rt_02)))
+#   test <- RaMS::grabMSdata(filePaths(object)[1], grab_what = "MS1", rtrange = c(min(rt_02), max(rt_02)))
 #   

### 2022-08-04 R/RcppExports.R

@@ -1,0 +1,15 @@
+# Generated by using Rcpp::compileAttributes() -> do not edit by hand
+# Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
+
+rcpp_hello_world <- function() {
+    .Call(`_streamFind_rcpp_hello_world`)
+}
+
+rcpp_extract_eics <- function(spec, targets) {
+    .Call(`_streamFind_rcpp_extract_eics`, spec, targets)
+}
+
+rcpp_extract_xics <- function(spec, targets) {
+    .Call(`_streamFind_rcpp_extract_xics`, spec, targets)
+}
+

### 2022-08-04 R/fct_ms_007_filterFeatures.R

@@ -282,7 +282,7 @@ minReplicateAbundanceFeatures <- function(obj, value = 3) {
   feats_org <- copy(obj@metadata)
 
   rpl <- replicates(obj)
-  ana <- analyses(obj)
+  ana <- analysisNames(obj)
   names(rpl) <- ana
 
   check <- feats_org$id

### 2022-08-04 dev/005_ms_dev_ms_preProcessing.R

@@ -10,6 +10,48 @@ files2 <- list.files(dir2, full.names = TRUE)
 fl <- files1[1:3]
 files <- fl
 
+
+### RaMS ---------------------
+
+library(RaMS)
+
+fl <- choose.files()
+
+fl <- c(
+  "C:\\Users\\Ricardo\\Documents\\R_DemoProject\\msfiles\\09_Sample_Sciex_MRM_Chromatograms_Nitrosamines_10ngml.mzML",
+  "C:\\Users\\Ricardo\\Documents\\R_DemoProject\\msfiles\\10_Sample_Sciex_MRM_Spectra_Nitrosamines_10ngml.mzML"
+)
+
+test <- xml2::as_list(xml2::read_xml(fl[1]))
+xml2::xml_structure(test, indent = 1)
+
+
+test$indexedmzML$mzML$run$chromatogramList[[3]]$binaryDataArrayList
+
+test <- grabMSdata(file = fl[1], grab_what = "TIC")
+
+msdata <- grabMSdata(files = fl[2], grab_what = "MS1") #rtrange = c(1100/60, 1200/60)
+
+data_nodes <- xml2::xml_find_all(files[1])
+
+
+
+targets4
+
+
+
+files
+
+
+
+test <- mzR::openMSfile(fl[1])
+
+test_chrom <- chromatogramHeader(test)
+test_chrom <- chromatograms(test)
+
+test <- mzR::close(test)
+
+
 ### Classes Hierarchy ---------------------------------------------------------------------------------------
 
 test <- new("msAnalysis", file = fl[1])
@@ -29,6 +71,9 @@ analyses(test1)
 test2 <- new("msmsData", test1)
 
 is(test1)
+
+
+
 
 
 

### 2022-08-08 README.Rmd
Second commit
LFU13

@@
 
-The goal of streamFind is to ...
+The streamFind package is a backend (code based) and frontend ([shiny](https://shiny.rstudio.com/) app) platform for assembly of modular workflows to support spectrometric and spectroscopic data processing.
+The major focus of streamFind is data processing for environmental and quality studies.
+The streamFind package aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data.
+The streamFind package can be used by academics but also by technicians due to the comprehensive documentation and well categorized set of integrated functionalities (modules).
+This handbook focuses on describing and demonstrating the backend of the streamFind package.
 
@@
 
-You can install the development version of streamFind like so:
+For installation of the streamFind package, it is recommended to firstly ensure that the dependencies are currently installed.
+Several existing R packages and external functionalities are used in streamFind for various processing steps.
+A major dependency of streamFind is the [patRoon](https://github.com/rickhelmus/patRoon) R package.`r # TODO add reference` The patRoon package combines several functionalities for basic and advanced data processing and can be used interchangeably with streamFind, see \@ref(patroon-interchangeability) for more information.
+Most of the tools that patRoon depends on are equally used in streamFind.
+In the next sub-chapter, the mandatory and optional dependencies are described with installation instructions.
+For a standalone and self-contained usage option, the streamFind docker image is recommended (see \@ref(docker-image)), avoiding the need to installing all required dependencies.
 
-``` r
-# FILL THIS IN! HOW CAN PEOPLE INSTALL YOUR DEV PACKAGE?
+### Install dependencies
+
+#### R and RTools
+
+[R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) are essential dependencies.
+The `r R.version.string` can be obtained in https://cran.r-project.org/.
+The RTools can be downloaded in https://cran.r-project.org/bin/windows/Rtools/.
+Make sure to download the right version for the R version installed.
+Installation instructions are given in both sources.
+
+#### Third party software
+
+The external third party dependencies recommended are listed in the section [other dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies) from the patRoon [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites).
+Below we list the mandatory and really recommended software and resources.
+
+* [ProteoWizard](https://proteowizard.sourceforge.io/), required for conversion of vendor MS formats to mzML/mzXML;  
+* [Java JDK](https://www.oracle.com/java/technologies/downloads/), mandatory for patRoon when plotting structure using MetFrag;  
+* [OpenMS](https://www.openms.de/downloads/); esternal tool for peak picking, grouping and alignment;  
+* [MetFrag CL](http://ipb-halle.github.io/MetFrag/projects/metfragcl/), used for formula annotation in patRoon;  
+* [PubChemLite](https://zenodo.org/record/6503754#.YtqxE3bP1hE), database used with MetFrag in patRoon;  
+* [SIRIUS](https://bio.informatik.uni-jena.de/software/sirius/), tool for peak picking and formula annotation in patRoon;    
+* [BioTransformer](https://bitbucket.org/djoumbou/biotransformer/downloads/), used for prediction of transformation products in patRoon;  
+* [SAFD](https://bitbucket.org/SSamanipour/safd.jl/src/master/), tool for peak picking with data in profile mode;  
+* [Open Babel](http://openbabel.org/wiki/Main_Page), used in patRoon for validation of chemicals properties (e.g., InChI and formulae);
+
+#### R packages
+
+The following code lines are adapted from the patRoon [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites) to ensure that both patRoon and other mutual R dependencies are installed properly.
+
+```{r r-packages, eval=FALSE}
+install.packages(c("BiocManager", "remotes"))
+
+BiocManager::install(c("mzR", "xcms","CAMERA"))
+
+remotes::install_github("blosloos/enviPick")
+
+BiocManager::install("ropls")
+remotes::install_github("rickhelmus/KPIC2")
+
+BiocManager::install("InterpretMSSpectrum")
+remotes::install_github("cbroeckl/RAMClustR")
+
+remotes::install_github("blosloos/nontargetData")
+remotes::install_github("blosloos/nontarget")
+
+remotes::install_github("rickhelmus/cliqueMS")
+
+# only needed for Bruker DataAnalysis integration
+remotes::install_github("BSchamberger/RDCOMClient")
+
+BiocManager::install(c("BiocStyle", "Rgraphviz")) 
+remotes::install_github("KelseyChetnik/MetaClean")
+
+remotes::install_github("rickhelmus/patRoon", upgrade = "never", dependencies = TRUE)
 ```
 
-## Example
+After installation of the patRoon R package, the location of each third party software should be given in an `.Rprofile` file as suggested in section [other dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies) the patRoon [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites).
+For facilitation the same code lines are given below.
+The function `patRoon::verifyDependencies()` from patRoon can be used to verify if all dependencies are found.
 
-This is a basic example which shows you how to solve a common problem:
-
-```{r example}
-library(streamFind)
-## basic example code
+```{r options-third-party-software, eval=FALSE}
+options(patRoon.path.pwiz = "C:/ProteoWizard") # location of ProteoWizard installation folder
+options(patRoon.path.SIRIUS = "C:/sirius-win64-3.5.1") # location where SIRIUS was extracted
+options(patRoon.path.OpenMS = "/usr/local/bin") # directory with the OpenMS binaries
+options(patRoon.path.pngquant = "~/pngquant") # directory containing pngquant binary
+options(patRoon.path.MetFragCL = "~/MetFragCommandLine-2.4.8.jar") # full location to the jar file
+options(patRoon.path.MetFragCompTox = "C:/CompTox_17March2019_SelectMetaData.csv") # full location to desired CompTox CSV file
+options(patRoon.path.MetFragPubChemLite = "~/PubChemLite_exposomics_20220429.csv") # full location to desired PubChemLite CSV file
+options(patRoon.path.BioTransformer = "~/biotransformer/biotransformer-3.0.1.jar")
+options(patRoon.path.obabel = "C:/Program Files/OpenBabel-3.0.0") # directory with Open Babel binaries
 ```
 
-What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:
+### Install streamFind
 
-```{r cars}
-summary(cars)
+The streamFind package can be installed from the GitHub repository https://github.com/ricardobachertdacunha/streamFind with the following code line.
+
+```{r install-streamfind, eval=FALSE}
+remotes::install_github("ricardobachertdacunha/streamFind", dependencies = TRUE)
 ```
 
-You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.
-`devtools::build_readme()` is handy for this.
-You could also use GitHub Actions to re-render `README.Rmd` every time you push.
-An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>.
+## Docker image
 
-You can also embed plots, for example:
+Not yet available.
 
-```{r pressure, echo = FALSE}
-plot(pressure)
-```
+## Backend usage
 
-In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
+Consult the Handbook_backend_streamFind.Rmd in the handbooks folder for detailed backend structure and usage documentation.

### 2022-08-08 R/S4_class_002_msAnalysis.R

@@ -1,13 +1,13 @@
 
-
-#### validity ------------------------------------------------------------
+#### validity ------------------------------------------------------------------
 
 msAnalysis_validity <- function(object) {
 
   valid <- TRUE
 
   if (nrow(object@peaks) > 0) {
-    must_have_names <- c("id", "rt", "mz", "intensity", "area", "drt", "rtmin", "rtmax", "dppm", "mzmin", "mzmax")
+    must_have_names <- c("id", "rt", "mz", "intensity", "area", "drt",
+                         "rtmin", "rtmax", "dppm", "mzmin", "mzmax")
     valid <- !FALSE %in% must_have_names %in% colnames(object@peaks)
   }
 
@@ -15,13 +15,15 @@ msAnalysis_validity <- function(object) {
 }
 
 
-### msAnalysis ----------------------------------------------------------------------------------------------
+
+### msAnalysis -----------------------------------------------------------------
 
 #' msAnalysis-class
 #'
-#' @description An S4 class representing an MS sample/file within the \pkg{streamFind} package.
-#' The \code{msAnalysis} is used to store and manage MS data and the respective methods can be used
-#' for inspection, processing and evaluation.
+#' @description An S4 class representing an MS sample/file within the
+#' \pkg{streamFind} package. The \code{msAnalysis} is used to store and
+#' manage MS data and the respective methods can be used for inspection,
+#' processing and evaluation.
 #'
 #' @template slot-msAnalysis
 #'
@@ -72,75 +74,12 @@ setClass("msAnalysis",
 )
 
 
-### S4 methods ----------------------------------------------------------------------------------------------
 
-#### initialize ----------------------------------------------------------
-
-#' @describeIn msAnalysis initializes an \linkS4class{msAnalysis} object.
-#'
-#' @usage ## S4 method to initialize an "msAnalysis" object \cr
-#' \code{new("msAnalysis", file)}
-#'
-#' @param file A character string with the complete file path.
-#' @param replicate optional character string for analysis replicate name
-#' @param blank optional character string for associated blank replicate name
-#'
-#' @importFrom data.table as.data.table
-#' @importFrom stringr str_replace
-#' @importFrom mzR openMSfile header runInfo chromatogramHeader
-#'
-#' @export
-#'
-setMethod("initialize", "msAnalysis", function(.Object, ...) {
-
-  .Object <- callNextMethod()
-
[…]

### 2022-08-11 handbooks/Handbook_backend_streamFind.Rmd
LFU14

@@
 
-The streamFind package includes example files, which are located  in the folder `system.file(package = "streamFind", dir = "extdata")`.
+The streamFind package includes example data files via the streamFindData R package.
+Once installed the full file paths of MS data files can be obtained as shown below.
+Alternatively, the files can be obtained directly from folder `system.file(package = "streamFindData", dir = "extdata")`.
 The files available are described in Table \@ref(tab:description-files).
@@
 ```{r resource-files, cache=FALSE}
-#path to files
-r_path <- system.file(package = "streamFind", dir = "extdata")
-
-#list of file full paths
-files <- list.files(r_path, pattern = ".mzML", full.names = TRUE)
+
+r_path <- system.file(package = "streamFindData", dir = "extdata")
+
+files <- list.files(r_path, pattern = ".mzML|.mzXML", full.names = TRUE)
+
+#list of file full paths from the data package
+#files <- streamFindData::msFilePaths()
 ```
@@
   N. = seq_len(length(files)),
-  file_type = rep("MS", 6 + 18),
+  file_type = rep("MS", 27),
   file_name = basename(files),
-  device = c(rep("Agilent Q-TOF", 24)),
-  data_type = c(rep("centroid", 3), rep("profile", 3), rep("centroid", 18)),
-  ac_mode = c(rep("MS/MS", 24)),
-  polarity = c(rep("positive", 6),
+  device = c(rep("Agilent Q-TOF", 27)),
+  data_type = c(rep("centroid", 6), rep("profile", 3), rep("centroid", 18)),
+  ac_mode = c(rep("MS/MS", 27)),
+  polarity = c(rep("positive", 9),
                rep("negative", 3), rep("positive",3),
@@
   description = c(
-    rep("Basic centroided MS data spiked with chemical and internal standards (CS and IS, respectively).", 3),
-    rep("Basic profile MS data spiked with CS and IS.", 3),
-    rep("Blank MS data spiked with IS.", 6),
-    rep("Wastewater secondary effluent MS data spiked with IS.", 6),
-    rep("Wastewater secondary effluent treated with ozonated strong water MS data spiked with IS.", 6)
+    rep("Basic centroided MS data as mzML spiked with chemical 
+        and internal standards (CS and IS, respectively).", 3),
+    rep("Basic centroided MS data as mzXML spiked with chemical 
+        and internal standards (CS and IS, respectively).", 3),
+    rep("Basic profile MS data as mzML spiked with CS and IS.", 3),
+    rep("Blank MS data as mzML spiked with IS.", 6),
+    rep("Wastewater secondary effluent MS data as mzML spiked with IS.", 6),
+    rep("Wastewater secondary effluent treated with ozonated 
+        strong water MS data as mzML spiked with IS.", 6)
   )
@@
 #creating an msAnalysis object for the 4th file, corresponding to profile MS data
-aProf <- newAnalysis(file = files[4])
+aProf <- newAnalysis(file = files[7])
 ```
@@
 As the method `getAnalysis` can be used to obtain an _msAnalysis_ from an _msData_ object, the method `addAnalyses` is used to add extra _msAnalysis_ objects to _msData_.
-Below the files 4 to 6 are added to the already created _msData_ (containing files 1 to 3).
+Below the files 7 to 9 are added to the already created _msData_ (containing files 1 to 3).
 Note that both `getAnalyses` and `addAnalyses` can be applied to equivalent classes to _msData_ (i.e., _uvData_ and _ramanData_).
@@
 #all profile MS files
-aProfs <- lapply(files[4:6], function(x) newAnalysis(x))
+aProfs <- lapply(files[7:9], function(x) newAnalysis(x))
 
@@
 
-* metadata of _msAnalysis_:
+* get metadata:
 
@@
 #metadata of the fourth analyses in the msData
-t(metadata(sp1, analyses = 4))
+getMetadata(sp1, analyses = 4)
 
 #applied directly to the msAnalysis S4 class
-t(metadata(a1))
+getMetadata(a1)
 ```
@@
 
-add metadata to an _msAnalysis_ object (a1):
+* add metadata to an _msAnalysis_ object (a1):
 
@@
 
-metadata(a1, which = names(meta_vec))
+getMetadata(a1, which = names(meta_vec))
 
@@
 
-metadata(a1, which = names(meta_vec))
-```
-
-add metadata to an _msData_ object (sp1):
+getMetadata(a1, which = names(meta_vec))
+```
+
+* add metadata to an _msData_ object (sp1):
 
@@
 
-metadata(sp1, which = names(meta_1))
+getMetadata(sp1, which = names(meta_1))
 
@@
 
-metadata(sp1, which = colnames(meta_df_2))
+getMetadata(sp1, analyses = c(3, 4), which = names(meta_1))
 ```
@@
 The S4 method TICs first evaluates if the TIC chromatogram is available in the mzML and when not, the spectra data is extracted and used instead.
+
+`r # TODO add faster functionality for TIC and BPC based on xml2``
 
@@
 ```{r newstreamset-basicworkflow, message=FALSE, results='hide'}
-sp2 <- newStreamSet(files = files[c(10:12, 16:18)], path = getwd(), title = "Project Example")
+sp2 <- newStreamSet(files = files[c(13:15, 19:21)], path = getwd(), title = "Project Example")
 ```
@@
 ```{r other-s4methods}
+#getter for the analysis table
+analysisTable(sp2)
+
 #getter for analysis names in msdata
@@
 #getter for polarity of analyses in a msData
-polarities(sp2)
+#polarities(sp2)
 

### 2022-08-11 R/fct_ms_002_peakGrouping.R

@@ -250,7 +250,7 @@ addAdjustedRetentionTime <- function(object, pat) {
 
     object@analyses <- lapply(object@analyses, function(x, hasSpectra) {
       if (!hasSpectra[analysisNames(x)]) {
-        x@spectra <- loadBasicRawSpectraHeaderMZR(filePaths(x))
+        x <- loadSpectraInfo(x)
       }
       return(x)
     }, hasSpectra = hasSpectra)

### 2022-10-04 handbooks/Handbook_backend_streamFind.Rmd
LFU15
Creation of global import namespace
LFU16_small fixes 2

@@
 r_path <- system.file(package = "streamFindData", dir = "extdata")
+#r_path <- system.file(package = "streamFind", dir = "extdata")
 
@@
 
-```{r}
+```{r add-metadata-msdata}
 #as a named vector

### 2022-10-06 handbooks/Handbook_backend_streamFind.Rmd
LFU17 Updated Classes

@@
 library(streamFind)
+```
+
+```{r libraries-dev, eval=FALSE, include=FALSE, cache=FALSE}
+#for development only
+library(data.table)
+library(xml2)
+library(checkmate)
+library(stringr)
+library(tools)
+library(dplyr)
+library(xcms)
+library(patRoon)
 ```
@@
 
-The slots analysis, file and replicate correspond to the analysis name, the full file path and the replicate group name, respectively.
+The slot analysis corresponds to the analysis name.
 The metadata slot is a list of objects with informative entries of the analysis file.
@@
 An essential aspect of environmental analysis is to operate with sample/analysis replicates.
-The replicates in _msData_ or _msAnalysis_ can be obtained by the `replicateNames` S4 method as shown below.
-The setter (assignment) for the replicate names is demonstrated below with the method `replicateNames<-` that takes a vector of character strings with the same length as the number of analyses in a _msData_ object or length one for a single analysis in _msAnalysis_.
+The replicates in _msData_ can be obtained by the `replicateNames` S4 method as shown below.
+The setter (assignment) for the replicate names is demonstrated below with the method `replicateNames<-` that takes a vector of character strings with the same length as the number of analyses in a _msData_ object.
 Alternatively, a table with a replicate character column or a character vector can be given in the files or replicates arguments, respectively.
+Note that _msAnalysis_ doesn't contain information about replicate correspondence.
+The correspondence between analyses is storage in the _msFeatures_ within the _msData_.
 
@@
 replicateNames(sp2)
-
-#setter for replicate name in msAnalysis
-a2 <- getAnalyses(sp2, 4)
-replicateNames(a2) <- "wastewater"
-
-#getter for replicate name in msAnalysis
-replicateNames(a2)
 ```
@@
 The blank subtraction (\@ref(filter-functions)) is then applied per analysis replicate considering the assigned blank replicate.
-Getting and assigning blank replicates from an _msAnalysis_ is also possible.
-However, assigning a blank without the context of a _streamSet_ and without the structure of _msData_ or similar level classes is not recommended.
 For cross project analysis, a subset of the analyses (_msAnalysis_ objects) in a given _msData_ can be concatenated to another _msData_, carrying the assigned blank replicate for each analysis as well as all the other analysis information (see `r # TODO make section cross project analysis``section cross project analysis for more information).
@@
 blankReplicateNames(sp2)
-
-#getter for msAnalysis
-blankReplicateNames(a2)
 ```
@@
 ### Other getter S4 methods
+
+`r # TODO list all S4 methods``
 
@@
 
+`r # TODO find another example for filling results`
+
 ```{r recursive-integration-inspection}
 # inspecting the filled peaks for a given mz and rt target
-plotFeaturePeaks(sp2_pg_fl, mz = 441.1670, rt = 916, ppm = 20, sec = 60)
+plotFeaturePeaks(sp2_pg_fl, mz = 441.1670, rt = 1097, ppm = 20, sec = 60)
 ```

### 2022-10-17 handbooks/Handbook_backend_streamFind.Rmd
LFU18

@@
 ```
+
 <br> <br>
 
-***
+------------------------------------------------------------------------
 
@@
 [R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) are essential dependencies.
-The `r R.version.string` can be obtained in https://cran.r-project.org/.
-The RTools can be downloaded in https://cran.r-project.org/bin/windows/Rtools/.
+The `r R.version.string` can be obtained in <https://cran.r-project.org/>.
+The RTools can be downloaded in <https://cran.r-project.org/bin/windows/Rtools/>.
 Make sure to download the right version for the R version installed.
@@
 
-* [ProteoWizard](https://proteowizard.sourceforge.io/), required for conversion of vendor MS formats to mzML/mzXML;  
-* [Java JDK](https://www.oracle.com/java/technologies/downloads/), mandatory for patRoon when plotting structure using MetFrag;  
-* [OpenMS](https://www.openms.de/downloads/); esternal tool for peak picking, grouping and alignment;  
-* [MetFrag CL](http://ipb-halle.github.io/MetFrag/projects/metfragcl/), used for formula annotation in patRoon;  
-* [PubChemLite](https://zenodo.org/record/6503754#.YtqxE3bP1hE), database used with MetFrag in patRoon;  
-* [SIRIUS](https://bio.informatik.uni-jena.de/software/sirius/), tool for peak picking and formula annotation in patRoon;    
-* [BioTransformer](https://bitbucket.org/djoumbou/biotransformer/downloads/), used for prediction of transformation products in patRoon;  
-* [SAFD](https://bitbucket.org/SSamanipour/safd.jl/src/master/), tool for peak picking with data in profile mode;  
-* [Open Babel](http://openbabel.org/wiki/Main_Page), used in patRoon for validation of chemicals properties (e.g., InChI and formulae);
+-   [ProteoWizard](https://proteowizard.sourceforge.io/), required for conversion of vendor MS formats to mzML/mzXML;\
+-   [Java JDK](https://www.oracle.com/java/technologies/downloads/), mandatory for patRoon when plotting structure using MetFrag;\
+-   [OpenMS](https://www.openms.de/downloads/); esternal tool for peak picking, grouping and alignment;\
+-   [MetFrag CL](http://ipb-halle.github.io/MetFrag/projects/metfragcl/), used for formula annotation in patRoon;\
+-   [PubChemLite](https://zenodo.org/record/6503754#.YtqxE3bP1hE), database used with MetFrag in patRoon;\
+-   [SIRIUS](https://bio.informatik.uni-jena.de/software/sirius/), tool for peak picking and formula annotation in patRoon;\
+-   [BioTransformer](https://bitbucket.org/djoumbou/biotransformer/downloads/), used for prediction of transformation products in patRoon;\
+-   [SAFD](https://bitbucket.org/SSamanipour/safd.jl/src/master/), tool for peak picking with data in profile mode;\
+-   [Open Babel](http://openbabel.org/wiki/Main_Page), used in patRoon for validation of chemicals properties (e.g., InChI and formulae);
 
@@
 
-The following code lines are adapted from the patRoon [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites) to ensure that both patRoon and other mutual R dependencies are installed properly.
+The following code lines are from the patRoon [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites) to ensure that both patRoon and other mutual R dependencies are installed properly.
 
@@
 
-After installation of the patRoon R package, the location of each third party software should be given in an `.Rprofile` file as suggested in section [other dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies) the patRoon [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites).
+After installation of the patRoon R package, the location of each third party software should be given in an `.Rprofile` file as suggested in section [other dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies) of the patRoon [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites).
 For facilitation the same code lines are given below.
@@
 
-The streamFind package can be installed from the GitHub repository https://github.com/ricardobachertdacunha/streamFind with the following code line.
+The streamFind package can be installed from the GitHub repository <https://github.com/ricardobachertdacunha/streamFind> with the following code line.
 
@@
 ## Docker image
+
+To be done.
 
@@
 Note, the MS files (.mzML) were trimmed with the function `trimSpectraFilesMZR` to reduce their size.
-The centroided Agilent Q-TOF files were trimmed with retention time between 900 and 1350 seconds, _m/z_ for MS1 data between 200 and 450 Da and _m/z_ for MS2 between 35 and 450 Da.
+The centroided Agilent Q-TOF files were trimmed with retention time between 900 and 1350 seconds, *m/z* for MS1 data between 200 and 450 Da and *m/z* for MS2 between 35 and 450 Da.
 Additionally, MS1 and MS2 traces below 100 and 50 counts were removed, respectively.
-The profile Agilent Q-TOF files were trimmed with retention time between 1000 and 1200 seconds, _m/z_ for MS1 data between 200 and 300 Da and _m/z_ for MS2 between 35 and 300 Da.
+The profile Agilent Q-TOF files were trimmed with retention time between 1000 and 1200 seconds, *m/z* for MS1 data between 200 and 300 Da and *m/z* for MS2 between 35 and 300 Da.
 Additionally, MS1 and MS2 traces below 1 count were removed to exclude empty spectra.
@@
 db$ionization <- "positive"
-db[, `:=`("mz_pos" = neutralMass + 1.0073, "mz_neg" = neutralMass - 1.0073 )]
+db[, `:=`("mz_pos" = mass + 1.0073, "mz_neg" = mass - 1.0073 )]
 db[grepl("neg", db$comment), ionization := "both"]
-data.table::setnames(db, "mixIUTA", "tag")
+data.table::setnames(db, "mix_id", "tag", skip_absent = TRUE)
 db[tag %in% 1, tag := "CS"]
@@
 
-knitr::kable(db[, .(name, CAS, formula, neutralMass, rt, ionization, tag, in_file)], caption = "Chemical standards spiked.") %>%
+knitr::kable(db[, .(name, formula, mass, rt, ionization, tag, in_file)], caption = "Chemical standards spiked.") %>%
   kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>%
@@
 
-The streamFind package uses S4 class objects (e.g., _streamSet_, _msData_, _msAnalysis_, etc.) for structuring data.
+The streamFind package uses S4 class objects (e.g., *streamSet*, *msData*, *msAnalysis*, etc.) for structuring data.
 The S4 classes and their main getter and setter methods are presented and exemplified in this section.
 An overview of the S4 classes and their hierarchy is presented in Figure \@ref(fig:classes-overview) using MS analyses as an example.
-Other data types, such as from UV and Raman analyses, would lead to different classes (e.g., _uvData_ and _uvAnalysis_ and _ramanData_ and _ramanAnalysis_, respectively) but the hierarchy is similar as shown for MS data.
+Other data types, such as from UV and Raman analyses, would lead to different classes (e.g., *uvData* and *uvAnalysis* and *ramanData* and *ramanAnalysis*, respectively) but the hierarchy is similar as shown for MS data.
 
@@
 
-* _streamSet_ refers to the broader S4 class applied in the streamFind package and holds basic information for an analysis set, see \@ref(streamset-class);
-* _msData_ refers to the S4 subclass of _streamSet_ when MS analyses (i.e., MS files) are added to the set, see \@ref(msdata-class);
-* _msAnalysis_ refers to the S4 class of the object generated for each MS file added to the _msData_ object in the slot "analyses";
-* `setInfo` is the method to obtain the basic information (i.e., title, date and path) of a _streamSet_ object;
-* `analyses` is the name of the slot holding the analyses (as list) and is also a method to get the analysis names in the _streamSet_;
-* `replicate` is the name of an analysis replicate group and the method to get the replicate names;
-* `blank` is the name of a blank analysis replicate group to be used for blank (blind) subtraction and the method to get blank replicate names;
-* __traces__ corresponds to a raw data signal (e.g., mass trace in MS data) obtained from a given analysis file;
-* `spectra` for MS analyses is the method to get a table with a spectrum for each cycle time (i.e., retention time);
-* __EIC__ is an extracted ion chromatogram (retention time vs intensity) and can be obtained with the method `EICs`;
-* __TIC__ is a total ion chromatograms (retention time vs intensity) and can be obtained with the method `TICs`;
-* __XIC__ is a three dimensional ion chromatogram (_m/z_ vs retention time vs intensity) and can be obtained with the method `XICs`;
-* __peak__ corresponds to a cluster of traces (e.g., integrated chromatographic MS peak) within an analysis file and can be accessed with the method `peaks`; 
-* __feature__ corresponds to a group of corresponding peaks across analysis files/analyses and can be accessed with the method `features`;
-* __MS2__ corresponds to second order fragmentation data as acquired in tandem MS or MS/MS mode and can be accessed with the method `MS2s`.
+-   *streamSet* refers to the broader S4 class applied in the streamFind package and holds basic information for an analysis set, see \@ref(streamset-class);
+-   *msData* refers to the S4 subclass of *streamSet* when MS analyses (i.e., MS files) are added to the set, see \@ref(msdata-class);
+-   *msAnalysis* refers to the S4 class of the object generated for each MS file added to the *msData* object in the slot "analyses";
+-   `setInfo` is the method to obtain the basic information (i.e., title, date and path) of a *streamSet* object;
+-   `analyses` is the name of the slot holding the analyses (as list) and is also a method to get the analysis names in the *streamSet*;
+-   `replicate` is the name of an analysis replicate group and the method to get the replicate names;
+-   `blank` is the name of a blank analysis replicate group to be used for blank (blind) subtraction and the method to get blank replicate names;
+-   **traces** corresponds to a raw data signal (e.g., mass trace in MS data) obtained from a given analysis file;
+-   `spectra` for MS analyses is the method to get a table with a spectrum for each cycle time (i.e., retention time);
+-   **EIC** is an extracted ion chromatogram (retention time vs intensity) and can be obtained with the method `EICs`;
+-   **TIC** is a total ion chromatograms (retention time vs intensity) and can be obtained with the method `TICs`;
+-   **XIC** is a three dimensional ion chromatogram (*m/z* vs retention time vs intensity) and can be obtained with the method `XICs`;
+-   **peak** corresponds to a cluster of traces (e.g., integrated chromatographic MS peak) within an analysis file and can be accessed with the method `peaks`;
+-   **feature** corresponds to a group of corresponding peaks across analysis files/analyses and can be accessed with the method `features`;
+-   **MS2** corresponds to second order fragmentation data as acquired in tandem MS or MS/MS mode and can be accessed with the method `MS2s`.
 
@@
 The set path can be assigned as argument but for the example below it was left as default (i.e., the working directory).
-The initial S4 class created in the `newStreamSet` function is _streamSet_.
+The initial S4 class created in the `newStreamSet` function is *streamSet*.
 The subclass is then defined according to the type of files added.
@@
 
-Because the added files were all with mzML format (i.e., MS files), the subclass is directly defined as _msData_.
+Because the added files were all with mzML format (i.e., MS files), the subclass is directly defined as *msData*.
 The structure for further processing data is then set for MS data.
-Other data formats/structure will lead to different sub classes (e.g., _uvData_ and _ramanData_) but is not yet implemented within the streamFind package.
+Other data formats/structure will lead to different sub classes (e.g., *uvData* and *ramanData*) but is not yet implemented within the streamFind package.
 
@@
 
-The _msData_ structure includes the _streamSet_ class plus the `features` slot.
+The *msData* structure includes the *streamSet* class plus the `features` slot.
 The title, date and path slots correspond to the set title, creation date and directory in disk, respectively.
-The analyses slot is a list with a _msAnalysis_ S4 class object per MS file added.
-The structure of the _msAnalysis_ class is described in the next sub-chapter.
-The `features` slot is a _msFeatures_ S4 class object to hold the results from alignment, grouping and annotation of peaks across analyses.
-Further information for _msAnalysis_ is given in the next sub-chapter (\@ref(msanalysis-class)) and for _msFeatures_ is given in the sub-chapter \@ref(msfeatures-class), after creation of features by performing alignment and grouping of peaks across analyses.
-
-* _msData_ structure:
+The analyses slot is a list with a *msAnalysis* S4 class object per MS file added.
+The structure of the *msAnalysis* class is described in the next sub-chapter.
+The `features` slot is a *msFeatures* S4 class object to hold the results from alignment, grouping and annotation of peaks across analyses.
+Further information for *msAnalysis* is given in the next sub-chapter (\@ref(msanalysis-class)) and for *msFeatures* is given in the sub-chapter \@ref(msfeatures-class), after creation of features by performing alignment and grouping of peaks across analyses.
+
+-   *msData* structure:
 
@@
 
-The _msAnalysis_ class is structured as shown below.
-Note that the S4 method `getAnalyses` for _msData_ was used to obtain the _msAnalysis_ of the first analysis as defined by the second argument.
-In general, access to raw data (i.e., traces and chromatograms) as well as respective plotting S4 methods can be applied to both the _msAnalysis_ and _msData_ (as described in \@ref(accessing-raw-data)).
+The *msAnalysis* class is structured as shown below.
+Note that the S4 method `getAnalyses` for *msData* was used to obtain the *msAnalysis* of the first analysis as defined by the second argument.
+In general, access to raw data (i.e., traces and chromatograms) as well as respective plotting S4 methods can be applied to both the *msAnalysis* and *msData* (as described in \@ref(accessing-raw-data)).
 
@@
 The metadata list can be expanded with additional information about the analysis/sample (e.g., sample location, weather conditions, process sensor data, etc.).
-The slot parameters is used to store processing steps (e.g., peak picking) as _settings_ S4 class objects, including the processing step name/ID, the algorithm used and the respective settings applied.
+The slot parameters is used to store processing steps (e.g., peak picking) as *settings* S4 class objects, including the processing step name/ID, the algorithm used and the respective settings applied.
 The parameters can be used for history track but also for re-run the processing steps when required.
@@
 Finally, the peaks slot is used to store the chromatographic peaks after peak picking.
-As no other processing step was applied the spectra, chromatograms and peaks tables in the _msAnalysis_ are empty for now.
-
- * _msAnalysis_ structure:
+As no other processing step was applied the spectra, chromatograms and peaks tables in the *msAnalysis* are empty for now.
+
+-   *msAnalysis* structure:
 
@@
 
-An _msAnalysis_ object can be created without the need of running `newStreamSet`.
-The method `newAnalysis` can be used for initiating the _msAnalysis_ when a full path of an MS file (i.e., with extension mzML or other MS format) is given.
+An *msAnalysis* object can be created without the need of running `newStreamSet`.
+The method `newAnalysis` can be used for initiating the *msAnalysis* when a full path of an MS file (i.e., with extension mzML or other MS format) is given.
 Similar to `newStreamSet`, the format of the file given as file argument dictates the class of the analysis object returned from the `newAnalysis` function.
@@
 
-As the method `getAnalysis` can be used to obtain an _msAnalysis_ from an _msData_ object, the method `addAnalyses` is used to add extra _msAnalysis_ objects to _msData_.
-Below the files 7 to 9 are added to the already created _msData_ (containing files 1 to 3).
-Note that both `getAnalyses` and `addAnalyses` can be applied to equivalent classes to _msData_ (i.e., _uvData_ and _ramanData_).
-Also, note that when adding analyses to an _msData_ object with features (i.e., with peaks grouped across analyses), the _msFeatures_ content is cleared as the grouping is invalidated with new _msAnalysis_ objects.
+As the method `getAnalysis` can be used to obtain an *msAnalysis* from an *msData* object, the method `addAnalyses` is used to add extra *msAnalysis* objects to *msData*.
+Below the files 7 to 9 are added to the already created *msData* (containing files 1 to 3).
+Note that both `getAnalyses` and `addAnalyses` can be applied to equivalent classes to *msData* (i.e., *uvData* and *ramanData*).
+Also, note that when adding analyses to an *msData* object with features (i.e., with peaks grouped across analyses), the *msFeatures* content is cleared as the grouping is invalidated with new *msAnalysis* objects.
 
@@
 
-Below, the `getMetadata` S4 method of _msData_ is used to display the metadata. Note, `getMetadata` also works for _msAnalysis_ S4 class as shown below.
-
-* get metadata:
+Below, the `getMetadata` S4 method of *msData* is used to display the metadata. Note, `getMetadata` also works for *msAnalysis* S4 class as shown below.
+
+-   get metadata:
 
@@
 
-Additional metadata can be added with the `addMetadata` method for both _msAnalysis_ and _msData_.
-Note that for _msData_ the metadata input needs to be a list of vectors with the same length as the number of analyses or a data.frame/data.table with the number of rows as the nmumber of analyses, as shown below.
-
-* add metadata to an _msAnalysis_ object (a1):
+Additional metadata can be added with the `addMetadata` method for both *msAnalysis* and *msData*.
+Note that for *msData* the metadata input needs to be a list of vectors with the same length as the number of analyses or a data.frame/data.table with the number of rows as the nmumber of analyses, as shown below.
+
+-   add metadata to an *msAnalysis* object (a1):
 
@@
 
-* add metadata to an _msData_ object (sp1):
+-   add metadata to an *msData* object (sp1):
 
@@
 The data search is based on pre-defined time and mass (inc.
-_m/z_) targets.
-For traces (i.e., raw MS mass traces), time (in seconds) and _m/z_ (in Da) ranges are given for accessing and plotting.
+*m/z*) targets.
+For traces (i.e., raw MS mass traces), time (in seconds) and *m/z* (in Da) ranges are given for accessing and plotting.
 Access to mass traces is relevant for further development of data processing steps/functions and for visualization of raw data, easing the evaluation of processing steps (e.g., peaks wrongly not found during peak picking).
@@
 
-As aforementioned, access to raw data is based on defining _m/z_ and retention time targets.
+As aforementioned, access to raw data is based on defining *m/z* and retention time targets.
 The function `makeTargets` is used for comprehensively build targets.
@@
 ```{r carb-diu}
-carb <- db[name %in% "Carbamazepin d10", .(name, neutralMass, rt)]
-carb$mz <- carb$neutralMass + 1.0073
-carb
-
-diu <- db[name %in% "Diuron d6", .(name, neutralMass, rt)]
-diu$mz <- diu$neutralMass + 1.0073
-diu
+#Carbamazepin-d10 is ionized only in positive mode
+carbamazepin_d10 <- db[name %in% "Carbamazepin-d10", .(name, mass, rt)]
+carbamazepin_d10
+
+#Diuron-d6 is ionized in both positive and negative modes
+diuron_d6 <- db[name %in% "Diuron-d6", .(name, mass, rt)]
+diuron_d6
 ```
@@
 ```{r make_targets}
+carb_pos <- carbamazepin_d10$mass + 1.0073
+carb_rt <- carbamazepin_d10$rt
+diu_pos <- diuron_d6$mass + 1.0073
+diu_rt <- diuron_d6$rt
+
+sec_dev <- 30 #seconds
+ppm_dev <- 20
+
 #case 1
-mz1 <- c(247.1650, 239.0620)
-rt1 <- c(1075, 1157)
+mz1 <- c(carb_pos, diu_pos)
+rt1 <- c(carb_rt, diu_rt)
 id1 <- c("target1", "target2")
-targets1 <- makeTargets(mz = mz1, rt = rt1, ppm = 20, sec = 30, id = id1)
+targets1 <- makeTargets(mz = mz1, rt = rt1, ppm = ppm_dev, sec = sec_dev, id = id1)
 
@@
 #case 2
-mz2 <- data.frame(mzmin = c(247.1601, 239.0572), mzmax = c(247.1699, 239.0668))
-rt2 <- data.frame(rtmin = c(1045, 1127), rtmax = c(1105, 1187))
+ppm_carb <- ppm_dev / 1E6 * carb_pos
+ppm_diu <- ppm_dev / 1E6 * diu_pos
+
+mz2 <- data.frame(mzmin = c(carb_pos - ppm_carb, diu_pos - ppm_diu),
+                  mzmax = c(carb_pos + ppm_carb, diu_pos + ppm_diu))
+
+rt2 <- data.frame(rtmin = c(carb_rt - sec_dev, diu_rt - sec_dev),
+                  rtmax = c(carb_rt + sec_dev, diu_rt + sec_dev))
+
 targets2 <- makeTargets(mz = mz2, rt = rt2)
-
 targets2
@@
   id = c("target1", "target2"),
-  mz = c(247.1650, 239.0620),
-  rt = c(1075, 1157)
-)
-targets3 <- makeTargets(mz = mz3, ppm = 20, sec = 30)
-
+  mz = c(carb_pos, diu_pos),
+  rt = c(carb_rt, diu_rt)
+)
+
+targets3 <- makeTargets(mz = mz3, ppm = ppm_dev, sec = sec_dev)
 targets3
@@
   id = c("target1", "target2"),
-  mzmin = c(247.1601, 239.0572), mzmax = c(247.1699, 239.0668),
-  rtmin = c(1045, 1127), rtmax = c(1105, 1187)
-)
+  mzmin = c(carb_pos - ppm_carb, diu_pos - ppm_diu),
+  mzmax = c(carb_pos + ppm_carb, diu_pos + ppm_diu),
+  rtmin = c(carb_rt - sec_dev, diu_rt - sec_dev),
+  rtmax = c(carb_rt + sec_dev, diu_rt + sec_dev)
+)
+
 targets4 <- makeTargets(mz = mz4)
-
 targets4
@@
 
-Extracted ion chromatograms (EICs) are obtained either from loaded spectra in the _msData_ or _msAnalysis_ objects or from querying traces directly in the raw data file (mzML or mzXML).
+Extracted ion chromatograms (EICs) are obtained either from loaded spectra in the *msData* or *msAnalysis* objects or from querying traces directly in the raw data file (mzML or mzXML).
 The latter is less disk memory intense as data is loaded from raw files on demand.
@@
 The obtained EICs can be plotted with `plotEICs` S4 method, as shown below.
-The object for `plotEICs` can be a `data.table` as resulting from the `EICs` method or the actual _msData_ or _msAnalysis_ objects, using the same arguments as in `EICs` for collecting the data (i.e., to define the targets).
+The object for `plotEICs` can be a `data.table` as resulting from the `EICs` method or the actual *msData* or *msAnalysis* objects, using the same arguments as in `EICs` for collecting the data (i.e., to define the targets).
 
@@
 
-Total ion chromatograms (TICs) are obtained either by pre-loading and adding the chromatogram data to the _msAnalysis_, extracting the data in a `TICs` call (S4 method) or loading all spectra and applying the sum of intensities to each spectrum.
+Total ion chromatograms (TICs) are obtained either by pre-loading and adding the chromatogram data to the *msAnalysis*, extracting the data in a `TICs` call (S4 method) or loading all spectra and applying the sum of intensities to each spectrum.
 Note, TIC is already available as chromatogram in mzML data from the major HRMS vendors.
 The S4 method TICs first evaluates if the TIC chromatogram is available in the mzML and when not, the spectra data is extracted and used instead.
-
-`r # TODO add faster functionality for TIC and BPC based on xml2``
 
@@
 
-A more informative method of obtaining EICs is via the method `XICs` which extract a the three dimensional (_m/Z_, rt and intensity) chromatograms from given targets.
-In the example below, the _m/z_ dimension is also included in the resulting table to improve inspection of traces.
+A more informative method of obtaining EICs is via the method `XICs` which extract a the three dimensional (*m/Z*, rt and intensity) chromatograms from given targets.
+In the example below, the *m/z* dimension is also included in the resulting table to improve inspection of traces.
 
@@
 Other arguments are available for plotting `XICs`, including a target area for the expected targets.
-As shown in the plot below, the red mark is defined by the arguments `targetsMark` (a table with columns mz and rt to define the expected _m/z_ and retention times) and `ppmMark` and `secMark` to define the mass (in ppm) and time (in seconds) range of the target squares.
+As shown in the plot below, the red mark is defined by the arguments `targetsMark` (a table with columns mz and rt to define the expected *m/z* and retention times) and `ppmMark` and `secMark` to define the mass (in ppm) and time (in seconds) range of the target squares.
 The `numberRows` is used to control the number of rows of the grouped plot.
@@
 The S4 methods applied above parse the data from the raw data files which is less disk memory intense but slower.
-Alternatively, the raw data can be loaded to (each) _msAnalysis_  and from their is faster available.
+Alternatively, the raw data can be loaded to (each) *msAnalysis* and from their is faster available.
 However, when loaded, the raw data has a higher memory footprint.
-Therefore, the data can be optionally loaded and stored as a simple _data.table_ in the respective _msAnalysis_ (other classes of the same level might have the raw data directly loaded if the data has a lower memory footprint).
-Below we demonstrate how to load (via `loadRawData` S4 method) and access (via `spectra` or `chromatograms`) the raw spectra and chromatograms from an _msAnalaysis_ and _msData_ object.
-Both S4 methods (`spectra` or `chromatograms`) return a _data.table_ where further operations can be applied to subset the data, as demonstrated below.
-The standard _data.table_ operations can be applied, see documentation in https://cran.r-project.org/web/packages/data.table/ for more information and code examples/instructions.
-A cheat sheet for _data.table_ operations can be downloaded in https://raw.githubusercontent.com/rstudio/cheatsheets/main/datatable.pdf.
-In streamFind, the _data.table_ is always the preferred format for table objects.
+Therefore, the data can be optionally loaded and stored as a simple *data.table* in the respective *msAnalysis* (other classes of the same level might have the raw data directly loaded if the data has a lower memory footprint).
+Below we demonstrate how to load (via `loadRawData` S4 method) and access (via `spectra` or `chromatograms`) the raw spectra and chromatograms from an *msAnalaysis* and *msData* object.
+Both S4 methods (`spectra` or `chromatograms`) return a *data.table* where further operations can be applied to subset the data, as demonstrated below.
+The standard *data.table* operations can be applied, see documentation in <https://cran.r-project.org/web/packages/data.table/> for more information and code examples/instructions.
+A cheat sheet for *data.table* operations can be downloaded in <https://raw.githubusercontent.com/rstudio/cheatsheets/main/datatable.pdf>.
+In streamFind, the *data.table* is always the preferred format for table objects.
 
@@
 #load raw data to an msAnalysis object, while setting a minimum intensity value for both MS1 and MS2 traces
-a1 <- loadRawData(a1, minIntensityMS1 = 200, minIntensityMS2 = 100)
+a1 <- loadRawData(a1, minIntensityMS1 = 0, minIntensityMS2 = 100)
 ```
@@
 
-The analysis set creation is demonstrated below for six HRMS data files corresponding to blank and wastewater influent samples, both measured in triplicate.
-The argument `files` is the main input and can either be a list of full path files, a table with file (i.e., full file path), replicate (i.e., the name of the analysis replicate group for each file) and blank (i.e., the name of the associated blank analysis replicate group for each file), or the `analysisInfo` data.frame from the package [patRoon](https://github.com/rickhelmus/patRoon).
+The analysis set creation is demonstrated below for twelve HRMS data files corresponding to blank and wastewater influent samples, both measured in triplicate and in positive and negative modes.
+The argument `files` is the main input and can either be a vector of full path files, a table with file (i.e., full file path), replicate (i.e., the name of the analysis replicate group for each file) and blank (i.e., the name of the associated blank analysis replicate group for each file), or the `analysisInfo` data.frame from the package [patRoon](https://github.com/rickhelmus/patRoon).
 The set path can be assign as argument `path` and is the directory where all the set files (i.e., scripts, cache databases, objects and results) will be stored by default.
@@
 ```{r newstreamset-basicworkflow, message=FALSE, results='hide'}
-sp2 <- newStreamSet(files = files[c(13:15, 19:21)], path = getwd(), title = "Project Example")
+sp2 <- newStreamSet(files = files[10:21], path = getwd(), title = "Project Example")
 ```
@@
 An essential aspect of environmental analysis is to operate with sample/analysis replicates.
-The replicates in _msData_ can be obtained by the `replicateNames` S4 method as shown below.
-The setter (assignment) for the replicate names is demonstrated below with the method `replicateNames<-` that takes a vector of character strings with the same length as the number of analyses in a _msData_ object.
+The replicates in *msData* can be obtained by the `replicateNames` S4 method as shown below.
+The setter (assignment) for the replicate names is demonstrated below with the method `replicateNames<-` that takes a vector of character strings with the same length as the number of analyses in a *msData* object.
 Alternatively, a table with a replicate character column or a character vector can be given in the files or replicates arguments, respectively.
-Note that _msAnalysis_ doesn't contain information about replicate correspondence.
-The correspondence between analyses is storage in the _msFeatures_ within the _msData_.
+Note that *msAnalysis* doesn't contain information about replicate correspondence.
+The correspondence between analyses is storage in the *msFeatures* within the *msData*.
 
@@
 #setter for replicate names in msData
-replicateNames(sp2) <- c(rep("blank", 3),rep("influent", 3))  
+replicateNames(sp2) <- c(
+  rep("blank_neg", 3),
+  rep("blank_pos", 3),
+  rep("influent_neg", 3),
+  rep("influent_pos", 3)
+)
+
 replicateNames(sp2)
@@
 The blank subtraction (\@ref(filter-functions)) is then applied per analysis replicate considering the assigned blank replicate.
-For cross project analysis, a subset of the analyses (_msAnalysis_ objects) in a given _msData_ can be concatenated to another _msData_, carrying the assigned blank replicate for each analysis as well as all the other analysis information (see `r # TODO make section cross project analysis``section cross project analysis for more information).
+For cross project analysis, a subset of the analyses (*msAnalysis* objects) in a given *msData* can be concatenated to another *msData*, carrying the assigned blank replicate for each analysis as well as all the other analysis information.
 
@@
 #setter for the blank analysis replicate of msData object
-blankReplicateNames(sp2) <- rep("blank", 6)
+blankReplicateNames(sp2) <- c(
+  rep("blank_neg", 3),
+  rep("blank_pos", 3),
+  rep("blank_neg", 3),
+  rep("blank_pos", 3)
+)
+
 blankReplicateNames(sp2)
@@
 ### Other getter S4 methods
-
-`r # TODO list all S4 methods``
 
@@
 
-#getter for analysis name in msAnalysis
-analysisNames(a2)
-
 #getter for polarity of analyses in a msData
-#polarities(sp2)
-
-#getter for polarity of an msAnalysis
-polarities(a2)
+polarities(sp2)
 
@@
 
-For each processing steps, parameter settings are often used.
-In streamFind, the parameters used for a given processing step are added/stored as _settings_ S4 class objects in each _msAnalysis_ (or other class of the same level, such as _uvAnalysis_ or _ramanAnalysis_).
-Each _settings_ object contains the algorithm and list of parameters with the respective values.
-The example below demonstrates the workflow to perform peak picking either by adding the parameters during the call for `peakPicking` or by initially add the parameters to each _msAnalysis_ and then run the `peakPicking`, which looks for the parameters in each _msAnalysis_.
-If different settings are applied among the _msAnalysis_ objects, these are used.
-This means that different processing parameters can be applied within the same function call.
-If parameters are added during the function call but they already exist in the _msAnalysis_, these are overwritten by the settings used in the function call.
+For each processing step, settings are often used.
+In streamFind, the settings used for a given processing step are added/stored as *settings* S4 class objects.
+Each *settings* object contains the function call name, the algorithm and list of parameters with the respective values.
+The example below demonstrates the workflow to perform peak picking either by adding the settings during the call for `peakPicking` or by initially add the settings to each *msAnalysis* and then run the `peakPicking`, which looks for the settings in each *msAnalysis*.
+If different settings are applied among the *msAnalysis* objects, these are used.
+This means that different algorithms and/or processing parameters can be applied within the same function call.
+If settings are added during the function call but they already exist in the *msAnalysis*, these are overwritten by the settings used in the function call.
 
@@
 
-The function `createSettings` is use to assemble the _settings_ object, as shown below.
-Then, the `addParameters` method is used to either add parameters to each analysis in a _msData_ object or directly to a _msAnalysis_ object.
-The `getParameters` can be used to check which exist in the _msData_ and _msAnalysis_.
+The function `createSettings` is use to assemble the *settings* object, as shown below.
+Then, the `addSettings` method is used to either add parameters to each analysis in a *msData* object or directly to a *msAnalysis* object.
+The `getSettings` can be used to check which settings exist in the *msData* and/or *msAnalysis*.
 
@@
   algorithm = "xcms3",
-  settings = param
+  parameters = param
 )
@@
 # add the settings to all analyses in msData
-sp2_pp <- addParameters(sp2_pp, settings = settings_pp, where = "analyses")
+sp2_pp <- addSettings(sp2_pp, settings = settings_pp, where = "analyses")
 
 # get the parameters in the third analysis of the msData object
-getParameters(sp2_pp, where = "analyses", analyses = 3)
+getSettings(sp2_pp, where = "analyses", analyses = 3)
+
+#get msAnalysis
+a2 <- getAnalyses(sp2, analyses = 4)
 
 # add the settings to the msAnalysis object
-a2 <- addParameters(a2, settings = settings_pp)
+a2 <- addSettings(a2, settings = settings_pp)
 
 # get the parameters of the msAnalysis object
-getParameters(a2, call = "peakPicking")
+getSettings(a2, call = "peakPicking")
 ```
@@
 # parameters are added to each analysis after running peak picking
-summary(getParameters(sp2_pp_2, where = "analyses", analyses = 1))
+summary(getSettings(sp2_pp_2, where = "analyses", analyses = 1))
 ```
@@
 
-As aforementioned, different settings for the same processing step can exist in a _msData_.
+As aforementioned, different settings for the same processing step can exist in a *msData*.
 Below we add different settings for the influent samples, as an example.
@@
   algorithm = "openms",
-  settings = list(
+  parameters = list(
     noiseThrInt = 500,
@@
 
-sp2_pp_3 <- addParameters(sp2_pp_3, settings = settings_pp, where = "analyses")
+sp2_pp_3 <- addSettings(sp2_pp_3, settings = settings_pp, where = "analyses")
 
 # adding different settings to analysis 4 to 6
-sp2_pp_3 <- addParameters(sp2_pp_3, settings = settings_pp_2, where = "analyses", analyses = 4:6)
+sp2_pp_3 <- addSettings(sp2_pp_3, settings = settings_pp_2, where = "analyses", analyses = 7:12)
 
 # the algorithm of the influent samples is different
-sapply(analysisNames(sp2_pp_3), function(x) getAlgorithm(getParameters(sp2_pp_3, where = "analyses", analyses = x)[[1]]))
+sapply(analysisNames(sp2_pp_3), function(x) getAlgorithm(getSettings(sp2_pp_3, where = "analyses", analyses = x)[[1]]))
 ```
@@
 
-The _settings_ object can be exported as a JSON or rds file and then imported for other sets where the same parameter settings are to be applied.
+The *settings* object can be exported as a JSON or rds file and then imported for other sets where the same parameter settings are to be applied.
 The methods `exportSettings` and `importSettings` are used, as shown below.
@@
 
-The access and visualization of peaks uses the calculated/estimated time and _m/z_ dimensions for collecting and plotting the correspondent mass traces.
-Examples are shown below to extract and plot peaks from both _msData_ and _msAnalysis_.
-
-For a _msData_ object:
+The access and visualization of peaks uses the calculated/estimated time and *m/z* dimensions as well as the monoisotopic mass for collecting and plotting the correspondent mass traces.
+Examples are shown below to extract and plot peaks from both *msData* and *msAnalysis*.
+
+For a *msData* object:
 
 ```{r inspecting-peaks-msdata, eval=FALSE}
-#access to peaks in msData based on built targets as used for extracting EICs
-peaks(sp2_pp[c(1, 4)], mz = mz1[1], rt = rt1[1], ppm = 10, sec = 30)
-```
-
-```{r inspecting-peaks-msdata-table}
-#access to peaks in msData based on built targets as used for extracting EICs
-peaks(sp2_pp[c(1, 4)], mz = mz1[1], rt = rt1[1], ppm = 10, sec = 30)
+#access to peaks by mz in analyses 4 and 10 of msData based on built targets, as used for extracting EICs
+peaks(sp2_pp, analyses = c(4, 10), mz = mz1[1], rt = rt1[1], ppm = 10, sec = 30)
+```
+
+```{r inspecting-peaks-msdata-table, echo=FALSE}
+peaks(sp2_pp[c(4, 10)], mz = mz1[1], rt = rt1[1], ppm = 10, sec = 30)
+```
+
+```{r inspecting-peaks-msdata-both, eval=FALSE}
+#access peaks for Diuron-d6 in both positive and negative influent samples using monoisotopic mass instead of mz
+peaks(sp2_pp,
+      analyses = grep("influent", analysisNames(sp2_pp)),
+      mass = diuron_d6, ppm = 10, sec = 15)
+```
+
+```{r inspecting-peaks-msdata-both-table, echo=FALSE}
+peaks(sp2_pp,
+      analyses = grep("influent", analysisNames(sp2_pp)),
+      mass = diuron_d6, ppm = 10, sec = 15)
 ```
@@
 #plotting peaks in msData
-plotPeaks(sp2_pp, mz = targets4[2, ], interactive = TRUE, colorBy = "analyses")
+plotPeaks(sp2_pp, mass = diuron_d6, interactive = TRUE, colorBy = "analyses")
 ```
@@
 ```{r inspecting-peaks-msdata-map}
-#map plot for peak time and m/z dimensions for msData
+#map plot for peak time and m/z dimensions for msData (the shading represents the peak m/z and time deviations)
 mapPeaks(sp2_pp, mz = targets4[2, ], colorBy = "replicates", xlim = 30, ylim = 0.001)
-```
-
-For a _msAnalysis_ object:
+
+#Note, peaks appear for the negative samples but clearly they are not the corresponding peaks of the given target for Diuron-d6.
+For instance, the visual evaluation with mapPeaks is useful to define/optimize settings for grouping peaks across samples.
+```
+
+For a *msAnalysis* object:
 
@@
 A retention time alignment can be applied to correct elution deviations across analyses.
-The result of peak grouping and alignment is added to the _msFeatures_ S4 class object already present in the _msData_.
+The result of peak grouping and alignment is added to the *msFeatures* S4 class object already present in the *msData*.
 
@@
 The processing parameters for grouping and alignment are added similarly to the processing settings for peak picking but the argument `where` is set to "features", as shown below.
-The settings are added to the slot parameters of the _msFeatures_ object and can be obtained by the method `getParameters`, as shown below.
+The settings are added to the slot parameters of the *msFeatures* object and can be obtained by the method `getParameters`, as shown below.
 
 ```{r creating-grouping-parameters}
-param_g <- list(
+
+# with xcms3
+settings_pg_xcms <- createSettings(
+  call = "peakGrouping",
+  algorithm = "xcms3",
+  parameters = list(
+    groupParam = xcms::PeakDensityParam(
+      sampleGroups = "holder",
+      bw = 10,
+      minFraction = 0.6,
+      minSamples = 2,
+      binSize = 0.008,
+      maxFeatures = 100
+    )
+  )
+)
+
+# with openms
+settings_pg_openms <- createSettings(
+  call = "peakGrouping",
+  algorithm = "openms",
+  parameters = list(
+    rtalign = FALSE,
+    QT = FALSE,
+    maxAlignRT = 10,
+    maxAlignMZ = 0.01,
+    maxGroupRT = 4,
+    maxGroupMZ = 0.01,
+    extraOptsRT = NULL,
+    extraOptsGroup = NULL
+  )
+)
+
+# add settings for peakGrouping to slot features in the msData
+sp2_pp <- addSettings(sp2_pp, settings = settings_pg_openms, where = "features")
+
+# get the settings applied/to be apliied for peakGrouping
+getSettings(sp2_pp, where = "features")
+```
+
+### Processing grouping and alignment
+
+The data processing for grouping and alignment follows the same principle as peak picking.
+Either the settings are added as arguments or the added parameters in the *msFeatures* object are used in the function call.
+Below, both cases are demonstrated below.
+
+```{r processing-grouping-run, results='hide', message=FALSE}
+#using the added parameter settings
+sp2_pg <- peakGrouping(object = sp2_pp)
+```
+
+```{r processing-grouping-show}
+#features added to msData
+sp2_pg
+```
+
+```{r processing_grouping-run2, results='hide', message=FALSE}
+#adding the parameter settings as arguments
+sp2_pg_2 <- peakGrouping(sp2_pp, settings = settings_pg_xcms)
+```
+
+```{r processing-grouping-show2}
+#features added to msData
+sp2_pg_2
+```
+
+### msFeatures class
+
+The structure of the *msFeatures* is shown below.
+The slot `analyses` is the `data.table` as obtained by the `analysisTable`, containing basic information about the analyses, such as file full path, replicate and associated blank replicates.
+The intensity slot contains the intensity of the feature in each analysis (i.e., corresponding peak intensity in each analysis).
+The slot metadata holds additional information about each feature, such as average *m/z*/ neutral mass, retention time and number of corresponding peaks.
+The slot settings has the list of settings used to build/process features, such as peakGrouping and filterFeatures.
+
+```{r msfeatures-class}
+str(sp2_pg@features, max.level = 2)
+#Note, accessing directly the features slot in a given msData object is not recommended.
+#Access to features should be done with the applied method, as shown below in inspecting features.
+```
+
+### Inspecting features
+
+Features can be accessed via the method `features` applied to a *msData* object.
+The argument `complete` can be set to `TRUE` for a complete list of features information.
+Also, the argument `average` can be set to `FALSE` for returning the intensity of the feature in each analysis than the average for each analysis replicate.
+
+```{r inspecting-features}
+#getter for feature intensities correspondent to the mass of Diuron in each sample by setting average to false
+features(sp2_pg, mass = diuron_d6, ppm = 10, sec = 20, average = FALSE)
+
+#getter for features intensities and intensity deviations using m/z; note that mass of targets m/z values will be calculated according to present adducts
+features(sp2_pg, mz = targets4)
+
+#getter for all the feature metadata by setting complete to TRUE
+t(features(sp2_pg, mass = carbamazepin_d10, ppm = 10, complete = TRUE, average = TRUE))
+
+#plotting features
+plotFeatures(sp2_pg,
+  mass = rbind(diuron_d6, carbamazepin_d10),
+  ppm = 10, sec = 10, colorBy = "targets",
+  interactive = FALSE)
+
+#plotting features for replicates and interactive
+plotFeatures(sp2_pg,
+  mass = rbind(diuron_d6, carbamazepin_d10),
+  ppm = 10, sec = 10, colorBy = "replicates",
+  interactive = TRUE)
+
+#plot the individual peaks in features and for each analysis shows the time deviation.
+plotFeaturePeaks(sp2_pg, mz = targets4)
+
+#plotFeaturePeaks(sp2_pg, mass = rbind(diuron_d6, carbamazepin_d10))
+```
+
+### Alignment results
+
+The alignment results can be inspected with the `plotAlignment` function.
+However, the time adjustment results are only available when using the algorithm "xcms3" for alignment with polarity positive or negative.
+Mixed positive and negative polarities does not support rt alignment.
+Below peaks from samples in positive mode only are grouped, including retention time alignment.
+
+```{r grouping-msdata-only-positive}
+
+sp2_pg_3 <- sp2_pp[c(4:6, 10:12)]
+
+settings_pg_xcms_2 <- list(
   rtalign = TRUE,
@@
 
-settings_pg <- createSettings(
+settings_pg_xcms_2 <- createSettings(
   call = "peakGrouping",
   algorithm = "xcms3",
-  settings = param_g
-)
-
-# add processing parameter for peakGrouping to features in the msData
-sp2_pp <- addParameters(sp2_pp, settings = settings_pg, where = "features")
-
-# get the processing parameters applied to features
-getParameters(sp2_pp, where = "features")
-```
-
-### Processing grouping and alignment
-
-The data processing for grouping and alignment follows the same principle as peak picking.
-Either the settings are added as arguments or the added parameters in the _msFeatures_ object are used in the function call.
-Below, both cases are demonstrated below.
-
-```{r processing-grouping-run, results='hide', message=FALSE}
-#using the added parameter settings
-sp2_pg <- peakGrouping(sp2_pp)
-```
-
-```{r processing-grouping-show}
-#features added to msData
-sp2_pg
-```
-
-```{r processing_grouping-run2, results='hide', message=FALSE}
-#adding the parameter settings as arguments
-sp2_pg_2 <- peakGrouping(sp2_pp, settings = settings_pg)
-```
-
-```{r processing-grouping-show2}
-#features added to msData
-sp2_pg_2
-```
-
-### msFeatures class
-
-The structure of the _msFeatures_ is shown below.
-The slot `analyses` is the `data.table` as obtained by the `analysisInfo`, containing basic information about the analyses, such as file, replicate and associeted blank.
-The intensity slot has the intensity of the feature in each analysis (i.e., peak intensity in each analysis).
-The slot metadata has other information about each feature, such as average _m/z_ and retention time.
-
-`r # TODO add annotation description`
-
-The slot parameters has the list of settings used to obtain/process the features.
-
-```{r msfeatures-class}
-#accessing directly the msFeatures slot, not recommended
-str(sp2_pg@features, max.level = 2)
-```
-
-### Inspecting features
-
-Features can be accessed via the method `features` applied to a _msData_ object.
-The argument `complete` can be set to `TRUE` for a complete list of features information.
-Also, the argument `average` can be set to `FALSE` for returning the intensity of the feature in each analysis.
-
-```{r inspecting-features}
-#getter for feature intensities and intensity deviations
-features(sp2_pg, mz = targets4)
-
-#getter for all the feature metadata by setting complete to TRUE
-t(features(sp2_pg_2, mz = targets4[1, ], complete = TRUE))
-
-#getter for features with intensities for each sample by setting average to FALSE
-features(sp2_pg_2, mz = targets4[1, ], average = FALSE)
-
-#plotting features
-plotFeatures(sp2_pg, mz = targets4, colorBy = "targets", interactive = FALSE)
-
-#plotting features for replicates and interactive
-plotFeatures(sp2_pg, mz = targets4, colorBy = "replicates", interactive = TRUE)
-
-#plot the individual peaks in features and for each analysis shows the time deviation.
-plotFeaturePeaks(sp2_pg, mz = targets4)
-```
-
-### Alignment results
-
-The alignment results can be inspected with the `plotAlignment` function. However, the time adjustment results are only available when using the algorithm "xcms3" for alignment.
+  parameters = settings_pg_xcms_2
+)
+
+sp2_pg_3 <- peakGrouping(sp2_pg_3, settings = settings_pg_xcms_2)
+
+peaks(sp2_pg_3)
+
+features(sp2_pg_3, mass = targets4, ppm = 10, sec = 10, complete = TRUE)
+
+plotAlignment(sp2_pg_3)
+```
 
 ```{r plot-alignment}
-plotAlignment(sp2_pg)
+plotAlignment(sp2_pg_3)
 ```
@@
 Below, an example of the recursive integration workflow using the function [fillChromPeaks](https://rdrr.io/bioc/xcms/man/fillChromPeaks.html) from the package [xcms](https://bioconductor.org/packages/release/bioc/html/xcms.html) is shown.
-The function `peakFilling` is used for recursive integration and a _settings_ S4 class object is given to define the algorithm and respective parameters.
-A _settings_ object can be obtained with `fillingSettingsDefaultXCMS` for a default _ChromPeakAreaParam_ from xcms.
-Note that the peakFilling settings are added to the parameters slot of the _msData_ object.
+The function `peakFilling` is used for recursive integration and a *settings* S4 class object is given to define the algorithm and respective parameters.
+A *settings* object can be obtained with `fillingSettingsDefaultXCMS` for a default *ChromPeakAreaParam* from xcms.
+Note that the peakFilling settings are added to the parameters slot of the *msData* object.
 
@@
 #filling peaks for msData
-sp2_pg_fl <- peakFilling(sp2_pg, settings = settings_fl)
+sp2_pg_3_fl <- peakFilling(sp2_pg_3, settings = settings_fl)
 ```
@@
 # inspecting the filled peaks for a given mz and rt target
-plotFeaturePeaks(sp2_pg_fl, mz = 441.1670, rt = 1097, ppm = 20, sec = 60)
+plotFeaturePeaks(sp2_pg_3_fl, mz = 441.1670, rt = 1097, ppm = 20, sec = 60)
 ```
@@
   algorithm = "camera",
-  settings = list(
+  parameters = list(
     ionization = NA_character_,
@@
 ```{r annotation-run, results='hide', message=FALSE}
-sp2_pg_fl_an <- peakAnnotation(sp2_pg_fl, settings = settings_an)
-```
-
-The results of annotation are stored under annotation of the _msFeatures_ S4 class and matches the [components](https://rickhelmus.github.io/patRoon/reference/components-class.html) S4 class of the [patRoon](https://github.com/rickhelmus/patRoon) package.
-This means that the feature components can be accessed via the `getAnnotation` method for an _msData_ object and inspected also with native methods from patRoon.
+sp2_pg_3_fl_an <- peakAnnotation(sp2_pg_3_fl, settings = settings_an)
+```
+
+The results of annotation are stored under annotation of the *msFeatures* S4 class and matches the [components](https://rickhelmus.github.io/patRoon/reference/components-class.html) S4 class of the [patRoon](https://github.com/rickhelmus/patRoon) package.
+This means that the feature components can be accessed via the `getAnnotation` method for an *msData* object and inspected also with native methods from patRoon.
 
@@
 
-The annotation can be inspected with the `annotation` S4 method for the _msData_ object, as shown below.
+The annotation can be inspected with the `annotation` S4 method for the *msData* object, as shown below.
 The method `plotAnnotation` can then be used to visualize the annotation of features, as shown below.
@@
 #isotopic and adduct features annotated with a given target mz and rt
-streamFind::annotation(sp2_pg_fl_an, mz = targets4[1, ], ppm = 20, rt = NULL, sec = 60, all = FALSE)
+annotation(sp2_pg_3_fl_an, mz = targets4[1, ], ppm = 20, rt = NULL, sec = 60, all = FALSE)
 
 #plotting the annotation of each feature in a given target
-plotAnnotation(sp2_pg_fl_an, mz = targets4[1, ], ppm = 20, rt = NULL, sec = 60, all = FALSE)
-```
-
-`r # TODO improve annotation of features/peaks``
+plotAnnotation(sp2_pg_3_fl_an, mz = targets4, ppm = 20, rt = NULL, sec = 60, all = FALSE, colorBy = "isotopes")
+```
 
@@
 
-The first is based on data sub-setting using the method `[` and is permanent, meaning that the peaks/features are completely removed from the _msAnalysis_/_msFeatures_.
+The first is based on data sub-setting using the method `[` and is permanent, meaning that the peaks/features are completely removed from the *msAnalysis*/*msFeatures*.
 An example is shown below for data filtration using the `[` method for different classes.
@@
 #sub-setting the msData class on analyses and features (the 3 last analyses and features annotated with a target)
-target_s <- streamFind::annotation(sp2_pg_fl_an, mz = targets4[1, ], ppm = 20, rt = NULL, sec = 5, all = FALSE)
-sp2_pg_fl_an[4:6, target_s$id]
+target_s <- annotation(sp2_pg_3_fl_an, mz = targets4[1, ], ppm = 20, rt = NULL, sec = 5, all = FALSE)
+sp2_pg_3_fl_an[4:6, target_s$id]
 
 #sub-seeting the msData class on analyses ([i, ]) and features ([, j]) by index
-sp2_pg_fl_an[1:2, 1:5]
+sp2_pg_3_fl_an[1:2, 1:5]
 
@@
 
-The second method uses filter functions (i.e., `filterPeaks` and `filterFeatures`) and is conservative, meaning that the data is conserved but tagged as filtered (i.e., the _filtered_ column is changed to `TRUE` and the respective filter tag is added to column _filter_, as shown below).
+The second method uses filter functions (i.e., `filterPeaks` and `filterFeatures`) and is conservative, meaning that the data is conserved but tagged as filtered (i.e., the *filtered* column is changed to `TRUE` and the respective filter tag is added to column *filter*, as shown below).
 Both filtered peaks and features can be permanently removed by running the methods `removeFilteredPeaks` or `removeFilteredFeatures`, respectively.
@@
 When removing peaks, the features left without peaks representation are removed as well.
-Also, the methods `peaks` and `features` have the argument _filtered_ to return filtered peaks/features.
-The default for the _filtered_ argument is `TRUE`, meaning that filtered peaks/features are returned by default for methods `peaks` and `features`.
-Simmilarly, to peakPicking, peak Grouping, etc., the filter methods use a _msSettings_ S4 class to store the filtering parameters.
-Also, the filtering _msSettings_ object when applied is stored in the parameters section opf the respective class (i.e., `filterPeaks` settings are stored in the _msAnalysis_ and `filterFeatures` settings are stored in the _msFeatures_).
+Also, the methods `peaks` and `features` have the argument *filtered* to return filtered peaks/features.
+The default for the *filtered* argument is `TRUE`, meaning that filtered peaks/features are returned by default for methods `peaks` and `features`.
+Simmilarly, to peakPicking, peak Grouping, etc., the filter methods use a *msSettings* S4 class to store the filtering parameters.
+Also, the filtering *msSettings* object when applied is stored in the parameters section opf the respective class (i.e., `filterPeaks` settings are stored in the *msAnalysis* and `filterFeatures` settings are stored in the *msFeatures*).
 
@@
   algorithm = "filter",
-  settings = filters
-)
-
-sp2_pg_fl_an_filFeat <- addParameters(sp2_pg_fl_an, where = "features", settings = settings_filFeat)
-
-sp2_pg_fl_an_filFeat <- filterFeatures(sp2_pg_fl_an_filFeat)
+  parameters = filters
+)
+
+sp2_pg_3_fl_an_filFeat <- addSettings(sp2_pg_3_fl_an, where = "features", settings = settings_filFeat)
+
+sp2_pg_3_fl_an_filFeat <- filterFeatures(sp2_pg_3_fl_an_filFeat)
 
 #number of features not filtered
-nrow(features(sp2_pg_fl_an, filtered = FALSE))
+nrow(features(sp2_pg_3_fl_an, filtered = FALSE))
 
 #number of features not filtered after applying the filters
-nrow(features(sp2_pg_fl_an_filFeat, filtered = FALSE))
+nrow(features(sp2_pg_3_fl_an_filFeat, filtered = FALSE))
 
 #remove permanently all filtered features from the msData object
-sp2_pg_fl_an_filFeat_rem <- removeFilteredFeatures(sp2_pg_fl_an_filFeat, which =  "all")
-
-sp2_pg_fl_an_filFeat_rem
+sp2_pg_3_fl_an_filFeat_rem <- removeFilteredFeatures(sp2_pg_3_fl_an_filFeat, which =  "all")
+
+sp2_pg_3_fl_an_filFeat_rem
 ```
@@
 Below an example of settings used for accessing and averaging MS2 data is given.
-The resulting _settings_ object is then used further in the call to `MS2s` or `plotMS2s`.
+The resulting *settings* object is then used further in the call to `MS2s` or `plotMS2s`.
 
@@
   algorithm = "streamFind",
-  settings = param_msn
+  parameters = param_msn
 )
@@
 When using streamFind basic workflow, the use of `MS2s` and `plotMS2s` is recommended.
-All the S4 methods/function can be applied for _msAnalysis_ and _msData_ objects.
+All the S4 methods/function can be applied for *msAnalysis* and *msData* objects.
 
 ```{r extract-ms2}
-#head of MS2 data for target 1 in a msAnalysis object
-head(MS2s(a1, mz = targets4[1, ], settings = settings_msn))
-
-#ploting the MS2 data of the first target in a msAnalysis object
-plotMS2s(a1, mz = targets4[1, ], settings = settings_msn, interactive = FALSE)
-
-#most intense MS2 data traces for both targets in a msData object
-s_data_3 <- MS2s(sp1, mz = targets4, settings = settings_msn)
-s_data_3 <- s_data_3[order(intensity, decreasing = TRUE), ]
-head(s_data_3)
-
-#ploting the MS2 data of the first target in a msAnalysis object
-plotMS2s(sp1, mz = targets4, settings = settings_msn, interactive = TRUE, colorBy = "targets")
+# #head of MS2 data for target 1 in a msAnalysis object
+# head(MS2s(a1, mz = targets4[1, ], settings = settings_msn))
+# 
+# #ploting the MS2 data of the first target in a msAnalysis object
+# plotMS2s(a1, mz = targets4[1, ], settings = settings_msn, interactive = FALSE)
+# 
+# #most intense MS2 data traces for both targets in a msData object
+# s_data_3 <- MS2s(sp1, mz = targets4, settings = settings_msn)
+# s_data_3 <- s_data_3[order(intensity, decreasing = TRUE), ]
+# head(s_data_3)
+# 
+# #ploting the MS2 data of the first target in a msAnalysis object
+# plotMS2s(sp1, mz = targets4, settings = settings_msn, interactive = TRUE, colorBy = "targets")
 ```
@@
 
-` # TODO create function for collecting MS2 data for peaks/features`
+`# TODO create function for collecting MS2 data for peaks/features`
 
@@
 ```{r democode, echo=FALSE, eval=FALSE, include=FALSE}
+
+# param_g <- list(
+#   rtalign = TRUE,
+#   loadRawData = TRUE,
+#   groupParam = xcms::PeakDensityParam(
+#     sampleGroups = "holder",
+#     bw = 3,
+#     minFraction = 0.6,
+#     minSamples = 2,
+#     binSize = 0.008,
+#     maxFeatures = 100),
+#   preGroupParam = xcms::PeakDensityParam(
+#     sampleGroups = "holder",
+#     bw = 5,
+#     minFraction = 1,
+#     minSamples = 3,
+#     binSize = 0.008,
+#     maxFeatures = 100),
+#   retAlignParam = xcms::PeakGroupsParam(
+#     minFraction = 1,
+#     extraPeaks = 0,
+#     smooth = "loess",
+#     span = 0.3,
+#     family = "gaussian")
+# )
 
@@
 
-***
-Insitut für Energie- und Umwelttechnik e.V. (IUTA)  
-<br>
-For further questions, please contact Ricardo Cunha ([cunha@iuta.de](cunha@iuta.de)).
+------------------------------------------------------------------------
+
+Insitut für Energie- und Umwelttechnik e.V. (IUTA)\
+<br> For further questions, please contact Ricardo Cunha ([cunha\@iuta.de](cunha@iuta.de)).

### 2022-10-17 R/S4_class_003_msFeatures.R

@@ -54,14 +54,14 @@ setClass("msFeatures",
            intensity = "data.table",
            metadata = "data.table",
            annotation = "list",
-           parameters = "list"
+           settings = "list"
          ),
          prototype = list(
            analyses = data.table(),
            intensity = data.table(),
            metadata = data.table(),
            annotation = list(),
-           parameters = list()
+           settings = list()
          ),
          validity = msFeatures_validity
 )
@@ -102,17 +102,17 @@ setMethod("replicateNames", "msFeatures", function(object) object@analyses$repli
 setMethod("blankReplicateNames", "msFeatures", function(object) object@analyses$blank)
 
 
-### addParameters ----------------------------------------------------------
+### addSettings ----------------------------------------------------------
 
-#' @describeIn msFeatures adds processing parameters for features.
+#' @describeIn msFeatures adds processing settings for features.
 #'
 #' @template args-single-settings
 #'
 #' @export
 #'
-#' @aliases addParameters,msFeatures,msFeatures-method
+#' @aliases addSettings,msFeatures,msFeatures-method
 #'
-setMethod("addParameters", "msFeatures", function(object, settings) {
+setMethod("addSettings", "msFeatures", function(object, settings) {
 
   valid <- checkmate::testClass(settings, "settings")
 
@@ -121,27 +121,27 @@ setMethod("addParameters", "msFeatures", function(object, settings) {
     return(object)
   }
 
-  object@features@parameters[[getCall(settings)]] <- settings
+  object@features@settings[[getCall(settings)]] <- settings
 
   return(object)
 })
 
-### getParameters ----------------------------------------------------------
+### getSettings ----------------------------------------------------------
 
-#' @describeIn msFeatures gets processing parameters.
+#' @describeIn msFeatures gets processing settings.
 #'
 #' @param call The call name of the settings to retrieve.
 #'
 #' @export
 #'
-#' @aliases getParameters,msFeatures,msFeatures-method
+#' @aliases getSettings,msFeatures,msFeatures-method
 #'
-setMethod("getParameters", "msFeatures", function(object, call = NULL) {
[…]

### 2022-10-17 R/fct_ms_007_filterFeatures.R

@@ -64,7 +64,7 @@ filterFeatures <- function(object, settings = NULL) {
 
 
   if (is.null(settings)) {
-    prs <- getParameters(obj, call = "filterFeatures")
+    prs <- getSettings(obj, call = "filterFeatures")
 
     if (length(prs) > 0) {
       if (length(prs) > 1) {
@@ -72,7 +72,7 @@ filterFeatures <- function(object, settings = NULL) {
       }
 
       algorithm = getAlgorithm(prs)
-      filterList = getSettings(prs)
+      filterList = getParameters(prs)
 
     } else {
 
@@ -82,7 +82,7 @@ filterFeatures <- function(object, settings = NULL) {
   } else if (testClass(settings, "settings")) {
 
     algorithm <- getAlgorithm(settings)
-    filterList <- getSettings(settings)
+    filterList <- getParameters(settings)
 
   } else {
 

### 2022-10-17 R/utils_ms_data_access.R

@@ -736,7 +736,7 @@ mzXML_loadSpectraInfo <- function(fl, rtr = NULL, levels = NULL) {
 
   xml_data <- read_xml(fl)
 
-  # TODO xcms error when removing empty septra
+  # TODO xcms error when removing empty septra, due to different scan numbers
   #scan_n <- xml_find_all(xml_data, xpath = "//d1:scan[@peaksCount>0]")
 
   scan_n <- xml_find_all(xml_data, xpath = "//d1:scan")
@@ -1142,7 +1142,7 @@ extractMSn <- function(
   if (checkmate::testClass(settings, "settings")) {
     call <- getCall(settings)
     algorithm <- getAlgorithm(settings)
-    settings <- getSettings(settings)
+    settings <- getParameters(settings)
 
   } else {
     warning("Invalid settings for getting MSn data!")

### 2022-10-18 handbooks/Handbook_backend_streamFind.Rmd
LFU19

@@
 ---
-title: "Handbook streamFind: backend"
+title: "Handbook streamFind: back end"
 author: "Ricardo Cunha"
@@
 
-The streamFind package is a backend (code based) and frontend ([shiny](https://shiny.rstudio.com/) app) platform for assembly of modular workflows to support spectrometric and spectroscopic data processing.
+The streamFind package is a back end (code based) and frontend ([shiny](https://shiny.rstudio.com/) app) platform for assembly of modular workflows to support spectrometric and spectroscopic data processing.
 The major focus of streamFind is data processing for environmental and quality studies.
@@
 The streamFind package can be used by academics but also by technicians due to the comprehensive documentation and well categorized set of integrated functionalities (modules).
-This handbook focuses on describing and demonstrating the backend of the streamFind package.
+This handbook focuses on describing and demonstrating the back end of the streamFind package.
 
@@
 Additionally, MS1 and MS2 traces below 1 count were removed to exclude empty spectra.
-The files are used within the document for demonstration of the backend.
+The files are used within the document for demonstration of the back end.
 
 ```{r resource-files, cache=FALSE}
-
-r_path <- system.file(package = "streamFindData", dir = "extdata")
-#r_path <- system.file(package = "streamFind", dir = "extdata")
-
-files <- list.files(r_path, pattern = ".mzML|.mzXML", full.names = TRUE)
+# 
+# r_path <- system.file(package = "streamFindData", dir = "extdata")
+# #r_path <- system.file(package = "streamFind", dir = "extdata")
+# 
+# files <- list.files(r_path, pattern = ".mzML|.mzXML", full.names = TRUE)
 
 #list of file full paths from the data package
-#files <- streamFindData::msFilePaths()
+files <- streamFindData::msFilePaths()
 ```
@@
 ```{r description-files, echo=FALSE}
-df_files_desc <- data.frame(
-  N. = seq_len(length(files)),
-  file_type = rep("MS", 27),
-  file_name = basename(files),
-  device = c(rep("Agilent Q-TOF", 27)),
-  data_type = c(rep("centroid", 6), rep("profile", 3), rep("centroid", 18)),
-  ac_mode = c(rep("MS/MS", 27)),
-  polarity = c(rep("positive", 9),
-               rep("negative", 3), rep("positive",3),
-               rep("negative", 3), rep("positive",3),
-               rep("negative", 3), rep("positive",3)),
-  description = c(
-    rep("Basic centroided MS data as mzML spiked with chemical 
-        and internal standards (CS and IS, respectively).", 3),
-    rep("Basic centroided MS data as mzXML spiked with chemical 
-        and internal standards (CS and IS, respectively).", 3),
-    rep("Basic profile MS data as mzML spiked with CS and IS.", 3),
-    rep("Blank MS data as mzML spiked with IS.", 6),
-    rep("Wastewater secondary effluent MS data as mzML spiked with IS.", 6),
-    rep("Wastewater secondary effluent treated with ozonated 
-        strong water MS data as mzML spiked with IS.", 6)
-  )
-)
+df_files_desc <- streamFindData::msFilesDescription()
 
@@
 ```{r load-table-chemicals}
-db <- paste0(r_path, "/spiked_chemicals_hrms.csv")
-db <- data.table::fread(db)
+db <- streamFindData::msSpikedChemicals()
 ```
@@
 ```{r show-table-chemicals, echo=FALSE}
-db$ionization <- "positive"
-db[, `:=`("mz_pos" = mass + 1.0073, "mz_neg" = mass - 1.0073 )]
-db[grepl("neg", db$comment), ionization := "both"]
-data.table::setnames(db, "mix_id", "tag", skip_absent = TRUE)
-db[tag %in% 1, tag := "CS"]
-db[, in_file := "1-6"]
-db[tag %in% "IS", in_file := "1-24"]
-
-knitr::kable(db[, .(name, formula, mass, rt, ionization, tag, in_file)], caption = "Chemical standards spiked.") %>%
+knitr::kable(db[, .(name, CAS, formula, mass, rt, ionization, tag, in_file)], caption = "Chemical standards spiked.") %>%
   kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>%
@@
 
-(...) `r # TODO add chromatograms entry and make method`
+(...)
 
@@
 The right column in Figure \@ref(fig:workflow-scheme illustrates the implementation of modular functionalities for assembly of data processing workflows.
-The backend for modular assembly and framework of workflows in streamFind is described and demonstrated in section \@ref(workflows).
+The back end for modular assembly and framework of workflows in streamFind is described and demonstrated in section \@ref(workflows).
 

### 2022-10-29 dev/template_report.Rmd
LFU20

--- 2022-07-01 Handbook_backend_streamFind.Rmd
@@
 ---
-title: "Handbook streamFind: Backend"
-author: "Ricardo Cunha"
-date: '2022-06-15'
+title: "ntsIUTA Report Template"
+author: Ricardo Cunha
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output:
@@
     theme: paper
+    code_folding: hide
+    includes:
+      in_header: !expr noquote(paste0(system.file(package = "streamFindData", dir = "extdata"), "\\logo_iuta_template_report.html"))
+editor_options: 
+  chunk_output_type: console
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 9) #results = "asis"
+
+knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 9, results = "asis")
+path <- getwd()
+
 ```
@@
 library(plotly)
-
-devtools::load_all()
+library(streamFind)
 
 ```
-
-# Objective
-
-# Install
-
-## Install dependencies
-
-## Install streamFind
-
-# Resources
-
-Trimmed example files are included in the streamFind package and are loaded below.
-The trimming was done with the function `trimSpectraFilesMZR` from streamFind which uses the package [mzR](https://bioconductor.org/packages/release/bioc/html/mzR.html).
-
-```{r resource_files}
-r_path <- system.file(package = "streamFind", dir = "extdata")
-files <- list.files(r_path, full.names = TRUE)
-```
-
-# S4 classes
-
-The streamFind package uses S4 class objects to define the data structure.
-The S4 classes, respective methods and usage are presented and exemplified through a demonstration of the basic workflow (pre-processing) for HRMS data.
-In short, the basic pre-processing workflow includes the following recommended steps:  (1) project creation, (2) data conversion (when needed), (3) replicate analyses assignment, (4) peak finding (or peak picking), alignment, grouping and isotopic and adduct annotation across analyses and (5) feature quality evaluation and (6) filtering.
-
-The semantics in streamFind package is described in details in section x. Yet, a brief summary is given below for the most relevant terms:  
-  * __trace__ correspond to a raw data signal (e.g., mass trace in MS data) obtained from a given analysis file;  
-  * __peak__ correspond to a cluster of traces (e.g., integrated chromatographic MS peak) within an analysis file;  
-  * __feature__ correspond to a group of corresponding peaks across analysis files;
-
-## streamProject
-
-The project creation is demonstrated below for three HRMS files (triplicate) where a set of substances were spiked.
-The project path can be assign as argument but left as default (i.e., the working directory) for this example.
-The initial S4 class created in the `newStreamProject` function is _streamProject_.
-
-```{r newStreamProject_call, message=FALSE, results='hide'}
-exS1 <- newStreamProject(files = files[1:3])
-```
-
-## msData
-
-Because the added files were all with mzML format, the sub-class is directly defined as _msData_, adding additional structure for further processing of MS data.
-Other data formats/structure will lead to different sub-classes (e.g., _uvData_ and _ramanData_) but is not yet implemented within streamFind.
-
-```{r msData_streamProject_classes}
-is(exS1)
-```
-
-The _msData_ structure includes the _streamProject_ class plus the `features` slot.
-The project, date and path slots correspond to the project title, creation date and directory in disk, respectively.
-The analyses slot is a list with a _msAnalysis_ S4 class object per MS file added.
-The structure of the _msAnalysis_ class is described in the next sub-chapter.
-
-The `features` `r # TODO make msFeatures after peaks extraction and storage`` 
-which includes the _msFeatures_ S4 class(presented in section X).
-
- * _msData_ structure:
-
-```{r structure_msData}
-str(exS1, max.level = 2)
-```
-
-## msAnalysis
-
-The _msAnalysis_ class is structured as shown below.
-Note that the S4 method `getAnalyses` for _msData_ was used to obtain the _msAnalysis_ of the first analysis as defined by the second argument.
-In general, access and plotting S4 methods can also be applied directly to the _msAnalysis_ class besides _msData_ (as described in \@ref(data-access-and-visualization)) but is not part of the basic workflow and mostly used for advanced data visualization and access of low level data (e.g., traces).
-
-The slots analysis, file and replicate correspond to the analysis name, the full file path and the replicate group name, respectively.
-The metadata slot is a list of objects with various informative entries of the analysis.
-The metadata list can be expanded with additional information about the analysis/sample (e.g., sample location, weather conditions, process sensor data, etc.).
-Below, the `getMetadata` S4 method of _msData_ is used to display the metadata of the first file.
-Note, `getMetadata` also works for _msAnalysis_ S4 class as shown below in the commented code.
-The slot parameters is used to store processing steps (e.g., peak picking) as _settingsSet_ S4 class objects, including the processing step name/ID, the algorithm used and the respective settings applied.
-The parameters can be applied for history track but also for re-run the processing when required.
-The spectra and chromatograms slots are used to store low level data (i.e., mass traces) from the MS files.
-Finally, the peaks slot is used to store the chromatographic peaks after peak picking.
-As no other processing step was applied the spectra, chromatograms and peaks tables are empty for now.
-
- * _msAnalysis_ structure:
-
-```{r msAnalysis_structure}
-exA1 <- getAnalyses(exS1, 1)
-str(exA1, max.level = 2)
-```
-
-* metadata of _msAnalysis_:
-
-```{r metadata_structure}
-t(metadata(exS1, analyses = 1))
-
-#applied directly to the msAnalysis S4 class
-t(metadata(exA1))
-```
-
-# Low level data access and plotting
-
-The tools for low level access and plotting (i.e., traces and chromatograms) share a target search approach.
-The data search is based on pre-defined time and mass (inc.
-_m/z_) targets.
-For traces (i.e., raw MS mass traces), time (seconds) and _m/z_ (Da) ranges are given for accessing and plotting.
-Access to mass traces is relevant for further development of data processing functionalists and for visualization of raw data to support evaluation of processing steps (e.g., peaks wrongly not found during peak picking).
-
-## Defining targets
-
-As aforementioned, access to raw data is based on defining time and _m/z_ targets.
-The function `makeTargets` is used for comprehensively build targets.
-As shown below, diverse ways can be used to assemble targets, similarly returning a uniform `data.table`.
-Accessing and plotting S4 methods in streamFind use the `makeTargets` function to collect data.
-Therefore, the arguments `mz`, `rt`, `ppm`, `sec` and `id` are present in most S4 methods for accessing and plotting data as demonstrated in the following sections.
-
-```{r make_targets}
-#case 1
-mz01 <- c(247.1651, 239.0628)
-rt01 <- c(839, 937)
-id01 <- c("target1", "target2")
-targets01 <- makeTargets(mz = mz01, rt = rt01, ppm = 20, sec = 60, id = id01)
-
-targets01
-
-#case 2
-mz02 <- data.frame(mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652))
-rt02 <- data.frame(rtmin = c(809, 907), rtmax = c(869, 967))
-targets02 <- makeTargets(mz = mz02, rt = rt02)
-
-targets02
-
-#case 3
-mz03 <- data.frame(
-  #id = c("target1", "target2"),
-  mz = c(247.1651, 239.0628),
-  rt = c(839, 937)
-)
-targets03 <- makeTargets(mz = mz03, ppm = 20, sec = 60)
-
-targets03
-
-#case 4
-mz04 <- data.frame(
-  id = c("target1", "target2"),
-  mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652),
-  rtmin = c(809, 907), rtmax = c(869, 967)
-)
-targets04 <- makeTargets(mz = mz04)
-
-targets04
-```
-
-## EICs
-
-`r # TODO remove zeros when returning EIC table?`
-`r # TODO add check for spectra/chroms is the object before querying the data?`
-
-Extracted ion chromatograms (EICs) are obtained either from loaded spectra in the _msData_ or _msAnalysis_ objects or from querying traces directly in the raw data file (mzML or mzXML).
-The latter is less disk memory intense as data is loaded from raw files on demand.
-The S4 method `EICs` is used to extract MS1 data from raw data files.
-
-```{r EICs_method}
-#EIC of targets from first analysis using msAnalysis class
-eic01 <- EICs(exA1, mz = targets04)
-head(eic01, 6)
-
-#EIC of targets from second analyses using msData class
-eic02 <- EICs(exS1, analyses = 2, mz = targets04)
-head(eic02, 6)
-```
-
-The obtained EICs can be plotted with `plotEICs` S4 method, as shown below.
-The object for `plotEICs` can be a `data.table` as resulting from the `EICs` method or the actual _msData_ or _msAnalysis_ objects, using the same arguments as in `EICs` for collecting the data (i.e., to define the targets).
-
-```{r plotEICs_static}
-#plot the produced EIC data.table
-plotEICs(eic01)
-```
-
-```{r plotEICs_interactive}
-#using the msData class object and targets but with interactive plot
-plotEICs(exS1, analyses = NULL, mz = targets04, colorBy = "analyses", interactive = TRUE)
-
-#the same plot as the first for data.table but interactive
-plotEICs(exA1, mz = targets04, interactive = TRUE)
-```
-
-## TICs
-
-Total ion chromatograms (TICs) are obtained either by pre-loading and adding the chromatogram data to the _msAnalysis_, extracting the data in a `TICs` call (S4 method) or loading all spectra and applying the sum of intensities to each spectrum.
-Note, TIC is already available as chromatogram in mzML data from the major HRMS vendors.
-The S4 method TICs first evaluates if the TIC chromatogram is available in the mzML and when not, the spectra data is extracted and used instead.
-
-```{r TICs_method}
-# TIC for both analyses in the msData class object
-tic01 <- TICs(exS1)
-tic01
-
-#TIC of the msAnalysis object
-head(TICs(exA1), 6)
-```
-
-Similarly to EICs, the TICs can be plotted with `plotTICs` S4 method, as shown below.
-
-```{r plotTICs_method}
-#using a data.table as obatined by TICs
-plotTICs(tic01, interactive = TRUE)
-
-#using the msData object
-plotTICs(exS1, colorBy = "replicates")
-
-#using the msAnalysis
-plotTICs(exA1)
-```
-
-## XICs
-
-A more informative method of obtaining EICs is via the method `XICs` which extract a the three dimensional (_m/Z_, rt and intensity) chromatograms from given targets.
-In the example below, the _m/z_ dimension is also included in the resulting table to improve inspection of traces.
-
-```{r XICs_method}
-#using the msData class object
-xic01 <- XICs(exS1, mz = targets04)
-xic01
-
-#with a msAnalysis class object
-xic02 <- XICs(exA1, mz = mz01, rt = rt01, ppm = 20, sec = 60)
-head(xic02, 3)
-```
-
-Similarly, the XICs can be plotted with the S4 method `plotXICs`, as shown below.
-Other arguments are available for plotting `XICs`, including a target area for the expected targets.
-As shown in the plot below, the red mark is defined by the arguments `targetsMark` (a table with columns mz and rt to define the expected _m/z_ and retention times) and `ppmMark` and `secMark` to define the mass (in ppm) and time (in seconds) range of the target squares.
-The `numberRows` is used to control the number of rows of the grouped plot.
-The concept for the XIC plot is adapted from the R package [MSnbase](https://bioconductor.org/packages/release/bioc/html/MSnbase.html).
-
-```{r plotXICs_method}
-#using the output of XICs S4 method (a data.table)
-plotXICs(xic01, legendNames = c("target number 1", "target number 2"), plotTargetMark = TRUE,
- targetsMark = targets04[, c("mz", "rt")], ppmMark = 5, secMark = 10, numberRows = 2)
-
-#using the msData class and plotting the first target from the first analysis
-plotXICs(exS1[1], mz = targets04[1, ])
-
-#for the msAnalysis, the second target with wide m/z and rt ranges
-plotXICs(exA1, mz = mz01[2], rt = rt01[2], ppm = 500, sec = 120)
-```
-
-`r # TODO add MSn low level access`
-
-# Basic workflow
-
-The basic workflow consists of essential steps for processing of raw data.
-
-`r # TODO make a scheme of the basic workflow, maybe from the PPT`
-
-## Project creation
-
-The project creation is demonstrated below for six HRMS data files corresponding to blank and wastewater influent samples, both measured in triplicate.
-The argument `files` is the main input and can either be a list of full path files, a table with file (i.e., full file path), replicate (i.e., the name of the analysis replicate group for each file) and blank (i.e., the name of the associated blank analysis replicate group for each file), or the `analysisInfo` data.frame from the package [patRoon](https://github.com/rickhelmus/patRoon).
-The project path can be assign as argument `path` and is the directory where all the project files (i.e., scripts, cache databases, objects and results) will be stored.
-
-```{r newStreamProject_basic_workflow, message=FALSE, results='hide'}
-exS2 <- newStreamProject(files = files[4:9], path = getwd(), title = "Project Example")
-```
-
-## Assign replicate names
-
-An essential aspect of environmental analysis is to operate with sample/analysis replicates.
-The replicates in _msData_ or _msAnalysis_ can be obtained by the `replicates` S4 method as shown below.
-The setter (assignment) for the replicate names is demonstrated below with the method `replicates<-` that takes a vector of character strings with the same length as the number of analyses in a _msData_ object or length one for a single analysis in _msAnalysis_.
-Alternatively, a table with a replicate character column or a character vector can be given in the files or replicates arguments, respectively.
-
-```{r assign_replicates}
-#getter for replicate names in msData
-replicates(exS2)
-
-#setter for replicate names in msData
-replicates(exS2) <- c(rep("blank", 3),rep("influent", 3))  
-replicates(exS2)
-
-#setter for replicate name in msAnalysis
-exA2 <- getAnalyses(exS2, 4)
-replicates(exA2) <- "wastewater"
-
-#getter for replicate name in msAnalysis
-replicates(exA2)
-```
-
-## Assign blanks
-
-The blank subtraction is another crucial aspect in environmental analysis.
-The getter and setter for blank subtraction works the same way as replicates.
-See examples below.
-The blank analysis replicate is then assign to the respective analysis.
-In the example, the first replicate group is assign to the influent samples.
-Note that the blank will also be assigned to itself.
-Although not shown, different blank analysis replicates can be assigned to different analysis.
-The subtraction is then applied per analysis considering the assigned blank replicate.
-Getting and assigning blank replicates from an _msAnalysis_ is also possible.
-However, assigning a blank without the context of a _streamProject_ and without the structure of _msData_ or similar level classes is not recommended.
-For cross project analysis, a subset of the analyses (_msAnalysis_ objects) in a given _msData_ can be concatenated to another _msData_, carrying the assigned blank replicate for each analysis as well as all the other analysis information (see section cross-project analysis for more information).
-
-```{r assign_blanks}
-#getter for blank replicates in each sample in a msData object
-blanks(exS2)
-
-#setter for the blank analysis replicate of msData object
-blanks(exS2) <- rep("blank", 6)
-blanks(exS2)
-
-#getter for msAnalysis
-blanks(exA2)
-```
-
-## Other getter S4 methods
-
-```{r other_S4_methods}
-#getter for analysis names in msdata
-analyses(exS2)
-
-#getter for analysis name in msAnalysis
-analyses(exA2)
-
-#getter for polarity of analyses in a msData
-polarities(exS2)
-
-#getter for polarity of an msAnalysis
-polarities(exA2)
-
-#subsetting an msData with `[`-method for the blank samples
-exS2[1:3]
-```
-
-## Peak picking
-
-The initial processing step of the basic workflow is the peak picking.
-Yet, other steps might be necessary prior the pick picking, such as data conversion, data calibration, etc.
-These are discussed later in section X.
-The basic workflow in streamFind is based on [patRoon](https://github.com/rickhelmus/patRoon) and most of the output in streamFind can be used/converted to enable the use of native [patRoon](https://github.com/rickhelmus/patRoon) functions and methods.
-
-### Processing parameters
-
-For each processing steps, parameter settings are often used.
-In streamFind, the parameters used for a given processing step are added/stored as _settings_ S4 class objects in each _msAnalysis_ (or other class of the same level, such as _uvAnalysis_ or _ramanAnalysis_).
-Each _settings_ object contains the algorithm and list of parameters with the respective values.
-The example below demonstrates the workflow to perform peak picking either by adding the parameters during the call for `peakPicking` or by initially add the parameters to each _msAnalysis_ and then run the `peakPicking`, which looks for the parameters in each _msAnalysis_.
-If different settings are applied among the _msAnalysis_ objects, these are used.
-This means that different processing parameters can be applied within the same function call.
-If parameters are added during the function call but they already exist in the _msAnalysis_, these are overwritten by the settings used in the function call.
-
-### Create and add parameters
-
-The function `createSettings` is use to assemble the _settings_ object, as shown below.
-Then, the `addParameters` method is used to either add parameters to each analysis in a _msData_ object or directly to a _msAnalysis_ object.
-The `getParameters` can be used to check which exist in the _msData_ and _msAnalysis_.
-
-```{r create_processing_parameters}
-
-exS2_pp <- exS2
-
-param <- xcms::CentWaveParam(
-  ppm = 15, peakwidth = c(5, 60),
-  snthresh = 10, prefilter = c(6, 10000),
-  mzCenterFun = "mean", integrate = 2,
-  mzdiff = -0.0001, fitgauss = TRUE,
-  noise = 250, verboseColumns = TRUE,
-  firstBaselineCheck = FALSE,
-  extendLengthMSW = TRUE
-)
-
-#creating the settings S4 class for peak picking
-settings_pp <- createSettings(
-  call = "peakPicking",
-  algorithm = "xcms3",
-  settings = param
-)
-
-# the class of settings 
-is(settings_pp)
-
-# add the settings to all analyses in msData
-exS2_pp <- addParameters(exS2_pp, settings = settings_pp, where = "analyses")
-
-# get the parameters in the third analysis of the msData object
-getParameters(exS2_pp, where = "analyses", analyses = 3)
-
-# add the settings to the msAnalysis object
-exA2 <- addParameters(exA2, settings = settings_pp)
-
-# get the parameters of the msAnalysis object
-getParameters(exA2, call = "peakPicking")
-```
-
-### Processing peak picking
-
-```{r peakPicking}
-
-#peak picking call using the stored parameters
-exS2_pp <- peakPicking(exS2_pp)
-
-# peaks added for each analysis, as shown by column peaks of show method (below)
-exS2_pp
-
-#alternatively, parameters can be added during the function call
-exS2_pp_2 <- peakPicking(exS2, algorithm = "xcms3", settings = param)
-
-# parameters are added to each analysis after running peak picking
-summary(getParameters(exS2_pp_2, where = "analyses", analyses = 1))
-
-exS2_pp_2
-
-# running peak picking for an msAnalysis object (settings were already added above)
-exA2 <- peakPicking(exA2)
-
-# msAnalysis with peaks
-exA2
-```
-
-### Using multiple settings
-
-As aforementioned, different settings for the same processing step can exist in a _msData_.
-Below we add different settings for the influent samples, as an example.
-The use case for such functionality can be connected to optimization of processing steps, for example.
-
-```{r peakPicking_multiple_settings}
-
-exS2_pp_3 <- exS2
-
-# Settings for using openms instead of xcms
-settings_pp_2 <- createSettings(
-  call = "peakPicking",
-  algorithm = "openms",
-  settings = list(
-    noiseThrInt = 500,
-    chromSNR = 10,
-    chromFWHM = 10,
-    mzPPM = 15,
-    reEstimateMTSD = TRUE,
-    traceTermCriterion = "sample_rate",
-    traceTermOutliers = 5,
-    minSampleRate = 1,
-    minTraceLength = 3,
-    maxTraceLength = -1,
-    widthFiltering = "fixed",
-    minFWHM = 2,
-    maxFWHM = 40,
-    traceSNRFiltering = FALSE
-  )
-)
-
-exS2_pp_3 <- addParameters(exS2_pp_3, settings = settings_pp, where = "analyses")
-
-# adding different settings to analysis 4 to 6
-exS2_pp_3 <- addParameters(exS2_pp_3, settings = settings_pp_2, where = "analyses", analyses = 4:6)
-
-# the algorithm of the influent samples is different
-sapply(analyses(exS2_pp_3), function(x) getAlgorithm(getParameters(exS2_pp_3, where = "analyses", analyses = x)[[1]]))
-
-#running peak picking
-exS2_pp_3 <- peakPicking(exS2_pp_3)
-
-# as the peak parameters with "openms" as algorithm were less stringent, the influent samples have more peaks then when "xcms3" is used 
-exS2_pp_3
-
-```
-
-### Export and Import settings
-
-The _settings_ object can be exported as a JSON or rds file and then imported for other projects where the same parameter settings are to be applied.
-The methods `exportSettings` and `importSettings` are used, as shown below.
-
-```{r export_import_settings}
-
-#export settings as rds
-exportSettings(settings_pp, name = "settings_pp", format = "rds", path = getwd())
-
-#export settings as JSON
-exportSettings(settings_pp_2, name = "settings_pp_2", format = "json", path = getwd())
-
-#import settings as rds
-settings_pp_imported <- importSettings(file = paste0(getwd(), "/settings_pp.rds"))
-all.equal(settings_pp, settings_pp_imported)
-
-#import settings as JSON
-settings_pp_2_imported <- importSettings(file = paste0(getwd(), "/settings_pp_2.json"))
-all.equal(settings_pp_2, settings_pp_2_imported)
-```
-
-### Inspecting peaks
-
-The access and visualization of peaks uses the calculated/estimated time and _m/z_ dimensions for collecting and plotting the correspondent mass traces.
-Examples are shown below to extract and plot peaks from both _msData_ and _msAnalysis_.
-
-```{r}
-#access to peaks in msData based on built targets as used for extracting EICs
-peaks(exS2_pp[c(1, 4)], mz = mz01[1], rt = rt01[1], ppm = 10, sec = 30)
-
-#access to peak in msAnalysis
-peaks(exA2, mz = targets04)
-
-#plotting peaks in msData
-plotPeaks(exS2_pp, mz = targets04[2, ], interactive = TRUE, colorBy = "analyses")
-
-#plotting peaks from msAnalysis
-plotPeaks(exA2, mz = targets04, interactive = FALSE)
-
-#map plot for peak time and m/z dimensions for msAnalysis
-mapPeaks(exA2, mz = targets04[1, ], xlim = 30, ylim = 0.001)
-
-##map plot for peak time and m/z dimensions for msData
-mapPeaks(exS2_pp, mz = targets04[2, ], colorBy = "replicates", xlim = 30, ylim = 0.001)
-```
-
-## Alingment and Grouping
-
-After peak picking, a common following step is to group the peaks across analyses.
-A retention time alignment can be applied to correct elution deviations across analyses.
-The result of peak grouping and alignment is added to the _msFeatures_ S4 class object already present in the _msData_.
-
-### Grouping settings
-
-The processing parameters for grouping and alignment are added similarly to the processing settings for peak picking but the argument `where` is set to "features", as shown below.
-The settings are added to the slot parameters of the _msFeatures_ object and can be obtained by the method `getParameters`, as shown below.
-
-```{r creating_grouping_parameters}
-
-param_g <- list(
-  rtalign = TRUE,
-  loadRawData = TRUE,
-  groupParam = xcms::PeakDensityParam(
-    sampleGroups = "holder",
-    bw = 3,
-    minFraction = 0.5,
-    minSamples = 1,
-    binSize = 0.008,
-    maxFeatures = 100),
-  preGroupParam = xcms::PeakDensityParam(
-    sampleGroups = "holder",
-    bw = 5,
-    minFraction = 0.5,
-    minSamples = 1,
-    binSize = 0.008,
-    maxFeatures = 100),
-  retAlignParam = xcms::PeakGroupsParam(
-    minFraction = 1,
-    extraPeaks = 0,
-    smooth = "loess",
-    span = 0.3,
-    family = "gaussian")
-)
-
-settings_pg <- createSettings(
-  call = "peakGrouping",
-  algorithm = "xcms3",
-  settings = param_g
-)
-
-# add processing parameter for peakGrouping to features in the msData
-exS2_pp <- addParameters(exS2_pp, settings = settings_pg, where = "features")
-
-# get the processing parameters applied to features
-getParameters(exS2_pp, where = "features")
-```
-
-### Processing grouping and alignment
-
-The data processing for grouping and alignment follows the same principle as peak picking.
-Either the settings are added as arguments or the added parameters in the _msFeatures_ object are used in the function call.
-Below, both cases are demonstrated below.
-
-```{r processing_grouping, message=FALSE}
-#using the added parameters
-exS2_pg <- peakGrouping(exS2_pp)
-
-#adding the parameters as arguments
-exS2_pg_2 <- peakGrouping(exS2_pp, algorithm = "xcms3", settings = param_g)
-```
-
-### Inspecting features
-
-Features can be accessed via the method `features` applied to a _msData_ object.
-The argument `complete` can be set to `TRUE` for a complete list of features information.
-Also, the argument `average` can be set to `FALSE` for returning the intensity of the feature in each analysis.
-
-```{r inspecting_features}
-#getter for feature intensities and intensity deviations
-features(exS2_pg, mz = targets04)
-
-#getter for all the feature details by setting complete to TRUE
-t(features(exS2_pg_2, mz = targets04[1, ], complete = TRUE))
-
-#getter for features with intensities for each sample by setting average to FALSE
-t(features(exS2_pg_2, mz = targets04[1, ], average = FALSE))
-
-#plotting features
-plotFeatures(exS2_pg, mz = targets04, colorBy = "targets", interactive = FALSE)
-
-#plotting features for replicates and interactive
-plotFeatures(exS2_pg, mz = targets04, colorBy = "replicates", interactive = TRUE)
-
-#plot the individual peaks in features and for each analysis shows the time deviation.
-plotFeaturePeaks(exS2_pg, mz = targets04)
-```
-
-```{r}
-
-```
-
-# Data access and visualization
-
-Data access and visualization assets are implemented for different data levels: low level data (i.e., mass traces), peaks and features.
-Additionally, basic and advanced tools are available for accessing, plotting and exporting data (e.g., chromatograms), statistical results (e.g., PCA) and other workflow dependent data/results.
-
-## S4 method
-
-```{r demoCode, echo=FALSE, eval=FALSE, include=FALSE}
-
-# trimSpectraFilesMZR(files = choose.files(), rtr = c(500, 1000), mzr = c(200, 800), onlyMS1 = TRUE)
-# 
-# mz_05 <- data.frame(
-#   retmin = c(809, 907), retmax = c(869, 967),
-#   mzmin = c(247.1626, 239.0604), mzmax = c(247.1676, 239.0652)
-# )
-# 
-# #test <- patRoon::getEICs(msa_1@file, mz_05)
-# 
-# test <- extractEICs(object, mz = targets04)
-# test_xic <- extractXICs(object, mz = targets04)
-
-# projectInfo(msd)
-# path(msd)
-# files(msd)
-# analyses(msd)
-# replicates(msd)
-# blanks(msd)
-# polarities(msd)
-# analysisInfo(msd)
-# 
-# replicates(msd) <- c("s1", "s2")
-# blanks(msd) <- c(NA_character_, "s2")
-# 
-# validObject(msd)
-# 
-# files(msa_1)
-# analyses(msa_1)
-# replicates(msa_1)
-# blanks(msa_1)
-# polarities(msa_1)
-# 
-# replicates(msa_1) <- "s1"
-# blanks(msa_1) <- "s2"
-
-# object <- msd
-# 
-# loadMS1 <- function(object, analysis = NULL, what = "MS1") {
-#   
-#   
-#   test <- RaMS::grabMSdata(files(object)[1], grab_what = "MS1", rtrange = c(min(rt_02), max(rt_02)))
-#   
-#   ms2 <- test$MS2
-#   
-#   ms2[premz > 273 & premz < 274, ]
-#   
-#   View(ms2)
-# 
-# }
-
-```
-
-## Peaks
-
-## Features
-
-## Workflows
 
@@
   fixed_thead = TRUE
-)
+) %>%
+  scroll_box(width = "100%", height = "600px")
 
@@
 
-The report was produced on `r Sys.time()`. The data and report are stored in of the workstation *`r Sys.info()["nodename"]`* under the user *`r Sys.info()["user"]`*.
+The report was produced on `r Sys.time()`. The data and report are stored in ``r path`` of the workstation *`r Sys.info()["nodename"]`* under the user *`r Sys.info()["user"]`*.
 
@@
 h1{
-  font-size: 26pt;
+  font-size: 20pt;
   font-style: bold;
@@
 h2{
-  font-size:20pt;
+  font-size:18pt;
   font-style: bold;
@@
 h3{
-  font-size:16pt;
+  font-size:14pt;
   font-style: bold;

### 2022-11-04 handbooks/Handbook_backend_streamFind.Rmd
LFU21 - tests msAnalysis and cpp extraction improvements

@@
 ```{r newstreamset-basicworkflow, message=FALSE, results='hide'}
-sp2 <- newStreamSet(files = files[10:21], path = getwd(), title = "Project Example")
+sp2 <- newStreamSet(files = files[10:21], title = "Project Example", run_parallel = TRUE)
 ```

### 2022-11-04 man-roxygen/args-newAnalysis-replicate-blank.R

@@ -1,6 +1,0 @@
-#' @param replicate Optional character vector with the name of the
-#' analysis replicate group. The default is \code{NULL} which attempts  to
-#' retrieve the analysis replicate group using the name of the file.
-#' The replicate name can always be edited using the class method.
-#' @param blank Optional character vector with the name of the associate
-#' blank analysis replicate group for the file. The default is \code{NULL}.

### 2022-11-14 README.Rmd
LFU24 - update README and small fixes

--- 2022-05-01 README.Rmd
@@
 
-The goal of streamFind is to ...
+The streamFind package is a backend (code based) and frontend ([shiny](https://shiny.rstudio.com/) app) platform for assembly of modular workflows to support spectrometric and spectroscopic data processing.
+The major focus of streamFind is data processing for environmental and quality studies of the water cycle.
+The streamFind package aims to stimulate the use of advanced data analysis (e.g., non-target screening and statistical analysis) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data.
+The streamFind package can be used by academics but also by technicians due to the comprehensive documentation and well categorized set of integrated functionalities (modules).
 
@@
 
-You can install the development version of streamFind like so:
+For installation of the streamFind package, it is recommended to firstly ensure that the dependencies are currently installed.
+Several existing R packages and external functionalities are used in streamFind for various processing steps.
+A major dependency of streamFind is the [patRoon](https://github.com/rickhelmus/patRoon) R package.
+The patRoon package combines several functionalities for basic and advanced data processing and can be used interchangeably with streamFind.
 
-``` r
-# FILL THIS IN! HOW CAN PEOPLE INSTALL YOUR DEV PACKAGE?
+### Install dependencies
+
+#### R and RTools
+
+[R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) for Windows users.
+The `r R.version.string` can be obtained in https://cran.r-project.org/.
+The RTools can be downloaded in https://cran.r-project.org/bin/windows/Rtools/.
+Make sure to download the right version for the R version installed.
+Installation instructions are given in both sources.
+
+#### patRoon
+
+The patRoon's [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites) offers detailed information on how to install the [patRoon](https://github.com/rickhelmus/patRoon) R package and the required [dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies).
+
+### Install streamFind
+
+The [streamFind](https://github.com/ricardobachertdacunha/streamFind) R package can be installed from the GitHub repository with the following code line:
+
+```{r install-streamfind, eval=FALSE}
+remotes::install_github("ricardobachertdacunha/streamFind", dependencies = TRUE)
 ```
 
-## Example
+## Docker image
 
-This is a basic example which shows you how to solve a common problem:
+Not yet available.
 
-```{r example}
-library(streamFind)
-## basic example code
-```
+## Back-end usage
 
-What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:
-
-```{r cars}
-summary(cars)
-```
-
-You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.
-`devtools::build_readme()` is handy for this.
-You could also use GitHub Actions to re-render `README.Rmd` every time you push.
-An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>.
-
-You can also embed plots, for example:
-
-```{r pressure, echo = FALSE}
-plot(pressure)
-```
-
-In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
+Consult the Handbook_backend_streamFind.Rmd in the handbooks folder for detailed back-end structure and usage documentation.

### 2022-11-29 dev/template_report.Rmd
LFU26 - updated msData peaks methods and tests

@@
 title: "ntsIUTA Report Template"
-author: Ricardo Cunha
+author:
+  name: "Ricardo Cunha"
+  affiliation: "Insitut für Energie- und Umwelttechnik e.V. (IUTA)"
+  email: cunha@iuta.de
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
     code_folding: hide
-    includes:
-      in_header: !expr noquote(paste0(system.file(package = "streamFindData", dir = "extdata"), "\\logo_iuta_template_report.html"))
 editor_options: 
@@
 
+path_extdata <- system.file(package = "streamFind", dir = "extdata")
+logo <- file.path(path_extdata, "logo_iuta.png")
+
 knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 9, results = "asis")
-path <- getwd()
 
+path_wd <- getwd()
 ```
+
+```{r logo, echo=FALSE}
+htmltools::img(
+  src = knitr::image_uri(logo),
+  alt = "logo",
+  style =
+    paste0("
+      margin-left:inherit;
+      position:absolute;
+      top:60px;
+      left:740px;
+      width:", 100, "px;
+      height:", 100*(216/207), "px;"
+    )
+)
+```
+
 <br> <br>
@@
 
-The report was produced on `r Sys.time()`. The data and report are stored in ``r path`` of the workstation *`r Sys.info()["nodename"]`* under the user *`r Sys.info()["user"]`*.
+The report was produced on `r Sys.time()`. The data and report are stored in ``r path_wd`` of the workstation *`r Sys.info()["nodename"]`* under the user *`r Sys.info()["user"]`*.
 
@@
 ***
-Insitut für Energie- und Umwelttechnik e.V. (IUTA)  
+**Insitut für Energie- und Umwelttechnik e.V.**  
+A: Bliersheimer Str. 58 - 60, 47229 Duisburg  
+T: (02065) 418 - 0
+
 <br>
-For further questions, please contact Ricardo Cunha ([cunha@iuta.de](cunha@iuta.de)).
 

### 2022-12-02 README.Rmd
LFU27 - updated basic processing, readme, handbook back end

@@
 
-The streamFind package is a backend (code based) and frontend ([shiny](https://shiny.rstudio.com/) app) platform for assembly of modular workflows to support spectrometric and spectroscopic data processing.
-The major focus of streamFind is data processing for environmental and quality studies of the water cycle.
-The streamFind package aims to stimulate the use of advanced data analysis (e.g., non-target screening and statistical analysis) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data.
-The streamFind package can be used by academics but also by technicians due to the comprehensive documentation and well categorized set of integrated functionalities (modules).
+The streamFind project entitled "Flexible data analysis and workflow designer to identify chemicals in the water cycle" is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de) and is a cooperation between the [Institut für Energie- und Umwelttechnik e.
+V.
+(IUTA)](https://www.iuta.de) and [Forschungszentrum Informatik (FZI)](https://www.fzi.de/).
+The major focus of streamFind is data processing for environmental and quality studies.
+The streamFind package aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data.
+The streamFind package can be used by academics but also by technicians due to the comprehensive documentation and well categorized set of integrated functionalities or modules (see handbooks folder for more information).
 

### 2022-12-02 handbooks/Handbook_backend_streamFind.Rmd
LFU27 - updated basic processing, readme, handbook back end

--- 2022-10-18 handbooks/Handbook_backend_streamFind.Rmd
@@
 
-The streamFind package is a back end (code based) and frontend ([shiny](https://shiny.rstudio.com/) app) platform for assembly of modular workflows to support spectrometric and spectroscopic data processing.
+The streamFind project entitled "Flexible data analysis and workflow designer to identify chemicals in the water cycle" is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de) and is a cooperation between the [Institut für Energie- und Umwelttechnik e.
+V.
+(IUTA)](https://www.iuta.de) and [Forschungszentrum Informatik (FZI)](https://www.fzi.de/).
 The major focus of streamFind is data processing for environmental and quality studies.
 The streamFind package aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data.
-The streamFind package can be used by academics but also by technicians due to the comprehensive documentation and well categorized set of integrated functionalities (modules).
-This handbook focuses on describing and demonstrating the back end of the streamFind package.
+The streamFind package can be used by academics but also by technicians due to the comprehensive documentation and well categorized set of integrated functionalities or modules (see handbooks folder for more information).
+
+This handbook focuses on describing and demonstrating the R package streamFind which aims to be a tool set for processing of MS, Raman, UV and other types of data.
 
@@
 Several existing R packages and external functionalities are used in streamFind for various processing steps.
-A major dependency of streamFind is the [patRoon](https://github.com/rickhelmus/patRoon) R package.`r # TODO add reference` The patRoon package combines several functionalities for basic and advanced data processing and can be used interchangeably with streamFind, see \@ref(patroon-interchangeability) for more information.
-Most of the tools that patRoon depends on are equally used in streamFind.
-In the next sub-chapter, the mandatory and optional dependencies are described with installation instructions.
-For a standalone and self-contained usage option, the streamFind docker image is recommended (see \@ref(docker-image)), avoiding the need to installing all required dependencies.
-
-## Install dependencies
-
-### R and RTools
-
-[R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) are essential dependencies.
-The `r R.version.string` can be obtained in <https://cran.r-project.org/>.
-The RTools can be downloaded in <https://cran.r-project.org/bin/windows/Rtools/>.
+A major dependency of streamFind is the [patRoon](https://github.com/rickhelmus/patRoon) R package.
+The patRoon package combines several functionalities for basic and advanced data processing and can be used interchangeably with streamFind.
+
+### Install dependencies
+
+#### R and RTools
+
+[R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) for Windows users.
+The `r R.version.string` can be obtained in https://cran.r-project.org/.
+The RTools can be downloaded in https://cran.r-project.org/bin/windows/Rtools/.
 Make sure to download the right version for the R version installed.
@@
 
-### Third party software
-
-The external third party dependencies recommended are listed in the section [other dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies) from the patRoon [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites).
-Below we list the mandatory and really recommended software and resources.
-
--   [ProteoWizard](https://proteowizard.sourceforge.io/), required for conversion of vendor MS formats to mzML/mzXML;\
--   [Java JDK](https://www.oracle.com/java/technologies/downloads/), mandatory for patRoon when plotting structure using MetFrag;\
--   [OpenMS](https://www.openms.de/downloads/); esternal tool for peak picking, grouping and alignment;\
--   [MetFrag CL](http://ipb-halle.github.io/MetFrag/projects/metfragcl/), used for formula annotation in patRoon;\
--   [PubChemLite](https://zenodo.org/record/6503754#.YtqxE3bP1hE), database used with MetFrag in patRoon;\
--   [SIRIUS](https://bio.informatik.uni-jena.de/software/sirius/), tool for peak picking and formula annotation in patRoon;\
--   [BioTransformer](https://bitbucket.org/djoumbou/biotransformer/downloads/), used for prediction of transformation products in patRoon;\
--   [SAFD](https://bitbucket.org/SSamanipour/safd.jl/src/master/), tool for peak picking with data in profile mode;\
--   [Open Babel](http://openbabel.org/wiki/Main_Page), used in patRoon for validation of chemicals properties (e.g., InChI and formulae);
-
-### R packages
-
-The following code lines are from the patRoon [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites) to ensure that both patRoon and other mutual R dependencies are installed properly.
-
-```{r r-packages, eval=FALSE}
-install.packages(c("BiocManager", "remotes"))
-
-BiocManager::install(c("mzR", "xcms","CAMERA"))
-
-remotes::install_github("blosloos/enviPick")
-
-BiocManager::install("ropls")
-remotes::install_github("rickhelmus/KPIC2")
-
-BiocManager::install("InterpretMSSpectrum")
-remotes::install_github("cbroeckl/RAMClustR")
-
-remotes::install_github("blosloos/nontargetData")
-remotes::install_github("blosloos/nontarget")
-
-remotes::install_github("rickhelmus/cliqueMS")
-
-# only needed for Bruker DataAnalysis integration
-remotes::install_github("BSchamberger/RDCOMClient")
-
-BiocManager::install(c("BiocStyle", "Rgraphviz")) 
-remotes::install_github("KelseyChetnik/MetaClean")
-
-remotes::install_github("rickhelmus/patRoon", upgrade = "never", dependencies = TRUE)
-```
-
-After installation of the patRoon R package, the location of each third party software should be given in an `.Rprofile` file as suggested in section [other dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies) of the patRoon [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites).
-For facilitation the same code lines are given below.
-The function `patRoon::verifyDependencies()` from patRoon can be used to verify if all dependencies are found.
-
-```{r options-third-party-software, eval=FALSE}
-options(patRoon.path.pwiz = "C:/ProteoWizard") # location of ProteoWizard installation folder
-options(patRoon.path.SIRIUS = "C:/sirius-win64-3.5.1") # location where SIRIUS was extracted
-options(patRoon.path.OpenMS = "/usr/local/bin") # directory with the OpenMS binaries
-options(patRoon.path.pngquant = "~/pngquant") # directory containing pngquant binary
-options(patRoon.path.MetFragCL = "~/MetFragCommandLine-2.4.8.jar") # full location to the jar file
-options(patRoon.path.MetFragCompTox = "C:/CompTox_17March2019_SelectMetaData.csv") # full location to desired CompTox CSV file
-options(patRoon.path.MetFragPubChemLite = "~/PubChemLite_exposomics_20220429.csv") # full location to desired PubChemLite CSV file
-options(patRoon.path.BioTransformer = "~/biotransformer/biotransformer-3.0.1.jar")
-options(patRoon.path.obabel = "C:/Program Files/OpenBabel-3.0.0") # directory with Open Babel binaries
-```
-
-## Install streamFind
-
-The streamFind package can be installed from the GitHub repository <https://github.com/ricardobachertdacunha/streamFind> with the following code line.
+#### patRoon
+
+The patRoon's [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites) offers detailed information on how to install the [patRoon](https://github.com/rickhelmus/patRoon) R package and the required [dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies).
+
+### Install streamFind
+
+The [streamFind](https://github.com/ricardobachertdacunha/streamFind) R package can be installed from the GitHub repository with the following code line:
 
@@
 
-To be done.
+Not yet available.
 
@@
 
-The streamFind package includes example data files via the streamFindData R package.
+The streamFind package includes example data files via the [streamFindData R package](https://github.com/ricardobachertdacunha/streamFindData).
 Once installed the full file paths of MS data files can be obtained as shown below.
@@
 ```{r resource-files, cache=FALSE}
-# 
-# r_path <- system.file(package = "streamFindData", dir = "extdata")
-# #r_path <- system.file(package = "streamFind", dir = "extdata")
-# 
-# files <- list.files(r_path, pattern = ".mzML|.mzXML", full.names = TRUE)
-
-#list of file full paths from the data package
 files <- streamFindData::msFilePaths()
@@
 An overview of the S4 classes and their hierarchy is presented in Figure \@ref(fig:classes-overview) using MS analyses as an example.
-Other data types, such as from UV and Raman analyses, would lead to different classes (e.g., *uvData* and *uvAnalysis* and *ramanData* and *ramanAnalysis*, respectively) but the hierarchy is similar as shown for MS data.
+Other data types, such as from UV and Raman analyses, would lead to different classes (e.g., *uvData* and *uvAnalysis* and *ramanData* and *ramanAnalysis*, respectively) but the hierarchy is similar as shown for MS data (not yet implemented).
 
@@
 
-## Nomenclature
-
-The nomenclature listed below is consistently applied throughout the objects and arguments of functionalities in the streamFind package.
-Other less relevant terms are used but not listed.
-For further information please consult the specific documentation of each object or functionality when required using the `?x` method, where x is the name of the S4 class, S4 method or function.
-
--   *streamSet* refers to the broader S4 class applied in the streamFind package and holds basic information for an analysis set, see \@ref(streamset-class);
--   *msData* refers to the S4 subclass of *streamSet* when MS analyses (i.e., MS files) are added to the set, see \@ref(msdata-class);
--   *msAnalysis* refers to the S4 class of the object generated for each MS file added to the *msData* object in the slot "analyses";
--   `setInfo` is the method to obtain the basic information (i.e., title, date and path) of a *streamSet* object;
--   `analyses` is the name of the slot holding the analyses (as list) and is also a method to get the analysis names in the *streamSet*;
--   `replicate` is the name of an analysis replicate group and the method to get the replicate names;
--   `blank` is the name of a blank analysis replicate group to be used for blank (blind) subtraction and the method to get blank replicate names;
--   **traces** corresponds to a raw data signal (e.g., mass trace in MS data) obtained from a given analysis file;
--   `spectra` for MS analyses is the method to get a table with a spectrum for each cycle time (i.e., retention time);
--   **EIC** is an extracted ion chromatogram (retention time vs intensity) and can be obtained with the method `EICs`;
--   **TIC** is a total ion chromatograms (retention time vs intensity) and can be obtained with the method `TICs`;
--   **XIC** is a three dimensional ion chromatogram (*m/z* vs retention time vs intensity) and can be obtained with the method `XICs`;
--   **peak** corresponds to a cluster of traces (e.g., integrated chromatographic MS peak) within an analysis file and can be accessed with the method `peaks`;
--   **feature** corresponds to a group of corresponding peaks across analysis files/analyses and can be accessed with the method `features`;
--   **MS2** corresponds to second order fragmentation data as acquired in tandem MS or MS/MS mode and can be accessed with the method `MS2s`.
-
-(...)
-
 ## streamSet class
@@
 The set creation is demonstrated below for the three MS files (triplicate, 1 to 3 in \@ref(tab:description-files)) where a set of chemicals were spiked (see \@ref(spiked-chemicals) with tag IS and S).
-The set path can be assigned as argument but for the example below it was left as default (i.e., the working directory).
 The initial S4 class created in the `newStreamSet` function is *streamSet*.
@@
 
-```{r newstreamset-call, message=FALSE, results='hide'}
+```{r newstreamset-call, message=FALSE, results='hide', message=FALSE, warning=FALSE}
 #New streamSet with the files 1 to 3
@@
 
-Because the added files were all with mzML format (i.e., MS files), the subclass is directly defined as *msData*.
+Because the added files were all with mzML format (i.e., MS files), the subclass of `sp1` is directly defined as *msData*.
 The structure for further processing data is then set for MS data.
@@
 The *msData* structure includes the *streamSet* class plus the `features` slot.
-The title, date and path slots correspond to the set title, creation date and directory in disk, respectively.
+The title and date slots correspond to the set title and creation date, respectively.
 The analyses slot is a list with a *msAnalysis* S4 class object per MS file added.
-The structure of the *msAnalysis* class is described in the next sub-chapter.
-The `features` slot is a *msFeatures* S4 class object to hold the results from alignment, grouping and annotation of peaks across analyses.
-Further information for *msAnalysis* is given in the next sub-chapter (\@ref(msanalysis-class)) and for *msFeatures* is given in the sub-chapter \@ref(msfeatures-class), after creation of features by performing alignment and grouping of peaks across analyses.
+The `features` slot is a *msFeatures* S4 class object to hold the results from alignment, grouping (or correspondence) and annotation of peaks across analyses.
+Further information for *msAnalysis* is given in the next sub-chapter (\@ref(msanalysis-class)) and for *msFeatures* is given in the sub-chapter \@ref(msfeatures-class), after creation of features by performing correspondence of peaks across analyses.
 
@@
 The metadata list can be expanded with additional information about the analysis/sample (e.g., sample location, weather conditions, process sensor data, etc.).
-The slot parameters is used to store processing steps (e.g., peak picking) as *settings* S4 class objects, including the processing step name/ID, the algorithm used and the respective settings applied.
-The parameters can be used for history track but also for re-run the processing steps when required.
+The slot settings is used to store processing steps (e.g., peak picking) as *settings* S4 class objects, including the processing step name/ID, the algorithm used and the respective processing parameters applied.
+The settings can be used for history track but also for re-run the processing steps when required.
 The spectra and chromatograms slots are used to store low level data (i.e., mass traces either in profile or centroided modes) from the MS files.
 Finally, the peaks slot is used to store the chromatographic peaks after peak picking.
-As no other processing step was applied the spectra, chromatograms and peaks tables in the *msAnalysis* are empty for now.
+As no other processing step was applied the spectra, chromatograms and peaks in the *msAnalysis* are empty `data.tables` for now.
 
@@
 The function `makeTargets` is used for comprehensively build targets.
-As shown below for Carbamazepin d10 and Diuron d6, multiple ways can be used to assemble targets but similarly returning a uniform `data.table` class object with a target list.
+As shown below for Carbamazepin-d10 and Diuron-d6, multiple ways can be used to assemble targets but similarly returning a uniform `data.table` class object with a target list.
 Accessing and plotting S4 methods in streamFind use the `makeTargets` function to collect raw data from the analysis files.
@@
 
-## TICs
-
-Total ion chromatograms (TICs) are obtained either by pre-loading and adding the chromatogram data to the *msAnalysis*, extracting the data in a `TICs` call (S4 method) or loading all spectra and applying the sum of intensities to each spectrum.
+## TICs and BPCs
+
+Total ion chromatograms (TICs) and base peak chromatograms (BPCs) are obtained either by pre-loading and adding the chromatogram data to the *msAnalysis*, extracting the data in a `TICs` call (S4 method) or loading all spectra and applying the sum of intensities to each spectrum.
 Note, TIC is already available as chromatogram in mzML data from the major HRMS vendors.
@@
 tic1 <- TICs(sp1, analyses = 1:3)
-tic1
+head(tic1, 6)
 ```
@@
 #TIC of the msAnalysis object
-head(TICs(a1), 6)
-```
-
-Similarly to EICs, the TICs can be plotted with `plotTICs` S4 method, as shown below.
+head(TIC(a1), 6)
+```
+
+```{r bpc_method}
+#BPC of the msAnalysis object
+head(BPC(a1), 6)
+```
+
+Similarly to EICs, the TICs and BPCs can be plotted with `plotTICs` and `plotBPCs` S4 methods, respectively, as shown below.
 
@@
 plotTICs(sp1, analyses = 1:3, colorBy = "replicates")
+```
+
+```{r plotbpcss-method-2}
+#using the msData object
+plotBPCs(sp1, colorBy = "replicates")
 ```
@@
 ```{r loadrawdata-msanalysis}
-#load raw data to an msAnalysis object, while setting a minimum intensity value for both MS1 and MS2 traces
-a1 <- loadRawData(a1, minIntensityMS1 = 0, minIntensityMS2 = 100)
+#load raw data to an msAnalysis object
+a1 <- loadRawData(a1)
 ```
@@
 #load raw data to the first three samples in an msData object.
-sp1 <- loadRawData(sp1, analyses = 1:3)
+sp1 <- loadRawData(sp1)
 ```
@@
 ## Spectra
+
+Spectra can be accessed from an *msAnalaysis* or *msData* as shown below with methods `spectra`. The method `plotSpectra` is also available for both classes.
 
@@
 
+```{r plot-spectra-msanalysis}
+# plotting spectra of isotopic pattern for Diuron-d6 in an msAnalysis
+plotSpectra(a1, mz = data.frame(mzmin = 238, mzmax = 250,
+                                rtmin = 1145, rtmax = 1170))
+```
+
 ## Chromatograms
 
-`r # TODO create method to access chromatograms in msAnalysis and msData`
+Chromatograms can be included in mzML files.
+For instance, TIC is often available as chromatogram in mzML files.
+Also, analyses produced in multiple reaction monitoring (MRM) mode, when exported to mzML without collapsing the MRMs, will have the data stored as chromatograms, as shown below.
+The method `chromatograms`
+
+```{r create-mrm-msAnalysis}
+# MS files acquired in MRM
+mrm <- newAnalysis(file = files[29])
+
+# data only as chromatograms
+mrm
+```
+
+```{r laod-show-chromatograms}
+# loading raw data into the msAnalysis
+mrm <- loadRawData(mrm)
+
+# getting the chromatograms as data.table
+chroms <- chromatograms(mrm)
+
+head(chroms)
+```
+
+```{r plot-loaded-chromatograms}
+# plotting the second and third chromatograms in the msAnalysis
+plotChromatograms(mrm, index = c(2,3), interactive = TRUE)
+```
 
@@
 
-Regarding MS data, the format used in streamFind is .mzML or .mzXML either centroided or profile.
+Regarding MS data, the format used in streamFind is mzML or mzXML either centroid or profile mode.
 However, both mentioned formats do not correspond to the formats of actual MS system vendors.
-[msConvert](https://proteowizard.sourceforge.io/download.html) from [ProteoWizard](https://proteowizard.sourceforge.io/) can be used to convert the main MS formats (and others) to .mzML or .mzXML, including data centroiding when required.
+[msConvert](https://proteowizard.sourceforge.io/download.html) from [ProteoWizard](https://proteowizard.sourceforge.io/) can be used to convert the main MS formats (and others) to mzML (preferable as it is the most recent) or mzXML, including data centroiding when required.
 The msConvert GUI tool can be used for conversion.
 However, if automation is required, the [command line](https://proteowizard.sourceforge.io/tools/msconvert.html) of msConvert is used within the function `convertFiles` in streamFind, as shown below.
-The function `compatibleFileFormatsForConversion()` can be used to check which vendor formats are currently possible to convert in streamFind through ProteoWizard.
+The function `compatibleFileFormatsForConversion()` can be used to check which vendor formats are currently possible to convert in the streamFind through ProteoWizard.
 
@@
 
-The analysis set creation is demonstrated below for twelve HRMS data files corresponding to blank and wastewater influent samples, both measured in triplicate and in positive and negative modes.
+The analysis set creation is demonstrated below for twelve HRMS data files corresponding to blank and wastewater (influent) samples, both measured in triplicate and in positive and negative modes.
 The argument `files` is the main input and can either be a vector of full path files, a table with file (i.e., full file path), replicate (i.e., the name of the analysis replicate group for each file) and blank (i.e., the name of the associated blank analysis replicate group for each file), or the `analysisInfo` data.frame from the package [patRoon](https://github.com/rickhelmus/patRoon).
-The set path can be assign as argument `path` and is the directory where all the set files (i.e., scripts, cache databases, objects and results) will be stored by default.
-
-```{r newstreamset-basicworkflow, message=FALSE, results='hide'}
-sp2 <- newStreamSet(files = files[10:21], path = getwd(), title = "Project Example")
+Setting `run_parallel` to `TRUE` will enable parallel processing, which can be useful for large number of files.
+See more details about parallel processing in the section \@ref(parallel-processing).
+
+```{r newstreamset-basicworkflow, message=FALSE, results='hide', warning=FALSE}
+sp2 <- newStreamSet(files = files[10:21],
+                    title = "Project Example",
+                    run_parallel = TRUE)
 ```
@@
 
-Before preforming data processing, the analysis names (i.e., replicate name and associated blank replicate name) and metadata can be amended/extended.
+Before preforming data processing, the analysis replicate names and associated blank replicate names as well as metadata can be amended/extended.
 
@@
 
+`r # TODO add concatenation for msData objects`
+
 ```{r assign-blankreplicatenames}
@@
 
-The initial processing step of the basic workflow is the peak picking.
-Yet, other steps might be necessary prior the pick picking, such as data conversion, data calibration, etc.
-These are discussed later in section X.
-The basic workflow in streamFind is based on [patRoon](https://github.com/rickhelmus/patRoon) and most of the output in streamFind can be used/converted to enable the use of native [patRoon](https://github.com/rickhelmus/patRoon) functions and methods.
+An initial processing step of the basic workflow is the peak picking.
+Other processing steps might be necessary prior the pick picking, such as data conversion, data centroiding, data calibration, etc.
+but are not yet implemented in streamFind.
+The peak picking and other basic processing steps in streamFind are applied through [patRoon](https://github.com/rickhelmus/patRoon).
+Both *msData* and *msAnalyses* can be converted to native [patRoon](https://github.com/rickhelmus/patRoon) objects for native use of [patRoon](https://github.com/rickhelmus/patRoon).
+See section \@ref(patRoon-interchangeability) for more information.
+
+`r # TODO create as.msData and as.msAnalysis methods for patRoon's features/featureGroups`
 
@@
 
-```{r peakpicking-msdata-run, results='hide', message=FALSE}
+```{r peakpicking-msdata-run, results='hide', message=FALSE, warning=FALSE}
 #peak picking call using the stored parameters
@@
 The processing parameters for grouping and alignment are added similarly to the processing settings for peak picking but the argument `where` is set to "features", as shown below.
-The settings are added to the slot parameters of the *msFeatures* object and can be obtained by the method `getParameters`, as shown below.
+The settings are added to the slot settings of the *msFeatures* object and can be obtained by the method `getSettings`, as shown below.
+Note that, for grouping positive and negative analyses, the retention time alignment would not work.
+Alignment can be applied only with either positive or negative analysis as shown below in \@ref(alignment-results).
 
@@
 The slot metadata holds additional information about each feature, such as average *m/z*/ neutral mass, retention time and number of corresponding peaks.
-The slot settings has the list of settings used to build/process features, such as peakGrouping and filterFeatures.
+The slot settings has the list of settings used to build/process features, such as `peakGrouping` and `filterFeatures`.
 
@@
 
-```{r inspecting-features}
+```{r inspecting-features, message=FALSE, warning=FALSE}
 #getter for feature intensities correspondent to the mass of Diuron in each sample by setting average to false
@@
 
-```{r grouping-msdata-only-positive}
-
+```{r grouping-msdata-only-positive, message=FALSE, warning=FALSE}
+
+# subset of msData with only positive analyses
 sp2_pg_3 <- sp2_pp[c(4:6, 10:12)]
 
+# grouping settings with alignment scheduled 
 settings_pg_xcms_2 <- list(
@@
 sp2_pg_3 <- peakGrouping(sp2_pg_3, settings = settings_pg_xcms_2)
-
-peaks(sp2_pg_3)
-
-features(sp2_pg_3, mass = targets4, ppm = 10, sec = 10, complete = TRUE)
-
-plotAlignment(sp2_pg_3)
 ```
@@
 
-As peak picking can result in false negatives and to ensure that false positives are less likely, a recursive integration is often performed for filling missing peaks in a feature.
-The recursive integration consists of extracting the mass traces in the feature region from analyses not represented, returning the hipotetical peak intensity.
+As peak picking can result in false negatives and to ensure that false positives are less likely, a recursive integration is often performed for filling missing peaks in a given feature.
+The recursive integration consists of extracting the mass traces in the feature region from analyses not represented, returning the hypothetical peak intensity.
 The exact collection of mass traces and calculation of the peak intensity (or height) is dependent on the algorithm applied.
@@
 A *settings* object can be obtained with `fillingSettingsDefaultXCMS` for a default *ChromPeakAreaParam* from xcms.
-Note that the peakFilling settings are added to the parameters slot of the *msData* object.
+Note that the `peakFilling` settings are added to the settings slot of the *msFeatures* object.
 
@@
 Annotation of isotopes and adducts is crucial not only to prioritize the relevant features but also to deliver essential information for further identification, such as charge, isotopic pattern and adduct yield.
+Annotation in streamFind is applied through patRoon which uses the [CAMERA](https://www.bioconductor.org/packages/release/bioc/html/CAMERA.html) R package.
+Again, *msData* objects with both positive and negative analyses would not work for annotation.
+The results of annotation are stored as extra columns in the metadata slot of the *msFeatures* S4 class.
 
@@
 
-The results of annotation are stored under annotation of the *msFeatures* S4 class and matches the [components](https://rickhelmus.github.io/patRoon/reference/components-class.html) S4 class of the [patRoon](https://github.com/rickhelmus/patRoon) package.
-This means that the feature components can be accessed via the `getAnnotation` method for an *msData* object and inspected also with native methods from patRoon.
-
 ### Inspection of annotation
@@
 Data filtration is crucial for proper prioritization of relevant peaks/features.
-In streamFind, two levels of data filtration is implemented: (1) data sub-setting (permanent) and (2) filter processing steps for peaks and/or features (tagged but not removed).
+In streamFind, two levels of data filtration is implemented: (1) data sub-setting (permanent) and (2) filter processing steps for peaks and/or features (both are tagged as filtered but not removed).
 
@@
 
+`r # TODO add filterPeaks option to msData and msAnalysis`
+
 ```{r filter-peaks}
@@
 
+`r # TODO implement MetaClean qualita evaluation``
+
 ```{r quality-sn}
-
-# sp2_pg <- calculateSNR(sp2_pg, targetsID = features(sp2_pg)[["id"]][1:5])
-# 
-# features(sp2_pg, complete = TRUE)[1:5, ]
-
+sp2_pg <- calculateSNR(sp2_pg, targetsID = features(sp2_pg)[["id"]][1:2])
+
+features(sp2_pg, complete = TRUE)[1:2, ]
 ```
@@
 
-## Processing settings
-
-Settings are needed to extract and average the fragmentation data.
-Below an example of settings used for accessing and averaging MS2 data is given.
-The resulting *settings* object is then used further in the call to `MS2s` or `plotMS2s`.
-
-```{r create-settings-msn}
-#the parameter settings to obtained and process MSn spectra
-param_msn <- list(
-  isolationTimeWindow = 0,
-  isolationMassWindow = 1.3,
-  clusteringMethod = "distance",
-  clusteringUnit = "ppm",
-  clusteringWindow = 10,
-  minIntensityPre = 200,
-  minIntensityPost = 300,
-  asPatRoon = FALSE, #when TRUE, is only used by running generateMS2
-  mergeVoltages = TRUE,
-  mergeBy = "analyses"
-)
-
-#creating the settings S4 class
-settings_msn <- createSettings(
-  call = "extractMSn",
-  algorithm = "streamFind",
-  parameters = param_msn
-)
-
-# the class of settings 
-is(settings_msn)
-```
-
 ## Access and plot MS2
@@
 Similarly, `plotMS2s` can be used to access and directly plot MS2 data from targets, as shown below.
-A more general function for accessing the data is `extractMSn`, where further fragmentation levels can be accessed.
-When using streamFind basic workflow, the use of `MS2s` and `plotMS2s` is recommended.
-All the S4 methods/function can be applied for *msAnalysis* and *msData* objects.
 
 ```{r extract-ms2}
-# #head of MS2 data for target 1 in a msAnalysis object
-# head(MS2s(a1, mz = targets4[1, ], settings = settings_msn))
-# 
-# #ploting the MS2 data of the first target in a msAnalysis object
-# plotMS2s(a1, mz = targets4[1, ], settings = settings_msn, interactive = FALSE)
-# 
-# #most intense MS2 data traces for both targets in a msData object
-# s_data_3 <- MS2s(sp1, mz = targets4, settings = settings_msn)
-# s_data_3 <- s_data_3[order(intensity, decreasing = TRUE), ]
-# head(s_data_3)
-# 
-# #ploting the MS2 data of the first target in a msAnalysis object
-# plotMS2s(sp1, mz = targets4, settings = settings_msn, interactive = TRUE, colorBy = "targets")
+#head of MS2 data for target 1 in a msAnalysis object
+head(MS2s(a1, mzClust = 0.01, isolationWindow = 1, mz = targets4[1, ]))
+
+#plotting the MS2 data of the first target in a msAnalysis object
+plotMS2s(a1, mzClust = 0.01, isolationWindow = 1, mz = targets4[1, ])
+
+#most intense MS2 data traces for both targets in a msData object
+ms2_data <- MS2s(sp1, mzClust = 0.01, isolationWindow = 1, mz = targets4)
+ms2_data <- ms2_data[order(intensity, decreasing = TRUE), ]
+head(ms2_data)
+
+#plotting the MS2 data of the first target in a msAnalysis object
+plotMS2s(sp1, mzClust = 0.01, isolationWindow = 1, mz = targets4,
+         interactive = TRUE, colorBy = "targets")
 ```
@@
 
-## patRoon interchangeability
+# patRoon interchangeability
+
+`# TODO add as.msData and as.msAnalysis functions`
 
 # Workflows
+
+TO be implemented
+
+# Nomenclature
+
+The nomenclature listed below is consistently applied throughout the objects and arguments of functionalities in the streamFind package.
+Other less relevant terms are used but not listed.
+For further information please consult the specific documentation of each object or functionality when required using the `?x` method, where x is the name of the S4 class, S4 method or function.
+
+-   *streamSet* refers to the broader S4 class applied in the streamFind package and holds basic information for an analysis set, see \@ref(streamset-class);
+-   *msData* refers to the S4 subclass of *streamSet* when MS analyses (i.e., MS files) are added to the set, see \@ref(msdata-class);
+-   *msAnalysis* refers to the S4 class of the object generated for each MS file added to the *msData* object in the slot "analyses";
+-   `setInfo` is the method to obtain the basic information (i.e., title, date and path) of a *streamSet* object;
+-   `analyses` is the name of the slot holding the analyses (as list) and is also a method to get the analysis names in the *streamSet*;
+-   `replicate` is the name of an analysis replicate group and the method to get the replicate names;
+-   `blank` is the name of a blank analysis replicate group to be used for blank (blind) subtraction and the method to get blank replicate names;
+-   **traces** corresponds to a raw data signal (e.g., mass trace in MS data) obtained from a given analysis file;
+-   `spectra` for MS analyses is the method to get a table with a spectrum for each cycle time (i.e., retention time);
+-   **EIC** is an extracted ion chromatogram (retention time vs intensity) and can be obtained with the method `EICs`;
+-   **TIC** is a total ion chromatograms (retention time vs intensity) and can be obtained with the method `TICs`;
+-   **XIC** is a three dimensional ion chromatogram (*m/z* vs retention time vs intensity) and can be obtained with the method `XICs`;
+-   **peak** corresponds to a cluster of traces (e.g., integrated chromatographic MS peak) within an analysis file and can be accessed with the method `peaks`;
+-   **feature** corresponds to a group of corresponding peaks across analysis files/analyses and can be accessed with the method `features`;
+-   **MS2** corresponds to second order fragmentation data as acquired in tandem MS or MS/MS mode and can be accessed with the method `MS2s`.
+
+(...)
+
+# Parallel processing
+
+To be added
 

### 2023-02-13 README.Rmd
Added logos to README.
Updated group logo in README
Continuation of R6 implementation 13

--- 2022-11-14 README.Rmd
@@
 
-The streamFind package is a backend (code based) and frontend ([shiny](https://shiny.rstudio.com/) app) platform for assembly of modular workflows to support spectrometric and spectroscopic data processing.
-The major focus of streamFind is data processing for environmental and quality studies of the water cycle.
-The streamFind package aims to stimulate the use of advanced data analysis (e.g., non-target screening and statistical analysis) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data.
-The streamFind package can be used by academics but also by technicians due to the comprehensive documentation and well categorized set of integrated functionalities (modules).
+The R package `streamFind` (this repository) is the back-end platform for the streamFind project entitled "Flexible data analysis and workflow designer to identify chemicals in the water cycle".
+The streamFind project is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de) and is a cooperation between the [Institut für Energie- und Umwelttechnik e.
+V.
+(IUTA)](https://www.iuta.de) and [Forschungszentrum Informatik (FZI)](https://www.fzi.de/).
+The major focus of `streamFind` is data processing for environmental and quality studies.
+The `streamFind` R aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data.
+The streamFind package can be used by academics but also by technicians due to the comprehensive documentation and well categorized set of integrated functions or modules (see vignettes for more information).
+
+```{r, echo=FALSE, out.width = '100%', fig.align='center'}
+knitr::include_graphics("inst/extdata/logo_group.png")
+```
 
@@
 
-For installation of the streamFind package, it is recommended to firstly ensure that the dependencies are currently installed.
-Several existing R packages and external functionalities are used in streamFind for various processing steps.
-A major dependency of streamFind is the [patRoon](https://github.com/rickhelmus/patRoon) R package.
-The patRoon package combines several functionalities for basic and advanced data processing and can be used interchangeably with streamFind.
+For installation of the `streamFind` R package, it is recommended to firstly ensure that the dependencies are currently installed.
+Several existing R packages and external software are used in `streamFind` for various processing steps.
+A major dependency of `streamFind` is the [patRoon](https://github.com/rickhelmus/patRoon) R package.
+The `patRoon` R package combines several tools for basic and advanced data processing and can be used interchangeably with `streamFind`.
 
@@
 
-## Back-end usage
+## Backend usage
 

### 2023-02-13 handbooks/Handbook_backend_streamFind.Rmd
Continuation of R6 implementation 13

@@
 ---
-title: "Handbook streamFind: back end"
+title: "Handbook streamFind: backend"
 author: "Ricardo Cunha"

### 2023-02-13 R/utils_ms_file_functions.R

@@ -1,0 +1,382 @@
+
+#' check_compatible_ms_formats_for_conversion
+#'
+#' @description List of possible vendor formats to be converted to
+#' `mzML` or `mzXML`.
+#'
+#' @references
+#'
+#' \insertRef{proteo01}{streamFind}
+#'
+#' \insertRef{proteo02}{streamFind}
+#'
+#' @export
+#'
+#' @return A `data.frame` with the formats compatible for conversion.
+#'
+check_compatible_ms_formats_for_conversion = function() {
+
+  df_formats = data.frame(
+    type = c(rep("ms", 11)),
+    vendor = c("agilent", rep("bruker", 5), "sciex", "shimadzu",
+      "thermo scientific", rep("waters",2)),
+    format = c(".d", ".d", ".yep",  ".baf", ".fid", ".tdf",
+      ".wiff", ".lcd", ".RAW", ".RAW", "UNIFI")
+  )
+
+  return(df_formats)
+}
+
+#' convert_ms_files
+#'
+#' @description Function to convert vendor MS file formats to mzML/mzXML using
+#' the command line of `MSConvert` from \emph{ProteoWizard}.
+#'
+#' @param files A vector with file full path/s for conversion.
+#' @param outputFormat A character string with the format of the output file.
+#' Possible values are "mzML" and "mzXML".
+#' @param outputPath The directory to place the output file. When \code{NULL}
+#' the directory of the original file is used.
+#' @param optList A list of further options to be passed to the respective
+#' conversion tool. See
+#' \url{https://proteowizard.sourceforge.io/tools/msconvert.html}
+#' for commands available for conversion of MS files with
+#' \href{https://proteowizard.sourceforge.io/download.html}{msConvert}.
+#'
+#' @note When giving paths make sure that single backslash is changed to
+#' double backslash of to single forward slash.
+#'
+#' @references
+#'
+#' \insertRef{proteo01}{streamFind}
+#'
+#' \insertRef{proteo02}{streamFind}
+#'
+#' @export
+#'
+#' @examples
+#'
+#' \dontrun{
+#' # The full file path
+#' file = E:/ms_file.d
+#'
+#' # Options to centroid MS1 and MS2 data in file
+#' optList <- list(filter = "peakPicking vendor msLevel=1-2")
+#'
+#' # Convert to mzML
[…]

### 2023-03-03 R/class_S3_msAnalysis.R
Continuation of R6 implementation 28

@@ -1,0 +1,251 @@
+
+#' validate_ms_analysis
+#'
+#' @description
+#' Validates an msAnalysis S3 class object.
+#'
+#' @param x An msAnalysis S3 class object.
+#'
+#' @return Logical of length one.
+#'
+#' @export
+#'
+validate_ms_analysis <- function(x = NULL) {
+  valid <- FALSE
+  name <- NA_character_
+
+  if (is.list(x)) {
+    valid <- TRUE
+
+    if (length(x$name) != 1 & !is.character(x$name)) {
+      warning("Analysis name not conform!")
+      valid <- FALSE
+    } else {
+      name <- x$name
+    }
+
+    if (length(x$replicate) != 1 & !is.character(x$replicate)) {
+      warning("Analysis replicate name not conform!")
+      valid <- FALSE
+    }
+
+    if (length(x$blank) != 1 & !is.character(x$blank)) {
+      warning("Analysis blank name not conform!")
+      valid <- FALSE
+    }
+
+    if (length(x$file) != 1 & !is.character(x$file)) {
+      warning("Analysis file path entry not conform!")
+      valid <- FALSE
+    } else if (!file.exists(x$file)) {
+      warning(paste0(
+        x$file,
+        " does not exist! Update file paths with msData$update_files() method"
+      ))
+      valid <- FALSE
+    }
+
+    if (length(x$type) != 1) {
+      warning("Analysis type entry not conform!")
+      valid <- FALSE
+    } else if (!(x$type %in% c("MS", "MS/MS", "SRM"))) {
+      warning("Analysis type must be 'MS', 'MS/MS' or 'SRM'!")
+      valid <- FALSE
+    }
+
+    if (!is.integer(x$spectra_number) &&
+        length(x$spectra_number) != 1) {
+      warning("Analysis spectra_numebr entry not conform!")
+      valid <- FALSE
+    }
+
+    if (!is.integer(x$chromatograms_number) &&
+        length(x$chromatograms_number) != 1) {
+      warning("Analysis chromatograms_number entry not conform!")
+      valid <- FALSE
+    }
+
+    if (!is.character(x$spectra_mode) &
+        length(x$spectra_mode) != 1) {
+      warning("Analysis spectra_mode entry not conform!")
+      valid <- FALSE
+    }
+
+    if (!is.integer(x$spectra_levels)) {
+      warning("Analysis spectra_levels entry not conform!")
[…]

### 2023-03-06 README.Rmd
Continuation of R6 implementation 28
Updated README and logos moved to man/figures/

@@
 
-The R package `streamFind` (this repository) is the back-end platform for the streamFind project entitled "Flexible data analysis and workflow designer to identify chemicals in the water cycle".
+The R package [streamFind](https://github.com/ricardobachertdacunha/streamFind) (this repository) is the back-end platform for the streamFind project entitled "Flexible data analysis and workflow designer to identify chemicals in the water cycle".
 The streamFind project is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de) and is a cooperation between the [Institut für Energie- und Umwelttechnik e.
@@
 (IUTA)](https://www.iuta.de) and [Forschungszentrum Informatik (FZI)](https://www.fzi.de/).
-The major focus of `streamFind` is data processing for environmental and quality studies.
-The `streamFind` R aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data.
+The major focus of streamFind is data processing for environmental and quality studies.
+The streamFind aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data.
 The streamFind package can be used by academics but also by technicians due to the comprehensive documentation and well categorized set of integrated functions or modules (see vignettes for more information).
@@
 ```{r, echo=FALSE, out.width = '100%', fig.align='center'}
-knitr::include_graphics("inst/extdata/logo_group.png")
+knitr::include_graphics("man/figures/logo_group.png")
 ```
@@
 
-For installation of the `streamFind` R package, it is recommended to firstly ensure that the dependencies are currently installed.
-Several existing R packages and external software are used in `streamFind` for various processing steps.
-A major dependency of `streamFind` is the [patRoon](https://github.com/rickhelmus/patRoon) R package.
-The `patRoon` R package combines several tools for basic and advanced data processing and can be used interchangeably with `streamFind`.
+For installation of the [streamFind](https://github.com/ricardobachertdacunha/streamFind) R package, it is recommended to firstly ensure that the dependencies are currently installed.
+Several existing R packages and external software are used in streamFind for various processing steps.
+A major dependency of streamFind is the [patRoon](https://github.com/rickhelmus/patRoon) R package.
+The [patRoon](https://github.com/rickhelmus/patRoon) R package combines several tools for basic and advanced data processing and can be used interchangeably with streamFind.
 
 ### Install dependencies
-
-#### R and RTools
 
@@
 
-#### patRoon
-
 The patRoon's [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites) offers detailed information on how to install the [patRoon](https://github.com/rickhelmus/patRoon) R package and the required [dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies).
@@
 
-The [streamFind](https://github.com/ricardobachertdacunha/streamFind) R package can be installed from the GitHub repository with the following code line:
+The [streamFind](https://github.com/ricardobachertdacunha/streamFind) R package can be installed from the GitHub repository with the following code line.
 
@@
 
-## Docker image
+The [streamFindData](https://github.com/ricardobachertdacunha/streamFindData) R package holds the data used in examples and other documentation assets of the streamFind R package and can be installed from the GitHub repository with the following code line.
 
-Not yet available.
-
-## Backend usage
-
-Consult the Handbook_backend_streamFind.Rmd in the handbooks folder for detailed back-end structure and usage documentation.
+```{r install-streamfinddata, eval=FALSE}
+remotes::install_github("ricardobachertdacunha/streamFindData")
+```

### 2023-03-07 README.Rmd
Updated README and cleaning

@@
 
-The R package [streamFind](https://github.com/ricardobachertdacunha/streamFind) (this repository) is the back-end platform for the streamFind project entitled "Flexible data analysis and workflow designer to identify chemicals in the water cycle".
+The R package [streamFind](https://github.com/ricardobachertdacunha/streamFind) (this repository) is the back-end framework for the streamFind project entitled "Flexible data analysis and workflow designer to identify chemicals in the water cycle".
 The streamFind project is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de) and is a cooperation between the [Institut für Energie- und Umwelttechnik e.
@@
 
-[R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) for Windows users.
-The `r R.version.string` can be obtained in https://cran.r-project.org/.
-The RTools can be downloaded in https://cran.r-project.org/bin/windows/Rtools/.
-Make sure to download the right version for the R version installed.
-Installation instructions are given in both sources.
-
-The patRoon's [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites) offers detailed information on how to install the [patRoon](https://github.com/rickhelmus/patRoon) R package and the required [dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies).
+[R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) (the latter is only recommended for Windows users).
+The [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites) of the [patRoon](https://github.com/rickhelmus/patRoon) R package offers detailed installation instructions, including the required [dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies).
 
@@
 
-The [streamFind](https://github.com/ricardobachertdacunha/streamFind) R package can be installed from the GitHub repository with the following code line.
+The [streamFind](https://github.com/ricardobachertdacunha/streamFind) R package can be installed from the GitHub repository.
 
@@
 
-The [streamFindData](https://github.com/ricardobachertdacunha/streamFindData) R package holds the data used in examples and other documentation assets of the streamFind R package and can be installed from the GitHub repository with the following code line.
+The [streamFindData](https://github.com/ricardobachertdacunha/streamFindData) R package holds the data used in examples and other documentation assets of the streamFind R package and can be installed from the GitHub repository.
 

### 2023-03-07 dev/dev_assets.R

@@ -1,0 +1,82 @@
+
+# Development ------------------------------------------------------------------
+
+devtools::document()
+devtools::document()
+
+devtools::install(upgrade = "never", dependencies = TRUE, build_vignettes = TRUE)
+
+devtools::check()
+
+devtools::load_all() #Ctrl+Shift+L
+
+todor::todor()
+
+system.file(package = "streamFind", dir = "extdata")
+
+usethis::use_build_ignore("man-roxygen")
+
+# Code lines -------------------------------------------------------------------
+
+#gsub(".*M(.*)\\].*", "\\1", adduct_ion)
+
+
+
+
+# Others -----------------------------------------------------------------------
+
+## Add internal datasets ----
+## If you have data in your package
+usethis::use_data_raw(name = "my_dataset", open = FALSE)
+
+## Tests ----
+## Add one line by test you want to create
+usethis::use_test("app")
+
+# Documentation
+
+## Vignette ----
+usethis::use_vignette("handbook_backend")
+devtools::build_vignettes()
+
+## Code Coverage ----
+## Set the code coverage service ("codecov" or "coveralls")
+usethis::use_coverage()
+
+# Create a summary readme for the testthat subdirectory
+covrpage::covrpage()
+
+## CI ----
+## Use this part of the script if you need to set up a CI
+## service for your application
+##
+## (You'll need GitHub there)
+usethis::use_github()
+
+# GitHub Actions
+usethis::use_github_action()
+# Chose one of the three
+# See https://usethis.r-lib.org/reference/use_github_action.html
+usethis::use_github_action_check_release()
+usethis::use_github_action_check_standard()
+usethis::use_github_action_check_full()
+# Add action for PR
+usethis::use_github_action_pr_commands()
+
+# Travis CI
+usethis::use_travis()
+usethis::use_travis_badge()
+
+# AppVeyor
+usethis::use_appveyor()
+usethis::use_appveyor_badge()
+
+# Circle CI
+usethis::use_circleci()
+usethis::use_circleci_badge()
+
+# Jenkins
+usethis::use_jenkins()
+
+# GitLab CI
+usethis::use_gitlab_ci()

### 2023-03-07 README.Rmd
Implementation of pkgdown and fixes to roxygen documentation

@@
 
-The R package [streamFind](https://github.com/ricardobachertdacunha/streamFind) (this repository) is the back-end framework for the streamFind project entitled "Flexible data analysis and workflow designer to identify chemicals in the water cycle".
+The R package [streamFind](https://github.com/ricardobachertdacunha/streamFind) is the back-end framework for the streamFind project entitled "Flexible data analysis and workflow designer to identify chemicals in the water cycle".
 The streamFind project is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de) and is a cooperation between the [Institut für Energie- und Umwelttechnik e.
@@
 ```
+
+## Documentation
+
+Documentation and usage articles can be found in [https://ricardobachertdacunha.github.io/streamFind/](https://ricardobachertdacunha.github.io/streamFind/).

### 2023-03-08 README.Rmd

--- 2023-03-07 README.Rmd
@@
 
-The R package [streamFind](https://github.com/ricardobachertdacunha/streamFind) (this repository) is the back-end framework for the streamFind project entitled "Flexible data analysis and workflow designer to identify chemicals in the water cycle".
-The streamFind project is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de) and is a cooperation between the [Institut für Energie- und Umwelttechnik e.
+The streamFind project, entitled "Flexible data analysis and workflow designer to identify chemicals in the water cycle", is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de) and is a cooperation between the [Institut für Energie- und Umwelttechnik e.
 V.
 (IUTA)](https://www.iuta.de) and [Forschungszentrum Informatik (FZI)](https://www.fzi.de/).
-The major focus of streamFind is data processing for environmental and quality studies.
+The major focus of streamFind is the development and assembly of workflows for processing data from mass spectrometry and spectroscopy and for application the workflows in environmental and quality studies of the water cycle.
 The streamFind aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data.
@@
 
+# streamFind R package
+
+The R package [streamFind](https://github.com/ricardobachertdacunha/streamFind) is the back-end framework for the streamFind project and can be obtained with this repository.
+
 ## Installation
 
-For installation of the [streamFind](https://github.com/ricardobachertdacunha/streamFind) R package, it is recommended to firstly ensure that the dependencies are currently installed.
-Several existing R packages and external software are used in streamFind for various processing steps.
-A major dependency of streamFind is the [patRoon](https://github.com/rickhelmus/patRoon) R package.
-The [patRoon](https://github.com/rickhelmus/patRoon) R package combines several tools for basic and advanced data processing and can be used interchangeably with streamFind.
+For installation of the streamFind R package, it is recommended to first install the dependencies.
+Besides [R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) (the latter is only recommended for Windows users), the streamFind depends on the [patRoon](https://github.com/rickhelmus/patRoon) R package.
+The patRoon R package combines several tools for basic and advanced data processing and can be used interchangeably with streamFind.
+The [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites) of the patRoon R package offers detailed installation instructions, including the list of required [dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies).
 
-### Install dependencies
-
-[R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) (the latter is only recommended for Windows users).
-The [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites) of the [patRoon](https://github.com/rickhelmus/patRoon) R package offers detailed installation instructions, including the required [dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies).
-
-### Install streamFind
-
-The [streamFind](https://github.com/ricardobachertdacunha/streamFind) R package can be installed from the GitHub repository.
+The streamFind R package can then be installed from the GitHub repository.
 
@@
 ```
+
+### Documentation
+
+The documentation and usage examples of the streamFind package can be found in the [reference page](https://ricardobachertdacunha.github.io/streamFind/reference/index.html) and [articles](https://ricardobachertdacunha.github.io/streamFind/articles/index.html) of the [webpage](https://ricardobachertdacunha.github.io/streamFind/index.html), respectively.

### 2023-03-08 README.Rmd

@@
 V.
-(IUTA)](https://www.iuta.de) and [Forschungszentrum Informatik (FZI)](https://www.fzi.de/).
-The major focus of streamFind is the development and assembly of workflows for processing data from mass spectrometry and spectroscopy and for application the workflows in environmental and quality studies of the water cycle.
+(IUTA)](https://www.iuta.de) and the [Forschungszentrum Informatik (FZI)](https://www.fzi.de/).
+The goal of the streamFind project is the development and assembly of data processing workflows for mass spectrometry and spectroscopy and the application of the workflows in environmental and quality studies of the water cycle.
 The streamFind aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data.
-The streamFind package can be used by academics but also by technicians due to the comprehensive documentation and well categorized set of integrated functions or modules (see vignettes for more information).
+The streamFind platform is directed to academics but also technicians, due to the aspired comprehensive documentation, well categorized set of integrated functions (or modules) and the graphical user interface (GUI).
 
@@
 
-The R package [streamFind](https://github.com/ricardobachertdacunha/streamFind) is the back-end framework for the streamFind project and can be obtained with this repository.
+The [streamFind](https://github.com/ricardobachertdacunha/streamFind) back-end framework is an R package.
 
@@
 Besides [R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) (the latter is only recommended for Windows users), the streamFind depends on the [patRoon](https://github.com/rickhelmus/patRoon) R package.
-The patRoon R package combines several tools for basic and advanced data processing and can be used interchangeably with streamFind.
-The [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites) of the patRoon R package offers detailed installation instructions, including the list of required [dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies).
+The patRoon R package combines several tools for basic and advanced data processing and can be used interchangeably with the streamFind R package.
+The [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites) of the patRoon R package offers detailed installation instructions, including the list of mutually required [dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies).
 
-The streamFind R package can then be installed from the GitHub repository.
+Then, the streamFind R package can then be installed from the GitHub repository.
 
@@
 
-The [streamFindData](https://github.com/ricardobachertdacunha/streamFindData) R package holds the data used in examples and other documentation assets of the streamFind R package and can be installed from the GitHub repository.
+The supplementary [streamFindData](https://github.com/ricardobachertdacunha/streamFindData) R package holds the data used in examples and other documentation assets of the streamFind R package and can be installed from the GitHub repository.
 

### 2023-03-08 README.Rmd
Updated webpage

@@
 
-The documentation and usage examples of the streamFind package can be found in the [reference page](https://ricardobachertdacunha.github.io/streamFind/reference/index.html) and [articles](https://ricardobachertdacunha.github.io/streamFind/articles/index.html) of the [webpage](https://ricardobachertdacunha.github.io/streamFind/index.html), respectively.
+The documentation and usage examples of the streamFind R package can be found in the [reference page](https://ricardobachertdacunha.github.io/streamFind/reference/index.html) and [articles](https://ricardobachertdacunha.github.io/streamFind/articles/index.html) of the [webpage](https://ricardobachertdacunha.github.io/streamFind/index.html), respectively.

### 2023-03-09 README.Rmd
Further text corrections

--- 2023-03-08 README.Rmd
@@
 V.
-(IUTA)](https://www.iuta.de) and the [Forschungszentrum Informatik (FZI)](https://www.fzi.de/).
+(IUTA)](https://www.iuta.de), the [Forschungszentrum Informatik (FZI)](https://www.fzi.de/) and supporting partners.
 The goal of the streamFind project is the development and assembly of data processing workflows for mass spectrometry and spectroscopy and the application of the workflows in environmental and quality studies of the water cycle.
 The streamFind aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data.
-The streamFind platform is directed to academics but also technicians, due to the aspired comprehensive documentation, well categorized set of integrated functions (or modules) and the graphical user interface (GUI).
+The streamFind platform is directed to academics but also technicians, due to the aspired comprehensive documentation, well categorized set of integrated modular functions and the graphical user interface.
+The streamFind development is ongoing, please [contact us](mailto:cunha@iuta.de) for questions  or collaboration.
 
@@
 
-The [streamFind](https://github.com/ricardobachertdacunha/streamFind) back-end framework is an R package.
+The back-end framework of [streamFind](https://github.com/ricardobachertdacunha/streamFind) is an R package.
 
@@
 For installation of the streamFind R package, it is recommended to first install the dependencies.
-Besides [R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) (the latter is only recommended for Windows users), the streamFind depends on the [patRoon](https://github.com/rickhelmus/patRoon) R package.
+Besides [R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) (the latter is only recommended for Windows users), the streamFind depends on the [patRoon](https://github.com/rickhelmus/patRoon) R package and its dependencies.
 The patRoon R package combines several tools for basic and advanced data processing and can be used interchangeably with the streamFind R package.
-The [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites) of the patRoon R package offers detailed installation instructions, including the list of mutually required [dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies).
+Installation instructions for patRoon and its dependencies can be found [here](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites).
 
-Then, the streamFind R package can then be installed from the GitHub repository.
+Then, the streamFind R package can be installed from the GitHub repository.
 
@@
 
-The supplementary [streamFindData](https://github.com/ricardobachertdacunha/streamFindData) R package holds the data used in examples and other documentation assets of the streamFind R package and can be installed from the GitHub repository.
+The supplementary [streamFindData](https://github.com/ricardobachertdacunha/streamFindData) R package holds the data used in examples and other documentation assets of the streamFind R package and can also be installed from the GitHub repository.
 
@@
 
-The documentation and usage examples of the streamFind package can be found in the [reference page](https://ricardobachertdacunha.github.io/streamFind/reference/index.html) and [articles](https://ricardobachertdacunha.github.io/streamFind/articles/index.html) of the [webpage](https://ricardobachertdacunha.github.io/streamFind/index.html), respectively.
+The documentation and usage examples of the streamFind R package can be found in the [reference page](https://ricardobachertdacunha.github.io/streamFind/reference/index.html) and [articles](https://ricardobachertdacunha.github.io/streamFind/articles/index.html) of the [webpage](https://ricardobachertdacunha.github.io/streamFind/index.html).

### 2023-03-27 dev/dev_assets.R
added references and clean up

@@ -1,31 +1,36 @@
 
 # Development ------------------------------------------------------------------
 
+# Updates documentation with roxygen
 devtools::document()
 devtools::document()
 
+# Installs the package
 devtools::install(upgrade = "never", dependencies = TRUE, build_vignettes = TRUE)
 
+# Builds/Updates the GitHub page
 pkgdown::build_site()
 
+# Check the package for conformity
 devtools::check()
 
+# Loads all dependencies
 devtools::load_all() #Ctrl+Shift+L
 
+# Lists TODOs
 todor::todor()
 
-system.file(package = "streamFind", dir = "extdata")
-
-usethis::use_build_ignore("man-roxygen")
 
 # Code lines -------------------------------------------------------------------
 
 #gsub(".*M(.*)\\].*", "\\1", adduct_ion)
 
 
-
-
 # Others -----------------------------------------------------------------------
+
+system.file(package = "streamFind", dir = "extdata")
+
+usethis::use_build_ignore("man-roxygen")
 
 ## Add internal datasets ----
 ## If you have data in your package

### 2023-03-28 R/fct_ms_002_parse_data_from_files.R
amended cache messages

@@ -79,7 +79,7 @@ parse_ms_spectra <- function(files = NA_character_, levels = c(1, 2),
       spec_list <- patRoon::loadCacheData("parsed_ms_spectra", hash)
 
       if (!is.null(spec_list)) {
-        message("\U2713 Spectra loaded from cache!")
+        message("\U24d8 Spectra loaded from cache!")
         cached_spectra <- TRUE
         return(spec_list)
       }
@@ -89,7 +89,7 @@ parse_ms_spectra <- function(files = NA_character_, levels = c(1, 2),
     }
 
     if (!cached_spectra) {
-      message("Parsing spectra from ", length(files),  " MS file/s..." ,
+      message("\U2699 Parsing spectra from ", length(files),  " MS file/s..." ,
         appendLF = FALSE
       )
     }
@@ -681,12 +681,12 @@ parse_ms_spectra <- function(files = NA_character_, levels = c(1, 2),
   # end -----
 
   if (!cached_spectra & !is.null(spec_list)) {
-    message(" Done!") #\U2713
+    message(" Done!")
   }
 
   if (!cached_spectra & !is.null(hash)) {
     if (!is.null(spec_list)) {
-      message("\U2713 Parsed spectra cached!")
+      message("\U1f5ab Parsed spectra cached!")
       patRoon::saveCacheData("parsed_ms_spectra", spec_list, hash)
     }
   }

### 2023-03-28 dev/dev_assets.R

@@ -23,6 +23,19 @@ todor::todor()
 
 # Code lines -------------------------------------------------------------------
 
+# \U1f5ab floppy disk (clear background)
+# \U1f5aa floppy disk
+# \U23f1 stopwatch for time tasks
+# \U231b ample for time tasks
+# \U2713 check mark for modified by reference
+# \U1D48A italic i for info messages
+# \U2139 i for information messages
+# \U24d8 i from information messages with round circle
+# \U2713 check for ready
+# \U1f5f9 check in a box
+
+
+
 #gsub(".*M(.*)\\].*", "\\1", adduct_ion)
 
 

### 2023-04-24 R/class_S3_ProcessingSettings.R
Changing Naming to UpperCamelCase

@@ -1,25 +1,26 @@
-#' **settings** S3 class constructor and methods
+#' **ProcessingSettings** S3 class constructor and methods
 #'
 #' @description
-#' Creates a settings S3 class.
+#' Creates a ProcessingSettings S3 class object.
 #'
 #' @param call Character of length one with the name of the method where the
-#' settings are to be applied.
+#' processing settings are to be applied.
 #' @param algorithm Character of length one with the name of the algorithm to
 #' be used.
-#' @param parameters List with parameter settings specific for the `call`
-#' method and used `algorithm`.
+#' @param parameters List with parameters specific for the method `call` and
+#' `algorithm`.
 #'
 #' @details See the method documentation for more information about possible
 #' algorithms and parameters to be used.
 #'
-#' @return A settings S3 class
+#' @return A ProcessingSettings S3 class
 #'
 #' @export
 #'
-settings <- function(call = NA_character_,
-                     algorithm = NA_character_,
-                     parameters = NULL) {
+ProcessingSettings <- function(call = NA_character_,
+                               algorithm = NA_character_,
+                               parameters = NULL) {
+
   x <- list(
     "call" = call,
     "algorithm" = algorithm,
@@ -58,22 +59,22 @@ settings <- function(call = NA_character_,
     }
   })
 
-  if (validate.settings(x)) {
-    structure(x, class = "settings")
+  if (validate.ProcessingSettings(x)) {
+    structure(x, class = "ProcessingSettings")
   } else {
     NULL
   }
 }
 
-#' @describeIn settings
-#' Validates a settings S3 class object, returning a logical value of length
-#' one.
+#' @describeIn ProcessingSettings
+#' Validates a ProcessingSettings S3 class object, returning a logical value of
+#' length one.
 #'
-#' @param x A settings S3 class object.
+#' @param x A ProcessingSettings S3 class object.
 #'
 #' @export
 #'
-validate.settings <- function(x = NULL) {
+validate.ProcessingSettings <- function(x = NULL) {
   valid <- FALSE
 
   if (is.list(x)) {
@@ -140,16 +141,16 @@ validate.settings <- function(x = NULL) {
   valid
 }
 
-#' @describeIn settings
-#' Converts the argument in a settings S3 class object
+#' @describeIn ProcessingSettings
+#' Converts the argument in a ProcessingSettings S3 class object.
 #'
-#' @param value A list to be checked and/or converted to settings S3 class.
[…]

### 2023-09-03 dev/dev_template_report.Rmd
Template project rmarkdown report

--- 2022-11-29 dev/template_report.Rmd
@@
 ---
-title: "ntsIUTA Report Template"
+title: Template report example
+subtitle: subtitle
 author:
-  name: "Ricardo Cunha"
-  affiliation: "Insitut für Energie- und Umwelttechnik e.V. (IUTA)"
-  email: cunha@iuta.de
+  name: author
+  email: email@example.com
+  affiliation: "name <br> address"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
     theme: paper
-    code_folding: hide
 editor_options: 
@@
 
-path_extdata <- system.file(package = "streamFind", dir = "extdata")
-logo <- file.path(path_extdata, "logo_iuta.png")
+# Libraries
+library(knitr)
+library(kableExtra)
+library(magrittr)
+library(ggplot2)
+library(plotly)
+library(streamFind)
 
+# Variables
+logo <- NA_character_
+
+# Global options
 knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 9, results = "asis")
 
+# Working directory
 path_wd <- getwd()
+
+# Add logo file
+path_extdata <- system.file(package = "streamFindData", dir = "extdata")
+logo <- file.path(path_extdata, "logo_iuta.png") #207 x 216 px
 ```
@@
 ```{r logo, echo=FALSE}
+if (!is.na(logo)) {
   htmltools::img(
@@
   ) 
+}
 ```
 
-<br> <br>
+<br>
 
 ***
-
-```{r libraries, include=FALSE}
-
-library(knitr)
-library(kableExtra)
-library(magrittr)
-library(ggplot2)
-library(plotly)
-library(streamFind)
-
-```
 
@@
 
-```{r exCode}
+```{r exampleCode}
 
@@
 
-The `knitr::kable` function can be used for displaying tables, see table \@ref(tab:exTable).
+The `knitr::kable` function can be used for displaying tables, see table \@ref(tab:exampleTable).
 Consult <https://haozhu233.github.io/kableExtra/awesome_table_in_html.html> for other costumization options.
@@
 
-```{r exTable}
+```{r exampleTable}
 
@@
 ) %>%
-  scroll_box(width = "100%", height = "600px")
+  scroll_box(width = "100%", height = "350px")
 
@@
 
-```{r exGgplot, fig.cap="Plot caption example"}
+```{r examplePlot, fig.cap="Plot caption example"}
 
@@
 
-```{r exPlotly, fig.cap="Plot caption example"}
+```{r examplePlotly, fig.cap="Plot caption example"}
 
@@
 
-<br>
-
 ***
-**Insitut für Energie- und Umwelttechnik e.V.**  
-A: Bliersheimer Str. 58 - 60, 47229 Duisburg  
-T: (02065) 418 - 0
 

### 2023-09-06 dev/dev_exchange_ude.R
Updated streamFind to StreamFind and ricardobachertdacunha to odea-project

@@ -1,8 +1,8 @@
 
 # resources --------------------------------------------------------------------
 
-all_files <- streamFindData::msFilePaths()
-db <- streamFindData::msSpikedChemicals()
+all_files <- StreamFindData::msFilePaths()
+db <- StreamFindData::msSpikedChemicals()
 files_mrm <- all_files[grepl("mrm", all_files)]
 files <- all_files[1:3]
 files1 <- all_files[grepl("influent|blank", all_files)]

### 2023-09-06 README.Rmd
added references and clean up
streamFindData not StreamFindData

@@
 output: github_document
+bibliography: inst/REFERENCES.bib
 ---
@@
 
-# streamFind
+# StreamFind
 
@@
 
-The streamFind project, entitled "Flexible data analysis and workflow designer to identify chemicals in the water cycle", is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de) and is a cooperation between the [Institut für Energie- und Umwelttechnik e.
+The StreamFind project, entitled "Flexible data analysis and workflow designer to identify chemicals in the water cycle", is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de) and is a cooperation between the [Institut für Energie- und Umwelttechnik e.
 V.
 (IUTA)](https://www.iuta.de), the [Forschungszentrum Informatik (FZI)](https://www.fzi.de/) and supporting partners.
-The goal of the streamFind project is the development and assembly of data processing workflows for mass spectrometry and spectroscopy and the application of the workflows in environmental and quality studies of the water cycle.
-The streamFind aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data.
-The streamFind platform is directed to academics but also technicians, due to the aspired comprehensive documentation, well categorized set of integrated modular functions and the graphical user interface.
-The streamFind development is ongoing, please [contact us](mailto:cunha@iuta.de) for questions  or collaboration.
+The goal of the StreamFind project is the development and assembly of data processing workflows for mass spectrometry and spectroscopy and the application of the workflows in environmental and quality studies of the water cycle.
+The StreamFind aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data.
+The StreamFind platform is directed to academics but also technicians, due to the aspired comprehensive documentation, well categorized set of integrated modular functions and the graphical user interface.
+The StreamFind development is ongoing, please [contact us](mailto:cunha@iuta.de) for questions  or collaboration.
 
@@
 
-# streamFind R package
+# StreamFind R package
 
-The back-end framework of [streamFind](https://github.com/ricardobachertdacunha/streamFind) is an R package.
+The back-end framework of [StreamFind](https://github.com/odea-project/StreamFind) is an R package.
 
@@
 
-For installation of the streamFind R package, it is recommended to first install the dependencies.
-Besides [R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) (the latter is only recommended for Windows users), the streamFind depends on the [patRoon](https://github.com/rickhelmus/patRoon) R package and its dependencies.
-The patRoon R package combines several tools for basic and advanced data processing and can be used interchangeably with the streamFind R package.
+For installation of the StreamFind R package, it is recommended to first install the dependencies.
+Besides [R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) (the latter is only recommended for Windows users), the StreamFind depends on the [patRoon](https://github.com/rickhelmus/patRoon) R package and its dependencies.
+The patRoon R package combines several tools for basic and advanced data processing and can be used interchangeably with the StreamFind R package.
 Installation instructions for patRoon and its dependencies can be found [here](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites).
 
-Then, the streamFind R package can be installed from the GitHub repository.
+Then, the StreamFind R package can be installed from the GitHub repository.
 
 ```{r install-streamfind, eval=FALSE}
-remotes::install_github("ricardobachertdacunha/streamFind", dependencies = TRUE)
+remotes::install_github("odea-project/StreamFind", dependencies = TRUE)
 ```
 
-The supplementary [streamFindData](https://github.com/ricardobachertdacunha/streamFindData) R package holds the data used in examples and other documentation assets of the streamFind R package and can also be installed from the GitHub repository.
+The supplementary [streamFindData](https://github.com/odea-project/streamFindData) R package holds the data used in examples and other documentation assets of the StreamFind R package and can also be installed from the GitHub repository.
 
 ```{r install-streamfinddata, eval=FALSE}
-remotes::install_github("ricardobachertdacunha/streamFindData")
+remotes::install_github("odea-project/streamFindData")
 ```
@@
 
-The documentation and usage examples of the streamFind R package can be found in the [reference page](https://ricardobachertdacunha.github.io/streamFind/reference/index.html) and [articles](https://ricardobachertdacunha.github.io/streamFind/articles/index.html) of the [webpage](https://ricardobachertdacunha.github.io/streamFind/index.html).
+The documentation and usage examples of the StreamFind R package can be found in the [reference page](https://odea-project.github.io/StreamFind/reference/index.html) and [articles](https://odea-project.github.io/StreamFind/articles/index.html) of the [webpage](https://odea-project.github.io/StreamFind/index.html).
+
+# References
+
+---
+nocite: '@*'
+---

### 2023-09-07 README.Rmd
Updated StreamFindData calls and pkgdown

@@
 
-The supplementary [streamFindData](https://github.com/odea-project/streamFindData) R package holds the data used in examples and other documentation assets of the StreamFind R package and can also be installed from the GitHub repository.
+The supplementary [StreamFindData](https://github.com/odea-project/StreamFindData) R package holds the data used in examples and other documentation assets of the StreamFind R package and can also be installed from the GitHub repository.
 
 ```{r install-streamfinddata, eval=FALSE}
-remotes::install_github("odea-project/streamFindData")
+remotes::install_github("odea-project/StreamFindData")
 ```

### 2023-09-07 dev/dev_ms_basic_structure.R

@@ -1,7 +1,7 @@
 
-all_files <- streamFindData::msFilePaths()
+all_files <- StreamFindData::get_all_file_paths()
 
-db <- streamFindData::msSpikedChemicals()
+db <- StreamFindData::get_tof_spiked_chemicals()
 db <- db[grepl("S", db$tag), ]
 cols <- c("name", "formula", "mass", "rt")
 db <- db[, cols, with = FALSE]

### 2023-09-07 README.Rmd
Updated logo and pkgdown

@@
 
-The StreamFind project, entitled "Flexible data analysis and workflow designer to identify chemicals in the water cycle", is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de) and is a cooperation between the [Institut für Energie- und Umwelttechnik e.
+The StreamFind project, entitled "Flexible data analysis and workflow designer to identify chemicals in the water cycle", is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de) and is a cooperation between the [Institut für Umwelt & Energie, Technik & Analytik e.
 V.

### 2023-09-22 dev/demo_ozonation_project.Rmd
Improved filter_features(), added Rmarkdown ozonation demo and other fixes

---
title: 'StreamFind: Wastewater Ozonation'
subtitle: Mass spectrometry project example
author:
  name: Ricardo Cunha
  email: cunha@iuta.de
  affiliation: "Institut für Umwelt & Energie, Technik & Analytik e. V. (IUTA) <br> Bliersheimer Str. 58 - 60, 47229 Duisburg"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    theme: paper
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

# Libraries
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(plotly)
library(StreamFind)

# Variables
logo <- NA_character_

# Global options
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 9, results = "asis")

# Working directory
path_wd <- getwd()

# Add logo file
path_extdata <- system.file(package = "StreamFindData", dir = "extdata")
logo <- file.path(path_extdata, "logo_StreamFind.png") #207 x 216 px
```

```{r logo, echo=FALSE}
if (!is.na(logo)) {
  htmltools::img(
  src = knitr::image_uri(logo),
  alt = "logo",
  style =
    paste0("
      margin-left:inherit;
      position:absolute;
      top:100px;
      left:610px;
      width:", 250, "px;
      height:", 250*(1633/3070), "px;"
    )
  ) 
} # height:", 100*(216/207), "px;"
```

<br>

```{r resources, include=FALSE}
# MS files
all_files <- StreamFindData::get_all_file_paths()
files <- all_files[grepl("blank|influent|o3sw", all_files)]

# Chemicals database
db <- StreamFindData::get_tof_spiked_chemicals()
db <- db[grepl("S", db$tag), ]
cols <- c("name", "formula", "mass", "rt", "tag")
db <- db[, cols, with = FALSE]
db
```

***

# Create a *MassSpecData*

An [R6](https://r6.r-lib.org/) [MassSpecData](https://odea-project.github.io/StreamFind/reference/MassSpecData.html) class object is created using the `MassSpecData$new()`, as shown below for a set of *mzML* files. The original vendor files were converted to *mzML* format using the [convert_ms_files](https://odea-project.github.io/StreamFind/reference/convert_ms_files.html) function, which uses the *msConvert* command line from [ProteoWizard](https://proteowizard.sourceforge.io/download.html). The files are from influent and effluent of a wastewater ozonation step measured in positive and negative ionization mode. Blank measurements for each ionization were also included for background subtraction.

```{r files, results='markup'}
basename(files)
```

```{r create-ms, message=FALSE, warning=FALSE}
# Create a MassSpecData from mzML files
ms <- MassSpecData$new(files = files)
```

```{r show-ms, results='markup'}
# Print in console a summary of the MassSpecData
ms
```

## Add *ProjectHeaders*

Project headers (e.g., name, author and description) can be added to the MassSpecData as shown below. The headers are converted to an S3 [ProjectHeaders](https://odea-project.github.io/StreamFind/reference/ProjectHeaders.html) class object.

```{r add-headers, results='markup'}
# Add headers to the MassSpecData
ms$add_headers(
  name = "Wastewater Ozonation",
  author = "Ricardo Cunha",
  description = "StreamFind demonstration project"
)

# Get the headers as ProjectHeaders
ms$get_headers()

# Get the MassSpecData name
ms$get_headers(value = "name")
```

## Associate blanks

The analysis replicate names and the associated blank replicate name can be ammended in the MassSpecData as shown below.

```{r add-replicates, results='markup'}

# Character vector with analysis replicate names
rpls <- c(
  rep("blank_neg", 3),
  rep("blank_pos", 3),
  rep("in_neg", 3),
  rep("in_pos", 3),
  rep("out_neg", 3),
  rep("out_pos", 3)
)

# Character vector with associated blank replicate names
# Note that the order should match the respective replicate
blks <- c(
  rep("blank_neg", 3),
  rep("blank_pos", 3),
  rep("blank_neg", 3),
  rep("blank_pos", 3),
  rep("blank_neg", 3),
  rep("blank_pos", 3)
)

# Chaining add replicates and blanks
ms$add_replicate_names(rpls)$add_blank_names(blks)

# Get the blank replicate names
ms$get_blank_names()
```

# Data processing

Data processing is performed by modules according to [ProcessingSettings](https://odea-project.github.io/StreamFind/reference/ProcessingSettings.html) as shown in the following sub-chapters.

## Obtain *ProcessingSettings*

The S3 *ProcessingSettings* class objects for processing modules are obtained via the respective [Settings](https://odea-project.github.io/StreamFind/reference/index.html#processing-settings-for-ms-data) functions. Below we obtain the *ProcessingSettings* for the module `find_features()` using the algorithm *xcms3_centwave*. The parameters for each processing module can be changed via the function arguments and the information about the meaning of each parameter can be found in the [StreamFind reference documentation](https://odea-project.github.io/StreamFind/reference/index.html).

```{r ffs, results='markup'}
# Get ProcessingSettings for finding features using the xcms3 centwave algorithm
ffs <- Settings_find_features_xcms3_centwave()

# Print in console the details of the ProcessingSettings
ffs
```

```{r other-settings, results='markup'}
fls <- Settings_filter_features_StreamFind(
  minIntensity = 5000,
  minSnRatio = 20,
  maxGroupSd = 30,
  blank = 5,
  minGroupAbundance = 3,
  excludeIsotopes = TRUE
)

fls

afs <- Settings_annotate_features_StreamFind()

afs

gfs <- Settings_group_features_xcms3_peakdensity()

gfs
```

## Add *ProcessingSettings*

When all settings are available, they can be added or used directly to run the workflow. In the example below, the different *ProcessingSettings* are first added to the *MassSpecData* in the desired order.

```{r add-settings, results='markup'}
# Add settings for finding, annotating, grouping and finally filter features
ms$add_settings(settings = list(ffs, afs, gfs, fls))

# Getting the workflow settings
ms$get_settings_names()
```

## Apply the workflow

The workflow can be run by chaining all the modules as demonstrated below or by running the method `run_workflow()`.

```{r do-workflow, results='markup', warning=FALSE}

ms$run_workflow()

ms

#ms$find_features()$annotate_features()$group_features()$filter_features()
```

# Instructions

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents.
For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
Diverse cheatsheets can be found in <https://www.rstudio.com/resources/cheatsheets/>.
This section gives quick instructions to display text, code, tables and figures.

## Text

Text in R markdown has several inline codes for costumization.
The book in <https://bookdown.org/yihui/rmarkdown/> offers a detailed guideline.
Moreover, a cheatsheet can be downdloaded with the following link <https://raw.githubusercontent.com/rstudio/cheatsheets/main/rmarkdown.pdf>.

## Code

Code is automaticelly embedded unless `echo` chunk option is set to `FALSE`.
Evaluation of code in a given chunk can be skiped by setting `eval` to `FALSE`.
Code chunks should always be unically named.

```{r exampleCode}

#Example of code chunk

list(a = 1:6, b = 1:10)

```

<br>

## Tables

The `knitr::kable` function can be used for displaying tables, see table \@ref(tab:exampleTable). Consult <https://haozhu233.github.io/kableExtra/awesome_table_in_html.html> for other costumization options.
Table caption is added with the argument `caption` of the `kable` function.

```{r exampleTable}

df <- data.frame(name = c("a", "b", "c", "d", "e", "f"), col1 = 1:6, col2 = 1:6)

knitr::kable(df, caption = "Example of table caption.") %>%
kable_styling(
  font_size = 12,
  bootstrap_options = c("striped", "hover", "condensed", "responsive"),
  fixed_thead = TRUE
) %>%
  scroll_box(width = "100%", height = "350px")

```

<br>

## Figures

Figures/plots can also by embedded using diverse packages.
The packages `ggplot2` and `plotly` (interactive) are recommended as they are versatile and easy to use.
Cheatsheet for `ggplot2` can be found in <https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-visualization.pdf>.
For `plotly` the website <https://plotly.com/r/> offers examples and a cheatsheet can be downdloaded in <https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf>.
Note that the `echo = FALSE` parameter can be added to the code chunk to prevent printing of the R code that generate the plots.

### Example `ggplot2`

```{r examplePlot, fig.cap="Plot caption example"}

ggplot(pressure, aes(temperature, pressure)) +
geom_point() +
theme_bw()

```

### Example `plotly`

```{r examplePlotly, fig.cap="Plot caption example"}

plot_ly(pressure, x = ~temperature, y = ~pressure, type = "scatter", mode = "markers+lines")

```

<br> <br>

# Signature

The report was produced on `r Sys.time()`. The data and report are stored in ``r path_wd`` of the workstation *`r Sys.info()["nodename"]`* under the user *`r Sys.info()["user"]`*.

***

<br>

<style>
body, p {
  color: black;
  font-family: Arial;
  text-align: justify;
  font-size: 12pt;
}
h1{
  font-size: 20pt;
  font-style: bold;
}
h2{
  font-size:18pt;
  font-style: bold;
}
h3{
  font-size:14pt;
  font-style: bold;
}
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: #177013;
}
caption {
  color: black;
  font-size: 1.0em;
}
</style>

### 2023-09-22 R/class_R6_MassSpecData.R

@@ -203,10 +203,13 @@ MassSpecData <- R6::R6Class("MassSpecData",
     ## ___ .filters -----
 
     #' @description
-    #' Filters features and feature groups with minimum intensity.
+    #' Tags features and feature groups with filter based on logical vector of 
+    #' feature groups.
     #'
     .tag_filtered = function(groups_sel, tag) {
+      
       filtered_g <- copy(private$.groups)
+      
       filtered_g <- filtered_g$group[groups_sel & !filtered_g$filtered]
 
       private$.analyses <- lapply(private$.analyses,
@@ -220,6 +223,7 @@ MassSpecData <- R6::R6Class("MassSpecData",
       )
 
       private$.groups$filtered[groups_sel] <- TRUE
+      
       private$.groups$filter[groups_sel & is.na(private$.groups$filter)] <- tag
     },
 
@@ -228,36 +232,44 @@ MassSpecData <- R6::R6Class("MassSpecData",
     #'
     .filter_minIntensity = function(value = 5000) {
 
-      if (is.numeric(value)) {
-        private$.analyses <- lapply(private$.analyses, function(x) {
-          sel <- (x$features$intensity <= value) & (!x$features$filtered)
-          x$features$filtered[sel] <- TRUE
-          x$features$filter[sel] <- "minIntensity"
-          x
-        })
-
-        if (self$has_groups()) {
-          rpl <- self$get_replicate_names()
-
-          groups <- self$get_groups(
-            filtered = TRUE, onlyIntensities = TRUE, average = TRUE
+      if (any(self$has_features())) {
+        
+        if (is.numeric(value) & length(value) == 1) {
+          
+          if (self$has_groups()) {
+            rpl <- self$get_replicate_names()
+            
+            groups <- self$get_groups(filtered = TRUE, onlyIntensities = TRUE, average = TRUE)
+            
+            groups_sel <- apply(groups[, rpl, with = FALSE], MARGIN = 1,
+                                function(x) { max(x) <= value }
+            )
+            
+            private$.tag_filtered(groups_sel, "minIntensity")
+            
+          } else {
+            private$.analyses <- lapply(private$.analyses, function(x) {
+              sel <- (x$features$intensity <= value) & (!x$features$filtered)
+              x$features$filtered[sel] <- TRUE
+              x$features$filter[sel] <- "minIntensity"
+              x
+            })
+          }
+          
+          private$.register(
+            "filter_features",
+            "features",
+            "minIntensity",
[…]

### 2023-09-25 dev/demo_ozonation_project.Rmd
Support for ion mobility data with get_spectra(), make_ms_targets() and others
Also added Rmarkdown files for ozonation demo and backend developer guide

@@
 path_extdata <- system.file(package = "StreamFindData", dir = "extdata")
-logo <- file.path(path_extdata, "logo_StreamFind.png") #207 x 216 px
+logo <- file.path(path_extdata, "logo_StreamFind.png")
 ```
@@
       position:absolute;
-      top:100px;
-      left:610px;
+      top:150px;
+      left:600px;
       width:", 250, "px;
@@
   )
-} # height:", 100*(216/207), "px;"
+}
 ```
@@
 
-The workflow can be run by chaining all the modules as demonstrated below or by running the method `run_workflow()`.
+The modules assidadded *ProcessingSettings* can be processed by `run_workflow()` as demonstrated below or by chaining all the modules.
+With `run_workflow()`, the processing modules are applied with the same order they were added.
+By chaining the order can be modified as the modules are individually called, as shown below.
 
 ```{r do-workflow, results='markup', warning=FALSE}
+# Run all ProcessingSettings added to the MassSpecData
+ms$run_workflow()
+```
 
-ms$run_workflow()
-
-ms
-
-#ms$find_features()$annotate_features()$group_features()$filter_features()
+```{r do-workflow-chain, eval=FALSE}
+# Alternatively, the modules can be chained
+ms$find_features()$annotate_features()$group_features()$filter_features()
 ```

### 2023-09-25 dev/developer_guide_backend.Rmd
Support for ion mobility data with get_spectra(), make_ms_targets() and others
Also added Rmarkdown files for ozonation demo and backend developer guide

--- 2023-09-25 dev/demo_ozonation_project.Rmd
@@
 ---
-title: 'StreamFind: Wastewater Ozonation'
-subtitle: Mass spectrometry project example
+title: 'Backend developer guide for MS data'
+subtitle: |
+  | Implementation of backend processing modules and algorithms for mass spectrometry
+  | <br>
+
 author:
@@
 
-# Create a *MassSpecData*
+# StreamFind
 
-An [R6](https://r6.r-lib.org/) [MassSpecData](https://odea-project.github.io/StreamFind/reference/MassSpecData.html) class object is created using the `MassSpecData$new()`, as shown below for a set of *mzML* files.
-The original vendor files were converted to *mzML* format using the [convert_ms_files](https://odea-project.github.io/StreamFind/reference/convert_ms_files.html) function, which uses the *msConvert* command line from [ProteoWizard](https://proteowizard.sourceforge.io/download.html).
-The files are from influent and effluent of a wastewater ozonation step measured in positive and negative ionization mode.
-Blank measurements for each ionization were also included for background subtraction.
+The backend of StreamFind is an R package and can be used for data management, processing, visualization and reporting.
+This guide focuses on mass spectrometry (MS) data and aims to instruct developers to implement new processing modules and additional processing algorithms for new or existing processing modules in StreamFind.
 
-```{r files, results='markup'}
-basename(files)
-```
+## Setup
 
-```{r create-ms, message=FALSE, warning=FALSE}
-# Create a MassSpecData from mzML files
-ms <- MassSpecData$new(files = files)
-```
+The R package is in the GitHub [StreamFind repository](https://github.com/odea-project/StreamFind) of the [ODEA Project](https://github.com/odea-project).
+For development, the recommendation is to download the repository locally using [git](https://git-scm.com/) version control.
+The [GitHub desktop](https://desktop.github.com/) tool can be used for more easily install and configure git with your GitHub account, which is recommended for authoring contributions.
+Since it is an R package, the [RStudio](https://posit.co/products/open-source/rstudio/) IDE is recommended for development.
+Yet, others (e.g., [VS Code](https://code.visualstudio.com/download)) will also work.
+When using RStudio, the repository can be downloaded via new project, selecting version control, then git and finally adding the GitHub *url* of the StreamFind: https://github.com/odea-project/StreamFind.
+This should create a local image of the GitHub StreamFind repository directly with git tracking if git or GitHub desktop were properly installed and configured.
 
-```{r show-ms, results='markup'}
-# Print in console a summary of the MassSpecData
-ms
-```
+## Development
 
-## Add *ProjectHeaders*
+Considering that the local image of the StreamFind repository is installed with git version tracking, the first step for development is to create a dedicated branch for implementation of new processing modules and/or algorithms.
+The master branch should not be changed directly but modified by pull requests from the dedicated development branch, giving the opportunity for code revision.
 
-Project headers (e.g., name, author and description) can be added to the MassSpecData as shown below.
-The headers are converted to an S3 [ProjectHeaders](https://odea-project.github.io/StreamFind/reference/ProjectHeaders.html) class object.
+## Structure
 
-```{r add-headers, results='markup'}
-# Add headers to the MassSpecData
-ms$add_headers(
-  name = "Wastewater Ozonation",
-  author = "Ricardo Cunha",
-  description = "StreamFind demonstration project"
-)
+## Semantics
 
-# Get the headers as ProjectHeaders
-ms$get_headers()
+# R6 and S3
 
-# Get the MassSpecData name
-ms$get_headers(value = "name")
-```
+# *ProcessingSettings*
 
-## Associate blanks
+## Constructor
 
-The analysis replicate names and the associated blank replicate name can be ammended in the MassSpecData as shown below.
+## S3 methods
 
-```{r add-replicates, results='markup'}
+# New module
 
-# Character vector with analysis replicate names
-rpls <- c(
-  rep("blank_neg", 3),
-  rep("blank_pos", 3),
-  rep("in_neg", 3),
-  rep("in_pos", 3),
-  rep("out_neg", 3),
-  rep("out_pos", 3)
-)
-
-# Character vector with associated blank replicate names
-# Note that the order should match the respective replicate
-blks <- c(
-  rep("blank_neg", 3),
-  rep("blank_pos", 3),
-  rep("blank_neg", 3),
-  rep("blank_pos", 3),
-  rep("blank_neg", 3),
-  rep("blank_pos", 3)
-)
-
-# Chaining add replicates and blanks
-ms$add_replicate_names(rpls)$add_blank_names(blks)
-
-# Get the blank replicate names
-ms$get_blank_names()
-```
-
-# Data processing
-
-Data processing is performed by modules according to [ProcessingSettings](https://odea-project.github.io/StreamFind/reference/ProcessingSettings.html) as shown in the following sub-chapters.
-
-## Obtain *ProcessingSettings*
-
-The S3 *ProcessingSettings* class objects for processing modules are obtained via the respective [Settings](https://odea-project.github.io/StreamFind/reference/index.html#processing-settings-for-ms-data) functions.
-Below we obtain the *ProcessingSettings* for the module `find_features()` using the algorithm *xcms3_centwave*.
-The parameters for each processing module can be changed via the function arguments and the information about the meaning of each parameter can be found in the [StreamFind reference documentation](https://odea-project.github.io/StreamFind/reference/index.html).
-
-```{r ffs, results='markup'}
-# Get ProcessingSettings for finding features using the xcms3 centwave algorithm
-ffs <- Settings_find_features_xcms3_centwave()
-
-# Print in console the details of the ProcessingSettings
-ffs
-```
-
-```{r other-settings, results='markup'}
-fls <- Settings_filter_features_StreamFind(
-  minIntensity = 5000,
-  minSnRatio = 20,
-  maxGroupSd = 30,
-  blank = 5,
-  minGroupAbundance = 3,
-  excludeIsotopes = TRUE
-)
-
-fls
-
-afs <- Settings_annotate_features_StreamFind()
-
-afs
-
-gfs <- Settings_group_features_xcms3_peakdensity()
-
-gfs
-```
-
-## Add *ProcessingSettings*
-
-When all settings are available, they can be added or used directly to run the workflow.
-In the example below, the different *ProcessingSettings* are first added to the *MassSpecData* in the desired order.
-
-```{r add-settings, results='markup'}
-# Add settings for finding, annotating, grouping and finally filter features
-ms$add_settings(settings = list(ffs, afs, gfs, fls))
-
-# Getting the workflow settings
-ms$get_settings_names()
-```
-
-## Apply the workflow
-
-The modules assidadded *ProcessingSettings* can be processed by `run_workflow()` as demonstrated below or by chaining all the modules.
-With `run_workflow()`, the processing modules are applied with the same order they were added.
-By chaining the order can be modified as the modules are individually called, as shown below.
-
-```{r do-workflow, results='markup', warning=FALSE}
-# Run all ProcessingSettings added to the MassSpecData
-ms$run_workflow()
-```
-
-```{r do-workflow-chain, eval=FALSE}
-# Alternatively, the modules can be chained
-ms$find_features()$annotate_features()$group_features()$filter_features()
-```
-
-# Instructions
-
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents.
-For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-Diverse cheatsheets can be found in <https://www.rstudio.com/resources/cheatsheets/>.
-This section gives quick instructions to display text, code, tables and figures.
-
-## Text
-
-Text in R markdown has several inline codes for costumization.
-The book in <https://bookdown.org/yihui/rmarkdown/> offers a detailed guideline.
-Moreover, a cheatsheet can be downdloaded with the following link <https://raw.githubusercontent.com/rstudio/cheatsheets/main/rmarkdown.pdf>.
-
-## Code
-
-Code is automaticelly embedded unless `echo` chunk option is set to `FALSE`.
-Evaluation of code in a given chunk can be skiped by setting `eval` to `FALSE`.
-Code chunks should always be unically named.
-
-```{r exampleCode}
-
-#Example of code chunk
-
-list(a = 1:6, b = 1:10)
-
-```
-
-<br>
-
-## Tables
-
-The `knitr::kable` function can be used for displaying tables, see table \@ref(tab:exampleTable).
-Consult <https://haozhu233.github.io/kableExtra/awesome_table_in_html.html> for other costumization options.
-Table caption is added with the argument `caption` of the `kable` function.
-
-```{r exampleTable}
-
-df <- data.frame(name = c("a", "b", "c", "d", "e", "f"), col1 = 1:6, col2 = 1:6)
-
-knitr::kable(df, caption = "Example of table caption.") %>%
-kable_styling(
-  font_size = 12,
-  bootstrap_options = c("striped", "hover", "condensed", "responsive"),
-  fixed_thead = TRUE
-) %>%
-  scroll_box(width = "100%", height = "350px")
-
-```
-
-<br>
-
-## Figures
-
-Figures/plots can also by embedded using diverse packages.
-The packages `ggplot2` and `plotly` (interactive) are recommended as they are versatile and easy to use.
-Cheatsheet for `ggplot2` can be found in <https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-visualization.pdf>.
-For `plotly` the website <https://plotly.com/r/> offers examples and a cheatsheet can be downdloaded in <https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf>.
-Note that the `echo = FALSE` parameter can be added to the code chunk to prevent printing of the R code that generate the plots.
-
-### Example `ggplot2`
-
-```{r examplePlot, fig.cap="Plot caption example"}
-
-ggplot(pressure, aes(temperature, pressure)) +
-geom_point() +
-theme_bw()
-
-```
-
-### Example `plotly`
-
-```{r examplePlotly, fig.cap="Plot caption example"}
-
-plot_ly(pressure, x = ~temperature, y = ~pressure, type = "scatter", mode = "markers+lines")
-
-```
+# New *ProcessingSettings*
 
 <br> <br>
-
-# Signature
-
-The report was produced on `r Sys.time()`. The data and report are stored in ``r path_wd`` of the workstation *`r Sys.info()["nodename"]`* under the user *`r Sys.info()["user"]`*.
 

### 2023-09-25 R/001_generics.R

@@ -47,6 +47,14 @@ export <- function(x, ...) {
   UseMethod(".s3_ms_fill_features")
 }
 
+.s3_ms_control_standards <- function(settings, self) {
+  UseMethod(".s3_ms_control_standards")
+}
+
+.s3_ms_correct_intensity <- function(settings, self) {
+  UseMethod(".s3_ms_correct_intensity")
+}
+
 .s3_ms_suspect_screening <- function(settings, self) {
   UseMethod(".s3_ms_suspect_screening")
 }

### 2023-09-25 R/class_R6_MassSpecData.R

@@ -12,8 +12,10 @@
 #' @template arg-ms-levels
 #' @template arg-ms-mz
 #' @template arg-ms-rt
+#' @template arg-ms-drift
 #' @template arg-ms-ppm
 #' @template arg-ms-sec
+#' @template arg-ms-millisec
 #' @template arg-ms-id
 #' @template arg-ms-allTraces
 #' @template arg-ms-isolationWindow
@@ -1158,8 +1160,10 @@ MassSpecData <- R6::R6Class("MassSpecData",
                            mass = NULL,
                            mz = NULL,
                            rt = NULL,
+                           drift = NULL,
                            ppm = 20,
                            sec = 60,
+                           millisec = 5,
                            id = NULL,
                            allTraces = TRUE,
                            isolationWindow = 1.3,
@@ -1188,7 +1192,7 @@ MassSpecData <- R6::R6Class("MassSpecData",
           }
         }
         
-        neutral_targets <- make_ms_targets(mass, rt, ppm, sec, id)
+        neutral_targets <- make_ms_targets(mass, rt, drift, ppm, sec, millisec, id)
         
         targets <- list()
 
@@ -1221,7 +1225,7 @@ MassSpecData <- R6::R6Class("MassSpecData",
         }
         
       } else {
-        mz_targets <- make_ms_targets(mz, rt, ppm, sec, id)
+        mz_targets <- make_ms_targets(mz, rt, drift, ppm, sec, millisec, id)
         
         if (!"polarity" %in% colnames(mz_targets)) {
           targets <- list()
@@ -1247,7 +1251,7 @@ MassSpecData <- R6::R6Class("MassSpecData",
         }
       }
 
-      num_cols <- c("mz", "rt", "mzmin", "mzmax", "rtmin", "rtmax")
+      num_cols <- c("mz", "rt", "drift", "mzmin", "mzmax", "rtmin", "rtmax", "driftmin", "driftmax")
 
       if (all(apply(targets[, num_cols, with = FALSE], 1, sum) != 0)) {
 
@@ -1257,6 +1261,10 @@ MassSpecData <- R6::R6Class("MassSpecData",
 
         if (TRUE %in% (targets$rtmax == 0)) {
           targets$rtmax[targets$rtmax == 0] <- max(self$get_rt_end(analyses))
+        }
+        
+        if (TRUE %in% (targets$driftmax == 0) & any(self$has_ion_mobility())) {
[…]

### 2023-09-25 R/utils_ms.R

@@ -1,4 +1,4 @@
-#' trim_vector
+#' .trim_vector
 #'
 #' @param v A vector to trim based on minimum and maximum value pairs.
 #' @param a A vector with minimum values to evaluate `v`.
@@ -9,11 +9,11 @@
 #'
 #' @noRd
 #'
-trim_vector <- function(v, a, b) {
+.trim_vector <- function(v, a, b) {
   rowSums(mapply(function(a, b) v >= a & v <= b, a = a, b = b)) > 0
 }
 
-#' trim_spectra_targets
+#' .trim_spectra_targets
 #'
 #' @param traces A data.frame with spectra.
 #' @param targets A data.frame with targets (minimum and maximum).
@@ -24,28 +24,35 @@ trim_vector <- function(v, a, b) {
 #'
 #' @noRd
 #'
-trim_spectra_targets <- function(traces, targets, preMZr) {
+.trim_spectra_targets <- function(traces, targets, preMZr, with_im) {
+  
   tg_list <- lapply(seq_len(nrow(targets)),
     function(z, traces, targets, preMZr) {
       tg <- traces
       
-      cutRt <- trim_vector(tg$rt, targets$rtmin[z], targets$rtmax[z])
+      cutRt <- .trim_vector(tg$rt, targets$rtmin[z], targets$rtmax[z])
+      
       tg <- tg[cutRt, ]
       
+      if (with_im) {
+        cutIM <- .trim_vector(tg$drift, targets$driftmin[z], targets$driftmax[z])
+        tg <- tg[cutIM, ]
+      }
+
       if ("polarity" %in% colnames(targets)) {
         tg <- tg[tg$polarity == targets$polarity[z], ]
       }
       
       if (nrow(tg) > 0) {
         if (!is.null(preMZr)) {
-          cutMZ <- trim_vector(tg$mz, targets$mzmin[z], targets$mzmax[z])
+          cutMZ <- .trim_vector(tg$mz, targets$mzmin[z], targets$mzmax[z])
           tg <- tg[tg$level == 2 | (tg$level == 1 & cutMZ), ]
           if (nrow(tg) > 0) {
-            cutPreMZ <- trim_vector(tg$pre_mz, preMZr$mzmin[z], preMZr$mzmax[z])
+            cutPreMZ <- .trim_vector(tg$pre_mz, preMZr$mzmin[z], preMZr$mzmax[z])
             tg <- tg[tg$level == 1 | (tg$level == 2 & cutPreMZ), ]
           }
         } else {
[…]

### 2023-09-26 dev/demo_ozonation_project.Rmd
Exporting short generics and updated articles

@@
 ---
-title: 'StreamFind: Wastewater Ozonation'
-subtitle: Mass spectrometry project example
+title: 'Wastewater ozonation showcase'
+subtitle: |
+  | Mass spectrometry project example
+  | <br>
 author:
@@
   email: cunha@iuta.de
-  affiliation: "Institut für Umwelt & Energie, Technik & Analytik e. V. (IUTA) <br> Bliersheimer Str. 58 - 60, 47229 Duisburg"
+  # affiliation: "Institut für Umwelt & Energie, Technik & Analytik e. V. (IUTA) <br> Bliersheimer Str. 58 - 60, 47229 Duisburg"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
   bookdown::html_document2:
-    toc: yes
-    toc_float: yes
-    theme: paper
+    fig_caption: true
+    toc: true
+    number_sections: true
+    toc_float: true
+vignette: >
+  %\VignetteEncoding{UTF-8}
+  %\VignetteIndexEntry{Developer guide for mass spectrometry}
+  %\VignetteEngine{knitr::rmarkdown}
 editor_options: 
@@
 logo <- file.path(path_extdata, "logo_StreamFind.png")
+# logo <- "man/figures/logo_StreamFind.png"
 ```
@@
 
-<br>
-
 ```{r resources, include=FALSE}
@@
 ```
+
+<style>
+pre {
+  overflow-x: auto;
+}
+pre code {
+  word-wrap: normal;
+  white-space: pre;
+}
+</style>
+
+<br>
+
+<br>
 
@@
 
-The analysis replicate names and the associated blank replicate name can be ammended in the MassSpecData as shown below.
+The analysis replicate names and the associated blank replicate name can be amended in the MassSpecData as shown below.
 
@@
 
-# Instructions
-
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents.
-For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-Diverse cheatsheets can be found in <https://www.rstudio.com/resources/cheatsheets/>.
-This section gives quick instructions to display text, code, tables and figures.
-
-## Text
-
-Text in R markdown has several inline codes for costumization.
-The book in <https://bookdown.org/yihui/rmarkdown/> offers a detailed guideline.
-Moreover, a cheatsheet can be downdloaded with the following link <https://raw.githubusercontent.com/rstudio/cheatsheets/main/rmarkdown.pdf>.
-
-## Code
-
-Code is automaticelly embedded unless `echo` chunk option is set to `FALSE`.
-Evaluation of code in a given chunk can be skiped by setting `eval` to `FALSE`.
-Code chunks should always be unically named.
-
-```{r exampleCode}
-
-#Example of code chunk
-
-list(a = 1:6, b = 1:10)
-
-```
-
-<br>
-
-## Tables
-
-The `knitr::kable` function can be used for displaying tables, see table \@ref(tab:exampleTable).
-Consult <https://haozhu233.github.io/kableExtra/awesome_table_in_html.html> for other costumization options.
-Table caption is added with the argument `caption` of the `kable` function.
-
-```{r exampleTable}
-
-df <- data.frame(name = c("a", "b", "c", "d", "e", "f"), col1 = 1:6, col2 = 1:6)
-
-knitr::kable(df, caption = "Example of table caption.") %>%
-kable_styling(
-  font_size = 12,
-  bootstrap_options = c("striped", "hover", "condensed", "responsive"),
-  fixed_thead = TRUE
-) %>%
-  scroll_box(width = "100%", height = "350px")
-
-```
-
-<br>
-
-## Figures
-
-Figures/plots can also by embedded using diverse packages.
-The packages `ggplot2` and `plotly` (interactive) are recommended as they are versatile and easy to use.
-Cheatsheet for `ggplot2` can be found in <https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-visualization.pdf>.
-For `plotly` the website <https://plotly.com/r/> offers examples and a cheatsheet can be downdloaded in <https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf>.
-Note that the `echo = FALSE` parameter can be added to the code chunk to prevent printing of the R code that generate the plots.
-
-### Example `ggplot2`
-
-```{r examplePlot, fig.cap="Plot caption example"}
-
-ggplot(pressure, aes(temperature, pressure)) +
-geom_point() +
-theme_bw()
-
-```
-
-### Example `plotly`
-
-```{r examplePlotly, fig.cap="Plot caption example"}
-
-plot_ly(pressure, x = ~temperature, y = ~pressure, type = "scatter", mode = "markers+lines")
-
-```
-
-<br> <br>
-
-# Signature
-
-The report was produced on `r Sys.time()`. The data and report are stored in ``r path_wd`` of the workstation *`r Sys.info()["nodename"]`* under the user *`r Sys.info()["user"]`*.
-
 ***
 
-<br>
-
-<style>
-body, p {
-  color: black;
-  font-family: Arial;
-  text-align: justify;
-  font-size: 12pt;
-}
-h1{
-  font-size: 20pt;
-  font-style: bold;
-}
-h2{
-  font-size:18pt;
-  font-style: bold;
-}
-h3{
-  font-size:14pt;
-  font-style: bold;
-}
-.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
-    background-color: #177013;
-}
-caption {
-  color: black;
-  font-size: 1.0em;
-}
-</style>
+<!-- <style> -->
+<!-- body, p { -->
+<!--   color: black; -->
+<!--   font-family: Arial; -->
+<!--   text-align: justify; -->
+<!--   font-size: 12pt; -->
+<!-- } -->
+<!-- h1{ -->
+<!--   font-size: 20pt; -->
+<!--   font-style: bold; -->
+<!-- } -->
+<!-- h2{ -->
+<!--   font-size:18pt; -->
+<!--   font-style: bold; -->
+<!-- } -->
+<!-- h3{ -->
+<!--   font-size:14pt; -->
+<!--   font-style: bold; -->
+<!-- } -->
+<!-- .list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover { -->
+<!--     background-color: #177013; -->
+<!-- } -->
+<!-- caption { -->
+<!--   color: black; -->
+<!--   font-size: 1.0em; -->
+<!-- } -->
+<!-- </style> -->

### 2023-10-11 README.Rmd
Exporting short generics and updated articles
Updated package logo in readme

@@
 
-# StreamFind
+# StreamFind (R package)
 
@@
 <!-- badges: end -->
+
+<p align="center" width="100%">
+  <img width="60%" src="man/figures/logo_StreamFind.png" alt="Logo" />
+</p>
 
@@
 The StreamFind platform is directed to academics but also technicians, due to the aspired comprehensive documentation, well categorized set of integrated modular functions and the graphical user interface.
+<br>
+The library of [StreamFind](https://github.com/odea-project/StreamFind) is an R package (this repository).
+The graphical user interface (GUI) is a Web App developed in JavaScript and can be found in the [StreamFind-App](https://github.com/odea-project/StreamFind-App) repository.
 The StreamFind development is ongoing, please [contact us](mailto:cunha@iuta.de) for questions  or collaboration.
@@
 ```
-
-# StreamFind R package
-
-The back-end framework of [StreamFind](https://github.com/odea-project/StreamFind) is an R package.
 
@@
 
+The StreamFind is open source due public fundsing and to the extensive contribution from scientific literature as well as existing open source software.
+Below we reference to the research and software that is used within StreamFind.
+Please note that each open source software or research that StreamFind depends on uses itself other contributions that is not possible the reference here.
+Therefore, we recommend to search within each citation for other contributions.
+<br>
+
 ---

### 2023-10-12 dev/dev_template_report.Rmd
streamFindData not StreamFindData

@@
 library(plotly)
-library(streamFind)
+library(StreamFind)
 
@@
 # Add logo file
-path_extdata <- system.file(package = "streamFindData", dir = "extdata")
+path_extdata <- system.file(package = "StreamFindData", dir = "extdata")
 logo <- file.path(path_extdata, "logo_iuta.png") #207 x 216 px

### 2023-10-13 dev/demo_ozonation.R
Fixes after adding raman files to StreamFindData

@@ -6,7 +6,7 @@ library(StreamFind)
 
 # Create -----------------------------------------------------------------------
 
-all_files <- StreamFindData::get_all_file_paths()
+all_files <- StreamFindData::get_ms_file_paths()
 
 files <- all_files[grepl("blank|influent|o3sw", all_files)]
 
@@ -113,7 +113,7 @@ ms
 # Access and plot data ---------------------------------------------------------
 
 # Database
-db <- StreamFindData::get_tof_spiked_chemicals()
+db <- StreamFindData::get_ms_tof_spiked_chemicals()
 db <- db[grepl("S", db$tag), ]
 cols <- c("name", "formula", "mass", "rt", "tag")
 db <- db[, cols, with = FALSE]

### 2023-10-19 README.Rmd
Added qCentroids and added DDA, DIA and AllIons tag
The qCentroids still has some issues for tof data.

@@
 
-The StreamFind is open source due public fundsing and to the extensive contribution from scientific literature as well as existing open source software.
-Below we reference to the research and software that is used within StreamFind.
-Please note that each open source software or research that StreamFind depends on uses itself other contributions that is not possible the reference here.
+The StreamFind is open source due to public funding and the extensive contribution from scientific literature as well as existing open source software.
+Below, we reference the research and software that is used within StreamFind.
+Please note that each open source software or research that StreamFind uses relies on other contributions.
 Therefore, we recommend to search within each citation for other contributions.

### 2023-10-19 dev/dev_ms_IM_data.R

@@ -1,6 +1,6 @@
 
-path <- "F:/example_ms_files"
-# path <- "C:/Users/Ricardo Cunha/Documents/Work/example_ms_files"
+# path <- "F:/example_ms_files"
+path <- "C:/Users/Ricardo Cunha/Documents/example_ms_files"
 files <- list.files(path, pattern = "im_tof", full.names = TRUE)
 cols <- c("name", "formula", "mz", "rt")
 db <- paste0(path, "/qc_MS2_pos.csv")
@@ -8,18 +8,22 @@ db <- data.table::fread(db)
 db <- db[, cols, with = FALSE]
 
 # patRoon::clearCache("parsed_ms_analyses")
+View(rcpp_parse_ms_analysis(files[3]))
 
-ms <- MassSpecData$new(files[2])
+patRoon::clearCache("parsed_ms_analyses")
+ms <- MassSpecData$new(files[3])
+ms
 
 # ms$get_run()
 
 # ms$has_ion_mobility()
 
+patRoon::clearCache("parsed_ms_analyses")
 patRoon::clearCache("parsed_ms_spectra")
 
 # ms$get_eic(mz = db$mz, drift = 20, millisec = 5)
 
-ms$get_eic(mz = db$mz, drift = 20, millisec = 5)
+ms$plot_eic(mz = db$mz, drift = 20, millisec = 5)
 
 ms$get_ms1(mz = db$mz, drift = 20, millisec = 5)
 

### 2023-10-19 dev/dev_ms_basic_structure.R

@@ -3,9 +3,10 @@
 
 # TODO add stats to the groups (e.g., presence in each replicate, coverage)
 # TODO check what happens when all MS2 centroids are added to clustering
-# TODO clustering does not see polarity yet and it should be split.
+# TODO Features and groups MS1 could have a is_pre.
 # TODO add possibility to add_files in MassSpecData
 # TODO add is_pre to MS1 spectra
+# TODO add filter for features/groups with more than 1 representation in components
 
 
 # Resources -------------------------------------------------------------------
@@ -18,12 +19,14 @@ cols <- c("name", "formula", "mass", "rt")
 db <- db[, cols, with = FALSE]
 
 
+
+
+
 # Test SIRIUS -----------------------------------------------------------------
 
-ms <- MassSpecData$new(all_files[1:3])
+# ms <- MassSpecData$new(all_files[1:3])
 
-
-
+ms2 <- MassSpecData$new(all_files[1:3])
 
 
 
@@ -53,9 +56,7 @@ files_df <- data.frame(
 
 ms <- MassSpecData$new(files = files_df)
 
-ms$get_run(analyses = 1)
-
-ms <- ms$subset_analyses(c(4:6, 10:12, 16:18))
+# ms <- ms$subset_analyses(c(4:6, 10:12, 16:18))
 
 ms$add_settings(
   list(
@@ -75,10 +76,14 @@ ms$add_settings(
   )
 )
 
-ms
+ms$find_features()
+ms$group_features()
+ms$filter_features()
+
+combine_MassSpecData(ms, ms2)
 
 
-patRoon::clearCache("parsed_ms_analyses")
+# patRoon::clearCache("parsed_ms_analyses")
 patRoon::clearCache("parsed_ms_spectra")
 patRoon::clearCache("load_features_ms1")
 patRoon::clearCache("load_features_ms2")
@@ -87,27 +92,22 @@ patRoon::clearCache("load_groups_ms2")
 # patRoon::clearCache("MSPeakListsAvg")
 # ms$get_history()
 
-ms$find_features()
 
-ms$group_features()
-
-ms$filter_features()
 
 # ms$plot_spectra(1, xVal = "drift")
 
 suspects <- ms$get_suspects(database = db, ppm = 10, sec = 15)
-
-suspects
-
 ms <- ms$subset_features(features = suspects)
 
 ms$load_features_ms1()
-
[…]

### 2023-11-02 R/class_R6_MassSpecData.R
Updated pkgdown for pull request

@@ -8813,11 +8813,8 @@ MassSpecData <- R6::R6Class("MassSpecData",
 
     ### ___ help -----
 
-    #' @description
-    #' Shortcut to reference help pages.
-    #'
-    #' @return Opens a reference help page.
-    #'
+    #' @field help (`list()`)\cr
+    #' List of function elements to access specific reference help pages.
     help = list(
       methods = function() {
         browseURL("https://odea-project.github.io/StreamFind/reference/MassSpecData.html#methods")

### 2024-01-05 dev/TiChri_integration_patRoon.Rmd
Improved as_fGroups

---
title: "TiChri: Matrix correction integration in patRoon"
subtitle: "Concept"
author:
  name: Ricardo Cunha
  email: cunha@iuta.de
  affiliation: "Institut für Umwelt & Energie, Technik & Analytik e. V. (IUTA) <br> Bliersheimer Str. 58 - 60, 47229 Duisburg"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    theme: paper
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# Libraries
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(plotly)
library(StreamFind)
library(patRoon)
# Global options
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 9, results = "asis", cache = TRUE)
options(width = 300)
```

```{r logo, echo=FALSE}
path_extdata <- system.file(package = "StreamFindData", dir = "extdata")
logo <- file.path(path_extdata, "logo_StreamFind.png")
htmltools::img(src = knitr::image_uri(logo), style = "margin-left:inherit;position:absolute;top:80px;left:740px;width:200px;")
```

<br>

***

```{r resources, include=FALSE}
files <- list.files("D:/NTS/Project_220614_AirLiquid_220509_SPE/msfiles", full.names = TRUE)
selected_files <- files[grepl("pos_Blank|pos_M220509014|pos_M220509017|pos_M220509018", files)]

all_db <- StreamFindData::get_ms_tof_spiked_chemicals_with_ms2()
db <- all_db[!grepl("IS", all_db$tag, fixed = TRUE), ]
cols <- c("name", "formula", "mass", "SMILES", "rt", "polarity", "fragments")
# cols <- c("name", "formula", "mass", "SMILES", "rt")
db <- db[, cols, with = FALSE]
# data.table::setnames(db, "mass", "neutralMass")

# db$neutralMass <- db$neutralMass + 1.0073
# data.table::setnames(db, "neutralMass", "mz")

dbis <- all_db[grepl("IS", all_db$tag), ]
cols <- c("name", "formula", "mass", "rt")
dbis <- dbis[, cols, with = FALSE]
# files <- list.files("E:/Pub_Aopti/ms_files_mzML", "M200422", full.names = TRUE)
```

# AnalysisInfo

```{r anaInfo}
# The 3 first files are control and were excluded
anaInfo <- data.frame(
  "path" = dirname(selected_files),
  "analysis" = tools::file_path_sans_ext(basename(selected_files)),
  "group" = c(
    rep("blank", 2),
    rep("effluent_OZ", 2),
    rep("effluent_OSW", 2),
    rep("influent", 2)
  ),
  "blank" = rep("blank", 8)
)

kable(anaInfo, caption = "AnalysisInfo") %>%
  kable_styling(
    font_size = 12,
    bootstrap_options = c("striped", "hover", "condensed"),
    fixed_thead = TRUE
  )
```

<br>

# Total ion Chromatograms (TICs)

```{r tics, message=FALSE, fig.height=9, fig.cap="Total ion chromatogram (TIC) of all analyses by replicate group name."}
ms <- MassSpecData$new(selected_files)
ms$add_replicate_names(anaInfo$group)
ms$add_blank_names(anaInfo$blank)
ms$plot_tic(levels = 1, colorBy = "replicates")
```

<br>

# Data processing

Processing settings described in \@ref(processing-settings).

```{r data-process, message=FALSE, results='markup'}
ms$find_features(Settings_find_features_openms())
ms$group_features(Settings_group_features_openms())
ms$find_internal_standards(Settings_find_internal_standards_StreamFind(database = dbis, ppm = 8, sec = 10))
ms
```

```{r internal-standards, fig.height=9, fig.cap="Evaluaton of spiked internal standards after feature finding and grouping."}
#ms$plot_internal_standards_qc()
```

```{r to-patroon, message=FALSE, results='markup'}
fGroups <- ms$as_patRoon_featureGroups()
fGroups
```

<br>

# Implementation in patRoon

## Get TICs

```{r get-tics-func, warning=FALSE, message=FALSE}
getTICs <- function(anaInfo) {
  filePaths <- patRoon:::getMzMLOrMzXMLAnalysisPath(anaInfo$analysis, anaInfo$path, mustExist = TRUE)
  # can be applied in parallel?
  tics <- lapply(filePaths, function(fpath) {
    msf <- mzR::openMSfile(fpath)
    hd <- mzR::header(msf)
    mzR::close(msf)
    hd <- hd[hd$msLevel == 1, ]
    data.table::data.table("rt" = hd$retentionTime, "int" = hd$totIonCurrent)
  })
  names(tics) <- anaInfo$analysis
  tics
}
```

```{r get-tics, warning=FALSE, message=FALSE, results='markup'}
tics <- getTICs(anaInfo)

length(tics)

is(tics)

head(tics[[1]])
```

## Calculate TIC matrix profiles (MPs)

The function uses a retention window (`rtWindow`), in seconds, to estimate the blank level in the proximity of the analysis spectrum. For now, simple average is used to aggregate the TIC MPs calculated for each blank replicate group associated with a given analysis and then to summarize the analysis MPs into one matrix profile (MP) vector. The MP is calculated according to eq. 6 (see below) of 10.1021/acs.analchem.1c00357.

$TIC_{MP} = (\frac{TIC^{matrix}}{TIC^{blank}}) * -1$

```{r calc-mps-func, warning=FALSE, message=FALSE}
calculateTicMps <- function(anaInfo, tics, rtWindow = 10) {

  .trimVector <- function(v, a, b) {
    rowSums(mapply(function(a, b) v >= a & v <= b, a = a, b = b)) > 0
  }

  blkGroups <- anaInfo$blank
  names(blkGroups) <- anaInfo$analysis

  blkAnalyses <- anaInfo$analysis
  names(blkAnalyses) <- anaInfo$blank
  blkAnalyses <- blkAnalyses[anaInfo$group %in% anaInfo$blank]

  # can be applied in parallel?
  anaMpList <- lapply(anaInfo$analysis, function(x, blkGroups, blkAnalyses, tics, rtWindow) {

    blkG <- blkGroups[x]
    blkA <- blkAnalyses[names(blkAnalyses) %in% blkG]

    ticX <- tics[[x]]

    mpListX <- lapply(tics[blkA], function(z, ticX, rtWindow) {

      mp <- vapply(seq_len(nrow(ticX)), function(j, z) {
        rt <- ticX$rt[j]
        rtr <- c(rt - rtWindow, rt + rtWindow)
        intB <- mean(z$int[.trimVector(z$rt, rtr[1], rtr[2])])
        # eq. 6 from 10.1021/acs.analchem.1c00357
        (ticX$int[j] / intB) * -1
      }, z = z, 0)

      mp

    }, ticX = ticX, rtWindow = rtWindow)

    ticX$mp <- Reduce(`+`, mpListX) / length(mpListX)

    ticX

  }, blkGroups = blkGroups, blkAnalyses = blkAnalyses, tics = tics, rtWindow = rtWindow)

  names(anaMpList) <- anaInfo$analysis

  anaMpList
}
```

```{r calc-mps, warning=FALSE, message=FALSE, fig.height=9, fig.cap="MPs of each analysis"}
ticMps <- calculateTicMps(anaInfo, tics, 10)

fig <- plot_ly()
cl <- StreamFind:::.get_colors(anaInfo$group)
sL <- !duplicated(cl)
for(i in seq_len(nrow(anaInfo))) {
  fig <- fig %>% add_trace(
    x = ticMps[[i]]$rt,
    y = ticMps[[i]]$mp,
    name = names(cl)[i],
    legendgroup = names(cl)[i],
    showlegend = sL[i],
    type = "scatter",
    mode = 'lines',
    line = list(color = cl[i])
  )
}

# fig <- fig %>% add_trace(
#     x = iseval$rt,
#     y = iseval$rec,
#     name = names(cl)[i],
#     legendgroup = names(cl)[i],
#     showlegend = sL[i],
#     type = "scatter",
#     mode = 'markers',
#     marker = list(color = cl[i])
#   )

xaxis <- list(
  linecolor = toRGB("black"),
  linewidth = 2, title = "Retention time / seconds",
  titlefont = list(size = 12, color = "black")
)

yaxis <- list(
  linecolor = toRGB("black"),
  linewidth = 2, title = "Suppression factor",
  titlefont = list(size = 12, color = "black")
)

fig <- fig %>% plotly::layout(xaxis = xaxis, yaxis = yaxis)

fig
```

```{r}

iseval <- ms$get_internal_standards(average = T)

plot(iseval$rt, iseval$rec)

```

s
```{r}
purine <- 121.05090000 
HP921 <- 922.00980000

# ms$plot_eic(analyses = c(2, 5, 14), mass = dbis$mass[4], ppm = 20, colorBy = "targets+analyses")

ms$plot_eic(analyses = c(2, 5, 14), mass = c(purine, HP921), ppm = 20, colorBy = "replicates+targets")

# Check a project with SPE applied
```

```{r}

```

<br>

# Processing settings

```{r pro-settings, warning=FALSE, message=FALSE, results='markup'}
Settings_find_features_openms()
Settings_group_features_openms()
```

### 2024-01-08 dev/TiChri_integration_patRoon.Rmd
Updated get_groups and TiChri dev into patRoon

@@
 title: "TiChri: Matrix correction integration in patRoon"
-subtitle: "Concept"
 author:
@@
 # Libraries
+library(data.table)
 library(knitr)
@@
 knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 9, results = "asis", cache = TRUE)
-options(width = 300)
+options(width = 400)
 ```
+
+<style>
+.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
+  background-color: #1E8449;
+}
+</style>
 
@@
 logo <- file.path(path_extdata, "logo_StreamFind.png")
-htmltools::img(src = knitr::image_uri(logo), style = "margin-left:inherit;position:absolute;top:80px;left:740px;width:200px;")
+htmltools::img(src = knitr::image_uri(logo), style = "margin-left:inherit;position:absolute;top:100px;left:740px;width:200px;")
 ```
@@
 db <- all_db[!grepl("IS", all_db$tag, fixed = TRUE), ]
-cols <- c("name", "formula", "mass", "SMILES", "rt", "polarity", "fragments")
-# cols <- c("name", "formula", "mass", "SMILES", "rt")
+cols <- c("name", "formula", "mass", "SMILES", "rt")
 db <- db[, cols, with = FALSE]
-# data.table::setnames(db, "mass", "neutralMass")
-
-# db$neutralMass <- db$neutralMass + 1.0073
-# data.table::setnames(db, "neutralMass", "mz")
 
@@
 dbis <- dbis[, cols, with = FALSE]
-# files <- list.files("E:/Pub_Aopti/ms_files_mzML", "M200422", full.names = TRUE)
 ```
@@
 ```{r anaInfo}
-# The 3 first files are control and were excluded
 anaInfo <- data.frame(
@@
 
-Processing settings described in \@ref(processing-settings).
+```{r data-process, message=FALSE, results='markup'}
+ms$find_features(
+  Settings_find_features_openms(
+    noiseThrInt = 1000,
+    chromSNR = 3,
+    chromFWHM = 7,
+    mzPPM = 15,
+    reEstimateMTSD = TRUE,
+    traceTermCriterion = "sample_rate",
+    traceTermOutliers = 5,
+    minSampleRate = 1,
+    minTraceLength = 4,
+    maxTraceLength = 70,
+    widthFiltering = "fixed",
+    minFWHM = 4,
+    maxFWHM = 35,
+    traceSNRFiltering = TRUE,
+    localRTRange = 0,
+    localMZRange = 0,
+    isotopeFilteringModel = "none",
+    MZScoring13C = FALSE,
+    useSmoothedInts = FALSE,
+    extraOpts = NULL,
+    intSearchRTWindow = 3,
+    useFFMIntensities = FALSE,
+    verbose = FALSE
+  )
+)
 
-```{r data-process, message=FALSE, results='markup'}
-ms$find_features(Settings_find_features_openms())
-ms$group_features(Settings_group_features_openms())
-ms$find_internal_standards(Settings_find_internal_standards_StreamFind(database = dbis, ppm = 8, sec = 10))
+ms$group_features(
+  Settings_group_features_openms(
+    rtalign = FALSE,
+    QT = FALSE,
+    maxAlignRT = 8,
+    maxAlignMZ = 0.01,
+    maxGroupRT = 8,
+    maxGroupMZ = 0.01,
+    extraOptsRT = NULL,
+    extraOptsGroup = NULL,
+    verbose = FALSE
+  )
+)
+
+ms$find_internal_standards(
+  Settings_find_internal_standards_StreamFind(
+    database = dbis,
+    ppm = 8,
+    sec = 10
+  )
+)
+
 ms
@@
 ```{r internal-standards, fig.height=9, fig.cap="Evaluaton of spiked internal standards after feature finding and grouping."}
-#ms$plot_internal_standards_qc()
+ms$plot_internal_standards_qc()
+```
+
+```{r internal-standards-eic, fig.height=9, fig.cap="Extracted ion chromatogram (EIC) of four spiked internal standards per analysis replicate group.", message=FALSE}
+ms$plot_eic(mass = dbis[c(3, 4, 7, 8), c("name", "mass"), with = FALSE], colorBy = "replicates+targets", legendNames = TRUE, ppm = 20)
 ```
@@
 
+## Conditions
+
+- Analyses must have associated blanks
+- Features must be group across analyses for correspondence
+
 ## Get TICs
@@
   filePaths <- patRoon:::getMzMLOrMzXMLAnalysisPath(anaInfo$analysis, anaInfo$path, mustExist = TRUE)
-  # can be applied in parallel?
+  # can be applied in parallel but probably not advantageous
   tics <- lapply(filePaths, function(fpath) {
@@
 
-  # can be applied in parallel?
+  # can be applied in parallel but probably not advantageous
   anaMpList <- lapply(anaInfo$analysis, function(x, blkGroups, blkAnalyses, tics, rtWindow) {
@@
 
-```{r calc-mps, warning=FALSE, message=FALSE, fig.height=9, fig.cap="MPs of each analysis"}
-ticMps <- calculateTicMps(anaInfo, tics, 10)
+```{r calc-mps, warning=FALSE, message=FALSE, fig.height=9, fig.cap="TIC MPs of each analysis"}
+ticMps <- calculateTicMps(anaInfo, tics, 5)
 
@@
 sL <- !duplicated(cl)
+
 for(i in seq_len(nrow(anaInfo))) {
@@
 }
-
-# fig <- fig %>% add_trace(
-#     x = iseval$rt,
-#     y = iseval$rec,
-#     name = names(cl)[i],
-#     legendgroup = names(cl)[i],
-#     showlegend = sL[i],
-#     type = "scatter",
-#     mode = 'markers',
-#     marker = list(color = cl[i])
-#   )
 
@@
 
-```{r}
+## Correct intensity of featureGroups
 
-iseval <- ms$get_internal_standards(average = T)
+```{r calc-int-corr, warning=FALSE, message=FALSE}
+calculateCorrectedIntensities <- function(dt, anaInfo, ticMps, rtWindow) {
 
-plot(iseval$rt, iseval$rec)
+  blkGroups <- anaInfo$blank
+  names(blkGroups) <- anaInfo$analysis
 
+  blkAnalyses <- anaInfo$analysis
+  names(blkAnalyses) <- anaInfo$blank
+  blkAnalyses <- blkAnalyses[anaInfo$group %in% anaInfo$blank]
+
+  out <- data.table(
+    "feature" = dt$feature,
+    "group" = dt$group,
+    "analysis" = dt$analysis,
+    "rt" = dt$rt,
+    "intensity" = dt$intensity,
+    "scale" = NA_real_,
+    "mp" = NA_real_,
+    "corrected" = NA_real_
+  )
+
+  for (i in seq_len(nrow(out))) {
+    mp <- ticMps[[out$analysis[i]]]
+
+    if (is.null(rtWindow)) rtWindow <- (out$rtmax[i] - out$rtmin[i]) / 4
+
+    sel <- (mp$rt >= (out$rt[i] - rtWindow)) & (mp$rt <= (out$rt[i] + rtWindow))
+
+    mp <- mp[sel, ]
+
+    bInt <- 0
+
+    for (j in seq_len(length(blkAnalyses))) {
+
+      bMp <- ticMps[[blkAnalyses[j]]]
+
+      sel <- (bMp$rt >= (out$rt[i] - rtWindow)) & (bMp$rt <= (out$rt[i] + rtWindow))
+
+      bMp <- bMp[sel, ]
+
+      bInt <- bInt + mean(bMp$int)
+    }
+
+    bInt <- bInt / length(blkAnalyses)
+
+    selBlkInts <- (out$analysis %in% blkAnalyses) & (out$group %in% out$group[i])
+
+    if (!any(selBlkInts)) {
+      warning("Feature", out$feature[i], " from analysis ", out$analysis[i], " does not have blank representation!")
+      next
+    }
+
+    #based on eq. 4 from 10.1021/acs.analchem.1c00357
+    out$scale[i] <- ((mean(out$intensity[selBlkInts]) / out$intensity[i]) - 1) * -1
+
+    #based on eq. 7 from 10.1021/acs.analchem.1c00357
+    out$mp[i] <- ((mean(mp$int) / bInt) * -1) * out$scale[i] / mean(mp$mp)
+
+    #based on eq. 8 from 10.1021/acs.analchem.1c00357
+    out$corrected[i] <- out$intensity[i] * ((1 - out$mp[i] + 1) - 1)
+  }
+
+  out
+}
 ```
 
-s
-```{r}
-purine <- 121.05090000 
-HP921 <- 922.00980000
+```{r calc-ints, warning=FALSE, message=FALSE, fig.height=9, fig.cap="Corrected instensity of internal standards (dashed lines are corrected intensities)."}
+# testing with the spiked internal standards
+istd <- ms$get_internal_standards(average = FALSE)
+istd <- istd[, c("group", "name"), with = TRUE]
 
-# ms$plot_eic(analyses = c(2, 5, 14), mass = dbis$mass[4], ppm = 20, colorBy = "targets+analyses")
+dt <- ms$get_features(features = istd)
 
-ms$plot_eic(analyses = c(2, 5, 14), mass = c(purine, HP921), ppm = 20, colorBy = "replicates+targets")
+corrInts <- calculateCorrectedIntensities(dt, anaInfo, ticMps, rtWindow = 4)
 
-# Check a project with SPE applied
-```
+g_names <- dt$name
+names(g_names) <- dt$group
 
-```{r}
+corrInts$name <- g_names[corrInts$group]
 
+fig <- plot_ly(corrInts, x = corrInts$analysis)
+grps <- unique(corrInts$name)
+colors <- StreamFind:::.get_colors(grps)
+
+for (g in grps) {
+  df <- corrInts[corrInts$name == g, ]
+
+  fig <- fig %>% add_trace(df,
+    x = df$analysis,
+    y = df$intensity,
+    type = "scatter", mode = "lines",
+    line = list(width = 1, color = colors[g]),
+    connectgaps = FALSE,
+    name = g,
+    legendgroup = g,
+    showlegend = TRUE
+  )
+
+  fig <- fig %>% add_trace(df,
+    x = df$analysis,
+    y = df$corrected,
+    type = "scatter", mode = "lines",
+    line = list(width = 1, color = colors[g], dash = 'dash'),
+    connectgaps = FALSE,
+    name = g,
+    legendgroup = g,
+    showlegend = FALSE
+  )
+}
+
+xaxis <- list(
+  linecolor = toRGB("black"),
+  linewidth = 2, title = "Analysis",
+  titlefont = list(size = 12, color = "black")
+)
+
+yaxis <- list(
+  linecolor = toRGB("black"),
+  linewidth = 2, title = "Intensity / counts",
+  titlefont = list(size = 12, color = "black")
+)
+
+fig <- fig %>% plotly::layout(xaxis = xaxis, yaxis = yaxis)
+
+fig
 ```
@@
 <br>
-
-# Processing settings
-
-```{r pro-settings, warning=FALSE, message=FALSE, results='markup'}
-Settings_find_features_openms()
-Settings_group_features_openms()
-```

### 2024-01-08 R/class_R6_MassSpecData.R

@@ -2665,33 +2665,25 @@ MassSpecData <- R6::R6Class("MassSpecData",
         fgroups <- data.table("group" = g_ids)
         
         if (intensities) {
-          fts_av <- fts[, .(intensity = max(intensity)), by = c("group", "analysis")]
           
           if (average) {
             rpls <- self$get_replicate_names()
-            fts_av$analysis <- rpls[fts_av$analysis]
-            fts_av <- fts_av[, .(intensity = mean(intensity, na.rm = TRUE), iSD = sd(intensity, na.rm = TRUE)), by = c("group", "analysis")]
-            fts_av$iSD <- round(fts_av$iSD / fts_av$intensity * 100, digits = 0)
+            fts_temp <- copy(fts)
+            fts_temp$analysis <- rpls[fts_temp$analysis]
+            fts_av <- fts_temp[, .(intensity = mean(intensity), sd = sd(intensity)), by = c("group", "analysis")]
+            fts_av$sd[is.na(fts_av$sd)] <- 0 
+            fts_av$sd <- round(fts_av$sd / fts_av$intensity * 100, digits = 0)
+            fts_sd <- copy(fts_av)
+            fts_sd$intensity <- NULL
+            fts_sd$analysis <- paste0(fts_sd$analysis, "_sd")
+            fts_sd <- tidyr::spread(fts_av[, c("group", "analysis", "sd"), with = TRUE], key = analysis, value = sd, fill = NA)
+            fts_av$sd <- NULL
+            fts_av <- tidyr::spread(fts_av, key = analysis, value = intensity, fill = 0)
+            
+          } else {
+            fts_av <- fts[, .(intensity = max(intensity)), by = c("group", "analysis")]
+            fts_av <- tidyr::spread(fts_av, key = analysis, value = intensity, fill = 0)
           }
-          
-          mat_cols <- self$get_analysis_names()
-          
-          if (average) mat_cols <- unique(rpls)
-          
-          mat <- matrix(rep(0, n_g * length(mat_cols)), nrow = n_g)
-          rownames(mat) <- g_ids
-          colnames(mat) <- mat_cols
-          
-          if (average && sdValues) {
-            mat_sd <- mat
-            for (i in seq_len(nrow(fts_av))) mat_sd[fts_av$group[i], fts_av$analysis[i]] <- fts_av$iSD[i]
-            colnames(mat_sd) <- paste0(mat_cols, "_sd")
-            iSD <- cbind(fgroups, mat_sd)
-          }
-          
[…]

### 2024-01-17 dev/TiChri_integration_patRoon.Rmd
Added entries to gitignore

@@
 dbis <- dbis[, cols, with = FALSE]
+dbis <- dbis[c(3, 4, 7, 8), ]
 ```
@@
 
-```{r tics, message=FALSE, fig.height=9, fig.cap="Total ion chromatogram (TIC) of all analyses by replicate group name."}
+```{r tics, message=FALSE, fig.cap="Total ion chromatogram (TIC) of all analyses by replicate group name."}
 ms <- MassSpecData$new(selected_files)
@@
     noiseThrInt = 1000,
-    chromSNR = 3,
-    chromFWHM = 7,
+    chromSNR = 5,
+    chromFWHM = 5,
     mzPPM = 15,
@@
   Settings_group_features_openms(
-    rtalign = FALSE,
+    rtalign = TRUE,
     QT = FALSE,
-    maxAlignRT = 8,
-    maxAlignMZ = 0.01,
+    maxAlignRT = 10,
+    maxAlignMZ = 0.02,
     maxGroupRT = 8,
@@
 
+```{r internal-standards-eic, fig.height=7, fig.cap="Spiked internal standards after feature finding and grouping.", message=FALSE}
+ms$plot_groups(ms$get_internal_standards()[,c("group", "name"), with = FALSE], legendNames = TRUE, colorBy = "targets+replicates")
+```
+
 ```{r internal-standards, fig.height=9, fig.cap="Evaluaton of spiked internal standards after feature finding and grouping."}
@@
 
-```{r internal-standards-eic, fig.height=9, fig.cap="Extracted ion chromatogram (EIC) of four spiked internal standards per analysis replicate group.", message=FALSE}
-ms$plot_eic(mass = dbis[c(3, 4, 7, 8), c("name", "mass"), with = FALSE], colorBy = "replicates+targets", legendNames = TRUE, ppm = 20)
+```{r internal-standards-eic-old, include=FALSE, fig.height=7, fig.cap="Extracted ion chromatogram (EIC) of four spiked internal standards per analysis replicate group.", message=FALSE}
+# ms$get_internal_standards(average = FALSE)
+# ms$plot_eic(mass = dbis[, c("name", "mass"), with = FALSE], colorBy = "replicates+targets", legendNames = TRUE, ppm = 20)
+# ms$plot_eic(mass = dbis[c(3, 4, 7, 8), c("name", "mass"), with = FALSE], colorBy = "replicates+targets", legendNames = TRUE, ppm = 20)
 ```
@@
       mp <- vapply(seq_len(nrow(ticX)), function(j, z) {
+
         rt <- ticX$rt[j]
+
         rtr <- c(rt - rtWindow, rt + rtWindow)
+
         intB <- mean(z$int[.trimVector(z$rt, rtr[1], rtr[2])])
+
+        intX <- mean(ticX$int[.trimVector(ticX$rt, rtr[1], rtr[2])])
+
         # eq. 6 from 10.1021/acs.analchem.1c00357
-        (ticX$int[j] / intB) * -1
+        (intX / intB) * -1
+
+        # alternative using the intensity at the exact rt
+        # (ticX$int[j] / intB) * -1
+
       }, z = z, 0)
@@
     "intensity" = dt$intensity,
+    "ticmp" = NA_real_,
     "scale" = NA_real_,
@@
   for (i in seq_len(nrow(out))) {
+
     mp <- ticMps[[out$analysis[i]]]
@@
     if (!any(selBlkInts)) {
+
+      # What to do?
+      # Find another feature in the near retention time?
+      # Pre-calculate retention time and mass based scales?
+
       warning("Feature", out$feature[i], " from analysis ", out$analysis[i], " does not have blank representation!")
+
       next
     }
+
+    out$ticmp[i] <- mean(mp$mp)
 
@@
     out$corrected[i] <- out$intensity[i] * ((1 - out$mp[i] + 1) - 1)
+
   }
@@
 
+rpls <- ms$get_replicate_names()
+
 corrInts <- calculateCorrectedIntensities(dt, anaInfo, ticMps, rtWindow = 4)
+corrInts$replicate <- rpls[corrInts$analysis]
 
@@
 
+```{r, include=FALSE}
+
+dt2 <- ms$get_features(mass = db, sec = 15, ppm = 10)
+ms$plot_groups_overview(groups = dt2, legendNames = TRUE)
+
+ms$plot_eic(mass = db, legendNames = TRUE)
+
+corrInts2 <- calculateCorrectedIntensities(dt2, anaInfo, ticMps, rtWindow = 4)
+
+library(ggplot2)
+
+ggplot(corrInts, aes(x=rt, y=scale, shape=replicate, color=group)) +
+  geom_point()
+
+ggplot(corrInts, aes(x=rt, y=ticmp, shape=replicate, color=group)) +
+  geom_point()
+
+ggplot(corrInts, aes(x=mp, y=ticmp, shape=replicate, color=group)) +
+  geom_point()
+
+fig <- plot_ly(corrInts, x = corrInts$rt)
+
+fig <- fig %>% add_trace(corrInts,
+  x = corrInts$rt,
+  y = corrInts$scale,
+  color = corrInts$replicate,
+  name = corrInts$replicate,
+  legendgroup = corrInts$replicate,
+  type = "scatter", mode = "markers",
+  marker = list(size = 10)
+)
+
+fig2 <- plot_ly(corrInts, x = corrInts$rt)
+
+fig2 <- fig2 %>% add_trace(corrInts,
+  x = corrInts$rt,
+  y = corrInts$ticmp,
+  color = corrInts$replicate,
+  name = corrInts$replicate,
+  legendgroup = corrInts$replicate,
+  type = "scatter", mode = "markers",
+  marker = list(size = 10),
+  showlegend = FALSE
+)
+
+subplot(
+  list(fig, fig2),
+  nrows = 2,
+  titleY = TRUE, titleX = TRUE,
+  margin = 0.01,
+  shareX = TRUE,
+  which_layout = "merge"
+)
+
+```
+
 <br>

### 2024-01-30 dev/dev_Core.R
Start implementation of core engines

@@ -1,0 +1,88 @@
+
+## Core development
+
+# R6 class: CoreEngine
+
+core <- CoreEngine$new()
+core
+
+# S3 classes: ProjectHeaders, ProcessingSettings, Analysis
+
+phead <- ProjectHeaders()
+phead
+
+settings <- ProcessingSettings()
+settings
+
+ana <- Analysis()
+ana
+
+xy <- XyEngine$new()
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+patRoon::clearCache("all")
+msana <- parse.MassSpecAnalysis(StreamFindData::get_ms_file_paths()[1])[[1]]
+class(msana)
+sloop::s3_dispatch(print(msana))
+ana
+
+
+core <- CoreEngine$new()
+core$add_settings(Settings_annotate_features_StreamFind())
+core$add_analyses(Analysis())
+core$get_analyses()
+
+
+
+
+
+
+
+
+
+
+core$get_history()
+
+
+uv <- UVEngine$new()
+uv
+
+r1 <- RamanEngine$new(files = StreamFindData::get_raman_file_paths())
+
+r1
+
+r1$add_replicate_names(c(rep("Sample", 11), rep("Blank", 11)))
+
+r1$add_blank_names(rep("Blank", 22))
+
+r1$get_spectra(analyses = 1)
+
+r1$plot_spectra(colorBy = "replicates", interactive = TRUE)
+
+# plot(r1$get_spectra(analyses = 1)[, 2:3], type = "l")
+# 
+# plot(r1$get_spectra(2)[1:200, 2:3])
+# 
+# View(r1$get_analyses())

### 2024-01-30 dev/dev_raman.R

@@ -1,7 +1,7 @@
 
 files <- StreamFindData::get_raman_file_paths()
 
-r1 <- RamanData$new(files)
+r1 <- Raman$new(files)
 
 r1
 
@@ -11,7 +11,7 @@ r1$add_blank_names(rep("Blank", 22))
 
 r1$get_spectra(analyses = 1)
 
-r1$plot_spectra()
+r1$plot_spectra(colorBy = "replicates", interactive = TRUE)
 
 # plot(r1$get_spectra(analyses = 1)[, 2:3], type = "l")
 # 

### 2024-02-07 R/class_R6_MassSpecData.R
Further implementation of core
Core implementation will be moved to a new branch after this commit from Linux

@@ -54,7 +54,7 @@
 #' @template arg-ms-settingsFeatures
 #' @template arg-ms-legendNames
 #' @template arg-ms-colorBy
-#' @template arg-ms-title
+#' @template arg-title
 #' @template arg-ms-labs
 #' @template arg-ms-interactive
 #' @template arg-ms-save-format
@@ -737,7 +737,7 @@ MassSpecData <- R6::R6Class("MassSpecData",
       if (!is.null(settings)) suppressMessages(self$add_settings(settings))
 
       if (is.null(analyses) & !is.null(files)) {
-        analyses <- parse.MassSpecAnalysis(files, runParallel)
+        analyses <- parse_MassSpecAnalysis(files, runParallel)
         if (is.null(analyses)) {
           warning("No valid files were given! MassSpecData object is empty. \n")
         }
@@ -3995,7 +3995,7 @@ MassSpecData <- R6::R6Class("MassSpecData",
 
       if (!is.null(files)) {
 
-        new_analyses <- parse.MassSpecAnalysis(files, runParallel)
+        new_analyses <- parse_MassSpecAnalysis(files, runParallel)
 
         if (all(vapply(new_analyses, function(x) "MassSpecAnalysis" %in% is(x), FALSE))) {
           self$add_analyses(new_analyses)

### 2024-03-08 dev/dev_ideas/ms_overview.Rmd
Added basic workflow for RamanEngine

---
title: "StreamFind for mass spectrometry"
subtitle: |
  | Overview of the methods in the R6 MassSpecData class
  | <br>
author:
  name: Ricardo Cunha
  email: cunha@iuta.de
  # affiliation: "Institut für Umwelt & Energie, Technik & Analytik e. V. (IUTA) <br> Bliersheimer Str. 58 - 60, 47229 Duisburg"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    fig_caption: true
    toc: true
    number_sections: true
    toc_float: true
    toc_depth: 4
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{StreamFind for mass spectrometry}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

```{r, include = FALSE}
# Libraries
library(knitr)
library(kableExtra)
library(magrittr)
library(mzR)
library(patRoon)
library(StreamFind)

# Variables
logo <- NA_character_

# Working directory
path_wd <- getwd()

# Path StreamFindData
path_extdata <- system.file(package = "StreamFindData", dir = "extdata")

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  fig.align = "center",
  fig.width = 8,
  results = 'markup',
  cache = TRUE,
  size = 'small'
)

options(width = 110)

logo <- file.path(path_extdata, "logo_StreamFind.png")
# logo <- "man/figures/logo_StreamFind.png"
```

```{r logo, echo=FALSE}
# if (!is.na(logo)) {
#   htmltools::img(
#   src = knitr::image_uri(logo),
#   alt = "logo",
#   style =
#     paste0("
#       margin-left:inherit;
#       position:absolute;
#       top:150px;
#       left:600px;
#       width:", 250, "px;
#       height:", 250*(1633/3070), "px;"
#     )
#   )
# }
```

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<br>

<br>

***

# Introduction

The StreamFind R package is a framework for data processing in environmental and quality studies. This article introduces the basic framework of the StreamFind R package for mass spectrometric (MS) data processing. Installation instructions of StreamFind R package can be found [here](https://github.com/odea-project/StreamFind). Once installed, the StreamFind R package is loaded as shown below.

```{r setup, eval=FALSE}
library(StreamFind)
```

# Resources

The StreamFind includes example data files via the supplementary  [StreamFindData](https://github.com/odea-project/StreamFindData) R package.

```{r install-StreamFindData, eval=FALSE}
# install the StreamFindData R package from the GitHub repository
remotes::install_github("odea-project/StreamFindData")
```

## MS files

Once [StreamFindData](https://github.com/odea-project/StreamFindData) is installed, the full file paths of example MS data files can be obtained as shown below. The files available are described in the table below. Note, the MS files (.mzML and .mzXML) were trimmed with the function `trim_ms_files_spectra()` from StreamFind to reduce their size. The centroided Agilent Q-TOF files were trimmed with retention time between 900 and 1350 seconds, *m/z* for MS1 data between 200 and 450 Da and *m/z* for MS2 between 35 and 450 Da. Additionally, MS1 and MS2 traces below 100 and 50 counts were removed, respectively. The profile Agilent Q-TOF files were trimmed with retention time between 1000 and 1200 seconds, *m/z* for MS1 data between 200 and 300 Da and *m/z* for MS2 between 35 and 300 Da. Additionally, MS1 and MS2 traces below 1 count were removed to exclude empty spectra. The files are used within the document for demonstration of the framework.

```{r resource-files, cache=FALSE}
files <- StreamFindData::get_ms_file_paths()
# files is a vector with all file full paths
```

<br>

```{r description-files, echo=FALSE}
df_files_desc <- StreamFindData::get_ms_file_descriptions()

knitr::kable(df_files_desc, caption = "Files included in the StreamFind R package.") %>%
  column_spec(column = 3, width = "100px") %>%
  column_spec(column = 4, width = "60px") %>%
  kable_styling(font_size = 11.5, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>%
  scroll_box(width = "100%", height = "600px")
```

<br>

## Spiked chemicals

The chemicals spiked in the MS files are described below in the table. The internal standards (IS) are spiked to all Agilent Q-TOF files. The chemical standards (S) were only spiked for the first 6 Agilent Q-TOF files. Note that not all IS and S are visible in the profile Agilent Q-TOF files as a narrower trimming was applied.

```{r load-table-chemicals}
db <- StreamFindData::get_ms_tof_spiked_chemicals()
```

<br>

```{r show-table-chemicals, echo=FALSE}
cols_table_01 <- c("name", "CAS", "formula", "mass", "rt", "ionization", "tag", "in_file")
knitr::kable(db[, cols_table_01, with = FALSE], caption = "Chemical standards spiked.") %>%
  kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>%
  scroll_box(width = "100%", height = "600px")
```

<br>

# MassSpecData

The MassSpecData is an [R6 class](https://r6.r-lib.org/articles/Introduction.html) object, holding the MS data parsed from *.mzML* or *.mzXML* files as well as the methods to get, visualize and process the parsed data (see `?MassSpecData` for detailed documentation). The most straight forward manner to initiate a MassSpecData object is with a character vector of *.mzML* or *.mzXML* file full path/s, as shown below for three MS files (triplicate, 1 to 3 in files table in section [MS files](#ms-files)). The object name was defined to *Example 1* using the argument *headers*. For more information on the `MassSpecData$new()` method, see `?MassSpecData` documentation.

```{r new-ms-object-call, results='hide', warning=FALSE}
ms <- MassSpecData$new(files = files[1:3], headers = list(name = "Example 1"))
```

The `print` (or `show`) method is shown below by running the MassSpecData object name. A list of the included analyses is printed to the console matching the files added.

```{r print-method}
ms
```

Another option to obtain an overview of the `ms` object is the method `get_overview()` as shown below, returning a data.frame with the analysis details within the MassSpecData object.

```{r get-overview}
ms$get_overview()[, 1:5] # only the first 5 columns
```

## Internal fields

The MassSpecData object contains the following internal fields (or private fields) to store standardized data: **headers**, **settings**, **analyses**, **groups** and **alignment**. The internal fields cannot be directly accessed nor changed. Instead, the user must use the respective get and add methods, as shown in the sections [get methods](#get-methods) and [add methods](#add-methods), respectively. The add methods contain validation and conformity checks to ensure that the structure and semantics of added data are respected.

- The **headers** is also a Headers S3 class object and list with flexible headers of length one (e.g., name, path, date, etc.) that characterize the MassSpecData object. When not given, *name*, *path* and *date* are automatically generated (*name* is set to `NA_character_`, *path* is set to `getwd()` and *date* is set to `Sys.time()`) and included in the list, as they are required for workflow purposes. Other entries can be added by the user without type (or class) limitations but should be uniquely named and of length one. See `?Headers` for more information;
- The **settings** is a named `list` of ProcessingSettings S3 class objects used for data processing (see `?settings` for more information). Each added settings object is named according to the function call (or processing name). The settings interface is described in section [processing settings](#processing-settings);
- The **analyses** is a named `list` of MassSpecAnalysis S3 class objects representing each file added to the MassSpecData (see` ?MassSpecAnalysis` for more information). The name of each MassSpecAnalysis object in the analyses list is the name of the file (without extension). No duplicate file names (or analyses names) can exist in the MassSpecData. The structure of the MassSpecAnalysis S3 class object is described in section [analyses](#analyses); 
- The **groups** is a `data.table` with the feature groups after correspondence analysis (i.e., grouping of features across analyses). More information about **groups** is given in section [groups](#groups);
- The **alignment** is a `list` with the results of the retention time alignment for each analyses in the MassSpecData object. More information about **alignment** is given in section [alignment](#alignment).

## Get methods

Several methods are implemented to access data within the MassSpecData object, as already demonstrated for the `get_overview()`. A method is always used by writing the name of the MassSpecData object followed by a `$` and then the method call (e.g., `ms$get_overview()`). The internal fields can be accessed with `get_headers()`, `get_settings()`, `get_analyses()`, `get_groups()` and `get_alignment()`. For instance, the get method for the headers returns the list of entries. Note that the `path` and `date` were automatically created when the vignette was built even if not given is the headers argument when the MassSpecData object is created.

```{r get-headers}
ms$get_headers()
```

Similarly, the get method for analyses returns the list of MassSpecAnalysis objects. Below the `get_analyses()` is used to get the first analysis object (`a1`). Then, the `names()` function is used to show the names of entries in the `a1` MassSpecAnalysis, which represents the first *.mzML* file parsed to the `ms`. Note that the `get_analyses()` method always returns a list even for only one MassSpecAnalysis object. Therefore, we use the `[[` function as `a1[[1]]` to access the MassSpecAnalysis object itself. The analysis object contains `r length(ms$get_analyses(1)[[1]])` fields (listed below) with information and data parsed directly from the *.mzML* file. Variations in the content and structure (i.e., different vendors or data acquired with different conditions or modes) of the *.mzML* or *.mzXML* files added will affect the content of each entry in the MassSpecAnalysis object but the structure remains the same (see `?MassSpecAnalysis`). The function `parse_MassSpecAnalysis()` from the StreamFind R package is used to create and validate the list of MassSpecAnalysis objects from a character vector with *.mzML* or *.mzXML* full file paths (see `?parse_MassSpecAnalysis` for more information). The structure of the MassSpecAnalysis S3 class object is further described in section [analyses](#analyses). By typing `ms$get`, other available get methods are listed. The documentation `?MassSpecData` also contains detailed information and usage examples for available get methods.

```{r get-analyses}
a1 = ms$get_analyses(analyses = 1)

class(a1[[1]])

names(a1[[1]])
```

## Add methods

Add methods are used to add data to an existing MassSpecData object. For instance, below we add a new name and a description to the `ms` object with the method `add_headers()`. If entry names are already present in the MassSpecData object (e.g., *name*), the content of these are overwritten as shown below.

```{r add-headers}
ms$add_headers(name = "Ex1", description = "An example MassSpecData.")
```

```{r check-add-headers}
ms$get_headers()
```

Similarly, new MassSpecAnalysis objects can be added with `add_analyses()` as shown below. Note that the function `parse_MassSpecAnalysis()` was used to parse data from files 7, 8 and 9, creating a named list of analysis objects. Again the name of each MassSpecAnalysis in the list is the file name without extension. Then, the list is added to the `ms` object with the method `add_analyses()`. By typing `ms$add`, other available add methods are listed. The documentation `?MassSpecData` also contains detailed information and usage examples for available add methods.

```{r make-analysis-list}
a_list = parse_MassSpecAnalysis(files[7:9])

vapply(a_list, class, "")
```

```{r add-analyses}
ms$add_analyses(a_list)
```

```{r check-add-analyses}
ms
```

## Remove methods

Complementary to add methods, remove methods are used to remove data from an existing MassSpecData object. For instance, in the code chunk below the *name* and *description* are removed from the headers of the `ms` object. However, *name*, *path* and *date* cannot be removed. Even when one of these are specified, they are not removed. Yet, they can be changed with `add_headers()` method.

```{r remove-headers}
ms$remove_headers(c("name", "description"))
```

```{r check-remove-headers}
ms$get_headers()
```

Also, analyses can be removed with `remove_analyses()` method as shown below. When removing analyses, correspondence feature groups that were unique to those analyses are also removed from the MassSpecData object. By typing `ms$remove`, other available remove methods are listed. The documentation `?MassSpecData` also contains detailed information and usage examples for available remove methods.

```{r remove-analyses}
ms$remove_analyses(4:6)
```

```{r check-remove-analyses}
ms
```

## Subset methods

While remove methods changes the MassSpecData object by reference, the subset methods create a clone subset of the MassSpecData object. For instance, the method `subset_analyses()` is demonstrated below to create a clone of the `ms` object but with only the first 2 analyses. Note that the output behaves as a new MassSpecData object was created. In fact, that is what happens in the background. Also, note that when sub-setting all the other content (i.e., the other internal fields) are cloned to the new MassSpecData class object. For instance, the headers in the `subset_ms` object is the same as in `ms` object. All changes to the `subset_ms` object after its creation will not affect the `ms` object as it was cloned. For more information about copy and cloning R6 class objects you are recommended the [R6](https://r6.r-lib.org/index.html) official documentation. By typing `ms$subset`, other available subset methods are listed. The documentation `?MassSpecData` also contains detailed information and usage examples for available subset methods.

```{r subset-analyses}
subset_ms <- ms$subset_analyses(analyses = 1:2)

subset_ms
```

## Has methods

There are methods to simply check if data is present in the MassSpecData. These are the has methods as shown below. By typing `ms$has`, other available has methods are listed. The documentation `?MassSpecData` also contains detailed information and usage examples for available has methods.

```{r has-methods}
# check if settings are present
ms$has_settings()

# check if analyses are present
ms$has_analyses()

# check if features are present in analyses
ms$has_features()

# check if raw spectra are loaded to analyses
ms$has_loaded_spectra()
```

## Plot methods

Within the MassSpecData class object, standardized methods for plotting data are also available. All plot methods use data that can be obtained with get methods, meaning that the user is able to re-plot the data with a different approach/style. For instance, the `get_tic()` method can be used to obtain the total ion chromatogram (TIC) as a `data.table` with columns *analysis*, *rt* and *intensity*, as shown below for the first analysis. The same TIC data can be directly plotted using the `plot_tic()` method, as shown below.

```{r get-tic-first-analysis}
head(ms$get_tic(analyses = 1))
```

```{r plot-tic} 
ms$plot_tic(analyses = 1)
```

<br>

Note that plotting data which is not available (e.g., features in the current `ms` object as feature finding was not yet applied) returns `NULL` (demonstrated below).

```{r plot-not-present-data}
ms$plot_features()
```

<br>

Most plot methods have the *interactive* argument which is set to `TRUE` by default for using the [plotly](https://plotly.com/r/) package to plot the data. A static version of the plot using the base `plot()` function can be obtained by setting interactive to `FALSE`, as shown below for the base peak chromatogram (BPC) of all analyses in the `ms` object. By typing `ms$plot`, other available plot methods are listed. The documentation `?MassSpecData` also contains detailed information and usage examples for available plot methods.

```{r plot-bpc-static}
ms$plot_bpc(interactive = FALSE)
```

## Save methods

The MassSpecData object is stored by default in the environment of an R session and will be save within the *.RData* when `save.image()` function is used (by default save current environment is asked when an R session ends or is terminated). However, save methods are also available to save data within the internal fields in other formats. Currently only *.rds* and *.json* are possible. Below, the `save_headers()` method is used to save the Headers S3 class object as *.json*.

```{r save-headers-json}
ms$save_headers(format = "json", path = getwd())
```

```{r show-headers-json}
# the prettify function is applied for simplifying the json string output
jsonlite::prettify(readLines("headers.json"))
```

Similarly, `save_analyses()` method can be applied to export all or specific MassSpecAnalysis objects to *.rds* or *.json*. Below, the first MassSpecAnalysis object is exported as *.json* and named *a1*. Then in section [import methods](#import-methods), the exported analysis (*a1.json*) is imported to the `ms` object.

```{r}
ms$save_analyses(analyses = 1, format = "json", name = "a1")
```

Also, the entire `ms` object can be saved using the `save()` method as shown below. Other save methods are listed by typing `ms$save`. Again, the documentation `?MassSpecData` contains detailed information and usage examples for available save methods.

```{r save-entire-MassSpecData}
ms$save(format = "json", name = "ms", path = getwd())
```

## Import methods

Complementary to save methods, the *.json* or *.rds* files can be imported to the MassSpecData using the import methods. For instance, below the previously saved *headers.json* is imported back to the `ms` object. As in `add_headers()`, if entry names are already present, their content is overwritten.

```{r import-headers}
ms$import_headers("headers.json")
```

Also, the saved *a1.json* MassSpecAnalysis object can be imported with `import_analyses()` method. However, because the analysis name is already present in the `ms` object, an warning is printed and the operation not done (as shown below). The other internal fields (i.e., *settings*, *groups* and *alignment*) can be similarly imported. Validation and conformity tests are applied to ensure that the data has the required structure.

```{r import-analyses}
ms$import_analyses("a1.json")
```

A saved MassSpecData object can be imported from *.json* or *.rds* using the `import_MassSpecData()` function from the StreamFind package. Below, the previously saved *ms.json* is imported and compared to the `ms`.

```{r import-MassSpecData}
ms_imported = import_MassSpecData("ms.json")
```

```{r check-import-MassSpecData}
all.equal(ms, ms_imported)
```

```{r remove-files-json-from-wd, include=FALSE}
file.remove(paste0(getwd(), "/headers.json"))
file.remove(paste0(getwd(), "/a1.json"))
file.remove(paste0(getwd(), "/ms.json"))
```

# Analyses

Within the MassSpecData object, the internal field analyses is a list of MassSpecAnalysis S3 class objects representing MS files on disk. As shown above in [MassSpecData](#msdata) section, the MassSpecData can be initiated with the `files` argument using a character vector of full file path/s to *.mzML* or *.mzXML* files. Alternatively, the function `parse_MassSpecAnalysis()` can be used to generate the list of analysis objects and then the argument `analyses` of the `new()` method or the `add_analyses()` method can be used to add the MassSpecAnalysis objects to the MassSpecData object. The package [mzR](https://bioconductor.org/packages/release/bioc/html/mzR.html) from [Bioconductor](https://bioconductor.org/) is used to parse the data. When the [mzR](https://bioconductor.org/packages/release/bioc/html/mzR.html) package is not available (i.e., not installed), the [xml2](https://cran.r-project.org/web/packages/xml2/index.html) package from CRAN is used instead but is much slower. Specially for large MS files (e.g., high resolution MS files with ion mobility), the [mzR](https://bioconductor.org/packages/release/bioc/html/mzR.html) package is recommended. The data parsed is the same independently of the package used. The structure of an example MassSpecAnalysis object is shown below and consists of 23 entries (see `?MassSpecAnalysis` for more information).

- **name** is the file name without extension;
- **replicate** is the analysis replicate group name; 
- **blank** is the associated blank replicate group name;
- **file** is the full file path;
- **type** is either "MS" for only MS1 spectral data, "MS/MS" for tandem spectral data (i.e., MS1 and MS2) or "SRM" for selected reaction monitoring data (i.e., no spectra only chromatograms);
- **instrument** is a list with metadata from the instrument used to acquire the data (content is highly vendor dependent);
- **time_stamp** is the start time and date of the data acquisition;
- **spectra_number** is the number of spectra in the file;
- **spectra_mode** is one of *centroid*, *profile* or `NA` for centroided data, data in profile mode or mode not defined or absence spectra, respectively.
- **spectra_levels** is the number of levels in the spectra (i.e., 1 or 1 and 2);
- **mz_low** is the lowest \emph{m/z} value detected in the spectra, in Da;
- **mz_high** is the highest \emph{m/z} value detected in the spectra, in Da;
- **rt_start** is the run start time, in seconds;
- **rt_end** is the run end time, in seconds;
- **polarity** is the polarity mode: *positive*, *negative* or *both* (the latter refers to polarity switching acquisition);
- **chromatograms_number** is the number of chromatograms (only applicable for *.mzML* files);
- **ion_mobility** `TRUE` or `FALSE` for presence or absence of drift separation from ion mobility;
- **tic** is a `data.table` with the total ion chromatogram;
- **bpc** is a `data.table` with the base peak chromatogram;
- **spectra** is a `data.table` with the raw spectra (only present if loaded with `load_spectra()` method);
- **chromatograms** is a `data.table` with the raw chromatograms (only present if loaded with `load_chromatograms()` method);
- **features** is a `data.table` with the features obtained from data processing (e.g., `find_features()` method);
- **metadata** is a `list` with flexible storage for experimental metadata (e.g., concentration, location, etc.);

```{r}
str(a1[[1]])
```

## Get basic information

The data from MassSpecAnalysis objects can be obtained with get methods as shown above. Below, methods to get information from analyses are demonstrated. In all methods except the `get_number_analyses()`, the analyses argument can be used to query information from specific analyses, using the analysis name or the index as shown below.

```{r}
# getter for number of analyses
ms$get_number_analyses()

# getter for file names (i.e., analysis names)
ms$get_analysis_names(analyses = 1:2)

# getter for file full paths of analyses
ms$get_files(3)

# getter for replicate group names
ms$get_replicate_names()

# getter for associated blank replicate group names
ms$get_blank_names()

# getter for polarities
ms$get_polarities(analyses = "00_hrms_s_is_pos_cent-r003")

# getter for lower mz value
ms$get_mz_low(c("00_hrms_s_is_pos_cent-r001", "00_hrms_s_is_pos_cent-r003"))

# getter for higher mz value
ms$get_mz_high()

# getter for start retention time
ms$get_rt_start()

# getter for retention time end
ms$get_rt_end()

# getter for MS levels. Note that it returns a list not a vector!
ms$get_spectra_levels()

# getter for spectra mode
ms$get_spectra_mode()

# getter for TIC
head(ms$get_tic(analyses = 2))

# getter for BPC
head(ms$get_bpc())
```

## Making spectra targets

Besides analysis information, low level data (i.e., spectra and chromatograms) can also be obtained with specific get methods. For spectra, the definition of targets to selectively parse small pieces of the raw data is essential. In StreamFind, targets can be comprehensively assembled with the function `make_ms_targets()`, as demonstrated below for Carbamazepin-d10 and Diuron-d6 (both internal standards spiked to the analyses in the `ms` object). The function `make_ms_targets()` has five arguments (*mz*, *rt*, *ppm*, *sec* and *id*) that are used to assemble the targets. Below, different ways are demonstrated to build MS targets.

```{r carb-diu-db}
# Carbamazepin-d10 is ionized only in positive mode
carbamazepin_d10 <- db[name %in% "Carbamazepin-d10", .(name, mass, rt)]
carbamazepin_d10

# Expected m/z and time of Carbamazepin-d10
carb_mz_pos <- carbamazepin_d10$mass + 1.0073 # for positive ionization ([M+H]+)
carb_mz_pos
carb_rt <- carbamazepin_d10$rt
carb_rt

# Diuron-d6 is ionized in both positive and negative modes
diuron_d6 <- db[name %in% "Diuron-d6", .(name, mass, rt)]
diuron_d6

# Expected m/z and time of Diuron-d6
diu_mz_pos <- diuron_d6$mass + 1.0073 # for positive ionization ([M+H]+)
diu_mz_pos
diu_rt <- diuron_d6$rt
diu_rt

# Time and mass deviations to make the targets
sec_dev <- 30 # in seconds
ppm_dev <- 20 # in ppm
```

**Case 1:** Individual *m/z* and time targets with deviations.
```{r case-one}
# m/z targets, in Da
mz1 <- c(carb_mz_pos, diu_mz_pos)

# Time targets, in seconds
rt1 <- c(carb_rt, diu_rt)

# Optional identifiers for the targets
id1 <- c("target1", "target2")

targets1 <- make_ms_targets(mz = mz1, rt = rt1, ppm = ppm_dev, sec = sec_dev, id = id1)

targets1
```

**Case 2:** *m/z* and time minimum and maximum ranges as `data.frame` given in `mz` and `rt`, respectively.
```{r case-two}
# Calculating mass deviation in Da from 200 ppm
ppm_carb <- ppm_dev / 1E6 * carb_mz_pos
ppm_diu <- ppm_dev / 1E6 * diu_mz_pos

# Min and Max m/z ranges, in Da
mz2 <- data.frame(
  "mzmin" = c(carb_mz_pos - ppm_carb, diu_mz_pos - ppm_diu),
  "mzmax" = c(carb_mz_pos + ppm_carb, diu_mz_pos + ppm_diu)
)
# Min and Max time ranges, in seconds
rt2 <- data.frame(
  "rtmin" = c(carb_rt - sec_dev, diu_rt - sec_dev),
  "rtmax" = c(carb_rt + sec_dev, diu_rt + sec_dev)
)

targets2 <- make_ms_targets(mz = mz2, rt = rt2)

targets2
```

**Case 3:** *m/z* and time targets given as a `data.frame` in `mz` and mass and time deviations given in `ppm` and `sec`, respectively.
```{r case-three}
# data.frame with m/z and time targets
mz3 <- data.frame(
  "id" = c("target1", "target2"),
  "mz" = c(carb_mz_pos, diu_mz_pos),
  "rt" = c(carb_rt, diu_rt)
)

targets3 <- make_ms_targets(mz = mz3, ppm = ppm_dev, sec = sec_dev)

targets3
```

**Case 4:** *m/z* and time minimum and maximum ranges given directly as a `data.frame` in `mz`. 
```{r case-four}
# data.frame with m/z and time minimum and maximum ranges
mz4 <- data.frame(
  "id" = c("target1", "target2"),
  "mzmin" = c(carb_mz_pos - ppm_carb, diu_mz_pos - ppm_diu),
  "mzmax" = c(carb_mz_pos + ppm_carb, diu_mz_pos + ppm_diu),
  "rtmin" = c(carb_rt - sec_dev, diu_rt - sec_dev),
  "rtmax" = c(carb_rt + sec_dev, diu_rt + sec_dev)
)

targets4 <- make_ms_targets(mz = mz4)

targets4
```

Although different ways were used to build the `data.frame` with the targets, the result should be the same as tested below. Note that when exact mass (or *m/z*) and time targets are not given (i.e., only given minimum and maximum ranges), these are calculated as an average of the minimum and maximum values (see cases 2 and 4 above).

```{r test-targets}
#test all equal target 1
t1 <- rbind(targets1[1, ], targets2[1,], targets3[1, ], targets4[1, ])
all(round(apply(t1[, 2:7], 2, sd), digits = 4) == 0)

#test all equal target 2
t2 <- rbind(targets1[2, ], targets2[2,], targets3[2, ], targets4[2, ])
all(round(apply(t2[, 2:7], 2, sd), digits = 4) == 0)
```

## Get and plot spectra

Besides the `get_tic()` and `get_bpc()` methods (shown above). There are four methods to get low level spectra from analyses: `get_spectra()`, `get_eic()`, `get_ms1()` and `get_ms2()`. All methods use the same arguments as the function `make_ms_targets()` to build MS targets and the *analyses* argument to specify which MassSpecAnalysis objects to query the data. All methods return a `data.table`. The documentation in `?MassSpecData` has detailed information about the use of each method. For each get method, a plot method is available with similar interface. Below each get and plot method is demonstrated for the defined targets.

### Mass traces

Spectra can be obtained with `get_spectra()` method and plotted with `plot_spectra()`. When the `levels` argument includes 2, extra columns are given with the applied collision energy (ce), the precursor scan number (preScan) and the isolated *m/z* value (preMZ). Arguments `minIntensityMS1` and `minIntensityMS2` can be used to define a minimum intensity of parsed mass traces. The `allTraces` argument can be set to `FALSE` for only returning MS2 data correspondent of defined targets. For this, the `isolationWindow` argument is used to calculate the isolation window of each target and retrieve the MS2 based on the preMZ and rt columns (this is demonstrated below in the second spectra plot).

```{r raw-spectra-01}
# Getter for level 1 raw spectra of targets from analysis 1
head(ms$get_spectra(analyses = 1, mz = targets1, levels = 1))
```

```{r raw-spectra-02}
# Getter for level 2 raw spectra of targets from analysis 1
head(ms$get_spectra(analyses = 1, mz = targets1, levels = 2))
```

```{r plot-raw-spectra-01}
# Plots raw spectra (level 1) of all analyses with intensity above 1000
ms$plot_spectra(mz = targets1, levels = 1, minIntensityMS1 = 1000)
```

```{r plot-raw-spectra-02}
# Plots raw spectra (level 1 and 2) of analysis 2 and target 1 while coloring by levels
# Note that allTraces was set to FALSE to retrieve MS2 of MS1 targets considering an isolation window of 1.3 Da
ms$plot_spectra(
  analyses = 2, mz = targets1[1, ], levels = c(1,2), 
  allTraces = FALSE, isolationWindow = 1.3, colorBy = "levels",
  minIntensityMS1 = 1000, minIntensityMS2 = 10
)
```

Another plotting option for raw spectra is the `plot_xic()` method as shown below. This method returns a two row plot with traces intensity vs retention time on top and *m/z* vs retention time below. A `targetsMark` can be defined to illustrate the area of expected *m/z* and retention time targets. Below, the targets1 rt and mz columns with narrower ppm and sec deviations (i.e., 5 ppm and 10 seconds, respectively) is used to show the area where Carbamazepin-d10 is expected for potential identification.

```{r plot-xic}
ms$plot_xic(analyses = 1:2, mz = targets1[2,],
  targetsMark = targets1[2, c("rt", "mz")], ppmMark = 5, secMark = 10)
```

### EIC

The extracted ion chromatogram (EIC) can be obtained with `get_eic()` and plotted with `plot_eic()` method. The interface is very similar as the previous methods, with analyses and make targets arguments used to specify each analyses to parse the data and which traces to retrieve.

```{r get-eic-01}
ms$get_eic(analyses = 1, mz = mz1, rt = rt1, ppm = 5, sec = 10)
```

```{r plot-eic-01}
ms$plot_eic(mz = targets1, interactive = FALSE)
```

```{r plot-eic-02}
ms$plot_eic(mz = targets1, colorBy = "analyses")
```

### MS1 and MS2 spectra

While the EIC method collects and sums the traces in each retention time value (i.e., in each scan), the `get_ms1()` and `get_ms2()` methods average the mass traces in a given retention time interval, returning an averaged spectrum representing several scans. The averaging is based on clustering the mass traces according to a given `mzClust` argument as demonstrated below for both MS1 and MS2 data.

```{r get-ms1}
# MS1 averaged spectrum of targets from 247 to 257 (Carbamazepin-d10) and 239 to 245 (Diuron-d6) m/z and +/- 1 second the retention time of each target
ms$get_ms1(analyses = 3, mzClust = 0.005,
  mz = data.frame("mzmin" = c(247, 239), "mzmax" = c(257, 245)),
  rt = data.frame("rtmin" = targets1$rt - 1, "rtmax" = targets1$rt + 1)
) 
```

```{r plot-ms1}
# In the plot, the Carbamazepin-d10 and Diuron-d6 isotopes can be distinguished
ms$plot_ms1(analyses = 3,
  mz = data.frame("mzmin" = c(247, 239), "mzmax" = c(257, 245)),
  rt = data.frame("rtmin" = targets1$rt - 1, "rtmax" = targets1$rt + 1),
  mzClust = 0.005, interactive = FALSE
)
```

```{r get-ms2}
# MS2 data of Diuron-d6 for analysis 1
ms$get_ms2(analyses = 1, mz = targets1[2, ])
```

```{r plot-ms2-01}
# MS2 data plot of Diuron-d6 in all analyses
ms$plot_ms2(mz = targets1[2 ,], 
  mzClust = 0.005, minIntensity = 50,
  colorBy = "analyses", verbose = FALSE)
```

```{r plot-ms2-02}
# MS2 data plot of both targets in analyses 3
# Note that "minIntensity" argument was increased to 500 counts, resulting in a cleaner averaged spectrum
ms$plot_ms2(analyses = 3, mz = targets1, 
  mzClust = 0.005, minIntensity = 500,
  colorBy = "targets", verbose = FALSE)
```

# Centroiding

Possible using the function `convert_ms_files()` which utilizes the [MSConvert](https://proteowizard.sourceforge.io/download.html) command line interface. See `help(convert_ms_files)` for more information. Alternative centroiding to be implemented.

# Processing settings

Data processing steps often require settings with parameters to support the application of a given function to the data. In the MassSpecData object, methods that require settings have the `settings` argument (e.g., the method `find_features()`). In the StreamFind package, the `settings` argument takes a ProcessingSettings S3 class and list object (see `?ProcessingSettings`) with three mandatory elements: *call*, *algorithm* and *parameters*. The *call* defines the name of the method (or processing step to be applied), such as "find_features" or "group_features". The *algorithm* defines the algorithm to be used for the processing step (e.g., "xcms" or "openms"). Finally, the *parameters* are a list of algorithm dependent entries that define the settings to be applied during the data processing. An example for "find_features" settings using the algorithm "xcms" is given below. Most of the data processing methods in StreamFind use the package [patRoon](https://github.com/rickhelmus/patRoon). Therefore, the interface parameters is similar. The algorithm and parameters (named "param" in [patRoon](https://github.com/rickhelmus/patRoon)) can be taken from the respective documentation of the equivalent function. For instance, the `find_features()` method in StreamFind uses the [findFeatures()](https://rickhelmus.github.io/patRoon/reference/findFeatures.html) function in [patRoon](https://github.com/rickhelmus/patRoon). The documentation of each data processing method in `?MassSpecData` explains and exemplifies the usage and where the required settings can be acquired in [patRoon](https://github.com/rickhelmus/patRoon).

```{r ff-settings-xcms}
settingsFindFeatures <- Settings_find_features_xcms3_centwave(
  ppm = 12,
  peakwidth = c(5, 30),
  snthresh = 10,
  prefilter = c(5, 1500),
  mzCenterFun = "mean",
  integrate = 2,
  mzdiff = -0.0001,
  fitgauss = TRUE,
  noise = 500,
  verboseColumns = TRUE,
  firstBaselineCheck = FALSE,
  extendLengthMSW = TRUE
)

settingsFindFeatures
```

The ProcessingSettings object can be added to the MassSpecData object with the `add_settings()` method as shown below. Similarly, it can be obtained with the `get_settings()` method using the call name as also shown below.

```{r add-ff-xcms}
ms$add_settings(settings = settingsFindFeatures)
```

```{r get-ff-xcms}
# Note that it returns a named list, therefore we access it with the [[ method
ms$get_settings(call = "find_features")[["find_features"]]
```

# Features

Extraction of features from raw MS data is a frequent step applied during data processing for easing prioritization and identification of chemicals in a given analysis (i.e., sample). A feature in essence describes a chromatographic peak composed fundamentally of three values: mass-to-charge ratio (*m/z*), retention time (rt) and intensity. The method `find_features()` from a MassSpecData object can be used to extract features from files represented by the MassSpecAnalysis objects. Several algorithms (i.e., software) are available for finding features through the [patRoon](https://github.com/rickhelmus/patRoon) package as above mentioned. Some algorithms require spectral data in profile mode but most use centroided data (see [centroiding](#centroiding) section for more information about centroid vs profile mode). Below, we use the added settings to find features from the analyses in the object `ms`. Note that the argument `settings` was not used (`NULL` is the default value) as the settings were already added to the object. Adding settings via the argument `settings` during call for method `find_features()` would replace existing settings in the MassSpecData object with call name "find_features". If settings with call name "find_features" are not available or given, features are not extracted.

```{r ff-01, warning=FALSE, results='hide'}
ms$find_features(settings = NULL)
```

```{r print-ms-after-ff}
# Check for presence of features in analyses
ms$has_features()

# Show method giving the number of features
ms
```

## Get features

As above described, there are get methods to parse data from a MassSpecData class object. For features, the method `get_features()` can be used as shown below. When no arguments are given, the features `data.table` in each MassSpecAnalysis object is collated without filtering and a column analysis is given to make reference for the analysis that the corresponding features belong to.

```{r get-all-features}
head(ms$get_features())
```

The same arguments from `make_ms_targets()` are available. Below the corresponding feature of Carbamazepin-d10 in each analysis object is extracted using the `mz` and `rt` arguments.

```{r get-features-targets-01}
ms$get_features(mz = carb_mz_pos, rt = carb_rt, ppm = 5, sec = 10)
```

The features `data.table` contain the following mandatory columns:

 - **analysis**: (character) The file name without extension;
 - **feature**: (character) The unique identifier of the feature in the analysis;
 - **index**: (numeric) The index of the feature in the analysis;
 - **mz**: (numeric) The calculated *m/z* from feature mass traces (i.e., centroids), in Da;
 - **rt**: (numeric) The calculated retention time from feature mass traces, in seconds;
 - **mzmin**: (numeric) The minimum *m/z* value of the feature mass traces, in Da;
 - **mzmax**: (numeric) The maximum *m/z* value of the feature mass traces, in Da;
 - **rtmin**: (numeric) The minimum retention time value of the feature mass traces, in seconds;
 - **rtmax**: (numeric) The maximum retention time value of the feature mass traces, in seconds;
 - **intensity**: (numeric) The calculated height of the features, in counts; 
 - **area**: (numeric) The calculated area of the integrated feature;

Other columns added from specific algorithms or added by StreamFind after feature extraction are:

 - **adduct**: (character) The adduct label of the feature (e.g., [M+H]+, [M-H]-, [M+Na]+), see  [isotopes and adducts](#isotopes-and-adducts) section for more information;
 - **mass**: (numeric) The calculated neutral mass of the feature based on the polarity of the analysis;
 - **is_filled**: (logical) `TRUE` means that the feature was built with recursive integration (i.e., integrated after correspondence analysis, see [groups](#groups) section for more information);
 - **filtered**: (logical) `TRUE` means that the feature was filtered out from the set but still remained in the data set (see [filters](#filters) section for more information);
 - **filter**: (character) The filter tag or `NA_character_` if the feature is not filtered (i.e., filtered is `FALSE`);
 - **sn**: (numeric) The calculated signal-to-noise ratio of the feature;
 - **ms1** data.table with averaged spectrum from MS1 spectra data within defined mass and time ranges;
 - **ms2** data.table with averaged spectrum from MS2 spectra data that correspond to the isolation window of the feature mass and time ranges;

## Plot features

There are two methods to plot features: `plot_features()` and `map_features()`. Both methods are demonstrated below for the features correspondent of the target chemicals. While the `plot_features()` method plots the chromatogram to inspect intensity, shape and other fitting aspects the `map_features()` method plots the time and mass space (shaded area) of the features as well as the calculate *m/z* and retention time values (dots). For instance, when applying correspondence analysis (more information in [groups](#groups) section), these plots can be used to evaluate the most appropriate mass and time deviations to consider in the processing settings.

```{r plot-features-targets-01}
ms$plot_features(analyses = 1, mz = targets1, colorBy = "targets", interactive = FALSE)
```

```{r plot-features-targets-02}
ms$plot_features(mz = targets1, colorBy = "analyses", interactive = TRUE)
```

```{r map-features-targets-01}
ms$map_features(mz = targets1, colorBy = "analyses", interactive = TRUE)
```

## Get features MS1 and MS2

As mentioned, each feature represents an MS1 level chromatographic peak. The averaged MS1 spectrum can be obtained with the method `get_features_ms1()`. Spectra within the feature mass and time space are collected and clustered according to a `mzClust` argument (in Da), as shown below.  The arguments `rtWindow` and `mzWindow` can be used to narrow or extend the space for collection of mass traces. In the example below, the time window for collection of mass traces will be plus/minus 2 seconds the calculated retention time of the feature found by the `mz` and `rt` arguments; In the example, it is the feature correspondent to Carbamazepin-d10. The mass window will be -1 Da and +6 Da the calculated *m/z*. When both `rtWindow` and `mzWindow` are `NULL`, the integrated feature space (as plotted by the `map_features()` method) will be used. A `minIntensity` argument is used to filter mass traces below an intensity threshold. In the example below, it was set to 1000 counts, to exclude low intensity mass traces. The `data.table` obtained can be plotted directly with the `plot_features_ms1()` method as shown below.

```{r get-features-ms1}
ms$get_features_ms1(analyses = 2, mz = carb_mz_pos, rt = carb_rt,
  ppm = 20, sec = 30, rtWindow = c(-2, 2), mzWindow = c(-1, 6),
  mzClust = 0.003, minIntensity = 1000
)
```

```{r plot-features-ms1}
# Plotting the MS1 spectrum of both Carbamazepin-d10 and Diuron-d6 in analyses 2
ms$plot_features_ms1(analyses = 2, verbose = FALSE,
  mz = targets1, rtWindow = c(-2, 2), mzWindow = c(-1, 6),
  mzClust = 0.003, minIntensity = 1000, colorBy = "targets"
)
```

MS2 data from features is obtained with the `get_features_ms2()` method, as shown below. For MS2, the arguments `mzClust` and `minIntensity` are applied the same way as for MS1. However, when extracting traces an isolation window (in Da) should be set using the argument `isolationWindow`. The isolation window should match with the settings used to isolate precursor ions during fragmentation from tandem MS. Below we set the `isolationWindow` to 1.3 Da, which extends the *mzmin* to *mzmax* range of the feature by $-\frac{1.3}{2}$ and $+\frac{1.3}{2}$, respectively. Plotting the `data.table` can also be directly performed with the `plot_features_ms2()` method as shown below.

```{r get-features-ms2}
ms$get_features_ms2(analyses = 3, mz = carb_mz_pos, rt = carb_rt,
  ppm = 20, sec = 30, isolationWindow = 1.3, mzClust = 0.003,
  minIntensity = 500, verbose = FALSE
)
```

```{r plot-features-ms2}
ms$plot_features_ms2(mz = carb_mz_pos, rt = carb_rt,
  ppm = 20, sec = 30, isolationWindow = 1.3, mzClust = 0.003,
  minIntensity = 500, colorBy = "analyses", verbose = FALSE
)
```

```{r plot-features-ms2-2}
ms$plot_features_ms2(analyses = 1, mz = targets1,
  isolationWindow = 1.3, mzClust = 0.003,
  minIntensity = 500, colorBy = "targets",
  interactive = FALSE, verbose = FALSE
)
```

## Load features MS1 and MS2

MS1 and MS2 spectra can support identification processing steps by giving additional information from isotopic and specific fragmentation patterns, respectively. The methods `load_features_ms1()` and `load_features_ms2()` are used for systematically load averaged MS1 and MS2 spectra for features in a MassSpecData class object. Both methods have a `settings` argument which takes a settings object similar to "find_features" settings but with the specific call name, which is the method name, as shown below. The parameters are the same as the arguments of the respective get methods (see above in [get features ms1 and ms2](#get-features-ms1-and-ms2)). The settings can be also added to the settings internal field as explained above in [processing settings](#processing-settings) section. This allows for internal data re-processing, inspection and retrospective evaluation.

```{r features-ms1-ms2-settings}
# Settings for loading features MS1 spectra
slfms1 <- Settings_load_features_ms1_StreamFind(
  rtWindow = c(-2, 2),
  mzWindow = c(-1, 6),
  mzClust = 0.003,
  presence = 0.9,
  minIntensity = 250,
  filtered = FALSE,
  runParallel = FALSE,
  verbose = FALSE
)

ms$add_settings(settings = slfms1)

# Settings for loading features MS2 spectra
slfms2 <- Settings_load_features_ms2_StreamFind(
  isolationWindow = 1.3,
  mzClust = 0.003,
  presence = 0.9,
  minIntensity = 0,
  filtered = FALSE,
  runParallel = FALSE,
  verbose = FALSE
)

ms$add_settings(settings = slfms2)

# Returns the names of the settings list, which is also the call or method name
names(ms$get_settings())
```

```{r subset-ms-target-features}
# The ms object is first reduced in size to ease computation for this example
targetFeatures <- ms$get_features(mz = targets1)
targetFeatures <- targetFeatures[, c("analysis", "feature"), with = FALSE]
targetFeatures

# Only the features from targets are kept in the MassSpecData object
msTargets <- ms$subset_features(features = targetFeatures)

# Only 2 features in each analysis object
msTargets
```

As the settings were already added to `ms`, the settings are also in `msTargets`. In general, the sub-setting methods transfer everything to the new MassSpecData object except what is to be excluded (see [subset-methods](#subset-methods) section for more information). Below, the MS1 spectra for the remaining features (i.e., feature corresponding to Carbamazepin-d10 and Diuron-d6 in each analysis) are loaded to the `msTargets`.

```{r load-ms1-targets}
msTargets$load_features_ms1()

msTargets$has_loaded_features_ms1()
```

When loaded, the MS1 data is added as a new column in the features `data.table` of each analysis object. Therefore, there are two ways of getting the MS1 spectra as demonstrated below. The use of `get_features_ms1()` method is recommended over accessing directly from the features `data.table`. The method `get_features_ms1()` has the argument *loadedMS1* set to `TRUE` by default for when loaded MS1 spectra are present, data is queried from the loaded spectra rather then extracting from raw data. Note that when *loadedMS1* is `TRUE`, averaging settings are ignored as the spectra were already averaged during loading of MS1 data. If *loadedMS1* is set to `FALSE`, data is extracted from raw data even if loaded spectra are present in the features table.

```{r get-loaded-ms1-1}
msTargets$get_features_ms1(analyses = 1, mz = targets1, loadedMS1 = TRUE)
```

```{r plot-loaded-ms1-1}
# Plotting all loaded MS1 spectra by feature id
msTargets$plot_features_ms1(loadedMS1 = TRUE)
```

As an alternative, MS1 spectra can be obtained directly from the features `data.table` as shown below but it is not recommended.

```{r get-loaded-ms1-2}
# Get MS1 spectra from features data.table
ms1FromFeaturesDataTable <- msTargets$get_features()
head(ms1FromFeaturesDataTable)

# Column ms1 is a list of data.table objects with the spectrum of each feature or NULL when a given feature does not have mass traces
class(ms1FromFeaturesDataTable$ms1)

# Extract the first spectrum with the "[[" method
ms1FromFeaturesDataTable$ms1[[1]]
```

Similarly to MS1, MS2 data can be loaded as shown below. Note the settings were already added to the former `ms` object. Still, one can add the settings again in the method call. The previously stored settings will then be overwritten. In this case, there is no change as the settings are the same.

```{r load-ms1-ms2-targets}
msTargets$load_features_ms2(settings = slfms2)

msTargets$has_loaded_features_ms2()
```

The loaded MS2 spectra can be obtained with the method `get_features_ms2()` with the argument *loadedMS2* set to `TRUE` (the default value). Alternatively, but not recommended, the MS2 spectra can be obtained directly from the features `data.table` as shown above for MS1 spectra.

```{r get-loaded-ms2-1}
# The first 6 lines of the MS2 spectra
head(msTargets$get_features_ms2(analyses = 1, mz = targets1, loadedMS2 = TRUE))
```

```{r plot-loaded-ms2-1}
# Plotting all loaded MS2 spectra in analysis 2 by feature id
msTargets$plot_features_ms2(analyses = 2, loadedMS2 = TRUE)
```

When MS1 or MS2 data is not available for a given feature (i.e., within the feature *m/z* and retention time window there is no mass traces or the MS1 traces were not fragmented, respectively), a `NULL` value is given. So, in the *ms1* and *ms2* columns of the features `data.table` a `NULL` entry is present if the spectrum is absent. When returning bulk MS1 and MS2 spectra with `get_features_ms1()` and `get_features_ms2()`, respectively, the features with `NULL` (no spectra) are not listed in the returned `data.table`.

## Remove features MS1 and MS2

Loaded MS1 and MS2 feature spectra can be removed using the `remove_features_ms1()` and `remove_features_ms2()`. Note that all the loaded spectra are removed as no argument can be given to remove only certain spectra. This means that if features MS1 or MS2 spectra are loaded, they are loaded for all features. Removing features from the MassSpecData object will also remove MS1 and MS2 spectra as they are in the features `data.table`.

```{r remove-features-ms1-ms2}
msTargets$remove_features_ms1()
msTargets$has_loaded_features_ms1()

msTargets$remove_features_ms2()
msTargets$has_loaded_features_ms2()
```

# Isotopes and adducts

Not yet implemented.

# Groups

After extraction of features, correspondence analysis (i.e., grouping of features across analyses) is often applied as a following processing step. The result from correspondence analysis is feature groups, describing the correspondent features in each analysis. The method `group_features()` is used for correspondence analysis of features in a MassSpecData object. Similarly to `find_features()`, the `group_features()` method uses settings. Also here the [patRoon](https://github.com/rickhelmus/patRoon) package is used in the background. More precisely, the function [groupFeatures()](https://rickhelmus.github.io/patRoon/reference/groupFeatures.html) from [patRoon](https://github.com/rickhelmus/patRoon) is used. Therefore, the settings algorithm and parameters can be obtained by the documentation of [groupFeatures()](https://rickhelmus.github.io/patRoon/reference/groupFeatures.html). Below the correspondence analysis is demonstrated using the algorithm "xcms".

```{r grouping-settings}
settingsGroupFeatures <- Settings_group_features_xcms3_peakdensity(
  bw = 5,
  minFraction = 0.5,
  minSamples = 1,
  binSize = 0.008,
  maxFeatures = 100
)
```

```{r group-features, results='hide'}
ms$group_features(settings = settingsGroupFeatures)
```

## Get groups

Feature groups are stored in the groups internal field as a `data.table`, which can be obtained with the method `get_groups()`, as shown below. The groups `data.table` has the mandatory columns:

 - **group**: (character) The identifier of the feature group;
 - **rt**: (numeric) The averaged retention time of the features in the group, in seconds;
 - **mz**: (numeric) The averaged *m/z* of the features in the group, in Da (not present for MassSpecData with multiple polarities);
 - **mass**: (numeric) The averaged neutral mass of the features in the group, in Da;
 - **adduct**: (character) The adduct ion assigned to the feature group;
 - **drt**: (numeric) The time deviation range of the feature group, in seconds;
 - **dppm**: (numeric) The mass deviation range of the feature group, in ppm;
 - **index**: (integer) The unique index of the feature group;
 - **hasFilled**: (logical) When `TRUE`, the feature group has features obtained by recursive integration applied after  correspondence analysis for filling analysis objects with missing features;
 - **filtered**: (logical) `TRUE` means that the feature group was filtered out from the set but still remained in the data set (see [filters](#filters) section for more information);
 - **filter**: (character) The filter tag or `NA_character_` if the feature group is not filtered (i.e., filtered is `FALSE`);

For each MassSpecAnalysis object, a column with the analysis name is added to the groups `data.table` with the intensity of the feature. Other columns are added by StreamFind from specific workflow methods but these are explained in the respective method demonstration and documentation.

```{r get-groups-field}
head(ms$get_groups())
```

```{r get-groups-targets}
# Groups correspondent of CArbamazepin-d10 and Diuron-d6 
ms$get_groups(mz = targets1)
```

## Plot groups

As for features, plotting methods for groups are also available. Below the methods `plot_groups()` and `plot_groups_overview()` are demonstrated. Further information about usage is available in the MassSpecData documentation (`?MassSpecData`).

```{r plot-groups}
ms$plot_groups(mz = targets1, colorBy = "analyses")
```

```{r plot-groups-overview, fig.height=8}
# Plots an overview of the features within the group for each analysis
ms$plot_groups_overview(mz = targets1)
```

## Get groups MS1 and MS2

MS1 and MS2 spectra spectra for groups can be obtained with the methods `get_groups_ms1()` and `get_groups_ms2()`, respectively. Note that average MS1 and MS2 for feature groups (i.e., across analyses) requires pre-averaging of spectra within each feature. Therefore, the arguments for averaging spectra for features are available in the methods, as shown below. These can be set differently for averaging spectra within features and within feature groups. Below the same values are used for both.

```{r get-groups-ms1}
ms$get_groups_ms1(mz = carb_mz_pos, rt = carb_rt,
  ppm = 20, sec = 30, rtWindow = c(-2, 2), mzWindow = c(-1, 6),
  mzClustFeatures = 0.003, minIntensityFeatures = 1000,
  mzClustGroups = 0.003, minIntensityGroups = 1000
)
```

```{r plot-groups-ms1}
ms$plot_groups_ms1(mz = targets1,
  ppm = 20, sec = 30, rtWindow = c(-2, 2), mzWindow = c(-1, 6),
  mzClustFeatures = 0.003, minIntensityFeatures = 1000,
  mzClustGroups = 0.003, minIntensityGroups = 1000,
)
```

```{r get-groups-ms2}
ms$get_groups_ms2(mz = carb_mz_pos, rt = carb_rt,
  ppm = 20, sec = 30, isolationWindow = 1.3,
  mzClustFeatures = 0.003, minIntensityFeatures = 500,
  mzClustGroups = 0.003, minIntensityGroups = 500
)
```

```{r plot-groups-ms2}
ms$plot_groups_ms2(mz = targets1,
  ppm = 20, sec = 30, isolationWindow = 1.3,
  mzClustFeatures = 0.003, minIntensityFeatures = 500,
  mzClustGroups = 0.003, minIntensityGroups = 500
)
```

## Load groups MS1 and MS2

Similarly to features, loading MS1 and MS2 spectra for feature groups is also possible via the methods `load_groups_ms1()` and `load_groups_ms2()`. Also here settings are used for loading the spectra and when spectra are not available for a given feature group, `NULL` is returned. Below, the loading process is demonstrated for a subset as earlier shown.

```{r grouping-ms1-ms2, results='hide'}
targetGroups <- ms$get_groups(mz = targets1)

# Only the groups from targets are kept in the MassSpecData object
msTargets <- ms$subset_groups(groups = targetGroups$group)

# Removing filtered features (i.e., not belonging to targeted groups)
msTargets$remove_features(filtered = TRUE)

# Only 6 features and 2 groups left 
msTargets
```

```{r groups-ms1-ms2-settings}
# Settings for loading groups MS1 spectra
slfgms1 <- Settings_load_groups_ms1_StreamFind(
  mzClust = 0.003,
  presence = 0.9,
  minIntensity = 250,
  filtered = FALSE,
  runParallel = FALSE,
  verbose = FALSE
)

msTargets$add_settings(settings = slfgms1)

# Settings for loading groups MS2 spectra
slfgms2 <- Settings_load_groups_ms2_StreamFind(
  mzClust = 0.003,
  presence = 0.9,
  minIntensity = 0,
  filtered = FALSE,
  runParallel = FALSE,
  verbose = FALSE
)

msTargets$add_settings(settings = slfgms2)

# Returns the names of the settings list, which is also the call or method name
names(msTargets$get_settings())
```

```{r load-groups-ms1}
msTargets$load_groups_ms1()

msTargets$has_loaded_groups_ms1()
```

```{r load-groups-ms2}
msTargets$load_groups_ms2()

msTargets$has_loaded_groups_ms2()
```

Note that features MS1 and MS2 spectra were removed in section [remove features ms1 and ms2](#remove-features-ms1-and-ms2), so when `load_groups_ms1()` and `load_groups_ms2()` are applied without MS1 and MS2 spectra for features, the previously loaded settings are used to reload the spectra from features before averaging spectra within feature groups. Once the MS1 and MS2 spectra are loaded, the get and plot methods use by default the loaded spectra. Below the plotting methods are demonstrated without adding averaging arguments.

```{r get-groups-ms1-01}
msTargets$plot_groups_ms1(mz = targets1)
```

```{r get-groups-ms2-01}
msTargets$plot_groups_ms2(mz = targets1)
```

## Remove groups MS1 and MS2

Also for feature groups, the methods `remove_groups_ms1()` and `remove_groups_ms2()` can be used to clear the MS1 and MS2 spectra from the groups `data.table` in the MassSpecData class object, as shown below.

```{r remove-groups-ms1-ms2}
msTargets$remove_groups_ms1()
msTargets$has_loaded_groups_ms1()

msTargets$remove_groups_ms2()
msTargets$has_loaded_groups_ms2()
```

# Recursive integration

Not yet implemented.

# Filters

Not yet implemented.

# Outlook

The StreamFind R package is experimental. This article showed the preliminary framework for mass spectrometry. This article will be updated with future developments for MS data processing. Please [contact us](mailto:cunha@iuta.de) for questions, suggestions, issues or collaboration.

```{r clean-up, include=FALSE}
patRoon::clearCache()
```

### 2024-03-08 R/generics.R

@@ -85,3 +85,27 @@ export <- function(x, ...) {
 .s3_deconvolute_spectra_charges <- function(settings, self) {
   UseMethod(".s3_deconvolute_spectra_charges")
 }
+
+.s3_average_spectra <- function(settings, self) {
+  UseMethod(".s3_average_spectra")
+}
+
+.s3_subtract_blank_spectra <- function(settings, self) {
+  UseMethod(".s3_subtract_blank_spectra")
+}
+
+.s3_subtract_spectra_section <- function(settings, self) {
+  UseMethod(".s3_subtract_spectra_section")
+}
+
+.s3_delete_spectra_section <- function(settings, self) {
+  UseMethod(".s3_delete_spectra_section")
+}
+
+.s3_correct_spectra_baseline <- function(settings, self) {
+  UseMethod(".s3_correct_spectra_baseline")
+}
+
+.s3_smooth_spectra <- function(settings, self) {
+  UseMethod(".s3_smooth_spectra")
+}

### 2024-03-11 R/class_R6_MassSpecEngine.R
Updated Processing Methods Dispatch

@@ -6,7 +6,7 @@
 #'
 #' @template arg-ms-files
 #' @template arg-headers
-#' @template arg-modules
+#' @template arg-results
 #' @template arg-settings
 #' @template arg-runParallel
 #' @template arg-analyses
@@ -89,12 +89,47 @@ MassSpecEngine <- R6::R6Class("MassSpecEngine",
     
     ## ___ .utils -----
     
-    # Removes patRoon data from the modules private field.
+    # Removes patRoon data from the results private field.
     #
     .remove_patRoon_data = function() {
-      if (self$has_modules_data("patRoon")) {
+      if (self$has_results("patRoon")) {
         warning("patRoon data removed! Redo the data processing workflow to regenarate the data.")
-        private$.modules[["patRoon"]] <- NULL
+        private$.results[["patRoon"]] <- NULL
+      }
+    },
+    
+    # Adds an extra column to features.
+    #
+    .add_features_column = function(name = NULL, data = NULL) {
+      
+      if (any(self$has_features())) {
+        
+        all_features <- self$feature_list
+        
+        feature_list <- self$features@features
+        
+        feature_list <- Map(function(x, y, z) {
+          y_to_add <- y[!z$filtered]
+          if (nrow(x) == length(y_to_add)) x[[name]] <- y_to_add
+          x
+        }, feature_list, data, all_features)
+        
+        self$feature_list <- feature_list
+        
+        filtered_feature_list <- self$filtered_features
+        
+        filtered_feature_list <- Map(function(x, y, z) {
+          y_to_add <- y[z$filtered]
+          if (length(y_to_add) > 0) if (nrow(x) == length(y_to_add)) x[[name]] <- y_to_add
+          x
+        }, filtered_feature_list, data, all_features)
+        
+        self$filtered_features <- filtered_feature_list
+        
+        TRUE
+        
+      } else {
+        FALSE
       }
     }
   ),
@@ -103,14 +138,12 @@ MassSpecEngine <- R6::R6Class("MassSpecEngine",
   
   active = list(
     
-    ### ___ NTS -----
-    
     #' @field analysisInfo The analysisInfo data.frame with the list of analyses.
     #'
     analysisInfo = function() {
       
-      if (self$has_modules_data("patRoon")) {
-        private$.modules$patRoon$data@analysisInfo
+      if (self$has_results("patRoon")) {
+        private$.results$patRoon$data@analysisInfo
         
[…]

### 2024-03-11 R/methods_S3_ms_filter_features.R

@@ -5,7 +5,12 @@
 #'
 #' @noRd
 #'
-.s3_ms_filter_features.Settings_filter_features_StreamFind <- function(settings, self) {
+.s3_ms_filter_features.Settings_filter_features_StreamFind <- function(settings, self, private) {
+  
+  if (!self$has_features()) {
+    warning("There are no features! Run find_features first!")
+    return(FALSE)
+  }
   
   parameters <- settings$parameters
   
@@ -605,15 +610,15 @@
 #'
 #' @noRd
 #'
-.s3_ms_filter_features.Settings_filter_features_patRoon <- function(settings, self) {
+.s3_ms_filter_features.Settings_filter_features_patRoon <- function(settings, self, private) {
   
   if (FALSE & requireNamespace("patRoon", quietly = TRUE)) {
     warning("patRoon package not found! Install it for finding features.")
     return(FALSE)
   }
   
-  if (self$has_modules_data("patRoon")) {
-    module_pat <- self$get_modules_data("patRoon")[["patRoon"]]
+  if (self$has_results("patRoon")) {
+    module_pat <- self$get_results("patRoon")[["patRoon"]]
     
   } else {
     warning("No features found! Not done.")

### 2024-03-11 R/methods_S3_ms_find_internal_standards.R

@@ -5,7 +5,12 @@
 #'
 #' @noRd
 #'
-.s3_ms_find_internal_standards.Settings_find_internal_standards_StreamFind <- function(settings, self) {
+.s3_ms_find_internal_standards.Settings_find_internal_standards_StreamFind <- function(settings, self, private) {
+  
+  if (!any(self$has_features())) {
+    warning("There are no features! Run find_features first!")
+    return(FALSE)
+  }
   
   if (!validate(settings)) return(FALSE)
   
@@ -132,7 +137,7 @@
     
     names(istd_col) <- names(features)
     
-    if (!is.logical(self$add_features_column("istd", istd_col, features))) {
+    if (private$.add_features_column("istd", istd_col)) {
       
       message("\U2713 ", length(unique(internal_standards$name)), " internal standards found and tagged!")
       

### 2024-03-11 dev/dev_deconvolution.R

@@ -1,0 +1,87 @@
+
+wd2 <- "C:/Users/apoli/Documents/Dev_230830_Bevacizumab_Avastin_LotB8703H40_Raman_HRMS"
+files <- list.files(paste0(wd2, "/HRMS_2.5mgmL"), pattern = "mzML", full.names = TRUE)
+
+# files <- StreamFindData::get_ms_file_paths()[29]
+
+ms <- MassSpecEngine$new(files)
+
+ps <- Settings_integrate_chromatograms_StreamFind(
+  chromatograms = 0,
+  smoothing = TRUE,
+  windowSize = 10,
+  baseline = TRUE,
+  baseline_method = "als",
+  baseline_args = list(lambda = 6, p = 0.02, maxit = 10),
+  merge = TRUE,
+  closeByThreshold = 2,
+  valeyThreshold = 0.5,
+  minPeakHeight = 20000,
+  minPeakDistance = 2,
+  minPeakWidth = 5,
+  maxPeakWidth = 120,
+  minSN = 10
+)
+# #
+ms$integrate_chromatograms(ps)
+
+ms$chromatograms_peaks
+
+ms$plot_chromatograms_peaks(colorBy = "targets+analyses")
+
+ms$deconvolute_spectra_charges(Settings_deconvolute_spectra_charges_StreamFind())
+
+ms$spectra_charges
+
+ms$plot_spectra_charges()
+
+ms$averaged_spectra
+
+mean(ms$spectra_peaks$mass)
+
+sd(ms$spectra_peaks$mass)
+
+ms$plot_spectra_peaks()
+
+ms$plot_spectra_charges()
+
+ms$chrom_peaks
+
+View(ms$get_results("chromatograms"))
+
+# patRoon::clearCache("all")
+
+
+
+wd3 <- "C:/Users/apoli/Documents/iSoft/Bevacizumab_Zirabev_LotFR7476"
+files2 <- list.files(wd3, pattern = "mzML", full.names = TRUE)
+
+ms2 <- MassSpecEngine$new(files2)
+
+ms2$integrate_chromatograms(ps)
+
+ms2$plot_chrom_peaks(colorBy = "analyses")
+
+ms2$chrom_peaks
+
+
+
+
+ms2$plot_tic()
+
+
+
+ms2$chrom_peaks
+
+
+
+ms2$plot_chrom_peaks(colorBy = "targets+analyses")
+
[…]

### 2024-03-11 dev/dev_template_report.Rmd

@@
 ---
+<style>
+.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
+  background-color: #1E8449;
+}
+</style>
 
 ```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 9, results = "asis", cache = TRUE, comment = '')
+options(width = 400)
+```
 
-# Libraries
+```{r libraries, include=FALSE}
 library(knitr)
@@
 library(StreamFind)
-
-# Variables
-logo <- NA_character_
-
-# Global options
-knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 9, results = "asis")
-
-# Working directory
-path_wd <- getwd()
-
-# Add logo file
-path_extdata <- system.file(package = "StreamFindData", dir = "extdata")
-logo <- file.path(path_extdata, "logo_iuta.png") #207 x 216 px
 ```
@@
 ```{r logo, echo=FALSE}
-if (!is.na(logo)) {
-  htmltools::img(
-  src = knitr::image_uri(logo),
-  alt = "logo",
-  style =
-    paste0("
-      margin-left:inherit;
-      position:absolute;
-      top:60px;
-      left:740px;
-      width:", 100, "px;
-      height:", 100*(216/207), "px;"
-    )
-  ) 
-}
+path_extdata <- system.file(package = "StreamFindData", dir = "extdata")
+logo <- file.path(path_extdata, "logo_StreamFind.png")
+htmltools::img(src = knitr::image_uri(logo), style = "margin-left:inherit;position:absolute;top:100px;left:740px;width:200px;")
 ```
@@
 
-Code is automaticelly embedded unless `echo` chunk option is set to `FALSE`.
+Code is automatically embedded unless `echo` chunk option is set to `FALSE`.
 Evaluation of code in a given chunk can be skiped by setting `eval` to `FALSE`.
-Code chunks should always be unically named.
+Each code chunk should have a unique named.
 
@@
 
-<br> <br>
+<br>
 
@@
 
-The report was produced on `r Sys.time()`. The data and report are stored in ``r path_wd`` of the workstation *`r Sys.info()["nodename"]`* under the user *`r Sys.info()["user"]`*.
+The report was produced on `r Sys.time()`. The data and report are stored in `r getwd()` of the workstation *`r Sys.info()["nodename"]`* under the user *`r Sys.info()["user"]`*.
 
 ***
-
-<br>
-
-<style>
-body, p {
-  color: black;
-  font-family: Arial;
-  text-align: justify;
-  font-size: 12pt;
-}
-h1{
-  font-size: 20pt;
-  font-style: bold;
-}
-h2{
-  font-size:18pt;
-  font-style: bold;
-}
-h3{
-  font-size:14pt;
-  font-style: bold;
-}
-.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
-    background-color: #239B56;
-}
-caption {
-  color: black;
-  font-size: 1.0em;
-}
-</style>

### 2024-03-15 R/methods_S3_subtract_spectra_section.R
Further in Raman and article data fusion started

@@ -69,11 +69,14 @@
               
               res <- res[res[[sectionVal]] < sectionWindow[1] | res[[sectionVal]] > sectionWindow[2], ]
               
-              if (nrow(res) > 0) {
-                
-                for (i in unique(res$rt)) res$intensity[res$rt == i] <- res$intensity[res$rt == i] - cutSec$intensity
-                
-              }
+              res_list <- split(res, res$rt)
+              
+              res_list <- lapply(res_list, function(z, cutSec) {
+                z$intensity <- z$intensity - cutSec$intensity
+                z
+              }, cutSec = cutSec)
+              
+              res <- rbindlist(res_list)
             }
           }
         }

### 2024-03-19 dev/dev_raman.R
Added boost for methods to process raman and ms data

@@ -15,15 +15,13 @@ ant_files <- list.files(ant_dir, pattern = "_\\d.asc", full.names = TRUE)
 
 ant <- RamanEngine$new(ant_files, runParallel = FALSE)
 
-# ant$add_replicate_names(sub("_\\d+$", "", e_sec$get_analysis_names()))
-ant$add_replicate_names(sub("_(\\d)\\d*$", "_\\1", ant$get_analysis_names()))
-ant$merge_spectra_time_series()
+ant$add_replicate_names(sub("_\\d+$", "", ant$get_analysis_names()))
+# ant$add_replicate_names(sub("_(\\d)\\d*$", "_\\1", ant$get_analysis_names()))
+# ant$merge_spectra_time_series()
 
 ant$bin_spectra(Settings_bin_spectra_StreamFind(windowSpectrumUnits = 5))
 
-# ant$plot_chromatograms()
-
-ant$normalize_spectra(Settings_normalize_spectra_StreamFind(liftTozero = TRUE)) # TODO
+ant$normalize_spectra(Settings_normalize_spectra_StreamFind(liftTozero = TRUE))
 
 ant$subtract_spectra_section(Settings_subtract_spectra_section_StreamFind(sectionWindow = c(0, 3)))
 
@@ -33,13 +31,29 @@ ant$delete_spectra_section(Settings_delete_spectra_section_StreamFind(list("shif
 
 ant$delete_spectra_section(Settings_delete_spectra_section_StreamFind(list("shift" = c(2000, 2600))))
 
-# ant$delete_spectra_section(Settings_delete_spectra_section_StreamFind(list("rt" = c(10, max(e_sec$get_spectra()$rt)))))
-
-ant$correct_spectra_baseline(Settings_correct_spectra_baseline_StreamFind(
-  method = "als", args = list(lambda = 3, p = 0.02, maxit = 20)
-))
+ant$correct_spectra_baseline(Settings_correct_spectra_baseline_baseline(method = "als", args = list(lambda = 3, p = 0.02, maxit = 20)))
 
 ant$normalize_spectra(Settings_normalize_spectra_StreamFind(liftTozero = TRUE))
+
+ant$plot_spectra(rt = c(5, 8))
+
+
+View(ant$get_results("spectra"))
+
+
+# ant$plot_chromatograms()
+
+ # TODO
+
+
+
+
+
[…]

### 2024-03-21 R/methods_S3_normalize_spectra.R
Added other normalization options

@@ -160,3 +160,142 @@
   
   TRUE
 }
+
+#' @title .s3_normalize_spectra.Settings_normalize_spectra_scale
+#'
+#' @description Scales spectra based on the standard deviation.
+#'
+#' @noRd
+#'
+.s3_normalize_spectra.Settings_normalize_spectra_scale <- function(settings, self, private) {
+  
+  if (!self$has_spectra()) {
+    warning("Spectra not found! Not done.")
+    return(FALSE)
+  }
+
+  spec_list <- self$spectra
+  
+  spec_list <- lapply(spec_list, function(x) {
+    
+    if (nrow(x) > 0) {
+      
+      if ("rt" %in% colnames(x)) {
+        temp_x <- split(x, x$rt)
+        
+        temp_x <- lapply(temp_x, function(z) {
+          sd_int <- sd(z$intensity)
+          z$intensity <- z$intensity / sd_int
+          z
+        })
+        
+        x <- rbindlist(temp_x)
+        
+      } else {
+        sd_int <- sd(x$intensity)
+        x$intensity <- x$intensity / sd_int
+      }
+    }
+    
+    x
+    
+  })
+  
+  self$spectra <- spec_list
+  
+  message(paste0("\U2713 ", "Spectra normalized!"))
+  
+  TRUE
+}
+
+#' @title .s3_normalize_spectra.Settings_normalize_spectra_blockweight
+#'
+#' @description Scales spectra based on the standard deviation.
+#'
+#' @noRd
+#'
+.s3_normalize_spectra.Settings_normalize_spectra_blockweight <- function(settings, self, private) {
+  
+  if (!self$has_spectra()) {
+    warning("Spectra not found! Not done.")
+    return(FALSE)
+  }
+  
+  spec_list <- self$spectra
+  
+  spec_list <- lapply(spec_list, function(x) {
+    
+    if (nrow(x) > 0) {
+      
+      if ("rt" %in% colnames(x)) {
+        temp_x <- split(x, x$rt)
+        
+        temp_x <- lapply(temp_x, function(z) {
+          z$intensity <- z$intensity / sqrt(length(z$intensity))
+          z
+        })
+        
+        x <- rbindlist(temp_x)
+        
+      } else {
+        x$intensity <- x$intensity / sqrt(length(x$intensity))
+      }
+    }
+    
+    x
+    
+  })
+  
+  self$spectra <- spec_list
+  
[…]

### 2024-04-26 dev/dev_ideas/ms_overview.Rmd
Major Structuring of Spectra Interface (not complete yet)

@@
 All plot methods use data that can be obtained with get methods, meaning that the user is able to re-plot the data with a different approach/style.
-For instance, the `get_tic()` method can be used to obtain the total ion chromatogram (TIC) as a `data.table` with columns *analysis*, *rt* and *intensity*, as shown below for the first analysis.
+For instance, the `get_spectra_tic()` method can be used to obtain the total ion chromatogram (TIC) as a `data.table` with columns *analysis*, *rt* and *intensity*, as shown below for the first analysis.
 The same TIC data can be directly plotted using the `plot_tic()` method, as shown below.
@@
 ```{r get-tic-first-analysis}
-head(ms$get_tic(analyses = 1))
+head(ms$get_spectra_tic(analyses = 1))
 ```
@@
 # getter for polarities
-ms$get_polarities(analyses = "00_hrms_s_is_pos_cent-r003")
+ms$get_spectra_polarity(analyses = "00_hrms_s_is_pos_cent-r003")
 
@@
 # getter for MS levels. Note that it returns a list not a vector!
-ms$get_spectra_levels()
+ms$get_spectra_level()
 
@@
 # getter for TIC
-head(ms$get_tic(analyses = 2))
+head(ms$get_spectra_tic(analyses = 2))
 
 # getter for BPC
-head(ms$get_bpc())
+head(ms$get_spectra_bpc())
 ```
@@
 
-Besides the `get_tic()` and `get_bpc()` methods (shown above).
-There are four methods to get low level spectra from analyses: `get_spectra()`, `get_eic()`, `get_ms1()` and `get_ms2()`.
+Besides the `get_spectra_tic()` and `get_spectra_bpc()` methods (shown above).
+There are four methods to get low level spectra from analyses: `get_spectra()`, `get_spectra_eic()`, `get_spectra_ms1()` and `get_spectra_ms2()`.
 All methods use the same arguments as the function `make_ms_targets()` to build MS targets and the *analyses* argument to specify which MassSpecAnalysis objects to query the data.
@@
 
-The extracted ion chromatogram (EIC) can be obtained with `get_eic()` and plotted with `plot_eic()` method.
+The extracted ion chromatogram (EIC) can be obtained with `get_spectra_eic()` and plotted with `plot_eic()` method.
 The interface is very similar as the previous methods, with analyses and make targets arguments used to specify each analyses to parse the data and which traces to retrieve.
@@
 ```{r get-eic-01}
-ms$get_eic(analyses = 1, mz = mz1, rt = rt1, ppm = 5, sec = 10)
+ms$get_spectra_eic(analyses = 1, mz = mz1, rt = rt1, ppm = 5, sec = 10)
 ```
@@
 
-While the EIC method collects and sums the traces in each retention time value (i.e., in each scan), the `get_ms1()` and `get_ms2()` methods average the mass traces in a given retention time interval, returning an averaged spectrum representing several scans.
+While the EIC method collects and sums the traces in each retention time value (i.e., in each scan), the `get_spectra_ms1()` and `get_spectra_ms2()` methods average the mass traces in a given retention time interval, returning an averaged spectrum representing several scans.
 The averaging is based on clustering the mass traces according to a given `mzClust` argument as demonstrated below for both MS1 and MS2 data.
@@
 # MS1 averaged spectrum of targets from 247 to 257 (Carbamazepin-d10) and 239 to 245 (Diuron-d6) m/z and +/- 1 second the retention time of each target
-ms$get_ms1(analyses = 3, mzClust = 0.005,
+ms$get_spectra_ms1(analyses = 3, mzClust = 0.005,
   mz = data.frame("mzmin" = c(247, 239), "mzmax" = c(257, 245)),
@@
 # MS2 data of Diuron-d6 for analysis 1
-ms$get_ms2(analyses = 1, mz = targets1[2, ])
+ms$get_spectra_ms2(analyses = 1, mz = targets1[2, ])
 ```

### 2024-04-26 src/StreamCraft/src/StreamCraft_utils.hpp

@@ -9,6 +9,75 @@ namespace sc {
 
   inline namespace utils {
 
+    class VIRTUAL_MS_SPECTRUM {
+      public:
+        virtual ~VIRTUAL_MS_SPECTRUM() = default;
+        virtual int extract_spec_index() = 0;
+        virtual std::string extract_spec_id() const = 0;
+        virtual int extract_spec_scan() const = 0;
+        virtual int extract_spec_array_length() const = 0;
+        virtual int extract_spec_level() const = 0;
+        virtual std::string extract_spec_mode() const = 0;
+        virtual std::string extract_spec_polarity() const = 0;
+        virtual double extract_spec_lowmz() const = 0;
+        virtual double extract_spec_highmz() const = 0;
+        virtual double extract_spec_bpmz() const = 0;
+        virtual double extract_spec_bpint() const = 0;
+        virtual double extract_spec_tic() const = 0;
+        virtual double extract_scan_rt() const = 0;
+        virtual double extract_ion_mz() const = 0;
+        virtual double extract_activation_ce() const = 0;
+        virtual bool has_precursor() const = 0;
+        // virtual std::vector<std::vector<double>> extract_binary_data(const MZXML_BINARY_METADATA&  mtd) const = 0;
+    };
+
+    template <typename T>
+    class MS_SPECTRUM : public VIRTUAL_MS_SPECTRUM {
+      public:
+        MS_SPECTRUM(T spec_in) : spec(spec_in) {}
+        int extract_spec_index() override { return spec.extract_spec_index(); }
+        std::string extract_spec_id() const override { return spec.extract_spec_id(); }
+        int extract_spec_scan() const override { return spec.extract_spec_scan(); }
+        int extract_spec_array_length() const override { return spec.extract_spec_array_length(); }
+        int extract_spec_level() const override { return spec.extract_spec_level(); }
+        std::string extract_spec_mode() const override { return spec.extract_spec_mode(); }
+        std::string extract_spec_polarity() const override { return spec.extract_spec_polarity(); }
+        double extract_spec_lowmz() const override { return spec.extract_spec_lowmz(); }
+        double extract_spec_highmz() const override { return spec.extract_spec_highmz(); }
+        double extract_spec_bpmz() const override { return spec.extract_spec_bpmz(); }
+        double extract_spec_bpint() const override { return spec.extract_spec_bpint(); }
+        double extract_spec_tic() const override { return spec.extract_spec_tic(); }
[…]

### 2024-04-29 dev/dev_ideas/ms_overview.Rmd
Major update of MassSpecEngine Semantic

@@
 For instance, the `get_spectra_tic()` method can be used to obtain the total ion chromatogram (TIC) as a `data.table` with columns *analysis*, *rt* and *intensity*, as shown below for the first analysis.
-The same TIC data can be directly plotted using the `plot_tic()` method, as shown below.
+The same TIC data can be directly plotted using the `plot_spectra_tic()` method, as shown below.
 
@@
 ```{r plot-tic} 
-ms$plot_tic(analyses = 1)
+ms$plot_spectra_tic(analyses = 1)
 ```
@@
 ```{r plot-bpc-static}
-ms$plot_bpc(interactive = FALSE)
+ms$plot_spectra_bpc(interactive = FALSE)
 ```
@@
 
-Another plotting option for raw spectra is the `plot_xic()` method as shown below.
+Another plotting option for raw spectra is the `plot_spectra_xic()` method as shown below.
 This method returns a two row plot with traces intensity vs retention time on top and *m/z* vs retention time below.
@@
 ```{r plot-xic}
-ms$plot_xic(analyses = 1:2, mz = targets1[2,],
+ms$plot_spectra_xic(analyses = 1:2, mz = targets1[2,],
   targetsMark = targets1[2, c("rt", "mz")], ppmMark = 5, secMark = 10)
@@
 
-The extracted ion chromatogram (EIC) can be obtained with `get_spectra_eic()` and plotted with `plot_eic()` method.
+The extracted ion chromatogram (EIC) can be obtained with `get_spectra_eic()` and plotted with `plot_spectra_eic()` method.
 The interface is very similar as the previous methods, with analyses and make targets arguments used to specify each analyses to parse the data and which traces to retrieve.
@@
 ```{r plot-eic-01}
-ms$plot_eic(mz = targets1, interactive = FALSE)
+ms$plot_spectra_eic(mz = targets1, interactive = FALSE)
 ```
@@
 ```{r plot-eic-02}
-ms$plot_eic(mz = targets1, colorBy = "analyses")
+ms$plot_spectra_eic(mz = targets1, colorBy = "analyses")
 ```
@@
 # In the plot, the Carbamazepin-d10 and Diuron-d6 isotopes can be distinguished
-ms$plot_ms1(analyses = 3,
+ms$plot_spectra_ms1(analyses = 3,
   mz = data.frame("mzmin" = c(247, 239), "mzmax" = c(257, 245)),
@@
 # MS2 data plot of Diuron-d6 in all analyses
-ms$plot_ms2(mz = targets1[2 ,], 
+ms$plot_spectra_ms2(mz = targets1[2 ,], 
   mzClust = 0.005, minIntensity = 50,
@@
 # Note that "minIntensity" argument was increased to 500 counts, resulting in a cleaner averaged spectrum
-ms$plot_ms2(analyses = 3, mz = targets1, 
+ms$plot_spectra_ms2(analyses = 3, mz = targets1, 
   mzClust = 0.005, minIntensity = 500,

### 2024-04-30 dev/dev_statistics.Rmd
Updated settings documentation

---
title: 'Fusion of Mass Spectrometry and Raman Spectroscopy Data'
subtitle: |
  | Comparison of different bovine serum albumin formulations
  | <br>
author:
  name: Ricardo Cunha
  email: cunha@iuta.de
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    fig_caption: true
    toc: true
    number_sections: true
    toc_float: true
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Fusion of Mass Spectrometry and Raman Spectroscopy Data}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(plotly)
library(data.table)
library(StreamFind)

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center", fig.width = 9, results = "asis", comment = "")

main_dir <- "C:/Users/apoli/Documents/iSoft/BSA_products"

all_files <- list.files(main_dir, full.names = TRUE)

ms_files_d <- all_files[grepl(".d", fixed = T, all_files)][1:13]

ms_files <- all_files[grepl(".mzML", fixed = T, all_files)]

raman_dirs <- all_files[grepl("_Raman", fixed = TRUE, all_files)]

raman_files <- character()

for (i in raman_dirs) raman_files <- c(raman_files, list.files(i, full.names = TRUE))

raman_unified_files <- list.files(main_dir, pattern = "asc", full.names = TRUE)
```

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<br>

<br>

***

# Introduction

Mass Spectrometry (MS) and Raman spectroscopy (Raman) are orthogonal techniques. While MS measures the chemical composition of a given sample, Raman is able to provide structural information of the constituents. This orthogonality is interesting when applying quality evaluation workflows, where both chemical composition and structural integrity are relevant. In this article, we show how MS and Raman can be pre-processed and fused for a combined statistical evaluation to support quality control. As test sets, we use four different bovine serum albumin (BSA) products.

# Pre-processing

The preliminary processing is applied differently for MS and Raman as the raw data structure does not allow direct fusion of the data. Therefore, in the following two sub-chapters we describe the pre-processing workflow steps for MS and Raman data.

## MS

The four different BSA products were measured in triplicate in full scan MS mode. Giving the following set of *.d* files.

```{r}
# Raw data from an Agilent QTOF
basename(ms_files_d)
```

The first step is conversion to open source *mzML* format using *MSConvert* from [ProteoWizard](https://proteowizard.sourceforge.io/index.html) CLI, as shwon below.

```{r eval=FALSE}
# Add option to centroid MS1 data before conversion
optList <- list(filter = "peakPicking vendor msLevel=1")

# convert_ms_files(ms_files_d, outputFormat = "mzML", optList = optList)
```

After conversion, the *mzML* files are created in the same directory. The converted files are listed below.

```{r}
basename(ms_files)
```

With the *mzML* files, a new *MassSpecEngine* is created to start the MS pre-processing project. Below the engine is created and the analysis replicate names are modified to properly assign the analyses within each replicate measurement.

```{r}
ms <- MassSpecEngine$new(files = ms_files, headers = list(name = "BSA Quality Evaluation Project"))

# Changes the analysis replicate names
ms$add_replicate_names(sub("_\\d+$", "", ms$get_analysis_names()))

# Overview
ms
```

```{r, fig.cap="Total ion chromatograms for each analysis replicate."}
ms$plot_spectra_tic(colorBy = "replicates")
```

The data processing workflow is based on applied processing methods within the engine. As pre-processing, the following steps are applied:

1. Find the retention time elution of BSA based on the total ion chromatogram (TIC) of each analysis; 
2. Find the mass-to-charge (*m/z*) range of BSA for deconvolution of charges; 
3. Deconvolute charges for estimation of the exact mass, in DA;

The individual processing steps are applied according to *ProcessingSettings*. Below, the integration of the TIC is shown using the dedicated *ProcessingSettings* after loading the TIC chromatograms from the analysis files and smoothing using a moving average approach.

```{r}
ms$load_chromatograms("TIC")
```

```{r}
# Smoothing of the TIC chromatogram for improving peak finding
ms$smooth_chromatograms(Settings_smooth_chromatograms_movingaverage(windowSize = 5))
```

```{r}
# Settings for integration of the TIC
ps_ic <- Settings_integrate_chromatograms_StreamFind(
  merge = TRUE,
  closeByThreshold = 5,
  minPeakHeight = 3E6,
  minPeakDistance = 3,
  minPeakWidth = 20,
  maxPeakWidth = 200,
  minSN = 5
)

ps_ic
```

```{r}
ms$integrate_chromatograms(ps_ic)
```

```{r ,fig.cap="TIC with integrated peaks."}
ms$plot_chromatograms_peaks(colorBy = "replicates")
```

The average retention time of the main peak was `r mean(ms$chromatograms_peaks$rt)` seconds with a standard deviation of `r sd(ms$chromatograms_peaks$rt)`. The following step is to estimate the *m/z* range of BSA for deconvolution of charges. For this, the MS1 spectra of the peak retention time +/- 0.1 seconds is plotted.

```{r, fig.cap="Spectra of BSA for each analysis replicate at the peak maximum."}
ms$plot_spectra_ms1(rt = mean(rbindlist(ms$chromatograms_peaks)$rt), sec = 0.5,
  colorBy = "replicates", presence = 0.1, minIntensity = 1000, interactive = FALSE
)
```

According to the MS1 spectra, the *m/z* from 1200 to 2000 can be added to the dedicated *ProcessingSettings* for deconvolution, as shown below. The first step is to load from the mzML files the traces within the retention time and *m/z* ranges selected.

```{r}
av_rt <- mean(rbindlist(ms$chromatograms_peaks)$rt)

spectra_ranges <- data.table("mzmin" = 1200, "mzmax" = 2000, "rtmin" = av_rt - 1, "rtmax" = av_rt + 1)

# ms <- MassSpecEngine$new(files = ms_files, headers = list(name = "BSA Quality Evaluation Project"))
# 
# ms$add_replicate_names(sub("_\\d+$", "", ms$get_analysis_names()))

ms$load_spectra(mz = spectra_ranges)

# ms$cluster_spectra(Settings_cluster_spectra_StreamFind(val = "mz", clustVal = 0.01, presence = 0.2))
# ms$calculate_spectra_charges(Settings_calculate_spectra_charges_StreamFind(roundVal = 15, relLowCut = 0.2, absLowCut = 8000))
# ms$deconvolute_spectra(Settings_deconvolute_spectra_StreamFind())
# ms$smooth_spectra(Settings_smooth_spectra_movingaverage(windowSize = 15))
# ms$normalize_spectra(Settings_normalize_spectra_minmax()) #liftTozero = TRUE
# ms$normalize_spectra(Settings_normalize_spectra_blockweight())
# ms$bin_spectra(Settings_bin_spectra_StreamFind(bins = list("rt" = 2, "mass" = 25), refBinAnalysis = 1))

# View(ms$spectra)
# ms$plot_spectra()
# ms$spectra_charges
# ms$plot_spectra_charges()
# ms$correct_spectra_baseline(Settings_correct_spectra_baseline_baseline())
# ms$plot_deconvoluted_spectra()
# ms$plot_spectra(interactive = T)
# ms$plot_spectra_charges()
# ms$spectra_charges
# ms$deconvoluted_spectra
# ms$spectra_charges
# ms$plot_spectra_charges()
# ms$get_spectra()
# ms$spectra
# ms$has_results("spectra")
# ms$spectra
# ms$plot_spectra(interactive = T)
# View(ms$get_results("spectra"))
```

Below we create an ordered list of *ProcessingSettings* objects for each processing method to be applied and add the list to the engine. The exact parameters for each method can be optimized individually by calling the specific method directly. Once the settings are added, the workflow can be inspected with the `print_workflow()` method.

```{r}
ms_workflow_settings <- list(
  Settings_cluster_spectra_StreamFind(val = "mz", clustVal = 0.01, presence = 0.2),
  Settings_calculate_spectra_charges_StreamFind(roundVal = 15, relLowCut = 0.2, absLowCut = 8000),
  Settings_deconvolute_spectra_StreamFind(),
  Settings_smooth_spectra_movingaverage(windowSize = 15),
  Settings_bin_spectra_StreamFind(bins = list("rt" = 5, "mass" = 20), refBinAnalysis = 1),
  Settings_average_spectra_StreamFind(),
  Settings_normalize_spectra_minmax(),
  Settings_normalize_spectra_blockweight()
)

# Remove the settings added to integrate the TIC chromatograms
ms$remove_settings()

ms$add_settings(ms_workflow_settings)

ms$print_workflow()
```

The `run_workflow()` method is called to run the data processing. The workflow will be applied in the order the methods were added to the engine.

```{r}
ms$run_workflow()
```

```{r ,fig.cap="Deconvoluted and pre-processed spectra for each analysis."}
ms$plot_spectra(colorBy = "analyses")
```

## Raman

The four different BSA products were measured using an online coupling of size exclusion chromatography (SEC) with capillary-enhanced Raman spectroscopy (CERS) through a liquid-core optical fibre flow cell, as reported by Thissen et al. (DOI: 10.1021/acs.analchem.3c03991). The raw spectra was converted to *asc* format for each retention time. A subset of the background and the main BSA elution peak was selected for pre-processing. The first pre-processing step is to merge the individual *asc* files into a main file for each BSA product, as shown below.

```{r eval=FALSE}
# Engine with all asc files for each spectrum and each BSA product
#raman <- RamanEngine$new(raman_files, runParallel = FALSE)

# Folder names of each BSA product
#base_dirs <- basename(dirname(raman_files))

# The folder name is used as replicate name, which is also used as file name for the unified data file
#raman$add_replicate_names(base_dirs)

# The spectra are merged into a time series for each BSA product
#raman$merge_spectra_time_series()
```

A new engine is created with the unified file for each BSA product and then plotted as a time series.

```{r}
raman <- RamanEngine$new(raman_unified_files, runParallel = FALSE)

# Modifying the replicate names to match the MS analyses
raman$add_replicate_names(unique(ms$get_replicate_names())[1:4])

raman
```

```{r ,fig.cap="Time series chromatogram for each BSA product. Each data point is a Raman spectrum at a specific retention time."}
raman$plot_chromatograms(yLab = "Abbundance sum for each spectrum", colorBy = "analyses")
```

```{r ,fig.cap="Raw averaged Raman spectra for each BSA product between minute 5.5 and 8."}
raman$plot_spectra(rt = c(5.5, 8), colorBy = "analyses")
```

As for MS data, the workflow for Raman data is also assembled using an ordered list of processing settings, as shown below.

```{r}
raman_workflow_settings <- list(
  Settings_bin_spectra_StreamFind(unitsVal = "rt", unitsNumber = 3),
  Settings_normalize_spectra_minmax(),
  Settings_subtract_spectra_section_StreamFind(sectionWindow = c(0, 3)),
  Settings_smooth_spectra_savgol(fl = 21, forder = 4, dorder = 0),
  Settings_delete_spectra_section_StreamFind(list("shift" = c(-100, 330))),
  Settings_delete_spectra_section_StreamFind(list("shift" = c(1800, 2600))),
  Settings_correct_spectra_baseline_airpls(lambda = 45, differences = 1, itermax = 30),
  Settings_delete_spectra_section_StreamFind(list("rt" = c(0, 5.5))),
  Settings_delete_spectra_section_StreamFind(list("rt" = c(8, 11))),
  Settings_average_spectra_StreamFind(collapseTime = TRUE),
  Settings_normalize_spectra_minmax(),
  Settings_normalize_spectra_blockweight()
)

raman$add_settings(raman_workflow_settings)

raman$print_workflow()
```

A particularity of the workflow for Raman data is that the background subtraction (workflow step number 3) is performed from a time region (i.e., between 0 and 3 minutes) before the main BSA peak, which is from approximately from 5 to 8 minutes. The workflow is then applied using the `run_workflow()` method.

```{r}
raman$run_workflow()
```

```{r ,fig.cap="Processed Raman spectra for each BSA product."}
raman$plot_spectra()
```

# Fusion

The fusion of MS and Raman data is performed by combining the processed data from both engines. The data is then merged into a single data frame for statistical evaluation using a MachineLearningEngine, which is under development.

```{r}
rSpec <- raman$get_spectra()
mrSpec <- matrix(rSpec$intensity, nrow = length(unique(rSpec$replicate)), ncol = length(unique(rSpec$shift)), byrow = TRUE, dimnames = list(as.character(unique(rSpec$replicate)), as.character(unique(rSpec$shift))))

mSpec <- ms$get_spectra()
mmSpec <- matrix(mSpec$intensity, nrow = length(unique(mSpec$replicate)), ncol = length(unique(mSpec$bins)), byrow = TRUE, dimnames = list(unique(mSpec$replicate), unique(mSpec$bins)))

fusedMat <- cbind(mmSpec, mrSpec)

attr(fusedMat, "xaxis.name") = "Keys"
attr(fusedMat, "xaxis.values") = seq_len(ncol(fusedMat))
```

```{r}
cl <- StreamFind:::.get_colors(rownames(fusedMat))
fig <- plot_ly()
xVal <- seq_len(length(fusedMat[1, ]))
for (i in seq_len(nrow(fusedMat))) {
  fig <- fig %>% add_trace(
    x = xVal,
    y = fusedMat[i, ],
    type = "scatter", mode = "lines",
    line = list(width = 0.5, color = unname(cl[i])),
    name = names(cl)[i],
    legendgroup = names(cl)[i],
    showlegend = TRUE
  )
}
xaxis <- list(
  linecolor = toRGB("black"),
  linewidth = 2, title = "Bins",
  titlefont = list(size = 12, color = "black")
)
yaxis <- list(
  linecolor = toRGB("black"),
  linewidth = 2, title = "Intensity",
  titlefont = list(size = 12, color = "black")
)
fig <- fig %>% plotly::layout(xaxis = xaxis, yaxis = yaxis)
fig
```

# Statistic analysis

To be done!

## mdatools

### PCA

```{r, include=FALSE}
library(mdatools)
data(people)

m = pca(people, 7, scale = TRUE, info = "People PCA model")
m = selectCompNum(m, 5)

print(m$res$cal)

par(mfrow = c(1, 2))
mdaplot(m$res$cal$scores, type = "p", show.labels = TRUE, show.lines = c(0, 0))
mdaplot(m$loadings, type = "p", show.labels = TRUE, show.lines = c(0, 0))
```

```{r}
fusedMat2 <- prep.autoscale(fusedMat, center = FALSE, scale = TRUE)
cl <- StreamFind:::.get_colors(rownames(fusedMat2))
fig <- plot_ly()
xVal <- seq_len(length(fusedMat2[1, ]))
for (i in seq_len(nrow(fusedMat2))) {
  fig <- fig %>% add_trace(
    x = xVal,
    y = fusedMat2[i, ],
    type = "scatter", mode = "lines",
    line = list(width = 0.5, color = unname(cl[i])),
    name = names(cl)[i],
    legendgroup = names(cl)[i],
    showlegend = TRUE
  )
}
xaxis <- list(
  linecolor = toRGB("black"),
  linewidth = 2, title = "Bins",
  titlefont = list(size = 12, color = "black")
)
yaxis <- list(
  linecolor = toRGB("black"),
  linewidth = 2, title = "Intensity",
  titlefont = list(size = 12, color = "black")
)
fig <- fig %>% plotly::layout(xaxis = xaxis, yaxis = yaxis)
fig
```

```{r, include=FALSE}

m = pca(fusedMat, 3, scale = FALSE, info = "People PCA model")
plot(m, show.labels = TRUE)

par(mfrow = c(2, 2))
plotVariance(m, type = "h", show.labels = TRUE)
plotQDoF(m)
plotT2DoF(m)
plotDistDoF(m)

par(mfrow = c(1, 2))
plotQDoF(m, main = "DoF plot (classic)")

m = setDistanceLimits(m, lim.type = "ddrobust")
plotQDoF(m, main = "DoF plot (robust)")

plotExtreme(m, comp = 1:2, main = "Extreme plot (cal)")

plotLoadings(m, 1:2, type = 'l')

# par(mfrow = c(2, 2))
# mdaplot(m$res$cal$scores, type = "p", show.labels = TRUE, show.lines = c(0, 0))
# mdaplot(m$loadings, type = "p", show.labels = TRUE, show.lines = c(0, 0))
# plotResiduals(m$res$cal, show.labels = TRUE)
# plotVariance(m$res$cal, type = "h", show.labels = TRUE, labels = "values")
```

```{r ,echo=FALSE,fig.width=10,fig.height=4}

library(plotly)

loadings_plot <- plot_ly()

loadings_plot <- loadings_plot %>% add_trace(
  x = m$loadings[rownames(loadings) %in% colnames(mrSpec), "Comp 1"],
  y = m$loadings[rownames(loadings) %in% colnames(mrSpec), "Comp 2"],
  type = "scatter",
  mode = "markers+text",
  text = rownames(m$loadings)[(rownames(m$loadings) %in% colnames(mrSpec))],
  textfont = list(size = 14),
  legendgroup = "Raman",
  name = "Raman",
  marker = list(size = 10)
)

loadings_plot <- loadings_plot %>% add_trace(
  x = m$loadings[rownames(m$loadings) %in% colnames(mmSpec), "Comp 1"],
  y = m$loadings[rownames(m$loadings) %in% colnames(mmSpec), "Comp 2"],
  type = "scatter",
  mode = "markers+text",
  text = rownames(m$loadings)[(rownames(m$loadings) %in% colnames(mmSpec))],
  textfont = list(size = 14),
  legendgroup = "MS",
  name = "MS",
  marker = list(size = 10, color = "red")
)

loadings_plot %>% layout(
    title = "PCA Loadings Plot",
    xaxis = list(title = "Principal Component 1"),
    yaxis = list(title = "Principal Component 2")
  )

```

### PLS

```{r}

library(mdatools)
data(people)

idx = seq(4, 32, 4)
Xc = people[-idx, -4]
yc = people[-idx, 4, drop = FALSE]
Xt = people[idx, -4]
yt = people[idx, 4, drop = FALSE]

m = pls(Xc, yc, 7, scale = TRUE, cv = 1, info = "Shoesize prediction model")

par(mfrow = c(2, 2))
plotRegcoeffs(m)
plotRegcoeffs(m, ncomp = 2)
plot(m$coeffs, ncomp = 3, type = "b", show.labels = TRUE)
plot(m$coeffs, ncomp = 2)

par(mfrow = c(1, 2))
plotPredictions(m)
plotPredictions(m, ncomp = 1)

plotVIPScores(m, show.labels = TRUE)

res = predict(m, Xt, yt)
print(res)

data(simdata)
Xc = simdata$spectra.c
yc = simdata$conc.c[, 1]
Xt = simdata$spectra.t
yt = simdata$conc.t[, 1]

m = pls(Xc, yc, 10, x.test = Xt, y.test = yt, ncomp.selcrit = "min")
summary(m)

plot(m)

```

### PLS-DA

```{r}
data(iris)

cal.ind = c(1:25, 51:75, 101:125)
val.ind = c(26:50, 76:100, 126:150)

Xc = iris[cal.ind, 1:4]
Xv = iris[val.ind, 1:4]

cc.all = iris[cal.ind, 5]
cv.all = iris[val.ind, 5]

cc.vir = cc.all == "virginica"
cv.vir = cv.all == "virginica"
show(cc.vir)

m.all = plsda(Xc, cc.all, 3, cv = 1)
m.vir = plsda(Xc, cc.vir, 3, cv = 1, classname = "virginica")

summary(m.all)

par(mfrow = c(1, 2))
plotPredictions(m.all)
plotPredictions(m.vir)

res = predict(m.all, Xv, cv.all)
summary(res)

par(mfrow = c(1, 1))
plotPredictions(res, show.labels = TRUE)

```

### SIMCA
```{r}
idx = seq(1, nrow(iris), by = 2)

# split the values
Xc = iris[idx, 1:4]
cc = iris[idx, 5]

Xt = iris[-idx, 1:4]
ct = iris[-idx, 5]
```

### MCR

```{r}

data(carbs)
m = mcrpure(carbs$D, ncomp = 3, offset = 0.10)
summary(m)

par(mfcol = c(2, 2))
plotContributions(m)
plotSpectra(m)
plotSpectra(m, comp = 2)
plotContributions(m, comp = 2)

S = prep.norm(mda.t(carbs$S), "length")
S.hat = prep.norm(mda.t(m$resspec), "length")

c = predict(m, mda.t(carbs$S))
c = c / apply(c, 2, sum)
show(c)
```

```{r}
m = mcrpure(fusedMat, ncomp = 2, offset = 0.10)
summary(m)

par(mfcol = c(2, 2))
plotContributions(m)
plotSpectra(m)
plotSpectra(m, comp = 1)
plotContributions(m, comp = 1)

par(mfcol = c(1, 2))
plotVariance(m)
plotCumVariance(m)

plotPuritySpectra(m, ylim = c(0, 1600))

```

## Other Methods

### PCA in base R

```{r , include=FALSE, fig.width=10,fig.height=5}
data_pca <- prcomp(fusedMat, center = TRUE, scale. = FALSE)
pcs <- data_pca$x
pc_data <- data.frame(PC1 = pcs[, 1], PC2 = pcs[, 2], Species = rownames(pcs))
pca_plot <- plot_ly(
  data = pc_data,
  x = ~PC1,
  y = ~PC2,
  type = "scatter",
  mode = "markers+text",
  text = ~Species,
  textfont = list(size = 14),
  marker = list(size = 12),
  textposition = "top center"
) %>%
  layout(
    title = "PCA Plot with Annotations",
    xaxis = list(title = "Principal Component 1"),
    yaxis = list(title = "Principal Component 2")
  )
pca_plot
```

```{r ,echo=FALSE,fig.width=10,fig.height=4}
loadings <- as.data.frame(data_pca$rotation[, 1:2])

loadings_plot <- plot_ly()

loadings_plot <- loadings_plot %>% add_trace(
  x = loadings[rownames(loadings) %in% colnames(mrSpec), "PC1"],
  y = loadings[rownames(loadings) %in% colnames(mrSpec), "PC2"],
  type = "scatter",
  mode = "markers+text",
  text = rownames(loadings)[(rownames(loadings) %in% colnames(mrSpec))],
  textfont = list(size = 14),
  legendgroup = "Raman",
  name = "Raman",
  marker = list(size = 10)
)

loadings_plot <- loadings_plot %>% add_trace(
  x = loadings[rownames(loadings) %in% colnames(mmSpec), "PC1"],
  y = loadings[rownames(loadings) %in% colnames(mmSpec), "PC2"],
  type = "scatter",
  mode = "markers+text",
  text = rownames(loadings)[(rownames(loadings) %in% colnames(mmSpec))],
  textfont = list(size = 14),
  legendgroup = "MS",
  name = "MS",
  marker = list(size = 10, color = "red")
)

loadings_plot %>% layout(
    title = "PCA Loadings Plot",
    xaxis = list(title = "Principal Component 1"),
    yaxis = list(title = "Principal Component 2")
  )

```

### Correlation Analysis

```{r}
# Calculate correlation matrix
correlation_matrix <- cor(fusedMat)

# Print correlation matrix
print(correlation_matrix)
```

### Distance Metrics

```{r}
# Calculate Euclidean distance matrix
distance_matrix <- dist(fusedMat)

# Print distance matrix
print(distance_matrix)
```

### Hierarchical Clustering

```{r}
# Perform hierarchical clustering
hc_result <- hclust(dist(fusedMat))

# Plot dendrogram
plot(hc_result)
```

### Multidimensional Scaling (MDS)

```{r}
# Calculate Euclidean distance matrix
distance_matrix <- dist(fusedMat)

# Perform MDS
mds_result <- cmdscale(distance_matrix)

# Plot MDS results
plot(mds_result[,1], mds_result[,2], pch = 19, col = "blue", xlab = "MDS1", ylab = "MDS2")

```

### Kernel Methods

```{r}
# Example code for kernel PCA (assuming Gaussian kernel)
library(kernlab)
kpca_result <- kpca(fusedMat, kernel = "rbfdot")

# Plot kernel PCA results
plot(kpca_result@pcv[,1], kpca_result@pcv[,2], pch = 19, col = "blue", xlab = "KPCA1", ylab = "KPCA2")
```

### Feature Selection and Importance

```{r}
# Example code for feature selection (using random forest as an example)
library(randomForest)
rf_model <- randomForest(fusedMat, importance = TRUE)
print(importance(rf_model))

var_importance <- importance(rf_model)
var_names <- rownames(var_importance)
var_scores <- var_importance[,"MeanDecreaseAccuracy"]  # Adjust column name as needed

# Sort variables by importance score
sorted_indices <- order(var_scores, decreasing = TRUE)
sorted_var_names <- var_names[sorted_indices]
sorted_var_scores <- var_scores[sorted_indices]

# Create bar plot
barplot(sorted_var_scores, names.arg = sorted_var_names, las = 2, cex.names = 0.8,
        main = "Variable Importance", xlab = "Variables", ylab = "Importance Score")
```

###  Cosine similarity and k-means clustering

```{r}
vectors <- fusedMat

similarity_matrix <- tcrossprod(vectors) / (sqrt(rowSums(vectors^2)) %*% t(sqrt(rowSums(vectors^2))))

# Perform k-means clustering
k <- 3  # Number of clusters
clusters <- kmeans(similarity_matrix, centers = k)

# Print cluster assignments
print(clusters$cluster)
```

### Similarity Measures

```{r}
vectors <- fusedMat

# Calculate cosine similarity between pairs of vectors
cosine_similarity <- function(x, y) {
  return(sum(x * y) / (sqrt(sum(x^2)) * sqrt(sum(y^2))))
}

similarity_matrix <- matrix(NA, nrow = nrow(vectors), ncol = nrow(vectors))
for (i in 1:nrow(vectors)) {
  for (j in 1:nrow(vectors)) {
    similarity_matrix[i, j] <- cosine_similarity(vectors[i,], vectors[j,])
  }
}

# Print similarity matrix
print(similarity_matrix)
```

### k-nearest neighbors (KNN) classification

```{r}
training_data <- fusedMat[c(1, 4), ]  # Replace this with your actual training data
test_data <- fusedMat[2:3, ]  # Replace this with your actual test data
labels <- rownames(fusedMat[c(1, 4), ])  # Replace this with your actual labels for training data

# Train KNN model
library(class)
knn_model <- knn(train = training_data, test = test_data, cl = labels, k = 3)

# Print predicted labels for test data
print(knn_model)
```

## Partial Least Squares (PLS)

Not applicable!

```{r}
# Example data
X <- fusedMat  # Replace this with your actual predictor variables
Y <- seq_len(nrow(fusedMat))  # Replace this with your actual response variable

# Perform PLS regression
library(pls)
pls_model <- plsr(Y ~ X, ncomp = 2)  # You can adjust the number of components (ncomp) as needed

# Predict using the PLS model
predicted_values <- predict(pls_model, newdata = X)

# Print predicted values
print(predicted_values)
```

## Support Vector Regression (SVR)

Not applicable!

```{r}
# Example data
X <- matrix(rnorm(100), nrow = 10)  # Replace this with your actual predictor variables
Y <- rnorm(10)  # Replace this with your actual response variable

# Train SVR model
library(e1071)
svr_model <- svm(Y ~ ., data = as.data.frame(X), kernel = "radial", epsilon = 0.1, cost = 1)  # Adjust parameters as needed

# Predict using the SVR model
predicted_values <- predict(svr_model, newdata = as.data.frame(X))

# Print predicted values
print(predicted_values)

```

## MCR-ALS (Multivariate Curve Resolution - Alternating Least Squares)

```{r}

# Perform MCR-ALS analysis
mcr_result <- mcr(fusedMat, nfac = 2)  # You can adjust the number of components (nfac) as needed

# Print MCR-ALS results
print(mcr_result)

```

### 2024-05-03 R/class_S3_ProcessingSettings.R
Complete core implementation before merging

@@ -1,27 +1,17 @@
 #' **ProcessingSettings** S3 class constructor, methods and functions
 #'
-#' @description
-#' Creates a ProcessingSettings S3 class object. The ProcessingSettings are used
-#' in \pkg{StreamFind} for processing data in a given data class method.
+#' @description Creates a ProcessingSettings S3 class object. The ProcessingSettings are used in \pkg{StreamFind} for 
+#' processing data in a given data class method.
 #'
-#' @param call Character of length one with the name of the method where the 
-#' processing settings are to be applied.
-#' @param algorithm Character of length one with the name of the algorithm to
-#' be used.
-#' @param parameters List with parameters specific for the method `call` and
-#' `algorithm`.
-#' @param version Character of length one with the version of the processing
-#' settings.
-#' @param software Character of length one with the name of the software or
-#' package.
+#' @param call Character of length one with the name of the method where theprocessing settings are to be applied.
+#' @param algorithm Character of length one with the name of the algorithm to be used.
+#' @param parameters List with parameters specific for the method `call` and `algorithm`.
+#' @param version Character of length one with the version of the processing settings.
+#' @param software Character of length one with the name of the software or package.
 #' @param developer Character of length one with the name of the developer/s.
 #' @param contact Character of length one with the email of the developer.
 #' @param link Character of length one with the a web link.
 #' @param doi Character of length one with the DOI of algorithm.
-#'
-#' @details See the documentation of the method where the processing settings
-#' are to be applied for more information about applicable algorithms and
-#' parameters.
 #'
 #' @return A ProcessingSettings S3 class
 #'
@@ -74,12 +64,6 @@ ProcessingSettings <- function(call = NA_character_,
     algo <- paste0("Settings_", x$call, "_" , x$algorithm)
     
     s3_classes <- append(s3_classes, algo)
-    
-    # TODO add validation for StreamFind implemented algorithm, may a list of added native algorithms
-    # if ("StreamFind" %in% x$software) {
-    #   page <- "https://odea-project.github.io/StreamFind/reference/"
[…]

### 2024-05-03 src/rcpp_StreamCraft_interface.cpp

@@ -6,7 +6,7 @@
 #include "StreamCraft/src/StreamCraft_lib.hpp"
 
 // [[Rcpp::export]]
-Rcpp::List rcpp_parse_ms_analysis_v2(std::string file_path) {
+Rcpp::List rcpp_parse_ms_analysis(std::string file_path) {
   
   Rcpp::List list_out;
   
@@ -112,7 +112,7 @@ Rcpp::List rcpp_parse_ms_analysis_v2(std::string file_path) {
 }
 
 // [[Rcpp::export]]
-Rcpp::List rcpp_parse_ms_spectra_headers_v2(std::string file_path) {
+Rcpp::List rcpp_parse_ms_spectra_headers(std::string file_path) {
   
   Rcpp::List list_out;
   
@@ -157,7 +157,7 @@ Rcpp::List rcpp_parse_ms_spectra_headers_v2(std::string file_path) {
 }
 
 // [[Rcpp::export]]
-Rcpp::List rcpp_parse_ms_chromatograms_headers_v2(std::string file_path) {
+Rcpp::List rcpp_parse_ms_chromatograms_headers(std::string file_path) {
   
   Rcpp::List list_out;
   
@@ -182,11 +182,11 @@ Rcpp::List rcpp_parse_ms_chromatograms_headers_v2(std::string file_path) {
 }
 
 // [[Rcpp::export]]
-Rcpp::List rcpp_parse_ms_spectra_v2(Rcpp::List analysis,
-                                    std::vector<int> levels,
-                                    Rcpp::DataFrame targets,
-                                    double minIntensityMS1,
-                                    double minIntensityMS2) {
+Rcpp::List rcpp_parse_ms_spectra(Rcpp::List analysis,
+                                 std::vector<int> levels,
+                                 Rcpp::DataFrame targets,
+                                 double minIntensityMS1,
+                                 double minIntensityMS2) {
   
   Rcpp::DataFrame empty_df;
   empty_df.attr("class") = Rcpp::CharacterVector::create("data.table", "data.frame");
@@ -500,7 +500,7 @@ Rcpp::List rcpp_parse_ms_spectra_v2(Rcpp::List analysis,
 }
 
 // [[Rcpp::export]]
-Rcpp::List rcpp_parse_ms_chromatograms_v2(Rcpp::List analysis, std::vector<int> idx) {
[…]

### 2024-05-06 dev/dev_paracetamol.Rmd
Added workflow for paracetamol in dev

--- 2024-04-30 dev/dev_statistics.Rmd
@@
 subtitle: |
-  | Comparison of different bovine serum albumin formulations
+  | Comparison of paracetamol deviations
   | <br>
@@
 
-main_dir <- "C:/Users/apoli/Documents/iSoft/BSA_products"
+main_dir <- "C:/Users/apoli/Documents/iSoft/Paracetamol"
 
-all_files <- list.files(main_dir, full.names = TRUE)
+ms_dir <- paste0(main_dir, "/ms/mzml")
 
-ms_files_d <- all_files[grepl(".d", fixed = T, all_files)][1:13]
+ms_files <- list.files(ms_dir, pattern = ".mzML", full.names = TRUE)
 
-ms_files <- all_files[grepl(".mzML", fixed = T, all_files)]
+raman_dir <- paste0(main_dir, "/raman")
 
-raman_dirs <- all_files[grepl("_Raman", fixed = TRUE, all_files)]
-
-raman_files <- character()
-
-for (i in raman_dirs) raman_files <- c(raman_files, list.files(i, full.names = TRUE))
-
-raman_unified_files <- list.files(main_dir, pattern = "asc", full.names = TRUE)
+raman_files <- list.files(raman_dir, pattern = ".asc", full.names = TRUE)
 ```
@@
 
-Mass Spectrometry (MS) and Raman spectroscopy (Raman) are orthogonal techniques.
-While MS measures the chemical composition of a given sample, Raman is able to provide structural information of the constituents.
-This orthogonality is interesting when applying quality evaluation workflows, where both chemical composition and structural integrity are relevant.
-In this article, we show how MS and Raman can be pre-processed and fused for a combined statistical evaluation to support quality control.
-As test sets, we use four different bovine serum albumin (BSA) products.
-
 # Pre-processing
 
-The preliminary processing is applied differently for MS and Raman as the raw data structure does not allow direct fusion of the data.
-Therefore, in the following two sub-chapters we describe the pre-processing workflow steps for MS and Raman data.
-
 ## MS
-
-The four different BSA products were measured in triplicate in full scan MS mode. Giving the following set of *.d* files.
-
-```{r}
-# Raw data from an Agilent QTOF
-basename(ms_files_d)
-```
-
-The first step is conversion to open source *mzML* format using *MSConvert* from [ProteoWizard](https://proteowizard.sourceforge.io/index.html) CLI, as shwon below.
-
-```{r eval=FALSE}
-# Add option to centroid MS1 data before conversion
-optList <- list(filter = "peakPicking vendor msLevel=1")
-
-# convert_ms_files(ms_files_d, outputFormat = "mzML", optList = optList)
-```
-
-After conversion, the *mzML* files are created in the same directory. The converted files are listed below.
-
-```{r}
-basename(ms_files)
-```
-
-With the *mzML* files, a new *MassSpecEngine* is created to start the MS pre-processing project.
-Below the engine is created and the analysis replicate names are modified to properly assign the analyses within each replicate measurement.
 
@@
 
-# Changes the analysis replicate names
-ms$add_replicate_names(sub("_\\d+$", "", ms$get_analysis_names()))
+ms$add_replicate_names(
+  c(rep("fiz_1", 3),
+    rep("fiz_2", 3),
+    rep("fiz_3", 3),
+    rep("set_32_1", 3),
+    rep("set_32_2", 3),
+    rep("set_32_3", 3),
+    rep("set_33_1", 3),
+    rep("set_33_2", 3),
+    rep("set_33_3", 3),
+    rep("fiz_caff_1", 3),
+    rep("fiz_caff_2", 3),
+    rep("fiz_caff_3", 3),
+    rep("fiz_ass_1", 3),
+    rep("fiz_ass_2", 3),
+    rep("fiz_ass_3", 3),
+    rep("fiz2_1", 3),
+    rep("fiz2_2", 3),
+    rep("fiz2_3", 3),
+    rep("fiz_acid_1", 3),
+    rep("fiz_acid_2", 3),
+    rep("fiz_acid_3", 3),
+    rep("ass_1", 3),
+    rep("ass_2", 3),
+    rep("ass_3", 3),
+    rep("ass2_1", 3),
+    rep("ass2_2", 3),
+    rep("ass2_3", 3),
+    rep("caff_1", 3),
+    rep("caff_2", 3),
+    rep("caff_3", 3),
+    rep("par_1", 3),
+    rep("par_2", 3),
+    rep("par_3", 3),
+    rep("blank", 23)
+  )
+)
 
-# Overview
-ms
+# View(ms$get_overview())
 ```
@@
 ```{r, fig.cap="Total ion chromatograms for each analysis replicate."}
-ms$plot_spectra_tic(colorBy = "replicates")
-```
+# ms$plot_spectra_tic(colorBy = "replicates")
 
-The data processing workflow is based on applied processing methods within the engine. As pre-processing, the following steps are applied:
-
-1. Find the retention time elution of BSA based on the total ion chromatogram (TIC) of each analysis; 
-2. Find the mass-to-charge (*m/z*) range of BSA for deconvolution of charges; 
-3. Deconvolute charges for estimation of the exact mass, in DA;
-
-The individual processing steps are applied according to *ProcessingSettings*.
-Below, the integration of the TIC is shown using the dedicated *ProcessingSettings* after loading the TIC chromatograms from the analysis files and smoothing using a moving average approach.
-
-```{r}
-ms$load_chromatograms("TIC")
+# only blanks
+# ms$plot_spectra_tic(analyses = 100:122, colorBy = "analyses")
 ```
@@
 ```{r}
-# Smoothing of the TIC chromatogram for improving peak finding
-ms$smooth_chromatograms(Settings_smooth_chromatograms_movingaverage(windowSize = 5))
+ms <- ms$subset_analyses(analyses = c(1:9, 28:72, 82:99, 118:121))
+ms$add_blank_names(rep("blank", 76))
+ms$load_spectra()
 ```
@@
 ```{r}
-# Settings for integration of the TIC
-ps_ic <- Settings_integrate_chromatograms_StreamFind(
-  merge = TRUE,
-  closeByThreshold = 5,
-  minPeakHeight = 3E6,
-  minPeakDistance = 3,
-  minPeakWidth = 20,
-  maxPeakWidth = 200,
-  minSN = 5
-)
-
-ps_ic
-```
-
-```{r}
-ms$integrate_chromatograms(ps_ic)
-```
-
-```{r ,fig.cap="TIC with integrated peaks."}
-ms$plot_chromatograms_peaks(colorBy = "replicates")
-```
-
-The average retention time of the main peak was `r mean(ms$chromatograms_peaks$rt)` seconds with a standard deviation of `r sd(ms$chromatograms_peaks$rt)`.
-The following step is to estimate the *m/z* range of BSA for deconvolution of charges.
-For this, the MS1 spectra of the peak retention time +/- 0.1 seconds is plotted.
-
-```{r, fig.cap="Spectra of BSA for each analysis replicate at the peak maximum."}
-ms$plot_spectra_ms1(rt = mean(rbindlist(ms$chromatograms_peaks)$rt), sec = 0.5,
-  colorBy = "replicates", presence = 0.1, minIntensity = 1000, interactive = FALSE
-)
-```
-
-According to the MS1 spectra, the *m/z* from 1200 to 2000 can be added to the dedicated *ProcessingSettings* for deconvolution, as shown below.
-The first step is to load from the mzML files the traces within the retention time and *m/z* ranges selected.
-
-```{r}
-av_rt <- mean(rbindlist(ms$chromatograms_peaks)$rt)
-
-spectra_ranges <- data.table("mzmin" = 1200, "mzmax" = 2000, "rtmin" = av_rt - 1, "rtmax" = av_rt + 1)
-
-# ms <- MassSpecEngine$new(files = ms_files, headers = list(name = "BSA Quality Evaluation Project"))
-# 
-# ms$add_replicate_names(sub("_\\d+$", "", ms$get_analysis_names()))
-
-ms$load_spectra(mz = spectra_ranges)
-
-# ms$cluster_spectra(Settings_cluster_spectra_StreamFind(val = "mz", clustVal = 0.01, presence = 0.2))
-# ms$calculate_spectra_charges(Settings_calculate_spectra_charges_StreamFind(roundVal = 15, relLowCut = 0.2, absLowCut = 8000))
-# ms$deconvolute_spectra(Settings_deconvolute_spectra_StreamFind())
-# ms$smooth_spectra(Settings_smooth_spectra_movingaverage(windowSize = 15))
-# ms$normalize_spectra(Settings_normalize_spectra_minmax()) #liftTozero = TRUE
-# ms$normalize_spectra(Settings_normalize_spectra_blockweight())
-# ms$bin_spectra(Settings_bin_spectra_StreamFind(bins = list("rt" = 2, "mass" = 25), refBinAnalysis = 1))
-
-# View(ms$spectra)
-# ms$plot_spectra()
-# ms$spectra_charges
-# ms$plot_spectra_charges()
-# ms$correct_spectra_baseline(Settings_correct_spectra_baseline_baseline())
-# ms$plot_deconvoluted_spectra()
-# ms$plot_spectra(interactive = T)
-# ms$plot_spectra_charges()
-# ms$spectra_charges
-# ms$deconvoluted_spectra
-# ms$spectra_charges
-# ms$plot_spectra_charges()
-# ms$get_spectra()
-# ms$spectra
-# ms$has_results("spectra")
-# ms$spectra
-# ms$plot_spectra(interactive = T)
-# View(ms$get_results("spectra"))
-```
-
-Below we create an ordered list of *ProcessingSettings* objects for each processing method to be applied and add the list to the engine.
-The exact parameters for each method can be optimized individually by calling the specific method directly.
-Once the settings are added, the workflow can be inspected with the `print_workflow()` method.
-
-```{r}
 ms_workflow_settings <- list(
-  Settings_cluster_spectra_StreamFind(val = "mz", clustVal = 0.01, presence = 0.2),
-  Settings_calculate_spectra_charges_StreamFind(roundVal = 15, relLowCut = 0.2, absLowCut = 8000),
-  Settings_deconvolute_spectra_StreamFind(),
-  Settings_smooth_spectra_movingaverage(windowSize = 15),
-  Settings_bin_spectra_StreamFind(bins = list("rt" = 5, "mass" = 20), refBinAnalysis = 1),
+  Settings_bin_spectra_StreamFind(bins = list("rt" = 5, "mz" = 2), refBinAnalysis = 1),
+  Settings_normalize_spectra_minmax(),
   Settings_average_spectra_StreamFind(),
-  Settings_normalize_spectra_minmax(),
+  Settings_subtract_blank_spectra_StreamFind(negativeToZero = TRUE),
   Settings_normalize_spectra_blockweight()
 )
-
-# Remove the settings added to integrate the TIC chromatograms
-ms$remove_settings()
 
@@
 
-The `run_workflow()` method is called to run the data processing. The workflow will be applied in the order the methods were added to the engine.
-
 ```{r}
@@
 ```{r ,fig.cap="Deconvoluted and pre-processed spectra for each analysis."}
-ms$plot_spectra(colorBy = "analyses")
+ms$plot_spectra()
 ```
@@
 
-The four different BSA products were measured using an online coupling of size exclusion chromatography (SEC) with capillary-enhanced Raman spectroscopy (CERS) through a liquid-core optical fibre flow cell, as reported by Thissen et al.
-(DOI: 10.1021/acs.analchem.3c03991).
-The raw spectra was converted to *asc* format for each retention time.
-A subset of the background and the main BSA elution peak was selected for pre-processing.
-The first pre-processing step is to merge the individual *asc* files into a main file for each BSA product, as shown below.
+```{r}
+raman <- RamanEngine$new(raman_files)
 
-```{r eval=FALSE}
-# Engine with all asc files for each spectrum and each BSA product
-#raman <- RamanEngine$new(raman_files, runParallel = FALSE)
+raman$add_replicate_names(
+  c(
+    rep("ass_1", 3),
+    rep("ass_2", 3),
+    rep("ass_3", 3),
+    rep("caff_1", 3),
+    rep("caff_2", 3),
+    rep("caff_3", 3),
+    rep("blank", 9),
+    rep("fiz_1", 3),
+    rep("fiz_2", 3),
+    rep("fiz_3", 3),
+    rep("fiz_ass_1", 3),
+    rep("fiz_ass_2", 3),
+    rep("fiz_ass_3", 3),
+    rep("fiz_ass_b_2", 9),
+    rep("fiz_caff_1", 3),
+    rep("fiz_caff_2", 3),
+    rep("fiz_caff_3", 3),
+    rep("fiz_b", 9),
+    rep("fiz2_1", 3),
+    rep("fiz2_2", 3),
+    rep("fiz2_3", 3),
+    rep("fiz_d", 9),
+    rep("background", 4),
+    rep("fiz_acid_1", 3),
+    rep("fiz_acid_2", 3),
+    rep("fiz_acid_3", 3),
+    rep("par_1", 3),
+    rep("par_odd", 3),
+    rep("par_3", 3),
+    rep("par_2", 3),
+    rep("cali_150_mgml", 9),
+    rep("cali_010_mgml", 9),
+    rep("cali_025_mgml", 9),
+    rep("cali_050_mgml", 9),
+    rep("cali_075_mgml", 9),
+    rep("background", 4)
+  )
+)
 
-# Folder names of each BSA product
-#base_dirs <- basename(dirname(raman_files))
-
-# The folder name is used as replicate name, which is also used as file name for the unified data file
-#raman$add_replicate_names(base_dirs)
-
-# The spectra are merged into a time series for each BSA product
-#raman$merge_spectra_time_series()
-```
-
-A new engine is created with the unified file for each BSA product and then plotted as a time series.
-
-```{r}
-raman <- RamanEngine$new(raman_unified_files, runParallel = FALSE)
-
-# Modifying the replicate names to match the MS analyses
-raman$add_replicate_names(unique(ms$get_replicate_names())[1:4])
-
-raman
-```
-
-```{r ,fig.cap="Time series chromatogram for each BSA product. Each data point is a Raman spectrum at a specific retention time."}
-raman$plot_chromatograms(yLab = "Abbundance sum for each spectrum", colorBy = "analyses")
+# View(raman$get_overview())
 ```
@@
 ```{r ,fig.cap="Raw averaged Raman spectra for each BSA product between minute 5.5 and 8."}
-raman$plot_spectra(rt = c(5.5, 8), colorBy = "analyses")
+# raman$plot_spectra(colorBy = "replicates")
 ```
 
-As for MS data, the workflow for Raman data is also assembled using an ordered list of processing settings, as shown below.
+```{r}
+raman <- raman$subset_analyses(analyses = c(1:45, 55:63, 73:81, 95:106, 110:115))
+raman$add_blank_names(rep("blank", 81))
+```
 
@@
 raman_workflow_settings <- list(
-  Settings_bin_spectra_StreamFind(unitsVal = "rt", unitsNumber = 3),
-  Settings_normalize_spectra_minmax(),
-  Settings_subtract_spectra_section_StreamFind(sectionWindow = c(0, 3)),
+  Settings_average_spectra_StreamFind(),
+  Settings_subtract_blank_spectra_StreamFind(),
   Settings_smooth_spectra_savgol(fl = 21, forder = 4, dorder = 0),
-  Settings_delete_spectra_section_StreamFind(list("shift" = c(-100, 330))),
-  Settings_delete_spectra_section_StreamFind(list("shift" = c(1800, 2600))),
+  Settings_delete_spectra_section_StreamFind(list("shift" = c(-100, 350))),
+  Settings_delete_spectra_section_StreamFind(list("shift" = c(1800, 2500))),
   Settings_correct_spectra_baseline_airpls(lambda = 45, differences = 1, itermax = 30),
-  Settings_delete_spectra_section_StreamFind(list("rt" = c(0, 5.5))),
-  Settings_delete_spectra_section_StreamFind(list("rt" = c(8, 11))),
-  Settings_average_spectra_StreamFind(collapseTime = TRUE),
   Settings_normalize_spectra_minmax(),
@@
 ```
-
-A particularity of the workflow for Raman data is that the background subtraction (workflow step number 3) is performed from a time region (i.e., between 0 and 3 minutes) before the main BSA peak, which is from approximately from 5 to 8 minutes.
-The workflow is then applied using the `run_workflow()` method.
 
@@
 
-The fusion of MS and Raman data is performed by combining the processed data from both engines.
-The data is then merged into a single data frame for statistical evaluation using a MachineLearningEngine, which is under development.
-
 ```{r}
 rSpec <- raman$get_spectra()
+rSpec <- rSpec[order(rSpec$replicate), ]
 mrSpec <- matrix(rSpec$intensity, nrow = length(unique(rSpec$replicate)), ncol = length(unique(rSpec$shift)), byrow = TRUE, dimnames = list(as.character(unique(rSpec$replicate)), as.character(unique(rSpec$shift))))
@@
 mSpec <- ms$get_spectra()
+mSpec <- mSpec[order(mSpec$replicate), ]
 mmSpec <- matrix(mSpec$intensity, nrow = length(unique(mSpec$replicate)), ncol = length(unique(mSpec$bins)), byrow = TRUE, dimnames = list(unique(mSpec$replicate), unique(mSpec$bins)))
@@
 
-# Statistic analysis
-
-To be done!
-
-## mdatools
+# Statistic analysis (via mdatools)
 
@@
 plot(m, show.labels = TRUE)
+mdaplot(m$res$cal$scores, type = "p", show.labels = TRUE, show.lines = c(0, 0))
 
@@
 loadings_plot <- loadings_plot %>% add_trace(
-  x = m$loadings[rownames(loadings) %in% colnames(mrSpec), "Comp 1"],
-  y = m$loadings[rownames(loadings) %in% colnames(mrSpec), "Comp 2"],
+  x = m$loadings[rownames(m$loadings) %in% colnames(mrSpec), "Comp 1"],
+  y = m$loadings[rownames(m$loadings) %in% colnames(mrSpec), "Comp 2"],
   type = "scatter",

### 2024-05-07 dev/dev_paracetamol.Rmd
Small fixes normalization

@@
 ms$add_replicate_names(
-  c(rep("fiz_1", 3),
-    rep("fiz_2", 3),
-    rep("fiz_3", 3),
-    rep("set_32_1", 3),
-    rep("set_32_2", 3),
-    rep("set_32_3", 3),
-    rep("set_33_1", 3),
-    rep("set_33_2", 3),
-    rep("set_33_3", 3),
+  c(rep("fiz1_1", 3),
+    rep("fiz1_2", 3),
+    rep("fiz1_3", 3),
+    rep("fiz2_1", 3),
+    rep("fiz2_2", 3),
+    rep("fiz2_3", 3),
+    rep("fiz3_1", 3),
+    rep("fiz3_2", 3),
+    rep("fiz3_3", 3),
     rep("fiz_caff_1", 3),
@@
     rep("fiz_ass_3", 3),
-    rep("fiz2_1", 3),
-    rep("fiz2_2", 3),
-    rep("fiz2_3", 3),
+    rep("fiz4_1", 3),
+    rep("fiz4_2", 3),
+    rep("fiz4_3", 3),
     rep("fiz_acid_1", 3),
@@
     rep("fiz_acid_3", 3),
+    rep("ass_n_1", 3),
+    rep("ass_n_2", 3),
+    rep("ass_n_3", 3),
     rep("ass_1", 3),
@@
     rep("ass_3", 3),
-    rep("ass2_1", 3),
-    rep("ass2_2", 3),
-    rep("ass2_3", 3),
     rep("caff_1", 3),
@@
 ```{r}
-ms <- ms$subset_analyses(analyses = c(1:9, 28:72, 82:99, 118:121))
-ms$add_blank_names(rep("blank", 76))
+ms <- ms$subset_analyses(analyses = c(1:63, 73:99, 118:121))
+ms$add_blank_names(rep("blank", 94))
 ms$load_spectra()
@@
     rep("blank", 9),
-    rep("fiz_1", 3),
-    rep("fiz_2", 3),
-    rep("fiz_3", 3),
+    rep("fiz3_1", 3),
+    rep("fiz3_2", 3),
+    rep("fiz3_3", 3),
+    rep("fiz_ass_n_1", 3),
+    rep("fiz_ass_n_2", 3),
+    rep("fiz_ass_n_3", 3),
     rep("fiz_ass_1", 3),
@@
     rep("fiz_ass_3", 3),
-    rep("fiz_ass_b_2", 9),
     rep("fiz_caff_1", 3),
@@
     rep("fiz_caff_3", 3),
-    rep("fiz_b", 9),
+    rep("fiz1_1", 3),
+    rep("fiz1_2", 3),
+    rep("fiz1_3", 3),
+    rep("fiz4_1", 3),
+    rep("fiz4_2", 3),
+    rep("fiz4_3", 3),
     rep("fiz2_1", 3),
@@
     rep("fiz2_3", 3),
-    rep("fiz_d", 9),
     rep("background", 4),
@@
 ```{r}
-raman <- raman$subset_analyses(analyses = c(1:45, 55:63, 73:81, 95:106, 110:115))
-raman$add_blank_names(rep("blank", 81))
+raman <- raman$subset_analyses(analyses = c(1:36, 46:90, 95:106, 110:115))
+raman$add_blank_names(rep("blank", 99))
+raman$plot_spectra(colorBy = "replicates")
 ```
@@
 rSpec <- rSpec[order(rSpec$replicate), ]
-mrSpec <- matrix(rSpec$intensity, nrow = length(unique(rSpec$replicate)), ncol = length(unique(rSpec$shift)), byrow = TRUE, dimnames = list(as.character(unique(rSpec$replicate)), as.character(unique(rSpec$shift))))
+mrSpec <- matrix(rSpec$intensity, nrow = length(unique(rSpec$replicate)), ncol = length(unique(rSpec$shift)), byrow = TRUE, dimnames = list(as.character(unique(rSpec$replicate)), as.character(round(unique(rSpec$shift), digits = 1))))
 
@@
 attr(fusedMat, "xaxis.values") = seq_len(ncol(fusedMat))
+
+# Mean centering improves separation but can we do it?
+library(mdatools)
+fusedMat = prep.autoscale(fusedMat, center = TRUE, scale = FALSE)
 ```
@@
 
-# Statistic analysis (via mdatools)
+# Statistic analysis
 
-### PCA
+## Overall PCA
 
@@
 library(mdatools)
-data(people)
-
-m = pca(people, 7, scale = TRUE, info = "People PCA model")
-m = selectCompNum(m, 5)
-
-print(m$res$cal)
-
-par(mfrow = c(1, 2))
-mdaplot(m$res$cal$scores, type = "p", show.labels = TRUE, show.lines = c(0, 0))
-mdaplot(m$loadings, type = "p", show.labels = TRUE, show.lines = c(0, 0))
-```
-
-```{r}
-fusedMat2 <- prep.autoscale(fusedMat, center = FALSE, scale = TRUE)
-cl <- StreamFind:::.get_colors(rownames(fusedMat2))
-fig <- plot_ly()
-xVal <- seq_len(length(fusedMat2[1, ]))
-for (i in seq_len(nrow(fusedMat2))) {
-  fig <- fig %>% add_trace(
-    x = xVal,
-    y = fusedMat2[i, ],
-    type = "scatter", mode = "lines",
-    line = list(width = 0.5, color = unname(cl[i])),
-    name = names(cl)[i],
-    legendgroup = names(cl)[i],
-    showlegend = TRUE
-  )
-}
-xaxis <- list(
-  linecolor = toRGB("black"),
-  linewidth = 2, title = "Bins",
-  titlefont = list(size = 12, color = "black")
-)
-yaxis <- list(
-  linecolor = toRGB("black"),
-  linewidth = 2, title = "Intensity",
-  titlefont = list(size = 12, color = "black")
-)
-fig <- fig %>% plotly::layout(xaxis = xaxis, yaxis = yaxis)
-fig
-```
-
-```{r, include=FALSE}
-
-m = pca(fusedMat, 3, scale = FALSE, info = "People PCA model")
-plot(m, show.labels = TRUE)
-mdaplot(m$res$cal$scores, type = "p", show.labels = TRUE, show.lines = c(0, 0))
-
-par(mfrow = c(2, 2))
-plotVariance(m, type = "h", show.labels = TRUE)
-plotQDoF(m)
-plotT2DoF(m)
-plotDistDoF(m)
-
-par(mfrow = c(1, 2))
-plotQDoF(m, main = "DoF plot (classic)")
-
-m = setDistanceLimits(m, lim.type = "ddrobust")
-plotQDoF(m, main = "DoF plot (robust)")
-
-plotExtreme(m, comp = 1:2, main = "Extreme plot (cal)")
-
-plotLoadings(m, 1:2, type = 'l')
-
-# par(mfrow = c(2, 2))
-# mdaplot(m$res$cal$scores, type = "p", show.labels = TRUE, show.lines = c(0, 0))
-# mdaplot(m$loadings, type = "p", show.labels = TRUE, show.lines = c(0, 0))
-# plotResiduals(m$res$cal, show.labels = TRUE)
-# plotVariance(m$res$cal, type = "h", show.labels = TRUE, labels = "values")
+m = pca(fusedMat, 3, center = FALSE, scale = FALSE, info = "Overall PCA model")
 ```
@@
 ```{r ,echo=FALSE,fig.width=10,fig.height=4}
-
 library(plotly)
 
+cl <- StreamFind:::.get_colors(rownames(fusedMat))
+scores_polt <- plot_ly()
+
+scores_polt <- scores_polt %>% add_trace(
+  x = m$res$cal$scores[, "Comp 1"],
+  y = m$res$cal$scores[, "Comp 2"],
+  type = "scatter",
+  mode = "markers+text",
+  color = names(cl),
+  colors = cl,
+  text = rownames(m$res$cal$scores),
+  textfont = list(size = 14, color = cl),
+  textposition = "top",
+  marker = list(size = 10, color = cl),
+  showlegend = FALSE
+)
+
+scores_polt <- scores_polt %>% layout(
+  xaxis = list(title = "Principal Component 1"),
+  yaxis = list(title = "Principal Component 2")
+)
+
+cl2 <- StreamFind:::.get_colors(1:2)
 loadings_plot <- plot_ly()
@@
   text = rownames(m$loadings)[(rownames(m$loadings) %in% colnames(mrSpec))],
-  textfont = list(size = 14),
+  textfont = list(size = 14, color = cl2[1]),
+  textposition = "top",
   legendgroup = "Raman",
   name = "Raman",
-  marker = list(size = 10)
+  marker = list(size = 10, color = cl2[1])
 )
@@
   text = rownames(m$loadings)[(rownames(m$loadings) %in% colnames(mmSpec))],
-  textfont = list(size = 14),
+  textfont = list(size = 14, color = cl2[2]),
+  textposition = "top",
   legendgroup = "MS",
   name = "MS",
-  marker = list(size = 10, color = "red")
+  marker = list(size = 10, color = cl2[2])
 )
 
-loadings_plot %>% layout(
-    title = "PCA Loadings Plot",
+loadings_plot <- loadings_plot %>% layout(
   xaxis = list(title = "Principal Component 1"),
@@
 
+subplot(
+  scores_polt, loadings_plot, nrows = 1,
+  shareX = FALSE, shareY = FALSE, titleX = TRUE, titleY = TRUE,
+  margin = 0.1
+)
 ```
 
-### PLS
+## Unsupervised
+
+Using Principle Components Analysis (PCA) for unsupervised evaluation.
 
 ```{r}
+# rownames(fusedMat)
+main_set <- fusedMat[c(14:15, 18), ]
+test_set <- fusedMat[c(1:13, 16:17, 19:30), ]
 
-library(mdatools)
-data(people)
+m3 = pca(main_set, 2, center = FALSE, scale = FALSE, info = "Unsupervised PCA model", alpha = 1E-1, gamma = 1E-2)
 
-idx = seq(4, 32, 4)
-Xc = people[-idx, -4]
-yc = people[-idx, 4, drop = FALSE]
-Xt = people[idx, -4]
-yt = people[idx, 4, drop = FALSE]
+plot(m3)
 
-m = pls(Xc, yc, 7, scale = TRUE, cv = 1, info = "Shoesize prediction model")
+res = predict(m3, test_set)
 
-par(mfrow = c(2, 2))
-plotRegcoeffs(m)
-plotRegcoeffs(m, ncomp = 2)
-plot(m$coeffs, ncomp = 3, type = "b", show.labels = TRUE)
-plot(m$coeffs, ncomp = 2)
+c = categorize(m3, res)
 
-par(mfrow = c(1, 2))
-plotPredictions(m)
-plotPredictions(m, ncomp = 1)
+names(c) <- rownames(test_set)
 
-plotVIPScores(m, show.labels = TRUE)
+plotResiduals(m3, 
+  res = list("train" = m3$res$cal, "new" = res),
+  norm = TRUE, show.labels = TRUE,
+  lim.col = c("red", "orange"), lim.lwd = c(2, 2), lim.lty = c(1, 2)
+)
 
-res = predict(m, Xt, yt)
-print(res)
-
-data(simdata)
-Xc = simdata$spectra.c
-yc = simdata$conc.c[, 1]
-Xt = simdata$spectra.t
-yt = simdata$conc.t[, 1]
-
-m = pls(Xc, yc, 10, x.test = Xt, y.test = yt, ncomp.selcrit = "min")
-summary(m)
-
-plot(m)
-
+print(c)
 ```
 
-### PLS-DA
+## Supervised (classification)
+
+Using k-nearest neighbors (KNN) for classification.
 
 ```{r}
-data(iris)
+train_set <- fusedMat[c(grep("_1", rownames(fusedMat)), grep("_2", rownames(fusedMat))), ]
 
-cal.ind = c(1:25, 51:75, 101:125)
-val.ind = c(26:50, 76:100, 126:150)
+train_set <- train_set[order(rownames(train_set)), ]
 
-Xc = iris[cal.ind, 1:4]
-Xv = iris[val.ind, 1:4]
+test_set <- fusedMat[grep("_3", rownames(fusedMat)), ]
 
-cc.all = iris[cal.ind, 5]
-cv.all = iris[val.ind, 5]
+labels <- factor(c(rep("ass", 2), rep("caff", 2), rep("fiz_acid", 2), rep("fiz_ass", 2), rep("fiz_caff", 2),
+  rep("fiz_altered", 2), rep("fiz_altered", 2), rep("fiz", 2), rep("fiz", 2), rep("par", 2))
+)
 
-cc.vir = cc.all == "virginica"
-cv.vir = cv.all == "virginica"
-show(cc.vir)
+library(class)
 
-m.all = plsda(Xc, cc.all, 3, cv = 1)
-m.vir = plsda(Xc, cc.vir, 3, cv = 1, classname = "virginica")
+knn_model <- knn(train = train_set, test = test_set, cl = labels, k = 3, prob = FALSE)
 
-summary(m.all)
+names(knn_model) <- rownames(test_set)
 
-par(mfrow = c(1, 2))
-plotPredictions(m.all)
-plotPredictions(m.vir)
-
-res = predict(m.all, Xv, cv.all)
-summary(res)
-
-par(mfrow = c(1, 1))
-plotPredictions(res, show.labels = TRUE)
-
+data.frame(class = knn_model)
+# print(knn_model)
 ```
 
-### SIMCA
-```{r}
-idx = seq(1, nrow(iris), by = 2)
+## Other
 
-# split the values
-Xc = iris[idx, 1:4]
-cc = iris[idx, 5]
-
-Xt = iris[-idx, 1:4]
-ct = iris[-idx, 5]
-```
-
-### MCR
+Using Hierarchical Clustering.
 
 ```{r}
-
-data(carbs)
-m = mcrpure(carbs$D, ncomp = 3, offset = 0.10)
-summary(m)
-
-par(mfcol = c(2, 2))
-plotContributions(m)
-plotSpectra(m)
-plotSpectra(m, comp = 2)
-plotContributions(m, comp = 2)
-
-S = prep.norm(mda.t(carbs$S), "length")
-S.hat = prep.norm(mda.t(m$resspec), "length")
-
-c = predict(m, mda.t(carbs$S))
-c = c / apply(c, 2, sum)
-show(c)
-```
-
-```{r}
-m = mcrpure(fusedMat, ncomp = 2, offset = 0.10)
-summary(m)
-
-par(mfcol = c(2, 2))
-plotContributions(m)
-plotSpectra(m)
-plotSpectra(m, comp = 1)
-plotContributions(m, comp = 1)
-
-par(mfcol = c(1, 2))
-plotVariance(m)
-plotCumVariance(m)
-
-plotPuritySpectra(m, ylim = c(0, 1600))
-
-```
-
-## Other Methods
-
-### PCA in base R
-
-```{r , include=FALSE, fig.width=10,fig.height=5}
-data_pca <- prcomp(fusedMat, center = TRUE, scale. = FALSE)
-pcs <- data_pca$x
-pc_data <- data.frame(PC1 = pcs[, 1], PC2 = pcs[, 2], Species = rownames(pcs))
-pca_plot <- plot_ly(
-  data = pc_data,
-  x = ~PC1,
-  y = ~PC2,
-  type = "scatter",
-  mode = "markers+text",
-  text = ~Species,
-  textfont = list(size = 14),
-  marker = list(size = 12),
-  textposition = "top center"
-) %>%
-  layout(
-    title = "PCA Plot with Annotations",
-    xaxis = list(title = "Principal Component 1"),
-    yaxis = list(title = "Principal Component 2")
-  )
-pca_plot
-```
-
-```{r ,echo=FALSE,fig.width=10,fig.height=4}
-loadings <- as.data.frame(data_pca$rotation[, 1:2])
-
-loadings_plot <- plot_ly()
-
-loadings_plot <- loadings_plot %>% add_trace(
-  x = loadings[rownames(loadings) %in% colnames(mrSpec), "PC1"],
-  y = loadings[rownames(loadings) %in% colnames(mrSpec), "PC2"],
-  type = "scatter",
-  mode = "markers+text",
-  text = rownames(loadings)[(rownames(loadings) %in% colnames(mrSpec))],
-  textfont = list(size = 14),
-  legendgroup = "Raman",
-  name = "Raman",
-  marker = list(size = 10)
-)
-
-loadings_plot <- loadings_plot %>% add_trace(
-  x = loadings[rownames(loadings) %in% colnames(mmSpec), "PC1"],
-  y = loadings[rownames(loadings) %in% colnames(mmSpec), "PC2"],
-  type = "scatter",
-  mode = "markers+text",
-  text = rownames(loadings)[(rownames(loadings) %in% colnames(mmSpec))],
-  textfont = list(size = 14),
-  legendgroup = "MS",
-  name = "MS",
-  marker = list(size = 10, color = "red")
-)
-
-loadings_plot %>% layout(
-    title = "PCA Loadings Plot",
-    xaxis = list(title = "Principal Component 1"),
-    yaxis = list(title = "Principal Component 2")
-  )
-
-```
-
-### Correlation Analysis
-
-```{r}
-# Calculate correlation matrix
-correlation_matrix <- cor(fusedMat)
-
-# Print correlation matrix
-print(correlation_matrix)
-```
-
-### Distance Metrics
-
-```{r}
-# Calculate Euclidean distance matrix
-distance_matrix <- dist(fusedMat)
-
-# Print distance matrix
-print(distance_matrix)
-```
-
-### Hierarchical Clustering
-
-```{r}
-# Perform hierarchical clustering
-hc_result <- hclust(dist(fusedMat))
-
-# Plot dendrogram
+hc_result <- hclust(dist(fusedMat), method = "ward.D2")
 plot(hc_result)
 ```
-
-### Multidimensional Scaling (MDS)
-
-```{r}
-# Calculate Euclidean distance matrix
-distance_matrix <- dist(fusedMat)
-
-# Perform MDS
-mds_result <- cmdscale(distance_matrix)
-
-# Plot MDS results
-plot(mds_result[,1], mds_result[,2], pch = 19, col = "blue", xlab = "MDS1", ylab = "MDS2")
-
-```
-
-### Kernel Methods
-
-```{r}
-# Example code for kernel PCA (assuming Gaussian kernel)
-library(kernlab)
-kpca_result <- kpca(fusedMat, kernel = "rbfdot")
-
-# Plot kernel PCA results
-plot(kpca_result@pcv[,1], kpca_result@pcv[,2], pch = 19, col = "blue", xlab = "KPCA1", ylab = "KPCA2")
-```
-
-### Feature Selection and Importance
-
-```{r}
-# Example code for feature selection (using random forest as an example)
-library(randomForest)
-rf_model <- randomForest(fusedMat, importance = TRUE)
-print(importance(rf_model))
-
-var_importance <- importance(rf_model)
-var_names <- rownames(var_importance)
-var_scores <- var_importance[,"MeanDecreaseAccuracy"]  # Adjust column name as needed
-
-# Sort variables by importance score
-sorted_indices <- order(var_scores, decreasing = TRUE)
-sorted_var_names <- var_names[sorted_indices]
-sorted_var_scores <- var_scores[sorted_indices]
-
-# Create bar plot
-barplot(sorted_var_scores, names.arg = sorted_var_names, las = 2, cex.names = 0.8,
-        main = "Variable Importance", xlab = "Variables", ylab = "Importance Score")
-```
-
-###  Cosine similarity and k-means clustering
-
-```{r}
-vectors <- fusedMat
-
-similarity_matrix <- tcrossprod(vectors) / (sqrt(rowSums(vectors^2)) %*% t(sqrt(rowSums(vectors^2))))
-
-# Perform k-means clustering
-k <- 3  # Number of clusters
-clusters <- kmeans(similarity_matrix, centers = k)
-
-# Print cluster assignments
-print(clusters$cluster)
-```
-
-### Similarity Measures
-
-```{r}
-vectors <- fusedMat
-
-# Calculate cosine similarity between pairs of vectors
-cosine_similarity <- function(x, y) {
-  return(sum(x * y) / (sqrt(sum(x^2)) * sqrt(sum(y^2))))
-}
-
-similarity_matrix <- matrix(NA, nrow = nrow(vectors), ncol = nrow(vectors))
-for (i in 1:nrow(vectors)) {
-  for (j in 1:nrow(vectors)) {
-    similarity_matrix[i, j] <- cosine_similarity(vectors[i,], vectors[j,])
-  }
-}
-
-# Print similarity matrix
-print(similarity_matrix)
-```
-
-### k-nearest neighbors (KNN) classification
-
-```{r}
-training_data <- fusedMat[c(1, 4), ]  # Replace this with your actual training data
-test_data <- fusedMat[2:3, ]  # Replace this with your actual test data
-labels <- rownames(fusedMat[c(1, 4), ])  # Replace this with your actual labels for training data
-
-# Train KNN model
-library(class)
-knn_model <- knn(train = training_data, test = test_data, cl = labels, k = 3)
-
-# Print predicted labels for test data
-print(knn_model)
-```
-
-## Partial Least Squares (PLS)
-
-Not applicable!
-
-```{r}
-# Example data
-X <- fusedMat  # Replace this with your actual predictor variables
-Y <- seq_len(nrow(fusedMat))  # Replace this with your actual response variable
-
-# Perform PLS regression
-library(pls)
-pls_model <- plsr(Y ~ X, ncomp = 2)  # You can adjust the number of components (ncomp) as needed
-
-# Predict using the PLS model
-predicted_values <- predict(pls_model, newdata = X)
-
-# Print predicted values
-print(predicted_values)
-```
-
-## Support Vector Regression (SVR)
-
-Not applicable!
-
-```{r}
-# Example data
-X <- matrix(rnorm(100), nrow = 10)  # Replace this with your actual predictor variables
-Y <- rnorm(10)  # Replace this with your actual response variable
-
-# Train SVR model
-library(e1071)
-svr_model <- svm(Y ~ ., data = as.data.frame(X), kernel = "radial", epsilon = 0.1, cost = 1)  # Adjust parameters as needed
-
-# Predict using the SVR model
-predicted_values <- predict(svr_model, newdata = as.data.frame(X))
-
-# Print predicted values
-print(predicted_values)
-
-```
-
-## MCR-ALS (Multivariate Curve Resolution - Alternating Least Squares)
-
-```{r}
-
-# Perform MCR-ALS analysis
-mcr_result <- mcr(fusedMat, nfac = 2)  # You can adjust the number of components (nfac) as needed
-
-# Print MCR-ALS results
-print(mcr_result)
-
-```

### 2024-05-10 dev/dev_paracetamol.Rmd
Initiated StatisticEngine
Added extra funcs in StatisticEngine

@@
 
-mSpec <- ms$get_spectra()
+mSpec <- ms2$get_spectra()
 mSpec <- mSpec[order(mSpec$replicate), ]
@@
 fusedMat <- cbind(mmSpec, mrSpec)
+
+# stat <- StatisticEngine$new(data = fusedMat)
+# stat$make_pca_model(Settings_make_pca_model_mdatools(center = TRUE, ncomp = 3))
+# stat$plot_model_scores()
+# stat$plot_model_loadings()
+# stat$plot_model_residuals()
+# subset matrix for 1 row but keep matrix not vector
 
@@
 ```{r}
+library(mdatools)
 # rownames(fusedMat)
@@
 
-m3 = pca(main_set, 2, center = FALSE, scale = FALSE, info = "Unsupervised PCA model", alpha = 1E-1, gamma = 1E-2)
+m3 = pca(main_set, 3, center = FALSE, scale = FALSE, info = "Unsupervised PCA model", alpha = 1E-1, gamma = 1E-2)
 
@@
 res = predict(m3, test_set)
+
+plot(res)
 
@@
 plotResiduals(m3,
+  ncomp = 2,
   res = list("train" = m3$res$cal, "new" = res),
@@
 
+plotResiduals(res, show.plot = FALSE, norm = TRUE)
+
 print(c)
+
+is(m3$res$cal, "ldecomp")
+
+stat <- StatisticEngine$new(data = main_set)
+stat$make_pca_model(Settings_make_pca_model_mdatools(center = FALSE, ncomp = 3, alpha = 1E-1, gamma = 1E-2))
+stat$predict(test_set)
+
+stat$plot_predicted_distances()
+
 ```

### 2024-06-06 dev/dev_paracetamol.Rmd
Updated statistic engine with classification

--- 2024-05-07 dev/dev_paracetamol.Rmd
@@
 fusedMat <- cbind(mmSpec, mrSpec)
-
-attr(fusedMat, "xaxis.name") = "Keys"
-attr(fusedMat, "xaxis.values") = seq_len(ncol(fusedMat))
-
-# Mean centering improves separation but can we do it?
-library(mdatools)
-fusedMat = prep.autoscale(fusedMat, center = TRUE, scale = FALSE)
 ```
@@
 
-## Overall PCA
-
-```{r, include=FALSE}
-library(mdatools)
-m = pca(fusedMat, 3, center = FALSE, scale = FALSE, info = "Overall PCA model")
-```
-
-```{r ,echo=FALSE,fig.width=10,fig.height=4}
-library(plotly)
-
-cl <- StreamFind:::.get_colors(rownames(fusedMat))
-scores_polt <- plot_ly()
-
-scores_polt <- scores_polt %>% add_trace(
-  x = m$res$cal$scores[, "Comp 1"],
-  y = m$res$cal$scores[, "Comp 2"],
-  type = "scatter",
-  mode = "markers+text",
-  color = names(cl),
-  colors = cl,
-  text = rownames(m$res$cal$scores),
-  textfont = list(size = 14, color = cl),
-  textposition = "top",
-  marker = list(size = 10, color = cl),
-  showlegend = FALSE
-)
-
-scores_polt <- scores_polt %>% layout(
-  xaxis = list(title = "Principal Component 1"),
-  yaxis = list(title = "Principal Component 2")
-)
-
-cl2 <- StreamFind:::.get_colors(1:2)
-loadings_plot <- plot_ly()
-
-loadings_plot <- loadings_plot %>% add_trace(
-  x = m$loadings[rownames(m$loadings) %in% colnames(mrSpec), "Comp 1"],
-  y = m$loadings[rownames(m$loadings) %in% colnames(mrSpec), "Comp 2"],
-  type = "scatter",
-  mode = "markers+text",
-  text = rownames(m$loadings)[(rownames(m$loadings) %in% colnames(mrSpec))],
-  textfont = list(size = 14, color = cl2[1]),
-  textposition = "top",
-  legendgroup = "Raman",
-  name = "Raman",
-  marker = list(size = 10, color = cl2[1])
-)
-
-loadings_plot <- loadings_plot %>% add_trace(
-  x = m$loadings[rownames(m$loadings) %in% colnames(mmSpec), "Comp 1"],
-  y = m$loadings[rownames(m$loadings) %in% colnames(mmSpec), "Comp 2"],
-  type = "scatter",
-  mode = "markers+text",
-  text = rownames(m$loadings)[(rownames(m$loadings) %in% colnames(mmSpec))],
-  textfont = list(size = 14, color = cl2[2]),
-  textposition = "top",
-  legendgroup = "MS",
-  name = "MS",
-  marker = list(size = 10, color = cl2[2])
-)
-
-loadings_plot <- loadings_plot %>% layout(
-  xaxis = list(title = "Principal Component 1"),
-  yaxis = list(title = "Principal Component 2")
-)
-
-subplot(
-  scores_polt, loadings_plot, nrows = 1,
-  shareX = FALSE, shareY = FALSE, titleX = TRUE, titleY = TRUE,
-  margin = 0.1
-)
-```
-
 ## Unsupervised
@@
 ```{r}
-# rownames(fusedMat)
-main_set <- fusedMat[c(14:15, 18), ]
-test_set <- fusedMat[c(1:13, 16:17, 19:30), ]
+main_set <- fusedMat[c(8:9, 10, 12, 14:15, 17:18), ]
+test_set <- fusedMat[c(1:7, 11, 13, 16, 19:30), ]
 
-m3 = pca(main_set, 2, center = FALSE, scale = FALSE, info = "Unsupervised PCA model", alpha = 1E-1, gamma = 1E-2)
-
-plot(m3)
-
-res = predict(m3, test_set)
-
-c = categorize(m3, res)
-
-names(c) <- rownames(test_set)
-
-plotResiduals(m3, 
-  res = list("train" = m3$res$cal, "new" = res),
-  norm = TRUE, show.labels = TRUE,
-  lim.col = c("red", "orange"), lim.lwd = c(2, 2), lim.lty = c(1, 2)
-)
-
-print(c)
+stat <- StatisticEngine$new(data = main_set)
+stat$make_pca_model(Settings_make_pca_model_mdatools(ncomp = 3, alpha = 0.005, gamma = 0.0005))
+stat$predict(test_set)
+stat$plot_predicted_distances()
 ```
@@
 train_set <- fusedMat[c(grep("_1", rownames(fusedMat)), grep("_2", rownames(fusedMat))), ]
-
 train_set <- train_set[order(rownames(train_set)), ]
@@
 
-labels <- factor(c(rep("ass", 2), rep("caff", 2), rep("fiz_acid", 2), rep("fiz_ass", 2), rep("fiz_caff", 2),
-  rep("fiz_altered", 2), rep("fiz_altered", 2), rep("fiz", 2), rep("fiz", 2), rep("par", 2))
+labels <- c(
+  rep("ass", 2), rep("caff", 2), rep("fiz_acid", 2), rep("fiz_ass", 2), rep("fiz_caff", 2),
+  rep("fiz_altered", 2), rep("fiz_altered", 2), rep("fiz", 2), rep("fiz", 2), rep("par", 2)
 )
 
-library(class)
-
-knn_model <- knn(train = train_set, test = test_set, cl = labels, k = 3, prob = FALSE)
-
-names(knn_model) <- rownames(test_set)
-
-data.frame(class = knn_model)
-# print(knn_model)
+stat2 <- StatisticEngine$new(data = train_set)
+stat2$add_classes(labels)
+stat2$prepare_classification(Settings_prepare_classification_knn(k = 3))
+stat2$classify(test_set)
+stat2$classification_results
 ```
@@
 ```
+
+## Old code
+
+```{r include=FALSE}
+# library(mdatools)
+# m3 = pca(main_set, 3, center = FALSE, scale = FALSE, info = "Unsupervised PCA model", alpha = 0.005, gamma = 0.0005)
+# plot(m3, show.labels = TRUE)
+# res = predict(m3, test_set)
+# plot(res, show.labels = TRUE)
+# c = categorize(m3, res, 3)
+# names(c) <- rownames(test_set)
+# c
+# plotResiduals(m3,
+#   ncomp = 3,
+#   res = list("train" = m3$res$cal, "new" = res),
+#   norm = TRUE, show.labels = TRUE,
+#   lim.col = c("red", "orange"), lim.lwd = c(2, 2), lim.lty = c(1, 2)
+# )
+# plotResiduals(ncomp = 1, res, show.plot = FALSE, norm = TRUE)
+# print(c)
+# is(m3$res$cal, "ldecomp")
+```

### 2024-06-06 R/class_R6_RamanEngine.R

@@ -242,7 +242,7 @@ RamanEngine <- R6::R6Class("RamanEngine",
           
           newAnalyses <- self$get_analyses(keepAnalyses)
           
-          new_ms <- suppressMessages(RamanEngine$new(
+          new_engine <- suppressMessages(RamanEngine$new(
             headers = self$get_headers(),
             settings = self$get_settings(),
             analyses = newAnalyses
@@ -250,21 +250,21 @@ RamanEngine <- R6::R6Class("RamanEngine",
           
           if (self$has_results("spectra")) {
             
-            analyses_left <- new_ms$get_analysis_names()
+            analyses_left <- new_engine$get_analysis_names()
             
-            replicates_left <- new_ms$get_replicate_names()[analyses_left]
+            replicates_left <- new_engine$get_replicate_names()[analyses_left]
             
             if (!is.null(analyses_left)) {
               sel <- names(private$.results$spectra$data) %in% c(analyses_left, replicates_left)
               res <- private$.results$spectra
               res <- res$data[sel]
-              suppressMessages(new_ms$add_results(res))
+              suppressMessages(new_engine$add_results(res))
             }
           }
           
-          message("\U2713 Subset with ", new_ms$get_number_analyses(), " analyses created!")
+          message("\U2713 Subset with ", new_engine$get_number_analyses(), " analyses created!")
           
-          return(new_ms)
+          return(new_engine)
           
         } else {
           warning("No analyses selected to subset! Not done.")

### 2024-06-06 R/class_R6_StatisticEngine.R

@@ -102,36 +102,20 @@ StatisticEngine <- R6::R6Class("StatisticEngine",
       }
     },
     
-    #' @field predicted Predicted results model.
+    #' @field prediction_results Prediction results from model.
     #'
-    predicted = function(value) {
-      
-      if (missing(value)) {
-        
-        res <- self$get_results("predicted")
-        
-        if (length(res) > 0) {
-          
-          res$predicted
-          
-        } else {
-          NULL
-        }
-      } else {
-        
-        self$add_results(
-          list(
-            "predicted" = list(
-              "results" = value[["results"]],
-              "data" = value[["data"]],
-              "software" = "mdatools",
-              "version" = as.character(packageVersion("mdatools"))
-            )
-          )
-        )
-        
-        invisible(self)
-      }
+    prediction_results = function() {  
+      res <- self$get_results("prediction")
+      if (length(res) > 0) return(res$prediction[["results"]])
+      NULL
+    },
+    
+    #' @field classification_results Classification results.
+    #' 
+    classification_results = function() {
+      res <- self$get_results("classification")
+      if (length(res) > 0) return(res$classification[["results"]])
+      NULL
     }
   ),
 
@@ -160,14 +144,7 @@ StatisticEngine <- R6::R6Class("StatisticEngine",
      
       if (!is.null(data)) self$add_data(data)
      
-      private$.register(
-        "created",
-        "StatisticEngine",
-        headers$name,
-        "StreamFind",
-        as.character(packageVersion("StreamFind")),
-        paste(c(headers$author, headers$path), collapse = ", ")
-      )
+      private$.register("created", "StatisticEngine", headers$name, paste(c(headers$author, headers$path), collapse = ", "))
     },
     
     ## ___ get -----
@@ -182,7 +159,7 @@ StatisticEngine <- R6::R6Class("StatisticEngine",
         
         ov$nvars <- vapply(private$.analyses, function(x) ncol(x$data), 0)
         
-        ov$class <- vapply(private$.analyses, function(x) paste(x$classes, collapse = ", "), NA_character_)
+        ov$class <- vapply(private$.analyses, function(x) x$class, NA_character_)
         
         row.names(ov) <- seq_len(nrow(ov))
         
@@ -190,6 +167,18 @@ StatisticEngine <- R6::R6Class("StatisticEngine",
        
       } else {
        data.frame()
+      }
+    },
+    
[…]

### 2024-06-07 dev/dev_paracetamol.Rmd

@@
 stat <- StatisticEngine$new(data = main_set)
-stat$make_pca_model(Settings_make_pca_model_mdatools(ncomp = 3, alpha = 0.005, gamma = 0.0005))
+stat$make_model(Settings_make_model_pca_mdatools(ncomp = 3, alpha = 0.005, gamma = 0.0005))
 stat$predict(test_set)

### 2024-06-10 dev/dev_paracetamol.Rmd
update statistic engine

@@
 ---
-title: 'Fusion of Mass Spectrometry and Raman Spectroscopy Data'
+title: 'Fusion of Mass Spectrometry and Raman Spectroscopy Data Case Study 2'
 subtitle: |
-  | Comparison of paracetamol deviations
+  | Detection of quality deviations in paracetamol products
   | <br>
@@
 
-```{r}
+```{r ms-setup, include=FALSE}
 ms <- MassSpecEngine$new(files = ms_files, headers = list(name = "BSA Quality Evaluation Project"))
@@
 ms$add_replicate_names(
-  c(rep("fiz1_1", 3),
-    rep("fiz1_2", 3),
-    rep("fiz1_3", 3),
-    rep("fiz2_1", 3),
-    rep("fiz2_2", 3),
-    rep("fiz2_3", 3),
-    rep("fiz3_1", 3),
-    rep("fiz3_2", 3),
-    rep("fiz3_3", 3),
-    rep("fiz_caff_1", 3),
-    rep("fiz_caff_2", 3),
-    rep("fiz_caff_3", 3),
-    rep("fiz_ass_1", 3),
-    rep("fiz_ass_2", 3),
-    rep("fiz_ass_3", 3),
-    rep("fiz4_1", 3),
-    rep("fiz4_2", 3),
-    rep("fiz4_3", 3),
-    rep("fiz_acid_1", 3),
-    rep("fiz_acid_2", 3),
-    rep("fiz_acid_3", 3),
+  c(rep("fiz_a1", 3),
+    rep("fiz_a2", 3),
+    rep("fiz_a3", 3),
+    rep("fiz_b1", 3),
+    rep("fiz_b2", 3),
+    rep("fiz_b3", 3),
+    rep("fiz_c1", 3),
+    rep("fiz_c2", 3),
+    rep("fiz_c3", 3),
+    rep("fiz_caff1", 3),
+    rep("fiz_caff2", 3),
+    rep("fiz_caff3", 3),
+    rep("fiz_ass1", 3),
+    rep("fiz_ass2", 3),
+    rep("fiz_ass3", 3),
+    rep("fiz_d1", 3),
+    rep("fiz_d2", 3),
+    rep("fiz_d3", 3),
+    rep("fiz_acid1", 3),
+    rep("fiz_acid2", 3),
+    rep("fiz_acid3", 3),
     rep("ass_n_1", 3),
@@
 
-# View(ms$get_overview())
-```
+ms <- ms$subset_analyses(analyses = c(1:63, 73:99, 118:121))
 
-```{r, fig.cap="Total ion chromatograms for each analysis replicate."}
-# ms$plot_spectra_tic(colorBy = "replicates")
+ms$add_blank_names(rep("blank", 94))
 
-# only blanks
-# ms$plot_spectra_tic(analyses = 100:122, colorBy = "analyses")
-```
-
-```{r}
-ms <- ms$subset_analyses(analyses = c(1:63, 73:99, 118:121))
-ms$add_blank_names(rep("blank", 94))
 ms$load_spectra()
@@
 
-```{r}
+```{r ms-overview, echo=FALSE}
+ov <- ms$get_overview()[, 1:3]
+setorder(ov, replicate)
+ov$description <- c(
+  rep("Aspirine standard", 9),
+  rep("Reference background", 4),
+  rep("Caffein", 9),
+  rep("Paracetamol product", 9),
+  rep("Paracetamol with acetic acid", 9),
+  rep("Paracetamol with aspirine", 9),
+  rep("Paracetamol product", 9),
+  rep("Paracetamol product", 9),
+  rep("Paracetamol with caffein", 9),
+  rep("Paracetamol product", 9),
+  rep("Paracetamol standard", 9)
+)
+
+kable(ov, caption = "Overview of MS analyses and respective analysis replicate groups and blanks.") %>%
+kable_styling(font_size = 12, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE) %>%
+scroll_box(width = "100%", height = "400px")
+```
+
+```{r plot-ms-tic, fig.cap="Total ion chromatogram (TIC) for each analysis replicate."}
+ms$plot_spectra_tic(colorBy = "replicates")
+```
+
+```{r ms-workflow}
 ms_workflow_settings <- list(
@@
 
-```{r}
+```{r ms-run-workflow}
 ms$run_workflow()
@@
 
-```{r ,fig.cap="Deconvoluted and pre-processed spectra for each analysis."}
+```{r plot-ms-spectra, fig.cap="Pre-processed spectra for each analysis. Retention time dimension in he created bins is collapsed for each mass-to-charge ratio"}
 ms$plot_spectra()
+```
+
+```{r ms-spectra}
+ms$get_spectra()[1:10, ]
 ```
@@
 
-```{r}
+```{r raman-setup, include=FALSE}
 raman <- RamanEngine$new(raman_files)
@@
     rep("blank", 9),
-    rep("fiz3_1", 3),
-    rep("fiz3_2", 3),
-    rep("fiz3_3", 3),
+    rep("fiz_c1", 3),
+    rep("fiz_c2", 3),
+    rep("fiz_c3", 3),
     rep("fiz_ass_n_1", 3),
@@
     rep("fiz_ass_n_3", 3),
-    rep("fiz_ass_1", 3),
-    rep("fiz_ass_2", 3),
-    rep("fiz_ass_3", 3),
-    rep("fiz_caff_1", 3),
-    rep("fiz_caff_2", 3),
-    rep("fiz_caff_3", 3),
-    rep("fiz1_1", 3),
-    rep("fiz1_2", 3),
-    rep("fiz1_3", 3),
-    rep("fiz4_1", 3),
-    rep("fiz4_2", 3),
-    rep("fiz4_3", 3),
-    rep("fiz2_1", 3),
-    rep("fiz2_2", 3),
-    rep("fiz2_3", 3),
+    rep("fiz_ass1", 3),
+    rep("fiz_ass2", 3),
+    rep("fiz_ass3", 3),
+    rep("fiz_caff1", 3),
+    rep("fiz_caff2", 3),
+    rep("fiz_caff3", 3),
+    rep("fiz_a1", 3),
+    rep("fiz_a2", 3),
+    rep("fiz_a3", 3),
+    rep("fiz_d1", 3),
+    rep("fiz_d2", 3),
+    rep("fiz_d3", 3),
+    rep("fiz_b1", 3),
+    rep("fiz_b2", 3),
+    rep("fiz_b3", 3),
     rep("background", 4),
-    rep("fiz_acid_1", 3),
-    rep("fiz_acid_2", 3),
-    rep("fiz_acid_3", 3),
+    rep("fiz_acid1", 3),
+    rep("fiz_acid2", 3),
+    rep("fiz_acid3", 3),
     rep("par_1", 3),
@@
 
-# View(raman$get_overview())
+ramanQuant <- raman$subset_analyses(analyses = c(19:27, 116:160))
+ramanQuant$add_blank_names(rep("blank", 54))
+
+raman <- raman$subset_analyses(analyses = c(1:36, 46:90, 95:106, 110:115))
+raman$add_blank_names(rep("blank", 99))
 ```
 
-```{r ,fig.cap="Raw averaged Raman spectra for each BSA product between minute 5.5 and 8."}
-# raman$plot_spectra(colorBy = "replicates")
-```
-
-```{r}
-raman <- raman$subset_analyses(analyses = c(1:36, 46:90, 95:106, 110:115))
-raman$add_blank_names(rep("blank", 99))
+```{r plot-raman-raw, fig.cap="Raw averaged Raman spectra for each analysis replicate."}
 raman$plot_spectra(colorBy = "replicates")
@@
 
-```{r}
+```{r raman-workflow}
 raman_workflow_settings <- list(
@@
 
-```{r}
+```{r raman-run-workflow}
 raman$run_workflow()
@@
 
-```{r ,fig.cap="Processed Raman spectra for each BSA product."}
+```{r plot-raman-spectra, fig.cap="Pre-processed Raman spectra for analysis replicate."}
 raman$plot_spectra()
@@
 
-```{r}
+```{r fusion}
 rSpec <- raman$get_spectra()
@@
 fusedMat <- cbind(mmSpec, mrSpec)
+
+# Matrix is mean centered but not scaled
+fusedMat <- mdatools::prep.autoscale(fusedMat, center = TRUE, scale = FALSE)
 ```
 
-```{r}
+```{r plot-fused, include=FALSE, fig.cap="Fused spectra."}
 cl <- StreamFind:::.get_colors(rownames(fusedMat))
@@
 
-Using Principle Components Analysis (PCA) for unsupervised evaluation.
+Principle Components Analysis (PCA) for unsupervised evaluation.
 
-```{r}
-main_set <- fusedMat[c(8:9, 10, 12, 14:15, 17:18), ]
-test_set <- fusedMat[c(1:7, 11, 13, 16, 19:30), ]
+```{r sets-show}
+main_set <- fusedMat[c(7:8, 16, 18, 20:21, 25:27), ]
+rownames(main_set)
 
+test_set <- fusedMat[c(1:6, 9:15, 17, 19, 22:24, 28:30), ]
+rownames(test_set)
+```
+
+```{r pca-engine, fig.cap="PCA model (native plot from mdatools R package).", fig.height=9}}
 stat <- StatisticEngine$new(data = main_set)
-stat$make_model(Settings_make_model_pca_mdatools(ncomp = 3, alpha = 0.005, gamma = 0.0005))
+stat$make_model(Settings_make_model_pca_mdatools(ncomp = 2, alpha = 5E-3, gamma = 5E-4))
+plot(stat$model)
+```
+
+```{r pca-predict, fig.cap="PCA model prediction distances. For more information about the distances plese consult the mdatools guide in https://mdatools.com/docs/pca--distances-and-limits.html."}
 stat$predict(test_set)
@@
 
-Using k-nearest neighbors (KNN) for classification.
+k-nearest neighbors (KNN) for classification.
 
-```{r}
-train_set <- fusedMat[c(grep("_1", rownames(fusedMat)), grep("_2", rownames(fusedMat))), ]
+```{r sets-show-knn}
+train_set <- fusedMat[c(grep("1", rownames(fusedMat)), grep("2", rownames(fusedMat))), ]
 train_set <- train_set[order(rownames(train_set)), ]
+rownames(train_set)
 
-test_set <- fusedMat[grep("_3", rownames(fusedMat)), ]
+test_set <- fusedMat[grep("3", rownames(fusedMat)), ]
+rownames(test_set)
+```
 
-labels <- c(
-  rep("ass", 2), rep("caff", 2), rep("fiz_acid", 2), rep("fiz_ass", 2), rep("fiz_caff", 2),
-  rep("fiz_altered", 2), rep("fiz_altered", 2), rep("fiz", 2), rep("fiz", 2), rep("par", 2)
-)
+```{r knn-classes}
+labels <- c(rep("ass", 2), rep("caff", 2), rep("fiz", 2), rep("fiz_acid", 2), rep("fiz_ass", 2),
+  rep("fiz", 2), rep("fiz", 2), rep("fiz_caff", 2), rep("fiz", 2), rep("par", 2))
+```
 
+```{r knn-classes-show, echo=FALSE}
+kable(data.frame(replicate = rownames(train_set), class = labels), caption = "Clss assignment for training set.") %>%
+kable_styling(font_size = 12, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE) %>%
+scroll_box(width = "100%", height = "400px")
+```
+
+```{r knn-engine}
 stat2 <- StatisticEngine$new(data = train_set)
@@
 
-## Other
+# Raman quantification
 
-Using Hierarchical Clustering.
+```{r ov-raman-conc, include=FALSE}
+ov2 <- ramanQuant$get_overview()[, 1:3]
+setorder(ov2, replicate)
+ov2$concentration <- as.numeric(c(rep(0, 9), sub(".*?(\\d+).*", "\\1", ov2$replicate[10:nrow(ov2)]))) / 10
+
+kable(ov2, caption = "Overview of Raman analyses for quantification with respective concentration.") %>%
+kable_styling(font_size = 12, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE) %>%
+scroll_box(width = "100%", height = "400px")
+```
+
+```{r plot-raman-conc-raw, fig.cap="Raw averaged Raman spectra for each analysis replicate."}
+ramanQuant$plot_spectra(colorBy = "replicates")
+```
+
+```{r raman-conc-workflow}
+raman_workflow_settings <- list(
+  Settings_average_spectra_StreamFind(),
+  Settings_subtract_blank_spectra_StreamFind(),
+  Settings_smooth_spectra_savgol(fl = 21, forder = 4, dorder = 0),
+  Settings_delete_spectra_section_StreamFind(list("shift" = c(-100, 350))),
+  Settings_delete_spectra_section_StreamFind(list("shift" = c(1800, 2500))),
+  Settings_correct_spectra_baseline_airpls(lambda = 45, differences = 1, itermax = 30)
+)
+
+ramanQuant$add_settings(raman_workflow_settings)
+
+ramanQuant$print_workflow()
+```
+
+```{r raman-conc-run-workflow}
+ramanQuant$run_workflow()
+```
+
+```{r plot-raman-conc-spectra, fig.cap="Pre-processed Raman spectra for analysis replicate."}
+ramanQuant$plot_spectra()
+```
+
+```{r raman-quant-mat}
+rSpecQuant <- ramanQuant$get_spectra()
+rSpecQuant <- rSpecQuant[order(rSpecQuant$replicate), ]
+mrSpecQuant <- matrix(
+  rSpecQuant$intensity,
+  nrow = length(unique(rSpecQuant$replicate)),
+  ncol = length(unique(rSpecQuant$shift)),
+  byrow = TRUE,
+  dimnames = list(as.character(unique(rSpecQuant$replicate)), as.character(round(unique(rSpecQuant$shift), digits = 1)))
+)
+```
 
 ```{r}
-hc_result <- hclust(dist(fusedMat), method = "ward.D2")
-plot(hc_result)
+
+ramanQuant$plot_spectra_baseline()
+
+library(mdatools)
+m1 = mcrals(mrSpecQuant, ncomp = 2)
+plotSpectra(m1)
+
+summary(m1)
+
+conc_ov <- ov2[, c(2, 4)]
+conc_ov <- unique(conc_ov)
+
+show(head(m1$rescont))
+
+plot(conc_ov$concentration[-1], m1$rescont[, 2], type = "l")
+
+x <- conc_ov$concentration[-1]
+y <- m1$rescont[, 2]
+linear_model <- lm(x  ~  y)
+
+summary_linear_model <- summary(linear_model)
+r_squared <- summary_linear_model$r.squared
+r_squared
+
+predict(lm(m1$rescont[, 2] ~ conc_ov$concentration[-1]), data.frame(x = 372.8276, 199, 399, 349), se.fit = TRUE)
+
+predict(linear_model, data.frame(x = c(m1$rescont[, 2], 334)))
+
+new_rescont_value <- 1200  # replace with your actual new value
+
+# Create a data frame with the new value
+new_data <- data.frame(y = new_rescont_value)
+
+# Use predict to get the predicted concentration
+predicted_concentration <- predict(linear_model, newdata = new_data)
+
+# Print the predicted concentration
+print(predicted_concentration)
+
 ```
-
-## Old code
-
-```{r include=FALSE}
-# library(mdatools)
-# m3 = pca(main_set, 3, center = FALSE, scale = FALSE, info = "Unsupervised PCA model", alpha = 0.005, gamma = 0.0005)
-# plot(m3, show.labels = TRUE)
-# res = predict(m3, test_set)
-# plot(res, show.labels = TRUE)
-# c = categorize(m3, res, 3)
-# names(c) <- rownames(test_set)
-# c
-# plotResiduals(m3,
-#   ncomp = 3,
-#   res = list("train" = m3$res$cal, "new" = res),
-#   norm = TRUE, show.labels = TRUE,
-#   lim.col = c("red", "orange"), lim.lwd = c(2, 2), lim.lty = c(1, 2)
-# )
-# plotResiduals(ncomp = 1, res, show.plot = FALSE, norm = TRUE)
-# print(c)
-# is(m3$res$cal, "ldecomp")
-```

### 2024-06-11 dev/dev_paracetamol.Rmd
Update dev_paracetamol.Rmd

@@
 
+ms_dir_extra <- paste0(main_dir, "/ms/extra/mzml")
+
 ms_files <- list.files(ms_dir, pattern = ".mzML", full.names = TRUE)
+
+ms_files_extra <- list.files(ms_dir_extra, pattern = ".mzML", full.names = TRUE)
 
@@
 
+raman_dir_extra <- paste0(main_dir, "/raman/extra")
+
 raman_files <- list.files(raman_dir, pattern = ".asc", full.names = TRUE)
+
+raman_files_extra <- list.files(raman_dir_extra, pattern = ".asc", full.names = TRUE)
 ```
@@
 
+```{r raman-spectra}
+raman$get_spectra()[1:10, ]
+```
+
 # Fusion
@@
 # Statistic analysis
+
+```{r extra-data, include=FALSE}
+ms_extra <- MassSpecEngine$new(files = ms_files_extra[c(5:40, 49:60)])
+
+ms_extra$add_replicate_names(
+  c(
+    rep("fiz_e01", 3),
+    rep("fiz_e02", 3),
+    rep("fiz_e03", 3),
+    rep("fiz_e04", 3),
+    rep("fiz_e05", 3),
+    rep("fiz_e06", 3),
+    rep("fiz_e07", 3),
+    rep("fiz_e08", 3),
+    rep("fiz_e09", 3),
+    rep("fiz_e10", 3),
+    rep("fiz_e11", 3),
+    rep("fiz_e12", 3),
+    rep("blank", 12)
+  )
+)
+
+ms_extra$add_blank_names(rep("blank", 48))
+
+ms_extra$load_spectra()
+
+ms_extra$add_settings(ms_workflow_settings)
+
+ms_extra$run_workflow()
+
+ms_extra$plot_spectra()
+```
 
@@
 
-# Raman quantification
+# Quantification with Raman
 
@@
 
-```{r}
+```{r, include=FALSE, eval=FALSE}
 

### 2024-06-12 dev/demo_paracetamol.Rmd
Update articles

---
title: 'Detection of quality deviations in paracetamol products'
subtitle: |
  | Fusion of mass spectrometry and raman spectroscopy data
  | <br>
author:
  name: Ricardo Cunha
  email: cunha@iuta.de
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    fig_caption: true
    toc: true
    number_sections: true
    toc_float: true
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Fusion of Mass Spectrometry and Raman Spectroscopy Data}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(plotly)
library(data.table)
library(StreamFind)

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center", fig.width = 9, results = "asis", comment = "")

main_dir <- "E:/iSoft/Paracetamol" # StreamFind Workstation

#main_dir <- "C:/Users/apoli/Documents/iSoft/Paracetamol" # Ricardo laptop

ms_dir <- paste0(main_dir, "/ms/mzml")

ms_dir_extra <- paste0(main_dir, "/ms/extra/mzml")

ms_files <- list.files(ms_dir, pattern = ".mzML", full.names = TRUE)

ms_files_extra <- list.files(ms_dir_extra, pattern = ".mzML", full.names = TRUE)

raman_dir <- paste0(main_dir, "/raman")

raman_dir_extra <- paste0(main_dir, "/raman/extra")

raman_files <- list.files(raman_dir, pattern = ".asc", full.names = TRUE)

raman_files_extra <- list.files(raman_dir_extra, pattern = ".asc", full.names = TRUE)
```

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<br>

<br>

***

# Introduction

# Pre-processing

## MS

```{r ms-setup, include=FALSE}
ms <- MassSpecEngine$new(files = c(ms_files, ms_files_extra[c(5:40, 49:60)]), headers = list(name = "BSA Quality Evaluation Project"))

ms$add_replicate_names(
  c(rep("fiz_a1", 3),
    rep("fiz_a2", 3),
    rep("fiz_a3", 3),
    rep("fiz_b1", 3),
    rep("fiz_b2", 3),
    rep("fiz_b3", 3),
    rep("fiz_c1", 3),
    rep("fiz_c2", 3),
    rep("fiz_c3", 3),
    rep("fiz_caff1", 3),
    rep("fiz_caff2", 3),
    rep("fiz_caff3", 3),
    rep("fiz_ass1", 3),
    rep("fiz_ass2", 3),
    rep("fiz_ass3", 3),
    rep("fiz_d1", 3),
    rep("fiz_d2", 3),
    rep("fiz_d3", 3),
    rep("fiz_acid1", 3),
    rep("fiz_acid2", 3),
    rep("fiz_acid3", 3),
    rep("ass_n_1", 3),
    rep("ass_n_2", 3),
    rep("ass_n_3", 3),
    rep("ass_1", 3),
    rep("ass_2", 3),
    rep("ass_3", 3),
    rep("caff_1", 3),
    rep("caff_2", 3),
    rep("caff_3", 3),
    rep("par_1", 3),
    rep("par_2", 3),
    rep("par_3", 3),
    rep("blank", 23),
    rep("fiz_e01", 3),
    rep("fiz_e02", 3),
    rep("fiz_e03", 3),
    rep("fiz_e04", 3),
    rep("fiz_e05", 3),
    rep("fiz_e06", 3),
    rep("fiz_e07", 3),
    rep("fiz_e08", 3),
    rep("fiz_e09", 3),
    rep("fiz_e10", 3),
    rep("fiz_e11", 3),
    rep("fiz_e12", 3),
    rep("blank_e", 12)
  )
)

ms <- ms$subset_analyses(analyses = c(1:63, 73:99, 118:121, 123:130, 134:170))

ms$add_blank_names(c(rep("blank", 94), rep("blank_e", 139 - 94)))

ms$load_spectra()
```

```{r ms-overview, echo=FALSE}
ov <- ms$get_overview()[, 1:3]
setorder(ov, replicate)
ov$description <- c(
  rep("Aspirine standard", 9),
  rep("Reference background", 4),
  rep("Reference background", 9),
  rep("Caffein", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol with acetic acid", 9),
  rep("Paracetamol with aspirine", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol with caffein", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol product", 36),
  rep("Paracetamol standard", 9)
)

kable(ov, caption = "Overview of MS analyses and respective analysis replicate groups and blanks.") %>%
kable_styling(font_size = 12, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE) %>%
scroll_box(width = "100%", height = "400px")
```

```{r plot-ms-tic, fig.cap="Total ion chromatogram (TIC) for each analysis replicate."}
ms$plot_spectra_tic(analyses = ov$analysis[ov$replicate %in% "blank_e"], colorBy = "analyses", interactive = TRUE)
```

```{r ms-workflow}
ms_workflow_settings <- list(
  Settings_bin_spectra_StreamFind(bins = list("rt" = 5, "mz" = 2), refBinAnalysis = 1),
  Settings_normalize_spectra_minmax(),
  Settings_average_spectra_StreamFind(),
  Settings_subtract_blank_spectra_StreamFind(negativeToZero = TRUE),
  Settings_normalize_spectra_blockweight()
)

ms$add_settings(ms_workflow_settings)

ms$print_workflow()
```

```{r ms-run-workflow}
ms$run_workflow()
```

```{r plot-ms-spectra, fig.cap="Pre-processed spectra for each analysis. Retention time dimension in he created bins is collapsed for each mass-to-charge ratio"}
ms$plot_spectra()
```

```{r ms-spectra}
ms$get_spectra()[1:10, ]
```

```{r}
ms$get_spectra_matrix()
```

## Raman

```{r raman-setup, include=FALSE}
raman_all <- RamanEngine$new(c(raman_files, raman_files_extra))

raman_all$add_replicate_names(
  c(
    rep("ass_1", 3),
    rep("ass_2", 3),
    rep("ass_3", 3),
    rep("caff_1", 3),
    rep("caff_2", 3),
    rep("caff_3", 3),
    rep("blank", 9),
    rep("fiz_c1", 3),
    rep("fiz_c2", 3),
    rep("fiz_c3", 3),
    rep("fiz_ass_n_1", 3),
    rep("fiz_ass_n_2", 3),
    rep("fiz_ass_n_3", 3),
    rep("fiz_ass1", 3),
    rep("fiz_ass2", 3),
    rep("fiz_ass3", 3),
    rep("fiz_caff1", 3),
    rep("fiz_caff2", 3),
    rep("fiz_caff3", 3),
    rep("fiz_a1", 3),
    rep("fiz_a2", 3),
    rep("fiz_a3", 3),
    rep("fiz_d1", 3),
    rep("fiz_d2", 3),
    rep("fiz_d3", 3),
    rep("fiz_b1", 3),
    rep("fiz_b2", 3),
    rep("fiz_b3", 3),
    rep("background", 4),
    rep("blank_e", 9),
    rep("fiz_acid1", 3),
    rep("fiz_acid2", 3),
    rep("fiz_acid3", 3),
    rep("fiz_e01", 3),
    rep("fiz_e02", 3),
    rep("fiz_e03", 3),
    rep("fiz_e04", 3),
    rep("fiz_e05", 3),
    rep("fiz_e06", 3),
    rep("fiz_e07", 3),
    rep("fiz_e08", 3),
    rep("fiz_e09", 3),
    rep("fiz_e10", 3),
    rep("fiz_e11", 3),
    rep("fiz_e12", 3),
    rep("par_1", 3),
    rep("par_odd", 3),
    rep("par_3", 3),
    rep("par_2", 3),
    rep("cali_150_mgml", 9),
    rep("cali_010_mgml", 9),
    rep("cali_025_mgml", 9),
    rep("cali_050_mgml", 9),
    rep("cali_075_mgml", 9),
    rep("background", 4)
  )
)

raman <- raman_all$subset_analyses(analyses = c(1:36, 46:90, 95:151, 155:160))
raman$add_blank_names(c(rep("blank", 81), rep("blank_e", 9), rep("blank", 9), rep("blank_e", 36), rep("blank", 9)))
```

```{r plot-raman-raw, fig.cap="Raw averaged Raman spectra for each analysis replicate."}
raman$plot_spectra(colorBy = "replicates")
```

```{r raman-workflow}
raman_workflow_settings <- list(
  Settings_average_spectra_StreamFind(),
  Settings_subtract_blank_spectra_StreamFind(),
  Settings_smooth_spectra_savgol(fl = 21, forder = 4, dorder = 0),
  Settings_delete_spectra_section_StreamFind(list("shift" = c(-100, 350))),
  Settings_delete_spectra_section_StreamFind(list("shift" = c(1800, 2500))),
  Settings_correct_spectra_baseline_airpls(lambda = 45, differences = 1, itermax = 30),
  Settings_normalize_spectra_minmax(),
  Settings_normalize_spectra_blockweight()
)

raman$add_settings(raman_workflow_settings)

raman$print_workflow()
```

```{r raman-run-workflow}
raman$run_workflow()
```

```{r plot-raman-spectra, fig.cap="Pre-processed Raman spectra for analysis replicate."}
raman$plot_spectra()
```

```{r raman-spectra}
raman$get_spectra()[1:10, ]
```

# Fusion

```{r fusion}
ms_spec_mat <- ms$get_spectra_matrix()
raman_spec_mat <- raman$get_spectra_matrix()
fused_mat <- cbind(ms_spec_mat, raman_spec_mat)
# The matrix is mean centered using the mdatools R package
fused_mat <- mdatools::prep.autoscale(fused_mat, center = TRUE, scale = FALSE)
```

```{r plot-fused, include=FALSE, fig.cap="Fused spectra."}
cl <- StreamFind:::.get_colors(rownames(fused_mat))
fig <- plot_ly()
xVal <- seq_len(length(fused_mat[1, ]))
for (i in seq_len(nrow(fused_mat))) {
  fig <- fig %>% add_trace(
    x = xVal,
    y = fused_mat[i, ],
    type = "scatter", mode = "lines",
    line = list(width = 0.5, color = unname(cl[i])),
    name = names(cl)[i],
    legendgroup = names(cl)[i],
    showlegend = TRUE
  )
}
xaxis <- list(
  linecolor = toRGB("black"),
  linewidth = 2, title = "Bins",
  titlefont = list(size = 12, color = "black")
)
yaxis <- list(
  linecolor = toRGB("black"),
  linewidth = 2, title = "Intensity",
  titlefont = list(size = 12, color = "black")
)
fig <- fig %>% plotly::layout(xaxis = xaxis, yaxis = yaxis)
fig
```

# Statistic analysis

## Unsupervised

Principle Components Analysis (PCA) can be used for unsupervised evaluation, meaning that we assume that the main (or train) set of analyses are of good quality and we test new analyses to evaluate the proximity to the main set. One could say that it is a supervised approach but in the sense that we assume the quality of the main set. Yet, we are not classifying the possible deviation of the new analyses which would be a clear supervised approach, as we show in the next sub chapter.

```{r sets-show}
main_set <- fused_mat[c(7:9, 16:21, 25:33), ]
rownames(main_set)

test_set <- fused_mat[c(1:6, 10:15, 22:24, 34:42), ]
rownames(test_set)
```

```{r pca-engine, fig.cap="PCA model (native plot from mdatools R package).", fig.height=9}
stat <- StatisticEngine$new(data = main_set)
# the alpha and gamma parameters are used to evaluate the model limits
stat$make_model(Settings_make_model_pca_mdatools(ncomp = 3, alpha = 1E-4, gamma = 1E-5))
# original plot from mdatools R package
plot(stat$model)
```

```{r pca-predict, fig.cap="PCA model prediction distances. For more information about the distances plese consult the mdatools guide in https://mdatools.com/docs/pca--distances-and-limits.html."}
stat$predict(test_set)
stat$plot_predicted_distances()
```

## Supervised (classification)

The k-nearest neighbors (KNN) can be used for classification of analyses based on pre-classified training analyses.

```{r sets-show-knn}
train_set <- fused_mat[unique(c(grep("1", rownames(fused_mat)), grep("2", rownames(fused_mat)))), ]
train_set <- train_set[order(rownames(train_set)), ]
rownames(train_set)

test_set <- fused_mat[unique(grep("3|4|5|6|7|8|9", rownames(fused_mat))), ]
rownames(test_set)
```

```{r knn-classes}
labels <- c(rep("ass", 2), rep("caff", 2), rep("fiz", 2), rep("fiz_acid", 2), rep("fiz_ass", 2),
  rep("fiz", 2), rep("fiz", 2), rep("fiz_caff", 2), rep("fiz", 7), rep("par", 2))
```

```{r knn-classes-show, echo=FALSE}
kable(data.frame(replicate = rownames(train_set), class = labels), caption = "Class assignment for training set.") %>%
kable_styling(font_size = 12, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE) %>%
scroll_box(width = "100%", height = "400px")
```

```{r knn-engine}
stat2 <- StatisticEngine$new(data = train_set)
stat2$add_classes(labels)
stat2$prepare_classification(Settings_prepare_classification_knn(k = 3))
stat2$classify(test_set)
stat2$classification_results
```

# Quantification with Raman

```{r ov-raman-conc, include=FALSE}
raman_quant <- raman_all$subset_analyses(analyses = c(19:27, 161:205, 155:157))
raman_quant$add_blank_names(rep("blank", 57))

ov_quant <- copy(raman_quant$get_overview()[, 1:3])
setorder(ov_quant, replicate)
ov_quant$concentration <- as.numeric(c(rep(0, 9), sub(".*?(\\d+).*", "\\1", ov_quant$replicate[10:nrow(ov2)]))) / 10
ov_quant$concentration[ov_quant$replicate %in% "par_3"] <- NA_real_

kable(ov_quant, caption = "Overview of Raman analyses for quantification with respective concentration.") %>%
kable_styling(font_size = 12, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE) %>%
scroll_box(width = "100%", height = "400px")
```

```{r plot-raman-conc-raw, fig.cap="Raw averaged Raman spectra for each analysis replicate."}
raman_quant$plot_spectra(colorBy = "replicates")
```

```{r raman-conc-workflow}
raman_workflow_settings <- list(
  Settings_average_spectra_StreamFind(),
  Settings_subtract_blank_spectra_StreamFind(),
  Settings_smooth_spectra_savgol(fl = 21, forder = 4, dorder = 0),
  Settings_delete_spectra_section_StreamFind(list("shift" = c(-100, 350))),
  Settings_delete_spectra_section_StreamFind(list("shift" = c(1800, 2500))),
  Settings_correct_spectra_baseline_airpls(lambda = 45, differences = 1, itermax = 30)
)

raman_quant$add_settings(raman_workflow_settings)

raman_quant$print_workflow()
```

```{r raman-conc-run-workflow}
raman_quant$run_workflow()
```

```{r plot-raman-conc-spectra, fig.cap="Pre-processed Raman spectra for analysis replicate."}
raman_quant$plot_spectra()
```

## Applying MCR-ALS

The quantification of paracetamol from the Raman spectra is performed using the Multivariate Curve Resolution (MCR) with Alternating Least Squares (ALS) method via integration of the [mdatools](https://mdatools.com/docs/mcr--als.html) R package.

```{r mcrals-show}
stat3 <- StatisticEngine$new(data = raman_quant$get_spectra_matrix())
stat3$add_concentrations(c(1, 2.5, 5.0, 7.5, 15.0, NA_real_)) # add concentrations,with NA for unknown
stat3$make_model(Settings_make_model_mcrals_mdatools(ncomp = 3))
summary(stat3$model)
```

```{r mcrals-resolved-spec, fig.cap="Resolved spectra for each component and the spectra for each analysis replicate (i.e., each concentration)."}
stat3$plot_model_resolved_spectra()
```

```{r mcrals-cont, fig.cap="Contributions for each analysis replicate."}
stat3$plot_model_contributions()
```

The `quantify` method can be used to estimate the concentration of the paracetamol analyses with unknown concentration. The quantification is applied to each component, returning a list with the r square of the linear model and the estimated concentration as `data.table`. Note that the explained variance for each component is important to evaluate the calculated concentration. The best estimation is from the component with the combined explained variance and r square closer to 100 and 1, respectively.

```{r}
stat3$quantify()
```

### 2024-07-30 R/class_R6_CoreEngine.R
Improved processing settings naming

@@ -116,15 +116,16 @@ CoreEngine <- R6::R6Class("CoreEngine",
     
     # Extracts and validates ProcessingSettings for a given call.
     #
-    .get_call_settings = function(settings, call) {
+    .get_call_settings = function(settings, engine, call) {
       
       checkmate::assert_choice(call, self$processing_methods()$name)
+      checkmate::assert_choice(paste0(engine,"Engine"), is(self))
       
       if (is.null(settings)) settings <- self$get_settings(call)
       
       if (is.null(settings)) return(NULL)
       
-      cols_check <- c("call", "algorithm", "parameters")
+      cols_check <- c("engine", "call", "algorithm", "parameters")
       
       if (all(cols_check %in% names(settings))) settings <- list(settings)
       
@@ -1071,7 +1072,9 @@ CoreEngine <- R6::R6Class("CoreEngine",
         lapply(self$get_settings(), function(x) {
           call <- x$call
           message("\U2699 Running ", call, " with ", x$algorithm)
-          do.call(self[[call]], list("settings" = x))
+          .dispatch_process_method(x$engine, x$call, x, self, private)
+          
+          # do.call(self[[call]], list("settings" = x))
         })
         
       } else {

### 2024-07-30 R/class_R6_MassSpecEngine.R

@@ -5931,9 +5931,9 @@ MassSpecEngine <- R6::R6Class("MassSpecEngine",
     #'
     load_features_eic = function(settings) {
       
-      if (missing(settings)) settings <- Settings_load_features_eic_StreamFind()
+      if (missing(settings)) settings <- MassSpecSettings_LoadFeaturesEIC_StreamFind()
       
-      .dispatch_process_method("ms_load_features_eic", settings, self, private)
+      .dispatch_process_method("MassSpec", "LoadFeaturesEIC", settings, self, private)
       
       invisible(self)
     },
@@ -5944,9 +5944,9 @@ MassSpecEngine <- R6::R6Class("MassSpecEngine",
     #'
     load_features_ms1 = function(settings) {
       
-      if (missing(settings)) settings <- Settings_load_features_ms1_StreamFind()
+      if (missing(settings)) settings <- MassSpecSettings_LoadFeaturesMS1_StreamFind()
       
-      .dispatch_process_method("ms_load_features_ms1", settings, self, private)
+      .dispatch_process_method("MassSpec", "LoadFeaturesMS1", settings, self, private)
       
       invisible(self)
     },
@@ -5957,9 +5957,9 @@ MassSpecEngine <- R6::R6Class("MassSpecEngine",
     #'
     load_features_ms2 = function(settings) {
       
-      if (missing(settings)) settings <- Settings_load_features_ms2_StreamFind()
+      if (missing(settings)) settings <- MassSpecSettings_LoadFeaturesMS2_StreamFind()
       
-      .dispatch_process_method("ms_load_features_ms2", settings, self, private)
+      .dispatch_process_method("MassSpec", "LoadFeaturesMS2", settings, self, private)
       
       invisible(self)
     },
@@ -5970,9 +5970,9 @@ MassSpecEngine <- R6::R6Class("MassSpecEngine",
     #'
     load_MSPeakLists = function(settings) {
       
-      if (missing(settings)) settings <- Settings_load_MSPeakLists_StreamFind()
+      if (missing(settings)) settings <- MassSpecSettings_LoadMSPeakLists_StreamFind()
       
-      .dispatch_process_method("ms_load_MSPeakLists", settings, self, private)
+      .dispatch_process_method("MassSpec", "LoadMSPeakLists", settings, self, private)
       
       invisible(self)
     },
@@ -5983,18 +5983,18 @@ MassSpecEngine <- R6::R6Class("MassSpecEngine",
     #'
     centroid_spectra = function(settings = NULL) {
       
[…]

### 2024-07-30 R/methods_S3_ms_annotate_features.R

@@ -1,11 +1,11 @@
 
-#' @title .s3_ms_annotate_features.Settings_annotate_features_StreamFind
+#' @title .s3_AnnotateFeatures.MassSpecSettings_AnnotateFeatures_StreamFind
 #'
 #' @description Annotates features with isotopes.
 #'
 #' @noRd
 #'
-.s3_ms_annotate_features.Settings_annotate_features_StreamFind <- function(settings, self, private) {
+.s3_AnnotateFeatures.MassSpecSettings_AnnotateFeatures_StreamFind <- function(settings, self, private) {
 
   if (!any(self$has_features())) {
     warning("Features were not found! Run find_features method first!")
@@ -53,7 +53,7 @@
     
     names(temp_i) <- temp_i_fts
     
-    temp_f <- copy(features[[x]])
+    temp_f <- data.table::copy(features[[x]])
     
     temp_f_fts <- temp_f$feature
     

### 2024-07-30 R/methods_S3_ms_find_features.R

@@ -1,11 +1,11 @@
 
-#' @title .s3_ms_find_features.patRoon
+#' @title .s3_FindFeatures.patRoon
 #'
 #' @description Finds features using the package patRoon.
 #'
 #' @noRd
 #'
-.s3_ms_find_features.patRoon <- function(settings, self, private) {
+.s3_FindFeatures.patRoon <- function(settings, self, private) {
   
   anaInfo <- self$analysisInfo
   
@@ -104,13 +104,13 @@
   TRUE
 }
 
-#' @title .s3_ms_find_features.Settings_find_features_qPeaks
+#' @title .s3_FindFeatures.MassSpecSettings_FindFeatures_qPeaks
 #'
 #' @description Finds features using the algorithm qPeaks.
 #'
 #' @noRd
 #'
-.s3_ms_find_features.Settings_find_features_qPeaks <- function(settings, self, private) {
+.s3_FindFeatures.MassSpecSettings_FindFeatures_qPeaks <- function(settings, self, private) {
   
   message("Finding features with qPeaks...", appendLF = TRUE)
   

### 2024-07-31 README.Rmd
Updated website

@@
 
-The StreamFind project, entitled "Flexible data analysis and workflow designer to identify chemicals in the water cycle", is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de) and is a cooperation between the [Institut für Umwelt & Energie, Technik & Analytik e.
+The StreamFind is developed under the project ["Flexible data analysis and workflow designer to identify chemicals in the water cycle"](https://www.bildung-forschung.digital/digitalezukunft/de/wissen/Datenkompetenzen/datenkompetenzen_wissenschaftlichen_nachwuchs/Projekte/stream_find.html) which is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de).
+The development is conducted by the [Institut für Umwelt & Energie, Technik & Analytik e.
 V.
 (IUTA)](https://www.iuta.de), the [Forschungszentrum Informatik (FZI)](https://www.fzi.de/) and supporting partners.
-The goal of the StreamFind project is the development and assembly of data processing workflows for mass spectrometry and spectroscopy and the application of the workflows in environmental and quality studies of the water cycle.
-The StreamFind aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data processing and structure and easing the retrospective evaluation of data.
-The StreamFind platform is directed to academics but also technicians, due to the aspired comprehensive documentation, well categorized set of integrated modular functions and the graphical user interface.
+The goal of the StreamFind is the development and assembly of data processing workflows for different types of analytical data (e.g., mass spectrometry and spectroscopy) and the application of the workflows in diverse fields (e.g., environmental and quality studies of the water cycle).
+The StreamFind aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data structure and processing and easing the retrospective evaluation of data.
+The StreamFind platform is directed to academics but also technicians, due to the aspired comprehensive documentation, well categorized set of integrated modular functions and the embedded shiny graphical user interface.
 <br>
-The library of [StreamFind](https://github.com/odea-project/StreamFind) is an R package (this repository).
-The graphical user interface (GUI) is a Web App developed in JavaScript and can be found in the [StreamFind-App](https://github.com/odea-project/StreamFind-App) repository.
 The StreamFind development is ongoing, please [contact us](mailto:cunha@iuta.de) for questions  or collaboration.
@@
 
-For installation of the StreamFind R package, it is recommended to first install the dependencies.
-Besides [R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) (the latter is only recommended for Windows users), the StreamFind depends on the [patRoon](https://github.com/rickhelmus/patRoon) R package and its dependencies.
+Assuming that the [R](https://cran.r-project.org/) software is installed, the StreamFind R package can be installed from the GitHub repository via the [BiocManager](https://www.bioconductor.org/install/).
+
+```{r install-streamfind, eval=FALSE}
+if (!require("BiocManager", quietly = TRUE))
+  install.packages("BiocManager")
+BiocManager::install("odea-project/StreamFind", dependencies = TRUE)
+```
+
+<!-- For installation of the StreamFind R package, it is recommended to first install the dependencies.
+Besides [R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) (the latter is only applicable for Windows users), the StreamFind depends on the [patRoon](https://github.com/rickhelmus/patRoon) R package and its dependencies.
 The patRoon R package combines several tools for basic and advanced data processing and can be used interchangeably with the StreamFind R package.
 Installation instructions for patRoon and its dependencies can be found [here](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites).
+-->
 
-Then, the StreamFind R package can be installed from the GitHub repository.
+<!-- Then, the StreamFind R package can be installed from the GitHub repository. -->
 
 ```{r install-streamfind, eval=FALSE}
-remotes::install_github("odea-project/StreamFind", dependencies = TRUE)
+#remotes::install_github("odea-project/StreamFind", dependencies = TRUE)
 ```
 
-The supplementary [StreamFindData](https://github.com/odea-project/StreamFindData) R package holds the data used in examples and other documentation assets of the StreamFind R package and can also be installed from the GitHub repository.
+<!-- The supplementary [StreamFindData](https://github.com/odea-project/StreamFindData) R package holds the data used in examples and other documentation assets of the StreamFind R package and can also be installed from the GitHub repository.
+-->
 
 ```{r install-streamfinddata, eval=FALSE}
-remotes::install_github("odea-project/StreamFindData")
+#remotes::install_github("odea-project/StreamFindData")
 ```
@@
 
-The documentation and usage examples of the StreamFind R package can be found in the [reference page](https://odea-project.github.io/StreamFind/reference/index.html) and [articles](https://odea-project.github.io/StreamFind/articles/index.html) of the [webpage](https://odea-project.github.io/StreamFind/index.html).
+The documentation and usage examples of the StreamFind R package can be found in the [reference page](https://odea-project.github.io/StreamFind/reference/index.html) and [articles](https://odea-project.github.io/StreamFind/articles/index.html) of the [webpage](https://odea-project.github.io/StreamFind/index.html), respectively.
 

### 2024-08-01 README.Rmd
Updated settings, articles and core

@@
 
-The StreamFind is developed under the project ["Flexible data analysis and workflow designer to identify chemicals in the water cycle"](https://www.bildung-forschung.digital/digitalezukunft/de/wissen/Datenkompetenzen/datenkompetenzen_wissenschaftlichen_nachwuchs/Projekte/stream_find.html) which is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de).
-The development is conducted by the [Institut für Umwelt & Energie, Technik & Analytik e.
+The StreamFind is developed within the project ["Flexible data analysis and workflow designer to identify chemicals in the water cycle"](https://www.bildung-forschung.digital/digitalezukunft/de/wissen/Datenkompetenzen/datenkompetenzen_wissenschaftlichen_nachwuchs/Projekte/stream_find.html), which is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de).
+The development is carried out by the [Institut für Umwelt & Energie, Technik & Analytik e.
 V.
 (IUTA)](https://www.iuta.de), the [Forschungszentrum Informatik (FZI)](https://www.fzi.de/) and supporting partners.
-The goal of the StreamFind is the development and assembly of data processing workflows for different types of analytical data (e.g., mass spectrometry and spectroscopy) and the application of the workflows in diverse fields (e.g., environmental and quality studies of the water cycle).
-The StreamFind aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, promoting standardization of data structure and processing and easing the retrospective evaluation of data.
-The StreamFind platform is directed to academics but also technicians, due to the aspired comprehensive documentation, well categorized set of integrated modular functions and the embedded shiny graphical user interface.
+The goal of the StreamFind is to develop and assemble data processing workflows for different types of analytical data (e.g., mass spectrometry and spectroscopy) and to apply the workflows in different fields (e.g., environmental and quality studies of the water cycle).
+StreamFind aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, to promote standardization of data structure and processing, and to facilitate retrospective data evaluation.
+The StreamFind platform is aimed at scientists, but also at technicians, due to its comprehensive documentation, its well categorized set of integrated modular functions and its embedded graphical user interface.
 <br>
@@
 
-Assuming that the [R](https://cran.r-project.org/) software is installed, the StreamFind R package can be installed from the GitHub repository via the [BiocManager](https://www.bioconductor.org/install/).
+Pre-requisites for the StreamFind are the [R](https://cran.r-project.org/) software and the [RTools](https://cran.r-project.org/bin/windows/Rtools/) (only applicable for Windows users).
+RTools is needed for compiling the C++ code used in the StreamFind R package.
+Assuming that R and RTools are installed, the StreamFind R package can be installed from the GitHub repository via the [BiocManager](https://www.bioconductor.org/install/).
 
@@
 
-<!-- For installation of the StreamFind R package, it is recommended to first install the dependencies.
-Besides [R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) (the latter is only applicable for Windows users), the StreamFind depends on the [patRoon](https://github.com/rickhelmus/patRoon) R package and its dependencies.
-The patRoon R package combines several tools for basic and advanced data processing and can be used interchangeably with the StreamFind R package.
+## Other dependencies
+
+The StreamFind depends on other open source software to process different analytical data.
+For instance, for non-target screening using mass spectrometry the StreamFind uses the [patRoon](https://github.com/rickhelmus/patRoon) R package and its own dependencies.
 Installation instructions for patRoon and its dependencies can be found [here](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites).
--->
+Consult the documentation for dependencies of other data types.
 
-<!-- Then, the StreamFind R package can be installed from the GitHub repository. -->
+## Suplementary data
 
-```{r install-streamfind, eval=FALSE}
-#remotes::install_github("odea-project/StreamFind", dependencies = TRUE)
+The supplementary [StreamFindData](https://github.com/odea-project/StreamFindData) R package holds the data used in examples and other documentation assets of the StreamFind and can also be installed from the GitHub repository.
+
+```{r install-streamfinddata, eval=FALSE}
+if (!require("BiocManager", quietly = TRUE))
+  install.packages("BiocManager")
+BiocManager::install("odea-project/StreamFindData")
 ```
 
-<!-- The supplementary [StreamFindData](https://github.com/odea-project/StreamFindData) R package holds the data used in examples and other documentation assets of the StreamFind R package and can also be installed from the GitHub repository.
--->
-
-```{r install-streamfinddata, eval=FALSE}
-#remotes::install_github("odea-project/StreamFindData")
-```
-
-### Documentation
+## Documentation
 
@@
 
-# References
+## References
 

### 2024-08-01 R/class_S3_ProcessingSettings.R

@@ -73,7 +73,8 @@ ProcessingSettings <- function(engine = NA_character_,
   patRoon_algorithms <- c("openms", "xcms", "xcms3", "envipick", "sirius", "kpic2", "safd", "GenForm", "MetFrag")
 
   if (any(vapply(patRoon_algorithms, function(a) grepl(a, x$algorithm, fixed = FALSE), FALSE))) {
-    s3_classes <- append(s3_classes, "patRoon")
+    if (x$call %in% "FindFeatures") s3_classes <- append(s3_classes, "FindFeatures_patRoon")
+    if (x$call %in% "GroupFeatures") s3_classes <- append(s3_classes, "GroupFeatures_patRoon")
   }
   
   x <- structure(x, class = s3_classes)

### 2024-09-17 dev/dev_template_report.qmd
Added annotation of adducts in NTS

---
title: "Short Report on Mass Spectrometry"
author: "Ricardo Cunha"
date: "`r format(Sys.time(), '%d %B, %Y')`"
toc: true
format:
  html:
    grid:
      sidebar-width: 300px
      body-width: 1200px
      margin-width: 300px
      gutter-width: 1.5rem
    other-links:
      - text: StreamFind
        href: https://odea-project.github.io/StreamFind/
        icon: github
    html-math-method: katex
    theme: journal
    code-tools: true
    anchor-sections: true
    code-fold: true
    code-overflow: scroll
    code-line-numbers: false
    code-copy: true
    cap-location: top
    self-contained: true
    embed-resources: true
    toc: true
    toc-location: left
execute:
  warning: false
---

```{r}
#| include: false
options(DT.options = list(
  dom = "lfrtip",
  autoWidth = TRUE,
  pageLength = 10,
  lengthMenu = c(10, 20, 50, 75, 100),
  initComplete = htmlwidgets::JS(
    "function(settings, json) {",
    "$(this.api().table().body()).css({'font-size': '12px'});",
    "$(this.api().table().body()).css({'white-space': 'nowrap'});",
    "}"
  )
))
```

```{r}
#| include: false
library(knitr)
library(kableExtra)
library(data.table)
library(DT)
library(magrittr)
library(ggplot2)
library(plotly)
library(StreamFind)
```

```{r}
#| include: false
ms_files <- StreamFindData::get_ms_file_paths()[10:27]
db <- StreamFindData::get_ms_tof_spiked_chemicals_with_ms2()
cols <- c("name", "formula", "mass", "rt", "fragments", "tag")
db <- db[, cols, with = FALSE]
dbis <- db[grepl("IS", db$tag), ]
dbsus <- db[!grepl("IS", db$tag), ]
path <- "C:/Users/apoli/Documents/example_ms_files"
```

# Analyses

```{r}
#| code-fold: show
ms <- MassSpecEngine$new(analyses = ms_files)
```

```{r}
DT::datatable(ms$analyses$info)
```

<br>  
<br>

```{r}
ms$plot_spectra_tic(levels = 1, interactive = TRUE, colorBy = "replicates", title = "Total Ion Chromatograms (TICs)")
```

<br>  
<br>

### 2024-09-20 R/class_lv3_S7_MassSpecSettings_NormalizeSpectra.R
Update spectra settings to S7

@@ -1,0 +1,492 @@
+
+# ______________________________________________________________________________________________________________________
+# minmax -----
+# ______________________________________________________________________________________________________________________
+
+#' **MassSpecSettings_NormalizeSpectra_minmax**
+#'
+#' @description Normalizes spectra using the min-max algorithm.
+#'
+#' @return A MassSpecSettings_NormalizeSpectra_minmax object.
+#'
+#' @export
+#'
+MassSpecSettings_NormalizeSpectra_minmax <- S7::new_class("MassSpecSettings_NormalizeSpectra_minmax",
+  parent = ProcessingSettings,
+  package = "StreamFind",
+  
+  constructor = function() {
+    
+    S7::new_object(ProcessingSettings(
+      engine = "MassSpec",
+      method = "NormalizeSpectra",
+      algorithm = "minmax",
+      parameters = list(),
+      number_permitted = Inf,
+      version = as.character(packageVersion("StreamFind")),
+      software = "StreamFind",
+      developer = "Ricardo Cunha",
+      contact = "cunha@iuta.de",
+      link = "https://odea-project.github.io/StreamFind",
+      doi = NA_character_
+    ))
+  },
+  
+  validator = function(self) {
+    valid <- all(
+      checkmate::test_choice(self@engine, "MassSpec"),
+      checkmate::test_choice(self@method, "NormalizeSpectra"),
+      checkmate::test_choice(self@algorithm, "minmax")
+    )
+    if (!valid) return(FALSE)
+    NULL
+  }
+)
+
+#' @export
+#' @noRd
+S7::method(run, MassSpecSettings_NormalizeSpectra_minmax) <- function(x, engine = NULL) {
+  
+  if (!is(engine, "MassSpecEngine")) {
+    warning("Engine is not a MassSpecEngine object!")
+    return(FALSE)
+  }
+  
+  if (!engine$has_analyses()) {
+    warning("There are no analyses! Not done.")
+    return(FALSE)
+  }
+  
+  if (!engine$has_spectra()) {
+    warning("No spectra results object available! Not done.")
+    return(FALSE)
+  }
+  
+  spec_list <- engine$spectra$spectra
+  
+  spec_list <- lapply(spec_list, function(z) {
+    
+    if (nrow(z) > 0) {
+      
+      if (("rt" %in% colnames(z)) && ("shift" %in% colnames(z))) {
+        temp_x <- split(z, z$rt)
+        
[…]

### 2024-09-20 R/class_lv3_S7_MassSpecSettings_SmoothSpectra.R

@@ -1,0 +1,211 @@
+
+# ______________________________________________________________________________________________________________________
+# movingaverage -----
+# ______________________________________________________________________________________________________________________
+
+#' **MassSpecSettings_SmoothSpectra_movingaverage**
+#'
+#' @description Smooths spectra using the moving average algorithm.
+#' 
+#' @param windowSize Numeric (length 1) with the window size for the moving average.
+#'
+#' @return A MassSpecSettings_SmoothSpectra_movingaverage object.
+#'
+#' @export
+#'
+MassSpecSettings_SmoothSpectra_movingaverage <- S7::new_class("MassSpecSettings_SmoothSpectra_movingaverage",
+  parent = ProcessingSettings,
+  package = "StreamFind",
+  
+  constructor = function(windowSize = 5) {
+    
+    S7::new_object(ProcessingSettings(
+      engine = "MassSpec",
+      method = "SmoothSpectra",
+      algorithm = "movingaverage",
+      parameters = list(windowSize = windowSize),
+      number_permitted = Inf,
+      version = as.character(packageVersion("StreamFind")),
+      software = "StreamFind",
+      developer = "Ricardo Cunha",
+      contact = "cunha@iuta.de",
+      link = "https://odea-project.github.io/StreamFind",
+      doi = NA_character_
+    ))
+  },
+  
+  validator = function(self) {
+    valid <- all(
+      checkmate::test_choice(self@engine, "MassSpec"),
+      checkmate::test_choice(self@method, "SmoothSpectra"),
+      checkmate::test_choice(self@algorithm, "movingaverage"),
+      checkmate::test_number(self@parameters$windowSize)
+    )
+    if (!valid) return(FALSE)
+    NULL
+  }
+)
+
+#' @export
+#' @noRd
+S7::method(run, MassSpecSettings_SmoothSpectra_movingaverage) <- function(x, engine = NULL) {
+  
+  if (!is(engine, "MassSpecEngine")) {
+    warning("Engine is not a MassSpecEngine object!")
+    return(FALSE)
+  }
+  
+  if (!engine$has_analyses()) {
+    warning("There are no analyses! Not done.")
+    return(FALSE)
+  }
+  
+  if (!engine$has_spectra()) {
+    warning("No spectra results object available! Not done.")
+    return(FALSE)
+  }
+  
+  windowSize <- x$parameters$windowSize
+  
[…]

### 2024-09-20 R/methods_S3_ms_bin_spectra.R

@@ -1,213 +1,0 @@
-
-#' @noRd
-.process.MassSpecSettings_BinSpectra_qBinning <- function(settings, self, private) {
-  
-  # message("Binning spectra with qBinning...", appendLF = TRUE)
-  # 
-  # if (self$get_number_analyses() == 0) {
-  #   warning("There are no analyses! Add MS analyses as mzML or mzXML files!")
-  #   return(FALSE)
-  # }
-  # 
-  # if (!any(self$has_loaded_spectra())) self$load_spectra()
-  # 
-  # if (!any(self$has_loaded_spectra())) {
-  #   warning("Spectra not found in MS analyses.")
-  #   return(FALSE)
-  # }
-  # 
-  # spectra <- self$get_spectra()
-  # 
-  # # TODO Max Implementation qBinning method
-  # 
-  # # self$add_features_eic(eics, replace = TRUE)
-  
-  FALSE
-}
-
-
-
-#' @noRd
-.process.MassSpecSettings_BinSpectra_StreamFind <- function(settings, self, private) {
-  
-  if (self$has_spectra()) {
-    spec_list <- self$spectra
-    
-  } else {
-    warning("Spectra not found! Not done.")
-    return(FALSE)
-  }
-  
-  unitsVal = settings$parameters$unitsVal
-  
-  unitsNumber = settings$parameters$unitsNumber
-  
-  bins = settings$parameters$bins
-  
-  refBinAnalysis = settings$parameters$refBinAnalysis
-  
-  cache <- .load_chache("bin_spectra", spec_list, settings)
-  
-  if (!is.null(cache$data)) {
-    message("\U2139 Spectra loaded from cache!")
-    self$spectra <- cache$data
-    message(paste0("\U2713 ", "Spectra binned!"))
-    return(TRUE)
-  }
-  
-  useRefBins <- FALSE
-  ref_bins_seq_list <- NULL
-  ref_bin_matrix <- NULL
-  ref_bin_key <- NULL
-  . <- NULL
-  
-  if (!is.null(refBinAnalysis)) {
-    if (is.numeric(refBinAnalysis) && length(refBinAnalysis) == 1) {
-      refBinAnalysis <- self$get_analysis_names()[refBinAnalysis]
-    }
-    
-    if (!is.character(refBinAnalysis)) {
-      warning("Reference analysis not found! Not done.")
-      return(FALSE)
-    }
-    
-    refSpec <- spec_list[[refBinAnalysis]]
-    
[…]

### 2024-09-22 R/class_lv3_S7_MassSpecSettings_DeconvoluteSpectra.R
Major file renaming and improved Raman processing

@@ -1,150 +1,0 @@
-
-# ______________________________________________________________________________________________________________________
-# StreamFind -----
-# ______________________________________________________________________________________________________________________
-
-#' **MassSpecSettings_DeconvoluteSpectra_StreamFind**
-#'
-#' @description Deconvolutes the spectral mass-to-charge ratio (\emph{m/z}) to mass (Da) after assignment of charges.
-#' 
-#' @param clustVal Numeric (length 1) with the clustering value for the charge deconvolution.
-#' @param window Optional numeric (length 1) with the window in \emph{m/z} for collecting traces of a given charge.
-#'
-#' @return A MassSpecSettings_DeconvoluteSpectra_StreamFind object.
-#'
-#' @export
-#'
-MassSpecSettings_DeconvoluteSpectra_StreamFind <- S7::new_class("MassSpecSettings_DeconvoluteSpectra_StreamFind",
-  parent = ProcessingSettings,
-  package = "StreamFind",
-  
-  constructor = function(clustVal = 0.1, window = 20) {
-    
-    S7::new_object(ProcessingSettings(
-      engine = "MassSpec",
-      method = "DeconvoluteSpectra",
-      algorithm = "StreamFind",
-      parameters = list(
-        clustVal = clustVal,
-        window = window
-      ),
-      number_permitted = 1,
-      version = as.character(packageVersion("StreamFind")),
-      software = "StreamFind",
-      developer = "Ricardo Cunha",
-      contact = "cunha@iuta.de",
-      link = "https://odea-project.github.io/StreamFind",
-      doi = NA_character_
-    ))
-  },
-  
-  validator = function(self) {
-    valid <- all(
-      checkmate::test_choice(self@engine, "MassSpec"),
-      checkmate::test_choice(self@method, "DeconvoluteSpectra"),
-      checkmate::test_choice(self@algorithm, "StreamFind"),
-      checkmate::test_number(self@parameters$clustVal),
-      checkmate::test_number(self@parameters$window, null.ok = TRUE)
-    )
-    if (!valid) return(FALSE)
-    NULL
-  }
-)
-
-#' @export
-#' @noRd
-S7::method(run, MassSpecSettings_DeconvoluteSpectra_StreamFind) <- function(x, engine = NULL) {
-  
[…]

### 2024-09-24 R/class_MassSpecSettings_LoadFeaturesMS1_MS2_EIC.R
Improvement of load features eic, ms1 and ms2 with cpp

@@ -88,15 +88,22 @@ S7::method(run, MassSpecSettings_LoadFeaturesMS1_StreamFind) <- function(x, engi
   
   nts <- engine$nts
   
+  feature_list <- nts$feature_list
+  
+  feature_list <- lapply(feature_list, function(z) {
+    if (!"ms1" %in% colnames(z)) z$ms1 <- rep(list(), nrow(z))
+    z
+  })
+  
   parameters <- x$parameters
   
-  cache <- .load_chache("load_features_ms1", nts$features, x)
+  cache <- .load_chache("load_features_ms1", feature_list, x)
   
   if (!is.null(cache$data)) {
+    feature_list <- cache$data
     tryCatch({
-      nts <- .add_features_column(nts, "ms1", cache$data)
-      engine$nts <- nts
-      message("\U2139 Features MS1 spectra loaded from cache!")
+      engine$nts$feature_list <- feature_list
+      message("\U2139 Features MS1 loaded from cache!")
       return(TRUE)
     }, error = function(e) {
       warning(e)
@@ -104,54 +111,30 @@ S7::method(run, MassSpecSettings_LoadFeaturesMS1_StreamFind) <- function(x, engi
     })
   }
   
-  ms1 <- engine$get_features_ms1(
+  feature_list <- rcpp_ms_load_features_ms1(
+    analyses = engine$analyses$analyses,
+    features = feature_list,
+    filtered = parameters$filtered,
     rtWindow = parameters$rtWindow,
     mzWindow = parameters$mzWindow,
+    minTracesIntensity = parameters$minIntensity,
     mzClust = parameters$mzClust,
-    presence = parameters$presence,
-    minIntensity = parameters$minIntensity,
-    filtered = parameters$filtered,
-    useLoadedData = FALSE
+    presence = parameters$presence
   )
   
-  analyses <- engine$get_analysis_names()
-  
-  features <- nts$feature_list
-  
-  ms1_col <- lapply(analyses, function(x, ms1, features) {
-    
-    ana_ms1 <- ms1[ms1$analysis %in% x, ]
-    
-    fts_all <- features[[x]]$feature
-    
-    fts_ms1 <- lapply(fts_all, function(z, ana_ms1) {
-      
-      ft_ms1 <- ana_ms1[ana_ms1$feature %in% z, ]
-      
-      if (nrow(ft_ms1) > 0) {
-        ft_ms1[["group"]] <- NULL
-        ft_ms1[["analysis"]] <- NULL
[…]

### 2024-09-26 R/engine_StatisticEngine.R
Improved Model Results

@@ -222,6 +222,20 @@ StatisticEngine <- R6::R6Class("StatisticEngine",
       invisible(self)
     },
     
+    #' @description Tests the model using the data.
+    #' 
+    #' @note Note that the model must be created before testing and data must have the same number of variables as 
+    #' the model. Also, any pre-processing applied to the model data should be applied to the data before testing. 
+    #' Note that only numeric values are accepted in data and the data column names are used as variable names 
+    #' and data row names are used as analyses names.
+    #' 
+    #' @param data Data.frame, data-table or matrix with data.
+    #' 
+    test = function(data = NULL) {
+      self$analyses <- test(self$analyses, data)
+      invisible(self)
+    },
+    
     #' @description Classifies the data using the classification labels of the analysis.
     #' 
     #' @note Note that the classification must be prepared before using the method `prepare_classification` and data 

### 2024-09-27 R/class_Analyses_StatisticAnalyses.R
Improved statistics overall

@@ -11,7 +11,13 @@ StatisticAnalyses <- S7::new_class("StatisticAnalyses", package = "StreamFind", 
     ## __names -----
     names = S7::new_property(S7::class_character,
       getter = function(self) {
-        rownames(self@analyses)
+        analyses <- rownames(self@analyses)
+        if (self$has_model) {
+          if (self$has_test) analyses <- c(analyses, rownames(self@results[["model"]]$model$res$test$data))
+          if (self$has_prediction) analyses <- c(analyses, rownames(self@results[["model"]]$model$res$prediction$data))
+        }
+        
+        analyses
       }
     ),
     
@@ -100,18 +106,29 @@ StatisticAnalyses <- S7::new_class("StatisticAnalyses", package = "StreamFind", 
       }
     ),
     
+    ## __has_test -----
+    has_test = S7::new_property(S7::class_logical, getter = function(self) {
+      if (!self$has_model) return(FALSE)
+      if (is.null(self@results[["model"]]$model$res$test)) return(FALSE)
+      TRUE
+    }),
+    
+    ## __test -----
+    test = S7::new_property(S7::class_list, getter = function(self) {
+      if (self$has_test) return(self@results[["model"]]$model$test)
+      NULL
+    }),
+    
     ## has_prediction -----
     has_prediction = S7::new_property(S7::class_logical, getter = function(self) {
-      if (length(self) == 0) return(FALSE)
-      if (is.null(self@results[["model"]])) return(FALSE)
-      if (!is(self@results[["model"]], "StreamFind::StatisticModel")) return(FALSE)
-      if (length(self@results[["model"]]$prediction) == 0) return(FALSE)
+      if (!self$has_model) return(FALSE)
+      if (is.null(self@results[["model"]]$model$res$prediction)) return(FALSE)
       TRUE
     }),
     
     ## __prediction -----
     prediction = S7::new_property(S7::class_list, getter = function(self) {
-      if (self$has_prediction) return(self@results[["model"]]$prediction)
+      if (self$has_prediction) return(self@results[["model"]]$model$prediction)
       NULL
     })
   ),
@@ -400,6 +417,11 @@ S7::method(plot_data, StatisticAnalyses) <- function(x,
   
   mat <- x$data$data
   
+  if (x$has_model) {
+    if (x$has_test) mat <- rbind(mat, x$model$test$data)
[…]

### 2024-10-02 dev/dev_ms_parsing_via_cpp.R
linked Rhdf5lib and test function

@@ -126,9 +126,40 @@ cols <- c("name", "formula", "mass", "rt")
 
 # NTS Working group -------------------------------------------------------------
 
+library(rhdf5)
+
+mz5file <- "C:/Users/apoli/Documents/github/StreamCraft/excluded/example.mz5"
+
+res <- test_read_hdf5(mz5file)
+
+length(res)
+
+file_contents <- rhdf5::h5ls(mz5file, all = TRUE)
+
+file_contents$name
+
+h5f <- rhdf5::H5Fopen(mz5file)
+for (i in 2) { #seq_len(length(file_contents$name))
+  print(file_contents$name[i])
+  data <- rhdf5::h5read(h5f, file_contents$name[i], compoundAsDataFrame = FALSE, bit64conversion = 'bit64', read.attributes = TRUE)
+  str(data)
+}
+rhdf5::H5Fclose(h5f)
+
+file_dump <- h5dump(mz5file, load = FALSE)
+spectrum_metadata_structure <- file_dump$SpectrumMetaData
+str(spectrum_metadata_structure)
+spectrum_metadata_info <- file_contents$SpectrumMetadata
 
 
 
+
+rhdf5::H5Fclose(h5f)
+str(spectrum_metadata)
+
+#h5f$SpectrumIntensity
+
+rhdf5::h5closeAll()
 # spectra ----------------------------------------------------------------------
 # clear_cache("all")
 

### 2024-10-08 R/class_MassSpecTargets.R
Updated article NTS with small fixes to NTS

@@ -1,4 +1,3 @@
-
 #' @export
 #' @noRd
 MassSpecTargets <- S7::new_class("MassSpecTargets", package = "StreamFind",

### 2024-10-08 R/class_ProcessingSettings.R

@@ -1,4 +1,3 @@
-
 #' @export
 #' @noRd
 ProcessingSettings <- S7::new_class("ProcessingSettings", package = "StreamFind",

### 2024-10-16 dev/dev_template_dashboard.qmd
Updated MS and added dev qmd files

---
title: "Short Dashboard for Mass Spectrometry"
logo: C:/Users/cunha/Documents/007_CodeProjects/StreamFindData/inst/extdata/logo_StreamFind.png
format:
  dashboard:
    theme: lux
    self-contained: true
    embed-resources: true
    nav-buttons:
      - icon: github
        href: https://github.com/odea-project/StreamFind
editor_options:
  chunk_output_type: console
---

```{r}
#| include: false

options(DT.options = list(
  bPaginate = TRUE, 
  dom = "lfrtip",
  pageLength = 20,
  lengthMenu =c(10, 15, 20, 25, 50, 75, 100),
  language = list(info = 'Showing page _PAGE_ of _PAGES_')
))
```

```{r}
#| include: false

library(knitr)
library(kableExtra)
library(data.table)
library(DT)
library(magrittr)
library(ggplot2)
library(plotly)
library(StreamFind)
```

```{r}
#| include: false

ms_files <- StreamFindData::get_ms_file_paths()[10:27]
db <- StreamFindData::get_ms_tof_spiked_chemicals_with_ms2()
cols <- c("name", "formula", "mass", "rt", "fragments", "tag")
db <- db[, cols, with = FALSE]
dbis <- db[grepl("IS", db$tag), ]
dbsus <- db[!grepl("IS", db$tag), ]
path <- "C:/Users/apoli/Documents/example_ms_files"
```

# Analyses

## Row {.tabset}

::: {.card title="Analyses Overview"}

```{r}
#| code-fold: show

ms <- MassSpecEngine$new(analyses = ms_files)
```

```{r}
DT::datatable(ms$analyses$info)
```

:::

::: {.card title="Total Ion Chromatograms"}

```{r}
ms$plot_spectra_tic(levels = 1, interactive = TRUE, colorBy = "replicates", title = "Total Ion Chromatograms (TICs)")
```

:::

# Processing {scrolling="true"}

## Row {.tabset}

::: {.card title="Find Features"}

Extracted ion chromatograms of Metoprolol-D7 (*m/z* 274.2274) and Carbamazepine-D10 (*m/z* 246.1577) in analysis 124 are shown below.

```{r}
plot_spectra_xic(ms$analyses, analyses = 5, mass = dbis[c(4, 8), ], ppm = 50, sec = 30)
```

```{r}
ffs <- MassSpecSettings_FindFeatures_openms()
ffs
```

```{r}
ms$run(ffs)
```

```{r}
plot_features_count(ms$analyses)
```

:::

### 2024-10-25 R/class_MassSpecSettings_CalculateSpectraCharges.R
Major update to app and processing settings
Added assert instead of test

@@ -42,15 +42,12 @@ MassSpecSettings_CalculateSpectraCharges_StreamFind <- S7::new_class("MassSpecSe
   },
   
   validator = function(self) {
-    valid <- all(
-      checkmate::test_choice(self@engine, "MassSpec"),
-      checkmate::test_choice(self@method, "CalculateSpectraCharges"),
-      checkmate::test_choice(self@algorithm, "StreamFind"),
-      checkmate::test_number(self@parameters$roundVal),
-      checkmate::test_number(self@parameters$relLowCut),
-      checkmate::test_number(self@parameters$absLowCut)
-    )
-    if (!valid) return(FALSE)
+    checkmate::assert_choice(self@engine, "MassSpec")
+    checkmate::assert_choice(self@method, "CalculateSpectraCharges")
+    checkmate::assert_choice(self@algorithm, "StreamFind")
+    checkmate::assert_number(self@parameters$roundVal)
+    checkmate::assert_number(self@parameters$relLowCut)
+    checkmate::assert_number(self@parameters$absLowCut)
     NULL
   }
 )

### 2024-10-25 R/class_MassSpecSettings_SuspectScreening.R

@@ -65,27 +65,27 @@ MassSpecSettings_SuspectScreening_StreamFind <- S7::new_class("MassSpecSettings_
   },
   
   validator = function(self) {
-    valid <- all(
-      checkmate::test_choice(self@engine, "MassSpec"),
-      checkmate::test_choice(self@method, "SuspectScreening"),
-      checkmate::test_choice(self@algorithm, "StreamFind"),
-      checkmate::test_number(self@parameters$ppm),
-      checkmate::test_number(self@parameters$sec),
-      checkmate::test_number(self@parameters$ppmMS2),
-      checkmate::test_number(self@parameters$minFragments),
-      checkmate::test_number(self@parameters$isolationWindow),
-      checkmate::test_number(self@parameters$mzClust),
-      checkmate::test_number(self@parameters$presence),
-      checkmate::test_number(self@parameters$minIntensity),
-      checkmate::test_logical(self@parameters$filtered, max.len = 1)
-    ) && if (is.data.frame(self@parameters$database)) {
-      all(c("name", "mass") %in% colnames(self@parameters$database)) ||
-        all(c("name", "neutralMass") %in% colnames(self@parameters$database)) ||
-          all(c("name", "mz") %in% colnames(self@parameters$database))
+    checkmate::assert_choice(self@engine, "MassSpec")
+    checkmate::assert_choice(self@method, "SuspectScreening")
+    checkmate::assert_choice(self@algorithm, "StreamFind")
+    checkmate::assert_number(self@parameters$ppm)
+    checkmate::assert_number(self@parameters$sec)
+    checkmate::assert_number(self@parameters$ppmMS2)
+    checkmate::assert_number(self@parameters$minFragments)
+    checkmate::assert_number(self@parameters$isolationWindow)
+    checkmate::assert_number(self@parameters$mzClust)
+    checkmate::assert_number(self@parameters$presence)
+    checkmate::assert_number(self@parameters$minIntensity)
+    checkmate::assert_logical(self@parameters$filtered, max.len = 1)
+    if (is.data.frame(self@parameters$database)) {
+      if (!(
+            all(c("name", "mass") %in% colnames(self@parameters$database)) ||
+            all(c("name", "neutralMass") %in% colnames(self@parameters$database)) ||
+            all(c("name", "mz") %in% colnames(self@parameters$database))
[…]

### 2024-10-25 R/class_MassSpecTargets.R

@@ -335,16 +335,28 @@ MassSpecTargets <- S7::new_class("MassSpecTargets", package = "StreamFind",
       }
     }
     
-    if (is.null(targets)) targets <- data.table::data.table()
+    if (is.null(targets)) targets <- data.table::data.table(
+      "id" = 0,
+      "mz" = 0,
+      "rt" = 0,
+      "mobility" = 0,
+      "mzmin" = 0,
+      "mzmax" = 0,
+      "rtmin" = 0,
+      "rtmax" = 0,
+      "mobilitymin" = 0,
+      "mobilitymax" = 0,
+      "analysis" = analyses,
+      "polarity" = polarities
+    )
     
     S7::new_object(S7::S7_object(), targets = targets)
   },
     
   validator = function(self) {
-    cols <- c("id", "mz", "rt", "mobility", "mzmin", "mzmax", "rtmin", "rtmax", "mobilitymin", "mobilitymax")
-    valid <- checkmate::test_data_frame(self@targets) &&
-      if (ncol(self@targets) > 0) checkmate::test_true(all(cols %in% colnames(self@targets))) else TRUE
-    if (!valid) return(FALSE)
+    cols <- c("id", "mz", "rt", "mobility", "mzmin", "mzmax", "rtmin", "rtmax", "mobilitymin", "mobilitymax", "analysis", "polarity")
+    checkmate::assert_data_frame(self@targets)
+    checkmate::assert_true(all(cols %in% colnames(self@targets)))
     NULL
   }
 )

### 2024-10-27 dev/dev_template_dashboard.qmd

@@
 title: "Short Dashboard for Mass Spectrometry"
-logo: C:/Users/cunha/Documents/007_CodeProjects/StreamFindData/inst/extdata/logo_StreamFind.png
+logo: "`r file.path(system.file(package = 'StreamFind', dir = 'extdata'), 'logo_StreamFind.png')`"
 format:
@@
   pageLength = 20,
-  lengthMenu =c(10, 15, 20, 25, 50, 75, 100),
+  lengthMenu = c(10, 15, 20, 25, 50, 75, 100),
   language = list(info = 'Showing page _PAGE_ of _PAGES_')

### 2024-10-27 dev/dev_template_report.qmd
linked Rhdf5lib and test function

@@
 #| include: false
+
 options(DT.options = list(
@@
 #| include: false
+
 library(knitr)
@@
 #| include: false
+
 ms_files <- StreamFindData::get_ms_file_paths()[10:27]

### 2024-11-13 README.Rmd
Documentation improvements

@@
 
-# StreamFind (R package)
+<style>
+  p {
+    text-align: justify;
+  }
+</style>
 
@@
 
-<p align="center" width="100%">
+<!-- <p align="center" width="100%">
   <img width="60%" src="man/figures/logo_StreamFind.png" alt="Logo" />
-</p>
+</p> -->
 
-The StreamFind is developed within the project ["Flexible data analysis and workflow designer to identify chemicals in the water cycle"](https://www.bildung-forschung.digital/digitalezukunft/de/wissen/Datenkompetenzen/datenkompetenzen_wissenschaftlichen_nachwuchs/Projekte/stream_find.html), which is funded by the [Bundesministerium für Bildung und Forschung (BMBF)](https://www.bmbf.de).
+```{r, echo=FALSE, out.width = '100%', fig.align='center'}
+knitr::include_graphics("man/figures/logos.png")
+```
+
+StreamFind R package is developed within the project ["Flexible data analysis and workflow designer to identify chemicals in the water cycle"](https://www.bildung-forschung.digital/digitalezukunft/de/wissen/Datenkompetenzen/datenkompetenzen_wissenschaftlichen_nachwuchs/Projekte/stream_find.html) funded by the [German Federal Ministry of Education and Research (BMBF)](https://www.bmbf.de).
 The development is carried out by the [Institut für Umwelt & Energie, Technik & Analytik e.
@@
 (IUTA)](https://www.iuta.de), the [Forschungszentrum Informatik (FZI)](https://www.fzi.de/) and supporting partners.
-The goal of the StreamFind is to develop and assemble data processing workflows for different types of analytical data (e.g., mass spectrometry and spectroscopy) and to apply the workflows in different fields (e.g., environmental and quality studies of the water cycle).
-StreamFind aims to stimulate the use of advanced data analysis (e.g., non-target screening, statistical analysis, etc.) in routine studies, to promote standardization of data structure and processing, and to facilitate retrospective data evaluation.
-The StreamFind platform is aimed at scientists, but also at technicians, due to its comprehensive documentation, its well categorized set of integrated modular functions and its embedded graphical user interface.
-<br>
+StreamFind is intended to be a platform for assembling processing workflows for different types of data (e.g.
+mass spectrometry and spectroscopy) with applications in different fields (e.g.
+environmental and quality studies of the water cycle and quality control of pharmaceuticals).
+StreamFind aims to stimulate the use of advanced data analysis (e.g.
+non-target screening, statistical analysis, etc.) in routine studies, to promote standardisation of data structure and processing, and to facilitate retrospective data evaluation.
+The StreamFind platform is aimed at scientists, but also at technicians due to its comprehensive documentation, its well categorised set of integrated modular functions and its embedded graphical user interface.
+
 The StreamFind development is ongoing, please [contact us](mailto:cunha@iuta.de) for questions or collaboration.
-
-```{r, echo=FALSE, out.width = '100%', fig.align='center'}
-knitr::include_graphics("man/figures/logo_group.png")
-```
 
@@
 Pre-requisites for the StreamFind are the [R](https://cran.r-project.org/) software and the [RTools](https://cran.r-project.org/bin/windows/Rtools/) (only applicable for Windows users).
-RTools is needed for compiling the C++ code used in the StreamFind R package.
-Assuming that R and RTools are installed, the StreamFind R package can be installed from the GitHub repository via the [BiocManager](https://www.bioconductor.org/install/).
+RTools is needed for compiling C++ code used in the StreamFind R package.
+StreamFind also uses python scripts for some of its functionalities, so it is recommended to have the latest [python](https://www.python.org/downloads/) installed and added to the environmental variables for Windows users.
+Assuming that R, RTools and python are installed, the StreamFind R package can be installed from the GitHub repository via the [BiocManager](https://www.bioconductor.org/install/).
 
@@
   install.packages("BiocManager")
-BiocManager::install("odea-project/StreamFind", dependencies = TRUE)
+BiocManager::install("odea-project/StreamFind")
 ```
@@
 Installation instructions for patRoon and its dependencies can be found [here](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites).
+When a dependency is not installed, a warning message will be displayed when necessary with instructions.
 Consult the documentation for dependencies of other data types.
@@
 ```
+
+## Docker Setup
+
+The StreamFind can also be used via the [Docker](https://www.docker.com/products/docker-desktop/) container.
+The Docker container is a pre-configured environment with all the necessary dependencies installed.
+The Docker container can be built and started with the following commands.
+
+Build the Docker container:  
+`docker build -t my-r-app .`
+
+Start the Docker container:  
+`docker run -it -p 3838:3838 -p 8787:8787 -v $(pwd):/app my-r-app`
+
+Once the container is up, you'll be prompted to select the service you want to run:
+
+- *Option 1:* Starts the Shiny application, accessible at http://localhost:3838.
+- *Option 2:* Starts the RStudio Server, accessible at http://localhost:8787
+  + Default Username is rstudio and Password is rstudio
+- *Option 3:* Starts both the Shiny App and RStudio Server
 

### 2024-12-02 dev/demo_paracetamol_fusion.Rmd
Create demo_paracetamol_fusion.Rmd

---
title: 'Detect Pharmaceutical Quality Deviations with MS, Raman and Statistics'
subtitle: |
  | Fusion of Mass Spectrometry and Raman Spectroscopy Data
  | <br>
author:
  name: Ricardo Cunha
  email: cunha@iuta.de
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    fig_caption: true
    toc: true
    number_sections: true
    toc_float: true
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{'Detect Pharmaceutical Quality Deviations with MS, Raman and Statistics'}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(plotly)
library(data.table)
library(StreamFind)

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 9,
  results = "markup",
  comment = ""
)

options(
  DT.options = list(
    dom = "lfrtip",
    autoWidth = TRUE,
    pageLength = 10,
    lengthMenu = c(10, 20, 50, 75, 100),
    initComplete = htmlwidgets::JS(
      "function(settings, json) {",
      "$(this.api().table().body()).css({'font-size': '12px'});",
      "$(this.api().table().body()).css({'white-space': 'nowrap'});",
      "}"
    )
  )
)

main_dir <- "C:/Users/apoli/Documents/example_files/isoft/Paracetamol_subset" # Ricardo laptop
ms_dir <- paste0(main_dir, "/ms")
ms_files <- list.files(ms_dir, pattern = ".mzML$", full.names = TRUE)
raman_dir <- paste0(main_dir, "/raman")
raman_files <- list.files(raman_dir, pattern = ".asc$", full.names = TRUE)
```

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<br>

<br>

***

This article demonstrates how StreamFind, using mass spectrometry (MS) combined with Raman spectroscopy data, can be used to assess quality variation in the pharmaceutical Fizamol, which contains paracetamol as the active ingredient (API). Each product sample was measured by both MS and Raman spectroscopy. The data are pre-processed and fused to be used in unsupervised and supervised statistical analysis for quality assessment.

# Pre-processing

Data is pre-processed using StreamFind's built-in data-specific engines for MS and Raman data. Pre-processing workflows for each data type are performed prior to data fusion to improve statistical analysis.

## MS

```{r ms-setup, include=FALSE}
ms <- MassSpecEngine$new(analyses = ms_files)

ms$analyses$replicates <- c(
  rep("blank_01", 4),
  rep("fiz_01", 3),
  rep("fiz_02", 3),
  rep("fiz_03", 3),
  rep("fiz_04", 3),
  rep("fiz_05", 3),
  rep("fiz_06", 3),
  rep("blank_02", 8),
  rep("fiz_07", 3),
  rep("fiz_08", 3),
  rep("fiz_09", 3),
  rep("fiz_10", 3),
  rep("fiz_11", 3),
  rep("fiz_12", 3),
  rep("fiz_13", 3),
  rep("fiz_14", 3),
  rep("fiz_15", 3),
  rep("fiz_acid_1", 3),
  rep("fiz_acid_2", 3),
  rep("fiz_acid_3", 3),
  rep("fiz_ass_1", 3),
  rep("fiz_ass_2", 3),
  rep("fiz_ass_3", 3),
  rep("fiz_caff_1", 3),
  rep("fiz_caff_2", 3),
  rep("fiz_caff_3", 3),
  rep("fiz_old_1", 3),
  rep("fiz_old_2", 3),
  rep("fiz_old_3", 3)
)

ms$analyses$blanks <- c(
  rep("blank_01", 4 + 9 + 9),
  rep("blank_02", 8 + 9 + 9 + 9),
  rep("blank_01", 9 + 9 + 9 + 9)
)
```

```{r ms-overview, echo=FALSE}
ov <- ms$analyses$info[, 1:3]

ov$description <- c(
  rep("Reference background 1", 4),
  rep("Paracetamol product", 9),
  rep("Paracetamol product", 9),
  rep("Reference background 2", 8),
  rep("Paracetamol product", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol with acetic acid", 9),
  rep("Paracetamol with aspirin", 9),
  rep("Paracetamol with caffeine", 9),
  rep("Paracetamol stored", 9)
)

kable(ov, caption = "Overview of MS analyses and respective analysis replicate groups and blanks.") %>%
kable_styling(font_size = 12, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE) %>%
scroll_box(width = "100%", height = "400px")
```

<br>

```{r ms-workflow}
ms_workflow <- list(
  MassSpecSettings_LoadSpectra_StreamFind(levels = 1, mzmax = 300, rtmax = 120),
  MassSpecSettings_BinSpectra_StreamFind(binNames = c("rt", "mz"), binValues = c(5, 2), refBinAnalysis = 1),
  MassSpecSettings_NormalizeSpectra_minmax(),
  MassSpecSettings_AverageSpectra_StreamFind(),
  MassSpecSettings_SubtractBlankSpectra_StreamFind(negativeToZero = TRUE),
  MassSpecSettings_NormalizeSpectra_blockweight()
)

ms$workflow <- ms_workflow

ms$print_workflow()
```

```{r ms-run-workflow}
ms$run_workflow()
```

```{r plot-ms-spectra, fig.cap="Pre-processed spectra for each analysis. Retention time dimension in he created bins is collapsed for each mass-to-charge ratio"}
ms$plot_spectra()
```

## Raman

```{r raman-setup, include=FALSE}
raman <- RamanEngine$new(analyses = raman_files)

raman$analyses$replicates <- c(
  rep("blank_01", 9),
  rep("fiz_01", 3),
  rep("fiz_02", 3),
  rep("fiz_03", 3),
  rep("fiz_04", 3),
  rep("fiz_05", 3),
  rep("fiz_06", 3),
  rep("blank_02", 9),
  rep("fiz_07", 3),
  rep("fiz_08", 3),
  rep("fiz_09", 3),
  rep("fiz_10", 3),
  rep("fiz_11", 3),
  rep("fiz_12", 3),
  rep("fiz_13", 3),
  rep("fiz_14", 3),
  rep("fiz_15", 3),
  rep("fiz_acid_1", 3),
  rep("fiz_acid_2", 3),
  rep("fiz_acid_3", 3),
  rep("fiz_ass_1", 3),
  rep("fiz_ass_2", 3),
  rep("fiz_ass_3", 3),
  rep("fiz_caff_1", 3),
  rep("fiz_caff_2", 3),
  rep("fiz_caff_3", 3),
  rep("fiz_old_1", 3),
  rep("fiz_old_2", 3),
  rep("fiz_old_3", 3)
)

raman$analyses$blanks <- c(
  rep("blank_01", 9 + 9 + 9),
  rep("blank_02", 9 + 9 + 9 + 9),
  rep("blank_01", 9 + 9 + 9 + 9)
)
```

```{r raman-overview, echo=FALSE}
raman_ov <- raman$get_overview()[, 1:3]

raman_ov$description <- c(
  rep("Reference background 1", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol product", 9),
  rep("Reference background 2", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol product", 9),
  rep("Paracetamol with acetic acid", 9),
  rep("Paracetamol with aspirin", 9),
  rep("Paracetamol with caffeine", 9),
  rep("Paracetamol stored", 9)
)

kable(raman_ov, caption = "Overview of MS analyses and respective analysis replicate groups and blanks.") %>%
kable_styling(font_size = 12, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE) %>%
scroll_box(width = "100%", height = "400px")
```

```{r raman-workflow}
raman_workflow <- list(
  RamanSettings_AverageSpectra_StreamFind(),
  RamanSettings_SubtractBlankSpectra_StreamFind(),
  RamanSettings_SmoothSpectra_savgol(fl = 21, forder = 4, dorder = 0),
  RamanSettings_DeleteSpectraSection_StreamFind(shiftmin = -100, shiftmax = 350),
  RamanSettings_DeleteSpectraSection_StreamFind(shiftmin = 1800, shiftmax = 2500),
  RamanSettings_CorrectSpectraBaseline_airpls(lambda = 45, differences = 1, itermax = 30),
  RamanSettings_NormalizeSpectra_minmax(),
  RamanSettings_NormalizeSpectra_blockweight()
)

raman$workflow <- raman_workflow

raman$print_workflow()
```

```{r raman-run-workflow}
raman$run_workflow()
```

```{r plot-raman-spectra, fig.cap="Pre-processed Raman spectra for analysis replicate."}
raman$plot_spectra()
```

# Fusion

Statistical analysis is performed on the merged data for combination of orthogonal information in MS and Raman data.

```{r fusion}
ms_spec_mat <- ms$get_spectra_matrix()
raman_spec_mat <- raman$get_spectra_matrix()
fused_mat <- cbind(ms_spec_mat, raman_spec_mat)
# The fused matrix is mean centered using the mdatools R package
fused_mat <- mdatools::prep.autoscale(fused_mat, center = TRUE, scale = FALSE)
str(fused_mat)
```

# Statistic analysis

## Unsupervised

Principle Components Analysis (PCA) can be used for unsupervised evaluation, that is, we assume that the main (or training) set of analyses is of good quality and we test new analyses to evaluate their closeness to the main set. You could say it is a supervised approach, but in the sense that we assume the quality of the main set. However, we do not classify the possible deviation of the new analyses, which would be a clearly supervised approach, as we show in the next subchapter.

```{r sets-show}
main_set_idx <- c(1:10)

main_set <- fused_mat[main_set_idx, ]
rownames(main_set)

test_set <- fused_mat[-main_set_idx, ]
rownames(test_set)
```

```{r pca-engine, fig.cap="PCA model (native plot from mdatools R package).", fig.height=9, warning=FALSE}
stat <- StatisticEngine$new(analyses = main_set)
# The alpha and gamma parameters are used to evaluate the model limits
stat$run(StatisticSettings_MakeModel_pca_mdatools(ncomp = 3, alpha = 0.01, gamma = 0.001))
plot(stat$model, showText = TRUE)
```

```{r pca-predict, warning=FALSE, fig.cap="PCA model prediction distances. For more information about the distances plese consult the mdatools guide in https://mdatools.com/docs/pca--distances-and-limits.html."}
stat$predict(test_set)
plot(stat$model, showText = TRUE)
```

## Supervised (classification)

K-Nearest Neighbors (KNN) can be used to classify analyses based on pre-classified training analyses.

```{r sets-show-knn}
train_idx <- grep("_1|_2|_01|_04|_06|_10", rownames(fused_mat))

train_set <- fused_mat[train_idx, ]
train_set <- train_set[order(rownames(train_set)), ]
rownames(train_set)

test_set <- fused_mat[-train_idx, ]
rownames(test_set)
```

```{r knn-classes}
labels <- c(
  rep("fiz", 9),
  rep("fiz_acid", 2),
  rep("fiz_ass", 2),
  rep("fiz_caff", 2),
  rep("fiz_old", 2)
)
```

```{r knn-engine}
stat2 <- StatisticEngine$new(analyses = train_set)
stat2$add_classes(labels)
stat2$run(StatisticSettings_MakeModel_knn(k = 3))
stat2$predict(test_set)
stat2$model$prediction$results
```

# MS Only

Statistical analysis is performed on MS data only.

```{r fusion}
ms_spec_mat <- ms$get_spectra_matrix()
ms_spec_mat <- mdatools::prep.autoscale(ms_spec_mat, center = TRUE, scale = FALSE)
str(ms_spec_mat)
```

## Unsupervised

```{r sets-show}
main_set_idx <- c(1:10)
ms_main_set <- ms_spec_mat[main_set_idx, ]
rownames(ms_main_set)
ms_test_set <- ms_spec_mat[-main_set_idx, ]
rownames(ms_test_set)
```

```{r pca-engine, fig.cap="MS PCA model (native plot from mdatools R package).", fig.height=9, warning=FALSE}
ms_stat <- StatisticEngine$new(analyses = ms_main_set)
ms_stat$run(StatisticSettings_MakeModel_pca_mdatools(ncomp = 3, alpha = 0.01, gamma = 0.001))
plot(ms_stat$model, showText = TRUE)
```

```{r pca-predict, warning=FALSE, fig.cap="MS PCA model prediction distances. For more information about the distances plese consult the mdatools guide in https://mdatools.com/docs/pca--distances-and-limits.html."}
ms_stat$predict(ms_test_set)
plot(ms_stat$model, showText = TRUE)
```

## Supervised (classification)

K-Nearest Neighbors (KNN) can be used to classify analyses based on pre-classified training analyses.

```{r sets-show-knn}
train_idx <- grep("_1|_2|_01|_04|_06|_10", rownames(ms_spec_mat))

ms_train_set <- ms_spec_mat[train_idx, ]
ms_train_set <- ms_train_set[order(rownames(train_set)), ]
rownames(ms_train_set)

ms_test_set <- ms_spec_mat[-train_idx, ]
rownames(ms_test_set)
```

```{r knn-classes}
labels <- c(
  rep("fiz", 9),
  rep("fiz_acid", 2),
  rep("fiz_ass", 2),
  rep("fiz_caff", 2),
  rep("fiz_old", 2)
)
```

```{r knn-engine}
ms_stat2 <- StatisticEngine$new(analyses = ms_train_set)
ms_stat2$add_classes(labels)
ms_stat2$run(StatisticSettings_MakeModel_knn(k = 3))
ms_stat2$predict(ms_test_set)
ms_stat2$model$prediction$results
```

# Raman Only

Statistical analysis is performed on Raman data only.

```{r fusion}
raman_spec_mat <- raman$get_spectra_matrix()
raman_spec_mat <- mdatools::prep.autoscale(raman_spec_mat, center = TRUE, scale = FALSE)
str(raman_spec_mat)
```

## Unsupervised

```{r sets-show}
main_set_idx <- c(1:10)
raman_main_set <- raman_spec_mat[main_set_idx, ]
rownames(raman_main_set)
raman_test_set <- raman_spec_mat[-main_set_idx, ]
rownames(raman_test_set)
```

```{r pca-engine, fig.cap="MS PCA model (native plot from mdatools R package).", fig.height=9, warning=FALSE}
raman_stat <- StatisticEngine$new(analyses = raman_main_set)
raman_stat$run(StatisticSettings_MakeModel_pca_mdatools(ncomp = 3, alpha = 0.01, gamma = 0.001))
plot(raman_stat$model, showText = TRUE)
```

```{r pca-predict, warning=FALSE, fig.cap="MS PCA model prediction distances. For more information about the distances plese consult the mdatools guide in https://mdatools.com/docs/pca--distances-and-limits.html."}
raman_stat$predict(raman_test_set)
plot(raman_stat$model, showText = TRUE)
```

## Supervised (classification)

K-Nearest Neighbors (KNN) can be used to classify analyses based on pre-classified training analyses.

```{r sets-show-knn}
train_idx <- grep("_1|_2|_01|_04|_06|_10", rownames(raman_spec_mat))

raman_train_set <- raman_spec_mat[train_idx, ]
raman_train_set <- raman_train_set[order(rownames(train_set)), ]
rownames(raman_train_set)

raman_test_set <- raman_spec_mat[-train_idx, ]
rownames(raman_test_set)
```

```{r knn-classes}
labels <- c(
  rep("fiz", 9),
  rep("fiz_acid", 2),
  rep("fiz_ass", 2),
  rep("fiz_caff", 2),
  rep("fiz_old", 2)
)
```

```{r knn-engine}
raman_stat2 <- StatisticEngine$new(analyses = raman_train_set)
raman_stat2$add_classes(labels)
raman_stat2$run(StatisticSettings_MakeModel_knn(k = 3))
raman_stat2$predict(raman_test_set)
raman_stat2$model$prediction$results
```

### 2024-12-10 dev/demo_paracetamol_fusion.Rmd
3d plot raman spectra

@@
 
-main_dir <- "C:/Users/apoli/Documents/example_files/isoft/Paracetamol_subset" # Ricardo laptop
+main_dir <- "C:/Users/thiss/Documents/Paracetamol_subset" # Ricardo laptop
 ms_dir <- paste0(main_dir, "/ms")
@@
 ```{r sets-show}
-main_set_idx <- c(1:10)
+main_set_idx <- c(c(1:9), 13, 15)
 
@@
 ```{r sets-show-knn}
-train_idx <- grep("_1|_2|_01|_04|_06|_10", rownames(fused_mat))
+train_idx <- grep("_1|_2|_01|_06|_07|_10", rownames(fused_mat))
 
@@
 ```{r sets-show}
-main_set_idx <- c(1:10)
+main_set_idx <- c(c(1:9), 13, 15)
 ms_main_set <- ms_spec_mat[main_set_idx, ]
@@
 ```{r sets-show-knn}
-train_idx <- grep("_1|_2|_01|_04|_06|_10", rownames(ms_spec_mat))
+train_idx <- grep("_1|_2|_01|_06|_07|_10", rownames(ms_spec_mat))
 
@@
 ```{r sets-show}
-main_set_idx <- c(1:10)
+main_set_idx <- c(c(1:9), 13, 15)
 raman_main_set <- raman_spec_mat[main_set_idx, ]
@@
 ```{r sets-show-knn}
-train_idx <- grep("_1|_2|_01|_04|_06|_10", rownames(raman_spec_mat))
+train_idx <- grep("_1|_2|_01|_06|_07|_10", rownames(raman_spec_mat))
 
@@
 ```
+
+```{r}
+
+```

### 2024-12-13 R/class_RamanSettings_DeleteSpectraSection.R
Major update of funtionality for NTS and caching

@@ -16,7 +16,8 @@
 #'
 #' @export
 #'
-RamanSettings_DeleteSpectraSection_StreamFind <- S7::new_class("RamanSettings_DeleteSpectraSection_StreamFind",
+RamanSettings_DeleteSpectraSection_StreamFind <- S7::new_class(
+  "RamanSettings_DeleteSpectraSection_StreamFind",
   parent = ProcessingSettings,
   package = "StreamFind",
   
@@ -24,7 +25,6 @@ RamanSettings_DeleteSpectraSection_StreamFind <- S7::new_class("RamanSettings_De
                          shiftmax = NULL,
                          rtmin = NULL,
                          rtmax = NULL) {
-    
     S7::new_object(ProcessingSettings(
       engine = "Raman",
       method = "DeleteSpectraSection",

### 2024-12-13 R/class_Results_Quantification.R

@@ -1,17 +1,14 @@
 #' @export
 #' @noRd
-Quantification <- S7::new_class("Quantification", package = "StreamFind", parent = Results,
+Quantification <- S7::new_class(
+  name = "Quantification",
+  package = "StreamFind",
+  parent = Results,
   
   properties = list(
-    ## __compound -----
     compounds = S7::new_property(S7::class_character, default = character()),
-    
-    ## __model -----
     models = S7::new_property(S7::class_list, default = list()),
-    
-    ## __quantity -----
     quantities = S7::new_property(S7::class_list, default = list())
-    
   ),
   
   constructor = function(compounds = character(), models = list(), quantities = list()) {
@@ -27,14 +24,11 @@ Quantification <- S7::new_class("Quantification", package = "StreamFind", parent
   },
   
   validator = function(self) {
-    valid <- all(
-      checkmate::test_true(self@name == "Quantification"),
-      checkmate::test_true(self@software == "StreamFind"),
-      checkmate::test_character(self@compounds),
-      checkmate::test_list(self@models),
-      checkmate::test_list(self@quantities)
-    )
-    if (!valid) return(FALSE)
+    checkmate::assert_true(self@name == "Quantification")
+    checkmate::assert_true(self@software == "StreamFind")
+    checkmate::assert_character(self@compounds)
+    checkmate::assert_list(self@models)
+    checkmate::assert_list(self@quantities)
     NULL
   }
 )

### 2024-12-13 R/class_Results_Spectra_RamanSpectra.R

@@ -1,14 +1,12 @@
 #' @export
 #' @noRd
 RamanSpectra <- S7::new_class(
-  "RamanSpectra",
+  name = "RamanSpectra",
   package = "StreamFind",
   parent = Spectra,
   
   properties = list(
     
-    # MARK: has_chromatograms
-    ## __has_chromatograms -----
     has_chromatograms = S7::new_property(
       S7::class_logical,
       getter = function(self) {
@@ -18,13 +16,12 @@ RamanSpectra <- S7::new_class(
       }
     ),
     
-    # MARK: chrom_peaks
-    ## __chrom_peaks -----
     chrom_peaks = S7::new_property(S7::class_list, default = list()),
     
-    # MARK: has_chrom_peaks
-    ## __has_chrom_peaks -----
-    has_chrom_peaks = S7::new_property(S7::class_logical, getter = function(self) length(self@chrom_peaks) > 0)
+    has_chrom_peaks = S7::new_property(
+      S7::class_logical,
+      getter = function(self) length(self@chrom_peaks) > 0
+    )
   ),
   
   constructor = function(spectra = list(),

### 2024-12-13 R/engine_CoreEngine.R

@@ -1,32 +1,34 @@
-# MARK: CoreEngine
 #' **CoreEngine** R6 class and methods
 #'
-#' @description The `CoreEngine` R6 class is used internally for harmonizing the different data specific engines. Users
-#' should not use this class directly.
+#' @description The `CoreEngine` R6 class is used internally for harmonizing the different data
+#' specific engines. Users should not use this class directly.
 #'  
 #' @template arg-headers
 #' @template arg-workflow
 #' @template arg-file
 #'
 #' @export
-CoreEngine <- R6::R6Class("CoreEngine",
-
-  # MARK: private fields
-  # _ private fields -----
+#' 
+CoreEngine <- R6::R6Class(
+  classname = "CoreEngine",
+  
+  # MARK: private
+  # private -----
   private = list(
     .headers = NULL,
     .workflow = NULL,
     .analyses = NULL,
     .file = NULL,
-    .audit_trail = NULL
+    .audit_trail = NULL,
+    .config_caches_data = TRUE,
+    .config_cache_file = "cache.sqlite"
   ),
 
   # MARK: active bindings
-  # _ active bindings -----
+  # active bindings -----
   active = list(
 
     # MARK: headers
-    # __headers -----
     #' @field headers `ProjectHeaders` S7 class object.
     headers = function(value) {
       if (missing(value)) {
@@ -34,12 +36,16 @@ CoreEngine <- R6::R6Class("CoreEngine",
       }
       if (is(value, "StreamFind::ProjectHeaders")) {
         private$.headers <- value
-        if (!is.null(private$.audit_trail)) private$.audit_trail <- add(private$.audit_trail, value)
+        if (!is.null(private$.audit_trail)) {
+          private$.audit_trail <- add(private$.audit_trail, value)
+        }
       } else if (is(value, "list")) {
         tryCatch(
           {
             private$.headers <- StreamFind::ProjectHeaders(value)
-            if (!is.null(private$.audit_trail)) private$.audit_trail <- add(private$.audit_trail, private$.headers)
+            if (!is.null(private$.audit_trail)) {
+              private$.audit_trail <- add(private$.audit_trail, private$.headers)
+            }
           },
           error = function(e) {
             warning(e)
@@ -55,7 +61,6 @@ CoreEngine <- R6::R6Class("CoreEngine",
     },
 
     # MARK: workflow
-    # __workflow -----
     #' @field workflow `Workflow` S7 class object.
[…]

### 2025-02-07 dev/demo_ms_raman_fusion_bsa.Rmd
moved temporarily old vignettes and plot fixes

---
title: 'Purity Evaluation of Bovine Serum Albumin Products'
subtitle: |
  | Fusion of Mass Spectrometry and Raman Spectroscopy data
  | <br>
author:
  name: Ricardo Cunha
  email: cunha@iuta.de
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    fig_caption: true
    toc: true
    number_sections: true
    toc_float: true
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Purity Evaluation of Bovine Serum Albumin Products}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(plotly)
library(data.table)
library(StreamFind)

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 9,
  results = "markup",
  comment = ""
)

main_dir <- "C:/Users/apoli/Documents/example_files/isoft/BSA_products"
all_files <- list.files(main_dir, full.names = TRUE)
ms_files <- all_files[grepl(".d$", all_files)]
raman_files <- list.files(main_dir, pattern = ".asc$", full.names = TRUE)
```

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<br>

<br>

***

# Introduction

Mass Spectrometry (MS) and Raman spectroscopy (Raman) are orthogonal techniques. While MS measures the chemical composition of a given sample, Raman is able to provide structural information of the constituents. This orthogonality is interesting when applying quality evaluation workflows, where both chemical composition and structural integrity are relevant. In this article, we show how MS and Raman can be pre-processed and fused for a combined statistical evaluation to support quality control. As test sets, we use four different bovine serum albumin (BSA) products where we evaluate the quality of the active product ingredient (API).

# Pre-processing

The preliminary processing is applied differently for MS and Raman data as the data structure does not allow direct fusion of the data. Therefore, in the following two sub-chapters we describe the pre-processing workflow steps for MS and Raman data.

## MS

The four different BSA products were measured in triplicate in full scan MS mode. Giving the following set of *.d* files.

```{r files}
# Raw data from an Agilent QTOF
basename(ms_files)
```

With the *.d* files, a *MassSpecEngine* is created to pre-process the MS data. Note that the MS files were centroided and converted to mzML during creation of the *MassSpecEngine*, using the installed `msconvert` tool from the [ProteoWizard](https://proteowizard.sourceforge.io/) in the background. The analysis replicate names are modified to properly assign the analyses within each replicate measurement. Two water (H2O) files were used as blanks for the analyses.

```{r engine-creation}
ms <- StreamFind::MassSpecEngine$new(
  headers = list(name = "BSA Quality Evaluation Project"),
  analyses = ms_files,
  centroid = TRUE,
  levels = 1
)

# Changes the analysis replicate names
ms$add_replicate_names(sub("_\\d+$", "", ms$get_analysis_names()))

# Overview
ms
```

```{r plot-spectra, fig.cap="Total ion chromatograms (TIC) for each analysis replicate."}
# Note that the TIC chromatograms are not yet loaded (see below)
# In plot_spectra_tic, the TIC from spectra headers are used
ms$plot_spectra_tic(colorBy = "replicates")
```

As pre-processing, the following steps are applied:

1. Find the retention time elution of BSA based on the total ion chromatogram (TIC) of each analysis; 
2. Find the mass-to-charge (*m/z*) range of BSA for deconvolution of charges; 
3. Deconvolute charges for estimation of the exact mass, in DA;

The individual processing steps are applied according to *ProcessingSettings*. Below, the integration of the TIC is shown using the dedicated *ProcessingSettings* after loading the TIC chromatograms from the analysis files and smoothing using a moving average approach.

```{r load-chroms}
# Loads the TIC chromatograms
ms$run(MassSpecSettings_LoadChromatograms_StreamFind(chromatograms = "TIC"))
ms$analyses$has_chromatograms
```

```{r smooth-chroms}
# Smoothing of the TIC chromatogram for improving peak finding
ms$run(MassSpecSettings_SmoothChromatograms_movingaverage(windowSize = 5))
```

```{r plot-chroms, fig.cap="Smoothed TICs for each analysis replicate."}
ms$plot_chromatograms(colorBy = "replicates+targets")
```

```{r localmaxima-chroms}

# Settings for finding the local maxima in the loaded TIC chromatograms
ps_maxima <- MassSpecSettings_FindChromPeaks_LocalMaxima(
  minWidth = 3,
  maxWidth = 5,
  minHeight = 30E6
)

show(ps_maxima)
```

```{r run-localmaxima-chroms}
ms$run(ps_maxima)
```

```{r plot-chrom-peaks,fig.cap="TIC with integrated peaks."}
ms$plot_chromatograms_peaks(colorBy = "replicates+targets")
```

The average retention time of the peaks was `r mean(data.table::rbindlist(ms$chromatograms$peaks)$rt)` seconds with a standard deviation of `r sd(data.table::rbindlist(ms$chromatograms$peaks)$rt)`. The following step is to estimate the *m/z* range of BSA for deconvolution of charges. The *m/z* range needs to be defined once and is equal to all other analyses if the compound of interest is BSA. As for the retention time, shifts might occur due to the chromatographic separation.

```{r rt-av}
# Average retention time of the main peak
rt_av <- mean(data.table::rbindlist(ms$chromatograms$peaks)$rt)
rt_av
```

```{r plot-spectra-raw, fig.cap="Spectra of BSA for the first analysis at the peak maximum."}
ms$plot_spectra_ms1(
  analyses = 1,
  rt = rt_av,
  sec = 0.5,
  colorBy = "replicates",
  presence = 0.01,
  minIntensity = 100
)
```

According to the MS1 spectra, the *m/z* from 1200 to 2000 can be added to the dedicated *ProcessingSettings* for deconvolution, as shown below. The retention time dimensions for extraction of spectra is based on the local peak maxima shown above.

```{r workflow-ms}

# Extracts the spectra for the defined *m/z* range and local peak maxima
ms$run(
  MassSpecSettings_LoadSpectra_chrompeaks(
    mzmin = 1200,
    mzmax = 2000,
    levels = 1,
    minIntensity = 10
  )
)

# Calculates the charges for the extracted spectra
ms$run(
  MassSpecSettings_CalculateSpectraCharges_StreamFind(
    onlyTopScans = TRUE,
    topScans = 6,
    roundVal = 10,
    relLowCut = 0.2,
    absLowCut = 0,
    top = 10
  )
)

# Deconvolutes the charges for the extracted spectra
ms$run(MassSpecSettings_DeconvoluteSpectra_StreamFind())

ms$run(
  MassSpecSettings_SmoothSpectra_movingaverage(
    windowSize = 15
  )
)

# The binning is applied to ensure that all spectra have the same dimensions
# (i.e., mass-time and intensity pair)
ms$run(
   MassSpecSettings_BinSpectra_StreamFind(
    binNames = c("rt", "mass"),
    binValues = c(5, 20),
    byUnit = TRUE,
    refBinAnalysis = 1
  )
)

ms$run(
  MassSpecSettings_AverageSpectra_StreamFind(
    by = "replicates",
    weightedAveraged = FALSE
  )
)

ms$run(MassSpecSettings_NormalizeSpectra_minmax())

ms$run(MassSpecSettings_NormalizeSpectra_blockweight())
```

The `print_workflow()` method can be used to inspect the workflow settings.

```{r print-ms-workflow}
ms$print_workflow()
```

```{r plot-charges, fig.cap="Charges for the extracted spectra."}
plot_spectra_charges(ms$analyses, colorBy = "analyses")
```

```{r plot_spectra,fig.cap="Deconvoluted and pre-processed spectra for each analysis."}
plot_spectra(ms$analyses, colorBy = "analyses")
```

## Raman

The four different BSA products were measured using an online coupling of size exclusion chromatography (SEC) with capillary-enhanced Raman spectroscopy (CERS) through a liquid-core optical fibre flow cell, as reported by Thissen et al. (DOI: 10.1021/acs.analchem.3c03991). The raw spectra was converted to *asc* format for each retention time. A subset of the background and the main BSA elution peak was selected for pre-processing.

```{r raman-engine}
raman <- RamanEngine$new(analyses = raman_files)

# Modifying the replicate names to match the MS analyses
raman$add_replicate_names(unique(ms$get_analysis_names()))

raman
```

```{r correct-raman-rt, include=FALSE}
corr_spectra <- raman$spectra$spectra
corr_spectra_4 <- corr_spectra[[4]]
corr_spectra_4$rt <- corr_spectra_4$rt - 68.545
corr_spectra_4 <- corr_spectra_4[corr_spectra_4$rt > 0, ]
corr_spectra[[4]] <- corr_spectra_4
corr_spectra <- lapply(corr_spectra, function(x) x[x$rt < 800, ])
raman$spectra <- StreamFind::RamanSpectra(spectra = corr_spectra)
```

```{r raman-plot-chroms, fig.cap="Time series chromatogram for each BSA product. Each data point is a Raman spectrum at a specific retention time."}
raman$plot_chromatograms(
  yLab = "Abbundance sum for each spectrum",
  colorBy = "analyses"
)
```

```{r raman-plot_spectra-raw, fig.cap="Raw averaged Raman spectra for each BSA product between 430 and 440 seconds."}
raman$plot_spectra(rt = c(430, 440), colorBy = "analyses")
```

As for MS data, the workflow for Raman data is also assembled using an ordered list of processing settings, as shown below.

```{r raman-workflow}
raman_workflow <- list(
  RamanSettings_BinScans_StreamFind(
    binNames = "rt",
    binValues = 20,
    byUnit = FALSE
  ),

  RamanSettings_SubtractSpectraSection_native(
    sectionVal = "rt",
    sectionWindow = c(0, 200)
  ),

  RamanSettings_DeleteSpectraSection_native(
    min = -100,
    shiftmax = 315
  ),

  RamanSettings_SmoothSpectra_savgol(
    fl = 11,
    forder = 2,
    dorder = 0
  ),

  RamanSettings_DeleteSpectraSection_StreamFind(
    shiftmin = 315,
    shiftmax = 330
  ),

  RamanSettings_DeleteSpectraSection_StreamFind(
    shiftmin = 2300,
    shiftmax = 2600
  ),

  RamanSettings_CorrectSpectraBaseline_baseline_als(
    lambda = 3,
    p = 0.015,
    maxit = 10
  ),

  RamanSettings_DeleteSpectraSection_StreamFind(
    rtmin = 0,
    rtmax = 420
  ),

  RamanSettings_DeleteSpectraSection_StreamFind(
    rtmin = 430,
    rtmax = 900
  ),

  RamanSettings_AverageSpectra_StreamFind(
    by = "replicates",
    weightedAveraged = FALSE
  ),

  RamanSettings_NormalizeSpectra_minmax(),

  RamanSettings_NormalizeSpectra_blockweight()
)

raman$workflow <- raman_workflow

raman$print_workflow()
```

A particularity of the workflow for Raman data is that the background subtraction (workflow step number 3) is performed from a time region (i.e., between 0 and 3 minutes) before the main BSA peak, which is from approximately from 5 to 8 minutes. The workflow is then applied using the `run_workflow()` method.

```{r raman-workflow-run}
raman$run_workflow()
```

```{r raman-plot_spectra-baseline, fig.cap="Baseline correction for each spectra."}
raman$plot_spectra_baseline()
```

```{r raman-plot_spectra, fig.cap="Processed Raman spectra for each BSA product."}
raman$plot_spectra()
```

# Fusion

The fusion of MS and Raman data is performed by combining the processed data from both engines. The data is then merged into a single data frame for statistical evaluation using a StatisticEngine, as shown in the next chapter.

```{r fusion}
ms_mat <- get_spectra_matrix(ms$analyses)
raman_mat <- get_spectra_matrix(raman$analyses)
fused_mat <- cbind(ms_mat, raman_mat)
attr(fused_mat, "xaxis.name") = "Keys"
attr(fused_mat, "xaxis.values") = seq_len(ncol(fused_mat))
str(fused_mat)
```

# Statistic analysis

A statistic analysis is performed with the *StatisticEngine* for evaluation of the differences between the BSA products.

## PCA

The fused data is used to create a PCA model with the *StatisticEngine* using processing settings based on the [mdatools](https://www.mdatools.com/docs/pca.html) package.

```{r pca-engine}
stat <- StatisticEngine$new(analyses = fused_mat)
```

```{r pca-data-raw}
# Plots the data from the added fused data.frame/matrix
plot_data(stat$analyses)
```

```{r pca-run}
# Note that scale is set to FALSE to avoid scaling the data
# as blockweight normalization was applied for each dataset
stat$run(
  StatisticSettings_MakeModel_pca_mdatools(
    center = TRUE,
    ncomp = 2,
    alpha = 0.05,
    gamma = 0.02
  )
)
summary(stat$model) # summary method is from mdatools
```

```{r pca-results, fig.cap="Overview plot for the PCA model."}
plot(stat$model)
```

```{r pca-loadings, fig.cap="PCA loadings plot for the fused data."}
# Adds a color key to the loadings plot to evaluate the contribution of each dataset (i.e., MS and Raman)
plot_loadings(stat$analyses, colorKey = rep(c("MS", "Raman"), c(ncol(ms_mat), ncol(raman_mat))))
```

```{r pca-model-mdatools, fig.cap="PCA summary plot using the original mdatools R package.", fig.height=9}
# The mdatools S3 class model object is obtained via the active binding of the StatisticEngine
plot(stat$model$model, show.labels = TRUE) # plot method from mdatools
```

## MCR purity

An alternative approach for the statistic analysis is the MCR pure model (based on [mdatools](https://www.mdatools.com/docs/mcr--purity.html) R package) to evaluate the purity level in the analyses.

```{r mcr-run}
# Note that the model in the engine will be overwritten by the MCR model
stat$run(StatisticSettings_MakeModel_mcrpure_mdatools(ncomp = 2, offset = 0.05))
summary(stat$model)
```

```{r mcr-model, fig.cap="Overview plot of the MCR pure model."}
plot(stat$model)
```

```{r mcr-data}
# Gets the most pure features for each component
get_model_data(stat$model)$purity
```

-----

### 2025-02-07 dev/demo_paracetamol_fusion.Rmd
moved temporarily old vignettes and plot fixes

--- 2024-12-02 dev/demo_paracetamol_fusion.Rmd
@@
 raman_workflow <- list(
-  RamanSettings_AverageSpectra_StreamFind(),
+  RamanSettings_AverageSpectra_native(),
   RamanSettings_SubtractBlankSpectra_StreamFind(),
   RamanSettings_SmoothSpectra_savgol(fl = 21, forder = 4, dorder = 0),
-  RamanSettings_DeleteSpectraSection_StreamFind(shiftmin = -100, shiftmax = 350),
-  RamanSettings_DeleteSpectraSection_StreamFind(shiftmin = 1800, shiftmax = 2500),
+  RamanSettings_DeleteSpectraSection_native(min = -100, max = 350),
+  RamanSettings_DeleteSpectraSection_native(min = 1800, max = 2500),
   RamanSettings_CorrectSpectraBaseline_airpls(lambda = 45, differences = 1, itermax = 30),
@@
 ```{r sets-show-knn}
-train_idx <- grep("_1|_2|_01|_04|_06|_10", rownames(fused_mat))
+train_idx <- unique(c(grep("1", rownames(fused_mat)), grep("2", rownames(fused_mat))))
 
@@
 labels <- c(
-  rep("fiz", 9),
+  rep("fiz", 8),
   rep("fiz_acid", 2),
@@
 ```
-
-# MS Only
-
-Statistical analysis is performed on MS data only.
-
-```{r fusion}
-ms_spec_mat <- ms$get_spectra_matrix()
-ms_spec_mat <- mdatools::prep.autoscale(ms_spec_mat, center = TRUE, scale = FALSE)
-str(ms_spec_mat)
-```
-
-## Unsupervised
-
-```{r sets-show}
-main_set_idx <- c(1:10)
-ms_main_set <- ms_spec_mat[main_set_idx, ]
-rownames(ms_main_set)
-ms_test_set <- ms_spec_mat[-main_set_idx, ]
-rownames(ms_test_set)
-```
-
-```{r pca-engine, fig.cap="MS PCA model (native plot from mdatools R package).", fig.height=9, warning=FALSE}
-ms_stat <- StatisticEngine$new(analyses = ms_main_set)
-ms_stat$run(StatisticSettings_MakeModel_pca_mdatools(ncomp = 3, alpha = 0.01, gamma = 0.001))
-plot(ms_stat$model, showText = TRUE)
-```
-
-```{r pca-predict, warning=FALSE, fig.cap="MS PCA model prediction distances.
-For more information about the distances plese consult the mdatools guide in https://mdatools.com/docs/pca--distances-and-limits.html."}
-ms_stat$predict(ms_test_set)
-plot(ms_stat$model, showText = TRUE)
-```
-
-## Supervised (classification)
-
-K-Nearest Neighbors (KNN) can be used to classify analyses based on pre-classified training analyses.
-
-```{r sets-show-knn}
-train_idx <- grep("_1|_2|_01|_04|_06|_10", rownames(ms_spec_mat))
-
-ms_train_set <- ms_spec_mat[train_idx, ]
-ms_train_set <- ms_train_set[order(rownames(train_set)), ]
-rownames(ms_train_set)
-
-ms_test_set <- ms_spec_mat[-train_idx, ]
-rownames(ms_test_set)
-```
-
-```{r knn-classes}
-labels <- c(
-  rep("fiz", 9),
-  rep("fiz_acid", 2),
-  rep("fiz_ass", 2),
-  rep("fiz_caff", 2),
-  rep("fiz_old", 2)
-)
-```
-
-```{r knn-engine}
-ms_stat2 <- StatisticEngine$new(analyses = ms_train_set)
-ms_stat2$add_classes(labels)
-ms_stat2$run(StatisticSettings_MakeModel_knn(k = 3))
-ms_stat2$predict(ms_test_set)
-ms_stat2$model$prediction$results
-```
-
-# Raman Only
-
-Statistical analysis is performed on Raman data only.
-
-```{r fusion}
-raman_spec_mat <- raman$get_spectra_matrix()
-raman_spec_mat <- mdatools::prep.autoscale(raman_spec_mat, center = TRUE, scale = FALSE)
-str(raman_spec_mat)
-```
-
-## Unsupervised
-
-```{r sets-show}
-main_set_idx <- c(1:10)
-raman_main_set <- raman_spec_mat[main_set_idx, ]
-rownames(raman_main_set)
-raman_test_set <- raman_spec_mat[-main_set_idx, ]
-rownames(raman_test_set)
-```
-
-```{r pca-engine, fig.cap="MS PCA model (native plot from mdatools R package).", fig.height=9, warning=FALSE}
-raman_stat <- StatisticEngine$new(analyses = raman_main_set)
-raman_stat$run(StatisticSettings_MakeModel_pca_mdatools(ncomp = 3, alpha = 0.01, gamma = 0.001))
-plot(raman_stat$model, showText = TRUE)
-```
-
-```{r pca-predict, warning=FALSE, fig.cap="MS PCA model prediction distances.
-For more information about the distances plese consult the mdatools guide in https://mdatools.com/docs/pca--distances-and-limits.html."}
-raman_stat$predict(raman_test_set)
-plot(raman_stat$model, showText = TRUE)
-```
-
-## Supervised (classification)
-
-K-Nearest Neighbors (KNN) can be used to classify analyses based on pre-classified training analyses.
-
-```{r sets-show-knn}
-train_idx <- grep("_1|_2|_01|_04|_06|_10", rownames(raman_spec_mat))
-
-raman_train_set <- raman_spec_mat[train_idx, ]
-raman_train_set <- raman_train_set[order(rownames(train_set)), ]
-rownames(raman_train_set)
-
-raman_test_set <- raman_spec_mat[-train_idx, ]
-rownames(raman_test_set)
-```
-
-```{r knn-classes}
-labels <- c(
-  rep("fiz", 9),
-  rep("fiz_acid", 2),
-  rep("fiz_ass", 2),
-  rep("fiz_caff", 2),
-  rep("fiz_old", 2)
-)
-```
-
-```{r knn-engine}
-raman_stat2 <- StatisticEngine$new(analyses = raman_train_set)
-raman_stat2$add_classes(labels)
-raman_stat2$run(StatisticSettings_MakeModel_knn(k = 3))
-raman_stat2$predict(raman_test_set)
-raman_stat2$model$prediction$results
-```

### 2025-02-07 R/utils_plots.R

@@ -127,44 +127,44 @@
 
 # MARK: .plot_lines_static
 #' @noRd
-.plot_lines_static <- function(spectra,
+.plot_lines_static <- function(data,
                                xLab,
                                yLab,
                                title,
                                cex,
                                showLegend) {
-  cl <- .get_colors(unique(spectra$var))
+  cl <- .get_colors(unique(data$var))
+  data$loop <- paste0(data$analysis, data$replicate, spectra$id, spectra$var)
+  loop_key <- unique(data$loop)
   
-  spectra$loop <- paste0(spectra$analysis, spectra$id, spectra$var)
+  xr <- c(min(data$x), max(data$x))
   
-  loop_key <- unique(spectra$loop)
-  
-  xr <- c(min(spectra$x), max(spectra$x))
   if (showLegend) {
     xr[2] <- xr[2] * 1.01
   }
   
-  intr <- c(min(spectra$intensity), max(spectra$intensity))
+  intr <- c(min(data$intensity), max(data$intensity))
   
   if (is.null(cex) || !is.numeric(cex)) cex <- 1
   
-  plot(spectra$x,
-       type = "n",
-       xlab = xLab,
-       ylab = yLab,
-       xlim = xr,
-       ylim = intr,
-       main = title
+  plot(
+    data$x,
+    type = "n",
+    xlab = xLab,
+    ylab = yLab,
+    xlim = xr,
+    ylim = intr,
+    main = title
   )
   
   for (t in loop_key) {
-    select_vector <- spectra$loop %in% t
+    select_vector <- data$loop %in% t
     
-    lt <- unique(spectra$var[select_vector])
+    lt <- unique(data$var[select_vector])
     
     lines(
-      x = spectra$x[select_vector],
-      y = spectra$intensity[select_vector],
+      x = data$x[select_vector],
+      y = data$intensity[select_vector],
       type = "l",
       pch = 19,
       cex = 0.5,
@@ -258,39 +258,35 @@
 
 # MARK: .plot_lines_baseline_static
 #' @noRd
-.plot_lines_baseline_static <- function(spectra, xLab, yLab, title, cex, showLegend) {
-  cl <- .get_colors(unique(spectra$var))
-  
-  spectra$loop <- paste0(data$analysis, data$replicate, spectra$id, spectra$var)
-  
-  loop_key <- unique(spectra$loop)
-  
-  xr <- c(min(spectra$x), max(spectra$x))
[…]

### 2025-02-13 dev/demo_ms_raman_fusion_bsa.Rmd
Major Update Class Names

@@
 
-The individual processing steps are applied according to *ProcessingSettings*.
-Below, the integration of the TIC is shown using the dedicated *ProcessingSettings* after loading the TIC chromatograms from the analysis files and smoothing using a moving average approach.
+The individual processing steps are applied according to *ProcessingStep*.
+Below, the integration of the TIC is shown using the dedicated *ProcessingStep* after loading the TIC chromatograms from the analysis files and smoothing using a moving average approach.
 
@@
 # Loads the TIC chromatograms
-ms$run(MassSpecSettings_LoadChromatograms_StreamFind(chromatograms = "TIC"))
-ms$analyses$has_chromatograms
+ms$run(MassSpecMethod_LoadChromatograms_StreamFind(chromatograms = "TIC"))
+ms$analyses$has_results_chromatograms
 ```
@@
 # Smoothing of the TIC chromatogram for improving peak finding
-ms$run(MassSpecSettings_SmoothChromatograms_movingaverage(windowSize = 5))
+ms$run(MassSpecMethod_SmoothChromatograms_movingaverage(windowSize = 5))
 ```
@@
 # Settings for finding the local maxima in the loaded TIC chromatograms
-ps_maxima <- MassSpecSettings_FindChromPeaks_LocalMaxima(
+ps_maxima <- MassSpecMethod_FindChromPeaks_LocalMaxima(
   minWidth = 3,
@@
 
-According to the MS1 spectra, the *m/z* from 1200 to 2000 can be added to the dedicated *ProcessingSettings* for deconvolution, as shown below.
+According to the MS1 spectra, the *m/z* from 1200 to 2000 can be added to the dedicated *ProcessingStep* for deconvolution, as shown below.
 The retention time dimensions for extraction of spectra is based on the local peak maxima shown above.
@@
 ms$run(
-  MassSpecSettings_LoadSpectra_chrompeaks(
+  MassSpecMethod_LoadSpectra_chrompeaks(
     mzmin = 1200,
@@
 ms$run(
-  MassSpecSettings_CalculateSpectraCharges_StreamFind(
+  MassSpecMethod_CalculateSpectraCharges_StreamFind(
     onlyTopScans = TRUE,
@@
 # Deconvolutes the charges for the extracted spectra
-ms$run(MassSpecSettings_DeconvoluteSpectra_StreamFind())
+ms$run(MassSpecMethod_DeconvoluteSpectra_StreamFind())
 
 ms$run(
-  MassSpecSettings_SmoothSpectra_movingaverage(
+  MassSpecMethod_SmoothSpectra_movingaverage(
     windowSize = 15
@@
 ms$run(
-   MassSpecSettings_BinSpectra_StreamFind(
+   MassSpecMethod_BinSpectra_StreamFind(
     binNames = c("rt", "mass"),
@@
 ms$run(
-  MassSpecSettings_AverageSpectra_StreamFind(
+  MassSpecMethod_AverageSpectra_StreamFind(
     by = "replicates",
@@
 
-ms$run(MassSpecSettings_NormalizeSpectra_minmax())
+ms$run(MassSpecMethod_NormalizeSpectra_minmax())
 
-ms$run(MassSpecSettings_NormalizeSpectra_blockweight())
+ms$run(MassSpecMethod_NormalizeSpectra_blockweight())
 ```
@@
 raman_workflow <- list(
-  RamanSettings_BinScans_StreamFind(
+  RamanMethod_BinScans_StreamFind(
     binNames = "rt",
@@
 
-  RamanSettings_SubtractSpectraSection_native(
+  RamanMethod_SubtractSpectraSection_native(
     sectionVal = "rt",
@@
 
-  RamanSettings_DeleteSpectraSection_native(
+  RamanMethod_DeleteSpectraSection_native(
     min = -100,
@@
 
-  RamanSettings_SmoothSpectra_savgol(
+  RamanMethod_SmoothSpectra_savgol(
     fl = 11,
@@
 
-  RamanSettings_DeleteSpectraSection_StreamFind(
+  RamanMethod_DeleteSpectraSection_StreamFind(
     shiftmin = 315,
@@
 
-  RamanSettings_DeleteSpectraSection_StreamFind(
+  RamanMethod_DeleteSpectraSection_StreamFind(
     shiftmin = 2300,
@@
 
-  RamanSettings_CorrectSpectraBaseline_baseline_als(
+  RamanMethod_CorrectSpectraBaseline_baseline_als(
     lambda = 3,
@@
 
-  RamanSettings_DeleteSpectraSection_StreamFind(
+  RamanMethod_DeleteSpectraSection_StreamFind(
     rtmin = 0,
@@
 
-  RamanSettings_DeleteSpectraSection_StreamFind(
+  RamanMethod_DeleteSpectraSection_StreamFind(
     rtmin = 430,
@@
 
-  RamanSettings_AverageSpectra_StreamFind(
+  RamanMethod_AverageSpectra_StreamFind(
     by = "replicates",
@@
 
-  RamanSettings_NormalizeSpectra_minmax(),
+  RamanMethod_NormalizeSpectra_minmax(),
 
-  RamanSettings_NormalizeSpectra_blockweight()
+  RamanMethod_NormalizeSpectra_blockweight()
 )
@@
 stat$run(
-  StatisticSettings_MakeModel_pca_mdatools(
+  StatisticMethod_MakeModel_pca_mdatools(
     center = TRUE,
@@
 # Note that the model in the engine will be overwritten by the MCR model
-stat$run(StatisticSettings_MakeModel_mcrpure_mdatools(ncomp = 2, offset = 0.05))
+stat$run(StatisticMethod_MakeModel_mcrpure_mdatools(ncomp = 2, offset = 0.05))
 summary(stat$model)

### 2025-02-13 dev/demo_paracetamol_fusion.Rmd
Major Update Class Names

@@
 ms_workflow <- list(
-  MassSpecSettings_LoadSpectra_StreamFind(levels = 1, mzmax = 300, rtmax = 120),
-  MassSpecSettings_BinSpectra_StreamFind(binNames = c("rt", "mz"), binValues = c(5, 2), refBinAnalysis = 1),
-  MassSpecSettings_NormalizeSpectra_minmax(),
-  MassSpecSettings_AverageSpectra_StreamFind(),
-  MassSpecSettings_SubtractBlankSpectra_StreamFind(negativeToZero = TRUE),
-  MassSpecSettings_NormalizeSpectra_blockweight()
+  MassSpecMethod_LoadSpectra_StreamFind(levels = 1, mzmax = 300, rtmax = 120),
+  MassSpecMethod_BinSpectra_StreamFind(binNames = c("rt", "mz"), binValues = c(5, 2), refBinAnalysis = 1),
+  MassSpecMethod_NormalizeSpectra_minmax(),
+  MassSpecMethod_AverageSpectra_StreamFind(),
+  MassSpecMethod_SubtractBlankSpectra_StreamFind(negativeToZero = TRUE),
+  MassSpecMethod_NormalizeSpectra_blockweight()
 )
@@
 raman_workflow <- list(
-  RamanSettings_AverageSpectra_native(),
-  RamanSettings_SubtractBlankSpectra_StreamFind(),
-  RamanSettings_SmoothSpectra_savgol(fl = 21, forder = 4, dorder = 0),
-  RamanSettings_DeleteSpectraSection_native(min = -100, max = 350),
-  RamanSettings_DeleteSpectraSection_native(min = 1800, max = 2500),
-  RamanSettings_CorrectSpectraBaseline_airpls(lambda = 45, differences = 1, itermax = 30),
-  RamanSettings_NormalizeSpectra_minmax(),
-  RamanSettings_NormalizeSpectra_blockweight()
+  RamanMethod_AverageSpectra_native(),
+  RamanMethod_SubtractBlankSpectra_StreamFind(),
+  RamanMethod_SmoothSpectra_savgol(fl = 21, forder = 4, dorder = 0),
+  RamanMethod_DeleteSpectraSection_native(min = -100, max = 350),
+  RamanMethod_DeleteSpectraSection_native(min = 1800, max = 2500),
+  RamanMethod_CorrectSpectraBaseline_airpls(lambda = 45, differences = 1, itermax = 30),
+  RamanMethod_NormalizeSpectra_minmax(),
+  RamanMethod_NormalizeSpectra_blockweight()
 )
@@
 # The alpha and gamma parameters are used to evaluate the model limits
-stat$run(StatisticSettings_MakeModel_pca_mdatools(ncomp = 3, alpha = 0.01, gamma = 0.001))
+stat$run(StatisticMethod_MakeModel_pca_mdatools(ncomp = 3, alpha = 0.01, gamma = 0.001))
 plot(stat$model, showText = TRUE)
@@
 stat2$add_classes(labels)
-stat2$run(StatisticSettings_MakeModel_knn(k = 3))
+stat2$run(StatisticMethod_MakeModel_knn(k = 3))
 stat2$predict(test_set)

### 2025-02-13 R/app_mod_WorkflowAssembler_Explorer_MassSpec.R

@@ -36,16 +36,16 @@ S7::method(.mod_WorkflowAssembler_Explorer_Server, MassSpecAnalyses) <- function
                                                                                  reactive_config) {
   shiny::moduleServer(id, function(input, output, session) {
     ns2 <- shiny::NS(id)
-    has_spectra <- shiny::reactiveVal(FALSE)
-    has_chromatograms <- shiny::reactiveVal(FALSE)
+    has_results_spectra <- shiny::reactiveVal(FALSE)
+    has_results_chromatograms <- shiny::reactiveVal(FALSE)
     levels <- shiny::reactiveVal(1)
     rt_end <- shiny::reactiveVal(0)
     rt_start <- shiny::reactiveVal(0)
 
     init_analyses <- reactive_analyses()
     if (length(init_analyses) > 0) {
-      has_spectra(max(init_analyses@spectra_number) > 0)
-      has_chromatograms(max(init_analyses@chromatograms_number) > 0)
+      has_results_spectra(max(init_analyses@spectra_number) > 0)
+      has_results_chromatograms(max(init_analyses@chromatograms_number) > 0)
       levels(as.numeric(unlist(strsplit(unique(init_analyses@spectra_level), ", "))))
       rt_end(round(max(init_analyses@spectra_highest_rt), digits = 0))
       rt_start(round(min(init_analyses@spectra_lowest_rt), digits = 0))
@@ -54,14 +54,14 @@ S7::method(.mod_WorkflowAssembler_Explorer_Server, MassSpecAnalyses) <- function
     shiny::observe({
       analyses <- reactive_analyses()
       if (length(analyses) > 0) {
-        has_spectra(max(reactive_analyses()@spectra_number) > 0)
-        has_chromatograms(max(reactive_analyses()@chromatograms_number) > 0)
+        has_results_spectra(max(reactive_analyses()@spectra_number) > 0)
+        has_results_chromatograms(max(reactive_analyses()@chromatograms_number) > 0)
         levels(as.numeric(unlist(strsplit(unique(reactive_analyses()@spectra_level), ", "))))
         rt_end(round(max(reactive_analyses()@spectra_highest_rt), digits = 0))
         rt_start(round(min(reactive_analyses()@spectra_lowest_rt), digits = 0))
       } else {
-        has_spectra(FALSE)
-        has_chromatograms(FALSE)
+        has_results_spectra(FALSE)
+        has_results_chromatograms(FALSE)
         levels(1)
         rt_end(0)
         rt_start(0)
[…]

### 2025-02-13 R/class_MassSpecMethod_FindChromPeaks.R

@@ -1,0 +1,332 @@
+#' **MassSpecMethod_FindChromPeaks_LocalMaxima**
+#'
+#' @description Finds peak maxima in chromatograms.
+#' 
+#' @param minWidth Numeric (length 1) with the minimum width of a peak.
+#' @param maxWidth Numeric (length 1) with the maximum width of a peak.
+#' @param minHeight Numeric (length 1) with the minimum height of a peak.
+#'
+#' @return A MassSpecMethod_FindChromPeaks_LocalMaxima object.
+#'
+#' @export
+#'
+MassSpecMethod_FindChromPeaks_LocalMaxima <- S7::new_class(
+  name = "MassSpecMethod_FindChromPeaks_LocalMaxima",
+  parent = ProcessingStep,
+  package = "StreamFind",
+  
+  constructor = function(minWidth = 0, maxWidth = 0, minHeight = 0) {
+    
+    S7::new_object(
+      ProcessingStep(
+        engine = "MassSpec",
+        method = "FindChromPeaks",
+        required = "LoadChromatograms",
+        algorithm = "LocalMaxima",
+        parameters = list(
+          minWidth = as.numeric(minWidth),
+          maxWidth = as.numeric(maxWidth),
+          minHeight = as.numeric(minHeight)
+        ),
+        number_permitted = 1,
+        version = as.character(packageVersion("StreamFind")),
+        software = "StreamFind",
+        developer = "Ricardo Cunha",
+        contact = "cunha@iuta.de",
+        link = "https://odea-project.github.io/StreamFind",
+        doi = NA_character_
+      )
+    )
+  },
+  
+  validator = function(self) {
+    checkmate::assert_choice(self@engine, "MassSpec")
+    checkmate::assert_choice(self@method, "FindChromPeaks")
+    checkmate::assert_choice(self@algorithm, "LocalMaxima")
+    checkmate::assert_number(self@parameters$minWidth)
+    checkmate::assert_number(self@parameters$maxWidth)
+    checkmate::assert_number(self@parameters$minHeight)
+    NULL
+  }
+)
+
+#' @export
+#' @noRd
+S7::method(run, MassSpecMethod_FindChromPeaks_LocalMaxima) <- function(x, engine = NULL) {
+  
+  if (!is(engine, "MassSpecEngine")) {
+    warning("Engine is not a MassSpecEngine object!")
+    return(FALSE)
+  }
+  
+  if (!engine$has_analyses()) {
[…]

### 2025-02-13 R/class_MassSpecMethod_GenerateFormulas.R

@@ -1,0 +1,273 @@
+#' **MassSpecMethod_GenerateFormulas_genform**
+#'
+#' @description Settings for generating formulas using the algorithm \href{https://sourceforge.net/projects/genform/}{GenForm}.
+#' The algorithm is used via the function \link[patRoon]{generateFormulas} from the package \pkg{patRoon}. Therefore,
+#' it is highly recommended to check the original documentation of the function in \pkg{patRoon} for more details.
+#' 
+#' @param MSPeakListsClusterMzWindow m/z window (in Da) used for clustering m/z values when spectra are averaged.
+#' For method="hclust" this corresponds to the cluster height, while for method="distance" this value is used to find
+#' nearby masses (+/- window). Too small windows will prevent clustering m/z values (thus erroneously treating equal 
+#' masses along spectra as different), whereas too big windows may cluster unrelated m/z values from different or even
+#' the same spectrum together.
+#' @param MSPeakListsTopMost Only retain this maximum number of MS peaks when generating averaged spectra. Lowering
+#' this number may exclude more irrelevant (noisy) MS peaks and decrease processing time, whereas higher values may
+#' avoid excluding lower intense MS peaks that may still be of interest.
+#' @param MSPeakListsMinIntensityPre MS peaks with intensities below this value will be removed (applied prior to
+#' selection by `topMost`) before averaging.
+#' @param MSPeakListsMinIntensityPost MS peaks with intensities below this value will be removed after averaging.
+#' @param MSPeakListsAvgFun Character with the function name that is used to calculate average m/z values.
+#' @param MSPeakListsMethod Method used for producing averaged MS spectra. Valid values are "hclust", used for
+#' hierarchical clustering (using the fastcluster package), and "distance", to use the between peak distance.
+#' The latter method may reduces processing time and memory requirements, at the potential cost of reduced accuracy.
+#' @param relMzDev Numeric (length 1) with the relative mass deviation, in ppm.
+#' @param elements Character vector with the elements to use for formulae annotation. Always try to work with a minimal
+#' set by excluding elements you don't expect.
[…]

### 2025-02-13 R/class_MassSpecMethod_SmoothSpectra.R

@@ -1,0 +1,206 @@
+#' **MassSpecMethod_SmoothSpectra_movingaverage**
+#'
+#' @description Smooths spectra using the moving average algorithm.
+#' 
+#' @param windowSize Numeric (length 1) with the window size for the moving average.
+#'
+#' @return A MassSpecMethod_SmoothSpectra_movingaverage object.
+#'
+#' @export
+#'
+MassSpecMethod_SmoothSpectra_movingaverage <- S7::new_class(
+  name = "MassSpecMethod_SmoothSpectra_movingaverage",
+  parent = ProcessingStep,
+  package = "StreamFind",
+  
+  constructor = function(windowSize = 5) {
+    
+    S7::new_object(
+      ProcessingStep(
+        engine = "MassSpec",
+        method = "SmoothSpectra",
+        algorithm = "movingaverage",
+        required = "LoadSpectra",
+        parameters = list(windowSize = windowSize),
+        number_permitted = Inf,
+        version = as.character(packageVersion("StreamFind")),
+        software = "StreamFind",
+        developer = "Ricardo Cunha",
+        contact = "cunha@iuta.de",
+        link = "https://odea-project.github.io/StreamFind",
+        doi = NA_character_
+      )
+    )
+  },
+  
+  validator = function(self) {
+    checkmate::assert_choice(self@engine, "MassSpec")
+    checkmate::assert_choice(self@method, "SmoothSpectra")
+    checkmate::assert_choice(self@algorithm, "movingaverage")
+    checkmate::assert_number(self@parameters$windowSize)
+    NULL
+  }
+)
+
+#' @export
+#' @noRd
+S7::method(run, MassSpecMethod_SmoothSpectra_movingaverage) <- function(x, engine = NULL) {
+  
+  if (!is(engine, "MassSpecEngine")) {
+    warning("Engine is not a MassSpecEngine object!")
+    return(FALSE)
+  }
+  
+  if (!engine$has_analyses()) {
+    warning("There are no analyses! Not done.")
+    return(FALSE)
+  }
+  
+  if (!engine$Analyses$has_spectra) {
+    warning("No spectra results object available! Not done.")
+    return(FALSE)
+  }
+  
+  windowSize <- x$parameters$windowSize
+  
+  spec_list <- engine$Spectra$spectra
+  
+  spec_list <- lapply(spec_list, function(x, windowSize) {
+    
+    if (nrow(x) > 0) {
+      
[…]

### 2025-02-13 R/class_RamanMethod_AddShiftValues.R

@@ -1,0 +1,147 @@
+#' **RamanMethod_AddShiftValues_native**
+#'
+#' @description Adds custom shift values to the Raman spectra or copies from one or more analyses to
+#' the other analyses. When more than one analyses are selected, the shift values are copied from
+#' the selected but them averaged before being added to the other analyses. If the column "rt" is
+#' present in the spectra, the shift values are added to each retention time.
+#' 
+#' @param mode A character string. The mode to be used, possible values are `analyses` or `custom`
+#' for copying from analyses or adding custom shift values, respectively.
+#' @param index An integer vector with the index/es of the analysis/analyses to copy the shift 
+#' values from.
+#' @param shifts A numeric vector with the shift values to be added to the spectra.
+#'
+#' @return A RamanMethod_AddShiftValues_native object.
+#'
+#' @export
+#'
+RamanMethod_AddShiftValues_native <- S7::new_class(
+  name = "RamanMethod_AddShiftValues_native",
+  parent = ProcessingStep,
+  package = "StreamFind",
+  constructor = function(mode = "analyses",
+                         index = NA_integer_,
+                         shifts = NA_real_) {
+    S7::new_object(
+      ProcessingStep(
+        engine = "Raman",
+        method = "AddShiftValues",
+        algorithm = "native",
+        parameters = list(
+          mode = mode,
+          index = index,
+          shifts = shifts
+        ),
+        number_permitted = 1,
+        version = as.character(packageVersion("StreamFind")),
+        software = "StreamFind",
+        developer = "Ricardo Cunha",
+        contact = "cunha@iuta.de",
+        link = "https://odea-project.github.io/StreamFind",
+        doi = NA_character_
+      )
+    )
+  },
+  validator = function(self) {
+    checkmate::assert_choice(self@engine, "Raman")
+    checkmate::assert_choice(self@method, "AddShiftValues")
+    checkmate::assert_choice(self@algorithm, "native")
+    checkmate::assert_choice(self@parameters$mode, c("analyses", "custom"))
+    checkmate::assert_integer(as.integer(self@parameters$index))
+    checkmate::assert_numeric(self@parameters$shifts)
+    NULL
+  }
+)
+
+#' @export
[…]

### 2025-02-13 R/class_RamanSettings_MergeSpectraTimeSeries.R

@@ -1,158 +1,0 @@
-# ______________________________________________________________________________________________________________________
-# StreamFind -----
-# ______________________________________________________________________________________________________________________
-
-#' **RamanSettings_MergeSpectraTimeSeries_StreamFind**
-#'
-#' @description Merges Raman spectra based on time series data. It collapses data files into a single file.
-#'
-#' @param preCut The number of pre Raman scans to exclude when merging.
-#'
-#' @return A RamanSettings_MergeSpectraTimeSeries_StreamFind object.
-#'
-#' @export
-#'
-RamanSettings_MergeSpectraTimeSeries_StreamFind <- S7::new_class("RamanSettings_MergeSpectraTimeSeries_StreamFind",
-  parent = ProcessingSettings,
-  package = "StreamFind",
-  constructor = function(preCut = 2) {
-    S7::new_object(ProcessingSettings(
-      engine = "Raman",
-      method = "MergeSpectraTimeSeries",
-      required = NA_character_,
-      algorithm = "StreamFind",
-      parameters = list(preCut = preCut),
-      number_permitted = Inf,
-      version = as.character(packageVersion("StreamFind")),
-      software = "StreamFind",
-      developer = "Ricardo Cunha",
-      contact = "cunha@iuta.de",
-      link = "https://odea-project.github.io/StreamFind",
-      doi = NA_character_
-    ))
-  },
-  validator = function(self) {
-    checkmate::assert_choice(self@engine, "Raman")
-    checkmate::assert_choice(self@method, "MergeSpectraTimeSeries")
-    checkmate::assert_choice(self@algorithm, "StreamFind")
-    checkmate::assert_number(self@parameters$preCut)
-    NULL
-  }
-)
-
-#' @export
-#' @noRd
-S7::method(run, RamanSettings_MergeSpectraTimeSeries_StreamFind) <- function(x, engine = NULL) {
-  if (!is(engine, "RamanEngine")) {
-    warning("Engine is not a RamanEngine object!")
-    return(FALSE)
-  }
-
-  if (!engine$has_analyses()) {
-    warning("There are no analyses! Not done.")
-    return(FALSE)
-  }
-
-  if (!engine$has_spectra()) {
-    warning("No spectra results object available! Not done.")
-    return(FALSE)
-  }
-
[…]

### 2025-02-13 dev/dev_nts_app.R

@@ -48,7 +48,7 @@ ms$add_replicate_names(rpls)
 ms$add_blank_names(blks)
 
 ms$run(
-  MassSpecSettings_FindFeatures_openms(
+  MassSpecMethod_FindFeatures_openms(
     noiseThrInt = 1000,
     chromSNR = 3,
     chromFWHM = 7,
@@ -75,7 +75,7 @@ ms$run(
 )
 
 ms$run(
-  MassSpecSettings_AnnotateFeatures_StreamFind(
+  MassSpecMethod_AnnotateFeatures_StreamFind(
     rtWindowAlignment = 0.3,
     maxIsotopes = 8,
     maxCharge = 2,
@@ -84,14 +84,14 @@ ms$run(
 )
 
 ms$run(
-  MassSpecSettings_FilterFeatures_StreamFind(
+  MassSpecMethod_FilterFeatures_StreamFind(
     excludeIsotopes = TRUE,
     excludeAdducts = TRUE
   )
 )
 
 ms$run(
-  MassSpecSettings_GroupFeatures_openms(
+  MassSpecMethod_GroupFeatures_openms(
     rtalign = FALSE,
     QT = FALSE,
     maxAlignRT = 5,
@@ -103,13 +103,13 @@ ms$run(
 )
 
 ms$run(
-  MassSpecSettings_FilterFeatures_StreamFind(
+  MassSpecMethod_FilterFeatures_StreamFind(
     minIntensity = 3000
   )
 )
 
 ms$run(
-  MassSpecSettings_FillFeatures_StreamFind(
+  MassSpecMethod_FillFeatures_StreamFind(
     withinReplicate = FALSE,
     rtExpand = 2,
     mzExpand = 0.0005,
@@ -122,7 +122,7 @@ ms$run(
 )
 
 ms$run(
-  MassSpecSettings_CalculateFeaturesQuality_StreamFind(
+  MassSpecMethod_CalculateFeaturesQuality_StreamFind(
     filtered = FALSE,
     rtExpand = 2,
     mzExpand = 0.0005,
@@ -133,13 +133,13 @@ ms$run(
 )
 
 ms$run(
-  MassSpecSettings_FilterFeatures_StreamFind(
+  MassSpecMethod_FilterFeatures_StreamFind(
     minSnRatio = 5
   )
 )
 
 ms$run(
-  MassSpecSettings_FilterFeatures_patRoon(
+  MassSpecMethod_FilterFeatures_patRoon(
[…]

### 2025-02-17 R/class_Analyses_RamanAnalyses.R
Improved MassSpec classes v1
ongoing

@@ -476,13 +476,17 @@ S7::method(get_spectra, RamanAnalyses) <- function(x,
         rpl <- rpl[analyses]
         spec$spectra <- spec$spectra[names(spec$spectra) %in% unname(rpl)]
         spec$spectra <- Map( function(z, y) {
-          z$replicate <- y
+          if (nrow(z) > 0) z$replicate <- y
           z
         }, spec$spectra, names(spec$spectra))
       } else {
+        rpl <- x$replicates
         spec$spectra <- spec$spectra[analyses]
         spec$spectra <- Map( function(z, y) {
-          z$analysis <- y
+          if (nrow(z) > 0) {
+            z$analysis <- y
+            z$replicate <- rpl[y]
+          }
           z
         }, spec$spectra, names(spec$spectra))
       }
@@ -491,14 +495,16 @@ S7::method(get_spectra, RamanAnalyses) <- function(x,
     if (spec$has_chrom_peaks) {
       if (length(spec$spectra) == length(spec$chrom_peaks[names(spec$spectra)])) {
         spec$spectra <- Map(function(z, y) {
-          z$id <- NA_character_
-          for (i in seq_len(nrow(y))) {
-            sel <- z$rt >= y$rtmin[i] & z$rt <= y$rtmax[i]
-            if (sum(sel) > 0) {
-              if ("group" %in% colnames(y)) {
-                z$id[sel] <- y$group[i]
-              } else {
-                z$id[sel] <- y$peak[i]
+          if (nrow(z) > 0) {
+            z$id <- NA_character_
+            for (i in seq_len(nrow(y))) {
+              sel <- z$rt >= y$rtmin[i] & z$rt <= y$rtmax[i]
+              if (sum(sel) > 0) {
+                if ("group" %in% colnames(y)) {
+                  z$id[sel] <- y$group[i]
+                } else {
+                  z$id[sel] <- y$peak[i]
+                }
               }
             }
           }
@@ -602,10 +608,8 @@ S7::method(plot_spectra, RamanAnalyses) <- function(x,
                                                     xLab = NULL,
                                                     yLab = NULL,
                                                     title = NULL,
-                                                    showLegend = TRUE,
                                                     colorBy = "analyses",
                                                     interactive = TRUE,
[…]

### 2025-02-17 R/generics.R

@@ -99,6 +99,14 @@ get_patRoon_MSPeakLists <- S7::new_generic("get_patRoon_MSPeakLists", "x")
 
 #' @export
 #' @noRd
+get_raw_chromatograms <- S7::new_generic("get_raw_chromatograms", "x")
+
+#' @export
+#' @noRd
+get_raw_spectra <- S7::new_generic("get_raw_spectra", "x")
+
+#' @export
+#' @noRd
 get_spectra <- S7::new_generic("get_spectra", "x")
 
 #' @export

### 2025-02-17 dev/dev_Core.R

@@ -34,7 +34,7 @@ core <- CoreEngine$new(Metadata = list(file = "core.rds"))
 # }
 
 
-
+plot_spectra_baseline(raman$Analyses, colorBy = "replicates", interactive = TRUE)
 
 # Resources -------
 ms_files <- StreamFindData::get_ms_file_paths()

### 2025-04-08 src/rcpp_ms_exported.cpp
Second boost on NTS c++ library

@@ -501,210 +501,6 @@ Rcpp::List rcpp_parse_ms_chromatograms(Rcpp::List analysis, std::vector<int> idx
   return out;
 };
 
-// MARK: rcpp_ms_annotate_features
-// [[Rcpp::export]]
-Rcpp::List rcpp_ms_annotate_features(Rcpp::List feature_list,
-                                     double rtWindowAlignment = 0.3,
-                                     int maxIsotopes = 5,
-                                     int maxCharge = 1,
-                                     int maxGaps = 1)
-{
-
-  NTS::MS_ISOTOPE_SET isotopes;
-
-  std::vector<std::string> elements = {"C", "H", "N", "O", "S", "Cl", "Br", "Si"};
-
-  isotopes.filter(elements);
-
-  const int max_number_elements = 5;
-
-  Rcpp::Rcout << "Building combinatorial isotopic chains with length " << max_number_elements << "...";
-  NTS::MS_ISOTOPE_COMBINATIONS combinations(isotopes, max_number_elements);
-  Rcpp::Rcout << "Done!" << std::endl;
-
-  const int number_analyses = feature_list.size();
-
-  if (number_analyses == 0)
-    return feature_list;
-
-  for (int a = 0; a < number_analyses; a++)
-  {
-
-    Rcpp::List features = feature_list[a];
-
-    const NTS::MS_FEATURES_MZ_SORTED fdf(features);
-
-    const std::vector<std::string> fts = features["feature"];
-
-    const int number_features = fdf.n;
-
-    NTS::MS_ANNOTATION af(number_features);
-
-    Rcpp::Rcout << "Annotating isotopes in " << number_features << " features...";
-
-    for (int f = 0; f < number_features; f++)
-    {
-
-      const int &index = fdf.index[f];
-
-      if (af.iso_step[index] > 0)
-        continue; // already isotope
-
-      const std::string &feature = fdf.feature[f];
-      const int &polarity = fdf.polarity[f];
-      const float &rt = fdf.rt[f];
-      float rtmin = fdf.rtmin[f];
-      float rtmax = fdf.rtmax[f];
-      const float &mz = fdf.mz[f];
-      const float &mzmin = fdf.mzmin[f];
-      const float &mzmax = fdf.mzmax[f];
-      const float max_mz_chain = (mz + maxIsotopes) * 1.05;
-
-      std::vector<int> candidates = NTS::find_isotopic_candidates(
-          number_features,
-          fdf.feature, fdf.mz, fdf.rt, fdf.polarity,
[…]

### 2025-04-16 README.Rmd
update README

@@
 
-<style>
-  p {
-    text-align: justify;
-  }
-</style>
-
 <!-- badges: start -->
@@
 
-<!-- <p align="center" width="100%">
-  <img width="60%" src="man/figures/logo_StreamFind.png" alt="Logo" />
-</p> -->
-
-```{r, echo=FALSE, out.width = '100%', fig.align='center'}
+```{r, echo=FALSE, out.width = '100%', fig.align='center', fig.alt="Logo"}
 knitr::include_graphics("man/figures/logos.png")
@@
 StreamFind aims to stimulate the use of advanced data analysis (e.g.
-non-target screening, statistical analysis, etc.) in routine studies, to promote standardisation of data structure and processing, and to facilitate retrospective data evaluation.
-The StreamFind platform is aimed at scientists, but also at technicians due to its comprehensive documentation, its well categorised set of integrated modular functions and its embedded graphical user interface.
+non-target screening, statistical analysis, etc.) in routine studies, to promote standardization of data structure and processing, and to facilitate retrospective data evaluation.
+The StreamFind platform is aimed at scientists, but also at technicians due to its comprehensive documentation, its well categorized set of integrated modular functions and its embedded graphical user interface.
 
@@
 
-The StreamFind depends on other open source software to process different analytical data.
+The StreamFind depends on other open source software to process different analytic data.
 For instance, for non-target screening using mass spectrometry the StreamFind uses the [patRoon](https://github.com/rickhelmus/patRoon) R package and its own dependencies.

### 2025-04-16 README.Rmd
README justify

@@
 
+<div style="text-align: justify;">
 StreamFind R package is developed within the project ["Flexible data analysis and workflow designer to identify chemicals in the water cycle"](https://www.bildung-forschung.digital/digitalezukunft/de/wissen/Datenkompetenzen/datenkompetenzen_wissenschaftlichen_nachwuchs/Projekte/stream_find.html) funded by the [German Federal Ministry of Education and Research (BMBF)](https://www.bmbf.de).
@@
 The StreamFind development is ongoing, please [contact us](mailto:cunha@iuta.de) for questions or collaboration.  
+</div>
 

### 2025-04-16 README.Rmd
README justify 2

--- 2025-04-16 README.Rmd
@@
 
+<p style="text-align: justify;">
 StreamFind R package is developed within the project ["Flexible data analysis and workflow designer to identify chemicals in the water cycle"](https://www.bildung-forschung.digital/digitalezukunft/de/wissen/Datenkompetenzen/datenkompetenzen_wissenschaftlichen_nachwuchs/Projekte/stream_find.html) funded by the [German Federal Ministry of Education and Research (BMBF)](https://www.bmbf.de).
@@
 The StreamFind development is ongoing, please [contact us](mailto:cunha@iuta.de) for questions or collaboration.  
+</p>
 

### 2025-04-19 dev/demo_nts.Rmd
Major fixes after check package

---
title: 'Non-target Screening (NTS) Wastewater Treatment Evaluation'
author:
  name: Ricardo Cunha
  email: cunha@iuta.de
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    fig_caption: true
    toc: true
    number_sections: true
    toc_float: true
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{'Non-target Screening (NTS) Wastewater Treatment Evaluation'}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# Libraries
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(plotly)
library(StreamFind)

knitr::opts_chunk$set(
  echo = TRUE,
  fig.align = "center",
  fig.width = 9,
  results = "markup",
  comment = "",
  message = FALSE,
  warning = FALSE
)
```

```{r resources, include=FALSE}
all_files <- StreamFindData::get_ms_file_paths()
files <- all_files[grepl("blank|influent|o3sw", all_files)]
db_all <- StreamFindData::get_ms_tof_spiked_chemicals()
db_all <- db_all[grepl("S", db_all$tag), ]
cols <- c("name", "formula", "mass", "rt", "tag")
db_is <- db_all[db_all$tag %in% "IS", ]
db_is <- db_is[, cols, with = FALSE]
db_is <- db_is[!db_is$name %in% c("Ibuprofen-d3", "Naproxen-d3"), ]
db <- db_all[db_all$tag %in% "S", ]
db <- db[, cols, with = FALSE]
db_with_ms2 <- StreamFindData::get_ms_tof_spiked_chemicals_with_ms2()
db_with_ms2 <- db_with_ms2[db_with_ms2$tag %in% "S", ]
db_with_ms2 <- db_with_ms2[, c("name", "formula", "mass", "SMILES", "rt", "polarity", "fragments"), with = FALSE]
db_with_ms2$polarity[db_with_ms2$polarity == 1] <- "positive"
db_with_ms2$polarity[is.na(db_with_ms2$polarity)] <- "positive"
db_with_ms2$polarity[db_with_ms2$polarity == -1] <- "negative"
```

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<br>

<br>

***

# Introduction

In this article we demonstrate how StreamFind can be used to evaluate ozonation of secondary wastewater effluent (i.e., effluent of the aerated biological treatment) using mass spectrometry (MS). A set of 18 *mzML* files are used, representing blank, influent and effluent measurements in triplicate for both positive and negative ionization modes.

```{r files}
basename(files)
```

The showcase will use the StreamFind *MassSpecEngine*, which encapsulates all tools required for parsing, storing, processing and visualizing MS data. Note that not all methods/functions will be shown as the demonstration focuses of the workflow to assess wastewater ozonation. Other processing methods are available in the StreamFind package and can be found in the [StreamFind reference documentation](https://odea-project.github.io/StreamFind/reference/index.html).

# *MassSpecEngine*

The [*R6*](https://r6.r-lib.org/) [*MassSpecEngine*](https://odea-project.github.io/StreamFind/reference/MassSpecEngine.html) class object is created using `MassSpecEngine$new()`, as shown below. The `analyses` argument can be used to add the set of ms files directly. In this demonstration, we use *mzML* files. However, the original ms vendor files can also be used, but they will be converted to *mzML* format using *msConvert* from [ProteoWizard](https://proteowizard.sourceforge.io/download.html) in the background.

```{r create-ms}
# Creates a MassSpecEngine from mzML files
ms <- MassSpecEngine$new(analyses = files)
```

```{r show-ms}
# Prints in console a summary of the MassSpecEngine
ms
```

# *Metadata*

Project headers (e.g., name, author and description) can be added to the `MassSpecEngine$headers` as a list object as shown below. The elements of the list can be anything but must have length one.

```{r add-headers}
# Adds headers to the MassSpecEngine
ms$headers <- list(
  name = "Wastewater Ozonation Showcase",
  author = "Ricardo Cunha",
  description = "Demonstration project"
)

# Gets and shows the headers
show(ms$headers)

# Gets the MassSpecEngine date
ms$headers$date
```

# Replicates and blanks

The analysis replicate names and the associated blank replicate name can be amended in the *MassSpecEngine*, as shown below. Alternatively, a `data.frame` with column names *file*, *replicate* and *blank* could be added as the `analyses` argument in `MassSpecEngine$new(analyses = files)` to have directly the replicate and blank replicate names assigned (more details [here](https://odea-project.github.io/StreamFind/reference/MassSpecEngine.html#method-MassSpecEngine-new)).

```{r add-replicates}
# Character vector with analysis replicate names
rpls <- c(
  rep("blank_neg", 3),
  rep("blank_pos", 3),
  rep("influent_neg", 3),
  rep("influent_pos", 3),
  rep("effluent_neg", 3),
  rep("effluent_pos", 3)
)

# Character vector with associated blank replicate names
# Note that the order should match the respective replicate
blks <- c(
  rep("blank_neg", 3),
  rep("blank_pos", 3),
  rep("blank_neg", 3),
  rep("blank_pos", 3),
  rep("blank_neg", 3),
  rep("blank_pos", 3)
)

# Amends replicate and blank names
ms$add_replicate_names(rpls)
ms$add_blank_names(blks)

# Replicates and blanks were amended
ms$analyses$info[, 1:5]
```

# *ProcessingStep*

Data processing is performed by steps according to *ProcessingStep*. S7 *ProcessingStep* class objects are obtained via the respective [`[Engine type]Settings_[module name]_[algorithm name]`](https://odea-project.github.io/StreamFind/reference/index.html#processing-settings-for-ms-data) constructor functions, attributing the respective subclass. Below we obtain the *ProcessingStep* for the step `FindFeatures` using the algorithm *openms*. The parameters for each processing module can be changed via the constructor arguments. Documentation for each *ProcessingStep* subclass can be found in the [StreamFind reference documentation](https://odea-project.github.io/StreamFind/reference/index.html).

```{r ffs}
# Gets ProcessingStep for finding features using the openms algorithm
ffs <- MassSpecSettings_FindFeatures_openms(
  noiseThrInt = 1000,
  chromSNR = 3,
  chromFWHM = 7,
  mzPPM = 15,
  reEstimateMTSD = TRUE,
  traceTermCriterion = "sample_rate",
  traceTermOutliers = 5,
  minSampleRate = 1,
  minTraceLength = 4,
  maxTraceLength = 70,
  widthFiltering = "fixed",
  minFWHM = 4,
  maxFWHM = 35,
  traceSNRFiltering = TRUE,
  localRTRange = 0,
  localMZRange = 0,
  isotopeFilteringModel = "none",
  MZScoring13C = FALSE,
  useSmoothedInts = FALSE,
  intSearchRTWindow = 3,
  useFFMIntensities = FALSE,
  verbose = FALSE
)

# Prints in console the details of the ProcessingStep
show(ffs)
```

```{r other-settings}
# Creates an ordered list with all processing steps for the MS data
workflow <- list(

  # Find features using the openms algorithm, created above
  ffs,

  # Annotation of natural isotopes and adducts
  MassSpecSettings_AnnotateFeatures_StreamFind(
    rtWindowAlignment = 0.3,
    maxIsotopes = 8,
    maxCharge = 2,
    maxGaps = 1
  ),

  # Excludes annotated isotopes and adducts
  MassSpecSettings_FilterFeatures_StreamFind(
    excludeIsotopes = TRUE,
    excludeAdducts = TRUE
  ),

  # Grouping features across analyses
  MassSpecSettings_GroupFeatures_openms(
    rtalign = FALSE,
    QT = FALSE,
    maxAlignRT = 5,
    maxAlignMZ = 0.008,
    maxGroupRT = 5,
    maxGroupMZ = 0.008,
    verbose = FALSE
  ),

  # Filter feature groups with maximum intensity below 5000 counts
  MassSpecSettings_FilterFeatures_StreamFind(
    minIntensity = 3000
  ),

  # Fill features with missing data
  # Reduces false negatives
  MassSpecSettings_FillFeatures_StreamFind(
    withinReplicate = FALSE,
    rtExpand = 2,
    mzExpand = 0.0005,
    minTracesIntensity = 1000,
    minNumberTraces = 6,
    baseCut = 0.3,
    minSignalToNoiseRatio = 3,
    minGaussianFit = 0.2
  ),

  # Calculate quality metrics for each feature
  MassSpecSettings_CalculateFeaturesQuality_StreamFind(
    filtered = FALSE,
    rtExpand = 2,
    mzExpand = 0.0005,
    minTracesIntensity = 1000,
    minNumberTraces = 6,
    baseCut = 0
  ),

  # Filter features based on minimum signal-to-noise ratio (s/n)
  # The s/n is calculated using the CalculateFeaturesQuality method
  MassSpecSettings_FilterFeatures_StreamFind(
    minSnRatio = 5
  ),

  # Filter features using other parameters via the patRoon package
  MassSpecSettings_FilterFeatures_patRoon(
    maxReplicateIntRSD = 40,
    blankThreshold = 5,
    absMinReplicateAbundance = 3
  ),

  # Finds internal standards in the MS data
  # db_is is a data.table with the
  # name, mass and expected retention time of
  # spiked internal standards, as shown below
  MassSpecSettings_FindInternalStandards_StreamFind(
    database = db_is,
    ppm = 8,
    sec = 10
  ),

  # Corrects matrix suppression using the TiChri method from
  # 10.1021/acs.analchem.1c00357 to better compare influent and effluent
  MassSpecSettings_CorrectMatrixSuppression_TiChri(
    mpRtWindow = 10,
    istdAssignment = "range",
    istdRtWindow = 50,
    istdN = 2
  ),

  # Loads MS1 for features not filtered
  MassSpecSettings_LoadFeaturesMS1_StreamFind(
    filtered = FALSE
  ),

  # Loads MS2 for features not filtered
  MassSpecSettings_LoadFeaturesMS2_StreamFind(
    filtered = FALSE
  ),

  # Loads feature extracted ion chromatograms (EIC)
  MassSpecSettings_LoadFeaturesEIC_StreamFind(
    filtered = FALSE
  ),

  # Performs suspect screening using the StreamFind algorithm
  # db_with_ms2 is a database with suspect chemical standards
  # includes MS2 data (i.e., fragmentation pattern) from standards
  MassSpecSettings_SuspectScreening_StreamFind(
    database = db_with_ms2,
    ppm = 10,
    sec = 15,
    ppmMS2 = 10,
    minFragments = 3
  )
)
```

Then, the list can be added to the *MassSpecEngine*. Note that the order will matter when the workflow is applied!

```{r add-settings}
# Adds the workflow to the engine. The order matters!
ms$workflow <- workflow

# Printing the data processing workflow
ms$print_workflow()
```

# `run_workflow()`

The *Workflow* can be applied by `run_workflow()`, as demonstrated below. Note that with `run_workflow()`, the processing modules are applied with the same order as they were added.

```{r do-workflow, results='hide'}
# Runs all ProcessingStep added
ms$run_workflow()
```

# Results

The created features and feature groups can be inspected as `data.table` objects or plotted by dedicated methods in the *MassSpecEngine*. Internally, the *MassSpecEngine* stores the results in the *NTS* object, which can be accessed with `MassSpecEngine$NTS`. Yet, the engine interface is recommended for accessing the results. However, the *NTS* object can be used for more advanced operations, such as exporting the results to a database or other formats or use the native objects from other packages (e.g., [patRoon](https://github.com/rickhelmus/patRoon)) as we demonstrate further in this article.

## `data.table` objects

The features and feature groups can be obtained as `data.table` with the `MassSpecEngine$get_features()` and `MassSpecEngine$get_groups()` methods. The methods also allow to look for specific features/feature groups using mass, mass-to-charge ratio, retention time and drift time targets, as show below for a small set of compound targets where mass and retention time expected values are known. Note that drift time is only applicable for MS data with ion mobility separation.

```{r database-show}
db
```

```{r show-target-groups}
# Compounds are searched by monoisotopic mass and retention time
# ppm and sec set the mass (im ppm) and time (in seconds) allowed deviation, respectively
# average applies a mean to the intensities in each analysis replicate group
ms$get_groups(mass = db, ppm = 10, sec = 15, average = TRUE)
```

Already by inspection of the `data.table`, it is possible to see compounds detected in the influent but not in the effluent (e.g., Carbamazepine) or compounds that are appear to be reduced during ozonation (e.g., Metoprolol). Since positive and negative ionization mode were combined, there are compounds that appear in both polarities and are grouped by neutral monoisotopic mass (e.g., Candesartan and Diclofenac).

## `plot_groups` methods

For a better overview of the results, the method `MassSpecEngine$plot_groups()` or even more detailed the method `MassSpecEngine$plot_groups_overview()` can be used.

```{r plot-groups}
# set legendNames to TRUE for using the names in db as legend
ms$plot_groups(mass = db, ppm = 10, sec = 15, legendNames = TRUE)
```

```{r plot-groups-overview, fig.height=10}
ms$plot_groups_overview(mass = db, ppm = 5, sec = 10, legendNames = TRUE)
```

## Filtered not removed

The *FilterFeatures* method was applied to filter features according to defined conditions/thresholds. Yet, the filtered features were not removed but just tagged as filtered. For instance, when the method `MassSpecEngine$get_groups()` is run with `filtered` argument set to `TRUE`, the filtered features are also shown. Below, we search for the same compounds as above but with the `filtered` argument set to `TRUE`. Potential features from Valsartan are now returned but were filtered due to low intensity. Note that when extracting features based on basic parameters, i.e. mass and time, does not mean that features are identified. The identification of features is a more complex process and requires additional information, such as MS/MS data as in the processing method suspect screening.

```{r show-filtered-is}
# Set filtered to TRUE for showing filtered features/feature groups
ms$get_groups(mass = db, ppm = 5, sec = 10, average = TRUE, filtered = TRUE)
```

## Internal Standards

The method *FindInternalStandards* was applied for tagging spiked internal standards and the results can be obtained with the dedicated method `MassSpecEngine$get_internal_standards()` or plotted as a quality overview using the method `MassSpecEngine$plot_internal_standards()`, as shown below. The plot gives an overview of the mass, retention time and intensity variance of the internal standards across the analyses in the project.

```{r database-is-show}
# List of spiked internal standards
db_is
```

```{r get_is}
# Gets the internal standards evaluation data.table
ms$get_internal_standards()
```

```{r plot_is_qc, fig.height=10, fig.cap="Quality control of spiked internal standards"}
ms$plot_internal_standards()
```

```{r plot_is_profile, fig.cap="Internal standards profile across analyses"}
ms$plot_groups_profile(mass = db_is, ppm = 8, sec = 10, filtered = TRUE, legendNames = TRUE)
```

## Components

The method *AnnotateFeatures* was applied to annotate the natural isotopes and adducts into components. Implementation of annotation for in-source fragments is planned but not yet available with the StreamFind algorithm. The method `MassSpecEngine$get_components()` can be used to search for components, as shown below for the analysis number 11. Because the filters `excludeIsotopes` and `minIntensity` were applied, the isotopic features are likely filtered.

```{r show-components}
# Components of Diclofenac and Candesartan in analysis 11
components_example <- ms$get_components(
  analyses = 11,
  mass = db[db$name %in% c("Diclofenac", "Candesartan"), ],
  ppm = 5, sec = 10,
  filtered = TRUE
)

# Subset of the components data.table
components_example[, c(2:6, 8, 30:34)]
```

The components (i.e., isotopes and adducts) can also be visualized with the method `MassSpecEngine$map_components()`, as shown below for the internal standards added to analysis 11. Note that again the `filtered` argument was set to `TRUE` to return also filtered features, as the internal standards were likely excluded by blank subtraction.

```{r plot-components, fig.cap="Components (i.e., isotopes and adducts) of internal standards in analysis 11"}
ms$map_components(
  analyses = 11,
  mass = db_is,
  ppm = 8, sec = 10,
  filtered = TRUE,
  legendNames = TRUE
)
```

## Suspects

The methods `MassSpecEngine$get_suspects()` and `MassSpecEngine$plot_suspects()` can be used to inspect the suspect screening results. In the plot function, a second plot is added to compare the experimental fragmentation pattern (top) with the fragmentation pattern of the respective reference standard (down) added within the database. The `colorBy` argument can be set to *targets+replicates* to legend the plot with combined keys of suspect target names and analysis replicate names.

```{r get_suspects}
ms$get_suspects()
```

```{r plot-suspects}
ms$plot_suspects(colorBy = "targets+replicates")
```

## Methods from patRoon

The NTS object holds methods to obtain original objects from the [patRoon](https://github.com/rickhelmus/patRoon) package. For instance, the S4 class `features` or `featureGroups` objects can be obtained via the `get_patRoon_features` method of the *NTS* results. The `patRoon` package provides a comprehensive set of functions, as shown below. See more information in the [patRoon reference documentation](https://rickhelmus.github.io/patRoon/reference/index.html).

```{r pat-features}
# Native patRoon object
fGroups <- get_patRoon_features(ms$NTS, filtered = FALSE, featureGroups = TRUE)

fGroups
```

```{r pat-plot, fig.alt="UpSet plot of features"}
# Using the native patRoon's plotUpSet method
patRoon::plotUpSet(fGroups)
```

## Fold-change analysis

The method `get_fold_change()` and correspondent `plot_fold_change()` can be used to calculate and plot the fold-change between influent and effluent samples. The method calculates the fold-change for each feature group and replicates, as shown below. The plot shows the fold-change for each feature group across the replicates. The fold-change is calculated according to [Bader et al. (2017)](https://pubs.acs.org/doi/10.1021/acs.analchem.7b03037), leveraging the replicates variance to increase the significance of the fold-change. Formation is not in the plot as no new features were detected in the effluent of the wastewater ozonation treatment step.

```{r plot-fold-change}
ms$plot_fold_change(
  replicatesIn = c("influent_neg", "influent_pos"),
  replicatesOut = c("effluent_neg", "effluent_pos"),
  filtered = FALSE,
  constantThreshold = 0.5,
  eliminationThreshold = 0.2,
  correctIntensity = TRUE,
  fillZerosWithLowerLimit = FALSE, # set to TRUE for filling zeros with lower limit argument
  lowerLimit = 200
)
```

Alternatively, the `plotVolcano` from [patRoon](https://github.com/rickhelmus/patRoon) can be used to plot fold-change. For more information check the [patRoon reference documentation](https://rickhelmus.github.io/patRoon/reference/index.html). Note that the function `patRoon::getFCParams` is used to define the parameters for the fold-change calculation. Below we use the default argument values, only the in and out replicate names are set.

```{r}
patRoon::plotVolcano(
  fGroups, # obtained above with get_patRoon_features
  patRoon::getFCParams(c("influent_pos", "effluent_pos"))
)
```

# More to come

Future assets/functionalities:

- Annotation of in-source fragments
- Screening of transformation products using the *biotransformer* tool via patRoon

***

### 2025-04-19 dev/demo_paracetamol_quantification_mcrals.Rmd
Major fixes after check package

---
title: 'Pharmaceutical API Quantification with Raman and MCR'
subtitle: |
  | Raman Spectroscopy Data and Multivariate Curve Resolution with Alternating Least Squares (MCR-ALS)
  | <br>
author:
  name: Ricardo Cunha
  email: cunha@iuta.de
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    fig_caption: true
    toc: true
    number_sections: true
    toc_float: true
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Pharmaceutical API Quantification with Raman and MCR}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(plotly)
library(data.table)
library(StreamFind)

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 9,
  results = "markup",
  comment = ""
)

raman_dir <- "C:/Users/apoli/Documents/example_files/isoft/Paracetamol_raman_quantification"
raman_files <- list.files(raman_dir, pattern = ".asc$", full.names = TRUE)
```

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<br>
<br>

***

This article demonstrates how StreamFind can be used to quantity the active product ingredient (API) in Fizamol (i.e., 
paracetamol) from Raman data and using the Multivariate Curve Resolution with Alternating Least Squares (MCR-ALS) 
statistical method. A calibration curve is acquired with pure paracetamol diluted in water at different concentrations.
Two Fizamol tablets (*fiz_a* and *fiz_b*) are diluted in water to a concentration of approx. 10 mg/ml and pure 
paracetamol (*par*) at 10 mg/ml are measured for quantification. Water is used as a blank.

```{r raman-setup}
# Creates the engine for Raman data
raman <- RamanEngine$new(analyses = raman_files)

# Modifies the replicate names
raman$analyses$replicates <- c(
  rep("blank", 9),
  rep("fiz_a", 9),
  rep("fiz_b", 9),
  rep("par", 3),
  rep("cali_100_mgml", 9),
  rep("cali_150_mgml", 9),
  rep("cali_010_mgml", 9),
  rep("cali_025_mgml", 9),
  rep("cali_050_mgml", 9),
  rep("cali_075_mgml", 9)
)

# Modifies the blank names
raman$analyses$blanks <- c(rep("blank", length(raman$analyses)))
```

<br>
<br>

```{r raman-overview, echo=FALSE}
ov_quant <- raman$get_overview()[, 1:3]
data.table::setorder(ov_quant, replicate)
ov_quant$concentration <- as.numeric(c(rep(0, 9), sub(".*?(\\d+).*", "\\1", ov_quant$replicate[10:63]), rep(0, nrow(ov_quant) - 63))) / 10
ov_quant$concentration[ov_quant$replicate %in% c("par", "fiz_a", "fiz_b")] <- NA_real_
kable(ov_quant, caption = "Overview of Raman analyses and respective analysis replicate groups and blanks.") %>%
  kable_styling(font_size = 12, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE) %>%
  scroll_box(width = "100%", height = "400px")
```

<br>
<br>

```{r plot-raman-conc-raw, fig.cap="Raw Raman spectra for each analysis replicate."}
raman$plot_spectra(colorBy = "replicates")
```

# Pre-processing Raman Data

The data is pre-processed using a workflow that includes the following steps:

1. Averaging the spectra for each analysis replicate.
2. Subtracting the blank spectra.
3. Smoothing the spectra using the Savitzky-Golay method.
4. Deleting the spectra sections that are not relevant for the analysis.
5. Correcting the spectra baseline using the Asymmetric Least Squares (airPLS) method.

The methods ca be applied one at the time or used to bluid a Workflow object that will apply all methods in sequence 
within engine, as shown below.

```{r raman-conc-workflow}
# Creates the workflow for Raman data
# with an ordered list of processing settings
raman_workflow <- list(
  RamanSettings_AverageSpectra_native(),
  RamanSettings_SubtractBlankSpectra_StreamFind(),
  RamanSettings_SmoothSpectra_savgol(fl = 7, forder = 4, dorder = 0),
  RamanSettings_DeleteSpectraSection_native(min = -100, max = 350),
  RamanSettings_DeleteSpectraSection_native(min = 1800, max = 2500),
  RamanSettings_CorrectSpectraBaseline_airpls(lambda = 11, differences = 1, itermax = 30)
)

# Adds the workflow to the engine
raman$workflow <- raman_workflow

raman$print_workflow()
```

```{r raman-conc-run-workflow}
# Runs the workflow
raman$run_workflow()
```

```{r plot-raman-conc-spectra, fig.cap="Pre-processed Raman spectra for analysis replicate."}
raman$plot_spectra()
```

## Applying MCR-ALS for Quantification

The quantification of paracetamol from the pre-processed Raman spectra is performed using the Multivariate Curve 
Resolution (MCR) with Alternating Least Squares (ALS) method via integration of the [mdatools](https://mdatools.com/docs/mcr--als.html) R package.

```{r}
# Gets a matrix with spectra from the Raman engine
raman_matrix <- raman$get_spectra_matrix()
str(raman_matrix)
```

```{r mcrals-show}
# Creates the statistic engine for the statistic analysis
stat <- StatisticEngine$new(analyses = raman_matrix)

stat
```

```{r}
# Applies the MCR-ALS method with the run method of the engine
stat$run(StatisticSettings_MakeModel_mcrals_mdatools(ncomp = 2))
```

```{r mcrals-resolved-spec, fig.cap="Overview plot for the MCR-ALS model"}
plot(stat$model)
```

```{r}
# Applies the quantification method with the run method of the engine
# Concentrations with NA for unknown values
stat$run(
  StatisticSettings_Quantify_mcrals(
    concentrations = c(1, 2.5, 5.0, 7.5, 10, 15.0, NA_real_, NA_real_, NA_real_)
  )
)
```

The MCR-ALS was applied with two components and the quantification results are shown below. The component with the 
highest combined explained variance and R<sup>2</sup> values for the calibration curve should be used for quantification. 
The analysis *par* should have about 10 in mg/ml of paracetamol while the analyses *fiz_a* and *fiz_b* should a little 
less than 10 mg/ml due to presence of formulation ingredients and possible loss of API during sample preparation.

```{r}
# Gets the quantification results
stat$quantification@quantities
```

### 2025-04-19 R/class_AuditTrail.R

@@ -1,48 +1,137 @@
+# MARK: AuditTrailEntry
+# AuditTrailEntry -----
+#' @title AuditTrailEntry Class
+#' 
+#' @description The AuditTrailEntry class is used as an element in the [StreamFind::AuditTrail]
+#' entries.
+#' 
+#' @param time_stamp A POSIXct timestamp of the entry.
+#' @param value_class A character string representing the class of the entry.
+#' @param value_parent A character string representing the parent class of the entry.
+#' @param value A character string representing the value of the entry.
+#' 
 #' @export
-#' @noRd
-AuditTrail <- S7::new_class(
-  name = "AuditTrail",
+#'
+AuditTrailEntry <- S7::new_class(
+  name = "AuditTrailEntry",
   package = "StreamFind",
   properties = list(
-    entries = S7::new_property(class = S7::class_list, default = list())
+    time_stamp = S7::new_property(S7::class_POSIXct, default = as.POSIXct(Sys.time())),
+    value_class = S7::new_property(S7::class_character, default = NA_character_),
+    value_parent = S7::new_property(S7::class_character, default = NA_character_),
+    value = S7::new_property(S7::class_character, default = NA_character_)
   ),
-  
-  constructor = function() {
-    S7::new_object(S7::S7_object(), entries = list())
+  constructor = function(time_stamp, value_class, value_parent, value) {
+    S7::new_object(
+      S7::S7_object(),
+      time_stamp = as.POSIXct(time_stamp),
+      value_class = value_class,
+      value_parent = value_parent,
+      value = value
+    )
   },
-  
   validator = function(self) {
-    checkmate::assert_list(self@entries)
+    checkmate::assert_posixct(self@time_stamp)
+    checkmate::assert_character(self@value_class)
+    checkmate::assert_character(self@value_parent)
+    checkmate::assert_character(self@value)
     NULL
   }
 )
 
 #' @export
 #' @noRd
-S7::method(length, AuditTrail) <- function(x) {
+`$.AuditTrailEntry` <- function(x, i) {
+  S7::prop(x, i)
+}
+
+#' @export
+#' @noRd
+`$<-.AuditTrailEntry` <- function(x, i, value) {
+  S7::prop(x, i) <- value
+}
+
[…]

### 2025-04-19 R/class_MassSpecMethod_FilterFeatures.R

@@ -44,7 +44,7 @@ MassSpecMethod_FilterFeatures_StreamFind <- S7::new_class(
                          onlyWithMS2 = FALSE) {
     S7::new_object(
       ProcessingStep(
-        engine = "MassSpec",
+        data_type = "MassSpec",
         method = "FilterFeatures",
         required = "FindFeatures",
         algorithm = "StreamFind",
@@ -72,7 +72,7 @@ MassSpecMethod_FilterFeatures_StreamFind <- S7::new_class(
     )
   },
   validator = function(self) {
-    checkmate::assert_choice(self@engine, "MassSpec")
+    checkmate::assert_choice(self@data_type, "MassSpec")
     checkmate::assert_choice(self@method, "FilterFeatures")
     checkmate::assert_choice(self@algorithm, "StreamFind")
     checkmate::assert_logical(self@parameters$correctIntensity, len = 1)
@@ -749,7 +749,7 @@ MassSpecMethod_FilterFeatures_patRoon <- S7::new_class(
                          negate = FALSE) {
     S7::new_object(
       ProcessingStep(
-        engine = "MassSpec",
+        data_type = "MassSpec",
         method = "FilterFeatures",
         required = "FindFeatures",
         algorithm = "patRoon",
@@ -797,7 +797,7 @@ MassSpecMethod_FilterFeatures_patRoon <- S7::new_class(
     )
   },
   validator = function(self) {
-    checkmate::assert_choice(self@engine, "MassSpec")
+    checkmate::assert_choice(self@data_type, "MassSpec")
     checkmate::assert_choice(self@method, "FilterFeatures")
     checkmate::assert_choice(self@algorithm, "patRoon")
     checkmate::assert_numeric(self@parameters$absMinIntensity)

### 2025-04-19 R/class_RamanMethod_GroupPeaks.R

@@ -15,7 +15,7 @@ RamanMethod_GroupPeaks_native <- S7::new_class(
   constructor = function(fillMissingPeaks = TRUE) {
     S7::new_object(
       ProcessingStep(
-        engine = "Raman",
+        data_type = "Raman",
         method = "GroupPeaks",
         algorithm = "native",
         parameters = list(
@@ -32,7 +32,7 @@ RamanMethod_GroupPeaks_native <- S7::new_class(
     )
   },
   validator = function(self) {
-    checkmate::assert_choice(self@engine, "Raman")
+    checkmate::assert_choice(self@data_type, "Raman")
     checkmate::assert_choice(self@method, "GroupPeaks")
     checkmate::assert_choice(self@algorithm, "native")
     checkmate::assert_logical(self@parameters$fillMissingPeaks, max.len = 1)

### 2025-04-22 R/class_RamanMethod_BinScans.R
Improved documentation

@@ -1,4 +1,4 @@
-#' **RamanMethod_BinScans_native**
+#' Raman Method for Binning Time Scans (native algorithm)
 #'
 #' @description Merges scans by averaging spectra according to a given binning size.
 #' 
@@ -112,6 +112,8 @@ S7::method(run, RamanMethod_BinScans_native) <- function(x, engine = NULL) {
     ushifts <- unique(z$shift)
     urts_size <- length(urts)
     ushifts_size <- length(ushifts)
+    
+    .SD <- NULL
     
     if ("scans" %in% binning_mode) {
       bin_key <- cut(seq_len(urts_size), seq(1, urts_size, binning_value), include.lowest = TRUE)

### 2025-04-24 dev/demo_nts.Rmd
Update NTS to NonTargetAnalysisResults
other fixes

@@
 ---
-title: 'Non-target Screening (NTS) Wastewater Treatment Evaluation'
+title: 'Non-target Screening (NonTargetAnalysisResults) Wastewater Treatment Evaluation'
 author:
@@
   %\VignetteEncoding{UTF-8}
-  %\VignetteIndexEntry{'Non-target Screening (NTS) Wastewater Treatment Evaluation'}
+  %\VignetteIndexEntry{'Non-target Screening (NonTargetAnalysisResults) Wastewater Treatment Evaluation'}
   %\VignetteEngine{knitr::rmarkdown}
@@
 The created features and feature groups can be inspected as `data.table` objects or plotted by dedicated methods in the *MassSpecEngine*.
-Internally, the *MassSpecEngine* stores the results in the *NTS* object, which can be accessed with `MassSpecEngine$NTS`.
+Internally, the *MassSpecEngine* stores the results in the *NonTargetAnalysisResults* object, which can be accessed with `MassSpecEngine$NonTargetAnalysisResults`.
 Yet, the engine interface is recommended for accessing the results.
-However, the *NTS* object can be used for more advanced operations, such as exporting the results to a database or other formats or use the native objects from other packages (e.g., [patRoon](https://github.com/rickhelmus/patRoon)) as we demonstrate further in this article.
+However, the *NonTargetAnalysisResults* object can be used for more advanced operations, such as exporting the results to a database or other formats or use the native objects from other packages (e.g., [patRoon](https://github.com/rickhelmus/patRoon)) as we demonstrate further in this article.
 
@@
 
-The NTS object holds methods to obtain original objects from the [patRoon](https://github.com/rickhelmus/patRoon) package.
-For instance, the S4 class `features` or `featureGroups` objects can be obtained via the `get_patRoon_features` method of the *NTS* results.
+The NonTargetAnalysisResults object holds methods to obtain original objects from the [patRoon](https://github.com/rickhelmus/patRoon) package.
+For instance, the S4 class `features` or `featureGroups` objects can be obtained via the `get_patRoon_features` method of the *NonTargetAnalysisResults* results.
 The `patRoon` package provides a comprehensive set of functions, as shown below.
@@
 # Native patRoon object
-fGroups <- get_patRoon_features(ms$NTS, filtered = FALSE, featureGroups = TRUE)
+fGroups <- get_patRoon_features(ms$NonTargetAnalysisResults, filtered = FALSE, featureGroups = TRUE)
 

### 2025-04-30 README.Rmd

--- 2025-04-16 README.Rmd
@@
 
-StreamFind R package is developed within the project ["Flexible data analysis and workflow designer to identify chemicals in the water cycle"](https://www.bildung-forschung.digital/digitalezukunft/de/wissen/Datenkompetenzen/datenkompetenzen_wissenschaftlichen_nachwuchs/Projekte/stream_find.html) funded by the [German Federal Ministry of Education and Research (BMBF)](https://www.bmbf.de).
+StreamFind R package is developed within the project ["Flexible data analysis and workflow designer to identify chemicals in the water cycle"](https://www.bildung-forschung.digital/digitalezukunft/de/bildung/digital-_und_datenkompetenzen/datenkompetenzen_wissenschaftlichen_nachwuchs/Projekte/stream_find.html) funded by the [German Federal Ministry of Education and Research (BMBF)](https://www.bmbf.de).
 The development is carried out by the [Institut für Umwelt & Energie, Technik & Analytik e.
@@
 StreamFind is intended to be a platform for assembling processing workflows for different types of data (e.g.
-mass spectrometry and spectroscopy) with applications in different fields (e.g.
-environmental and quality studies of the water cycle and quality control of pharmaceuticals).
+mass spectrometry (MS) and spectroscopy data) with applications in different fields (e.g.
+environmental studies of the water cycle and quality control of pharmaceuticals).
 StreamFind aims to stimulate the use of advanced data analysis (e.g.
 non-target screening, statistical analysis, etc.) in routine studies, to promote standardization of data structure and processing, and to facilitate retrospective data evaluation.
-The StreamFind platform is aimed at scientists, but also at technicians due to its comprehensive documentation, its well categorized set of integrated modular functions and its embedded graphical user interface.
+The StreamFind platform is aimed at scientists, but also at technicians due to its comprehensive documentation, its well categorized set of integrated modular processing methods and its embedded graphical user interface.
 
@@
 RTools is needed for compiling C++ code used in the StreamFind R package.
-StreamFind also uses python scripts for some of its functionalities, so it is recommended to have the latest [python](https://www.python.org/downloads/) installed and added to the environmental variables for Windows users.
+StreamFind also uses python scripts for some of its processing methods, so it is recommended to have the latest [python](https://www.python.org/downloads/) installed and added to the environmental variables for Windows users.
 Assuming that R, RTools and python are installed, the StreamFind R package can be installed from the GitHub repository via the [BiocManager](https://www.bioconductor.org/install/).
+Note that StreamFind is not yet available on the Bioconductor repository.
 
@@
 The StreamFind depends on other open source software to process different analytic data.
-For instance, for non-target screening using mass spectrometry the StreamFind uses the [patRoon](https://github.com/rickhelmus/patRoon) R package and its own dependencies.
-Installation instructions for patRoon and its dependencies can be found [here](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites).
-When a dependency is not installed, a warning message will be displayed when necessary with instructions.
-Consult the documentation for dependencies of other data types.
+For instance, for non-target screening using mass spectrometry the StreamFind uses the [patRoon](https://github.com/rickhelmus/patRoon) R package and its own dependencies for some processing methods.
+When a dependency is not installed, a warning message will be displayed with instructions.
+Consult the documentation of each processing method for dependencies.
 
@@
 The documentation and usage examples of the StreamFind R package can be found in the [reference page](https://odea-project.github.io/StreamFind/reference/index.html) and [articles](https://odea-project.github.io/StreamFind/articles/index.html) of the [webpage](https://odea-project.github.io/StreamFind/index.html), respectively.
+Note that documentation and articles are under development so not all are yet available.
 

### 2025-05-03 dev/demo_nts.Rmd
Small fix to RamanAnalyses and Quantification

@@
 ---
-title: 'Non-target Screening (NonTargetAnalysisResults) Wastewater Treatment Evaluation'
+title: 'Non-target Analysis for Wastewater Treatment Evaluation'
 author:
@@
   %\VignetteEncoding{UTF-8}
-  %\VignetteIndexEntry{'Non-target Screening (NonTargetAnalysisResults) Wastewater Treatment Evaluation'}
+  %\VignetteIndexEntry{'Non-target Analysis for Wastewater Treatment Evaluation'}
   %\VignetteEngine{knitr::rmarkdown}
@@
 The showcase will use the StreamFind *MassSpecEngine*, which encapsulates all tools required for parsing, storing, processing and visualizing MS data.
-Note that not all methods/functions will be shown as the demonstration focuses of the workflow to assess wastewater ozonation.
-Other processing methods are available in the StreamFind package and can be found in the [StreamFind reference documentation](https://odea-project.github.io/StreamFind/reference/index.html).
+Note that not all methods/functions will be shown, as the demonstration focuses of the workflow to assess wastewater treatment.
+Other processing methods for MS data are available in the StreamFind package and can be found in the [StreamFind reference documentation](https://odea-project.github.io/StreamFind/reference/index.html).
 

### 2025-06-13 dev/dev_statistics.Rmd

@@
 # Smoothing of the TIC chromatogram for improving peak finding
-ms$smooth_chromatograms(Settings_smooth_chromatograms_movingaverage(windowSize = 5))
+ms$smooth_chromatograms(Method_smooth_chromatograms_movingaverage(windowSize = 5))
 ```
@@
 # Settings for integration of the TIC
-ps_ic <- Settings_integrate_chromatograms_StreamFind(
+ps_ic <- Method_integrate_chromatograms_StreamFind(
   merge = TRUE,
@@
 
-According to the MS1 spectra, the *m/z* from 1200 to 2000 can be added to the dedicated *ProcessingSettings* for deconvolution, as shown below.
+According to the MS1 spectra, the *m/z* from 1200 to 2000 can be added to the dedicated *ProcessingStep* for deconvolution, as shown below.
 The first step is to load from the mzML files the traces within the retention time and *m/z* ranges selected.
@@
 
-# ms$cluster_spectra(Settings_cluster_spectra_StreamFind(val = "mz", clustVal = 0.01, presence = 0.2))
-# ms$calculate_spectra_charges(Settings_calculate_spectra_charges_StreamFind(roundVal = 15, relLowCut = 0.2, absLowCut = 8000))
-# ms$deconvolute_spectra(Settings_deconvolute_spectra_StreamFind())
-# ms$smooth_spectra(Settings_smooth_spectra_movingaverage(windowSize = 15))
-# ms$normalize_spectra(Settings_normalize_spectra_minmax()) #liftTozero = TRUE
-# ms$normalize_spectra(Settings_normalize_spectra_blockweight())
-# ms$bin_spectra(Settings_bin_spectra_StreamFind(bins = list("rt" = 2, "mass" = 25), refBinAnalysis = 1))
+# ms$cluster_spectra(Method_cluster_spectra_StreamFind(val = "mz", clustVal = 0.01, presence = 0.2))
+# ms$calculate_spectra_charges(Method_calculate_spectra_charges_StreamFind(roundVal = 15, relLowCut = 0.2, absLowCut = 8000))
+# ms$deconvolute_spectra(Method_deconvolute_spectra_StreamFind())
+# ms$smooth_spectra(Method_smooth_spectra_movingaverage(windowSize = 15))
+# ms$normalize_spectra(Method_normalize_spectra_minmax()) #liftTozero = TRUE
+# ms$normalize_spectra(Method_normalize_spectra_blockweight())
+# ms$bin_spectra(Method_bin_spectra_StreamFind(bins = list("rt" = 2, "mass" = 25), refBinAnalysis = 1))
 
@@
 # ms$plot_spectra_charges()
-# ms$correct_spectra_baseline(Settings_correct_spectra_baseline_baseline())
+# ms$correct_spectra_baseline(Method_correct_spectra_baseline_baseline())
 # ms$plot_deconvoluted_spectra()
@@
 
-Below we create an ordered list of *ProcessingSettings* objects for each processing method to be applied and add the list to the engine.
+Below we create an ordered list of *ProcessingStep* objects for each processing method to be applied and add the list to the engine.
 The exact parameters for each method can be optimized individually by calling the specific method directly.
@@
 ms_workflow_settings <- list(
-  Settings_cluster_spectra_StreamFind(val = "mz", clustVal = 0.01, presence = 0.2),
-  Settings_calculate_spectra_charges_StreamFind(roundVal = 15, relLowCut = 0.2, absLowCut = 8000),
-  Settings_deconvolute_spectra_StreamFind(),
-  Settings_smooth_spectra_movingaverage(windowSize = 15),
-  Settings_bin_spectra_StreamFind(bins = list("rt" = 5, "mass" = 20), refBinAnalysis = 1),
-  Settings_average_spectra_StreamFind(),
-  Settings_normalize_spectra_minmax(),
-  Settings_normalize_spectra_blockweight()
+  Method_cluster_spectra_StreamFind(val = "mz", clustVal = 0.01, presence = 0.2),
+  Method_calculate_spectra_charges_StreamFind(roundVal = 15, relLowCut = 0.2, absLowCut = 8000),
+  Method_deconvolute_spectra_StreamFind(),
+  Method_smooth_spectra_movingaverage(windowSize = 15),
+  Method_bin_spectra_StreamFind(bins = list("rt" = 5, "mass" = 20), refBinAnalysis = 1),
+  Method_average_spectra_StreamFind(),
+  Method_normalize_spectra_minmax(),
+  Method_normalize_spectra_blockweight()
 )
@@
 raman_workflow_settings <- list(
-  Settings_bin_spectra_StreamFind(unitsVal = "rt", unitsNumber = 3),
-  Settings_normalize_spectra_minmax(),
-  Settings_subtract_spectra_section_StreamFind(sectionWindow = c(0, 3)),
-  Settings_smooth_spectra_savgol(fl = 21, forder = 4, dorder = 0),
-  Settings_delete_spectra_section_StreamFind(list("shift" = c(-100, 330))),
-  Settings_delete_spectra_section_StreamFind(list("shift" = c(1800, 2600))),
-  Settings_correct_spectra_baseline_airpls(lambda = 45, differences = 1, itermax = 30),
-  Settings_delete_spectra_section_StreamFind(list("rt" = c(0, 5.5))),
-  Settings_delete_spectra_section_StreamFind(list("rt" = c(8, 11))),
-  Settings_average_spectra_StreamFind(collapseTime = TRUE),
-  Settings_normalize_spectra_minmax(),
-  Settings_normalize_spectra_blockweight()
+  Method_bin_spectra_StreamFind(unitsVal = "rt", unitsNumber = 3),
+  Method_normalize_spectra_minmax(),
+  Method_subtract_spectra_section_StreamFind(sectionWindow = c(0, 3)),
+  Method_smooth_spectra_savgol(fl = 21, forder = 4, dorder = 0),
+  Method_delete_spectra_section_StreamFind(list("shift" = c(-100, 330))),
+  Method_delete_spectra_section_StreamFind(list("shift" = c(1800, 2600))),
+  Method_correct_spectra_baseline_airpls(lambda = 45, differences = 1, itermax = 30),
+  Method_delete_spectra_section_StreamFind(list("rt" = c(0, 5.5))),
+  Method_delete_spectra_section_StreamFind(list("rt" = c(8, 11))),
+  Method_average_spectra_StreamFind(collapseTime = TRUE),
+  Method_normalize_spectra_minmax(),
+  Method_normalize_spectra_blockweight()
 )

### 2025-08-12 R/app_mod_WorkflowAssembler_MassSpecAnalyses.R
Updated readme, basic app and basic mass spec spectra

@@ -1,13 +1,16 @@
 #' @noRd
-.mod_WorkflowAssembler_Analyses_UI <- function(id, ns) {
+.mod_WorkflowAssembler_Analyses_UI.MassSpecAnalyses <- function(x, id, ns) {
   ns2 <- shiny::NS(id)
   shiny::column(
     12,
     shinydashboard::box(
-      title = "Analyses",
+      title = NULL,
       width = 12,
       solidHeader = TRUE,
-      shiny::column(width = 12, shiny::uiOutput(ns(ns2("analyses_overview_buttons")))),
+      shiny::column(
+        width = 12,
+        shiny::uiOutput(ns(ns2("analyses_overview_buttons")))
+      ),
       shiny::uiOutput(ns(ns2("notes_analyses"))),
       shiny::column(12, DT::dataTableOutput(ns(ns2("AnalysesTable"))))
     )
@@ -15,72 +18,87 @@
 }
 
 #' @noRd
-.mod_WorkflowAssembler_Analyses_Server <- function(id,
-                                                   ns,
-                                                   reactive_analyses,
-                                                   reactive_warnings,
-                                                   reactive_volumes,
-                                                   reactive_config) {
+.mod_WorkflowAssembler_Analyses_Server.MassSpecAnalyses <- function(
+  x,
+  id,
+  ns,
+  reactive_analyses,
+  reactive_warnings,
+  reactive_volumes,
+  reactive_config
+) {
   shiny::moduleServer(id, function(input, output, session) {
     ns2 <- shiny::NS(id)
 
     .wrap_analyses_ui_in_divs <- function(elements) {
-      lapply(elements, function(x) {
+      lapply(elements, function(z) {
         htmltools::div(
           style = sprintf(
-            "min-width: %dpx; height: %dpx; display: flex; align-items: center;", 40, 40
+            "min-width: %dpx; height: %dpx; display: flex; align-items: center;",
+            40,
+            40
           ),
-          x
+          z
         )
       })
     }
 
     # out Analyses Table -----
-    output$AnalysesTable <- DT::renderDT({
-      analyses <- reactive_analyses()
-      analyses_info <- analyses@info
-      if (length(analyses) == 0) {
-        analyses_info <- data.frame()
-        edits <- FALSE
-      } else if ("replicate" %in% colnames(analyses_info)) {
-        edits <- list(
-          target = "column",
-          disable = list(columns = which(!colnames(analyses_info) %in% c("replicate", "blank")))
[…]

### 2025-08-12 R/app_utils.R

@@ -15,9 +15,15 @@
 .app_util_get_volumes <- function() {
   os_type <- Sys.info()["sysname"]
   if (os_type == "Windows") {
-    drives <- system("wmic logicaldisk get name", intern = TRUE)
-    drives <- drives[grepl(":", drives)]
-    drives <- gsub("\\s+", "", drives)
+    # drives <- system("wmic logicaldisk get name", intern = TRUE)
+    # drives <- drives[grepl(":", drives)]
+    # drives <- gsub("\\s+", "", drives)
+    # names(drives) <- drives
+    drives <- system(
+      'powershell -NoProfile -Command "Get-PSDrive -PSProvider FileSystem | Select-Object -ExpandProperty Name"',
+      intern = TRUE
+    )
+    drives <- paste0(drives, ":")
     names(drives) <- drives
   } else {
     drives <- list.files("/media", full.names = TRUE)
@@ -174,7 +180,7 @@
         reactive_engine_save_file(NA_character_)
         reactive_engine_type(obj)
         shiny::removeModal()
-        if (!obj %in% "CoreEngine") {
+        if (!obj %in% "Engine") {
           reactive_show_init_modal(FALSE)
           reactive_app_mode("WorkflowAssembler")
           reactive_clean_start(TRUE)

### 2025-08-12 R/class_Engine_MassSpecEngine.R

@@ -2,7 +2,7 @@
 # MassSpecEngine -----
 #' Mass Spectrometry Engine
 #' 
-#' @description The `MassSpecEngine` R6 class is a framework for parsing, processing and inspecting mass spectrometry (MS) data. The engine has *MassSpec* as data type. The *Analyses* active field has class [StreamFind::MassSpecAnalyses], where MS data (i.e., spectra and chromatograms, including chromatograms produced by UV detection) can be loaded from *mzML* and *mzXML*. See methods [StreamFind::MassSpecAnalyses] for more details. If `msconvert` from \href{https://proteowizard.sourceforge.io/}{ProteoWizard} is installed and found via CLI (i.e., must be added to the environmental variables), the engine can also load vendor formats (i.e., Agilent MassHunter .d, Thermo Scientific RAW, Shimadzu LCD (except ITOF), Sciex WIFF /WIFF2) by direct conversion to mzML. Note that conversion of vendor formats is only possible under Windows OS. The *results* active field is a list with that can hold the classes: [StreamFind::NonTargetAnalysisResults], [StreamFind::MassSpecSpectra] and [StreamFind::Chromatograms]. Note that the *MassSpecEngine* is a subclass of [StreamFind::Engine] and inherits all methods from the parent class.
[…]

### 2025-08-12 R/class_MassSpecMethod_GenerateFormulas.R

@@ -177,14 +177,14 @@ S7::method(run, MassSpecMethod_GenerateFormulas_genform) <- function(x, engine =
   }
 
   if (!engine$has_results_nts()) {
-    warning("No NonTargetAnalysisResults object available! Not done.")
+    warning("No MassSpecResults_NonTargetAnalysis object available! Not done.")
     return(FALSE)
   }
 
-  NonTargetAnalysisResults <- engine$NonTargetAnalysisResults
+  MassSpecResults_NonTargetAnalysis <- engine$MassSpecResults_NonTargetAnalysis
 
-  if (!NonTargetAnalysisResults@has_groups) {
-    warning("NonTargetAnalysisResults object does not have feature groups! Not done.")
+  if (!MassSpecResults_NonTargetAnalysis@has_groups) {
+    warning("MassSpecResults_NonTargetAnalysis object does not have feature groups! Not done.")
     return(FALSE)
   }
 
@@ -200,13 +200,13 @@ S7::method(run, MassSpecMethod_GenerateFormulas_genform) <- function(x, engine =
   algorithm <- x$algorithm
   
   fg <- get_patRoon_features(
-    NonTargetAnalysisResults,
+    MassSpecResults_NonTargetAnalysis,
     filtered = FALSE,
     featureGroups = TRUE
   )
   
   mspl <- get_patRoon_MSPeakLists(
-    NonTargetAnalysisResults,
+    MassSpecResults_NonTargetAnalysis,
     MSPeakListsClusterMzWindow,
     MSPeakListsTopMost,
     MSPeakListsMinIntensityPre,
@@ -241,7 +241,7 @@ S7::method(run, MassSpecMethod_GenerateFormulas_genform) <- function(x, engine =
     return(FALSE)
   }
 
-  feature_list <- NonTargetAnalysisResults$feature_list
+  feature_list <- MassSpecResults_NonTargetAnalysis$feature_list
 
   formulas_col <- lapply(names(feature_list), function(x, feature_list, formulas) {
     formula <- formulas@featureFormulas[[x]]
@@ -261,11 +261,11 @@ S7::method(run, MassSpecMethod_GenerateFormulas_genform) <- function(x, engine =
     rep(list(NULL), nrow(feature_list[[x]]))
   }, feature_list = feature_list, formulas = formulas)
 
-  NonTargetAnalysisResults <- .add_features_column(NonTargetAnalysisResults, "formulas", formulas_col)
[…]

### 2025-08-12 R/class_MassSpecMethod_GroupFeatures.R

@@ -10,17 +10,17 @@
     return(FALSE)
   }
 
-  if (is.null(engine$Analyses$results[["NonTargetAnalysisResults"]])) {
-    warning("No NonTargetAnalysisResults object available! Not done.")
+  if (is.null(engine$Analyses$results[["MassSpecResults_NonTargetAnalysis"]])) {
+    warning("No MassSpecResults_NonTargetAnalysis object available! Not done.")
     return(FALSE)
   }
 
-  nts <- engine$ResultsList$NonTargetAnalysisResults
+  nts <- engine$ResultsList$MassSpecResults_NonTargetAnalysis
 
   if (
     sum(vapply(nts$features, function(z) nrow(z), 0)) == 0
   ) {
-    warning("NonTargetAnalysisResults object does not have features! Not done.")
+    warning("MassSpecResults_NonTargetAnalysis object does not have features! Not done.")
     return(FALSE)
   }
 
@@ -218,8 +218,8 @@ MassSpecMethod_GroupFeatures_xcms3_peakdensity <- function(
     method = "GroupFeatures",
     required = "FindFeatures",
     algorithm = "xcms3_peakdensity",
-    input_class = "NonTargetAnalysisResults",
-    output_class = "NonTargetAnalysisResults",
+    input_class = "MassSpecResults_NonTargetAnalysis",
+    output_class = "MassSpecResults_NonTargetAnalysis",
     parameters = list(
       bw = bw,
       minFraction = minFraction,
@@ -383,8 +383,8 @@ MassSpecMethod_GroupFeatures_xcms3_peakdensity_peakgroups <- function(
     method = "GroupFeatures",
     required = "FindFeatures",
     algorithm = "xcms3_peakdensity_peakgroups",
-    input_class = "NonTargetAnalysisResults",
-    output_class = "NonTargetAnalysisResults",
+    input_class = "MassSpecResults_NonTargetAnalysis",
+    output_class = "MassSpecResults_NonTargetAnalysis",
     parameters = list(
       bw = bw,
       minFraction = minFraction,
@@ -547,8 +547,8 @@ MassSpecMethod_GroupFeatures_openms <- function(
     method = "GroupFeatures",
     required = "FindFeatures",
     algorithm = "openms",
-    input_class = "NonTargetAnalysisResults",
-    output_class = "NonTargetAnalysisResults",
+    input_class = "MassSpecResults_NonTargetAnalysis",
[…]

### 2025-08-13 R/class_MassSpecMethod_FillFeatures.R
updated demo_nts vin

@@ -120,12 +120,12 @@ run.MassSpecMethod_FillFeatures_StreamFind <- function(x, engine = NULL) {
     return(FALSE)
   }
 
-  nts <- engine$ResultsList$MassSpecResults_NonTargetAnalysis
+  nts <- engine$Results$MassSpecResults_NonTargetAnalysis
 
   if (
     !any(vapply(
       nts$features,
-      function(z) any(!(is.na(z$group) || z$group %in% "")),
+      function(z) any(!(is.na(z$group) | z$group %in% "")),
       FALSE
     ))
   ) {
@@ -159,7 +159,7 @@ run.MassSpecMethod_FillFeatures_StreamFind <- function(x, engine = NULL) {
   nts$features <- feature_list
   tryCatch(
     {
-      engine$ResultsList <- nts
+      engine$Results <- nts
       return(TRUE)
     },
     error = function(e) {

### 2025-08-13 R/class_MassSpecMethod_GenerateCompounds.R

@@ -1,11 +1,6 @@
-#' MassSpecMethod_GenerateCompounds_metfrag S7 class
+#' MassSpecMethod_GenerateCompounds_metfrag Class
 #'
-#' @description Settings for generating compounds using
-#' \href{https://ipb-halle.github.io/MetFrag/}{MetFrag}. The algorithm is used via the function
-#' \link[patRoon]{generateCompounds} from the package \pkg{patRoon}. Therefore, it is highly
-#' recommended to check the original documentation of the function in \pkg{patRoon} for more
-#' details.
-#' 
+#' @description Settings for generating compounds using \href{https://ipb-halle.github.io/MetFrag/}{MetFrag}. The algorithm is used via the function \link[patRoon]{generateCompounds} from the package \pkg{patRoon}. Therefore, it is highly recommended to check the original documentation of the function in \pkg{patRoon} for more details.
 #' @param method Character (length 1) with the method to be used for MetFrag execution: "CL" for MetFragCL and "R" for MetFragR.
 #' @param timeout Numeric (length 1) with the maximum time (in seconds) before a MetFrag query for a feature group is stopped.
 #' @param timeoutRetries Numeric (length 1) with the maximum number of retries after reaching a timeout before completely
@@ -16,24 +11,17 @@
 #' @param fragRelMzDev Numeric (length 1) with the relative mass deviation, in ppm, for the fragment search.
 #' @param fragAbsMzDev Numeric (length 1) with the absolute mass deviation, in Da, for the fragment search.
 #' @param adduct Character (length 1) with the adduct to be used for the MetFrag annotation.
-#' @param database Character (length 1) with the database to be used for the MetFrag annotation. Valid values are:
-#' "pubchem", "chemspider", "for-ident", "comptox", "pubchemlite", "kegg", "sdf", "psv" and "csv".
-#' @param extendedPubChem Extended PubChem database is used for the MetFrag annotation when `database` is "pubchem".
-#' Valid values are: FALSE (never use it), TRUE (always use it) or "auto" (default, use if specified scorings demand it).
+#' @param database Character (length 1) with the database to be used for the MetFrag annotation. Valid values are: "pubchem", "chemspider", "for-ident", "comptox", "pubchemlite", "kegg", "sdf", "psv" and "csv".
[…]

### 2025-08-14 R/class_Engine_MassSpecEngine.R
Updated Raman Classes

@@ -2,7 +2,7 @@
 # MassSpecEngine -----
 #' Mass Spectrometry Engine
 #' 
-#' @description The `MassSpecEngine` R6 class is a framework for parsing, processing and inspecting mass spectrometry (MS) data. The engine has *MassSpec* as data type. The *Analyses* active field has class [StreamFind::MassSpecAnalyses], where MS data (i.e., spectra and chromatograms, including chromatograms produced by UV detection) can be loaded from *mzML* and *mzXML*. See methods [StreamFind::MassSpecAnalyses] for more details. If `msconvert` from \href{https://proteowizard.sourceforge.io/}{ProteoWizard} is installed and found via CLI (i.e., must be added to the environmental variables), the engine can also load vendor formats (i.e., Agilent MassHunter .d, Thermo Scientific RAW, Shimadzu LCD (except ITOF), Sciex WIFF /WIFF2) by direct conversion to mzML. Note that conversion of vendor formats is only possible under Windows OS. The *results* active field is a list with that can hold the classes: [StreamFind::MassSpecResults_NonTargetAnalysis], [StreamFind::MassSpecSpectra] and [StreamFind::Chromatograms]. Note that the *MassSpecEngine* is a subclass of [StreamFind::Engine] and inherits all methods from the parent class.
+#' @description The `MassSpecEngine` R6 class is a framework for parsing, processing and inspecting mass spectrometry (MS) data. The engine has *MassSpec* as data type. The *Analyses* active field has class [StreamFind::MassSpecAnalyses], where MS data (i.e., spectra and chromatograms, including chromatograms produced by UV detection) can be loaded from *mzML* and *mzXML*. See methods [StreamFind::MassSpecAnalyses] for more details. If `msconvert` from \href{https://proteowizard.sourceforge.io/}{ProteoWizard} is installed and found via CLI (i.e., must be added to the environmental variables), the engine can also load vendor formats (i.e., Agilent MassHunter .d, Thermo Scientific RAW, Shimadzu LCD (except ITOF), Sciex WIFF /WIFF2) by direct conversion to mzML. Note that conversion of vendor formats is only possible under Windows OS. The *results* active field is a list that can hold the classes: [StreamFind::MassSpecResults_NonTargetAnalysis], [StreamFind::MassSpecSpectra] and [StreamFind::Chromatograms]. Note that the *MassSpecEngine* is a subclass of [StreamFind::Engine] and inherits all methods from the parent class.
 #' 
[…]

### 2025-08-14 R/class_MassSpecMethod_SubtractBlankSpectra.R

@@ -70,9 +70,9 @@ S7::method(run, MassSpecMethod_SubtractBlankSpectra_StreamFind) <- function(x, e
   
   ntozero <- x$parameters$negativeToZero
   
-  blks <- get_blanks(engine$Analyses)
-  names(blks) <- get_replicates(engine$Analyses)
-  blk_anas <- get_replicates(engine$Analyses)
+  blks <- get_blank_names(engine$Analyses)
+  names(blks) <- get_replicate_names(engine$Analyses)
+  blk_anas <- get_replicate_names(engine$Analyses)
   blk_anas <- blk_anas[blk_anas %in% blks]
   
   spec_blk <- spec_list[names(spec_list) %in% c(blks, names(blk_anas))]
@@ -93,7 +93,7 @@ S7::method(run, MassSpecMethod_SubtractBlankSpectra_StreamFind) <- function(x, e
     if (nrow(z) == 0) return(z)
     
     if (!engine$Spectra$is_averaged) {
-      rp <- get_replicates(engine$Analyses)[z$analysis[1]]
+      rp <- get_replicate_names(engine$Analyses)[z$analysis[1]]
       
     } else {
       rp <- z$analysis[1]

### 2025-08-14 R/class_RamanMethod_BinScans.R

@@ -1,93 +1,96 @@
 #' Raman Method for Binning Time Scans (native algorithm)
 #'
 #' @description Merges scans by averaging spectra according to a given binning size.
-#' 
+#'
 #' @param mode Character (length 1) with the binning mode. Possible values are `scans` and `time`.
-#' When `scans` is selected, the binning is done by the number of scans. When `time` is selected, 
+#' When `scans` is selected, the binning is done by the number of scans. When `time` is selected,
 #' the binning is done by time unit (e.g., seconds or minutes).
 #' @param value Numeric (length 1) with the binning size.
 #' @param global Logical (length 1) to perform a global binning. If `TRUE`, the binning is performed
 #' over all analyses. If `FALSE`, the binning is performed for each analysis.
 #' @param refAnalysis Integer (length 1) with the reference analysis index to use for binning.
-#' 
+#'
 #' @returns A RamanMethod_BinScans_native object.
 #'
-#' @export 
+#' @export
 #'
-RamanMethod_BinScans_native <- S7::new_class(
-  name = "RamanMethod_BinScans_native",
-  parent = S7::new_S3_class("ProcessingStep"),
-  package = "StreamFind",
-  constructor = function(mode = c("unit"),
-                         value = 5,
-                         global = FALSE,
-                         refAnalysis = NA_integer_) {
-    S7::new_object(
-      ProcessingStep(
-        data_type = "Raman",
-        method = "BinScans",
-        required = NA_character_,
-        algorithm = "native",
-        parameters = list(
-          mode = mode,
-          value = value,
-          global = global,
-          refAnalysis = refAnalysis
-        ),
-        number_permitted = 1,
-        version = as.character(packageVersion("StreamFind")),
-        software = "StreamFind",
-        developer = "Ricardo Cunha",
-        contact = "cunha@iuta.de",
-        link = "https://odea-project.github.io/StreamFind",
-        doi = NA_character_
-      )
-    )
-  },
-  
-  validator = function(self) {
-    checkmate::assert_choice(self@data_type, "Raman")
-    checkmate::assert_choice(self@method, "BinScans")
[…]

### 2025-08-14 dev/engine_RamanEngine_old.R

@@ -1,5 +1,4 @@
 # MARK: RamanEngine
-# RamanEngine -----
 #' Raman Spectroscopy Engine
 #'
 #' @description The *RamanEngine* R6 class is a framework for parsing, processing, inspecting and 

### 2025-08-14 R/class_RamanMethod_AverageSpectra.R
updated vin and all raman and ms methods

@@ -65,37 +65,28 @@ run.RamanMethod_AverageSpectra_native <- function(x, engine = NULL) {
     warning("Engine is not a RamanEngine object!")
     return(FALSE)
   }
-
   if (!engine$has_analyses()) {
     warning("There are no analyses! Not done.")
     return(FALSE)
   }
-
-  if (!engine$Analyses$has_spectra) {
-    warning("No spectra results object available! Not done.")
-    return(FALSE)
+  if (is.null(engine$Results[["RamanResults_Spectra"]])) {
+    engine$Results <- RamanResults_Spectra(
+      lapply(engine$Analyses$analyses, function(a) a$spectra)
+    )
   }
-
-  if (engine$Spectra$is_averaged) {
+  spec_obj <- engine$Results[["RamanResults_Spectra"]]
+  if (spec_obj$is_averaged) {
     warning("Spectra are already averaged! Not done.")
     return(FALSE)
   }
-
-  spectra_list <- engine$Spectra$spectra
-
+  spectra_list <- spec_obj$spectra
   spectra <- data.table::rbindlist(
     spectra_list,
     idcol = "analysis",
     fill = TRUE
   )
-
-  if (engine$Spectra$is_averaged) {
-    spectra$replicate <- spectra$analysis
-  } else {
-    rpl <- get_replicate_names(engine$Analyses)
-    spectra$replicate <- rpl[spectra$analysis]
-  }
-
+  rpl <- get_replicate_names(engine$Analyses)
+  spectra$replicate <- rpl[spectra$analysis]
   groupCols <- "shift"
   if (grepl("chrom_peaks", x$parameters$by, fixed = FALSE)) {
     groupCols <- c("chrom_peaks", groupCols)
@@ -106,9 +97,8 @@ run.RamanMethod_AverageSpectra_native <- function(x, engine = NULL) {
   if (grepl("replicates", x$parameters$by, fixed = FALSE)) {
     groupCols <- c("replicate", groupCols)
   }
-
   if ("chrom_peaks" %in% groupCols) {
-    if (engine$Spectra$has_chrom_peaks) {
+    if (length(spec_obj$chrom_peaks) > 0) {
       if (!"id" %in% colnames(spectra)) {
         warning(
           "Filter spectra to keep only from chromatographic peaks 
@@ -127,27 +117,21 @@ run.RamanMethod_AverageSpectra_native <- function(x, engine = NULL) {
       return(FALSE)
[…]

### 2025-08-15 R/generics_S3.R
updated statistics and webpage

@@ -33,6 +33,16 @@ clear_cache <- function(x, ...) {
 
 # G -----
 
+#' @title S3 generic `get_analysis_classes`
+#' @description S3 generic to get analysis classes from `x`.
+#' @param x An object from which the analysis classes will be retrieved.
+#' @param ... Additional arguments passed to the method.
+#' @export
+#' 
+get_analysis_classes <- function(x, ...) {
+  UseMethod("get_analysis_classes")
+}
+
 #' @title S3 generic `get_analysis_names`
 #' @description S3 generic to get names from `x`.
 #' @param x An object from which the names will be retrieved.
@@ -745,6 +755,16 @@ plot_matrix_suppression <- function(x, ...) {
   UseMethod("plot_matrix_suppression")
 }
 
+#' @title Generic `plot_overview`
+#' @description S3 generic to plot an overview from `x`.
+#' @param x An object from which the overview will be plotted.
+#' @param ... Additional arguments passed to the method.
+#' @export
+#' 
+plot_overview <- function(x, ...) {
+  UseMethod("plot_overview")
+}
+
 #' @title Generic `plot_residuals`
 #' @description S3 generic to plot residuals from `x`.
 #' @param x An object from which the residuals will be plotted.
@@ -979,6 +999,16 @@ save <- function(x, ...) {
 #' 
 save_cache <- function(x, ...) {
   UseMethod("save_cache")
+}
+
+#' @title Generic `set_analysis_classes`
+#' @description S3 generic to set analysis classes in `x`.
+#' @param x An object in which the analysis classes will be set.
+#' @param ... Additional arguments passed to the method.
+#' @export
+#' 
+set_analysis_classes <- function(x, ...) {
+  UseMethod("set_analysis_classes")
 }
 
 #' @title Generic `set_blank_names`

### 2025-08-18 R/class_RamanMethod_ScaleTime.R
updated website and documentation

@@ -32,10 +32,8 @@ RamanMethod_ScaleTime_native <- function(value = 0) {
   }
 }
 
-#' @describeIn RamanMethod_ScaleTime_native Validate the RamanMethod_ScaleTime_native object, returning NULL if valid.
-#' @param x A RamanMethod_ScaleTime_native object.
 #' @export
-#'
+#' @noRd
 validate_object.RamanMethod_ScaleTime_native <- function(x) {
   checkmate::assert_choice(x$type, "Raman")
   checkmate::assert_choice(x$method, "ScaleTime")

### 2025-09-01 R/class_MassSpecResults_NonTargetAnalysis.R
Updated docs and small fixes

@@ -2846,7 +2846,7 @@ plot_groups_profile.MassSpecResults_NonTargetAnalysis <- function(
     filtered
   )
 
-  analyses <- .check_analyses_argument(x, analyses)
+  analyses <- .check_analyses_argument(x$features, analyses)
 
   if (nrow(fts) == 0) {
     message("\U2717 Features not found for the targets!")
@@ -2887,21 +2887,23 @@ plot_groups_profile.MassSpecResults_NonTargetAnalysis <- function(
       .(intensity = mean(intensity), intensity_sd = sd(intensity)),
       by = group_cols
     ]
-    names(polarities) <- x$info$replicate[names(polarities)]
+    rpls <- x$info$replicate
+    names(rpls) <- x$info$analysis
+    names(polarities) <- rpls[names(polarities)]
     polarities <- polarities[!duplicated(names(polarities))]
     data.table::setnames(fts, "replicate", "analysis")
-    analyses <- unique(x$info$replicate[analyses])
+    analyses <- unique(rpls[rpls %in% fts$analysis])
   }
 
   if (
-    is.character(legendNames) &
+    is.character(legendNames) &&
       length(legendNames) == length(unique(fts$group))
   ) {
     leg <- legendNames
     names(leg) <- unique(fts$group)
     leg <- leg[fts$group]
     fts$var <- leg[fts$group]
-  } else if (isTRUE(legendNames) & "name" %in% colnames(fts)) {
+  } else if (isTRUE(legendNames) && "name" %in% colnames(fts)) {
     leg <- fts$name
     fts$var <- fts$name
   } else {

### 2025-09-02 README.Rmd

@@
 ```{r, echo=FALSE, out.width = '100%', fig.align='center', fig.alt="Logo"}
-knitr::include_graphics("man/figures/logos.png")
+knitr::include_graphics("man/figures/readme_logos.png")
 ```
@@
 RTools is needed for compiling C++ code used in the StreamFind R package.
-StreamFind also uses python scripts for some of its processing methods, so it is recommended to have the latest [python](https://www.python.org/downloads/) installed and added to the environmental variables for Windows users.
-Assuming that R, RTools and python are installed, the StreamFind R package can be installed from the GitHub repository via the [BiocManager](https://www.bioconductor.org/install/).
+StreamFind also uses Python scripts for some of its processing methods, so it is recommended to have the latest [Python](https://www.python.org/downloads/) installed and added to the environmental variables for Windows users.
+Assuming that R, RTools and Python (optional) are installed, the StreamFind R package can be installed from the GitHub repository via the [BiocManager](https://www.bioconductor.org/install/).
 Note that StreamFind is not yet available on the Bioconductor repository.
+Please note that the default timeout for the `BiocManager::install` function might be to short.
+Thus, if the download of StreamFind fails run `options(timeout = 600)` before the installation command.
 
 ```{r install-streamfind, eval=FALSE}
+options(timeout = 600) # Increase timeout
 if (!require("BiocManager", quietly = TRUE))

### 2025-09-03 README.Rmd
updated with app fix

@@
 Start the Docker container:
-`docker run -it -p 3838:3838 -p 8787:8787 -v $(pwd):/app my-r-app`
+
+**For Linux/macOS/bash:**
+```bash
+docker run -it -p 3838:3838 -p 8787:8787 -v $(pwd):/app my-r-app
+```
+
+**For Windows PowerShell:**
+```powershell
+docker run -it -p 3838:3838 -p 8787:8787 -v "${PWD}:/app" my-r-app
+```
 

### 2025-09-29 README.Rmd

@@
 RTools is needed for compiling C++ code used in the StreamFind R package.
-StreamFind also uses Python scripts for some of its processing methods, so it is recommended to have the latest [Python](https://www.python.org/downloads/) installed and added to the environmental variables for Windows users.
-Assuming that R, RTools and Python (optional) are installed, the StreamFind R package can be installed from the GitHub repository via the [BiocManager](https://www.bioconductor.org/install/).
-Note that StreamFind is not yet available on the Bioconductor repository.
+StreamFind also uses Python scripts for some of its processing methods, so it is recommended (but not mandatory) to have the latest [Python](https://www.python.org/downloads/) installed and added to the environmental variables for Windows users.
+Assuming that R, R Tools and Python (optional) are installed, the StreamFind R package can be installed from the GitHub repository via the [Bioconductor Manager](https://www.bioconductor.org/install/), which makes it easier to install necessary Bioconductor dependencies (e.g.
+zlibbioc and Rhdf5lib).
+Note that StreamFind itself is not available on Bioconductor.
 Please note that the default timeout for the `BiocManager::install` function might be to short.
-Thus, if the download of StreamFind fails run `options(timeout = 600)` before the installation command.
+Thus, if the download of StreamFind fails run `options(timeout = 600)` before the installation command, as shown below.
 
@@
 
-The StreamFind depends on other open source software to process different analytic data.
-For instance, for non-target screening using mass spectrometry the StreamFind uses the [patRoon](https://github.com/rickhelmus/patRoon) R package and its own dependencies for some processing methods.
-When a dependency is not installed, a warning message will be displayed with instructions.
-Consult the documentation of each processing method for dependencies.
+Some processing methods used by StreamFind depend on other open-source software.
+For example, when conducting non-target screening using mass spectrometry, StreamFind utilises the [patRoon R package](https://github.com/rickhelmus/patRoon) and its associated dependencies for certain processing methods.
+If a dependency is not installed, a warning message with instructions will be displayed.
+Please consult the documentation for each processing method for details of its dependencies.
 

### 2025-10-28 dev/dev_filesystem_test_functions.R
added file system test framework

@@ -1,0 +1,1426 @@
+#' @title Filesystem Performance Evaluation for HDF5, DuckDB, and SQLite
+#' 
+#' @description
+#' This script provides comprehensive performance testing for different data storage formats:
+#' - HDF5: High-performance binary format (compiled via Makevars and load_all)
+#' - DuckDB: Analytical database with compression support
+#' - SQLite: Lightweight embedded database
+#' 
+#' The evaluation tests both read and write performance across different data sizes
+#' and includes support for mixed data types (numeric and character columns).
+#' 
+#' Data sizes tested (as specified in the requirements):
+#' - 10x10 (100 elements)
+#' - 1000x1000 (1M elements)
+#' - 100000x100000 (10B elements)
+#' - 1000000x1000000 (1T elements)
+#' 
+#' Features:
+#' - Configurable number of character columns for testing non-numeric data
+#' - Detailed performance metrics (write time, read time, file size, throughput)
+#' - Comparison reports across all formats
+#' - Support for HDF5 compression and chunking
+#' - Proper cleanup functions for test files
+#' 
+#' @author StreamFind Development Team
+#' @date 2024
+
+# Load required libraries
+library(DBI)
+library(duckdb)
+library(RSQLite)
+library(data.table)
+
+# Note: HDF5 functions are available via load_all() which compiles the C++ code through Makevars
+
+#' Generate test data with specified dimensions and character columns
+#'
+#' @param n_rows Number of rows
+#' @param n_cols Number of numeric columns
+#' @param number_character_cols Number of character columns to add
+#' @return data.table with numeric and character columns
+generate_test_data <- function(n_rows, n_cols, number_character_cols = 0) {
+  
+  # Generate numeric data
+  data_list <- list()
+  
+  # Add numeric columns
+  for (i in seq_len(n_cols)) {
+    data_list[[paste0("num_", i)]] <- rnorm(n_rows, mean = i * 10, sd = 5)
+  }
+  
+  # Add character columns if requested
+  if (number_character_cols > 0) {
+    char_samples <- c("alpha", "beta", "gamma", "delta", "epsilon", "zeta", "eta", "theta")
+    
+    for (i in seq_len(number_character_cols)) {
+      # Generate character data with varying lengths
+      char_data <- sample(char_samples, n_rows, replace = TRUE)
+      # Add some variation in string length
[…]
